### 2019-10-14 Nicolas.tex
Initial commit

\documentclass{aa}
\usepackage{preambule}
\usepackage[varg]{txfonts}

\begin{document}
\title{Redshift evolution of the SN stretch distribution}

%\subtitle{II. An example text with infinitesimal scientific value whose title
%and subtitle may also be split}

\author{N. Nicolas\inst{1}
    \and M. Rigault\inst{2}
}

%\offprints{R. Plemmons, \email{plemmons@...}}

\institute{Université de Lyon, F-69622, Lyon, France; Université de Lyon
    1, Villeurbanne; CNRS/IN2P3, Institut de Physique des Deux Infinis, Lyon
    \and Université Clermont Auvergne, CNRS/IN2P3, Laboratoire de
    Physique de Clermont, F-63000 Clermont-Ferrand, France.
}

\date{Received 2 November 1992 / Accepted 7 January 1993}

\abstract{Type Ia supernovae (SNe Ia) allow for the construction of the Hubble
    diagram, giving us information about the Universe's expansion and its
    fondamental components, one of which is dark energy. But systematic
    uncertainties are now starting to be limiting in our ability to measure
    those parameters. In particular, the physics of SNe Ia is still mostly
    unknown, and is thought not to change in time/with the redshift.}
    {In an attempt to reduce those uncertainties, we try to find an empirical
    law describing SNe Ia's length of explosion (stretch) evolution with the
redshift.}
    {We started by getting a complete sample representing all of the stretch
        distribution that Nature can give us, before using LsSFR measurments, an
        age tracer which evolution with redshift is known, that has been shown
        to have a strong correlation with the stretch. We compare their AICc, an
    estimator of the relative quality of statistical models that includes the
number of free parameters, to determine which ones describe besto the data.}
    {Models with an evolution of the stretch with the redshift have a better
    AICc than the ones without.}
    {We find that implementing these models allows us to fit the data better
    than models without stretch evolution.}

\keywords{Cosmology -- Type Ia Supernova -- Systematic uncertainties}
\maketitle

\section{Introduction}

Type Ia supernovae (SNe Ia) are now well-known for their capacity to determine
cosmological parameters: their study led to the discovery of the accelerated
expansion of the Universe (Riess 98, Perlmutter 99) through the name of "dark
energy", and they have been used continuously for better measurments since then
(Betoule 2014). They are acquired through their lightcurves, giving the
evolution of their luminosity from the time of explosion, in different
wavelength. 3 parameters are used to describe those: an amplitude, a width
(named "stretch") and a color (magnitude difference in the B and V bands).

The simple use of those is not enough for our aim, as SNe Ia have an intrinsec
dispersion of their luminosity of $\approx \SI{0.4}{mag}$ that gives a huge
uncertainty on the determination of their distance modulus. Henceforth they are
standardized using the "brigther-slower" and "brighter-bluer" relations
(Philipps 93, Riess 96, Tripp 98) in the SALT2 algorithm (Guy 2007, 2010) that
fits the distance modulus which is expressed as
\begin{center}
    $ \displaystyle \m = m_b + \a x_1 - \b c - M$
\end{center}
with $m_B$ the logarithm of their flux, $x_1$ their stretch, $c$ their color and
$M$ their intrinsec magnitude. This relation lowered their magnitude dispersion
to $\approx \SI{0.15}{mag}$, allowing for previously mentioned accelerated
expansion to be discovered.

This Tripp estimator lies on the idea that this standardization doesn't change
with the redshift. However, Rigault 2015 showed that SNe Ia depend on their
environments, and these environments' properties evolve with the redshift. In
this Letter, we try to determine whether a stretch evolution with the redshift
allows for a better description of the collected data.

\section{Sample}
We use data from 5 different surveys, 3 of which had selection effects that we
tried to remove with a statistical approach in lack of precise data concerning
the instruments' capacity to acquire fluxes. Here it is:

\section{Method}
We used LsSFR measurments which evolution with the redshift is analytically
known (references) and stretch measurments made by SNf to try an correlate these
to parameters. We fit different model, and compare their AICc.

\section{Results}
We find that every model lacking an evolution of the stretch with the redshift
is systematically worse than those that implement it.

\section{Conclusion}
Stretch evolution with the redshift is a thing. Need to see if it has an impact
on the cosmology though.

\end{document}

### 2019-10-14 Nicolas.tex
Petites modif, gitignore

@@
 \section{Sample}
-We use data from 5 different surveys, 3 of which had selection effects that we
-tried to remove with a statistical approach in lack of precise data concerning
-the instruments' capacity to acquire fluxes. Here it is:
+We use data from 5 different surveys: Hubble Space Telescope (HST, REF),
+Supernova Nearby Factory (SNf, REF), Supernova Legacy Survey (SNLS, REF), Sloan
+DSS (SDSS, REF) and Panstarr-1 (PS1, REF). 3 of which had selection effects that
+we tried to remove with a statistical approach in lack of precise data
+concerning the instruments' capacity to acquire fluxes. Here it is:
 

### 2019-10-14 Nicolas.tex
First draft sec:sam

--- 2019-10-14 Nicolas.tex
@@
 
-\section{Sample}
-We use data from 5 different surveys, 3 of which had selection effects that we
-tried to remove with a statistical approach in lack of precise data concerning
-the instruments' capacity to acquire fluxes. Here it is:
+\section{Sample}\label{sec:sam}
+
+We use data from 5 different surveys: Hubble Space Telescope (HST, REF),
+Supernova Nearby Factory (SNf, REF), Supernova Legacy Survey (SNLS, Aster 2006,
+Betoule 2014), Sloan DSS (SDSS, REF) and Panstarr-1 (PS1, REF). The last 3 had
+selection effects that come from the "brighter-slower" and "brighter-bluer"
+relations: the first SNe Ia that are missed by the instruments are the redder
+and faster ones. That implies that taking all of these surveys' data would not
+represent the stretch distribution that Nature could give us.
+
+As such, we determined the maximum redshift at which we would cut these samples
+to get complete ones. We used a statistical approach in lack of precise data
+concerning the instruments' capacity to acquire fluxes. We proceeded by making
+histograms of the surveys, and defining a function of the expected SNe Ia rate:
+\begin{center}
+    $ \displaystyle N\St{SNe Ia} = a \times V\z$
+\end{center}
+with $a$ the density of SNe Ia and $V\z$ the volume of Universe in each bin.
+Before selection effects, the histograms are supposed to grow proportionnaly to
+$V\z$ assuming an homogeneous repartition of SNe Ia. When we start to miss some
+supernovae, this growth is not followed anymore.
+
+To determine this limit, we first took the observed counts in each bin of the
+histograms. We then fitted $a$ on these counts for different numbers of bins
+used to fit: the quality of the fit diminishes when we use bins that include
+selection effects. To quantize this, we computed the expected counts from the
+fitted $a$ in the same bins we used for the observed counts. That allowed us to
+compute the poissonian probability in each bin, given by
+\begin{center}
+    $ \displaystyle p(k) = \frac{\lambda^k}{k!}e^{-\lambda}$
+\end{center}
+with $k$ the expected counts, $\lambda$ the observed counts. This work is done 100
+times with a random number of total bins, $a$ being fitted 4 times on a
+different and random number of bins used to fit each time. We interpolate these
+400 poissonian evolutions to get one final result that is as close as possible
+to accurately represent how the probability to follow the expected rate evolves.
+
+To get the maximum redshift after which we stop following the expected rate, we
+took the redshift for which we are at a 0.3 on the interpolated cumulative
+distribution function. We chose that value to have a systematic analytical
+result that matches what we expect from the histograms; its impact on the work
+after is to be studied.
+
+\begin{table}[htbp!]
+    \centering
+    \caption{Number of SNe Ia effectively used for each survey.}
+    \label{tab:sample}
+    \begin{tabularx}{\linewidth}{|Y*{1}{|Y}|}\hline
+        \rowcolor{gray!15} Survey & Number of SNe Ia \\\hline\hline
+        SNf & 141 \\\hline
+        PS1 & 178 \\\hline
+        SDSS & 206 \\\hline
+        SNLS & 138 \\\hline
+        HST & 26 \\\hline
+    \end{tabularx}
+\end{table}
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/zmax_determination.pdf}
+    \caption{$z\St{max}$ determination from the interpolated poissonian
+    evolutions.}
+    \label{fig:zdet}
+\end{figure}
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/surveys_cuts_x.pdf}
+    \caption{SDSS, SNLS and PS1 samples cut at $z\St{max}$ as explained in
+    section \ref{sec:sam}. The data we use is in plain markers.}
+    \label{fig:surveys_cuts}
+\end{figure}
 

### 2019-10-18 Nicolas.tex
Updated plots from variaIa update

@@
 \usepackage[varg]{txfonts}
+
+\def\l{\lambda}\def\L{\Lambda}
 
@@
 \begin{center}
-    $ \displaystyle p(k) = \frac{\lambda^k}{k!}e^{-\lambda}$
+    $ \displaystyle p(k) = \frac{\l^k}{k!}e^{-\l}$
 \end{center}
-with $k$ the expected counts, $\lambda$ the observed counts. This work is done 100
-times with a random number of total bins, $a$ being fitted 4 times on a
+with $k$ the expected counts, $\l$ the observed counts. This work is done 100
+times with a random number of total bins, $a$ being fitted 10 times on a
 different and random number of bins used to fit each time. We interpolate these
-400 poissonian evolutions to get one final result that is as close as possible
+1000 poissonian evolutions to get one final result that is as close as possible
 to accurately represent how the probability to follow the expected rate evolves.
@@
     \centering
-    \caption{Number of SNe Ia effectively used for each survey.}
+    \caption{Number of SNe Ia effectively used for each survey, totalling 689
+    Sne Ia. TO CORRECT FOR NEW ZMAX}
     \label{tab:sample}
@@
     \centering
-    \includegraphics[width=\linewidth]{Article_figures/zmax_determination.pdf}
-    \caption{$z\St{max}$ determination from the interpolated poissonian
-    evolutions.}
-    \label{fig:zdet}
+    \subfigure[Method]{\label{fig:zdet}
+        \includegraphics[width=\linewidth]{Article_figures/zmax_determination.pdf}}
+    \subfigure[Result]{\label{fig:surveys_cuts}
+    \includegraphics[width=\linewidth]{Article_figures/surveys_cuts_x.pdf}}
+    \caption{(a) $z\St{max}$ determination from the interpolated poissonian
+    evolutions; (b) SDSS, SNLS and PS1 samples cut at $z\St{max}$ as explained in
+    section \ref{sec:sam}. The data we use is in plain markers.}
 \end{figure}
+
+\section{Method}
+
+We used the LsSFR and stretch measurments from the SNf sample. The LsSFR has an
+evolution with the redshift that is analytically known: calling $\de\z$ the
+fraction of young stars and $\p\z$ the fraction of old ones, Rigault 2018 et +
+find
+\begin{equation}
+    \frac{\de\z}{\p\z} \equiv LsSFR\z = K\times \LF{1+z}^\f
+\end{equation}
+with $\f = 2.8$, and knowing $\de\z + \p\z = 1$:
+\begin{align}
+    \de\z & = \LF{K^{-1}\times \LF{1+z}^{-\f} +1}^{-1}\\
+    \p\z & = \LF{K\times \LF{1+z}^{+\f} + 1}\um
+\end{align}
+
+The goal is to find how the stretch depends on the LsSFR, distinguishing old and
+young SNe which fractions evolve with the redshift, in order to have an
+analytical law for the mean redshift evolution of the stretch. The measured data
+on which we based our first model is plotted in figure \ref{fig:mod_first};
+based of the shape of the $x_1$ vs LsSFR cloud, we implemented the following
+model:
+\begin{itemize}
+    \item \textbf{young}: a gaussian of mean $\m_1$ and standard deviation
+        $\s_1$, namely $\Nc_1 \equiv \Nc \LF{\m_1,\s_1}$;
+    \item \textbf{old}: a linear combination between $\Nc_1$ and another
+        gaussian $\Nc_2 \equiv \Nc \LF{\m_2,\s_2}$
+\end{itemize}
+As such, the probability to observe a young SNe Ia labeled "i" with a stretch
+$x_1^i$ and error $\d x_1^i$ is
+\begin{equation}
+    p \LF{x_1^i, \d x_1{}^i|\m_1,\s_1} = \Nc \LF{\m_1, \Sq{\s_1{}^2 + \d
+    x_1^i{}^2}}(x_1^i)
+\end{equation}
+and for an old one:
+\begin{align}
+    p(x_1^i, \d x_1^i | \m_{1,2}, \s_{1,2}, a) = a & \times \Nc \LF{\m_1,
+    \sqrt{\s_1{}^2 + \d x_1^i{}^{2}}} (x_1^i) \ + \\ (1-a) & \times \Nc
+    \LF{\m_2, \sqrt{ \s_2{}^2 + \d x_1^i{}^{2}}} (x_1^i), \nonumber 
+\end{align}
+We fit different model, and compare their AICc.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{Article_figures/surveys_cuts_x.pdf}
-    \caption{SDSS, SNLS and PS1 samples cut at $z\St{max}$ as explained in
-    section \ref{sec:sam}. The data we use is in plain markers.}
-    \label{fig:surveys_cuts}
+    \subfigure[Raw data]{\label{fig:snf_data}
+        \includegraphics[width=\linewidth]{Article_figures/BiGaussian_onlydata.pdf}}
+    \subfigure[First implemented model]{\label{fig:3G2M2SSNF}
+        \includegraphics[width=\linewidth]{Article_figures/BiGaussian.pdf}}
+    \caption{Stretch des supernovae étudiées par la collaboration SNF en
+    fonction de log(LsSFR). La couleur représente la probabilité pour une
+supernova d'être issue d'un jeune progéniteur.}
+    \label{fig:mod_first}
 \end{figure}
-
-\section{Method}
-We used LsSFR measurments which evolution with the redshift is analytically
-known (references) and stretch measurments made by SNf to try an correlate these
-to parameters. We fit different model, and compare their AICc.
 

### 2019-10-21 Nicolas.tex
Results. Need to add table pages. Preambule shortened.

@@
 The simple use of those is not enough for our aim, as SNe Ia have an intrinsec
-dispersion of their luminosity of $\approx \SI{0.4}{mag}$ that gives a huge
+dispersion of their luminosity of $\approx 0.4~mag$ that gives a huge
 uncertainty on the determination of their distance modulus. Henceforth they are
@@
 fits the distance modulus which is expressed as
-\begin{center}
-    $ \displaystyle \m = m_b + \a x_1 - \b c - M$
-\end{center}
+\begin{equation}
+    \m = m_b + \a x_1 - \b c - M
+\end{equation}
 with $m_B$ the logarithm of their flux, $x_1$ their stretch, $c$ their color and
 $M$ their intrinsec magnitude. This relation lowered their magnitude dispersion
-to $\approx \SI{0.15}{mag}$, allowing for previously mentioned accelerated
+to $\approx 0.15~mag$, allowing for previously mentioned accelerated
 expansion to be discovered.
@@
 histograms of the surveys, and defining a function of the expected SNe Ia rate:
-\begin{center}
-    $ \displaystyle N\St{SNe Ia} = a \times V\z$
-\end{center}
+\begin{equation}
+    N\St{SNe Ia} = a \times V\z
+\end{equation}
 with $a$ the density of SNe Ia and $V\z$ the volume of Universe in each bin.
@@
 compute the poissonian probability in each bin, given by
-\begin{center}
-    $ \displaystyle p(k) = \frac{\l^k}{k!}e^{-\l}$
-\end{center}
+\begin{equation}
+    p(k) = \frac{\l^k}{k!}e^{-\l}
+\end{equation}
 with $k$ the expected counts, $\l$ the observed counts. This work is done 100
@@
 \begin{equation}
-    \frac{\de\z}{\p\z} \equiv LsSFR\z = K\times \LF{1+z}^\f
+    \frac{\de\z}{\p\z} \equiv \mathrm{LsSFR}\z = K\times \LF{1+z}^\f
 \end{equation}
@@
 \end{align}
+
+\begin{figure}[htbp!]
+    \centering
+    \subfigure[Raw data]{\label{fig:snf_data}
+        \includegraphics[width=\linewidth]{Article_figures/BiGaussian_onlydata.pdf}}
+    \subfigure[First implemented model]{\label{fig:3G2M2SSNF}
+        \includegraphics[width=\linewidth]{Article_figures/BiGaussian.pdf}}
+    \caption{Stretch des supernovae étudiées par la collaboration SNF en
+    fonction de log(LsSFR). La couleur représente la probabilité pour une
+supernova d'être issue d'un jeune progéniteur.}
+    \label{fig:mod_first}
+\end{figure}
 
@@
 \end{align}
-We fit different model, and compare their AICc.
 
-\begin{figure}[htbp!]
+\noindent where $a$ is the relative amplitude between the two gaussians.
+Finally, the normalized stretch distribution at a given $z$ $\D\LF{x_1|z}$ is
+the weigthed sum of both young and old stretch distributions given their
+relative fraction $\de\z$:
+\begin{equation}
+    \D\LF{x_1|z} = \de\z\times\Nc_1 +
+    \LF{1-\de\z} \times \LF{a\Nc_1 + (1-a)\Nc_2}
+\end{equation}
+
+We fitted it on SNf data, giving the results table \ref{tab:val}. For clarity
+with the next models, we named it 3G2M2S$\St{SNf}$ for it has a total of 3
+gaussians but with only 2 means and 2 standard deviations, and has been fitted
+on SNf data only.
+
+We implemented and compared 10 models in total, 4 of which have an
+evolution with the redshift from $\de\z$, and 6 don't ($\de\z = f =
+\mathrm{cst}$). The ones with and evolution are:
+\begin{itemize}
+    \item 3G2M2S, the one we described (but fitted on all the data);
+    \item 3G2M1S, where this time $\s_1 \equiv \s_2$;
+    \item 2G2M2S, model taken from HOWELL 2009 where we added $\de\z$;
+    \item 3G3M3S, with three independent gaussians.
+\end{itemize}
+The ones without a stretch evolution are the same ones but with an "F" implying
+we set $\de\z = f = \mathrm{cst}$, and two others:
+\begin{itemize}
+    \item 1G1M1S, where there is no distinction between old and young SNe;
+    \item 1G1M2S, taken from KESLLER 2017 and used in recent cosmological
+        analysis SCOLNIC 2018. It's an asymetric model where
+\end{itemize}
+
+\begin{align}
+ p(x_1^i, \d x_1^i | \m, \s_-, \s_+) = 
+    \begin{cases}
+        \mathcal{N} \LF{\m, \sqrt{\s_-{}^2 + \d x_1^i{}^{2}}} (x_1^i) & \text{if
+        } x_1^i\geq \m\\
+        \mathcal{N} \LF{\m, \sqrt{\s_+{}^2 + \d x_1^i{}^{2}}} (x_1^i), &
+        \text{else}
+    \end{cases}
+\end{align}
+
+The fitted parameters are showed table \ref{tab:val}.
+
+\begin{table}[htbp!]
     \centering
-    \subfigure[Raw data]{\label{fig:snf_data}
-        \includegraphics[width=\linewidth]{Article_figures/BiGaussian_onlydata.pdf}}
-    \subfigure[First implemented model]{\label{fig:3G2M2SSNF}
-        \includegraphics[width=\linewidth]{Article_figures/BiGaussian.pdf}}
-    \caption{Stretch des supernovae étudiées par la collaboration SNF en
-    fonction de log(LsSFR). La couleur représente la probabilité pour une
-supernova d'être issue d'un jeune progéniteur.}
-    \label{fig:mod_first}
-\end{figure}
+    \caption{Valeurs des paramètres pour différents modèles. En rouge les
+    données aberrantes.}
+    \label{tab:val}
+    \begin{tabularx}{\linewidth}{|Y*{6}{|Y}|}\hline
+
+        \rowcolor{gray!15} Modèle & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ &
+        $\s_2$ \\\hline\hline
+
+        3G2M2S$_{\mathrm{SNf}}$ & $0.48 \pm 0.07$ & none & $0.39 \pm 0.07$ &
+        $0.56 \pm 0.05$ & $-1.5 \pm 0.1$ & $0.52 \pm 0.09$ \\\hline
+
+        3G2M2S & $0.48 \pm 0.17$ & none & $0.36 \pm 0.08$ & $0.61 \pm 0.05$ &
+        $-1.3 \pm 0.2$ & $0.60 \pm 0.12$ \\\hline
+
+        3G2M2SF & \textcolor{red}{$0.1 \pm 0.6$} & \textcolor{red}{$0.2 \pm
+        0.6$} & $-0.9 \pm 0.7$ & $0.7 \pm 0.3$ & $0.5 \pm 0.2$ & $0.6 \pm 0.1$
+        \\\hline
+
+        3G2M1S & $0.47 \pm 0.07$ & none & $0.35 \pm 0.04$ & $0.61 \pm 0.03$ &
+        $-1.25 \pm 0.10$ & $\s_1$ \\\hline
+
+        3G2M1SF & \textcolor{red}{$0.2 \pm 0.9$} & $0.7 \pm 0.3$ & $0.36 \pm
+        0.04$ & $0.60 \pm 0.03$ & $-1.23 \pm 0.10$ & $\s_1$ \\\hline
+
+        2G2M2S & none & none & $0.49 \pm 0.04$ & $0.54 \pm 0.03$ & $-0.72 \pm
+        0.08$ & $0.83 \pm 0.07$ \\\hline
+
+        2G2M2SF & none & $0.3 \pm 0.2$ & $-0.9 \pm 0.6$ & $0.7 \pm 0.2$ & $0.5
+        \pm 0.2$ & $0.56 \pm 0.09$ \\\hline
+
+    \end{tabularx} \bigbreak
+
+\begin{tabularx}{\linewidth}{|Y*{2}{|Y}|}\hline
+
+    \rowcolor{gray!15} Modèle & $\m$ & $\s$ \\\hline\hline
+
+    1G1M1S & $0.01 \pm 0.04$ & $0.90 \pm 0.03$ \\\hline
+
+\end{tabularx} \bigbreak
+
+\begin{tabularx}{\linewidth}{|Y*{3}{|Y}|}\hline
+
+    \rowcolor{gray!15} Modèle & $\m$ & $\s_-$ & $\s_+$ \\\hline\hline
+
+    1G1M2S & $0.16617 \pm 0.00004$ & $1.07 \pm 0.04$ & $0.69 \pm 0.03$ \\\hline
+
+\end{tabularx} \bigbreak
+
+\begin{tabularx}{\linewidth}{|Y*{8}{|Y}|}\hline
+
+    \rowcolor{gray!15} Modèle & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ & $\s_2$ &
+    $m_3$ & $\s_3$ \\\hline\hline
+
+    3G3M3S & $0.14 \pm 0.08$ & none & $0.51 \pm 0.06$ & $0.54 \pm 0.04$ & $-1.9
+    \pm 0.2$ & $0.29 \pm 0.11$ & $-0.55 \pm 0.12$ & $0.67 \pm 0.15$ \\\hline
+
+    3G3M3SF & $0.2 \pm 0.2 $ & $0.10 \pm 0.04 $ & $-1.7 \pm 0.2$ & $0.4 \pm 0.1$
+            & $0.9 \pm 0.1$ & $0.3 \pm 0.2$ & $0.0 \pm 0.2$ & $0.7 \pm 0.1$
+            \\\hline
+
+\end{tabularx} \bigbreak
+
+\end{table}
 
 \section{Results}
+To compare all these models, we use the Akaike Information Criterion corrected
+for sample size (BURNHAM 2002), that penalises the increase of free parameters
+in order to discourage overfitting:
+\begin{equation}
+    \mathrm{AICc} = \mathrm{AIC} + \LR{2k\LF{k+1}}{n - k - 1}
+\end{equation}
+with $\mathrm{AIC} = 2k - 2\ln\LF{\mathcal{L}}$, where $k$ is the number of free
+parameters and $\Lc$ the likelihood. The probability for a model to be as
+representative as the "best" one is given by:
+\begin{equation}
+    p(\mathrm{other} > \mathrm{best}) = \exp\LF{\D\mathrm{AICc}/2}
+\end{equation}
+As such, we obtain the results table \ref{tab:comp}.
+
+\begin{table}[htbp!]
+    \centering
+    \caption{Comparaison des modèles. NR représente les modèles implémentés
+             durant ce stage. (F) indique les modèles pour lesquels il n'y a pas
+             d'évolution de la fraction de SNe~Ia jeunes et vieilles en fonction
+             du redshift.}
+    \label{tab:comp}
+    \begin{tabularx}{\linewidth}{|Y*{6}{|Y}|}\hline
+
+        \rowcolor{gray!15} Name & Description & Free param &
+        $\ln\mathcal{L}$ & $AICc$ & $\D AICc$ & Proba \\\hline\hline
+
+        3G2M1S & NR 1S & 4 & 1815 & 1823 & 0.0 & $1.0$ \\\hline
+
+        3G2M2S & NR 2S & 5 & 1815  & 1825 & -2.0 & $3.6\times10^{-1}$ \\\hline
+
+        2G2M2S & Howell & 4 & 1818  & 1826 & -3.4 & $1.8\times10^{-1}$ \\\hline
+
+        3G3M3S & NR 3S & 7 & 1812  & 1826 & -3.6 & $1.6\times10^{-1}$ \\\hline
+
+        3G3M3SF & NR 3S (F) & 8 & 1813 & 1829 & -6.3  & $4.3\times10^{-2}$
+        \\\hline
+
+        2G2M2SF & Howell (F) & 5 & 1823  & 1833 & -9.9 & $7.0\times10^{-3}$
+        \\\hline
+
+        3G2M1SF & NR 1S (F) & 5 & 1823  & 1833 & -10.4 & $5.5\times10^{-3}$
+        \\\hline
+
+        3G2M2SF & NR 2S (F) & 6 & 1823  & 1835 & -12.0 & $2.5\times10^{-3}$
+        \\\hline
+
+        1G1M2S & Kessler (F) & 3 & 1837  & 1843 & -20.2 & $4.1\times10^{-5}$
+        \\\hline
+
+        1G1M1S & 1 gauss. (F) & 2 & 1872 & 1876 & -53.5 &
+        $2.4\times10^{-12}$\\\hline
+
+    \end{tabularx}
+\end{table}
+
 We find that every model lacking an evolution of the stretch with the redshift

### 2019-10-31 Nicolas.tex
Little changes

@@
 \usepackage{preambule}
+\usepackage{graphicx}
 \usepackage[varg]{txfonts}
@@
 \section{Introduction}
+
+\begin{figure*}[htbp!]
+    \centering
+    \subfigure[Method]{\label{fig:zdet}
+        \includegraphics[width=8.5cm]{Article_figures/zmax_determination.pdf}}
+    \subfigure[Result]{\label{fig:surveys_cuts}
+    \includegraphics[width=8.5cm]{Article_figures/surveys_cuts_x.pdf}}
+    \caption{(a) $z\St{max}$ determination from the interpolated poissonian
+    evolutions; (b) SDSS, SNLS and PS1 samples cut at $z\St{max}$ as explained in
+    section \ref{sec:sam}. The data we use is in plain markers.}
+\end{figure*}
 
@@
     \label{tab:sample}
-    \begin{tabularx}{\linewidth}{|Y*{1}{|Y}|}\hline
-        \rowcolor{gray!15} Survey & Number of SNe Ia \\\hline\hline
-        SNf & 141 \\\hline
-        PS1 & 178 \\\hline
-        SDSS & 206 \\\hline
-        SNLS & 138 \\\hline
+    \begin{tabular}{c c}\hline
+        \rowcolor{gray!15} Survey & Number of SNe Ia \\\hline
+        SNf & 141 \\
+        PS1 & 178 \\
+        SDSS & 206 \\
+        SNLS & 138 \\
         HST & 26 \\\hline
-    \end{tabularx}
+    \end{tabular}
 \end{table}
-
-\begin{figure}[htbp!]
-    \centering
-    \subfigure[Method]{\label{fig:zdet}
-        \includegraphics[width=\linewidth]{Article_figures/zmax_determination.pdf}}
-    \subfigure[Result]{\label{fig:surveys_cuts}
-    \includegraphics[width=\linewidth]{Article_figures/surveys_cuts_x.pdf}}
-    \caption{(a) $z\St{max}$ determination from the interpolated poissonian
-    evolutions; (b) SDSS, SNLS and PS1 samples cut at $z\St{max}$ as explained in
-    section \ref{sec:sam}. The data we use is in plain markers.}
-\end{figure}
 
@@
 
-\begin{figure}[htbp!]
+\begin{figure*}[htbp!]
     \centering
     \subfigure[Raw data]{\label{fig:snf_data}
-        \includegraphics[width=\linewidth]{Article_figures/BiGaussian_onlydata.pdf}}
+        \includegraphics[width=8.5cm]{Article_figures/BiGaussian_onlydata.pdf}}
     \subfigure[First implemented model]{\label{fig:3G2M2SSNF}
-        \includegraphics[width=\linewidth]{Article_figures/BiGaussian.pdf}}
+        \includegraphics[width=8.5cm]{Article_figures/BiGaussian.pdf}}
     \caption{Stretch des supernovae étudiées par la collaboration SNF en
@@
     \label{fig:mod_first}
-\end{figure}
+\end{figure*}
 
@@
 \end{itemize}
-
 \begin{align}
@@
 
-\begin{table}[htbp!]
+\begin{table*}[htbp!]
     \centering
@@
     \label{tab:val}
-    \begin{tabularx}{\linewidth}{|Y*{6}{|Y}|}\hline
+    \begin{tabular}{c c c c c c c}\hline\hline
 
         \rowcolor{gray!15} Modèle & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ &
-        $\s_2$ \\\hline\hline
+        $\s_2$ \\\hline
 
         3G2M2S$_{\mathrm{SNf}}$ & $0.48 \pm 0.07$ & none & $0.39 \pm 0.07$ &
-        $0.56 \pm 0.05$ & $-1.5 \pm 0.1$ & $0.52 \pm 0.09$ \\\hline
+        $0.56 \pm 0.05$ & $-1.5 \pm 0.1$ & $0.52 \pm 0.09$ \\
 
         3G2M2S & $0.48 \pm 0.17$ & none & $0.36 \pm 0.08$ & $0.61 \pm 0.05$ &
-        $-1.3 \pm 0.2$ & $0.60 \pm 0.12$ \\\hline
+        $-1.3 \pm 0.2$ & $0.60 \pm 0.12$ \\
 
@@
         0.6$} & $-0.9 \pm 0.7$ & $0.7 \pm 0.3$ & $0.5 \pm 0.2$ & $0.6 \pm 0.1$
-        \\\hline
+        \\
 
         3G2M1S & $0.47 \pm 0.07$ & none & $0.35 \pm 0.04$ & $0.61 \pm 0.03$ &
-        $-1.25 \pm 0.10$ & $\s_1$ \\\hline
+        $-1.25 \pm 0.10$ & $\s_1$ \\
 
         3G2M1SF & \textcolor{red}{$0.2 \pm 0.9$} & $0.7 \pm 0.3$ & $0.36 \pm
-        0.04$ & $0.60 \pm 0.03$ & $-1.23 \pm 0.10$ & $\s_1$ \\\hline
+        0.04$ & $0.60 \pm 0.03$ & $-1.23 \pm 0.10$ & $\s_1$ \\
 
         2G2M2S & none & none & $0.49 \pm 0.04$ & $0.54 \pm 0.03$ & $-0.72 \pm
-        0.08$ & $0.83 \pm 0.07$ \\\hline
+        0.08$ & $0.83 \pm 0.07$ \\
 
@@
 
-    \end{tabularx} \bigbreak
+    \end{tabular} \bigbreak
 
-\begin{tabularx}{\linewidth}{|Y*{2}{|Y}|}\hline
+\begin{tabular}{c c c}\hline\hline
 
-    \rowcolor{gray!15} Modèle & $\m$ & $\s$ \\\hline\hline
+    \rowcolor{gray!15} Modèle & $\m$ & $\s$ \\\hline
 
@@
 
-\end{tabularx} \bigbreak
+\end{tabular} \bigbreak
 
-\begin{tabularx}{\linewidth}{|Y*{3}{|Y}|}\hline
+\begin{tabular}{c c c c c}\hline\hline
 
-    \rowcolor{gray!15} Modèle & $\m$ & $\s_-$ & $\s_+$ \\\hline\hline
+    \rowcolor{gray!15} Modèle & $\m$ & $\s_-$ & $\s_+$ \\\hline
 
@@
 
-\end{tabularx} \bigbreak
+\end{tabular} \bigbreak
 
-\begin{tabularx}{\linewidth}{|Y*{8}{|Y}|}\hline
+\begin{tabular}{c c c c c c c c c}\hline\hline
 
     \rowcolor{gray!15} Modèle & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ & $\s_2$ &
-    $m_3$ & $\s_3$ \\\hline\hline
+    $m_3$ & $\s_3$ \\\hline
 
     3G3M3S & $0.14 \pm 0.08$ & none & $0.51 \pm 0.06$ & $0.54 \pm 0.04$ & $-1.9
-    \pm 0.2$ & $0.29 \pm 0.11$ & $-0.55 \pm 0.12$ & $0.67 \pm 0.15$ \\\hline
+    \pm 0.2$ & $0.29 \pm 0.11$ & $-0.55 \pm 0.12$ & $0.67 \pm 0.15$ \\
 
@@
 
-\end{tabularx} \bigbreak
+\end{tabular} \bigbreak
 
-\end{table}
+\end{table*}
 
@@
 \end{equation}
-As such, we obtain the results table \ref{tab:comp}.
+The results are showed table \ref{tab:comp}. We find that every model lacking an
+evolution of the stretch with the redshift is systematically worse than those
+that implement it.
 
-\begin{table}[htbp!]
+\begin{table*}[htbp!]
     \centering
@@
     \label{tab:comp}
-    \begin{tabularx}{\linewidth}{|Y*{6}{|Y}|}\hline
+    \begin{tabular}{c c c c c c c}\hline\hline
 
         \rowcolor{gray!15} Name & Description & Free param &
-        $\ln\mathcal{L}$ & $AICc$ & $\D AICc$ & Proba \\\hline\hline
+        $\ln\mathcal{L}$ & $AICc$ & $\D AICc$ & Proba \\\hline
 
-        3G2M1S & NR 1S & 4 & 1815 & 1823 & 0.0 & $1.0$ \\\hline
+        3G2M1S & NR 1S & 4 & 1815 & 1823 & 0.0 & $1.0$ \\
 
-        3G2M2S & NR 2S & 5 & 1815  & 1825 & -2.0 & $3.6\times10^{-1}$ \\\hline
+        3G2M2S & NR 2S & 5 & 1815  & 1825 & -2.0 & $3.6\times10^{-1}$ \\
 
-        2G2M2S & Howell & 4 & 1818  & 1826 & -3.4 & $1.8\times10^{-1}$ \\\hline
+        2G2M2S & Howell & 4 & 1818  & 1826 & -3.4 & $1.8\times10^{-1}$ \\
 
-        3G3M3S & NR 3S & 7 & 1812  & 1826 & -3.6 & $1.6\times10^{-1}$ \\\hline
+        3G3M3S & NR 3S & 7 & 1812  & 1826 & -3.6 & $1.6\times10^{-1}$ \\
 
         3G3M3SF & NR 3S (F) & 8 & 1813 & 1829 & -6.3  & $4.3\times10^{-2}$
-        \\\hline
+        \\
 
         2G2M2SF & Howell (F) & 5 & 1823  & 1833 & -9.9 & $7.0\times10^{-3}$
-        \\\hline
+        \\
 
         3G2M1SF & NR 1S (F) & 5 & 1823  & 1833 & -10.4 & $5.5\times10^{-3}$
-        \\\hline
+        \\
 
         3G2M2SF & NR 2S (F) & 6 & 1823  & 1835 & -12.0 & $2.5\times10^{-3}$
-        \\\hline
+        \\
 
         1G1M2S & Kessler (F) & 3 & 1837  & 1843 & -20.2 & $4.1\times10^{-5}$
-        \\\hline
+        \\
 
@@
 
-    \end{tabularx}
-\end{table}
-
-We find that every model lacking an evolution of the stretch with the redshift
-is systematically worse than those that implement it.
+    \end{tabular}
+\end{table*}
 

### 2019-11-21 Nicolas.tex
First draft of mag-limited zmax

@@
 fits the distance modulus which is expressed as
-\begin{equation}
+\begin{equation}\label{eq:tripp}
     \m = m_b + \a x_1 - \b c - M
@@
 relations: the first SNe Ia that are missed by the instruments are the redder
-and faster ones. That implies that taking all of these surveys' data would not
-represent the stretch distribution that Nature could give us.
+and faster-declining ones. That implies that taking all of these surveys' data
+would not represent the stretch distribution that Nature could give us.
+
+We used the 5$\sigma$ depth of each survey to know what was the smallest
+apparent magnitude they could acquire. We want to find the maximum redshift
+(or equivalently distance modulus assuming Planck15 cosmology) for which all
+combination of stretch and color are visible. We computed the absolute magnitude
+5 days prior peak (taken at $-19.3\,$mag) for each value of
+$(x_1, c)$ in an $([-3, 3], [-0.3, 0.3])$ ellipse, finding which combination had
+the smallest one, using a SALT2 model in SNCosmo. Using $\a$ and $\b$ given in
+Scolnic 2018a, we can then compute the distance modulus corresponding to that
+minimum absolute magnitude and verifying \eqref{eq:tripp}.
 

### 2019-11-27 Nicolas.tex
Finished zmax inputs

--- 2019-10-31 Nicolas.tex
@@
-\documentclass{aa}
-\usepackage{preambule}
-\usepackage{graphicx}
-\usepackage[varg]{txfonts}
+\documentclass[bibnumber]{aa}
+
+\usepackage{hyperref}
+
+\usepackage{color}
+\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
+\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
 
@@
     \and M. Rigault\inst{2}
+    \and R. Graziani\inst{2}     
+    \and M. Briday\inst{1}
+    \and Y. Copin\inst{1}    
+    \and Y. Kim\inst{1}
 }
@@
 
-\begin{figure*}[htbp!]
-    \centering
-    \subfigure[Method]{\label{fig:zdet}
-        \includegraphics[width=8.5cm]{Article_figures/zmax_determination.pdf}}
-    \subfigure[Result]{\label{fig:surveys_cuts}
-    \includegraphics[width=8.5cm]{Article_figures/surveys_cuts_x.pdf}}
-    \caption{(a) $z\St{max}$ determination from the interpolated poissonian
-    evolutions; (b) SDSS, SNLS and PS1 samples cut at $z\St{max}$ as explained in
-    section \ref{sec:sam}. The data we use is in plain markers.}
-\end{figure*}
+%\begin{figure*}
+%    \centering
+%    \subfigure[Method]{\label{fig:zdet}
+%        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}
+%    \subfigure[Result]{\label{fig:surveys_cuts}
+%    \includegraphics[width=8.5cm]{A    \subfigure[Method]{\label{fig:zdet}
+%rticle_figures/surveys_cuts_x.pdf}}
+%    \subfigure[Method]{\label{fig:zdet}
 
-Type Ia supernovae (SNe Ia) are now well-known for their capacity to determine
-cosmological parameters: their study led to the discovery of the accelerated
-expansion of the Universe (Riess 98, Perlmutter 99) through the name of "dark
-energy", and they have been used continuously for better measurments since then
-(Betoule 2014). They are acquired through their lightcurves, giving the
-evolution of their luminosity from the time of explosion, in different
-wavelength. 3 parameters are used to describe those: an amplitude, a width
-(named "stretch") and a color (magnitude difference in the B and V bands).
+%    \caption{(a) $z\St{max}$ determination from the magnitude limited data; (b) SDSS, %SNLS and PS1 samples cut at $z\St{max}$ as explained in
+%    section \ref{sec:sam}. The data we use is in plain markers.}
+%\end{figure*}
 
-The simple use of those is not enough for our aim, as SNe Ia have an intrinsec
-dispersion of their luminosity of $\approx 0.4~mag$ that gives a huge
-uncertainty on the determination of their distance modulus. Henceforth they are
-standardized using the "brigther-slower" and "brighter-bluer" relations
-(Philipps 93, Riess 96, Tripp 98) in the SALT2 algorithm (Guy 2007, 2010) that
-fits the distance modulus which is expressed as
-\begin{equation}
-    \m = m_b + \a x_1 - \b c - M
-\end{equation}
-with $m_B$ the logarithm of their flux, $x_1$ their stretch, $c$ their color and
-$M$ their intrinsec magnitude. This relation lowered their magnitude dispersion
-to $\approx 0.15~mag$, allowing for previously mentioned accelerated
-expansion to be discovered.
+Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
+have enabled the discovery of the acceleration of the Universe's expansion
+\citep{riess1998, permutter1999}. They remain today a key comological probe to
+understand the properties of dark energy (DE) as it is the only tool able to
+precisely map the recent expansion rate $z<0.5$, when DE is driving it
+\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
+Hubble Constant (H$_0$) provided one can calibrate their absolute magnitude
+\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
+the SNeIa are anchored on Cepheids \citep[the SH0ES
+project][]{riess2009,riess2016} is $\sim5~\sigma$ higher than what is predicted
+by cosmic microwave background (CMB) data measured by Planck assuming the
+standard $\L$CDM or when the SN luminosity is anchored at intermediate redshift
+by the baryon accoustic oscillation (BAO) scale \citep{riess2019, ried2019,
+planck, feeney}. While using the tip of the red giant branch technique in place
+of the Cepheids seem to favor lower values of H$_0$ \citep{freedman,
+recalibrationpaper} time delay measurements from strong lensing seem to also
+favor high H$_0$ values \citep{wong2019}.
 
-This Tripp estimator lies on the idea that this standardization doesn't change
-with the redshift. However, Rigault 2015 showed that SNe Ia depend on their
-environments, and these environments' properties evolve with the redshift. In
-this Letter, we try to determine whether a stretch evolution with the redshift
-allows for a better description of the collected data.
+The H$_0$ tension has received a lot of attention as it could be a sign of new
+fundamental physics. Yet, no simple solution is able to accomodate this H$_0$
+tension when accounting for all other probes and the current most promising
+scenario appears to be a burst of expansion at the matter-radiation decoupling
+moment caused by a (fine tuned) early dark energy \citep{poulin}.
+
+Alternatively, systematic effects affecting one or several of the aforementioned
+analysis could also explain the tension. In \cite{rigault2015} we suggested that
+SNe Ia from the cepheid calibrator sample differ significantly to the one from
+the Hubble flow sample that enables to derive H$_0$. Indeed, Cepheid are very
+young stars and the existence of Cepheids in a galaxy implies that the host is
+star forming. In \citep{Rigault2015} we claimed that a more than 90\% of the
+calibrator SNe Ia arise from a young-progenitor propulation while this
+population only accounts for about half of the Hubble flow sample. Following
+\cite{sullivan2010} and \citep{rigault2013}, we further showed that young and
+old progenitor SNe Ia have a different magnitude of about 0.15 mag. This has
+been confirmed for modern standardization technique
+\citep[SALT2.4][]{guy2010,betoule2014} with high significance ($\sim 6\sigma$)
+in \citep{rigault2018} as well as independently by \cite{roman2019} at $\sim
+7\sigma$. As detailed in \citep{Rigault2015}, if the magnitude difference
+between young and old SNe Ia is indeed of $0.15$ mag and if SNe Ia from the
+calibrator sample is fully dominated by young SNeIa, the effect on H$_0$ of the
+existance of these two SNe Ia populations is of 3.5\%. Accounting for the fact
+that another environmental effect (the mass-step) is already included in the
+SH0ES analysis, the net bias in the SH0ES analysis is expected closer to 3\%
+\citep[see table 6 of ][]{rigault2015}.
+
+SHOULD BE SHORTENED A LOT: However, the importance of this astrophysical bias in
+the direct measurement of H$_0$ is still highly debated. We also highlight that
+it could not explain the high value of H$_0$ derived by strong lensing. First,
+\cite{jones2015} did not find any environmental bias when extending the
+\citep{rigault2015} study. It was more recently reported in \citep{jones2019}
+that indeed local host analysis has significant influence on the SN magnitude
+but much more importantly than claimed in \citep{rigault2018} and with a
+connection with the progenitor physics and/or a line of sight effect that is
+unclear. More importantly, \cite{riess2016} attended to mimic the cepheid sample
+selection function on the Hubble flow sample by removing SNe Ia from the latter
+if they classificed their host as early-type or non-star forming. They measured
+that doing so has no effect on the measurement of H$_0$.
+
+% SERIES OF PAPER
+This paper is part of a series of papers where we test the validity of the two
+SNe Ia population models developped in \citep{rigault2018}.Briday et al. in prep
+will present how using different environmental tracer is affecting our ability
+to distinguish the two populations, and Rigault et al. in prep will further
+study the connection with H$_0$. Here we analyze predictions concerning the
+SNe Ia redshift drifts: (1) that the distribution of the SNIa stretch, a purely
+intrinsic SNIa property, is evolving as a function of redshift and (2) that this
+population-drift follows the relation that \citep{rigault2018} derived assuming
+that one population is “young“ with a rate proportional to the star formation
+rate in the Universe and the second is "old" with a rate proportional to the
+stellar mass.
+
+The concept of the SNeIa age dichotomy arised when the SNIa rate has first been
+studied. \cite{Mannucci et al. 2005; Scannapieco & Bildsten 2005; Sullivan et
+al. 2006; Aubourg et al. 2008} have shown that the relative SNe Ia rate in
+galaxies could only be explained if two populations existed, one young,
+following the host star formation activity, and one old following the host
+stellar mass (the so called “prompt and delayed“ or “A+B“ model). In
+\cite{rigault2018} we used the specific star formation rate at the SN location
+(Local sSFR or LsSFR) to classify which are the prompts (those with a high
+LsSFR) and which are the delayed (those with low LsSFR). Since the first SNe Ia
+host analysis, the SN strech has been known to be strongly correlated with the
+SN host properties \citep{Hamuy1996, Hamuy2000} and it has been extensively
+confirmed since \citep[e.g.][]{Neill2009,Sullivan2010; Lampeitl2010; Kelly2010;
+Gupta2011,D’Andrea2011; Childress2013,Rigault2013,Pan2014}. Following the “A+B“
+model and the connection between SN stretch and host properties,
+\citep{Howell20XX} first discussed the potential redshift drift of the SN strech
+distribution. In this paper we revisit this analysis with the most recent SNe Ia
+dataset and we test the validity of the two SNIa age populations model
+developped by \citep{rigault2018}.
+
+We present in section~\ref{sec:data} the sample we are using for this analysis,
+which is based on the pantheon dataset \citep{scolnic2018a}. We discuss the
+importance of obtaining a "complete" sample, i.e. reprensentative of the true
+underlying SNe Ia distribution, and how we build one from the Pantheon sample.
+We then present in section~\ref{sec:model} our modeling of the distribution of
+stretch as a function of redshift based on the relative stretch distrubtion of
+young and old SNeIa. Our results are presented in section~\ref{sec:result} where
+we test if the distribution of stretch does evolve as a function of redshift and
+if the age-model is in good agreement with this evolution. We discuss our
+results and conclude in section~\ref{sec:conclusion}.
+
+Throughout the paper we use the Planck 2015 cosmology from the astropy library.
 
@@
 
-We use data from 5 different surveys: Hubble Space Telescope (HST, REF),
-Supernova Nearby Factory (SNf, REF), Supernova Legacy Survey (SNLS, Aster 2006,
-Betoule 2014), Sloan DSS (SDSS, REF) and Panstarr-1 (PS1, REF). The last 3 had
-selection effects that come from the "brighter-slower" and "brighter-bluer"
-relations: the first SNe Ia that are missed by the instruments are the redder
-and faster ones. That implies that taking all of these surveys' data would not
-represent the stretch distribution that Nature could give us.
+We base our analysis on the latest SNe Ia compilation: the Pantheon catalog from
+\cite{scolnic2018a}. Using this dataset, we aim at probing the evolution of the
+distribution of SNe Ia stretch in nature as a function of redshift. Because the
+observed magnitude of a SN Ia is correlated with its stretch and its color, in a
+magnitude limited survey such as those consituing most of the Pantheon sample,
+the distribution of these lightcurve parameters is not representative of the
+true underlying distribution above a certain limiting redshift. The first SNe Ia
+that such a survey will miss because they will be too faint are indeed the
+reddest and the lowest-stretch SNe Ia. If such a selection effect is not
+accounted for, we might confuse true population drift we selection function or
+conversally.
 
-As such, we determined the maximum redshift at which we would cut these samples
-to get complete ones. We used a statistical approach in lack of precise data
-concerning the instruments' capacity to acquire fluxes. We proceeded by making
-histograms of the surveys, and defining a function of the expected SNe Ia rate:
-\begin{equation}
-    N\St{SNe Ia} = a \times V\z
-\end{equation}
-with $a$ the density of SNe Ia and $V\z$ the volume of Universe in each bin.
-Before selection effects, the histograms are supposed to grow proportionnaly to
-$V\z$ assuming an homogeneous repartition of SNe Ia. When we start to miss some
-supernovae, this growth is not followed anymore.
+The figure~\ref{fig:maglim} shows the distribution of stretch for the magnitude
+limited samples of the Pantheon data, i.e. PanStarrs \citep[PS1][]{ref}, the
+Sloan Digital Sky Survey \citep[SDSS][]{ref} and the SuperNovae Legacy Survey
+\citep[SNLS][]{astier2006}. An ellipse in the salt2.4 $(x_1, c)$ plan with major
+axis $x_1=3$ and manor axis $c=0.3$ incapsulate the full distribution \citep[see
+also][for a similar contour but using a less conservative
+$c=0.2$]{campbell2013}. Assuming the average SN absolute magnitude ($x_1=0$,
+$c=0$) is $M=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute magnitude of
+SNeIa along this ellipse given the standardisation coefficient from
+\citep{scolnic2018a}: $\alpha=0.16$ and $\beta=3.14$.
 
-To determine this limit, we first took the observed counts in each bin of the
-histograms. We then fitted $a$ on these counts for different numbers of bins
-used to fit: the quality of the fit diminishes when we use bins that include
-selection effects. To quantize this, we computed the expected counts from the
-fitted $a$ in the same bins we used for the observed counts. That allowed us to
-compute the poissonian probability in each bin, given by
-\begin{equation}
-    p(k) = \frac{\l^k}{k!}e^{-\l}
-\end{equation}
-with $k$ the expected counts, $\l$ the observed counts. This work is done 100
-times with a random number of total bins, $a$ being fitted 10 times on a
-different and random number of bins used to fit each time. We interpolate these
-1000 poissonian evolutions to get one final result that is as close as possible
-to accurately represent how the probability to follow the expected rate evolves.
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/zmax_maglim_all.pdf}
+    \caption{SALT2.4 stretch and color lightcurve parameters of SNeIa from the
+        SDSS, PS1 and SNLS samples from the pantheon catalog. The individual SN
+        data are shown as blue dots and a 2D histogram is shown in grey to
+        highlight point density. The ellipse encapsulating all the SN data is
+        displayed colored by the effective standardized absolute magnitude using
+    the $\alpha$ and $\beta$ standardisation coefficents from
+\citep{scolnic2018}.}
+    \label{fig:maglim}
+\end{figure}
 
-To get the maximum redshift after which we stop following the expected rate, we
-took the redshift for which we are at a 0.3 on the interpolated cumulative
-distribution function. We chose that value to have a systematic analytical
-result that matches what we expect from the histograms; its impact on the work
-after is to be studied.
+The faintest SNIa is that with $x_1=-1.66$ and $c=0.25$ and has an absolute \mr{standardized} magnitude at peak \mr{in Bessel-b band} of $-18.31$ mag, which leads to limiting asbolute magnitude of
+$-18.00$ if we want to be able to observe it typically a week before and 10 days after peak.  
+Therefore, given the $5\sigma$ point source detection magnitude
+limit of a magnitude limited survey, one can derive the maximum redshift above which the faintest SNeIa will start to be missed.
 
-\begin{table}[htbp!]
+\mr{SNLS typically acquires SNeIa in the redshift range $0.4<z<0.8$.
+At these redshifts the rest-frame Bessel-b band roughly corresponds to the SNLS-$i$ filter, that has a 24.8 mag $5\sigma$ depth \footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT final release website.}}.
+This converts to a $z_{lim}=0.60$, in perfect agreement with \citep{neill2006} and \citep{perrett2010}.
+Similarly, PS1 observing SNeIa in the range $0.2<z<0.4$, their $g$-band $5\sigma$ depth is 23.1 mag \citep{rest2014}, corresponding $z_{lim}=0.30$.
+The fig.~6 of \citep{scolnic2018} suggests a more conservative $z_{lim}$ of 0.27 for the PS1 catalog.
+In similar redshift range, SDSS has a limiting magnitude of 22.5 \citep{dilday2008,sake2008,sako2014}, which would lead to a $z_max=0.24$.
+However, the SDSS survey were more sensitive to limited spectroscopic ressources.
+\cite{kessler2009} presents that during year-1 of SDSS, SNeIa with $r-mag<20.5$ where favored for spectroscopic follow up, corresponding to the redshift cut at $0.15$.
+For the rest of the SDSS survey, additional spectroscopic ressources where used, such that \cite{kessler2009} and  \cite{dilday2008} of show a relative completness up to $z_{lim}=0.2$.
+For the rest of the analysis, we will use $z_{lim}=0.2$ as baseline magnitude.}
+
+\mr{In Section~\ref{sec:testzmax}, we test the impact on using more conservative redshift limits ($z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS and $z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}).
+}
+
+In addition, we use the SNe Ia sample of from the nearby supernovae factory
+\citep[SNfactory][]{aldering2004} published in \citep{rigault2018}. As the SNe
+search were much deeper than the spectrophotometric follow-up, SNfactory SNe Ia
+within a redshift range of $0.02<z<0.09$ should also be a random sampling of the
+underlying SN population. The HST sample from
+Pantheon follows the same logic and we therefore kept it entirely \citep{FIND BACK THE REF}.
+
+Our "complete sample" is therefore made of 656 SNeIa whose composition is
+detailed in table~\ref{tab:sample}.
+
+\begin{table}
     \centering
-    \caption{Number of SNe Ia effectively used for each survey, totalling 689
-    Sne Ia. TO CORRECT FOR NEW ZMAX}
+    \caption{Origin of the SNeIa in our sample and redshift limit (if
+    applicable) of each surveys.}
     \label{tab:sample}
-    \begin{tabular}{c c}\hline
-        \rowcolor{gray!15} Survey & Number of SNe Ia \\\hline
-        SNf & 141 \\
-        PS1 & 178 \\
-        SDSS & 206 \\
-        SNLS & 138 \\
-        HST & 26 \\\hline
+    \begin{tabular}{l c c}
+    \hline\hline\\[-0.8em]
+        Survey & N$_{\mathrm{SN}}$  & z$_{lim}$ \\[0.15em]
+        \hline\\[-0.8em]
+        SNf & 141 & --\\[0.30em]
+        SDSS & 167 & 0.20\\[0.30em]
+        PS1 & 160 & 0.30 \\[0.30em]        
+        SNLS & 102 & 0.60\\[0.30em]
+        HST & 26 & --\\[0.30em]
+        \hline
     \end{tabular}
 \end{table}
+
+%\mri{COMPARE HERE WITH LITERATURE. E.G. FIG. 6 OF SCOLNIC 2018, PS zmax AROUND0.27}
+
+%\mri{Perrett et al. 2010 | SNLS | zmax ="z $\sim$ 0.6"}
+
+%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, fig. 10 suggest still okat z=0.2}
+
+%\mri{Kessler 2009 "For the SDSS-II, the image-subtraction
+%pipeline efficiency (	subtr) is complete up to a redshift of z ∼ 0.2"}
+
+%\mri{Kesslet 2009 | "For the SDSS-II and SNLS samples, the cutoff redshifts are
+%estimated to be 0.15 and 0.65, respectively."}
+% ADDIND: PS cut from pantheon close to 0.27 see Fig. 6 of Scolnic 2018
+
+\mri{CHECK LITERATURE CAN WE USE THE CfA SAMPLES ???}
 
@@
 \begin{equation}
-    \frac{\de\z}{\p\z} \equiv \mathrm{LsSFR}\z = K\times \LF{1+z}^\f
+    \frac{\delta\z}{\psi\z} \equiv \mathrm{LsSFR}\z = K\times \LF{1+z}^\varphi
 \end{equation}
@@
 \begin{align}
-    \de\z & = \LF{K^{-1}\times \LF{1+z}^{-\f} +1}^{-1}\\
-    \p\z & = \LF{K\times \LF{1+z}^{+\f} + 1}\um
+    \delta\z & = \left{K^{-1}\times \LF{1+z}^{-\f} +1\right}^{-1}\\
+    \psi\z & = \LF{K\times \LF{1+z}^{+\f} + 1}\um
 \end{align}
 
-\begin{figure*}[htbp!]
-    \centering
-    \subfigure[Raw data]{\label{fig:snf_data}
-        \includegraphics[width=8.5cm]{Article_figures/BiGaussian_onlydata.pdf}}
-    \subfigure[First implemented model]{\label{fig:3G2M2SSNF}
-        \includegraphics[width=8.5cm]{Article_figures/BiGaussian.pdf}}
-    \caption{Stretch des supernovae étudiées par la collaboration SNF en
-    fonction de log(LsSFR). La couleur représente la probabilité pour une
-supernova d'être issue d'un jeune progéniteur.}
-    \label{fig:mod_first}
-\end{figure*}
+%\begin{figure*}
+%    \centering
+%    \subfigure[Raw data]{\label{fig:snf_data}
+%        \includegraphics[width=8.5cm]{Article_figures/BiGaussian_onlydata.pdf}}
+%    \subfigure[First implemented model]{\label{fig:3G2M2SSNF}
+%        \includegraphics[width=8.5cm]{Article_figures/BiGaussian.pdf}}
+%    \caption{Stretch des supernovae étudiées par la collaboration SNF en
+%    fonction de log(LsSFR). La couleur représente la probabilité pour une
+%supernova d'être issue d'un jeune progéniteur.}
+%    \label{fig:mod_first}
+%\end{figure*}
 
@@
 
-        \rowcolor{gray!15} Modèle & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ &
+        Model & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ &
         $\s_2$ \\\hline
@@
 
-    \rowcolor{gray!15} Modèle & $\m$ & $\s$ \\\hline
+    Model & $\m$ & $\s$ \\\hline
 
@@
 
-    \rowcolor{gray!15} Modèle & $\m$ & $\s_-$ & $\s_+$ \\\hline
+    Model & $\m$ & $\s_-$ & $\s_+$ \\\hline
 
@@
 
-    \rowcolor{gray!15} Modèle & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ & $\s_2$ &
+    Model & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ & $\s_2$ &
     $m_3$ & $\s_3$ \\\hline
@@
 
-        \rowcolor{gray!15} Name & Description & Free param &
+        Name & Description & Free param &
         $\ln\mathcal{L}$ & $AICc$ & $\D AICc$ & Proba \\\hline

### 2019-12-05 Nicolas.tex
Integrated plots and updates

@@
-\documentclass[bibnumber]{aa}
+\documentclass[]{aa} %bibnumber
 
-\usepackage{hyperref}
+%\usepackage{hyperref}
+\usepackage{txfonts}
+\usepackage{amsmath}
+\usepackage[colorlinks,breaklinks]{hyperref}
+\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}
 
@@
 %    section \ref{sec:sam}. The data we use is in plain markers.}
+%\subfigure[Method]{\label{fig:zdet}
+
 %\end{figure*}
@@
 We base our analysis on the latest SNe Ia compilation: the Pantheon catalog from
-\cite{scolnic2018a}. Using this dataset, we aim at probing the evolution of the
-distribution of SNe Ia stretch in nature as a function of redshift. Because the
+\cite{scolnic2018}. Using this dataset, we aim at probing the evolution of the
+distribution of SNe Ia stretch in nature as a function of redshift. \mr{We thus have to }
+
+Because the
 observed magnitude of a SN Ia is correlated with its stretch and its color, in a
@@
 
-\mr{In Section~\ref{sec:testzmax}, we test the impact on using more conservative redshift limits ($z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS and $z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}).
-}
+\mr{In Section~\ref{sec:testzmax}, we test the impact on using more conservative redshift limits ($z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS and $z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}).} The cuts on the raw data of those three samples are shown figure \red{fig:cuts}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/hist_surveys_cuts.pdf}
+    \caption{Histograms of the 3 surveys on which we applied cuts to get rid of selection effects, as discussed in section \ref{sec:sam}.}
+    \label{fig:cuts}
+\end{figure}
 
@@
 
-We fitted it on SNf data, giving the results table \ref{tab:val}. For clarity
+We fitted it on SNf data, giving the results table \ref{tab:val}, with a graphical representation shown figure \ref{fig:snf_model}. For clarity
 with the next models, we named it 3G2M2S$\St{SNf}$ for it has a total of 3
@@
 on SNf data only.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/stretchevol_model_maglim-cuts.pdf}
+    \caption{Result of the model fitted on SNF data and mean of stretch and redshift of the other surveys used for this analysis. Conservative data is shown in transparent markers.}
+    \label{fig:snf_model}
+\end{figure}
 
@@
 
+To test the coherence of this model, we also fitted it using all the data from the surveys indistinctly. These results are shown figures \ref{fig:model_all} and \ref{fig:model_all_cons}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/stretchevol_all_vs_snf_maglim_sup.pdf}
+    \caption{Result of the model fitted on all surveys data cut at the non-conservative $z_{\mathrm{max}}$.
+We chose the bins to compute the means of stretch and redshift by making them equal in number of SNe Ia.}
+    \label{fig:model_all}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/stretchevol_all_vs_snf_maglim_inf.pdf}
+    \caption{Result of the model fitted on all surveys data cut at the conservative $z_{\mathrm{max}}$.
+We chose the bins to compute the means of stretch and redshift by making them equal in number of SNe Ia.}
+    \label{fig:model_all}
+\end{figure}
+
 \section{Results}
@@
 
+\begin{acknowledgements}
+    This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement n 759194 - USNAC).
+Support has been provided by the Institut Universitaire de France, the CNES, and the region Auvergne-Rhone-Alpes.
+\end{acknowledgements}
+
+\bibliographystyle{aa} % style aa.bst
+\begin{thebibliography}{} 
+% A
+\bibitem[Abbott et al.(2018)]{abbott2018} Abbott, T.~M.~C., Abdalla, F.~B., Alarcon, A., et al.\ 2018, \prd, 98, 043526
+
+\bibitem[Ata et al.(2017)]{ata2017} Ata, M., Kitaura, F.-S., Chuang, C.-H., et al.\ 2017, \mnras, 467, 3993
+
+% B
+% C
+\bibitem[Chabanier et al.(2019)]{chabanier2019} Chabanier, S., Millea, M., \& Palanque-Delabrouille, N.\ 2019, \mnras, 489, 2247
+\bibitem[Coles, \& Jones(1991)]{coles1991} Coles, P., \& Jones, B.\ 1991, \mnras, 248, 1
+
+% D
+% E
+% F
+\bibitem[Freedman et al.(2019)]{freedman2019} Freedman, W.~L., Madore, B.~F., Hatt, D., et al.\ 2019, \apj, 882, 34
+
+% G
+
+%\bibitem[Graziani et al.(2019)]{graziani2019} Graziani, R., Courtois, H.~M., Lavaux, G., et al.\ 2019, \mnras, 488, 5438
+
+% H
+% I 
+% J
+\bibitem[Jones et al.(2018)]{jones2018} Jones, D.~O., Riess, A.~G., Scolnic, D.~M., et al.\ 2018, \apj, 867, 108
+
+% K
+
+\bibitem[Knox, \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv e-prints, arXiv:1908.03663
+
+% L
+
+% M
+% N
+% O
+% P
+\bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T., et al.\ 2019, \prl, 122, 221301
+
+\bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration, Aghanim, N., Akrami, Y., et al.\ 2018, arXiv e-prints, arXiv:1807.06209
+
+% Q
+% R
+\bibitem[Reid et al.(2019)]{reid2019} Reid, M.~J., Pesce, D.~W., \& Riess, A.~G.\ 2019, arXiv e-prints, arXiv:1908.05625
+
+\bibitem[Riess et al.(2016)]{riess2016} Riess, A.~G., Macri, L.~M., Hoffmann, S.~L., et al.\ 2016, \apj, 826, 56
+
+\bibitem[Riess et al.(2019)]{riess2019} Riess, A.~G., Casertano, S., Yuan, W., et al.\ 2019, \apj, 876, 85
+
+\bibitem[{Rigault {et~al.}(2013)}]{rigault2013}
+Rigault, M., Copin, Y., Aldering, G., {et~al.} 2013, \aap, 560, A66
+
+\bibitem[Rigault et al.(2015)]{rigault2015} Rigault, M., Aldering, G., Kowalski, M., et al.\ 2015, \apj, 802, 20
+
+\bibitem[Rigault et al.(2018)]{rigault2018} Rigault, M.,
+  Brinnel, V., Aldering, G., et al.\ 2018, arXiv:1806.03849
+% S
+\bibitem[Scolnic et al.(2018)]{scolnic2018} Scolnic, D.~M., Jones, D.~O., Rest, A., et al.\ 2018, \apj, 859, 101
+
+% T
+% U 
+% V
+% W
+\bibitem[Wong et al.(2019)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F., et al.\ 2019, arXiv e-prints, arXiv:1907.04869
+
+% X
+% Y
+ Z
+\end{thebibliography}
+
 \end{document}

### 2019-12-06 Nicolas.tex
Update

@@
 As such, the probability to observe a young SNe Ia labeled "i" with a stretch
-$x_1^i$ and error $\d x_1^i$ is
+$x_1^i$ and error $\mathrm{d} x_1^i$ is
 \begin{equation}
-    p \LF{x_1^i, \d x_1{}^i|\m_1,\s_1} = \Nc \LF{\m_1, \Sq{\s_1{}^2 + \d
-    x_1^i{}^2}}(x_1^i)
+    p \LF{x_1^i, \mathrm{d} x_1{}^i|\mu_1,\sigma_1} = \mathcal{N} \left(\mu_1,
+    \sqrt{\sigma_1{}^2 + \mathrm{d} x_1^i{}^2}\right)(x_1^i)
 \end{equation}
@@
 \begin{align}
-    p(x_1^i, \d x_1^i | \m_{1,2}, \s_{1,2}, a) = a & \times \Nc \LF{\m_1,
-    \sqrt{\s_1{}^2 + \d x_1^i{}^{2}}} (x_1^i) \ + \\ (1-a) & \times \Nc
-    \LF{\m_2, \sqrt{ \s_2{}^2 + \d x_1^i{}^{2}}} (x_1^i), \nonumber 
+    p(x_1^i, \mathrm{d} x_1^i | \mu_{1,2}, \sigma_{1,2}, a) = a & \times
+    \mathcal{N} \left(\mu_1, \sqrt{\sigma_1{}^2 + \mathrm{d} x_1^i{}^{2}}\right) (x_1^i) \ + \\ (1-a) & \times \mathcal{N}
+    \left(\mu_2, \sqrt{ \sigma_2{}^2 + \mathrm{d} x_1^i{}^{2}}\right) (x_1^i), \nonumber 
 \end{align}
@@
 \noindent where $a$ is the relative amplitude between the two gaussians.
-Finally, the normalized stretch distribution at a given $z$ $\D\LF{x_1|z}$ is
+Finally, the normalized stretch distribution at a given $z$
+$\Delta\left(x_1|z\right)$ is
 the weigthed sum of both young and old stretch distributions given their
-relative fraction $\de\z$:
+relative fraction $\delta(z)$:
 \begin{equation}
-    \D\LF{x_1|z} = \de\z\times\Nc_1 +
-    \LF{1-\de\z} \times \LF{a\Nc_1 + (1-a)\Nc_2}
+    \Delta\left(x_1|z\right) = \delta(z)\times\mathcal{N}_1 +
+    \left(1-\delta(z)\right) \times \left(a\mathcal{N}_1 +
+    (1-a)\mathcal{N}_2\right)
 \end{equation}
@@
 We fitted it on SNf data, giving the results table \ref{tab:val}, with a graphical representation shown figure \ref{fig:snf_model}. For clarity
-with the next models, we named it 3G2M2S$\St{SNf}$ for it has a total of 3
+with the next models, we named it 3G2M2S$_{\text{SNf}}$ for it has a total of 3
 gaussians but with only 2 means and 2 standard deviations, and has been fitted
@@
 We implemented and compared 10 models in total, 4 of which have an
-evolution with the redshift from $\de\z$, and 6 don't ($\de\z = f =
+evolution with the redshift from $\delta(z)$, and 6 don't ($\delta(z) = f =
 \mathrm{cst}$). The ones with and evolution are:
@@
     \item 3G2M2S, the one we described (but fitted on all the data);
-    \item 3G2M1S, where this time $\s_1 \equiv \s_2$;
-    \item 2G2M2S, model taken from HOWELL 2009 where we added $\de\z$;
+    \item 3G2M1S, where this time $\sigma_1 \equiv \sigma_2$;
+    \item 2G2M2S, model taken from HOWELL 2009 where we added $\delta(z)$;
     \item 3G3M3S, with three independent gaussians.
@@
 The ones without a stretch evolution are the same ones but with an "F" implying
-we set $\de\z = f = \mathrm{cst}$, and two others:
+we set $\delta(z) = f = \mathrm{cst}$, and two others:
 \begin{itemize}
@@
 \begin{align}
- p(x_1^i, \d x_1^i | \m, \s_-, \s_+) = 
+ p(x_1^i, \mathrm{d} x_1^i | \mu, \sigma_-, \sigma_+) = 
     \begin{cases}
-        \mathcal{N} \LF{\m, \sqrt{\s_-{}^2 + \d x_1^i{}^{2}}} (x_1^i) & \text{if
-        } x_1^i\geq \m\\
-        \mathcal{N} \LF{\m, \sqrt{\s_+{}^2 + \d x_1^i{}^{2}}} (x_1^i), &
+        \mathcal{N} \left(\mu, \sqrt{\sigma_-{}^2 + \mathrm{d} x_1^i{}^{2}}\right) (x_1^i) & \text{if
+        } x_1^i\geq \mu\\
+        \mathcal{N} \left(\mu, \sqrt{\sigma_+{}^2 + \mathrm{d} x_1^i{}^{2}}\right) (x_1^i), &
         \text{else}
@@
 
-        Model & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ &
-        $\s_2$ \\\hline
+        Model & $a$ & $f$ & $\mu_1$ & $\sigma_1$ & $\mu_2$ &
+        $\sigma_2$ \\\hline
 
@@
         3G2M1S & $0.47 \pm 0.07$ & none & $0.35 \pm 0.04$ & $0.61 \pm 0.03$ &
-        $-1.25 \pm 0.10$ & $\s_1$ \\
+        $-1.25 \pm 0.10$ & $\sigma_1$ \\
 
         3G2M1SF & \textcolor{red}{$0.2 \pm 0.9$} & $0.7 \pm 0.3$ & $0.36 \pm
-        0.04$ & $0.60 \pm 0.03$ & $-1.23 \pm 0.10$ & $\s_1$ \\
+        0.04$ & $0.60 \pm 0.03$ & $-1.23 \pm 0.10$ & $\sigma_1$ \\
 
@@
 
-    Model & $\m$ & $\s$ \\\hline
+    Model & $\mu$ & $\sigma$ \\\hline
 
@@
 
-    Model & $\m$ & $\s_-$ & $\s_+$ \\\hline
+    Model & $\mu$ & $\sigma_-$ & $\sigma_+$ \\\hline
 
@@
 
-    Model & $a$ & $f$ & $\m_1$ & $\s_1$ & $\m_2$ & $\s_2$ &
-    $m_3$ & $\s_3$ \\\hline
+    Model & $a$ & $f$ & $\mu_1$ & $\sigma_1$ & $\mu_2$ & $\sigma_2$ &
+    $\mu_3$ & $\sigma_3$ \\\hline
 

### 2019-12-19 Nicolas.tex
Big, BIG update

@@
 \newcommand{\mri}[1]{{\textcolor{red}{#1}}}
+\newcommand{\nn}[1]{{\textcolor[rgb]{1, 0.27, 0}{#1}}}
 
 \def\l{\lambda}\def\L{\Lambda}
+
+\newcommand{\prob}[2]{\mathcal{P}\left( #1 \mid #2\right)}
 
@@
     diagram, giving us information about the Universe's expansion and its
-    fondamental components, one of which is dark energy. But systematic
+    fundamental components, one of which is dark energy. But systematic
     uncertainties are now starting to be limiting in our ability to measure
-    those parameters. In particular, the physics of SNe Ia is still mostly
-    unknown, and is thought not to change in time/with the redshift.}
+those parameters. In particular, the physics of SNe Ia is still mostly unknown,
+and is thought not to change in time/with the redshift.}
     {In an attempt to reduce those uncertainties, we try to find an empirical
@@
     {We started by getting a complete sample representing all of the stretch
-        distribution that Nature can give us, before using LsSFR measurments, an
-        age tracer which evolution with redshift is known, that has been shown
-        to have a strong correlation with the stretch. We compare their AICc, an
-    estimator of the relative quality of statistical models that includes the
-number of free parameters, to determine which ones describe besto the data.}
+        distribution that Nature can give us, before using LsSFR measurements,
+        an age tracer which evolution with redshift is known, that has been
+        shown to have a strong correlation with the stretch. We compare their
+        AICc, an estimator of the relative quality of statistical models that
+    includes the number of free parameters, to determine which ones describe
+best the data.}
     {Models with an evolution of the stretch with the redshift have a better
@@
 %    \centering
-%    \subfigure[Method]{\label{fig:zdet}
+%    \subfigure[Method]{\label{fig:zdet}%        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}
+
 %        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}
@@
 have enabled the discovery of the acceleration of the Universe's expansion
-\citep{riess1998, permutter1999}. They remain today a key comological probe to
+\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
 understand the properties of dark energy (DE) as it is the only tool able to
@@
 standard $\L$CDM or when the SN luminosity is anchored at intermediate redshift
-by the baryon accoustic oscillation (BAO) scale \citep{riess2019, ried2019,
-planck, feeney}. While using the tip of the red giant branch technique in place
-of the Cepheids seem to favor lower values of H$_0$ \citep{freedman,
-recalibrationpaper} time delay measurements from strong lensing seem to also
-favor high H$_0$ values \citep{wong2019}.
+by the baryon acoustic oscillation (BAO) scale \citep{riess2019, reid2019,
+planck2016, feeney} \nn{changed ried to reid; changed planck to planck2016;
+feeney = https://tinyurl.com/vg6uufm ?}. While using the tip of the red giant
+branch technique in place of the Cepheids seem to favor lower values of H$_0$
+\citep{freedman2019, recalibrationpaper} \nn{changed "freedman" to
+"freedman2019" ; recalibrationpaper = ?} time delay measurements from strong
+lensing seem to also favor high H$_0$ values \citep{wong2019}.
 
 The H$_0$ tension has received a lot of attention as it could be a sign of new
-fundamental physics. Yet, no simple solution is able to accomodate this H$_0$
+fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
 tension when accounting for all other probes and the current most promising
 scenario appears to be a burst of expansion at the matter-radiation decoupling
-moment caused by a (fine tuned) early dark energy \citep{poulin}.
+moment caused by a (fine tuned) early dark energy \citep{poulin2019} \nn{changed
+"poulin" to "poulin 2019"}.
 
@@
 analysis could also explain the tension. In \cite{rigault2015} we suggested that
-SNe Ia from the cepheid calibrator sample differ significantly to the one from
+SNe Ia from the Cepheid calibrator sample differ significantly to the one from
 the Hubble flow sample that enables to derive H$_0$. Indeed, Cepheid are very
 young stars and the existence of Cepheids in a galaxy implies that the host is
-star forming. In \citep{Rigault2015} we claimed that a more than 90\% of the
-calibrator SNe Ia arise from a young-progenitor propulation while this
-population only accounts for about half of the Hubble flow sample. Following
-\cite{sullivan2010} and \citep{rigault2013}, we further showed that young and
-old progenitor SNe Ia have a different magnitude of about 0.15 mag. This has
-been confirmed for modern standardization technique
+star forming. In \citep{rigault2015} we claimed that a more than 90\% of the
+calibrator SNe Ia arise from a young-progenitor population while this population
+only accounts for about half of the Hubble flow sample. Following
+\cite{sullivan2010} and \citep{rigault2013} \nn{should use cite or citep?}, we
+further showed that young and old progenitor SNe Ia have a different magnitude
+of about 0.15 mag. This has been confirmed for modern standardization technique
 \citep[SALT2.4][]{guy2010,betoule2014} with high significance ($\sim 6\sigma$)
-in \citep{rigault2018} as well as independently by \cite{roman2019} at $\sim
-7\sigma$. As detailed in \citep{Rigault2015}, if the magnitude difference
-between young and old SNe Ia is indeed of $0.15$ mag and if SNe Ia from the
-calibrator sample is fully dominated by young SNeIa, the effect on H$_0$ of the
-existance of these two SNe Ia populations is of 3.5\%. Accounting for the fact
-that another environmental effect (the mass-step) is already included in the
-SH0ES analysis, the net bias in the SH0ES analysis is expected closer to 3\%
-\citep[see table 6 of ][]{rigault2015}.
+in \citep{rigault2018} as well as independently by \cite{roman2019} \nn{not
+found on ADS: https://tinyurl.com/uemattg} at $\sim 7\sigma$. As detailed in
+\citep{rigault2015}, if the magnitude difference between young and old SNe Ia is
+indeed of $0.15$ mag and if SNe Ia from the calibrator sample is fully dominated
+by young SNeIa, the effect on H$_0$ of the existence of these two SNe Ia
+populations is of 3.5\%. Accounting for the fact that another environmental
+effect (the mass-step) is already included in the SH0ES analysis, the net bias
+in the SH0ES analysis is expected closer to 3\% \citep[see table 6
+of][]{rigault2015}.
 
@@
 it could not explain the high value of H$_0$ derived by strong lensing. First,
-\cite{jones2015} did not find any environmental bias when extending the
-\citep{rigault2015} study. It was more recently reported in \citep{jones2019}
-that indeed local host analysis has significant influence on the SN magnitude
-but much more importantly than claimed in \citep{rigault2018} and with a
-connection with the progenitor physics and/or a line of sight effect that is
-unclear. More importantly, \cite{riess2016} attended to mimic the cepheid sample
+\cite{jones2015} \nn{put the ref but nos sure it's the correct one} did not find
+any environmental bias when extending the \citep{rigault2015} study. It was more
+recently reported in \citep{jones2019} \nn{put the ref but nos sure it's the
+correct one} that indeed local host analysis has significant influence on the SN
+magnitude but much more importantly than claimed in \citep{rigault2018} and with
+a connection with the progenitor physics and/or a line of sight effect that is
+unclear. More importantly, \cite{riess2016} attended to mimic the Cepheid sample
 selection function on the Hubble flow sample by removing SNe Ia from the latter
-if they classificed their host as early-type or non-star forming. They measured
+if they classified their host as early-type or non-star forming. They measured
 that doing so has no effect on the measurement of H$_0$.
 
-% SERIES OF PAPER
 This paper is part of a series of papers where we test the validity of the two
-SNe Ia population models developped in \citep{rigault2018}.Briday et al. in prep
+SNe Ia population models developed in \citep{rigault2018}.Briday et al. in prep
 will present how using different environmental tracer is affecting our ability
 to distinguish the two populations, and Rigault et al. in prep will further
-study the connection with H$_0$. Here we analyze predictions concerning the
-SNe Ia redshift drifts: (1) that the distribution of the SNIa stretch, a purely
+study the connection with H$_0$. Here we analyze predictions concerning the SNe
+Ia redshift drifts: (1) that the distribution of the SNIa stretch, a purely
 intrinsic SNIa property, is evolving as a function of redshift and (2) that this
@@
 The concept of the SNeIa age dichotomy arised when the SNIa rate has first been
-studied. \cite{Mannucci et al. 2005; Scannapieco & Bildsten 2005; Sullivan et
-al. 2006; Aubourg et al. 2008} have shown that the relative SNe Ia rate in
-galaxies could only be explained if two populations existed, one young,
-following the host star formation activity, and one old following the host
-stellar mass (the so called “prompt and delayed“ or “A+B“ model). In
-\cite{rigault2018} we used the specific star formation rate at the SN location
-(Local sSFR or LsSFR) to classify which are the prompts (those with a high
-LsSFR) and which are the delayed (those with low LsSFR). Since the first SNe Ia
-host analysis, the SN strech has been known to be strongly correlated with the
-SN host properties \citep{Hamuy1996, Hamuy2000} and it has been extensively
-confirmed since \citep[e.g.][]{Neill2009,Sullivan2010; Lampeitl2010; Kelly2010;
-Gupta2011,D’Andrea2011; Childress2013,Rigault2013,Pan2014}. Following the “A+B“
-model and the connection between SN stretch and host properties,
-\citep{Howell20XX} first discussed the potential redshift drift of the SN strech
-distribution. In this paper we revisit this analysis with the most recent SNe Ia
-dataset and we test the validity of the two SNIa age populations model
-developped by \citep{rigault2018}.
+studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
+shown that the relative SNe Ia rate in galaxies could only be explained if two
+populations existed, one young, following the host star formation activity, and
+one old following the host stellar mass (the so called “prompt and delayed“ or
+“A+B“ model). In \cite{rigault2018} we used the specific star formation rate at
+the SN location (Local sSFR or LsSFR) to classify which are the prompts (those
+with a high LsSFR) and which are the delayed (those with low LsSFR). Since the
+first SNe Ia host analysis, the SN stretch has been known to be strongly
+correlated with the SN host properties \citep{hamuy1996, hamuy2000} \nn{put
+hamuy1996 as most cited of https://tinyurl.com/sjlbdz8} and it has been
+extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
+lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
+pan2014} \nn{added all the refs}. Following the “A+B“ model and the connection
+between SN stretch and host properties, \citep{howell2009} \nn{changed
+howell20xx to howell2009
+(https://ui.adsabs.harvard.edu/abs/2009ApJ...691..661H/abstract)} first
+discussed the potential redshift drift of the SN stretch distribution. In this
+paper we revisit this analysis with the most recent SNe Ia dataset and we test
+the validity of the two SNIa age populations model developed by
+\citep{rigault2018}.
 
-We present in section~\ref{sec:data} the sample we are using for this analysis,
-which is based on the pantheon dataset \citep{scolnic2018a}. We discuss the
-importance of obtaining a "complete" sample, i.e. reprensentative of the true
-underlying SNe Ia distribution, and how we build one from the Pantheon sample.
-We then present in section~\ref{sec:model} our modeling of the distribution of
-stretch as a function of redshift based on the relative stretch distrubtion of
-young and old SNeIa. Our results are presented in section~\ref{sec:result} where
-we test if the distribution of stretch does evolve as a function of redshift and
-if the age-model is in good agreement with this evolution. We discuss our
-results and conclude in section~\ref{sec:conclusion}.
+We present in section~\ref{sec:sample} the sample we are using for this
+analysis, which is based on the pantheon dataset \citep{scolnic2018a}. We
+discuss the importance of obtaining a "complete" sample, i.e. representative of
+the true underlying SNe Ia distribution, and how we build one from the Pantheon
+sample.  We then present in section~\ref{sec:modeling} our modeling of the
+distribution of stretch as a function of redshift based on the relative stretch
+distribution of young and old SNeIa. Our results are presented in
+section~\ref{sec:results} where we test if the distribution of stretch does
+evolve as a function of redshift and if the age-model is in good agreement with
+this evolution. We discuss our results and conclude in section~\ref{sec:ccl}.
 
@@
 
-\section{Sample}\label{sec:sam}
+\section{A "complete" Sample}
+\label{sec:sample}
 
 We base our analysis on the latest SNe Ia compilation: the Pantheon catalog from
-\cite{scolnic2018}. Using this dataset, we aim at probing the evolution of the
-distribution of SNe Ia stretch in nature as a function of redshift. \mr{We thus have to }
+\cite{scolnic2018a}.  \mr{TO BE REPHRASED} Using this dataset, we aim at probing
+if the average stretch, an intrinsic SN Ia property, evolves as a function of
+redshift or not; the underlying question being that of the SNIa population
+redshift drift.  The stretch has been shown to depend on the SN environment star
+formation activity
+\citep[e.g.][]{howell2007,sullivan2010,childress2013,rigault2013,rigault2018}
+and because the star formation rate is about 40 times larger at the high
+redshift end of the pantheon sample than at its lower redshifts
+\citep[e.g.,][and references therein]{tasca2015}, one might think that simply
+comparing the redshifts distribution in a few bins of redshift could be
+sufficient to probe any SN stretch redshift drift.  In practice, selection
+effects are affecting the observed SN stretch distributions, which could
+significantly differ from the true underlying one if these selection effects
+become significant. Indeed, because the observed SNIa magnitude correlates with
+the lightcurve stretches (and colors) the first SNIa that such a survey will
+miss will the reddest and the lowest-stretch ones. Consequently, if selection
+effects is not accounted for, one might confuse true population drift with
+selection effect and conversely.
 
-Because the
-observed magnitude of a SN Ia is correlated with its stretch and its color, in a
-magnitude limited survey such as those consituing most of the Pantheon sample,
-the distribution of these lightcurve parameters is not representative of the
-true underlying distribution above a certain limiting redshift. The first SNe Ia
-that such a survey will miss because they will be too faint are indeed the
-reddest and the lowest-stretch SNe Ia. If such a selection effect is not
-accounted for, we might confuse true population drift we selection function or
-conversally.
+In a magnitude limited survey, providing sufficient (and random) spectroscopic
+follow-up for acquiring typing and host redshift, the selection effects should
+be negligible bellow a given redshift where all SNe Ia, even the faintest, could
+be observed.  In contrast, targeted surveys have highly complex selection
+functions and we will discard them from our analysis. Fortunately, modern SN
+cosmology sample such has the pantheon dataset are now dominated by magnitude
+limited survey.
 
-The figure~\ref{fig:maglim} shows the distribution of stretch for the magnitude
-limited samples of the Pantheon data, i.e. PanStarrs \citep[PS1][]{ref}, the
-Sloan Digital Sky Survey \citep[SDSS][]{ref} and the SuperNovae Legacy Survey
-\citep[SNLS][]{astier2006}. An ellipse in the salt2.4 $(x_1, c)$ plan with major
-axis $x_1=3$ and manor axis $c=0.3$ incapsulate the full distribution \citep[see
-also][for a similar contour but using a less conservative
-$c=0.2$]{campbell2013}. Assuming the average SN absolute magnitude ($x_1=0$,
-$c=0$) is $M=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute magnitude of
-SNeIa along this ellipse given the standardisation coefficient from
-\citep{scolnic2018a}: $\alpha=0.16$ and $\beta=3.14$.
+We present fig.~\ref{fig:maglim} the lightcurve stretch and color distributions
+of SNeIa from these surveys, namely: PanStarrs \citep[PS1][]{ref}, the Sloan
+Digital Sky Survey \citep[SDSS][]{ref} and the SuperNovae Legacy Survey
+\citep[SNLS][]{astier2006}.  An ellipse in the salt2.4 $(x_1, c)$ plan with
+major axis $x_1=3$ and manor axis $c=0.3$ encapsulates the full distribution
+(see also \citealt{bazin2011} and \citealt{campbell2013} for a similar contour
+but using a less conservative $c=0.2$ for the latter). Assuming the SN absolute
+magnitude with $x_1=0$ and  $c=0$ is $M=-19.36$ \citep{kessler2009,scolnic2014}
+we can derive the absolute standardized magnitude at maximum of light along the
+aforementioned ellipse given the standardization coefficient $\alpha=0.16$ and
+$\beta=3.14$ from \cite{scolnic2018a}.  The faintest SNIa is that with
+$x_1=-1.66$ and $c=0.25$ and has an absolute standardized magnitude at peak in
+Bessel-b band of $-18.31\,\mathrm{mag}$. Since we need to see this object
+typically a week before and 10 days after peak to build a suitable lightcurve,
+the effective absolute limiting magnitude therefore is
+$\sim-18.00\,\mathrm{mag}$.  Hence, given the $5\sigma$ point source detection
+magnitude limit of a magnitude limited survey, one can derive the maximum
+redshift above which the faintest SNeIa will start to be missed.
 
@@
         SDSS, PS1 and SNLS samples from the pantheon catalog. The individual SN
-        data are shown as blue dots and a 2D histogram is shown in grey to
+        data are shown as blue dots and a 2D histogram is shown in gray to
         highlight point density. The ellipse encapsulating all the SN data is
         displayed colored by the effective standardized absolute magnitude using
-    the $\alpha$ and $\beta$ standardisation coefficents from
-\citep{scolnic2018}.}
+    the $\alpha$ and $\beta$ standardization coefficients from
+\citep{scolnic2018a}.}
     \label{fig:maglim}
@@
 
-The faintest SNIa is that with $x_1=-1.66$ and $c=0.25$ and has an absolute \mr{standardized} magnitude at peak \mr{in Bessel-b band} of $-18.31$ mag, which leads to limiting asbolute magnitude of
-$-18.00$ if we want to be able to observe it typically a week before and 10 days after peak.  
-Therefore, given the $5\sigma$ point source detection magnitude
-limit of a magnitude limited survey, one can derive the maximum redshift above which the faintest SNeIa will start to be missed.
-
-\mr{SNLS typically acquires SNeIa in the redshift range $0.4<z<0.8$.
-At these redshifts the rest-frame Bessel-b band roughly corresponds to the SNLS-$i$ filter, that has a 24.8 mag $5\sigma$ depth \footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT final release website.}}.
-This converts to a $z_{lim}=0.60$, in perfect agreement with \citep{neill2006} and \citep{perrett2010}.
-Similarly, PS1 observing SNeIa in the range $0.2<z<0.4$, their $g$-band $5\sigma$ depth is 23.1 mag \citep{rest2014}, corresponding $z_{lim}=0.30$.
-The fig.~6 of \citep{scolnic2018} suggests a more conservative $z_{lim}$ of 0.27 for the PS1 catalog.
-In similar redshift range, SDSS has a limiting magnitude of 22.5 \citep{dilday2008,sake2008,sako2014}, which would lead to a $z_max=0.24$.
-However, the SDSS survey were more sensitive to limited spectroscopic ressources.
-\cite{kessler2009} presents that during year-1 of SDSS, SNeIa with $r-mag<20.5$ where favored for spectroscopic follow up, corresponding to the redshift cut at $0.15$.
-For the rest of the SDSS survey, additional spectroscopic ressources where used, such that \cite{kessler2009} and  \cite{dilday2008} of show a relative completness up to $z_{lim}=0.2$.
-For the rest of the analysis, we will use $z_{lim}=0.2$ as baseline magnitude.}
-
-\mr{In Section~\ref{sec:testzmax}, we test the impact on using more conservative redshift limits ($z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS and $z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}).} The cuts on the raw data of those three samples are shown figure \red{fig:cuts}.
+SNLS typically acquires SNeIa in the redshift range $0.4<z<0.8$. At these
+redshifts the rest-frame Bessel-b band roughly corresponds to the SNLS-$i$
+filter that has a 24.8 mag $5\sigma$
+depth\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
+final release website.}}. This converts to a $z_{lim}=0.60$, in perfect
+agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.
+Similarly, PS1 observes SNeIa in the range $0.2<z<0.4$, their $g$-band $5\sigma$
+depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$.  Figure~6 of
+\citep{scolnic2018a} suggests a more conservative $z_{lim}$ of 0.27 for the PS1
+catalog.  In similar redshift range, SDSS has a limiting magnitude of 22.5
+\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
+SDSS survey were more sensitive to limited spectroscopic resources.
+\cite{kessler2009} present that during year-1 of SDSS, SNeIa with $r-mag<20.5$
+were favored for spectroscopic follow up, corresponding to the redshift cut at
+$0.15$. For the rest of the SDSS survey, additional spectroscopic resources were
+used, such that \cite{kessler2009} and \cite{dilday2008} show a relative
+completeness up to $z_{lim}=0.2$.  Following these analysis we will use
+$z_{lim}=0.2$ as the baseline SDSS redshift limit for the rest of the analysis.
+We could have chosen more conservative redshift limits for each of these
+surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS and
+$z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}, and we will
+also present results obtained when doing so.  We present the redshift
+distribution of these three surveys in fig.~\ref{fig:cuts}, where we see that
+the redshift limit roughly correspond to the pick of these histograms, as
+expected.
 
@@
     \includegraphics[width=\linewidth]{Article_figures/hist_surveys_cuts.pdf}
-    \caption{Histograms of the 3 surveys on which we applied cuts to get rid of selection effects, as discussed in section \ref{sec:sam}.}
+    \caption{Histograms of the 3 surveys on which we applied cuts to get rid of
+    selection effects, as discussed in section \ref{sec:sample}. The SNe Ia
+between the conservative and non-conservative cuts are shown in transparent.}
     \label{fig:cuts}
@@
 
-In addition, we use the SNe Ia sample of from the nearby supernovae factory
-\citep[SNfactory][]{aldering2004} published in \citep{rigault2018}. As the SNe
-search were much deeper than the spectrophotometric follow-up, SNfactory SNe Ia
-within a redshift range of $0.02<z<0.09$ should also be a random sampling of the
-underlying SN population. The HST sample from
-Pantheon follows the same logic and we therefore kept it entirely \citep{FIND BACK THE REF}.
-
-Our "complete sample" is therefore made of 656 SNeIa whose composition is
-detailed in table~\ref{tab:sample}.
+In addition, we use the SNeIa from the nearby supernovae factory
+\citep[SNfactory][]{aldering2004} \nn{added ref assuming
+https://ui.adsabs.harvard.edu/abs/2004APS..APR.V4003A/abstract} published in
+\citep{rigault2018} and that have been discoverd from no-targeted survey (114
+SNeIa, see their section~3 and 4.2.2). As the SNe search were much deeper than
+the spectrophotometric follow-up, SNfactory SNe Ia within a redshift range of
+$0.02<z<0.09$ should also be a random sampling of the underlying SN population.
+Data from \citep{rigault2018} are within these redshift boundaries. The HST
+sample from Pantheon follows the same logic and we therefore kept it entirely
+\citep{FIND BACK THE REF} \nn{lots of choices: https://tinyurl.com/rcsx7cw}.
 
@@
     \centering
-    \caption{Origin of the SNeIa in our sample and redshift limit (if
-    applicable) of each surveys.}
+    \caption{Source surveys of SNeIa used in our "complete" sample. Conservative
+    cuts are indicated in brackets.}
     \label{tab:sample}
@@
     \hline\hline\\[-0.8em]
-        Survey & N$_{\mathrm{SN}}$  & z$_{lim}$ \\[0.15em]
+        Survey & N$_{\mathrm{SN}}$  & z$_{max}$ \\[0.15em]
         \hline\\[-0.8em]
-        SNf & 141 & --\\[0.30em]
-        SDSS & 167 & 0.20\\[0.30em]
-        PS1 & 160 & 0.30 \\[0.30em]        
-        SNLS & 102 & 0.60\\[0.30em]
+        SNf & 114 & --\\[0.30em]
+        SDSS & 167(82) & 0.20(0.15)\\[0.30em]
+        PS1 & 160(122) & 0.30(0.27) \\[0.30em]        
+        SNLS & 102(78) & 0.60(0.55)\\[0.30em]
         HST & 26 & --\\[0.30em]
@@
 
-%\mri{COMPARE HERE WITH LITERATURE. E.G. FIG. 6 OF SCOLNIC 2018, PS zmax AROUND0.27}
+%%%%%%%%%%%%%%%%%%%%%%%
+%   LITERATURE        %
+%%%%%%%%%%%%%%%%%%%%%%%
+%\mri{COMPARE HERE WITH LITERATURE. E.G. FIG. 6 OF SCOLNIC 2018, PS zmax
+%AROUND0.27}
 
@@
 
-%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, fig. 10 suggest still okat z=0.2}
+%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, fig. 10 suggest still
+%okat z=0.2}
 
@@
 
-\mri{CHECK LITERATURE CAN WE USE THE CfA SAMPLES ???}
+%%%%%%%%%%%%%%%%%%%%%%%
+%       INFO          %
+%%%%%%%%%%%%%%%%%%%%%%%
+%Fraction of random SDSS stretchs being < mean_conservative =  0.299
+%Fraction of random PS1 stretchs being < mean_conservative =  0.254
+%Fraction of random SNLS stretchs being < mean_conservative =  0.417
 
-\section{Method}
+\section{Modeling the redshift drift}
+\label{sec:modeling}
 
-We used the LsSFR and stretch measurments from the SNf sample. The LsSFR has an
-evolution with the redshift that is analytically known: calling $\de\z$ the
-fraction of young stars and $\p\z$ the fraction of old ones, Rigault 2018 et +
-find
-\begin{equation}
-    \frac{\delta\z}{\psi\z} \equiv \mathrm{LsSFR}\z = K\times \LF{1+z}^\varphi
-\end{equation}
-with $\f = 2.8$, and knowing $\de\z + \p\z = 1$:
+In \cite{rigault2018} we presented a modeling of the evolution of the fraction
+of prompt and delayed SNeIa as a function of redshift following former work on
+rates and delay time distributions; see e.g. the prompt and delayed SNeIa model,
+aka the "A+B" model,
+\cite{mannucci2005,scannapieco2005,sullivan2006,aubourg2008} and
+\cite{maozmannucci2014} for a review on delay time distributions of SNeIa. In
+short, we assume that the number of prompt SNeIa follows the star formation
+activity in the Universe while the number of delayed SNeIa follows the number of
+Gyr-old stars in the Universe, i.e. the stellar mass. Hence, if we denote
+calling $\delta(z)$ (resp. $\phi(z)$) the fraction of young (resp. old) SNeIa in
+the Universe as a function of redshift, then their ratio is expected to follow
+the evolution of the specific star formation rate which goes as $(1+z)^{2.8}$
+until $z\sim2$ \citep{tasca2015}. Hence since there is  $\sim50\%$ of young
+SNeIa at $z\sim0.05$ \citep{rigault2013,rigault2018}, in agreement with rate
+expectations \citep{mannucci2006,rodney2014}, then, we find in
+\cite{rigault2018}:
 \begin{align}
-    \delta\z & = \left{K^{-1}\times \LF{1+z}^{-\f} +1\right}^{-1}\\
-    \psi\z & = \LF{K\times \LF{1+z}^{+\f} + 1}\um
+    \label{eq:delta}
+    \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
+                  \mathrm{or,}\\
+    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1},
 \end{align}
 
-%\begin{figure*}
-%    \centering
-%    \subfigure[Raw data]{\label{fig:snf_data}
-%        \includegraphics[width=8.5cm]{Article_figures/BiGaussian_onlydata.pdf}}
-%    \subfigure[First implemented model]{\label{fig:3G2M2SSNF}
-%        \includegraphics[width=8.5cm]{Article_figures/BiGaussian.pdf}}
-%    \caption{Stretch des supernovae étudiées par la collaboration SNF en
-%    fonction de log(LsSFR). La couleur représente la probabilité pour une
-%supernova d'être issue d'un jeune progéniteur.}
-%    \label{fig:mod_first}
-%\end{figure*}
+where $K=0.87$ given the fraction of young in the SNfactory sample. This
+modelisation is comparable to the evolution predicted by \cite{childress2014}
+based on SN rates in galaxies depending on their quenching time as a function of
+their stellar mass.
 
-The goal is to find how the stretch depends on the LsSFR, distinguishing old and
-young SNe which fractions evolve with the redshift, in order to have an
-analytical law for the mean redshift evolution of the stretch. The measured data
-on which we based our first model is plotted in figure \ref{fig:mod_first};
-based of the shape of the $x_1$ vs LsSFR cloud, we implemented the following
-model:
-\begin{itemize}
-    \item \textbf{young}: a gaussian of mean $\m_1$ and standard deviation
-        $\s_1$, namely $\Nc_1 \equiv \Nc \LF{\m_1,\s_1}$;
-    \item \textbf{old}: a linear combination between $\Nc_1$ and another
-        gaussian $\Nc_2 \equiv \Nc \LF{\m_2,\s_2}$
-\end{itemize}
-As such, the probability to observe a young SNe Ia labeled "i" with a stretch
-$x_1^i$ and error $\mathrm{d} x_1^i$ is
+\subsection{With LsSFR data}
+\label{sec:modelpy}
+
+In \cite{rigault2018} we present that the SN stretch is strongly correlated with
+the LsSFR, which traces age. We present this correlation in
+fig.~\ref{fig:stretchlssfr}. To model the evolution of the SN stretch as a
+function of redshift given our aforementioned model of the evolution of the
+fraction of young (old) SNeIa, we need to model the SN stretch distribution of
+each age sample. Given the structure of the stretch--LsSFR scatter plot show
+fig.~\ref{fig:stretchlssfr}, we define our model as follows: for the young
+population the underlying stretch distribution is modeled as $\mathcal{N}(\mu_1,
+\sigma_1)$, i.e. a normal distribution centered on $\mu_1$ with a $\sigma_1$
+width ; the old population stretch distribution is  modeled as $a\times
+\mathcal{N}(\mu_1, \sigma_1) + (1-a)\times \mathcal{N}(\mu_2, \sigma_2)$, i.e. a
+gaussian mixture where one mode is the same as the young population one.  The
+estimation of the 5 free parameters
+($\theta\equiv{\mu_1,\mu_2,\sigma_1,\sigma_2,a}$) of the model is given by
+minimizing:
 \begin{equation}
-    p \LF{x_1^i, \mathrm{d} x_1{}^i|\mu_1,\sigma_1} = \mathcal{N} \left(\mu_1,
-    \sqrt{\sigma_1{}^2 + \mathrm{d} x_1^i{}^2}\right)(x_1^i)
+    \label{eq:likelihood}
+    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{\theta\ ;
+                     \mathrm{d}x^{i}_{1}, p(y)^{i}}\right),
 \end{equation}
-and for an old one:
+with
 \begin{align}
-    p(x_1^i, \mathrm{d} x_1^i | \mu_{1,2}, \sigma_{1,2}, a) = a & \times
-    \mathcal{N} \left(\mu_1, \sqrt{\sigma_1{}^2 + \mathrm{d} x_1^i{}^{2}}\right) (x_1^i) \ + \\ (1-a) & \times \mathcal{N}
-    \left(\mu_2, \sqrt{ \sigma_2{}^2 + \mathrm{d} x_1^i{}^{2}}\right) (x_1^i), \nonumber 
+    \label{eq:likelihoodsnf}
+    \prob{x^{i}_{1}}{ \theta\ ; \mathrm{d}x^{i}_{1}, p(y)^{i}} = 
+    p(y)^{i}\,\times\,&\mathcal{N}\left(\mu_1, \sqrt{\sigma_1{}^{2}
+    +\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1})\,+\nonumber\\
+    (1-p(y)^{i}) \times [a\,\times\,& \mathcal{N}\left(\mu_1,
+    \sqrt{\sigma_1{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1})\,
+    + \nonumber\\
+    (1-a)\,\times\,&\mathcal{N}\left(\mu_2, \sqrt{\sigma_2{}^{2}
+    +\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1})\,],
+\end{align}
+where $i$ is the index of the SNIa, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
+$p(y)^{i}$ are the salt2.4 stretch, error of the salt2.4 stretch and the
+probability that the SN is prompt, respectively.  In practice, since "$a$" is
+defined between 0 and 1, we fit for $\alpha$ such that
+$a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.
+
+\begin{figure*}
+    \label{fig:stretchlssfr}
+    \centering
+      \includegraphics[width=\linewidth]{Article_figures/BiGaussian_hist.pdf}
+    \caption{SNF data on which we based our models of stretch distribution
+    between young and old}
+\end{figure*}
+
+The fit results are shown table~\ref{tab:modelresults} and the best fit model is
+illustrated in fig.~\ref{fig:stretchlssfr}.
+
+\begin{table*}
+    \centering
+    \caption{Best fit value of the stretch distribution model. The "SNf" results
+    are derived using eq.~\ref{eq:likelihoodsnf} solely using the SNf data for
+which we have $p(y)$. The "All" and "All(cons)" results are derived using
+eq.~\ref{eq:likelihood} without the age information (see
+section~\ref{sec:modelnopy}) ; "All(cons)" correspond to the "complete sample"
+using conservative redshift cut (see section~\ref{sec:sample}).}
+    \label{tab:modelresults}
+    \begin{tabular}{l c c c c c}
+    \hline\hline\\[-0.8em]
+        Sample & $\mu_1$  & $\sigma_1$ &$\mu_2$  & $\sigma_2$ & a \\[0.15em]
+        \hline\\[-0.8em]
+        SNf & $0.41 \pm 0.08$ & $0.55 \pm 0.06$ & $-1.38 \pm 0.10$ &
+        $0.44 \pm 0.08$ & $0.48^{+0.08}_{-0.08}$ \\[0.15em]
+        All & $0.43 \pm 0.09$ & $0.58 \pm 0.05$ & $-1.05 \pm 0.28$ &
+        $0.63 \pm 0.13$ & $0.51^{+0.09}_{-0.10}$ \\[0.15em]
+        All(cons) & $0.38 \pm 0.05$ & $0.60 \pm 0.06$ & $-1.26 \pm 0.13$
+                  & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\[0.15em]
+        \hline
+    \end{tabular}
+\end{table*}
+
+When combining our prediction of the evolution of young SNeIa as a function of
+redshift eq.~\ref{eq:delta}, and the stretch distribution of the both the young
+and the old SNeIa, our model for the underlying distribution of stretch
+$\Delta\left(x_1|z\right)$ as a function of redshift is:
+\begin{align}
+    \label{eq:stretchz}
+    \Delta\left(x_1\,|\,z \right) = 
+    &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1)\,+\nonumber\\
+    (1-&\,\delta(z)) \times \left[a\times\mathcal{N}(\mu_1,\sigma_1)
+    + (1-a)\times\mathcal{N}(\mu_2,\sigma_2)\right]
 \end{align}
 
-\noindent where $a$ is the relative amplitude between the two gaussians.
-Finally, the normalized stretch distribution at a given $z$
-$\Delta\left(x_1|z\right)$ is
-the weigthed sum of both young and old stretch distributions given their
-relative fraction $\delta(z)$:
+\subsection{Without LsSFR data}
+\label{sec:modelnopy}
+
+Given eq.~\ref{eq:stretchz}, we can extend the analysis by fitting the free
+parameters of the model ($\theta\equiv{\mu_1,\mu_2,\sigma_1,\sigma_2,a}$)
+assuming the evolution of the fraction of young SNeIa as a function of redshift
+given by $\delta(z)$ and given, this time, $x_1^{i}$,  $\mathrm{d}x^{i}_{1}$ and
+$z^{i}$, the stretch, error on the stretch and redshift of any given SN $i$. In
+practice, it means that we are still minimizing for the same
+eq.~\ref{eq:likelihood}, but replacing $p(y)^{i}$ by $\delta(z^{i})$
+eq.~\ref{eq:likelihoodsnf}.
+
+The fit results when using the conservative redshift cuts or not are shown
+table~\ref{tab:modelresults} and the best fit models are illustrated in
+fig.~\ref{fig:modelall}.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/stretchevol_all_vs_snf_maglim-cuts.pdf}
+    \caption{Result of the model fitted on all the surveys' data, cut at the
+    conservative $z_{\mathrm{max}}$ in transparent and non-conservative
+$z_{\mathrm{max}}$ in full markers. The number of SNe Ia in each bin is
+indicated just above the x-axis in transparent text; between brackets are the
+numbers for the conservative cuts.}
+    \label{fig:modelall}
+\end{figure*}
+
+\subsection{Other modelings}
+\label{sec:othermodel}
+
+In section~\ref{sec:modelpy} and \ref{sec:modelnopy} we have modeled the
+underlying stretch distribution as a single gaussian for the young SNe Ia and a
+combination of two gaussians for the old SNe Ia population, i.e., the same as
+the young-population one plus another for the fast declining SNeIa that seem to
+only exist in old populations.
+
+\cite{howell2007} used a simpler modelisation: the underlying stretch of each
+age population is a single gaussian. This rises the question of the accuracy of
+our modeling choice. To test this, we have implemented a suite of extra
+parametrisation that we also fit to the data following the procedure described
+section~\ref{sec:modelnopy}. Namely we consider:
+\begin{itemize}
+
+    \item "Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$": base model but
+        where the high-stretch Gaussian of the old population is not set as that
+        of the young population ;
+
+    \item "Base$-(\sigma_2)$": base model but where $\sigma_2$ is fixed at the
+        same value as $\sigma_1$, since both width seem not significantly
+        different in Table~\ref{tab:modelresults};
+
+    \item "Howell": one single Gaussian per age group as in \citep{howell2007},
+        but with the $\delta(z)$ factor.
+
+\end{itemize}
+
+Furthermore, since we aim at probing the SN redshift drift, we are also fitting
+the "Base" model and the three aforementioned extensions by fixing the
+population drift parameter $\delta(z)$ to a free parameter constant
+$\delta(z)=f$. Doing so enables us to consider two last models. The first is
+modeling the underlying SNe Ia stretch distribution by a simple Gaussian, namely
+"Gaussian (f)". The second is using an asymmetric Gaussian distribution as
+currently done by the BBC formalism introduced by \cite{kessler2017,
+scolnic2018b} for modeling the sample selection functions in SN Cosmology. This
+last model is particularly interesting to test as it is currently used in the
+most recent SN Cosmological analyses \cite{riess2016, riess2018, PSpaper,
+scolnic2018a, DESpaper} \nn{supposed riess2018 to be
+https://ui.adsabs.harvard.edu/abs/2018ApJ...861..126R/abstract}; see further
+discussion concerning this issue in section~\ref{sec:bbc}.
+
+The ability of the model to explain the data is detailed in
+section~\ref{sec:results}.
+
+\section{Results}
+\label{sec:results}
+
+We applied each model on both the conservative and non-conservative sample (cf.
+section \ref{sec:sample}. We want to test whether including an evolution of the
+stretch with the redshift ($\delta(z)$ factor) would better describe the data.
+As the various models presented section~\ref{sec:modeling}, have various degrees
+of freedom. To compare them all, we use the Akaike Information Criterion
+corrected for sample size (AICc), that penalizes extra degrees of freedom to
+avoid over-fitting the data. The AICc is defined as follow:
 \begin{equation}
-    \Delta\left(x_1|z\right) = \delta(z)\times\mathcal{N}_1 +
-    \left(1-\delta(z)\right) \times \left(a\mathcal{N}_1 +
-    (1-a)\mathcal{N}_2\right)
+    \mathrm{AICc} = \mathrm{AIC} + \frac{2k\left(k+1\right)}{n - k - 1}
 \end{equation}
 
-We fitted it on SNf data, giving the results table \ref{tab:val}, with a graphical representation shown figure \ref{fig:snf_model}. For clarity
-with the next models, we named it 3G2M2S$_{\text{SNf}}$ for it has a total of 3
-gaussians but with only 2 means and 2 standard deviations, and has been fitted
-on SNf data only.
+with $\mathrm{AIC} = 2k + \mathcal{L}$, where $k$ is the number of free
+parameters and $\mathcal{L}$ is defined eq~\eqref{eq:likelihood}.  The best
+model is the one with the smaller AICc and the probability for another model to
+be at least as representative of the data as this one is given by:
+\begin{equation}
+    p(\mathrm{other} > \mathrm{best}) = \exp\left(\Delta\mathrm{AICc}/2\right)
+\end{equation}
+The results are summarized in table \ref{tab:comp} and represented figure
+\ref{fig:mod_comp}.
 
-\begin{figure}
+\begin{table*}
     \centering
-    \includegraphics[width=\linewidth]{Article_figures/stretchevol_model_maglim-cuts.pdf}
-    \caption{Result of the model fitted on SNF data and mean of stretch and redshift of the other surveys used for this analysis. Conservative data is shown in transparent markers.}
-    \label{fig:snf_model}
-\end{figure}
+    \caption{Comparison of quality of fit for all the models. Base indicates the
+    models designed for this study. (F) indicates models without an evolution of
+the fraction of young SNe Ia as a function of the redshift.}
+    \label{tab:comp}
+    \begin{tabular}{c c c  c c c c  c c c c}\hline\hline\\[-0.8em]
+        & & & \multicolumn{4}{c}{All SNeIa (569)} &
+        \multicolumn{4}{c}{All SNeIa (conservative ; 422)} \\
+        %\cline{2-6}
+        Name & drift(?) & Free param &
+        $\ln\mathcal{L}$ & $\mathrm{AICc}$ & $\Delta \mathrm{AICc}$ & Proba &
+        $\ln\mathcal{L}$ & $\mathrm{AICc}$ & $\Delta \mathrm{AICc}$ & Proba \\
+        [0.15em]\hline\\[-0.8em]
 
-We implemented and compared 10 models in total, 4 of which have an
-evolution with the redshift from $\delta(z)$, and 6 don't ($\delta(z) = f =
-\mathrm{cst}$). The ones with and evolution are:
+        Base$-(\sigma_2)$ & $\delta(z)$ & 4
+        & 1456.9 & 1464.9 & -- &  --
+        & 1079.9 & 1088.0 & -- & -- \\[0.15em]
+
+        Base & $\delta(z)$ & 5
+        & 1456.7  & 1466.8 & -1.8 & $4.0\times10^{-1}$
+        & 1079.5 & 1089.6 & -1.7 & $4.4\times10^{-1}$\\[0.15em]
+
+        Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$ & $\delta(z)$ & 7
+        & 1453.1  & 1467.3 & -2.4 & $3.1\times10^{-1}$
+        & 1074.5 & 1088.8 & -0.8 & $6.7\times10^{-1}$\\[0.15em]
+
+        Howell & $\delta(z)$ & 4
+        & 1463.3 & 1471.3 & -6.4 & $4.1\times10^{-2}$
+        & 1088.2 & 1096.3 & -8.3 & $1.5\times10^{-2}$\\[0.15em]
+
+        Assymetric & $f$ & 3
+        & 1485.2 & 1491.3 & -26.3 & $1.9\times10^{-6}$
+        & 1101.3 & 1107.4 & -19.4 & $6.2\times10^{-5}$\\[0.15em]
+
+        Howell & $f$ & 5
+        & 1484.2 & 1494.3 & -29.4 & $4.2\times10^{-7}$
+        & 1101.2 & 1111.3 & -23.4 & $8.5\times10^{-6}$\\[0.15em]
+
+        Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$ & $f$ & 8
+        & 1478.9 & 1495.2 & -30.2 & $2.8\times10^{-7}$
+        & 1095.5 & 1111.9 & -23.9 & $6.5\times10^{-6}$\\[0.15em]
+
+        Base & $f$ & 6
+        & 1484.2 & 1496.4 & -31.4 & $1.5\times10^{-7}$
+        & 1101.2 & 1113.4 & -25.4 & $3.0\times10^{-6}$\\[0.15em]
+
+        Gaussian & -- & 2
+        & 1521.8 & 1525.8 & -60.9 & $6.0\times10^{-14}$
+        & 1142.6 & 1146.6 & -58.6 & $1.9\times10^{-13}$ \\[0.15em]
+
+        Base$-(\sigma_2)$ & $f$ & 5
+        & 1521.8 & 1531.9 & -67.0 & $2.9\times10^{-15}$
+        & 1142.6 & 1152.7 & -64.7 & $8.8\times10^{-15}$\\\hline
+
+    \end{tabular}
+\end{table*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/mod_comp.pdf}
+    \caption{$\Delta$AICc of the models fitted on all the surveys' data, cut at
+    the conservative $z_{\mathrm{max}}$ in transparent and non-conservative
+$z_{\mathrm{max}}$ in full markers.}
+    \label{fig:mod_comp}
+\end{figure*}
+
+For the non-conservative part, we find that every model lacking an evolution of
+the stretch with the redshift is systematically worse than those that implement
+it, and that removing the $\sigma_2$ parameter gives better results than letting
+it free, or adding $(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$. However, the
+base model with 2 gaussians for the old population is roughly the same as the
+Howell model with one gaussian per population. As for the conservative part, all
+the models but the asymmetric and the gaussian ones correctly fit the data. No
+fixed model fit better than those with evolution, but some come close. It would
+seem that taking SNe Ia close to us doesn't give us a sample where the evolution
+is visible, justifying the study done in section~\ref{sec:sample}.
+
+\section{discussion}
+
+\subsection{BBC and assumptions on underlying SN property}
+\label{sec:bbc}
+BBC use the asymmetric model. When we do it, we get the table TABLE
+
+\begin{table*}
+    \centering
+    \caption{Fit results of the asymmetric and base models for the 3 surveys
+    having selection effects.}
+    \label{tab:bbc}
+    \begin{tabular}{c c c c c c c c c c}\hline\hline\\[-0.8em]
+    Asymetric & Free param & $\sigma^-$ & $\sigma^-_{\mathrm{err}}$ &
+    $\sigma^+$ & $\sigma^+_{\mathrm{err}}$ & $\mu^0$ & $\mu^0_{\mathrm{err}}$
+               & ln $\mathcal{L}$ & AICc \\\hline
+    SDSS & 3 & 1.25 & 0.08 & 0.39 & 0.07 & 0.69 & 0.10 & 893.9 & 900.0 \\
+    PS1 & 3 & 1.00 & 0.09 & 0.49 & 0.09 & 0.42 & 0.12 & 702.0 & 708.1 \\
+    SNLS & 3 & 1.18 & 0.11 & 0.39 & 0.11 & 0.81 & 0.15 & 594.8 & 600.9 \\\hline
+    Total & 9 & -- & -- & -- & -- & -- & -- & 2190.7 & 2209.0\\\hline
+    \end{tabular}
+
+    \vspace*{12pt}
+
+    \begin{tabular}{c c c c c}\hline\hline\\[-0.8em]
+    Base &  Free param & ln $\mathcal{L}$ & AICc \\\hline
+    SDSS & 0 & 395.2 & 405.6 \\
+    PS1 & 0 & 444.7 & 455.0 \\
+    SNLS & 0 & 254.1 & 264.7 \\\hline
+    Total & 0 & 1094.0 & 1125.4 \\\hline
+    \end{tabular}
+\end{table*}
+
+\section{Conclusion}
+\label{sec:ccl}
+Stretch evolution with the redshift is a thing. Need to see if it has an impact
+on the cosmology though.
+
+\begin{acknowledgements}
+This project has received funding from the European Research Council (ERC) under
+the European Union's Horizon 2020 research and innovation programme (grant
+agreement n 759194 - USNAC).
+
+    % How has received that support for this project ? 
+    %Support has been provided by the Institut Universitaire de France, the CNES, and the region Auvergne-Rhone-Alpes.
+\end{acknowledgements}
+
+\bibliographystyle{aa} % style aa.bst
+\begin{thebibliography}{} 
+% A
+\bibitem[Abbott et al.(2018)]{abbott2018} Abbott, T.~M.~C., Abdalla, F.~B., Alarcon, A., et al.\ 2018, \prd, 98, 043526
+
+\bibitem[Aldering(2004)]{aldering2004} Aldering, G.\ 2004, APS April Meeting Abstracts 2004, V4.003
+
+\bibitem[Astier et al.(2006)]{astier2006} Astier, P., Guy, J., Regnault, N., et al.\ 2006, \aap, 447, 31
+
+\bibitem[Ata et al.(2017)]{ata2017} Ata, M., Kitaura, F.-S., Chuang, C.-H., et al.\ 2017, \mnras, 467, 3993
+
+\bibitem[Aubourg et al.(2008)]{aubourg2008} Aubourg, {\'E}.,
+  Tojeiro, R., Jimenez, R., et al.\ 2008, \aap, 492, 631
+
+% B
+\bibitem[Betoule et al.(2014)]{betoule2014} Betoule, M., Kessler, R., Guy, J., et al.\ 2014, \aap, 568, A22
+
+\bibitem[Bazin et al.(2011)]{bazin2011} Bazin, G., Ruhlmann-Kleider, V., Palanque-Delabrouille, N., et al.\ 2011, \aap, 534, A43
+
+% C
+\bibitem[Campbell et al.(2013)]{campbell2013} Campbell, H., D'Andrea, C.~B., Nichol, R.~C., et al.\ 2013, \apj, 763, 88
+
+\bibitem[Chabanier et al.(2019)]{chabanier2019} Chabanier, S., Millea, M., \& Palanque-Delabrouille, N.\ 2019, \mnras, 489, 2247
+
+\bibitem[Childress et al.(2013)]{childress2013} Childress, M., Aldering, G., Antilogus, P., et al.\ 2013, \apj, 770, 108
+
+\bibitem[Childress et al.(2014)]{childress2014} Childress, M.~J., Wolf, C., \& Zahid, H.~J.\ 2014, \mnras, 445, 1898
+
+\bibitem[Coles, \& Jones(1991)]{coles1991} Coles, P., \& Jones, B.\ 1991, \mnras, 248, 1
+
+% D
+\bibitem[D'Andrea et al.(2011)]{dandrea2011} D'Andrea, C.~B., Gupta, R.~R., Sako, M., et al.\ 2011, \apj, 743, 172
+
+\bibitem[Dilday et al.(2008)]{dilday2008} Dilday, B., Kessler, R., Frieman, J.~A., et al.\ 2008, \apj, 682, 262
+
+% E
+% F
+\bibitem[Freedman et al.(2019)]{freedman2019} Freedman, W.~L., Madore, B.~F., Hatt, D., et al.\ 2019, \apj, 882, 34
+
+% G
+\bibitem[Gupta et al.(2011)]{gupta2011} Gupta, R.~R., D'Andrea, C.~B., Sako, M., et al.\ 2011, \apj, 740, 92
+
+\bibitem[Guy et al.(2010)]{guy2010} Guy, J., Sullivan, M., Conley, A., et al.\ 2010, \aap, 523, A7
+
+%\bibitem[Graziani et al.(2019)]{graziani2019} Graziani, R., Courtois, H.~M., Lavaux, G., et al.\ 2019, \mnras, 488, 5438
+
+% H
+\bibitem[Hamuy et al.(1996)]{hamuy1996} Hamuy, M., Phillips, M.~M., Suntzeff, N.~B., et al.\ 1996, \aj, 112, 2391
+
+\bibitem[Hamuy et al.(2000)]{hamuy2000} Hamuy, M., Trager, S.~C., Pinto, P.~A., et al.\ 2000, \aj, 120, 1479
+
+\bibitem[Howell et al.(2007)]{howell2007} Howell, D.~A., Sullivan, M., Conley, A., et al.\ 2007, \apjl, 667, L37
+
+\bibitem[Howell et al.(2009)]{howell2009} Howell, D.~A., Sullivan, M., Brown, E.~F., et al.\ 2009, \apj, 691, 661
+
+% I 
+% J
+\bibitem[Jones et al.(2015)]{jones2015} Jones, D.~O., Riess, A.~G., \& Scolnic, D.~M.\ 2015, \apj, 812, 31
+
+\bibitem[Jones et al.(2018)]{jones2018} Jones, D.~O., Riess, A.~G., Scolnic, D.~M., et al.\ 2018, \apj, 867, 108
+
+\bibitem[Jones et al.(2019)]{jones2019} Jones, D.~O., Scolnic, D.~M., Foley, R.~J., et al.\ 2019, \apj, 881, 19
+
+% K
+\bibitem[Kelly et al.(2010)]{kelly2010} Kelly, P.~L., Hicken, M., Burke, D.~L., et al.\ 2010, \apj, 715, 743
+
+\bibitem[Kessler et al.(2009)]{kessler2009} Kessler, R., Becker, A.~C., Cinabro, D., et al.\ 2009, \apjs, 185, 32
+
+\bibitem[Kessler, \& Scolnic(2017)]{kessler2017} Kessler, R., \& Scolnic, D.\ 2017, \apj, 836, 56
+
+\bibitem[Knox, \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv e-prints, arXiv:1908.03663
+
+% L
+\bibitem[Lampeitl et al.(2010)]{lampeitl2010} Lampeitl, H., Smith, M., Nichol, R.~C., et al.\ 2010, \apj, 722, 566
+
+% M
+\bibitem[Mannucci et al.(2005)]{mannucci2005} Mannucci, F.,
+  Della Valle, M., Panagia, N., et al.\ 2005, \aap, 433, 807 
+\bibitem[Mannucci et al.(2006)]{mannucci2006} Mannucci, F.,
+  Della Valle, M., \& Panagia, N.\ 2006, \mnras, 370, 773
+
+\bibitem[Maoz et al.(2014)]{maozmannucci2014} Maoz, D., Mannucci,
+  F., \& Nelemans, G.\ 2014, \araa, 52, 107
+
+% N
+\bibitem[Neill et al.(2006)]{neill2006} Neill, J.~D., Sullivan, M., Balam, D., et al.\ 2006, \aj, 132, 1126
+
+\bibitem[Neill et al.(2009)]{neill2009} Neill, J.~D., Sullivan, M., Howell, D.~A., et al.\ 2009, \apj, 707, 1449
+
+% O
+% P
+\bibitem[Pan et al.(2014)]{pan2014} Pan, Y.-C., Sullivan, M., Maguire, K., et al.\ 2014, \mnras, 438, 1391
+
+\bibitem[Perlmutter et al.(1999)]{perlmutter1999} Perlmutter, S., Aldering, G., Goldhaber, G., et al.\ 1999, \apj, 517, 565
+
+\bibitem[Perrett et al.(2010)]{perrett2010} Perrett, K., Balam, D., Sullivan, M., et al.\ 2010, \aj, 140, 518
+
+\bibitem[Planck Collaboration et al.(2016)]{planck2016} Planck Collaboration, Ade, P.~A.~R., Aghanim, N., et al.\ 2016, \aap, 594, A13
+
+\bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T., et al.\ 2019, \prl, 122, 221301
+
+\bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration, Aghanim, N., Akrami, Y., et al.\ 2018, arXiv e-prints, arXiv:1807.06209
+
+% Q
+% R
+\bibitem[Reid et al.(2019)]{reid2019} Reid, M.~J., Pesce, D.~W., \& Riess, A.~G.\ 2019, arXiv e-prints, arXiv:1908.05625
+
+\bibitem[Rest et al.(2014)]{rest2014} Rest, A., Scolnic, D., Foley, R.~J., et al.\ 2014, \apj, 795, 44
+
+\bibitem[Riess et al.(1998)]{riess1998} Riess, A.~G., Filippenko, A.~V., Challis, P., et al.\ 1998, \aj, 116, 1009
+
+\bibitem[Riess et al.(2009)]{riess2009} Riess, A.~G., Macri, L., Casertano, S., et al.\ 2009, \apj, 699, 539
+
+\bibitem[Riess et al.(2016)]{riess2016} Riess, A.~G., Macri, L.~M., Hoffmann, S.~L., et al.\ 2016, \apj, 826, 56
+
+\bibitem[Riess et al.(2018)]{riess2018} Riess, A.~G., Casertano, S., Yuan, W., et al.\ 2018, \apj, 861, 126
+
+\bibitem[Riess et al.(2019)]{riess2019} Riess, A.~G., Casertano, S., Yuan, W., et al.\ 2019, \apj, 876, 85
+
+\bibitem[{Rigault {et~al.}(2013)}]{rigault2013}
+Rigault, M., Copin, Y., Aldering, G., {et~al.} 2013, \aap, 560, A66
+
+\bibitem[Rigault et al.(2015)]{rigault2015} Rigault, M., Aldering, G., Kowalski, M., et al.\ 2015, \apj, 802, 20
+
+\bibitem[Rigault et al.(2018)]{rigault2018} Rigault, M.,
+  Brinnel, V., Aldering, G., et al.\ 2018, arXiv:1806.03849
+
+\bibitem[Rodney et al.(2014)]{rodney2014} Rodney, S.~A.,
+  Riess, A.~G., Strolger, L.-G., et al.\ 2014, \aj, 148, 13
+
+% S
+\bibitem[Sako et al.(2008)]{sako2008} Sako, M., Bassett, B., Becker, A., et al.\ 2008, \aj, 135, 348
+
+\bibitem[Scannapieco \& Bildsten(2005)]{scannapieco2005} Scannapieco, E., \& Bildsten, L.\ 2005, \apjl, 629, L85
+
+\bibitem[Scolnic et al.(2014)]{scolnic2014} Scolnic, D., Rest, A., Riess, A., et al.\ 2014, \apj, 795, 45
+
+\bibitem[Scolnic et al.(2018)]{scolnic2018a} Scolnic, D.~M., Jones, D.~O., Rest, A., et al.\ 2018a, \apj, 859, 101
+
+\bibitem[Scolnic et al.(2018)]{scolnic2018b} Scolnic, D.~M., Lochner, M., Gris, P., et al.\ 2018, arXiv e-prints, arXiv:1812.00516
+
+\bibitem[Scolnic et al.(2019)]{scolnicastro2020} Scolnic, D., Perlmutter, S., Aldering, G., et al.\ 2019, Astro2020: Decadal Survey on Astronomy and Astrophysics, 2020, 270
+
+\bibitem[Sullivan et al.(2006)]{sullivan2006} Sullivan, M., Le  Borgne, D., Pritchet, C.~J., et al.\ 2006, \apj, 648, 868
+
+\bibitem[Sullivan et al.(2010)]{sullivan2010} Sullivan, M., Conley, A., Howell, D.~A., et al.\ 2010, \mnras, 406, 782
+
+% T
+\bibitem[Tasca et al.(2015)]{tasca2015} Tasca, L.~A.~M., Le F{\`e}vre, O., Hathi, N.~P., et al.\ 2015, \aap, 581, A54
+
+% U 
+% V
+% W
+\bibitem[Wong et al.(2019)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F., et al.\ 2019, arXiv e-prints, arXiv:1907.04869
+
+% X
+% Y
+ Z
+\end{thebibliography}
+
+\end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%  BACKUP 
+%
+%%%%%%%%%%%%%%%%%%%%%%%%
+
+%For clarity with the next models, we named it 3G2M2S$_{\text{SNf}}$ for it has a total of 3 gaussians but with only 2 means and 2 standard deviations, and has been fitted on SNf data only.
+
+%We implemented and compared 10 models in total, 4 of which have an
+%evolution with the redshift from $\delta(z)$, and 6 don't ($\delta(z) = f =\mathrm{cst}$). The ones with and evolution are:
 \begin{itemize}
@@
     \centering
-    \caption{Valeurs des paramètres pour différents modèles. En rouge les
-    données aberrantes.}
+    \caption{Values of the best-it parameters for all the models.
+In red are the non-coherent ones which coincides with the models without an evolution of the fraction of young SNe Ia with the redshift.}
     \label{tab:val}
@@
 \end{tabular} \bigbreak
-
 \end{table*}
 
-To test the coherence of this model, we also fitted it using all the data from the surveys indistinctly. These results are shown figures \ref{fig:model_all} and \ref{fig:model_all_cons}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{Article_figures/stretchevol_all_vs_snf_maglim_sup.pdf}
-    \caption{Result of the model fitted on all surveys data cut at the non-conservative $z_{\mathrm{max}}$.
-We chose the bins to compute the means of stretch and redshift by making them equal in number of SNe Ia.}
-    \label{fig:model_all}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{Article_figures/stretchevol_all_vs_snf_maglim_inf.pdf}
-    \caption{Result of the model fitted on all surveys data cut at the conservative $z_{\mathrm{max}}$.
-We chose the bins to compute the means of stretch and redshift by making them equal in number of SNe Ia.}
-    \label{fig:model_all}
-\end{figure}
-
-\section{Results}
-To compare all these models, we use the Akaike Information Criterion corrected
-for sample size (BURNHAM 2002), that penalises the increase of free parameters
-in order to discourage overfitting:
-\begin{equation}
-    \mathrm{AICc} = \mathrm{AIC} + \LR{2k\LF{k+1}}{n - k - 1}
-\end{equation}
-with $\mathrm{AIC} = 2k - 2\ln\LF{\mathcal{L}}$, where $k$ is the number of free
-parameters and $\Lc$ the likelihood. The probability for a model to be as
-representative as the "best" one is given by:
-\begin{equation}
-    p(\mathrm{other} > \mathrm{best}) = \exp\LF{\D\mathrm{AICc}/2}
-\end{equation}
-The results are showed table \ref{tab:comp}. We find that every model lacking an
-evolution of the stretch with the redshift is systematically worse than those
-that implement it.
-
-\begin{table*}[htbp!]
-    \centering
-    \caption{Comparaison des modèles. NR représente les modèles implémentés
-             durant ce stage. (F) indique les modèles pour lesquels il n'y a pas
-             d'évolution de la fraction de SNe~Ia jeunes et vieilles en fonction
-             du redshift.}
-    \label{tab:comp}
-    \begin{tabular}{c c c c c c c}\hline\hline
-
-        Name & Description & Free param &
-        $\ln\mathcal{L}$ & $AICc$ & $\D AICc$ & Proba \\\hline
-
-        3G2M1S & NR 1S & 4 & 1815 & 1823 & 0.0 & $1.0$ \\
-
-        3G2M2S & NR 2S & 5 & 1815  & 1825 & -2.0 & $3.6\times10^{-1}$ \\
-
-        2G2M2S & Howell & 4 & 1818  & 1826 & -3.4 & $1.8\times10^{-1}$ \\
-
-        3G3M3S & NR 3S & 7 & 1812  & 1826 & -3.6 & $1.6\times10^{-1}$ \\
-
-        3G3M3SF & NR 3S (F) & 8 & 1813 & 1829 & -6.3  & $4.3\times10^{-2}$
-        \\
-
-        2G2M2SF & Howell (F) & 5 & 1823  & 1833 & -9.9 & $7.0\times10^{-3}$
-        \\
-
-        3G2M1SF & NR 1S (F) & 5 & 1823  & 1833 & -10.4 & $5.5\times10^{-3}$
-        \\
-
-        3G2M2SF & NR 2S (F) & 6 & 1823  & 1835 & -12.0 & $2.5\times10^{-3}$
-        \\
-
-        1G1M2S & Kessler (F) & 3 & 1837  & 1843 & -20.2 & $4.1\times10^{-5}$
-        \\
-
-        1G1M1S & 1 gauss. (F) & 2 & 1872 & 1876 & -53.5 &
-        $2.4\times10^{-12}$\\\hline
-
-    \end{tabular}
-\end{table*}
-
-\section{Conclusion}
-Stretch evolution with the redshift is a thing. Need to see if it has an impact
-on the cosmology though.
-
-\begin{acknowledgements}
-    This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement n 759194 - USNAC).
-Support has been provided by the Institut Universitaire de France, the CNES, and the region Auvergne-Rhone-Alpes.
-\end{acknowledgements}
-
-\bibliographystyle{aa} % style aa.bst
-\begin{thebibliography}{} 
-% A
-\bibitem[Abbott et al.(2018)]{abbott2018} Abbott, T.~M.~C., Abdalla, F.~B., Alarcon, A., et al.\ 2018, \prd, 98, 043526
-
-\bibitem[Ata et al.(2017)]{ata2017} Ata, M., Kitaura, F.-S., Chuang, C.-H., et al.\ 2017, \mnras, 467, 3993
-
-% B
-% C
-\bibitem[Chabanier et al.(2019)]{chabanier2019} Chabanier, S., Millea, M., \& Palanque-Delabrouille, N.\ 2019, \mnras, 489, 2247
-\bibitem[Coles, \& Jones(1991)]{coles1991} Coles, P., \& Jones, B.\ 1991, \mnras, 248, 1
-
-% D
-% E
-% F
-\bibitem[Freedman et al.(2019)]{freedman2019} Freedman, W.~L., Madore, B.~F., Hatt, D., et al.\ 2019, \apj, 882, 34
-
-% G
-
-%\bibitem[Graziani et al.(2019)]{graziani2019} Graziani, R., Courtois, H.~M., Lavaux, G., et al.\ 2019, \mnras, 488, 5438
-
-% H
-% I 
-% J
-\bibitem[Jones et al.(2018)]{jones2018} Jones, D.~O., Riess, A.~G., Scolnic, D.~M., et al.\ 2018, \apj, 867, 108
-
-% K
-
-\bibitem[Knox, \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv e-prints, arXiv:1908.03663
-
-% L
-
-% M
-% N
-% O
-% P
-\bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T., et al.\ 2019, \prl, 122, 221301
-
-\bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration, Aghanim, N., Akrami, Y., et al.\ 2018, arXiv e-prints, arXiv:1807.06209
-
-% Q
-% R
-\bibitem[Reid et al.(2019)]{reid2019} Reid, M.~J., Pesce, D.~W., \& Riess, A.~G.\ 2019, arXiv e-prints, arXiv:1908.05625
-
-\bibitem[Riess et al.(2016)]{riess2016} Riess, A.~G., Macri, L.~M., Hoffmann, S.~L., et al.\ 2016, \apj, 826, 56
-
-\bibitem[Riess et al.(2019)]{riess2019} Riess, A.~G., Casertano, S., Yuan, W., et al.\ 2019, \apj, 876, 85
-
-\bibitem[{Rigault {et~al.}(2013)}]{rigault2013}
-Rigault, M., Copin, Y., Aldering, G., {et~al.} 2013, \aap, 560, A66
-
-\bibitem[Rigault et al.(2015)]{rigault2015} Rigault, M., Aldering, G., Kowalski, M., et al.\ 2015, \apj, 802, 20
-
-\bibitem[Rigault et al.(2018)]{rigault2018} Rigault, M.,
-  Brinnel, V., Aldering, G., et al.\ 2018, arXiv:1806.03849
-% S
-\bibitem[Scolnic et al.(2018)]{scolnic2018} Scolnic, D.~M., Jones, D.~O., Rest, A., et al.\ 2018, \apj, 859, 101
-
-% T
-% U 
-% V
-% W
-\bibitem[Wong et al.(2019)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F., et al.\ 2019, arXiv e-prints, arXiv:1907.04869
-
-% X
-% Y
- Z
-\end{thebibliography}
-
-\end{document}
+To test the coherence of this model, we also fitted it using all the data from the surveys indistinctly. These results are shown figure \ref{fig:model_all}.

### 2020-01-27 Nicolas.tex
Big update. Version 1.2.

@@
 \begin{document}
-\title{Redshift evolution of the SN stretch distribution}
+%\title{Evolution of the Type Ia Supernova Underlying Stretch Distribution as a Function of Redshift}
+\title{Redshift Evolution of the Type Ia Supernova Population Stretch Distribution}
 
@@
 
-\author{N. Nicolas\inst{1}
-    \and M. Rigault\inst{2}
+\author{N. Nicolas
+\thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
+\inst{1}
+    \and M. Rigault
+\thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
+\inst{2}
     \and R. Graziani\inst{2}     
@@
 
-\abstract{Type Ia supernovae (SNe Ia) allow for the construction of the Hubble
+\abstract{\mr{ABSTRACT NOT DONE YET.}
+Type Ia supernovae (SNe Ia) allow for the construction of the Hubble
     diagram, giving us information about the Universe's expansion and its
@@
     uncertainties are now starting to be limiting in our ability to measure
-those parameters. In particular, the physics of SNe Ia is still mostly unknown,
-and is thought not to change in time/with the redshift.}
+    those parameters. In particular, the physics of SNe Ia is still mostly
+    unknown, and is thought not to change in time/with the redshift.}
     {In an attempt to reduce those uncertainties, we try to find an empirical
@@
     {We started by getting a complete sample representing all of the stretch
-        distribution that Nature can give us, before using LsSFR measurements,
-        an age tracer which evolution with redshift is known, that has been
-        shown to have a strong correlation with the stretch. We compare their
-        AICc, an estimator of the relative quality of statistical models that
-    includes the number of free parameters, to determine which ones describe
-best the data.}
+        distribution that Nature can give us, before using LsSFR measurements, an
+        age tracer which evolution with redshift is known, that has been shown
+        to have a strong correlation with the stretch. We compare their AIC, an
+    estimator of the relative quality of statistical models that includes the
+number of free parameters, to determine which ones describe best the data.}
     {Models with an evolution of the stretch with the redshift have a better
-    AICc than the ones without.}
+    AIC than the ones without.}
     {We find that implementing these models allows us to fit the data better
@@
 \citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
-Hubble Constant (H$_0$) provided one can calibrate their absolute magnitude
+Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
 \citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
-the SNeIa are anchored on Cepheids \citep[the SH0ES
-project][]{riess2009,riess2016} is $\sim5~\sigma$ higher than what is predicted
+the SNe~Ia are anchored on Cepheids \citep[the SH0ES
+project][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
 by cosmic microwave background (CMB) data measured by Planck assuming the
-standard $\L$CDM or when the SN luminosity is anchored at intermediate redshift
-by the baryon acoustic oscillation (BAO) scale \citep{riess2019, reid2019,
-planck2016, feeney} \nn{changed ried to reid; changed planck to planck2016;
-feeney = https://tinyurl.com/vg6uufm ?}. While using the tip of the red giant
-branch technique in place of the Cepheids seem to favor lower values of H$_0$
-\citep{freedman2019, recalibrationpaper} \nn{changed "freedman" to
-"freedman2019" ; recalibrationpaper = ?} time delay measurements from strong
-lensing seem to also favor high H$_0$ values \citep{wong2019}.
+standard $\L$CDM\nn{,} or when the SN luminosity is anchored at intermediate
+redshift by the baryon acoustic oscillation (BAO) scale
+\citep{riess2019,reid2019,planck2016, feeney2019}.  While using the tip of the
+red giant branch technique in place of the Cepheids seem to favor an
+\mr{intermediate} value of H$_0$ \citep{freedman2019}, time delay measurements
+from strong lensing seem to also favor high H$_0$ values \citep{wong2019}.
 
@@
 fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
-tension when accounting for all other probes and the current most promising
-scenario appears to be a burst of expansion at the matter-radiation decoupling
-moment caused by a (fine tuned) early dark energy \citep{poulin2019} \nn{changed
-"poulin" to "poulin 2019"}.
+tension when accounting for all other probes \citep{knox2019} and the current
+most promising scenario appears to be a burst of expansion at the
+matter-radiation decoupling moment caused by a (fine tuned) early dark energy
+\citep{poulin2019}.
 
 Alternatively, systematic effects affecting one or several of the aforementioned
-analysis could also explain the tension. In \cite{rigault2015} we suggested that
-SNe Ia from the Cepheid calibrator sample differ significantly to the one from
-the Hubble flow sample that enables to derive H$_0$. Indeed, Cepheid are very
-young stars and the existence of Cepheids in a galaxy implies that the host is
-star forming. In \citep{rigault2015} we claimed that a more than 90\% of the
-calibrator SNe Ia arise from a young-progenitor population while this population
-only accounts for about half of the Hubble flow sample. Following
-\cite{sullivan2010} and \citep{rigault2013} \nn{should use cite or citep?}, we
-further showed that young and old progenitor SNe Ia have a different magnitude
-of about 0.15 mag. This has been confirmed for modern standardization technique
-\citep[SALT2.4][]{guy2010,betoule2014} with high significance ($\sim 6\sigma$)
-in \citep{rigault2018} as well as independently by \cite{roman2019} \nn{not
-found on ADS: https://tinyurl.com/uemattg} at $\sim 7\sigma$. As detailed in
-\citep{rigault2015}, if the magnitude difference between young and old SNe Ia is
-indeed of $0.15$ mag and if SNe Ia from the calibrator sample is fully dominated
-by young SNeIa, the effect on H$_0$ of the existence of these two SNe Ia
-populations is of 3.5\%. Accounting for the fact that another environmental
-effect (the mass-step) is already included in the SH0ES analysis, the net bias
-in the SH0ES analysis is expected closer to 3\% \citep[see table 6
-of][]{rigault2015}.
-
-SHOULD BE SHORTENED A LOT: However, the importance of this astrophysical bias in
-the direct measurement of H$_0$ is still highly debated. We also highlight that
-it could not explain the high value of H$_0$ derived by strong lensing. First,
-\cite{jones2015} \nn{put the ref but nos sure it's the correct one} did not find
-any environmental bias when extending the \citep{rigault2015} study. It was more
-recently reported in \citep{jones2019} \nn{put the ref but nos sure it's the
-correct one} that indeed local host analysis has significant influence on the SN
-magnitude but much more importantly than claimed in \citep{rigault2018} and with
-a connection with the progenitor physics and/or a line of sight effect that is
-unclear. More importantly, \cite{riess2016} attended to mimic the Cepheid sample
-selection function on the Hubble flow sample by removing SNe Ia from the latter
-if they classified their host as early-type or non-star forming. They measured
-that doing so has no effect on the measurement of H$_0$.
-
-This paper is part of a series of papers where we test the validity of the two
-SNe Ia population models developed in \citep{rigault2018}.Briday et al. in prep
-will present how using different environmental tracer is affecting our ability
-to distinguish the two populations, and Rigault et al. in prep will further
-study the connection with H$_0$. Here we analyze predictions concerning the SNe
-Ia redshift drifts: (1) that the distribution of the SNIa stretch, a purely
-intrinsic SNIa property, is evolving as a function of redshift and (2) that this
-population-drift follows the relation that \citep{rigault2018} derived assuming
-that one population is “young“ with a rate proportional to the star formation
-rate in the Universe and the second is "old" with a rate proportional to the
-stellar mass.
-
-The concept of the SNeIa age dichotomy arised when the SNIa rate has first been
-studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
-shown that the relative SNe Ia rate in galaxies could only be explained if two
-populations existed, one young, following the host star formation activity, and
-one old following the host stellar mass (the so called “prompt and delayed“ or
-“A+B“ model). In \cite{rigault2018} we used the specific star formation rate at
-the SN location (Local sSFR or LsSFR) to classify which are the prompts (those
-with a high LsSFR) and which are the delayed (those with low LsSFR). Since the
-first SNe Ia host analysis, the SN stretch has been known to be strongly
-correlated with the SN host properties \citep{hamuy1996, hamuy2000} \nn{put
-hamuy1996 as most cited of https://tinyurl.com/sjlbdz8} and it has been
-extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
-lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
-pan2014} \nn{added all the refs}. Following the “A+B“ model and the connection
-between SN stretch and host properties, \citep{howell2009} \nn{changed
-howell20xx to howell2009
-(https://ui.adsabs.harvard.edu/abs/2009ApJ...691..661H/abstract)} first
-discussed the potential redshift drift of the SN stretch distribution. In this
-paper we revisit this analysis with the most recent SNe Ia dataset and we test
-the validity of the two SNIa age populations model developed by
-\citep{rigault2018}.
+analysis could also explain the tension.  In \cite{rigault2015} we suggested
+that SNe Ia from the Cepheid calibrator sample differs by construction from  the
+Hubble flow as it strongly favors young stellar population environments where
+one could find Cepheids. This selection effect would impact the derivation of
+H$_0$ if SNe~Ia from young and older environments differ in average magnitudes.
+
+For the last decades, numerous analysis have studied the relation between SNe~Ia
+and host properties\nn{,} finding first that the standardized SNe~Ia magnitude
+significantly depends on the host stellar-mass, SNe~Ia from high-mass host
+\nn{being brighter on average}
+\cite[e.g.][]{kelly2010,sullivan2010,childress2013,betoule2014,rigault2018}.
+This mass-step correction is currently used in cosmological analyses
+\citep[e.g.][]{betoule2014,scolnic2018a}, including for deriving H$_0$
+\citep{riess2016,riess2019}. Yet, the underlying connection between the SN and
+the host remains unclear when using global properties such as the host stellar
+mass, which raises the question of the accuracy of the correction. More
+recently, studies have used the local SN environment to probe more direct
+connections between the SN and its environments \citep{rigault2013}\nn{,}
+showing that local age-tracers such as the local specific star formation rate or
+the local color are more strongly correlated with the standardized SN magnitude
+\citep{rigault2018,roman2018}, suggesting the age to be the driving parameter
+underlying the mass-step. If true, this would have a significant impact for
+cosmology because the redshift drift magnitude correction to apply might
+strongly vary \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
+importance of local SN environmental study remains highly debated
+\cite[e.g.][]{jones2015,jones2019} and especially the impact it has on the
+derivation of H$_0$ \cite{jones2015,riess2016,riess2018,rose2019}.
+
+In this paper, we take a step aside to probe the validity of our modeling of the
+SN population that we claim to be constituted of two age-populations
+\citep{rigault2013,rigault2015,rigault2018}: one old and one young, the former
+having on average lower lightcurve stretches and being brighter after
+standardization. We indeed use the correlation between the SN age probed with
+local specific star formation rate in \cite{rigault2018} and the SN stretch to
+model the expected evolution of the underlying SN stretch distribution as a
+function of redshift. This modeling relies on three assumptions: (1) there
+\nn{are indeed} 2 populations of SNe~Ia; (2) the relative fraction of each of
+these populations as a function of redshift follows the age-modeling made in
+\citep{rigault2018} and (3) the underlying distribution of stretch for each
+age-sample can be modeled and do not significantly drift. This paper aims at
+testing this with data from the literature.
+
+% SERIES OF PAPER
+%This paper is part of a series of papers where we test the validity of the two
+%SNe Ia population models developed in \citep{rigault2018}. Briday et al. in
+%prep will present how using different environmental tracer is affecting our
+%ability to distinguish the two populations, and Rigault et al. in prep will
+%further study the connection with H$_0$. Here we analyze predictions concerning
+%the SNe Ia redshift drifts: (1) that the distribution of the SN~Ia stretch, a
+%purely intrinsic SN~Ia property, is evolving as a function of redshift and (2)
+%that this population-drift follows the relation that \citep{rigault2018}
+%derived assuming that one population is “young“ with a rate proportional to the
+%star formation rate in the Universe and the second is "old" with a rate
+%proportional to the stellar mass.
+
+The concept of the SNe~Ia age dichotomy arised when the SN~Ia rate has first
+been studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008}
+have shown that the relative SNe~Ia rate in galaxies could only be explained if
+two populations existed, one young, following the host star formation activity,
+and one old following the host stellar mass (the so called “prompt and delayed“
+or “A+B“ model). In \cite{rigault2018} we used the specific star formation rate
+at the SN location (Local sSFR or LsSFR) to classify which are the prompts
+(those with a high LsSFR) and which are the delayed (those with low LsSFR).
+Furthermore, since the first SNe~Ia host analysis, the SN stretch has been known
+to be strongly correlated with the SN host properties \citep{hamuy1996,
+hamuy2000}, which has been extensively confirmed since \citep[e.g.][]{neill2009,
+sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
+rigault2013, pan2014}. Following the “A+B“ model and the connection between SN
+stretch and host properties, \cite{howell2007} first discussed the potential
+redshift drift of the SN stretch distribution. In this paper we revisit this
+analysis with the most recent SNe~Ia dataset..
 
 We present in section~\ref{sec:sample} the sample we are using for this
-analysis, which is based on the pantheon dataset \citep{scolnic2018a}. We
+analysis, which is based on the Pantheon dataset \citep{scolnic2018a}. We
 discuss the importance of obtaining a "complete" sample, i.e. representative of
@@
 sample.  We then present in section~\ref{sec:modeling} our modeling of the
-distribution of stretch as a function of redshift based on the relative stretch
-distribution of young and old SNeIa. Our results are presented in
-section~\ref{sec:results} where we test if the distribution of stretch does
-evolve as a function of redshift and if the age-model is in good agreement with
-this evolution. We discuss our results and conclude in section~\ref{sec:ccl}.
-
-Throughout the paper we use the Planck 2015 cosmology from the astropy library.
-
-\section{A "complete" Sample}
+distribution of stretch as a function of redshift based stretch distribution of
+young and old SNe~Ia. Our results are presented in section~\ref{sec:results}
+where we test if the SN stretch does evolve as a function of redshift and if the
+age model is in good agreement with this evolution. Consequence for cosmology of
+our results are briefly discussed in section~\ref{sec:bbc} and we conclude in
+section~\ref{sec:ccl}.
+
+%Throughout the paper we use the Planck 2015 cosmology from the astropy library.
+
+\section{A ''complete'' Sample}
 \label{sec:sample}
 
-We base our analysis on the latest SNe Ia compilation: the Pantheon catalog from
-\cite{scolnic2018a}.  \mr{TO BE REPHRASED} Using this dataset, we aim at probing
-if the average stretch, an intrinsic SN Ia property, evolves as a function of
-redshift or not; the underlying question being that of the SNIa population
-redshift drift.  The stretch has been shown to depend on the SN environment star
-formation activity
-\citep[e.g.][]{howell2007,sullivan2010,childress2013,rigault2013,rigault2018}
-and because the star formation rate is about 40 times larger at the high
-redshift end of the pantheon sample than at its lower redshifts
-\citep[e.g.,][and references therein]{tasca2015}, one might think that simply
-comparing the redshifts distribution in a few bins of redshift could be
-sufficient to probe any SN stretch redshift drift.  In practice, selection
-effects are affecting the observed SN stretch distributions, which could
-significantly differ from the true underlying one if these selection effects
-become significant. Indeed, because the observed SNIa magnitude correlates with
-the lightcurve stretches (and colors) the first SNIa that such a survey will
-miss will the reddest and the lowest-stretch ones. Consequently, if selection
-effects is not accounted for, one might confuse true population drift with
-selection effect and conversely.
-
-In a magnitude limited survey, providing sufficient (and random) spectroscopic
-follow-up for acquiring typing and host redshift, the selection effects should
-be negligible bellow a given redshift where all SNe Ia, even the faintest, could
-be observed.  In contrast, targeted surveys have highly complex selection
-functions and we will discard them from our analysis. Fortunately, modern SN
-cosmology sample such has the pantheon dataset are now dominated by magnitude
-limited survey.
+We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
+\cite{scolnic2018a}.  A naive approach to test the SN stretch redshift drift
+would be to simply compare the observed SN stretch distributions in a few bins
+of redshift.  In practice, selection effects are affecting the observed SN
+stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
+with the lightcurve stretch (and color), the first SN~Ia that a
+magnitude-limited survey will miss are the reddest and the lowest-stretch ones.
+Consequently, if selection effects are not accounted for, one might confuse true
+population drift with selection effects, and conversely.
+
+\nn{Assuming} sufficient (and random) spectroscopic follow-up for acquiring
+typing and host redshift, the selection effects of magnitude-limited survey
+should be negligible bellow a given redshift where even the faintest \nn{SN~Ia}
+could be observed.  In contrast, targeted surveys have highly complex selection
+functions and will be discarded from our analysis. Fortunately, modern SN
+cosmology samples such as the pantheon dataset are now dominated by magnitude
+limited surveys.
 
 We present fig.~\ref{fig:maglim} the lightcurve stretch and color distributions
-of SNeIa from these surveys, namely: PanStarrs \citep[PS1][]{ref}, the Sloan
-Digital Sky Survey \citep[SDSS][]{ref} and the SuperNovae Legacy Survey
-\citep[SNLS][]{astier2006}.  An ellipse in the salt2.4 $(x_1, c)$ plan with
-major axis $x_1=3$ and manor axis $c=0.3$ encapsulates the full distribution
-(see also \citealt{bazin2011} and \citealt{campbell2013} for a similar contour
-but using a less conservative $c=0.2$ for the latter). Assuming the SN absolute
-magnitude with $x_1=0$ and  $c=0$ is $M=-19.36$ \citep{kessler2009,scolnic2014}
-we can derive the absolute standardized magnitude at maximum of light along the
-aforementioned ellipse given the standardization coefficient $\alpha=0.16$ and
-$\beta=3.14$ from \cite{scolnic2018a}.  The faintest SNIa is that with
-$x_1=-1.66$ and $c=0.25$ and has an absolute standardized magnitude at peak in
-Bessel-b band of $-18.31\,\mathrm{mag}$. Since we need to see this object
-typically a week before and 10 days after peak to build a suitable lightcurve,
-the effective absolute limiting magnitude therefore is
-$\sim-18.00\,\mathrm{mag}$.  Hence, given the $5\sigma$ point source detection
-magnitude limit of a magnitude limited survey, one can derive the maximum
-redshift above which the faintest SNeIa will start to be missed.
+of SNe~Ia from these surveys, namely: PanStarrs \citep[PS1][]{rest2014}, the
+Sloan Digital Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy
+Survey \citep[SNLS][]{astier2006}. An ellipse in the salt2.4 $(x_1, c)$ plan
+with semi-major axis $x_1=3$ and semi-minor axis $c=0.3$ encapsulates the full
+distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
+\citet{campbell2013} for similar contours, the second using a more conservative
+$c=0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and  $c=0$ is
+$M_0=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute
+standardized magnitude at maximum of light along the aforementioned ellipse
+given the standardization coefficient $\alpha=0.156$ and $\beta=3.14$ from
+\cite{scolnic2018a}. The faintest SN~Ia is that with $x_1=-1.66$ and $c=0.25$
+and has an absolute standardized magnitude at peak in Bessel-b band of
+$-18.31\,\mathrm{mag}$. Since one need to detect this object typically a week
+before and 10 days after peak to build a suitable lightcurve, the limiting
+\nn{standardized} absolute magnitude effectively is approximately
+$-18.00\,\mathrm{mag}$. Hence, given the $5\sigma$ point source detection
+magnitude limit of a magnitude limited survey $m_{lim}$, one can derive the
+maximum redshift above which the faintest SNe~Ia will be missed \nn{using the
+link between apparent magnitude, redshift and absolute magnitude given by:}
+\begin{equation}\label{eq:m_lim}
+    \nn{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\,\text{mag}}
+    < m_{lim}}
+\end{equation}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{Article_figures/zmax_maglim_all.pdf}
-    \caption{SALT2.4 stretch and color lightcurve parameters of SNeIa from the
+    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_all.pdf}
+    \caption{SALT2.4 stretch and color lightcurve parameters of SNe~Ia from the
         SDSS, PS1 and SNLS samples from the pantheon catalog. The individual SN
         data are shown as blue dots and a 2D histogram is shown in gray to
-        highlight point density. The ellipse encapsulating all the SN data is
-        displayed colored by the effective standardized absolute magnitude using
-    the $\alpha$ and $\beta$ standardization coefficients from
-\citep{scolnic2018a}.}
+        highlight point density.  The ellipse ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$)
+    is displayed, colored by the corresponding standardized absolute magnitude
+using the $\alpha$ and $\beta$ coefficients from \citep{scolnic2018a}.}
     \label{fig:maglim}
@@
 
-SNLS typically acquires SNeIa in the redshift range $0.4<z<0.8$. At these
-redshifts the rest-frame Bessel-b band roughly corresponds to the SNLS-$i$
+SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
+redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
 filter that has a 24.8 mag $5\sigma$
@@
 agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.
-Similarly, PS1 observes SNeIa in the range $0.2<z<0.4$, their $g$-band $5\sigma$
-depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$.  Figure~6 of
-\citep{scolnic2018a} suggests a more conservative $z_{lim}$ of 0.27 for the PS1
-catalog.  In similar redshift range, SDSS has a limiting magnitude of 22.5
-\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
-SDSS survey were more sensitive to limited spectroscopic resources.
-\cite{kessler2009} present that during year-1 of SDSS, SNeIa with $r-mag<20.5$
-were favored for spectroscopic follow up, corresponding to the redshift cut at
-$0.15$. For the rest of the SDSS survey, additional spectroscopic resources were
-used, such that \cite{kessler2009} and \cite{dilday2008} show a relative
-completeness up to $z_{lim}=0.2$.  Following these analysis we will use
-$z_{lim}=0.2$ as the baseline SDSS redshift limit for the rest of the analysis.
-We could have chosen more conservative redshift limits for each of these
-surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS and
-$z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}, and we will
-also present results obtained when doing so.  We present the redshift
-distribution of these three surveys in fig.~\ref{fig:cuts}, where we see that
-the redshift limit roughly correspond to the pick of these histograms, as
-expected.
+Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
+$5\sigma$ depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
+agreement with, e.g., Figure~6 of \cite{scolnic2018a}. This figure could also
+suggest a more conservative $z_{lim}$ of 0.27; both will be considered as
+discussed below.  In similar redshift range, SDSS has a limiting magnitude of
+22.5 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However,
+the SDSS surveys were more sensitive to limited spectroscopic resources.
+\cite{kessler2009} \nn{pointed out} that during year-1 of SDSS, SNe~Ia with
+$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to the
+redshift cut at $0.15$. For the rest of the SDSS survey, additional
+spectroscopic resources were used, such that \cite{kessler2009} and
+\cite{dilday2008} show a relative completeness up to $z_{lim}=0.2$.  Following
+these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit for
+the rest of \nn{this study}.  For the rest of the analysis, we will consider
+both cases, the \nn{aforementioned} redshift limits as well as more conservative
+cuts for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$
+for PS1 and $z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}.
+
+The complete sample selection is summarized table~\ref{tab:sample} and the
+redshift distribution of these three surveys in are shown fig.~\ref{fig:cuts}.
+We see in this figure that the redshift limit we have selected roughly
+correspond to the peak of these histograms as expected.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{Article_figures/hist_surveys_cuts.pdf}
-    \caption{Histograms of the 3 surveys on which we applied cuts to get rid of
-    selection effects, as discussed in section \ref{sec:sample}. The SNe Ia
-between the conservative and non-conservative cuts are shown in transparent.}
+    \includegraphics[width=0.9\linewidth]{Article_figures/hist_surveys_cuts.pdf}
+    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
+    SNLS from the pantheon dataset, respectively. The colored part of the
+distribution shows which SNe~Ia has been kept in our analysis for they are
+supposedly free from selection bias (see section~\ref{sec:sample}). The
+full(light) color responds to the conservative(nominal) cut.}
     \label{fig:cuts}
@@
 
-In addition, we use the SNeIa from the nearby supernovae factory
-\citep[SNfactory][]{aldering2004} \nn{added ref assuming
-https://ui.adsabs.harvard.edu/abs/2004APS..APR.V4003A/abstract} published in
-\citep{rigault2018} and that have been discoverd from no-targeted survey (114
-SNeIa, see their section~3 and 4.2.2). As the SNe search were much deeper than
-the spectrophotometric follow-up, SNfactory SNe Ia within a redshift range of
-$0.02<z<0.09$ should also be a random sampling of the underlying SN population.
-Data from \citep{rigault2018} are within these redshift boundaries. The HST
-sample from Pantheon follows the same logic and we therefore kept it entirely
-\citep{FIND BACK THE REF} \nn{lots of choices: https://tinyurl.com/rcsx7cw}.
+In addition, we use the SNe~Ia from the Nearby Supernova Factory
+\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
+been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
+4.2.2). As the searches were much deeper than the spectrophotometric follow-up
+made by SNfactory \nn{(SNf)}, SNe~Ia within a redshift range of $0.02<z<0.09$
+\citep[as in ][]{rigault2018} should also be a random sampling of the underlying
+SN population. The SNf data is particularly useful for studying SN property
+drift as it enable us to have a large SN~Ia sample at $z<0.1$.  The HST sample
+from Pantheon follows the same logic of having a search deeper than the
+follow-up and we therefore kept it entirely \citep{FIND BACK THE REF}, see
+table~\ref{tab:sample}.
 
@@
     \centering
-    \caption{Source surveys of SNeIa used in our "complete" sample. Conservative
-    cuts are indicated in brackets.}
+    \caption{\nn{Source surveys and number of SNe~Ia used in our "complete"
+    sample. Conservative cuts are indicated in brackets.}}
     \label{tab:sample}
@@
     \hline\hline\\[-0.8em]
-        Survey & N$_{\mathrm{SN}}$  & z$_{max}$ \\[0.15em]
+        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
         \hline\\[-0.8em]
-        SNf & 114 & --\\[0.30em]
-        SDSS & 167(82) & 0.20(0.15)\\[0.30em]
-        PS1 & 160(122) & 0.30(0.27) \\[0.30em]        
-        SNLS & 102(78) & 0.60(0.55)\\[0.30em]
-        HST & 26 & --\\[0.30em]
+        SNf & -- & 114\\[0.30em]
+        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
+        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
+        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
+        HST & -- & 26\\[0.30em]
+        Total & -- & 569(422)\\[0.30em]
         \hline
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-%Fraction of random SDSS stretchs being < mean_conservative =  0.299
-%Fraction of random PS1 stretchs being < mean_conservative =  0.254
-%Fraction of random SNLS stretchs being < mean_conservative =  0.417
+%Fraction of random SDSS stretches being < mean_conservative =  0.299
+%Fraction of random PS1 stretches being < mean_conservative =  0.254
+%Fraction of random SNLS stretches being < mean_conservative =  0.417
 
@@
 In \cite{rigault2018} we presented a modeling of the evolution of the fraction
-of prompt and delayed SNeIa as a function of redshift following former work on
-rates and delay time distributions; see e.g. the prompt and delayed SNeIa model,
-aka the "A+B" model,
-\cite{mannucci2005,scannapieco2005,sullivan2006,aubourg2008} and
-\cite{maozmannucci2014} for a review on delay time distributions of SNeIa. In
-short, we assume that the number of prompt SNeIa follows the star formation
-activity in the Universe while the number of delayed SNeIa follows the number of
-Gyr-old stars in the Universe, i.e. the stellar mass. Hence, if we denote
-calling $\delta(z)$ (resp. $\phi(z)$) the fraction of young (resp. old) SNeIa in
-the Universe as a function of redshift, then their ratio is expected to follow
-the evolution of the specific star formation rate which goes as $(1+z)^{2.8}$
-until $z\sim2$ \citep{tasca2015}. Hence since there is  $\sim50\%$ of young
-SNeIa at $z\sim0.05$ \citep{rigault2013,rigault2018}, in agreement with rate
-expectations \citep{mannucci2006,rodney2014}, then, we find in
-\cite{rigault2018}:
+of prompt and delayed SNe~Ia as a function of redshift following former work on
+rates and delay time distributions \citep[e.g.,][]{mannucci2005,
+scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
+In short, we assumed that the number of prompt SNe~Ia follows the star formation
+activity in the Universe while the number of delayed SNe~Ia follows the number
+of Gyr-old stars in the Universe, i.e. the stellar mass. Hence, if we denote
+$\delta(z)$ (resp. $\psi(z)$) the fraction of young (resp. old) SNe~Ia in the
+Universe as a function of redshift, then their ratio $\delta/\psi$ is expected
+to follow the evolution of the specific star formation rate which goes as
+$(1+z)^{2.8}$ until $z\sim2$ \citep[e.g.,][]{tasca2015}. Hence since
+\nn{$\delta(0.05) \sim \psi(0.05)$} \citep{rigault2013,rigault2018,wiseman2020},
+in agreement with rate expectations \citep{mannucci2006,rodney2014}, \nn{and
+$\delta + \psi = 1$,} then we find in \cite{rigault2018}:
 \begin{align}
@@
     \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
-                  \mathrm{or,}\\
-    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1},
+    \mathrm{and}\\
+    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1}
 \end{align}
-
-where $K=0.87$ given the fraction of young in the SNfactory sample. This
-modelisation is comparable to the evolution predicted by \cite{childress2014}
-based on SN rates in galaxies depending on their quenching time as a function of
-their stellar mass.
-
-\subsection{With LsSFR data}
-\label{sec:modelpy}
-
-In \cite{rigault2018} we present that the SN stretch is strongly correlated with
-the LsSFR, which traces age. We present this correlation in
-fig.~\ref{fig:stretchlssfr}. To model the evolution of the SN stretch as a
-function of redshift given our aforementioned model of the evolution of the
-fraction of young (old) SNeIa, we need to model the SN stretch distribution of
-each age sample. Given the structure of the stretch--LsSFR scatter plot show
-fig.~\ref{fig:stretchlssfr}, we define our model as follows: for the young
-population the underlying stretch distribution is modeled as $\mathcal{N}(\mu_1,
-\sigma_1)$, i.e. a normal distribution centered on $\mu_1$ with a $\sigma_1$
-width ; the old population stretch distribution is  modeled as $a\times
-\mathcal{N}(\mu_1, \sigma_1) + (1-a)\times \mathcal{N}(\mu_2, \sigma_2)$, i.e. a
-gaussian mixture where one mode is the same as the young population one.  The
-estimation of the 5 free parameters
+where $K=0.87$. This modelisation is comparable to the evolution predicted by
+\cite{childress2014} based on SN rates in galaxies depending on their quenching
+time as a function of their stellar mass.
+
+\subsection{\nn{Base underlying distribution}}
+\label{sec:basemodel}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
+    \caption{\textit{Main}: Salt2.4 lightcurve stretch ($x_1$) as a function of
+        the local specific star formation rate (LsSFR) for SNf data used in this
+        analysis. The color corresponds to the probability for the SNe~Ia to be
+        above $\log(\mathrm{LsSFR}=-10.82)$, which defines $p(y)$; see
+        \cite{rigault2018}. \textit{Right}: histogram of $p(y)$-weighted
+        marginalization of the SN stretch.  On both panel, we illustrate the
+        Base model fitted on the SNf data. The young and old population
+        contributions are shown in purple and yellow, respectively.}
+    \label{fig:stretchlssfr}
+\end{figure*}
+
+\nn{In \cite{rigault2018} we presented the relation between SNe stretches and
+LsSFR measurements, which trace progenitor age, from the SNf sample, and is
+shown fig.~\ref{fig:stretchlssfr}}. To model the evolution of the SN stretch as
+a function of redshift given our aforementioned model of the evolution of the
+fraction of young and old SNe~Ia with redshift, we need to model the SN stretch
+distribution for each age subsample. Given the structure of the stretch--LsSFR
+scatter plot shown in fig.~\ref{fig:stretchlssfr}, we define our model as
+follows: for the young population the underlying stretch distribution is modeled
+as $\mathcal{N}(\mu_1, \sigma_1^2)$, i.e. a normal distribution centered on
+$\mu_1$ with a $\sigma_1$ width; the old population stretch distribution is
+modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
+\mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a gaussian mixture where one mode is the
+same as the young population one.  \nn{When combining our prediction of the
+evolution of young SNe~Ia as a function of redshift eq.~\ref{eq:delta}, and the
+stretch distribution of both young and SNe~Ia, our model for the underlying
+distribution of stretch $X_1\left(z\right)$ as a function of redshift is:}
+
+\begin{align}
+    \label{eq:stretchz}
+    \nn{X_1\left(z \right) =
+    }&\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
+    (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
+    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
+\end{align}
+
+The estimation of the 5 free parameters
 ($\theta\equiv{\mu_1,\mu_2,\sigma_1,\sigma_2,a}$) of the model is given by
-minimizing:
+minimizing a pseudo-$\chi^2$ $\mathcal{L}$ such that:
+
 \begin{equation}
     \label{eq:likelihood}
-    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{\theta\ ;
-                     \mathrm{d}x^{i}_{1}, p(y)^{i}}\right),
+    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
+    \mathrm{d}x^{i}_{1}, p(\nn{\text{young}})^{i}}\right),
 \end{equation}
-with
+
+\nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
+$p(\text{young})^{i}$ are the salt2.4 stretch, error of the salt2.4 stretch and
+the probability that the SN is young, respectively. Depending on the data
+available, there are two ways to fit these 5 free parameters}.
+
+\subsubsection{\nn{Parameters estimation with LsSFR data}}
+\label{sec:modelpy}
+
+\nn{When using SNf data, we choose $p(\text{young})$ to be their given $p(y)$,
+computed from their LsSFR data (see fig.~\ref{fig:stretchlssfr}). Thus, we would
+have $\mathcal{P}$ from eq.~\ref{eq:likelihood} to be}
+
 \begin{align}
     \label{eq:likelihoodsnf}
-    \prob{x^{i}_{1}}{ \theta\ ; \mathrm{d}x^{i}_{1}, p(y)^{i}} = 
-    p(y)^{i}\,\times\,&\mathcal{N}\left(\mu_1, \sqrt{\sigma_1{}^{2}
-    +\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1})\,+\nonumber\\
-    (1-p(y)^{i}) \times [a\,\times\,& \mathcal{N}\left(\mu_1,
-    \sqrt{\sigma_1{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1})\,
-    + \nonumber\\
-    (1-a)\,\times\,&\mathcal{N}\left(\mu_2, \sqrt{\sigma_2{}^{2}
-    +\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1})\,],
+    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, p(\nn{\text{young}})^{i}} =
+    p(y)^{i} \times &\mathcal{N}\left(\mu_1,
+    \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
+    \nonumber\\
+    (1-p(y)^{i}) \times \nn{\bigg[} a \times & \mathcal{N}\left(\mu_1,
+        \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
+        \nonumber\\
+     (1-a) \times &\mathcal{N}\left(\mu_2,
+ \sqrt{\sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) \nn{\bigg]}
 \end{align}
-where $i$ is the index of the SNIa, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
-$p(y)^{i}$ are the salt2.4 stretch, error of the salt2.4 stretch and the
-probability that the SN is prompt, respectively.  In practice, since "$a$" is
-defined between 0 and 1, we fit for $\alpha$ such that
-$a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.
-
-\begin{figure*}
-    \label{fig:stretchlssfr}
-    \centering
-      \includegraphics[width=\linewidth]{Article_figures/BiGaussian_hist.pdf}
-    \caption{SNF data on which we based our models of stretch distribution
-    between young and old}
-\end{figure*}
-
-The fit results are shown table~\ref{tab:modelresults} and the best fit model is
-illustrated in fig.~\ref{fig:stretchlssfr}.
+
+In practice, since "$a$" is defined between 0 and 1, we fit for $\alpha$ such
+that $a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.
+\mr{Results on fitting the SNf data with this model are shown
+table~\ref{tab:modelresults} and illustrated in fig.~\ref{fig:stretchlssfr}.}
 
@@
     \centering
-    \caption{Best fit value of the stretch distribution model. The "SNf" results
-    are derived using eq.~\ref{eq:likelihoodsnf} solely using the SNf data for
-which we have $p(y)$. The "All" and "All(cons)" results are derived using
-eq.~\ref{eq:likelihood} without the age information (see
-section~\ref{sec:modelnopy}) ; "All(cons)" correspond to the "complete sample"
-using conservative redshift cut (see section~\ref{sec:sample}).}
+    \caption{Best fit values of the parameters for the Base stretch distribution
+    model when applied to the SNf dataset only (114 SNe~Ia), the nominal 569
+SN~Ia sample or the conservative one (422).}
     \label{tab:modelresults}
@@
         \hline\\[-0.8em]
-        SNf & $0.41 \pm 0.08$ & $0.55 \pm 0.06$ & $-1.38 \pm 0.10$ &
-        $0.44 \pm 0.08$ & $0.48^{+0.08}_{-0.08}$ \\[0.15em]
-        All & $0.43 \pm 0.09$ & $0.58 \pm 0.05$ & $-1.05 \pm 0.28$ &
-        $0.63 \pm 0.13$ & $0.51^{+0.09}_{-0.10}$ \\[0.15em]
-        All(cons) & $0.38 \pm 0.05$ & $0.60 \pm 0.06$ & $-1.26 \pm 0.13$
-                  & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\[0.15em]
+        SNf & $0.41 \pm 0.08$  & $0.55 \pm 0.06$ & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$ & $0.48^{+0.08}_{-0.08}$ \\[0.15em]
+        All & $0.37 \pm 0.05$  & $0.61 \pm 0.04$ & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$ & $0.51^{+0.09}_{-0.10}$ \\[0.15em]
+        All(cons) & $0.38 \pm 0.05$  & $0.60 \pm 0.04$ & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\[0.15em]
         \hline
@@
 
-When combining our prediction of the evolution of young SNeIa as a function of
-redshift eq.~\ref{eq:delta}, and the stretch distribution of the both the young
-and the old SNeIa, our model for the underlying distribution of stretch
-$\Delta\left(x_1|z\right)$ as a function of redshift is:
+\subsubsection{Without LsSFR data}
+\label{sec:modelnopy}
+
+Given eq.~\ref{eq:stretchz}, we can extend the analysis to non-SNfactory sample
+by fitting the free parameters of the model
+($\theta\equiv{\mu_1,\mu_2,\sigma_1,\sigma_2,a}$) assuming the evolution of the
+fraction of young SNe~Ia as a function of redshift given by $\delta(z)$ and
+given, this time, $x_1^{i}$,  $\mathrm{d}x^{i}_{1}$ and $z^{i}$, the stretch,
+error on the stretch and redshift of any given SN $i$. In practice, it means
+that we are still minimizing for the same eq.~\ref{eq:likelihood}, but replacing
+$p(y)^{i}$ by $\delta(z^{i})$ eq.~\ref{eq:likelihoodsnf}.
+
+\mr{For the rest of the analysis, we will therefore fit for
+eq.~\ref{eq:likelihood} using $p(y)^{i}$ --~the probability for the SN $i$ to be
+young~-- when available (i.e. for SNf dataset) and $\delta(z^{i})$ --~the
+expected fraction of young SNe~Ia at the SN redshift $z^{i}$~-- otherwise.}
+
+\mr{Results on fitting the all the 569(422) SNe~Ia data (with conservative cuts)
+with this model are given table~\ref{tab:modelresults} and illustrated in
+fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
+stretch per bin of redshift follow our redshift drift modeling. That is, when
+considering selection-bias free SN~Ia samples, SNe~Ia at higher redshift have on
+average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
+redshift ($-0.17\pm 0.10$ at $z\sim0.05$). This is indeed what is expected if
+old environments favor low SN stretch \citep[e.g.]{howell2007} \textit{and} if
+the fraction of old SNe~Ia reduces as a function of redshift. See
+section~\ref{sec:results} for a more quantitative discussion.}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
+    \caption{Evolution of the mean SN salt2.4 stretch ($x_1$) as a function of
+        redshift.  Markers show the mean stretch measured in redshift bins of
+        equal sample size indicated in light gray at the bottom of each redshift
+        bins. Full and light markers are used when considering the nominal or
+        the conservative sample, respectively, and are colored by their mean
+        redshift.  The orange horizontal line represent the mean redshift of the
+    nominal sample illustrating the expectation if the SN stretch distribution
+is not drifting. Best fits of our Base drifting model are shown as blue,
+dashed-blue and gray, when fitted on the nominal, the conservative or the SNf
+dataset, respectively; all are similar. The light-blue illustrate the amplitude
+of the error of the best fit model when considering the nominal dataset.}
+    \label{fig:modelall}
+\end{figure*}
+
+\subsection{Other modelings}
+\label{sec:othermodel}
+
+In section~\nn{\ref{sec:basemodel}} we have modeled the underlying stretch
+distribution as a single gaussian for the young SNe Ia and a combination of two
+gaussians for the old SNe Ia population, i.e., the same as the young-population
+one plus another for the fast declining SNe~Ia that seem to only exist in old
+populations. This is our “Base” model.
+
+\cite{howell2007} used a simpler modelisation: a single gaussian for each age
+sub-group. To test \nn{different modeling choices}, we have implemented a suite
+of extra parametrisation that we also fit to the data following the procedure
+described section~\ref{sec:modelnopy}.
+
+%Namely we consider:
+%\begin{itemize}
+ %   \item “Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$”: Base model but where the high-stretch Gaussian of the old population is not set as that of the young population (+2 degrees of freedom);
+ %   \item “Base$-(\sigma_2)$”: Base model but where $\sigma_2$ is fixed at the same value as $\sigma_1$ (-1 degree of freedom);
+ %   \item “Howell”: one single Gaussian per age group as in \cite{howell2007}.
+%\end{itemize}
+
+\nn{We thus consider a model ''Howell+drift'', with one single gaussian per age
+group as in \citet{howell2007}}. In addition, since we aim at probing the SN
+redshift drift, we are also fitting the ''Base'' model and \nn{this
+aforementioned extension} by fixing the population drift parameter $\delta(z)$
+to a free parameter constant $\delta(z)=f$.
+
+We are also considering alternative non-drifting models, and notably the one
+developed for the Beam with Bias Correction \cite[BBC,][]{scolnic2016,
+kessler2017}, currently used by all recent SN Cosmological analyses
+\cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to account
+for Malmquist biases. The BBC formalism assumes sample-based (hence
+non-drifting) asymmetric Gaussian distributions ($\mathcal{N}\left(\mu,
+\sigma^{-}{}^2\, \mathrm{if}\,x_1<\mu\,\mathrm{else}\, \sigma^{+}\right)$). The
+idea behind the sample-based approach is twofold: (1) Malmquist biases are
+driven by survey properties and (2) because current surveys cover limited
+redshift ranges, doing so contains some potential redshift evolution information
+\citep{scolnic2016, scolnic2018a}. See further discussion concerning the BBC
+modeling in section~\ref{sec:bbc}. For the sake of the comparison, we are also
+running Gaussian and asymmetric Gaussian non-drifting models.
+
+The ability these models to explain the data is detailed in
+section~\ref{sec:results}.
+
+\section{Results}
+\label{sec:results}
+We applied each model on both the conservative and non-conservative samples (cf.
+section \ref{sec:sample}). Because the models presented
+section~\ref{sec:modeling} have various degrees of freedom, we use the Akaike
+Information Criterion (AIC) \citep{burnham2004} to compare them. This estimator
+penalizes extra degrees of freedom to avoid over-fitting the data. It is defined
+as follow:
+\begin{equation}
+    \nn{\mathrm{AIC} = 2k - 2\ln L}
+\end{equation}
+where $k$ is the number of free parameters and \nn{$-2\ln L = \mathcal{L}$} is
+the pseudo-$\chi^2$ defined eq.~\eqref{eq:likelihood}. The \nn{reference} model
+is the one with the smaller AIC and the probability for another model to be at
+least as representative of the data as this one is given by:
+\begin{equation}
+    p(\mathrm{other} \geq \mathrm{ref}) =
+    \exp\left(\Delta\mathrm{AIC}/2\right)
+\end{equation}
+Results are presented in table \ref{tab:comp} and are illustrated
+fig.~\ref{fig:mod_comp}.
+
+\begin{table*}
+    \centering
+    \caption{Comparison relative ability for a model to describe the data using
+    AIC criterion. See the model definition section~\ref{sec:modeling}. The
+drift column indicates if the model contains the age-drifting model
+($\delta(z)$) or not (f, i.e. non-drifting). The BBC modeling is sample based
+which somewhat contains redshift dependency information. $\mathcal{L}$ is the
+pseudo-$chi^{2}$ defined eq.~\ref{eq:likelihood} and Proba is the probability
+that a model is at least as good as the reference model to describe the data.
+See also fig.~\ref{fig:mod_comp}.}
+    \label{tab:comp}
+    \begin{tabular}{c c c  |c c c c|  c c c c}\hline\hline\\[-0.8em]
+        & & & \multicolumn{4}{|c}{All SNe~Ia (569)} & \multicolumn{4}{|c}{All SNe~Ia (conservative; 422)} \\
+        %\cline{2-6}
+        Name & drift & Free param &
+        $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba & $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba \\[0.15em]
+        \hline\\[-0.8em]
+
+        Base & $\delta(z)$ & 5
+        & 1456.7 & 1466.7 & -- & --
+        & 1079.5 & 1089.5 & -- & --\\[0.15em]
+
+        Howell & $\delta(z)$ & 4
+        & 1463.3 & 1471.3 & -4.6 & $1.0\times10^{-1}$
+        & 1088.2 & 1096.2 & -6.7 & $3.4\times10^{-2}$\\[0.15em]
+
+        Asymmetric & -- & 3
+        & 1485.2 & 1491.2 & -24.5 & $4.7\times10^{-6}$
+        & 1101.3 & 1107.3 & -17.8 & $1.4\times10^{-4}$\\[0.15em]
+
+        Howell & $f$ & 5
+        & 1484.2 & 1494.2 & -27.5 & $1.0\times10^{-6}$
+        & 1101.2 & 1111.2 & -21.7 & $1.9\times10^{-5}$\\[0.15em]
+
+        Base & $f$ & 6
+        & 1484.2 & 1496.2 & -29.5 & $3.9\times10^{-7}$
+        & 1101.2 & 1113.2 & -23.7 & $7.1\times10^{-6}$\\[0.15em]
+
+        BBC-modeling & per sample & 3x5
+        & 1468.2 & 1498.2 & -31.5 & $1.5\times10^{-7}$
+        & 1083.6 & 1113.6 & -24.1 & $5.7\times10^{-6}$\\[0.15em]
+
+        Gaussian & -- & 2
+        & 1521.8 & 1525.8 & -59.1 & $1.5\times10^{-13}$
+        & 1142.6 & 1146.6 & -57.1 & $4.0\times10^{-13}$ \\\hline\hline
+    \end{tabular}
+\end{table*}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/mod_comp.pdf}
+    \caption{$\Delta$AIC between the reference model and the other considered
+    models(see also table~\ref{tab:comp}). The full and open blue markers
+correspond to model with and without redshift drift, respectively. Light markers
+show the results when this analysis is done on the conservative sample.
+Color-bands illustrate the validity of the models, from Acceptable to Excluded,
+see inside text. All non-drifting models are excluded.}
+    \label{fig:mod_comp}
+\end{figure}
+
+\mr{We find that underlying stretch distributions with no redshift drift are all
+excluded to describe the data as well as the best redshift models (i.e. Base and
+Base$-(\sigma_2)$). In fact, the best non drifting model (Asymmetric) has
+$2\times10^{-4}\%$ of chance to describe the data as well as Base$-(\sigma_2)$,
+the best model.  This is visible fig~\ref{fig:modelall} where the mean stretches
+per bin of redshift of equal sample size suggest a redshift evolution following
+our Base model --~fitted either on SNf, on the nominal 569 SNe~Ia dataset or the
+conservative one~-- rather than having a constant mean value.}
+
+\mr{As just mentioned, the best model is Base$-(\sigma_2)$ where both the
+    low-stretch (mode 2 that only exists in old environments) and high-stretch
+    (mode 1 that exists in all environments) normal distributions share the same
+    width. It is marginally better than the Base model where both distributions
+    have their own ($\Delta \mathrm{AICc}<2$). It is unclear which physical
+    mechanism would translate into a simple distribution shift of mode 1 to a
+    mode 2 that would only exist in old populations. Two distinct underlying
+    distributions, as assumed by the Base model, seems more natural, yet, more
+    data would be necessary to discriminate between these models.  Similarly,
+letting the high-stretch distribution to vary between the young and the old
+population (Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$) only marginally
+improves the fit ($\Delta\mathcal{L}=3.6$ in comparison to Base at a cost of 2
+more degrees of freedom, making this mode less likely. This suggests that,
+indeed, the high-stretch mode (mode 1) exists in all kind of age-populations.}
+\mr{We also remark that considering the stretch distribution of each
+age-population as independent normal distributions, as suggested by
+\cite{howell2007}, is 10 times less likely than our best model and is therefore
+unfavored.}
+
+\mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
+    distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
+    modelisation.  While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the best of all
+    the non-drift model, but still $\Delta \mathcal{L}=11.3$ worse than our best
+    model, it is penalized by requiring 15 free parameters, 3 per sample.
+    \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
+    highlighted that, because current surveys span limited redshift ranges, the
+    per sample approach somewhat account for unmodeled redshift drifts. We
+    stress here that, as measure of modern surveys cover larger redshift ranges
+    to reduce calibration systematic  uncertainties, this is becoming less true,
+    notably for PS1, DES and, soon, LSST. Already with the current surveys, the
+    BBC sample-based technique is excluded to be an as good representation of
+    the data as our best model ($p=4\times 10^{-8}$).  A more detailed
+discussion of the consequence of this results for cosmology follows in
+section~\ref{sec:bbc}}
+
+\mr{We report in table~\ref{tab:bbc} the sample's $\sigma^-$,  $\sigma^+$ and
+    $\mu^0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
+    SDSS and with \cite{scolnic2018a} for PS1. This  validates our somewhat
+    simple construction of our selection-bias free sample, see
+    section~\ref{sec:sample}.  If we were to use the \cite{scolnic2016} and
+    \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
+    the $\Delta\mathrm{AICc}$ between our Base drifting model and the BBC
+    modeling would go from -32 to -47.  Furthermore, fixing the Base model
+    parameters to those derived using SNf data only (top row
+    table.~\ref{tab:modelresults}) and using the BBC parameters from the
+    aforementioned literature for PS1, SDSS and SNLS, we find
+    $\Delta\mathrm{AICc}=31.7$. Removing the SNf data reduces this difference to
+    $\Delta\mathrm{AICc}=5$.  Such a reduction is expected for two reasons.
+    First, it removes the nearby SNe~Ia, and thereby significantly reduces the
+    redshift level arm to probe the impact of potential redshift drifts.
+    Second, the SNf data are better at discriminating the individual stretch
+    distributions by age since the probability for a SN to be young $p(y^i)$ is
+    known while other samples rely on $\delta(z^i)$ which only is the fraction
+    of young SNe~Ia at the SN redshift, i.e. a proxy for $p(y^i)$.  While in any
+case the Base-model is favored, the relative weight of a SNf sample highlight
+the  importance of addition non-targeted nearby surveys such the Zwicky
+Transient Facility \cite[ZTF,][]{bellm2019,graham2019} to unambiguously confirm
+our result.}
+
+%Using S18/SK16 values of \sigma_-, \sigma_+ and \mu_0, we find that SDSS, PS1 and SNLS's $\mathcal{L}$ to be 457.9, 397.6 and 254.5 respectively, and their AICc to be 464.0, 403.7 and 260.8 respectively, systematically higher than those we find in this study (see table \ref{tab:bbc_comp}).
+
+\begin{table}
+    \centering
+    \caption{Best fit parameters for our sample-based asymmetric modeling of the
+    underlying stretch distribution.}
+    \label{tab:bbc}
+    \begin{tabular}{c c c c c c c}\hline\hline\\[-0.8em]
+    Asymmetric & $\sigma^-$ & $\sigma^+$ & $\mu^0$ \\\hline\\[-0.8em]
+    SNf &  1.34 $\pm$ 0.13 & 0.41 $\pm$ 0.10 & 0.68 $\pm$ 0.15\\[0.15em]
+    SDSS & 1.31 $\pm$ 0.11 & 0.42 $\pm$ 0.09 & 0.72 $\pm$ 0.13 \\[0.15em]
+    PS1 &  1.01 $\pm$ 0.11 & 0.52 $\pm$ 0.12 & 0.38 $\pm$ 0.16 \\[0.15em]
+    SNLS & 1.41 $\pm$ 0.13 & 0.15 $\pm$ 0.13 & 1.22 $\pm$ 0.15 \\[0.15em]
+    HST &  0.76 $\pm$ 0.36 & 0.79 $\pm$ 0.35 & 0.11 $\pm$ 0.44 \\\hline\hline
+    \end{tabular}
+\end{table}
+
+\mr{Considering or not the conservative sample has no effect on our conclusion
+as illustrated fig.~\ref{fig:modelall} and presented table~\ref{tab:comp}.}
+
+\section{Discussion}
+\label{sec:bbc}
+\mr{Our results strongly suggest that the underlying distribution of SN~Ia
+stretch evolves as a function of redshift. Our redshift drift modeling presented
+section~\ref{sec:modeling} is a much better representation of the pantheon data
+(see section~\ref{sec:sample}) than any non-drifting models, including the
+sample-based BBC one.}
+
+\mr{To the best of our knowledge, the SN~Ia stretch redshift drift has never
+    been explicitly accounted for in cosmological analyses. Not doing so is a
+    second order issue for SN cosmology as it only affects the way one does
+    Malmquist bias correction. Indeed, as long as the Phillip's relation
+    standardization parameter $\alpha$ is not redshift dependent (study behind
+    the scope of this paper, but see e.g. \citealt{scolnic2018a}), if a sample
+    is free from selection effects, the stretch-corrected SNe~Ia magnitudes used
+    for cosmology are blind to the underlying stretch distribution.  However,
+    since the first missed SNe~Ia are the faintest, which are also those with
+    the lowest stretch, once we include SNe~Ia affected by selection effects we
+    need to account for the fact that only the brightest SNe~Ia exist in the
+    low-stretch end of the sample after standardization.  The modeling of the SN
+stretch then enters the analysis to understand which SNe~Ia are missing given
+the selection functions to mimic a selection-bias free sample.}
+
+\mr{The state-of-the-art tool for doing so is the BBC described in
+\cite{scolnic2016} and \cite{kessler2017}, which is used in all recent SN
+Cosmological analyses
+\citep{jones2018b,scolnic2018a,brout2019,descosmopaper2019} including the direct
+measurement of $H_0$ \citep{riess2016,riess2019}. As already discussed, it
+currently assumes an asymmetric stretch distribution per sample and we have
+shown in section~\ref{sec:results} that this modeling is excluded to be an as
+good description of the data than our age-based redshift drift modeling
+eq.~\ref{eq:stretchz}.}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/bbc_comp_PS1.pdf}
+    \caption{Underlying SN salt2.4 stretch ($x_1$) distributions for the PS1
+    sample. The Green line(band) shows our best fitted asymmetric Gaussian
+distribution (its error) mimicking the BBC approach. The dashed line shows that
+estimated by \cite{scolnic2018a}. The blue to red curves display our Base
+drifting stretch model estimated at each PS SN redshift; bluer for lower
+redshifts.}
+    \label{fig:bbc_pdf_ps1}
+\end{figure}
+
+\mr{We illustrate fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
+underlying stretch distribution between the BBC modeling and our Base drifting
+model for the PS1 sample. By definition, our model is bimodal and the relative
+amplitude of each mode depends on the fraction of old SNe~Ia in the sample
+--~the highest the fraction of old SNe~Ia, the highest the amplitude of the
+low-stretch mode~-- and is therefore redshift dependent.  As expected the two
+approaches strongly differ in modeling the negative part of the SN stretch
+distribution. The BBC asymmetric distribution goes in the middle of the bimodal
+distribution, largely over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
+largely under-estimating it at $x_1\sim-1.7$ for typical PS1 SN redshifts. This
+means that the bias-correction of the individual SN stretch
+$\overline{\delta}_{x_1}$ \citep{kessler2017} and consequently the magnitude
+bias correction $\mu_B$ is most likely inaccurate. However, given the complexity
+of the BBC analysis, a full study using our Base model given
+eq.~\ref{eq:stretchz} in place of the sample-based asymmetric modeling would be
+necessary to have an exact understanding of the effect this inaccurate modeling
+has on cosmology.}
+
+\mr{Nonetheless, to have a rough estimation of the expected amplitude of the
+effect, we show fig.~\ref{fig:magdrift} the difference of mean stretch
+correction $\alpha\times\left(\langle x_1 \rangle_{\mathrm{BBC}} - \langle x_1
+\rangle_{\mathrm{Base}}\right)$ as a function of redshift, using $\alpha=0.156$
+from \cite{scolnic2018a}. We see in this figure that the amplitude of the
+magnitude bias caused by an accurate modeling of the underlying stretch
+distribution is of the order of few tens of millimag, which is the typical scale
+exotic forms of dark energy could have on $\mu(z)$. In the era of modern
+cosmology where we aim at probing $w_0$ at a sub-percent and $w_a$ at a few
+percent \citep[e.g.,][]{lsstpaper}, it is therefore of paramount importance to
+further understand the exact modeling of the lightcurve parameters and
+especially their potential redshift drift; see also discussions about the impact
+of stretch and color distribution modeling for cosmology in \citealt{rubin2015}
+and \citealt{rubin2016}.}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/BBC_distmod_w0wa.pdf}
+    \caption{ Difference of mean stretch standardization (in mag) between the
+        BBC algorithm used in \cite{scolnic2018a} and our Base drifting model.
+        The blue markers represent the mean evolution in 30 regular redshift
+        bins, while the small colored points show the SDSS, PS1 and SNLS
+        samples. For comparison, we show in gray bands the expected magnitude
+        difference between $\Lambda$CDM  and $w$CDM cosmologies ($w=-1\pm
+        10,\,5\,\mathrm{and}\,1\%$) or, in orange, more exotic $w_0,w_a$CDM
+        examples.  This figure illustrates that the amplitude of the potential
+    magnitude bias caused by inaccurate modeling of the underlying stretch
+distribution is comparable to the observational effect of exotic form of dark
+energy.}
+    \label{fig:magdrift}
+\end{figure}
+
+\section{Conclusion}
+\label{sec:ccl}
+We have presented a study of the drift of the underlyin SNe~Ia stretch
+distribution as a function of redshift.  We used SNe~Ia from magnitude-limited
+surveys from the pantheon data set \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
+well as HST to which we added SNfactory data from \cite{rigault2018} for the
+low-redshift bin. We only considered the supernovae that have been discovered
+in redshift ranges of each surveys where selection effect are negligible. This
+way the observed SNe~Ia stretch are random sampling of the true underlying
+distribution. This corresponds the 569 SNe~Ia (422 when considering more
+conservative cuts).
+
+Following observations made in \cite{rigault2018} we introduced a redshift drift
+modeling that depend on the expected fraction of young and old SNe~Ia as a
+function of redshift, each age-population having its own underlying stretch
+distribution. We studied various suites of distribution as well as non-drifting
+modeling. We also closely studied the stretch distribution modeling made in
+current version of the BCC algorithm used for Malquist bias correction by all
+recent Dark energy or Hubble constant cosmological analyses.
+
+Our conclusion are the following:
+\begin{enumerate}
+
+    \item Non-drifting models are excluded to be as good description of the data
+        in comparison to our Base modeling.  This model assumes that: (1) the
+        young population has a unimodal Gaussian stretch distribution while the
+        old population stretch one is bimodal, one mode been the young one; (2)
+        the evolution of the relative fraction of young and old SNe~Ia follows
+        the prediction made in \cite{rigault2018}.
+
+    \item Given 1. we conclude that the SNe~Ia is indeed drifting, as already
+        suggested by \citep[e.g.][]{howell2007}.
+
+    \item The BBC approach, which assumes each survey to have its own asymmetric
+        Gaussian distribution, is excluded to be an as good description of the
+        data as our drifting model ($p=4\times10^{-8}$). Hence, the BBC
+        sample-based approach does not accurately account for redshift drift as
+        suggested by \cite{scolnic2016} and \cite{scolnic2018a}.
+
+    \item An inaccurate underlying stretch distribution modeling is estimated to
+        bias the mean standardized SNe~Ia on the order of a few percent, which
+        is comparable in scale to the observational signature of exotic forms of
+        dark energy.
+
+    \item We suggest to use the following stretch population modeling:
     \begin{align}
-    \label{eq:stretchz}
+    \label{eqconclusion:stretchz}
         \Delta\left(x_1\,|\,z \right) =
         &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1)\,+\nonumber\\
-    (1-&\,\delta(z)) \times \left[a\times\mathcal{N}(\mu_1,\sigma_1)
-    + (1-a)\times\mathcal{N}(\mu_2,\sigma_2)\right]
+        (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1) +
+        (1-a)\times\mathcal{N}(\mu_2,\sigma_2)\right]\nonumber
     \end{align}
-
-\subsection{Without LsSFR data}
-\label{sec:modelnopy}
-
-Given eq.~\ref{eq:stretchz}, we can extend the analysis by fitting the free
-parameters of the model ($\theta\equiv{\mu_1,\mu_2,\sigma_1,\sigma_2,a}$)
-assuming the evolution of the fraction of young SNeIa as a function of redshift
-given by $\delta(z)$ and given, this time, $x_1^{i}$,  $\mathrm{d}x^{i}_{1}$ and
-$z^{i}$, the stretch, error on the stretch and redshift of any given SN $i$. In
-practice, it means that we are still minimizing for the same
-eq.~\ref{eq:likelihood}, but replacing $p(y)^{i}$ by $\delta(z^{i})$
-eq.~\ref{eq:likelihoodsnf}.
-
-The fit results when using the conservative redshift cuts or not are shown
-table~\ref{tab:modelresults} and the best fit models are illustrated in
-fig.~\ref{fig:modelall}.
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{Article_figures/stretchevol_all_vs_snf_maglim-cuts.pdf}
-    \caption{Result of the model fitted on all the surveys' data, cut at the
-    conservative $z_{\mathrm{max}}$ in transparent and non-conservative
-$z_{\mathrm{max}}$ in full markers. The number of SNe Ia in each bin is
-indicated just above the x-axis in transparent text; between brackets are the
-numbers for the conservative cuts.}
-    \label{fig:modelall}
-\end{figure*}
-
-\subsection{Other modelings}
-\label{sec:othermodel}
-
-In section~\ref{sec:modelpy} and \ref{sec:modelnopy} we have modeled the
-underlying stretch distribution as a single gaussian for the young SNe Ia and a
-combination of two gaussians for the old SNe Ia population, i.e., the same as
-the young-population one plus another for the fast declining SNeIa that seem to
-only exist in old populations.
-
-\cite{howell2007} used a simpler modelisation: the underlying stretch of each
-age population is a single gaussian. This rises the question of the accuracy of
-our modeling choice. To test this, we have implemented a suite of extra
-parametrisation that we also fit to the data following the procedure described
-section~\ref{sec:modelnopy}. Namely we consider:
-\begin{itemize}
-
-    \item "Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$": base model but
-        where the high-stretch Gaussian of the old population is not set as that
-        of the young population ;
-
-    \item "Base$-(\sigma_2)$": base model but where $\sigma_2$ is fixed at the
-        same value as $\sigma_1$, since both width seem not significantly
-        different in Table~\ref{tab:modelresults};
-
-    \item "Howell": one single Gaussian per age group as in \citep{howell2007},
-        but with the $\delta(z)$ factor.
-
-\end{itemize}
-
-Furthermore, since we aim at probing the SN redshift drift, we are also fitting
-the "Base" model and the three aforementioned extensions by fixing the
-population drift parameter $\delta(z)$ to a free parameter constant
-$\delta(z)=f$. Doing so enables us to consider two last models. The first is
-modeling the underlying SNe Ia stretch distribution by a simple Gaussian, namely
-"Gaussian (f)". The second is using an asymmetric Gaussian distribution as
-currently done by the BBC formalism introduced by \cite{kessler2017,
-scolnic2018b} for modeling the sample selection functions in SN Cosmology. This
-last model is particularly interesting to test as it is currently used in the
-most recent SN Cosmological analyses \cite{riess2016, riess2018, PSpaper,
-scolnic2018a, DESpaper} \nn{supposed riess2018 to be
-https://ui.adsabs.harvard.edu/abs/2018ApJ...861..126R/abstract}; see further
-discussion concerning this issue in section~\ref{sec:bbc}.
-
-The ability of the model to explain the data is detailed in
-section~\ref{sec:results}.
-
-\section{Results}
-\label{sec:results}
-
-We applied each model on both the conservative and non-conservative sample (cf.
-section \ref{sec:sample}. We want to test whether including an evolution of the
-stretch with the redshift ($\delta(z)$ factor) would better describe the data.
-As the various models presented section~\ref{sec:modeling}, have various degrees
-of freedom. To compare them all, we use the Akaike Information Criterion
-corrected for sample size (AICc), that penalizes extra degrees of freedom to
-avoid over-fitting the data. The AICc is defined as follow:
-\begin{equation}
-    \mathrm{AICc} = \mathrm{AIC} + \frac{2k\left(k+1\right)}{n - k - 1}
-\end{equation}
-
-with $\mathrm{AIC} = 2k + \mathcal{L}$, where $k$ is the number of free
-parameters and $\mathcal{L}$ is defined eq~\eqref{eq:likelihood}.  The best
-model is the one with the smaller AICc and the probability for another model to
-be at least as representative of the data as this one is given by:
-\begin{equation}
-    p(\mathrm{other} > \mathrm{best}) = \exp\left(\Delta\mathrm{AICc}/2\right)
-\end{equation}
-The results are summarized in table \ref{tab:comp} and represented figure
-\ref{fig:mod_comp}.
-
-\begin{table*}
-    \centering
-    \caption{Comparison of quality of fit for all the models. Base indicates the
-    models designed for this study. (F) indicates models without an evolution of
-the fraction of young SNe Ia as a function of the redshift.}
-    \label{tab:comp}
-    \begin{tabular}{c c c  c c c c  c c c c}\hline\hline\\[-0.8em]
-        & & & \multicolumn{4}{c}{All SNeIa (569)} &
-        \multicolumn{4}{c}{All SNeIa (conservative ; 422)} \\
-        %\cline{2-6}
-        Name & drift(?) & Free param &
-        $\ln\mathcal{L}$ & $\mathrm{AICc}$ & $\Delta \mathrm{AICc}$ & Proba &
-        $\ln\mathcal{L}$ & $\mathrm{AICc}$ & $\Delta \mathrm{AICc}$ & Proba \\
-        [0.15em]\hline\\[-0.8em]
-
-        Base$-(\sigma_2)$ & $\delta(z)$ & 4
-        & 1456.9 & 1464.9 & -- &  --
-        & 1079.9 & 1088.0 & -- & -- \\[0.15em]
-
-        Base & $\delta(z)$ & 5
-        & 1456.7  & 1466.8 & -1.8 & $4.0\times10^{-1}$
-        & 1079.5 & 1089.6 & -1.7 & $4.4\times10^{-1}$\\[0.15em]
-
-        Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$ & $\delta(z)$ & 7
-        & 1453.1  & 1467.3 & -2.4 & $3.1\times10^{-1}$
-        & 1074.5 & 1088.8 & -0.8 & $6.7\times10^{-1}$\\[0.15em]
-
-        Howell & $\delta(z)$ & 4
-        & 1463.3 & 1471.3 & -6.4 & $4.1\times10^{-2}$
-        & 1088.2 & 1096.3 & -8.3 & $1.5\times10^{-2}$\\[0.15em]
-
-        Assymetric & $f$ & 3
-        & 1485.2 & 1491.3 & -26.3 & $1.9\times10^{-6}$
-        & 1101.3 & 1107.4 & -19.4 & $6.2\times10^{-5}$\\[0.15em]
-
-        Howell & $f$ & 5
-        & 1484.2 & 1494.3 & -29.4 & $4.2\times10^{-7}$
-        & 1101.2 & 1111.3 & -23.4 & $8.5\times10^{-6}$\\[0.15em]
-
-        Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$ & $f$ & 8
-        & 1478.9 & 1495.2 & -30.2 & $2.8\times10^{-7}$
-        & 1095.5 & 1111.9 & -23.9 & $6.5\times10^{-6}$\\[0.15em]
-
-        Base & $f$ & 6
-        & 1484.2 & 1496.4 & -31.4 & $1.5\times10^{-7}$
-        & 1101.2 & 1113.4 & -25.4 & $3.0\times10^{-6}$\\[0.15em]
-
-        Gaussian & -- & 2
-        & 1521.8 & 1525.8 & -60.9 & $6.0\times10^{-14}$
-        & 1142.6 & 1146.6 & -58.6 & $1.9\times10^{-13}$ \\[0.15em]
-
-        Base$-(\sigma_2)$ & $f$ & 5
-        & 1521.8 & 1531.9 & -67.0 & $2.9\times10^{-15}$
-        & 1142.6 & 1152.7 & -64.7 & $8.8\times10^{-15}$\\\hline
-
-    \end{tabular}
-\end{table*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{Article_figures/mod_comp.pdf}
-    \caption{$\Delta$AICc of the models fitted on all the surveys' data, cut at
-    the conservative $z_{\mathrm{max}}$ in transparent and non-conservative
-$z_{\mathrm{max}}$ in full markers.}
-    \label{fig:mod_comp}
-\end{figure*}
-
-For the non-conservative part, we find that every model lacking an evolution of
-the stretch with the redshift is systematically worse than those that implement
-it, and that removing the $\sigma_2$ parameter gives better results than letting
-it free, or adding $(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$. However, the
-base model with 2 gaussians for the old population is roughly the same as the
-Howell model with one gaussian per population. As for the conservative part, all
-the models but the asymmetric and the gaussian ones correctly fit the data. No
-fixed model fit better than those with evolution, but some come close. It would
-seem that taking SNe Ia close to us doesn't give us a sample where the evolution
-is visible, justifying the study done in section~\ref{sec:sample}.
-
-\section{discussion}
-
-\subsection{BBC and assumptions on underlying SN property}
-\label{sec:bbc}
-BBC use the asymmetric model. When we do it, we get the table TABLE
-
-\begin{table*}
-    \centering
-    \caption{Fit results of the asymmetric and base models for the 3 surveys
-    having selection effects.}
-    \label{tab:bbc}
-    \begin{tabular}{c c c c c c c c c c}\hline\hline\\[-0.8em]
-    Asymetric & Free param & $\sigma^-$ & $\sigma^-_{\mathrm{err}}$ &
-    $\sigma^+$ & $\sigma^+_{\mathrm{err}}$ & $\mu^0$ & $\mu^0_{\mathrm{err}}$
-               & ln $\mathcal{L}$ & AICc \\\hline
-    SDSS & 3 & 1.25 & 0.08 & 0.39 & 0.07 & 0.69 & 0.10 & 893.9 & 900.0 \\
-    PS1 & 3 & 1.00 & 0.09 & 0.49 & 0.09 & 0.42 & 0.12 & 702.0 & 708.1 \\
-    SNLS & 3 & 1.18 & 0.11 & 0.39 & 0.11 & 0.81 & 0.15 & 594.8 & 600.9 \\\hline
-    Total & 9 & -- & -- & -- & -- & -- & -- & 2190.7 & 2209.0\\\hline
-    \end{tabular}
-
-    \vspace*{12pt}
-
-    \begin{tabular}{c c c c c}\hline\hline\\[-0.8em]
-    Base &  Free param & ln $\mathcal{L}$ & AICc \\\hline
-    SDSS & 0 & 395.2 & 405.6 \\
-    PS1 & 0 & 444.7 & 455.0 \\
-    SNLS & 0 & 254.1 & 264.7 \\\hline
-    Total & 0 & 1094.0 & 1125.4 \\\hline
-    \end{tabular}
-\end{table*}
-
-\section{Conclusion}
-\label{sec:ccl}
-Stretch evolution with the redshift is a thing. Need to see if it has an impact
-on the cosmology though.
+    with: $a=0.50$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
+    $\sigma_2=0.56$, see table~\ref{tab:modelresults} and using the
+    age-population drift modelization $\delta(z)$ defined in \cite{rigault2018}
+    with $K=0.87$:
+    \begin{align}
+        \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\nonumber
+    \end{align}
+\end{enumerate}
 
 \begin{acknowledgements}
-This project has received funding from the European Research Council (ERC) under
-the European Union's Horizon 2020 research and innovation programme (grant
-agreement n 759194 - USNAC).
-
+    This project has received funding from the European Research Council (ERC)
+    under the European Union's Horizon 2020 research and innovation programme
+    (grant agreement n 759194 - USNAC).
     % How has received that support for this project ? 
@@
 
+\bibitem[Abbott et al.(2019)]{descosmopaper2019} Abbott, T.~M.~C., Allam, S., Andersen, P., et al.\ 2019, \apjl, 872, L30
+
 \bibitem[Aldering(2004)]{aldering2004} Aldering, G.\ 2004, APS April Meeting Abstracts 2004, V4.003
@@
 % B
+\bibitem[Bazin et al.(2011)]{bazin2011} Bazin, G., Ruhlmann-Kleider, V., Palanque-Delabrouille, N., et al.\ 2011, \aap, 534, A43
+
+\bibitem[Bellm et al.(2019)]{bellm2019} Bellm, E.~C., Kulkarni, S.~R., Graham, M.~J., et al.\ 2019, \pasp, 131, 018002
+
 \bibitem[Betoule et al.(2014)]{betoule2014} Betoule, M., Kessler, R., Guy, J., et al.\ 2014, \aap, 568, A22
 
-\bibitem[Bazin et al.(2011)]{bazin2011} Bazin, G., Ruhlmann-Kleider, V., Palanque-Delabrouille, N., et al.\ 2011, \aap, 534, A43
+\bibitem[Brout et al.(2019)]{brout2019} Brout, D., Scolnic, D., Kessler, R., et al.\ 2019, \apj, 874, 150
+
+\bibitem[Burnham \& Anderson(2004)]{burnham2004} Burnham, K., Anderson, D., \
+2004, Sociological Methods \& Research, 33, 2
 
@@
 
+%\bibitem[Conley et al.(2011)]{conley2011} Conley, A., Guy, J., Sullivan, M., et al.\ 2011, \apjs, 192, 1
+
 % D
@@
 % F
+\bibitem[Feeney et al.(2019)]{feeney2019} Feeney, S.~M., Peiris, H.~V., Williamson, A.~R., et al.\ 2019, \prl, 122, 061105
+
 \bibitem[Freedman et al.(2019)]{freedman2019} Freedman, W.~L., Madore, B.~F., Hatt, D., et al.\ 2019, \apj, 882, 34
 
+\bibitem[Frieman et al.(2008)]{frieman2008} Frieman, J.~A., Bassett, B., Becker, A., et al.\ 2008, \aj, 135, 338
+
 % G
+\bibitem[Graham et al.(2019)]{graham2019} Graham, M.~J., Kulkarni, S.~R., Bellm, E.~C., et al.\ 2019, \pasp, 131, 078001
+
 \bibitem[Gupta et al.(2011)]{gupta2011} Gupta, R.~R., D'Andrea, C.~B., Sako, M., et al.\ 2011, \apj, 740, 92
 
-\bibitem[Guy et al.(2010)]{guy2010} Guy, J., Sullivan, M., Conley, A., et al.\ 2010, \aap, 523, A7
+\bibitem[Guy et al.(2007)]{guy2007} Guy, J., Astier, P., Baumont, S., et al.\ 2007, \aap, 466, 11
+
+%\bibitem[Guy et al.(2010)]{guy2010} Guy, J., Sullivan, M., Conley, A., et al.\ 2010, \aap, 523, A7
 
@@
 
-\bibitem[Howell et al.(2009)]{howell2009} Howell, D.~A., Sullivan, M., Brown, E.~F., et al.\ 2009, \apj, 691, 661
+%\bibitem[Howell et al.(2009)]{howell2009} Howell, D.~A., Sullivan, M., Brown, E.~F., et al.\ 2009, \apj, 691, 661
 
@@
 % J
-\bibitem[Jones et al.(2015)]{jones2015} Jones, D.~O., Riess, A.~G., \& Scolnic, D.~M.\ 2015, \apj, 812, 31
+\bibitem[Jones et al.(2015)]{jones2015} Jones, D.~O., Riess, A.~G., \& Scolnic, D.~M.\ 2015, \apj, 812, 3
+1
 
 \bibitem[Jones et al.(2018)]{jones2018} Jones, D.~O., Riess, A.~G., Scolnic, D.~M., et al.\ 2018, \apj, 867, 108
+
+\bibitem[Jones et al.(2018)b]{jones2018b} Jones, D.~O., Scolnic, D.~M., Riess, A.~G., et al.\ 2018, \apj, 857, 51
 
@@
 
-\bibitem[Kessler, \& Scolnic(2017)]{kessler2017} Kessler, R., \& Scolnic, D.\ 2017, \apj, 836, 56
-
-\bibitem[Knox, \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv e-prints, arXiv:1908.03663
+\bibitem[Kessler \& Scolnic(2017)]{kessler2017} Kessler, R., \& Scolnic, D.\ 2017, \apj, 836, 56
+
+\bibitem[Knox \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv e-prints, arXiv:1908.03663
 
@@
 
+\bibitem[Roman et al.(2018)]{roman2018} Roman, M., Hardin, D., Betoule, M., et al.\ 2018, \aap, 615, A68
+
+\bibitem[Rose et al.(2019)]{rose2019} Rose, B.~M., Garnavich, P.~M., \& Berg, M.~A.\ 2019, \apj, 874, 32
+
+\bibitem[Rubin et al.(2015)]{rubin2015} Rubin, D., Aldering, G., Barbary, K., et al.\ 2015, \apj, 813, 137
+
+\bibitem[Rubin \& Hayden(2016)]{rubin2016} Rubin, D., \& Hayden, B.\ 2016, \apjl, 833, L30
+
 % S
@@
 \bibitem[Scolnic et al.(2014)]{scolnic2014} Scolnic, D., Rest, A., Riess, A., et al.\ 2014, \apj, 795, 45
+
+\bibitem[Scolnic \& Kessler(2016)]{scolnic2016} Scolnic, D., \& Kessler, R.\ 2016, \apjl, 822, L35
 
@@
 % W
+\bibitem[Wiseman et al.(2020)]{wiseman2020} Wiseman, P., Smith, M., Childress, M., et al.\ 2020, arXiv e-prints, arXiv:2001.02640
+
 \bibitem[Wong et al.(2019)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F., et al.\ 2019, arXiv e-prints, arXiv:1907.04869
@@
 \end{thebibliography}
-
 \end{document}
@@
 
-To test the coherence of this model, we also fitted it using all the data from the surveys indistinctly. These results are shown figure \ref{fig:model_all}.
+To test the coherence of this model, we also fitted it using all the data from
+the surveys indistinctly. These results are shown figure \ref{fig:model_all}.
+
+\begin{table*}
+    \centering
+    \caption{BBC comparison}
+    \label{tab:bbc_comp}
+    \begin{tabular}{c c |c c c |c c c |c c c}\hline\hline\\[-1.05em]
+     &  & \multicolumn{3}{|c}{Asym (this)} & \multicolumn{3}{|c}{Asym (SK18/S18)} & \multicolumn{3}{|c}{Base SNf} \\
+    N$_{\mathrm{SN}}$ & Survey & $\mathcal{L}$ & AICc & Free param & $\mathcal{L}$ & AICc & Free param & $\mathcal{L}$ & AICc & Free param \\\hline
+    167 & SDSS & 445.0 & 451.2 & 3 & 457.9 & 464.0 & 3 & 452.9 & 452.9 & 0\\
+    160 & PS1 & 397.4 & 403.6 & 3 & 397.6 & 403.7 & 3 & 405.2 & 405.2 & 0\\
+    102 & SNLS & 252.7 & 259.0 & 3 & 254.5 & 260.8 & 3 & 269.5 & 269.5 & 0\\
+    114 & SNf & 299.3 & 305.5 & 3 & -- & -- & -- & 268.6 & 279.1 & 5\\
+    543 & Total & 1394.5 & 1419.1 & 12 & -- & -- & -- & 1396.0 & 1406.2 & 5\\
+    429 & Total-SNf & 1095.2 & 1113.6 & 9 & 1110.0 & 1128.4 & 9 & 1127.5 & 1127.5 & 0\\\hline\hline
+    \end{tabular}
+\end{table*}
+
+\begin{table*}
+    \centering
+    \caption{BBC comparison}
+    \label{tab:bbc_comp}
+    \begin{tabular}{c |c c c |c c c}\hline\hline\\[-1.05em]
+     & \multicolumn{3}{|c}{Asym (this)}  & \multicolumn{3}{|c}{Base SNf} \\
+    Survey & $\mathcal{L}$ & AICc & Free param & $\mathcal{L}$ & AICc & Free param \\\hline
+    SNf & 299.3 & 305.5 & 3 & 268.6 & 279.1 & 5\\
+    SDSS & 445.0 & 451.2 & 3 & 452.9 & 452.9 & 0\\
+    PS1 & 397.4 & 403.6 & 3 & 405.2 & 405.2 & 0\\
+    SNLS & 252.7 & 259.0 & 3  & 269.5 & 269.5 & 0\\
+    Total & 1394.5 & 1419.1 & 12 & 1396.0 & 1406.2 & 5\\\hline\hline
+    \end{tabular}
+\end{table*}\documentclass[]{aa} %bibnumber

### 2020-01-27 Nicolas.tex
Big changes. Version 1.5 at least!

@@
 \begin{document}
-%\title{Evolution of the Type Ia Supernova Underlying Stretch Distribution as a Function of Redshift}
-\title{Redshift Evolution of the Type Ia Supernova Population Stretch Distribution}
-
-%\subtitle{II. An example text with infinitesimal scientific value whose title
-%and subtitle may also be split}
+\title{Redshift Evolution of the Type Ia Supernova Population Stretch
+Distribution}
 
@@
 
-%\offprints{R. Plemmons, \email{plemmons@...}}
-
 \institute{Université de Lyon, F-69622, Lyon, France; Université de Lyon
@@
     {We started by getting a complete sample representing all of the stretch
-        distribution that Nature can give us, before using LsSFR measurements, an
-        age tracer which evolution with redshift is known, that has been shown
-        to have a strong correlation with the stretch. We compare their AIC, an
-    estimator of the relative quality of statistical models that includes the
-number of free parameters, to determine which ones describe best the data.}
+        distribution that Nature can give us, before using LsSFR measurements,
+        an age tracer which evolution with redshift is known, that has been
+        shown to have a strong correlation with the stretch. We compare their
+        AIC, an estimator of the relative quality of statistical models that
+    includes the number of free parameters, to determine which ones describe
+best the data.}
     {Models with an evolution of the stretch with the redshift have a better
@@
 \section{Introduction}
-
-%\begin{figure*}
-%    \centering
-%    \subfigure[Method]{\label{fig:zdet}%        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}
-
-%        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}
-%    \subfigure[Result]{\label{fig:surveys_cuts}
-%    \includegraphics[width=8.5cm]{A    \subfigure[Method]{\label{fig:zdet}
-%rticle_figures/surveys_cuts_x.pdf}}
-%    \subfigure[Method]{\label{fig:zdet}
-
-%    \caption{(a) $z\St{max}$ determination from the magnitude limited data; (b) SDSS, %SNLS and PS1 samples cut at $z\St{max}$ as explained in
-%    section \ref{sec:sam}. The data we use is in plain markers.}
-%\subfigure[Method]{\label{fig:zdet}
-
-%\end{figure*}
-
 Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
@@
 
-We present fig.~\ref{fig:maglim} the lightcurve stretch and color distributions
-of SNe~Ia from these surveys, namely: PanStarrs \citep[PS1][]{rest2014}, the
-Sloan Digital Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy
-Survey \citep[SNLS][]{astier2006}. An ellipse in the salt2.4 $(x_1, c)$ plan
-with semi-major axis $x_1=3$ and semi-minor axis $c=0.3$ encapsulates the full
-distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
-\citet{campbell2013} for similar contours, the second using a more conservative
-$c=0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and  $c=0$ is
-$M_0=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute
+We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
+distributions of SNe~Ia from these surveys, namely: PanStarrs
+\citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
+\citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
+\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
+c)$ plan with semi-major axis $x_1=3$ and semi-minor axis $c=0.3$ encapsulates
+the full distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011}
+and \citet{campbell2013} for similar contours, the second using a more
+conservative $c=0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and
+$c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute
 standardized magnitude at maximum of light along the aforementioned ellipse
@@
 \cite{scolnic2018a}. The faintest SN~Ia is that with $x_1=-1.66$ and $c=0.25$
-and has an absolute standardized magnitude at peak in Bessel-b band of
+and has an absolute standardized magnitude at peak in Bessel-B band of
 $-18.31\,\mathrm{mag}$. Since one need to detect this object typically a week
@@
 maximum redshift above which the faintest SNe~Ia will be missed \nn{using the
-link between apparent magnitude, redshift and absolute magnitude given by:}
+relation between apparent magnitude, redshift and absolute magnitude given by:}
 \begin{equation}\label{eq:m_lim}
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_all.pdf}
-    \caption{SALT2.4 stretch and color lightcurve parameters of SNe~Ia from the
-        SDSS, PS1 and SNLS samples from the pantheon catalog. The individual SN
-        data are shown as blue dots and a 2D histogram is shown in gray to
-        highlight point density.  The ellipse ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$)
-    is displayed, colored by the corresponding standardized absolute magnitude
-using the $\alpha$ and $\beta$ coefficients from \citep{scolnic2018a}.}
+    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
+    \caption{\textsc{\texttt{SALT2.4}} stretch and color lightcurve parameters
+        of SNe~Ia from the SDSS, PS1 and SNLS samples from the pantheon catalog.
+        The individual SN data are shown as blue dots and a 2D histogram is
+        shown in gray to highlight point density. The ellipse
+        ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$) is displayed, colored by the
+        corresponding standardized absolute magnitude using the $\alpha$ and
+        $\beta$ coefficients from \citep{scolnic2018a}. \nn{The lines represent
+        the $(x_1, c)$ evolution for $m = m_{lim}$ in eq.~\ref{eq:m_lim},
+    for $z=0.60$ and $z=0.70$ using SNLS's $m_{lim}$ of $24.8\,$mag. The data
+    above these lines are those for which $m > m_{lim}$, meaning they can't be
+    observed.}}
     \label{fig:maglim}
@@
 these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit for
-the rest of \nn{this study}.  For the rest of the analysis, we will consider
-both cases, the \nn{aforementioned} redshift limits as well as more conservative
-cuts for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$
-for PS1 and $z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}.
+the rest of \nn{this study}. For the rest of the analysis, we will consider both
+cases, the \nn{aforementioned} redshift limits as well as more conservative cuts
+for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for
+PS1 and $z_{lim}=0.55$ for SNLS (following Fig.~14 of \citealt{perrett2010}).
 
 The complete sample selection is summarized table~\ref{tab:sample} and the
-redshift distribution of these three surveys in are shown fig.~\ref{fig:cuts}.
-We see in this figure that the redshift limit we have selected roughly
-correspond to the peak of these histograms as expected.
+redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
+We see in this figure that the redshift limits we have selected roughly
+correspond to the peak of these histograms\nn{. As the peaks indicate that we
+are starting to miss the faintest SNe~Ia, it seems logical to have the redshift
+cuts fall near said peaks.}
 
@@
 been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
-4.2.2). As the searches were much deeper than the spectrophotometric follow-up
-made by SNfactory \nn{(SNf)}, SNe~Ia within a redshift range of $0.02<z<0.09$
-\citep[as in ][]{rigault2018} should also be a random sampling of the underlying
+4.2.2). \nn{Their spectroscopic follow-up were only done on a redshift range of
+$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
+deeper. As such, these SNe~Ia} should also be a random sampling of the underlying
 SN population. The SNf data is particularly useful for studying SN property
@@
     \centering
-    \caption{\nn{Source surveys and number of SNe~Ia used in our "complete"
+    \caption{\nn{Source surveys and number of SNe~Ia used in our ''complete''
     sample. Conservative cuts are indicated in brackets.}}
@@
 
-%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, fig. 10 suggest still
+%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, Fig. 10 suggest still
 %okat z=0.2}
@@
 
-In \cite{rigault2018} we presented a modeling of the evolution of the fraction
+In~\cite{rigault2018} we presented a modeling of the evolution of the fraction
 of prompt and delayed SNe~Ia as a function of redshift following former work on
@@
 In short, we assumed that the number of prompt SNe~Ia follows the star formation
-activity in the Universe while the number of delayed SNe~Ia follows the number
-of Gyr-old stars in the Universe, i.e. the stellar mass. Hence, if we denote
-$\delta(z)$ (resp. $\psi(z)$) the fraction of young (resp. old) SNe~Ia in the
-Universe as a function of redshift, then their ratio $\delta/\psi$ is expected
-to follow the evolution of the specific star formation rate which goes as
-$(1+z)^{2.8}$ until $z\sim2$ \citep[e.g.,][]{tasca2015}. Hence since
-\nn{$\delta(0.05) \sim \psi(0.05)$} \citep{rigault2013,rigault2018,wiseman2020},
-in agreement with rate expectations \citep{mannucci2006,rodney2014}, \nn{and
-$\delta + \psi = 1$,} then we find in \cite{rigault2018}:
+activity \nn{(SFR)} in the Universe while the number of delayed SNe~Ia follows
+the number of Gyr-old stars in the Universe, i.e. the stellar mass \nn{(M)}.
+Hence, if we denote $\delta(z)$ (resp. $\psi(z)$) the fraction of young (resp.
+old) SNe~Ia in the Universe as a function of redshift, then their ratio
+$\delta/\psi$ is expected to follow the evolution of the specific star formation
+rate \nn{(SFR/M)} which goes as $(1+z)^{2.8}$ until $z\sim2$
+\citep[e.g.,][]{tasca2015}. Hence since \nn{$\delta(0.05) \sim \psi(0.05)$}
+\citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
+\citep{mannucci2006,rodney2014}, \nn{and $\delta + \psi = 1$,} then we find in
+\cite{rigault2018}:
 \begin{align}
@@
     \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
-    \caption{\textit{Main}: Salt2.4 lightcurve stretch ($x_1$) as a function of
+    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$) as a function of
         the local specific star formation rate (LsSFR) for SNf data used in this
@@
 
-\nn{In \cite{rigault2018} we presented the relation between SNe stretches and
+\nn{In~\cite{rigault2018} we presented the relation between SNe stretches and
 LsSFR measurements, which trace progenitor age, from the SNf sample, and is
-shown fig.~\ref{fig:stretchlssfr}}. To model the evolution of the SN stretch as
-a function of redshift given our aforementioned model of the evolution of the
-fraction of young and old SNe~Ia with redshift, we need to model the SN stretch
-distribution for each age subsample. Given the structure of the stretch--LsSFR
-scatter plot shown in fig.~\ref{fig:stretchlssfr}, we define our model as
-follows: for the young population the underlying stretch distribution is modeled
-as $\mathcal{N}(\mu_1, \sigma_1^2)$, i.e. a normal distribution centered on
-$\mu_1$ with a $\sigma_1$ width; the old population stretch distribution is
-modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
+shown in Fig.~\ref{fig:stretchlssfr}}. To model the evolution of the SN stretch
+as a function of redshift, given our aforementioned model of the evolution of
+the fraction of young and old SNe~Ia with redshift, we need to model the SN
+stretch distribution for each age subsample. Given the structure of the
+stretch--LsSFR scatter plot shown in Fig.~\ref{fig:stretchlssfr}, we define our
+model as follows: for the young population the underlying stretch distribution
+is modeled as $\mathcal{N}(\mu_1, \sigma_1^2)$, i.e. a normal distribution
+centered on $\mu_1$ with a $\sigma_1$ width; the old population stretch
+distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
 \mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a gaussian mixture where one mode is the
 same as the young population one. \nn{When combining our prediction of the
-evolution of young SNe~Ia as a function of redshift eq.~\ref{eq:delta}, and the
-stretch distribution of both young and SNe~Ia, our model for the underlying
-distribution of stretch $X_1\left(z\right)$ as a function of redshift is:}
-
+evolution of young SNe~Ia as a function of redshift (eq.~\ref{eq:delta}),
+and the stretch distribution of both young and old SNe~Ia, our model for the
+underlying distribution of stretch $X_1\left(z\right)$ as a function of
+redshift is:}
 \begin{align}
@@
     \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
-    \mathrm{d}x^{i}_{1}, p(\nn{\text{young}})^{i}}\right),
+    \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}}\right),
 \end{equation}
@@
 \nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
-$p(\text{young})^{i}$ are the salt2.4 stretch, error of the salt2.4 stretch and
-the probability that the SN is young, respectively. Depending on the data
-available, there are two ways to fit these 5 free parameters}.
+    $p_y{}^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
+    \textsc{\texttt{SALT2.4}} stretch and the probability that the SN is young,
+    respectively. Depending on the data available, there are two ways to fit
+these 5 free parameters}.
 
@@
 
-\nn{When using SNf data, we choose $p(\text{young})$ to be their given $p(y)$,
-computed from their LsSFR data (see fig.~\ref{fig:stretchlssfr}). Thus, we would
-have $\mathcal{P}$ from eq.~\ref{eq:likelihood} to be}
+\nn{When using SNf data, we choose $p_y{}^i$ to be their given
+$p_y(\mathrm{LsSFR}^i)$, computed from their LsSFR data (see
+Fig.~\ref{fig:stretchlssfr}). Thus, we would have $\mathcal{P}$ from
+eq.~\ref{eq:likelihood} to be}
 
@@
     \label{eq:likelihoodsnf}
-    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, p(\nn{\text{young}})^{i}} =
-    p(y)^{i} \times &\mathcal{N}\left(\mu_1,
+    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}} =
+    \nn{p_y(\mathrm{LsSFR}^i)} \times &\mathcal{N}\left(\mu_1,
     \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
     \nonumber\\
-    (1-p(y)^{i}) \times \nn{\bigg[} a \times & \mathcal{N}\left(\mu_1,
+    (1-\nn{p_y(\mathrm{LsSFR}^i)}) \times \nn{\bigg[} a \times &
+        \mathcal{N}\left(\mu_1,
         \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
@@
 
-In practice, since "$a$" is defined between 0 and 1, we fit for $\alpha$ such
+In practice, since ''$a$'' is defined between 0 and 1, we fit for $\alpha$ such
 that $a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.
 \mr{Results on fitting the SNf data with this model are shown
-table~\ref{tab:modelresults} and illustrated in fig.~\ref{fig:stretchlssfr}.}
+table~\ref{tab:modelresults} and illustrated in Fig.~\ref{fig:stretchlssfr}.}
 
@@
 Given eq.~\ref{eq:stretchz}, we can extend the analysis to non-SNfactory sample
-by fitting the free parameters of the model
-($\theta\equiv{\mu_1,\mu_2,\sigma_1,\sigma_2,a}$) assuming the evolution of the
-fraction of young SNe~Ia as a function of redshift given by $\delta(z)$ and
-given, this time, $x_1^{i}$,  $\mathrm{d}x^{i}_{1}$ and $z^{i}$, the stretch,
-error on the stretch and redshift of any given SN $i$. In practice, it means
-that we are still minimizing for the same eq.~\ref{eq:likelihood}, but replacing
-$p(y)^{i}$ by $\delta(z^{i})$ eq.~\ref{eq:likelihoodsnf}.
+by fitting the free parameters of the model ($\theta\equiv{\mu_1, \mu_2,
+\sigma_1, \sigma_2, a}$) assuming the evolution of the fraction of young SNe~Ia
+as a function of redshift given by $\delta(z)$ and given, this time, $x_1^{i}$,
+$\mathrm{d}x^{i}_{1}$ and $z^{i}$, the stretch, error on the stretch and
+redshift of any given SN $i$. In practice, it means that we are still minimizing
+for the same eq.~\ref{eq:likelihood}, but \nn{using $p_y{}^i = \delta(z^{i})$,
+which is equivalent to replacing $p_y(\mathrm{LsSFR}^i)$ with $\delta(z^i)$ in
+eq.~\ref{eq:likelihoodsnf}.}
 
 \mr{For the rest of the analysis, we will therefore fit for
-eq.~\ref{eq:likelihood} using $p(y)^{i}$ --~the probability for the SN $i$ to be
-young~-- when available (i.e. for SNf dataset) and $\delta(z^{i})$ --~the
-expected fraction of young SNe~Ia at the SN redshift $z^{i}$~-- otherwise.}
+eq.~\ref{eq:likelihood} using \nn{$p_y(\mathrm{LsSFR}^i)$} --~the
+probability for the SN $i$ to be young~-- when available (i.e. for SNf
+dataset) and $\delta(z^{i})$ --~the expected fraction of young SNe~Ia at the
+SN redshift $z^{i}$~-- otherwise.}
 
@@
 with this model are given table~\ref{tab:modelresults} and illustrated in
-fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
+Fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
 stretch per bin of redshift follow our redshift drift modeling. That is, when
@@
     \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
-    \caption{Evolution of the mean SN salt2.4 stretch ($x_1$) as a function of
-        redshift.  Markers show the mean stretch measured in redshift bins of
-        equal sample size indicated in light gray at the bottom of each redshift
-        bins. Full and light markers are used when considering the nominal or
-        the conservative sample, respectively, and are colored by their mean
-        redshift.  The orange horizontal line represent the mean redshift of the
-    nominal sample illustrating the expectation if the SN stretch distribution
-is not drifting. Best fits of our Base drifting model are shown as blue,
-dashed-blue and gray, when fitted on the nominal, the conservative or the SNf
-dataset, respectively; all are similar. The light-blue illustrate the amplitude
-of the error of the best fit model when considering the nominal dataset.}
+    \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
+        as a function of redshift.  Markers show the mean stretch measured in
+        redshift bins of equal sample size indicated in light gray at the bottom
+        of each redshift bins. Full and light markers are used when considering
+        the nominal or the conservative sample, respectively, and are colored by
+        their mean redshift.  The orange horizontal line represent the mean
+        redshift of the nominal sample illustrating the expectation if the SN
+        stretch distribution is not drifting. Best fits of our Base drifting
+        model are shown as blue, dashed-blue and gray, when fitted on the
+    nominal, the conservative or the SNf dataset, respectively; all are similar.
+The light-blue illustrate the amplitude of the error of the best fit model when
+considering the nominal dataset.}
     \label{fig:modelall}
@@
 Results are presented in table \ref{tab:comp} and are illustrated
-fig.~\ref{fig:mod_comp}.
+Fig.~\ref{fig:mod_comp}.
 
@@
 that a model is at least as good as the reference model to describe the data.
-See also fig.~\ref{fig:mod_comp}.}
+See also Fig.~\ref{fig:mod_comp}.}
     \label{tab:comp}
@@
 \mr{We find that underlying stretch distributions with no redshift drift are all
-excluded to describe the data as well as the best redshift models (i.e. Base and
-Base$-(\sigma_2)$). In fact, the best non drifting model (Asymmetric) has
-$2\times10^{-4}\%$ of chance to describe the data as well as Base$-(\sigma_2)$,
-the best model.  This is visible fig~\ref{fig:modelall} where the mean stretches
-per bin of redshift of equal sample size suggest a redshift evolution following
-our Base model --~fitted either on SNf, on the nominal 569 SNe~Ia dataset or the
-conservative one~-- rather than having a constant mean value.}
-
-\mr{As just mentioned, the best model is Base$-(\sigma_2)$ where both the
-    low-stretch (mode 2 that only exists in old environments) and high-stretch
-    (mode 1 that exists in all environments) normal distributions share the same
-    width. It is marginally better than the Base model where both distributions
-    have their own ($\Delta \mathrm{AICc}<2$). It is unclear which physical
-    mechanism would translate into a simple distribution shift of mode 1 to a
-    mode 2 that would only exist in old populations. Two distinct underlying
-    distributions, as assumed by the Base model, seems more natural, yet, more
-    data would be necessary to discriminate between these models.  Similarly,
-letting the high-stretch distribution to vary between the young and the old
-population (Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$) only marginally
-improves the fit ($\Delta\mathcal{L}=3.6$ in comparison to Base at a cost of 2
-more degrees of freedom, making this mode less likely. This suggests that,
-indeed, the high-stretch mode (mode 1) exists in all kind of age-populations.}
-\mr{We also remark that considering the stretch distribution of each
-age-population as independent normal distributions, as suggested by
-\cite{howell2007}, is 10 times less likely than our best model and is therefore
-unfavored.}
-
+    excluded to describe the data as well as the Base redshift model. In fact,
+    the best non drifting model (Asymmetric) has $2\times10^{-4}\%$ of chance to
+    describe the data as well as Base. This is visible fig~\ref{fig:modelall}
+    where the mean stretches per bin of redshift of equal sample size suggest a
+    redshift evolution following our Base model --~fitted either on SNf, on the
+    nominal 569 SNe~Ia dataset or the conservative one~-- rather than having a
+constant mean value.}
+
+\mr{As just mentioned, the \nn{reference model is Base}, but we also implemented
+    a Base$-(\sigma_2)$ model, where both the low-stretch (mode 2 that only
+    exists in old environments) and high-stretch (mode 1 that exists in all
+    environments) normal distributions share the same width. It is marginally
+    better than the Base model where both distributions have their own
+    \nn{dispersion} ($\Delta \mathrm{AIC}<2$). Similarly, letting the
+    high-stretch distribution to vary between the young and the old population
+    (\nn{model we call} Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$) only
+    marginally improves the fit ($\Delta\mathcal{L}=3.6$ in comparison to Base)
+    at a cost of 2 more degrees of freedom, making this mode less likely. This
+    suggests that, indeed, the high-stretch mode (mode 1) exists in all kind of
+    age-populations. We also remark that considering the stretch distribution of
+    each age-population as independent normal distributions, as suggested by
+\cite{howell2007}, is \nn{significantly disfavored}.}
+
+\mr{Our results strongly suggest that the underlying distribution of SN~Ia
+    stretch evolves as a function of redshift. Our redshift drift modeling
+    presented section~\ref{sec:modeling} is a much better representation of the
+    pantheon data (see section~\ref{sec:sample}) than any non-drifting models,
+    including the sample-based BBC one. Considering or not the conservative
+sample has no effect on our conclusion as illustrated Fig.~\ref{fig:modelall}
+and presented table~\ref{tab:comp}.}
+
+\mr{To the best of our knowledge, the SN~Ia stretch redshift drift has never
+    been explicitly accounted for in cosmological analyses. Not doing so is a
+    second order issue for SN cosmology as it only affects the way one does
+    Malmquist bias correction. Indeed, as long as the Phillip's relation
+    standardization parameter $\alpha$ is not redshift dependent (study behind
+    the scope of this paper, but see e.g. \citealt{scolnic2018a}), if a sample
+    is free from selection effects, the stretch-corrected SNe~Ia magnitudes used
+    for cosmology are blind to the underlying stretch distribution.  However,
+    since the first missed SNe~Ia are the faintest, which are also those with
+    the lowest stretch, once we include SNe~Ia affected by selection effects we
+    need to account for the fact that only the brightest SNe~Ia exist in the
+    low-stretch end of the sample after standardization. The modeling of the SN
+    stretch then enters the analysis to understand which SNe~Ia are missing
+given the selection functions to mimic a selection-bias free sample.}
+
+\mr{The state-of-the-art tool for doing so is the BBC described in
+    \cite{scolnic2016} and \cite{kessler2017}, which is used in all recent SN
+    Cosmological analyses \citep{jones2018b, scolnic2018a, brout2019,
+    descosmopaper2019} including the direct measurement of H$_0$
+\citep{riess2016,riess2019}. As already discussed, it currently assumes an
+asymmetric stretch distribution per sample. \nn{A more detailed discussion on
+this modeling is presented section \ref{sec:bbc}.}}
+
+\section{\nn{BBC discussion}}
+\label{sec:bbc}
 \mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
@@
     modelisation. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the best of all
-    the non-drift model, but still $\Delta \mathcal{L}=11.3$ worse than our best
-    model, it is penalized by requiring 15 free parameters, 3 per sample.
+    the non-drift model, but still $\Delta \mathcal{L}=-11.5$ worse than our
+    best model, it is penalized by requiring 15 free parameters, 3 per sample.
     \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
@@
     BBC sample-based technique is excluded to be an as good representation of
-    the data as our best model ($p=4\times 10^{-8}$).  A more detailed
-discussion of the consequence of this results for cosmology follows in
+    the data as our best model ($p=4\times 10^{-8}$). A more detailed discussion
+    of the consequence of this results for cosmology follows in
 section~\ref{sec:bbc}}
 
-\mr{We report in table~\ref{tab:bbc} the sample's $\sigma^-$,  $\sigma^+$ and
+\mr{We report in table~\ref{tab:bbc} the samples' $\sigma^-$, $\sigma^+$ and
     $\mu^0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
@@
     \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
-    the $\Delta\mathrm{AICc}$ between our Base drifting model and the BBC
-    modeling would go from -32 to -47.  Furthermore, fixing the Base model
-    parameters to those derived using SNf data only (top row
-    table.~\ref{tab:modelresults}) and using the BBC parameters from the
+    the $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC
+    modeling would go from -32 to -47. \nn{It would seem that our minimization
+    of these parameters is closer to our Base model than theirs}. Furthermore,
+    fixing the Base model parameters to those derived using SNf data only (top
+    row table.~\ref{tab:modelresults}) and using the BBC parameters from the
     aforementioned literature for PS1, SDSS and SNLS, we find
-    $\Delta\mathrm{AICc}=31.7$. Removing the SNf data reduces this difference to
-    $\Delta\mathrm{AICc}=5$.  Such a reduction is expected for two reasons.
+    $\Delta\mathrm{AIC}=-31.7$. Removing the SNf data reduces this difference to
+    $\Delta\mathrm{AIC}=-5$. Such a reduction is expected for two reasons.
     First, it removes the nearby SNe~Ia, and thereby significantly reduces the
@@
     Second, the SNf data are better at discriminating the individual stretch
-    distributions by age since the probability for a SN to be young $p(y^i)$ is
+    distributions by age since the probability for a SN to be young $p(y)^i$ is
     known while other samples rely on $\delta(z^i)$ which only is the fraction
-    of young SNe~Ia at the SN redshift, i.e. a proxy for $p(y^i)$.  While in any
-case the Base-model is favored, the relative weight of a SNf sample highlight
-the  importance of addition non-targeted nearby surveys such the Zwicky
-Transient Facility \cite[ZTF,][]{bellm2019,graham2019} to unambiguously confirm
-our result.}
-
-%Using S18/SK16 values of \sigma_-, \sigma_+ and \mu_0, we find that SDSS, PS1 and SNLS's $\mathcal{L}$ to be 457.9, 397.6 and 254.5 respectively, and their AICc to be 464.0, 403.7 and 260.8 respectively, systematically higher than those we find in this study (see table \ref{tab:bbc_comp}).
+    of young SNe~Ia at the SN redshift, i.e. a probabilistic proxy for $p(y)^i$.
+    While in any case the Base-model is favored, the relative weight of a SNf
+    sample highlight the importance of addition non-targeted nearby surveys such
+    the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,graham2019} to
+unambiguously confirm our result.}
+
+%Using S18/SK16 values of \sigma_-, \sigma_+ and \mu_0, we find that SDSS, PS1 and SNLS's $\mathcal{L}$ to be 457.9, 397.6 and 254.5 respectively, and their AIC to be 464.0, 403.7 and 260.8 respectively, systematically higher than those we find in this study (see table \ref{tab:bbc_comp}).
 
@@
 
-\mr{Considering or not the conservative sample has no effect on our conclusion
-as illustrated fig.~\ref{fig:modelall} and presented table~\ref{tab:comp}.}
-
-\section{Discussion}
-\label{sec:bbc}
-\mr{Our results strongly suggest that the underlying distribution of SN~Ia
-stretch evolves as a function of redshift. Our redshift drift modeling presented
-section~\ref{sec:modeling} is a much better representation of the pantheon data
-(see section~\ref{sec:sample}) than any non-drifting models, including the
-sample-based BBC one.}
-
-\mr{To the best of our knowledge, the SN~Ia stretch redshift drift has never
-    been explicitly accounted for in cosmological analyses. Not doing so is a
-    second order issue for SN cosmology as it only affects the way one does
-    Malmquist bias correction. Indeed, as long as the Phillip's relation
-    standardization parameter $\alpha$ is not redshift dependent (study behind
-    the scope of this paper, but see e.g. \citealt{scolnic2018a}), if a sample
-    is free from selection effects, the stretch-corrected SNe~Ia magnitudes used
-    for cosmology are blind to the underlying stretch distribution.  However,
-    since the first missed SNe~Ia are the faintest, which are also those with
-    the lowest stretch, once we include SNe~Ia affected by selection effects we
-    need to account for the fact that only the brightest SNe~Ia exist in the
-    low-stretch end of the sample after standardization.  The modeling of the SN
-stretch then enters the analysis to understand which SNe~Ia are missing given
-the selection functions to mimic a selection-bias free sample.}
-
-\mr{The state-of-the-art tool for doing so is the BBC described in
-\cite{scolnic2016} and \cite{kessler2017}, which is used in all recent SN
-Cosmological analyses
-\citep{jones2018b,scolnic2018a,brout2019,descosmopaper2019} including the direct
-measurement of $H_0$ \citep{riess2016,riess2019}. As already discussed, it
-currently assumes an asymmetric stretch distribution per sample and we have
-shown in section~\ref{sec:results} that this modeling is excluded to be an as
-good description of the data than our age-based redshift drift modeling
-eq.~\ref{eq:stretchz}.}
-
 \begin{figure}
@@
     \includegraphics[width=\linewidth]{Article_figures/bbc_comp_PS1.pdf}
-    \caption{Underlying SN salt2.4 stretch ($x_1$) distributions for the PS1
-    sample. The Green line(band) shows our best fitted asymmetric Gaussian
-distribution (its error) mimicking the BBC approach. The dashed line shows that
-estimated by \cite{scolnic2018a}. The blue to red curves display our Base
-drifting stretch model estimated at each PS SN redshift; bluer for lower
-redshifts.}
+    \caption{Underlying SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
+        distributions for the PS1 sample. The Green line(band) shows our best
+        fitted asymmetric Gaussian distribution (its error) mimicking the BBC
+        approach. The dashed line shows that estimated by \cite{scolnic2018a}.
+    The blue to red curves display our Base drifting stretch model estimated at
+each PS SN redshift; bluer for lower redshifts.}
     \label{fig:bbc_pdf_ps1}
@@
 
-\mr{We illustrate fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
+\mr{We illustrate Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
 underlying stretch distribution between the BBC modeling and our Base drifting
@@
 \mr{Nonetheless, to have a rough estimation of the expected amplitude of the
-effect, we show fig.~\ref{fig:magdrift} the difference of mean stretch
+effect, we show Fig.~\ref{fig:magdrift} the difference of mean stretch
 correction $\alpha\times\left(\langle x_1 \rangle_{\mathrm{BBC}} - \langle x_1
@@
      &  & \multicolumn{3}{|c}{Asym (this)} & \multicolumn{3}{|c}{Asym (SK18/S18)} & \multicolumn{3}{|c}{Base SNf} \\
-    N$_{\mathrm{SN}}$ & Survey & $\mathcal{L}$ & AICc & Free param & $\mathcal{L}$ & AICc & Free param & $\mathcal{L}$ & AICc & Free param \\\hline
+    N$_{\mathrm{SN}}$ & Survey & $\mathcal{L}$ & AIC & Free param & $\mathcal{L}$ & AIC & Free param & $\mathcal{L}$ & AIC & Free param \\\hline
     167 & SDSS & 445.0 & 451.2 & 3 & 457.9 & 464.0 & 3 & 452.9 & 452.9 & 0\\
@@
      & \multicolumn{3}{|c}{Asym (this)}  & \multicolumn{3}{|c}{Base SNf} \\
-    Survey & $\mathcal{L}$ & AICc & Free param & $\mathcal{L}$ & AICc & Free param \\\hline
+    Survey & $\mathcal{L}$ & AIC & Free param & $\mathcal{L}$ & AIC & Free param \\\hline
     SNf & 299.3 & 305.5 & 3 & 268.6 & 279.1 & 5\\
@@
     \end{tabular}
-\end{table*}\documentclass[]{aa} %bibnumber
+\end{table*}

### 2020-01-30 Nicolas.tex
Small changes; notations

@@
     \nn{X_1\left(z \right) =
-    }&\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
+    } &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
     (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
@@
     \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
-    \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}}\right),
+    \mathrm{d}x^{i}_{1}, \nn{y^i}}\right),
 \end{equation}
@@
 \nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
-    $p_y{}^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
+    $y^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
     \textsc{\texttt{SALT2.4}} stretch and the probability that the SN is young,
@@
 
-\nn{When using SNf data, we choose $p_y{}^i$ to be their given
-$p_y(\mathrm{LsSFR}^i)$, computed from their LsSFR data (see
-Fig.~\ref{fig:stretchlssfr}). Thus, we would have $\mathcal{P}$ from
-eq.~\ref{eq:likelihood} to be}
+\nn{When using SNf data, we choose $y^i$ to be their given $p_y{}^i$, computed
+from their LsSFR data (see Fig.~\ref{fig:stretchlssfr}). \nn{It represents the
+probability to be above $\log(\mathrm{LsSFR} = -10.82)$.} Thus, we would have
+$\mathcal{P}$ from eq.~\ref{eq:likelihood} to be}
 
@@
     \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}} =
-    \nn{p_y(\mathrm{LsSFR}^i)} \times &\mathcal{N}\left(\mu_1,
+    \nn{p_y{}^i} \times &\mathcal{N}\left(\mu_1,
     \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
     \nonumber\\
-    (1-\nn{p_y(\mathrm{LsSFR}^i)}) \times \nn{\bigg[} a \times &
+        (1-\nn{p_y{}^i}) \times \nn{\bigg[} a \times &
         \mathcal{N}\left(\mu_1,
@@
 redshift of any given SN $i$. In practice, it means that we are still minimizing
-for the same eq.~\ref{eq:likelihood}, but \nn{using $p_y{}^i = \delta(z^{i})$,
-which is equivalent to replacing $p_y(\mathrm{LsSFR}^i)$ with $\delta(z^i)$ in
+for the same eq.~\ref{eq:likelihood}, but \nn{using $y^i = \delta(z^i)$, which
+is equivalent to replacing $p_y{}^i$ with $\delta(z^i)$ in
 eq.~\ref{eq:likelihoodsnf}.}
@@
 \mr{For the rest of the analysis, we will therefore fit for
-eq.~\ref{eq:likelihood} using \nn{$p_y(\mathrm{LsSFR}^i)$} --~the
+eq.~\ref{eq:likelihood} using \nn{$p_y{}^i$} --~the
 probability for the SN $i$ to be young~-- when available (i.e. for SNf
@@
 
-\mr{Results on fitting the all the 569(422) SNe~Ia data (with conservative cuts)
+\mr{Results on fitting all the 569(422) SNe~Ia data (with conservative cuts)
 with this model are given table~\ref{tab:modelresults} and illustrated in
 Fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
-stretch per bin of redshift follow our redshift drift modeling. That is, when
+stretch per bin of redshift follows our redshift drift modeling. That is, when
 considering selection-bias free SN~Ia samples, SNe~Ia at higher redshift have on
@@
         the nominal or the conservative sample, respectively, and are colored by
-        their mean redshift.  The orange horizontal line represent the mean
+        their mean redshift. The orange horizontal line represents the mean
         redshift of the nominal sample illustrating the expectation if the SN
@@
     nominal, the conservative or the SNf dataset, respectively; all are similar.
-The light-blue illustrate the amplitude of the error of the best fit model when
+The light-blue illustrates the amplitude of the error of the best fit model when
 considering the nominal dataset.}
@@
 
-The ability these models to explain the data is detailed in
+The ability of these models to explain the data is detailed in
 section~\ref{sec:results}.
@@
     \centering
-    \caption{Comparison relative ability for a model to describe the data using
-    AIC criterion. See the model definition section~\ref{sec:modeling}. The
-drift column indicates if the model contains the age-drifting model
-($\delta(z)$) or not (f, i.e. non-drifting). The BBC modeling is sample based
-which somewhat contains redshift dependency information. $\mathcal{L}$ is the
-pseudo-$chi^{2}$ defined eq.~\ref{eq:likelihood} and Proba is the probability
-that a model is at least as good as the reference model to describe the data.
+    \caption{Comparison of the relative ability for a model to describe the data
+        using AIC criterion. See the model definition
+        section~\ref{sec:modeling}. The drift column indicates if the model
+        contains the age-drifting model ($\delta(z)$) or not (f, i.e.
+        non-drifting). The BBC modeling is sample based which somewhat contains
+        redshift dependency information. $\mathcal{L}$ is the pseudo-$\chi^{2}$
+        defined eq.~\ref{eq:likelihood} and Proba is the probability that a
+        model is at least as good as the reference model to describe the data.
     See also Fig.~\ref{fig:mod_comp}.}
@@
     highlighted that, because current surveys span limited redshift ranges, the
-    per sample approach somewhat account for unmodeled redshift drifts. We
-    stress here that, as measure of modern surveys cover larger redshift ranges
-    to reduce calibration systematic  uncertainties, this is becoming less true,
-    notably for PS1, DES and, soon, LSST. Already with the current surveys, the
-    BBC sample-based technique is excluded to be an as good representation of
-    the data as our best model ($p=4\times 10^{-8}$). A more detailed discussion
-    of the consequence of this results for cosmology follows in
-section~\ref{sec:bbc}}
+    per sample approach somewhat accounts for unmodeled redshift drifts. We
+    stress here that, as measurements of modern surveys cover larger redshift
+    ranges to reduce calibration systematic  uncertainties, this is becoming
+    less true, notably for PS1, DES and, soon, LSST. Already with the current
+    surveys, the BBC sample-based technique is excluded to be an as good
+    representation of the data as our best model ($p=4\times 10^{-8}$). A more
+    detailed discussion of the consequence of this results for cosmology follows
+in section~\ref{sec:bbc}}
 
@@
     $\mu^0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
-    SDSS and with \cite{scolnic2018a} for PS1. This validates our somewhat
+    SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{the} somewhat
     simple construction of our selection-bias free sample, see
@@
     Second, the SNf data are better at discriminating the individual stretch
-    distributions by age since the probability for a SN to be young $p(y)^i$ is
+    distributions by age since the probability for a SN to be young $p_y{}^i$ is
     known while other samples rely on $\delta(z^i)$ which only is the fraction
-    of young SNe~Ia at the SN redshift, i.e. a probabilistic proxy for $p(y)^i$.
-    While in any case the Base-model is favored, the relative weight of a SNf
-    sample highlight the importance of addition non-targeted nearby surveys such
-    the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,graham2019} to
-unambiguously confirm our result.}
+    of young SNe~Ia at the SN redshift, i.e. a probabilistic proxy for
+    $p_y{}^i$. While in any case the Base-model is favored, the relative weight
+    of a SNf sample highlights the importance of adding non-targeted nearby
+surveys, such as the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,
+graham2019} to unambiguously confirm our result.}
 
@@
         fitted asymmetric Gaussian distribution (its error) mimicking the BBC
-        approach. The dashed line shows that estimated by \cite{scolnic2018a}.
-    The blue to red curves display our Base drifting stretch model estimated at
-each PS SN redshift; bluer for lower redshifts.}
+        approach. The dashed line shows the noe estimated by
+    \cite{scolnic2018a}. The blue to red curves display our Base drifting
+stretch model estimated at each PS SN redshift; bluer for lower redshifts.}
     \label{fig:bbc_pdf_ps1}
@@
 distribution is of the order of few tens of millimag, which is the typical scale
-exotic forms of dark energy could have on $\mu(z)$. In the era of modern
+that exotic forms of dark energy could have on $\mu(z)$. In the era of modern
 cosmology where we aim at probing $w_0$ at a sub-percent and $w_a$ at a few
@@
     \includegraphics[width=\linewidth]{Article_figures/BBC_distmod_w0wa.pdf}
-    \caption{ Difference of mean stretch standardization (in mag) between the
-        BBC algorithm used in \cite{scolnic2018a} and our Base drifting model.
-        The blue markers represent the mean evolution in 30 regular redshift
-        bins, while the small colored points show the SDSS, PS1 and SNLS
-        samples. For comparison, we show in gray bands the expected magnitude
-        difference between $\Lambda$CDM  and $w$CDM cosmologies ($w=-1\pm
-        10,\,5\,\mathrm{and}\,1\%$) or, in orange, more exotic $w_0,w_a$CDM
-        examples.  This figure illustrates that the amplitude of the potential
-    magnitude bias caused by inaccurate modeling of the underlying stretch
-distribution is comparable to the observational effect of exotic form of dark
-energy.}
+    \caption{Difference of mean stretch standardization (in mag) between the BBC
+        algorithm used in \cite{scolnic2018a} and our Base drifting model.
+        \nn{The small colored points show the SDSS, PS1 and SNLS evolution,
+            while the orange markers represent their mean evolution in 30
+            regular redshift bins.} For comparison, we show in gray bands the
+            expected magnitude difference between $\Lambda$CDM and $w$CDM
+            cosmologies ($w=-1\pm 10,\,5\,\mathrm{and}\,1\%$) or, in orange,
+            more exotic $w_0,w_a$CDM examples. This figure illustrates that the
+            amplitude of the potential magnitude bias caused by inaccurate
+        modeling of the underlying stretch distribution is comparable to the
+    observational effect of exotic form of dark energy.}
     \label{fig:magdrift}
@@
 \label{sec:ccl}
-We have presented a study of the drift of the underlyin SNe~Ia stretch
+We have presented a study of the drift of the underlying SNe~Ia stretch
 distribution as a function of redshift. We used SNe~Ia from magnitude-limited
-surveys from the pantheon data set \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
+surveys from the pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
 well as HST to which we added SNfactory data from \cite{rigault2018} for the
@@
 way the observed SNe~Ia stretch are random sampling of the true underlying
-distribution. This corresponds the 569 SNe~Ia (422 when considering more
+distribution. This corresponds to 569 SNe~Ia (422 when considering more
 conservative cuts).
@@
 Following observations made in \cite{rigault2018} we introduced a redshift drift
-modeling that depend on the expected fraction of young and old SNe~Ia as a
+modeling that depends on the expected fraction of young and old SNe~Ia as a
 function of redshift, each age-population having its own underlying stretch
@@
 modeling. We also closely studied the stretch distribution modeling made in
-current version of the BCC algorithm used for Malquist bias correction by all
+current version of the BCC algorithm used for Malmquist bias correction by all
 recent Dark energy or Hubble constant cosmological analyses.
 
-Our conclusion are the following:
+Our conclusions are the following:
 \begin{enumerate}
 
-    \item Non-drifting models are excluded to be as good description of the data
-        in comparison to our Base modeling.  This model assumes that: (1) the
-        young population has a unimodal Gaussian stretch distribution while the
-        old population stretch one is bimodal, one mode been the young one; (2)
-        the evolution of the relative fraction of young and old SNe~Ia follows
-        the prediction made in \cite{rigault2018}.
+    \item Non-drifting models are excluded to be as good descriptions of the
+        data \nn{as} our Base modeling. This model assumes that: (1) the young
+        population has an unimodal Gaussian stretch distribution while the old
+        population stretch one is bimodal, one mode being the young one; (2) the
+        evolution of the relative fraction of young and old SNe~Ia follows the
+        prediction made in \cite{rigault2018}.
 

### 2020-02-03 Nicolas.tex
New plot

@@
     \and Y. Copin\inst{1}
-    \and Y. Kim\inst{1}
+    \and Y.-L. Kim\inst{1}
 }
@@
 \nn{being brighter on average}
-\cite[e.g.][]{kelly2010,sullivan2010,childress2013,betoule2014,rigault2018}.
+\cite[e.g.][]{kelly2010, sullivan2010, childress2013, betoule2014, rigault2018, kim19}.
 This mass-step correction is currently used in cosmological analyses
-\citep[e.g.][]{betoule2014,scolnic2018a}, including for deriving H$_0$
-\citep{riess2016,riess2019}. Yet, the underlying connection between the SN and
+\citep[e.g.][]{betoule2014, scolnic2018a}, including for deriving H$_0$
+\citep{riess2016, riess2019}. Yet, the underlying connection between the SN and
 the host remains unclear when using global properties such as the host stellar
@@
 the local color are more strongly correlated with the standardized SN magnitude
-\citep{rigault2018,roman2018}, suggesting the age to be the driving parameter
+\citep{rigault2018, roman2018, kim18}, suggesting the age to be the driving parameter
 underlying the mass-step. If true, this would have a significant impact for
@@
 \cite[e.g.][]{jones2015,jones2019} and especially the impact it has on the
-derivation of H$_0$ \cite{jones2015,riess2016,riess2018,rose2019}.
+derivation of H$_0$ \citep{jones2015,riess2016,riess2018,rose2019}.
 
@@
 sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
-rigault2013, pan2014}. Following the “A+B“ model and the connection between SN
+rigault2013, pan2014, kim19}. Following the “A+B“ model and the connection between SN
 stretch and host properties, \cite{howell2007} first discussed the potential
@@
 Consequently, if selection effects are not accounted for, one might confuse true
-population drift with selection effects, and conversely.
+population drift with selection effects, and conversely \nn{\citep[e.g.,][]{kim19}}.
 
@@
 limited surveys.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
+    \caption{\textsc{\texttt{SALT2.4}} stretch and color lightcurve parameters
+        of SNe~Ia from the SDSS, PS1 and SNLS samples from the pantheon catalog.
+        The individual SN data are shown as blue dots and a 2D histogram is
+        shown in gray to highlight point density. The ellipse
+        ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$) is displayed, colored by the
+        corresponding standardized absolute magnitude using the $\alpha$ and
+        $\beta$ coefficients from \cite{scolnic2018a}. \nn{The lines represent
+        the $(x_1, c)$ evolution for $m = m_{lim}$ in eq.~\ref{eq:m_lim},
+    for $z$ between $0.60$ and $z=1.00$ using SNLS's $m_{lim}$ of $24.8\,$mag. The data
+    above these lines are those for which $m > m_{lim}$, meaning they can't be
+    observed.}}
+    \label{fig:maglim}
+\end{figure}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
-    \caption{\textsc{\texttt{SALT2.4}} stretch and color lightcurve parameters
-        of SNe~Ia from the SDSS, PS1 and SNLS samples from the pantheon catalog.
-        The individual SN data are shown as blue dots and a 2D histogram is
-        shown in gray to highlight point density. The ellipse
-        ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$) is displayed, colored by the
-        corresponding standardized absolute magnitude using the $\alpha$ and
-        $\beta$ coefficients from \citep{scolnic2018a}. \nn{The lines represent
-        the $(x_1, c)$ evolution for $m = m_{lim}$ in eq.~\ref{eq:m_lim},
-    for $z=0.60$ and $z=0.70$ using SNLS's $m_{lim}$ of $24.8\,$mag. The data
-    above these lines are those for which $m > m_{lim}$, meaning they can't be
-    observed.}}
-    \label{fig:maglim}
-\end{figure}
-
 SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{Article_figures/hist_surveys_cuts.pdf}
+    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55.pdf}
     \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
-    SNLS from the pantheon dataset, respectively. The colored part of the
+    SNLS from the pantheon dataset. The colored part of the
 distribution shows which SNe~Ia has been kept in our analysis for they are
@@
 from Pantheon follows the same logic of having a search deeper than the
-follow-up and we therefore kept it entirely \citep{FIND BACK THE REF}, see
-table~\ref{tab:sample}.
+follow-up and we therefore kept it entirely \citep{strolger04}, see
+table~\ref{tab:sample}. \nn{We present the stretch distribution, redshift histogram
+and stretch cdf of these five surveys Fig.~\ref{fig:sample}}.
 
@@
 \end{table}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75.pdf}
+    \caption{\nn{\textit{Main}: Stretches of each survey in full and empty markers for the
+    conservative and nominal cuts, respectively, as a function of redshift. \textit{Right}:
+    normed CDF of each survey's stretch distribution. \textit{Top}: combined redshift 
+    histogram of all the data, in dark gray for the conservative cuts;
+    in light gray for the nominal cuts.}}
+    \label{fig:sample}
+\end{figure}
 
@@
         analysis. The color corresponds to the probability for the SNe~Ia to be
-        above $\log(\mathrm{LsSFR}=-10.82)$, which defines $p(y)$; see
-        \cite{rigault2018}. \textit{Right}: histogram of $p(y)$-weighted
+        above $\log(\mathrm{LsSFR}=-10.82)$, which defines $p_y$; see
+        \cite{rigault2018}. \textit{Right}: histogram of $p_y$-weighted
         marginalization of the SN stretch. On both panel, we illustrate the
@@
 \end{equation}
-
 \nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
@@
     less true, notably for PS1, DES and, soon, LSST. Already with the current
-    surveys, the BBC sample-based technique is excluded to be an as good
+    surveys, the BBC sample-based technique is \nn{ruled out} to be an as good
     representation of the data as our best model ($p=4\times 10^{-8}$). A more
@@
     $\mu^0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
-    SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{the} somewhat
-    simple construction of our selection-bias free sample, see
+    SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{our} construction
+    of \nn{a} selection-bias free sample, see
     section~\ref{sec:sample}. If we were to use the \cite{scolnic2016} and
@@
     First, it removes the nearby SNe~Ia, and thereby significantly reduces the
-    redshift level arm to probe the impact of potential redshift drifts.
+    redshift \nn{lever} arm to probe the impact of redshift drifts.
     Second, the SNf data are better at discriminating the individual stretch
     distributions by age since the probability for a SN to be young $p_y{}^i$ is
-    known while other samples rely on $\delta(z^i)$ which only is the fraction
+    known while other samples rely on $\delta(z^i)$ which is the fraction
     of young SNe~Ia at the SN redshift, i.e. a probabilistic proxy for
     $p_y{}^i$. While in any case the Base-model is favored, the relative weight
-    of a SNf sample highlights the importance of adding non-targeted nearby
+    of the SNf sample highlights the importance of adding non-targeted nearby
 surveys, such as the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,
-graham2019} to unambiguously confirm our result.}
+graham2019} to unambiguously confirm our \nn{model}.}
 
@@
 underlying stretch distribution between the BBC modeling and our Base drifting
-model for the PS1 sample. By definition, our model is bimodal and the relative
+model for the PS1 sample. Our model is bimodal and the relative
 amplitude of each mode depends on the fraction of old SNe~Ia in the sample
@@
 Following observations made in \cite{rigault2018} we introduced a redshift drift
-modeling that depends on the expected fraction of young and old SNe~Ia as a
+modeling \nn{which} depends on the expected fraction of young and old SNe~Ia as a
 function of redshift, each age-population having its own underlying stretch
-distribution. We studied various suites of distribution as well as non-drifting
-modeling. We also closely studied the stretch distribution modeling made in
-current version of the BCC algorithm used for Malmquist bias correction by all
+distribution. We have studied various suites of distribution\nn{, including} non-drifting
+modeling. We also studied the stretch distribution modeling \nn{of the}
+current version of the BBC algorithm used for Malmquist bias correction by all
 recent Dark energy or Hubble constant cosmological analyses.
@@
     \item Non-drifting models are excluded to be as good descriptions of the
-        data \nn{as} our Base modeling. This model assumes that: (1) the young
+        data \nn{as} our Base model. This model assumes that: (1) the young
         population has an unimodal Gaussian stretch distribution while the old
@@
 
-    \item Given 1. we conclude that the SNe~Ia is indeed drifting, as already
-        suggested by \citep[e.g.][]{howell2007}.
+    \item Given 1. we conclude that the SNe~Ia \nn{stretch} is indeed drifting, as
+        already suggested by \citep[e.g.][]{howell2007}.
 
@@
         data as our drifting model ($p=4\times10^{-8}$). Hence, the BBC
-        sample-based approach does not accurately account for redshift drift as
-        suggested by \cite{scolnic2016} and \cite{scolnic2018a}.
+        sample-based approach does not accurately account for redshift drift \nn{contrary
+        to the suggestion of} \cite{scolnic2016} and \cite{scolnic2018a}.
 
     \item An inaccurate underlying stretch distribution modeling is estimated to
-        bias the mean standardized SNe~Ia on the order of a few percent, which
+        bias the mean standardized SNe~Ia \nn{to} the order of a few percents, which
         is comparable in scale to the observational signature of exotic forms of
@@
             \label{eqconclusion:stretchz}
-        \Delta\left(x_1\,|\,z \right) =
-        &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1)\,+\nonumber\\
-        (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1) +
-        (1-a)\times\mathcal{N}(\mu_2,\sigma_2)\right]\nonumber
+            \nn{X_1\left(z \right) =
+            } &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
+            (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
+            (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
         \end{align}
@@
 
+\bibitem[Kim et al.(2018)]{kim18} Kim, Y.-L., Smith, M., Sullivan, M., et al.\ 2018, \apj, 854, 24
+
+\bibitem[Kim et al.(2019)]{kim19} Kim, Y.-L., Kang, Y., \& Lee, Y.-W.\ 2019, Journal of Korean Astronomical Society, 52, 181
+
 \bibitem[Knox \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv e-prints, arXiv:1908.03663
@@
 \bibitem[Scolnic et al.(2019)]{scolnicastro2020} Scolnic, D., Perlmutter, S., Aldering, G., et al.\ 2019, Astro2020: Decadal Survey on Astronomy and Astrophysics, 2020, 270
+
+\bibitem[Strolger et al.(2004)]{strolger04} Strolger, L.-G., Riess, A.~G., Dahlen, T., et al.\ 2004, \apj, 613, 200
 

### 2020-02-17 Nicolas.tex
Started YC changes

@@
 \usepackage{txfonts}
-\usepackage{amsmath}
+\usepackage{amsmath, mathtools}
 \usepackage[colorlinks,breaklinks]{hyperref}
+\usepackage[normalem]{ulem}
+\usepackage{cancel}
 \hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}
@@
 \newcommand{\nn}[1]{{\textcolor[rgb]{1, 0.27, 0}{#1}}}
+\newcommand{\yc}[1]{{\textcolor{blue}{#1}}}
 
@@
 age-sample can be modeled and do not significantly drift. This paper aims at
-testing this with data from the literature.
+testing this with datasets from the literature.
 
@@
 
-The concept of the SNe~Ia age dichotomy arised when the SN~Ia rate has first
+The concept of the SNe~Ia age dichotomy aroused when the SN~Ia rate has first
 been studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008}
@@
 and one old following the host stellar mass (the so called “prompt and delayed“
-or “A+B“ model). In \cite{rigault2018} we used the specific star formation rate
+or “A+B“ model). In \cite{rigault2018} we used the specific Star Formation Rate
 at the SN location (Local sSFR or LsSFR) to classify which are the prompts
@@
 redshift drift of the SN stretch distribution. In this paper we revisit this
-analysis with the most recent SNe~Ia dataset..
+analysis with the most recent SNe~Ia dataset.
 
@@
 
-\section{A ''complete'' Sample}
+\section{Complete Sample Construction}
 \label{sec:sample}
@@
 with the lightcurve stretch (and color), the first SN~Ia that a
-magnitude-limited survey will miss are the reddest and the lowest-stretch ones.
+magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
 Consequently, if selection effects are not accounted for, one might confuse true
@@
 
-\nn{Assuming} sufficient (and random) spectroscopic follow-up for acquiring
-typing and host redshift, the selection effects of magnitude-limited survey
+\nn{Assuming} sufficient (and random \yc{unbiased?}) spectroscopic follow-up for acquiring
+typing and host redshift, the selection effects of magnitude-limited surveys
 should be negligible bellow a given redshift where even the faintest \nn{SN~Ia}
-could be observed. In contrast, targeted surveys have highly complex selection
+can be observed. In contrast, targeted surveys have highly complex selection
 functions and will be discarded from our analysis. Fortunately, modern SN
-cosmology samples such as the pantheon dataset are now dominated by magnitude
+cosmology samples such as the Pantheon dataset are now dominated by magnitude
 limited surveys.
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
-    \caption{\textsc{\texttt{SALT2.4}} stretch and color lightcurve parameters
-        of SNe~Ia from the SDSS, PS1 and SNLS samples from the pantheon catalog.
-        The individual SN data are shown as blue dots and a 2D histogram is
+    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$) lightcurve parameters
+        of SNe~Ia from the SDSS, PS1 and SNLS samples from the Pantheon catalog.
+        The individual SNe are shown as blue dots and a 2D histogram is
         shown in gray to highlight point density. The ellipse
-        ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$) is displayed, colored by the
+        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the
         corresponding standardized absolute magnitude using the $\alpha$ and
         $\beta$ coefficients from \cite{scolnic2018a}. \nn{The lines represent
-        the $(x_1, c)$ evolution for $m = m_{lim}$ in eq.~\ref{eq:m_lim},
-    for $z$ between $0.60$ and $z=1.00$ using SNLS's $m_{lim}$ of $24.8\,$mag. The data
-    above these lines are those for which $m > m_{lim}$, meaning they can't be
+        the $(x_1, c)$ evolution for $m = m_{lim}$ in Eq.~\ref{eq:m_lim},
+    for $z$ between $0.60$ and $z=1.00$ using SNLS's $m_{lim}$ of $24.8$~mag. The SNe
+    above these lines are those for which $m > m_{lim}$, meaning they cannot be
     observed.}}
@@
 We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
-distributions of SNe~Ia from these surveys, namely: PanStarrs
+distributions of SNe~Ia from \nn{the following surveys}: PanStarrs
 \citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
@@
 \citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
-c)$ plan with semi-major axis $x_1=3$ and semi-minor axis $c=0.3$ encapsulates
+c)$ plan with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates
 the full distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011}
 and \citet{campbell2013} for similar contours, the second using a more
-conservative $c=0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and
-$c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute
-standardized magnitude at maximum of light along the aforementioned ellipse
+conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and
+$c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
+standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$ along the aforementioned ellipse
 given the standardization coefficient $\alpha=0.156$ and $\beta=3.14$ from
-\cite{scolnic2018a}. The faintest SN~Ia is that with $x_1=-1.66$ and $c=0.25$
-and has an absolute standardized magnitude at peak in Bessel-B band of
-$-18.31\,\mathrm{mag}$. Since one need to detect this object typically a week
+\cite{scolnic2018a}: the faintest SN~Ia is that with $(x_1=-1.66, c=0.25)$
+and an absolute standardized magnitude at peak in Bessel-B band of
+$M^{t_0}_{min} = -18.31$~mag. Since one need to detect this object typically a week
 before and 10 days after peak to build a suitable lightcurve, the limiting
 \nn{standardized} absolute magnitude effectively is approximately
-$-18.00\,\mathrm{mag}$. Hence, given the $5\sigma$ point source detection
-magnitude limit of a magnitude limited survey $m_{lim}$, one can derive the
-maximum redshift above which the faintest SNe~Ia will be missed \nn{using the
+$M_{min} = -18.00$~mag. Hence, given the \nn{\sout{$5\sigma$ point source detection}}
+magnitude limit $m_{lim}$ of a magnitude limited survey, one can derive the
+maximum redshift $z_{lim}$ above which the faintest SNe~Ia will be missed \nn{using the
 relation between apparent magnitude, redshift and absolute magnitude given by:}
 \begin{equation}\label{eq:m_lim}
-    \nn{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\,\text{mag}}
-    < m_{lim}}
+    \nn{\cancel{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\;\text{mag}}
+    < m_{lim}}}
+\end{equation}
+\begin{equation}\label{eq:m_lim}
+    \yc{\mu(z_{lim}) = m_{lim} - M_{min}} \quad\text{min vs. lim???}
 \end{equation}
@@
 redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
-filter that has a 24.8 mag $5\sigma$
-depth\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
-final release website.}}. This converts to a $z_{lim}=0.60$, in perfect
-agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.
+filter that has a $5\sigma$ depth of 
+24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT final release website.}}. 
+This converts to a $z_{lim}=0.60$, in perfect agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. \nn{Fig.~14 of \citealt{perrett2010}
+suggests a lower limit of $z_{lim}=0.55$; both will be considered as discussed below.}
+
 Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
 $5\sigma$ depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
-agreement with, e.g., Figure~6 of \cite{scolnic2018a}. This figure could also
-suggest a more conservative $z_{lim}$ of 0.27; both will be considered as
-discussed below.  In similar redshift range, SDSS has a limiting magnitude of
+agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure could also
+suggest a more conservative $z_{lim}$ of 0.27.
+
+In similar redshift range, SDSS has a limiting magnitude of
 22.5 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However,
@@
 \cite{kessler2009} \nn{pointed out} that during year-1 of SDSS, SNe~Ia with
-$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to the
+$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to a
 redshift cut at $0.15$. For the rest of the SDSS survey, additional
-spectroscopic resources were used, such that \cite{kessler2009} and
+spectroscopic resources were used, and \cite{kessler2009} and
 \cite{dilday2008} show a relative completeness up to $z_{lim}=0.2$. Following
-these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit for
-the rest of \nn{this study}. For the rest of the analysis, we will consider both
-cases, the \nn{aforementioned} redshift limits as well as more conservative cuts
+these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit \sout{for
+the rest of \nn{this study}}.
+
+For the rest of the analysis, we will consider \nn{\sout{both
+cases,}} the \nn{aforementioned} redshift limits as well as more conservative cuts
 for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for
-PS1 and $z_{lim}=0.55$ for SNLS (following Fig.~14 of \citealt{perrett2010}).
-
-The complete sample selection is summarized table~\ref{tab:sample} and the
+PS1 and $z_{lim}=0.55$ for SNLS.
+
+The complete sample selection is summarized in Table~\ref{tab:sample}, and the
 redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
-We see in this figure that the redshift limits we have selected roughly
-correspond to the peak of these histograms\nn{. As the peaks indicate that we
+We see there that the redshift limits we have selected roughly
+correspond to the peak of these histograms. \nn{As the peaks indicate that we
 are starting to miss the faintest SNe~Ia, it seems logical to have the redshift
@@
     \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
-    SNLS from the pantheon dataset. The colored part of the
+    SNLS from the Pantheon dataset. The colored part of the
 distribution shows which SNe~Ia has been kept in our analysis for they are
 supposedly free from selection bias (see section~\ref{sec:sample}). The
-full(light) color responds to the conservative(nominal) cut.}
+dark (resp. light) color responds to the conservative (resp. fiducial) selection cut.}
     \label{fig:cuts}
@@
 been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
-4.2.2). \nn{Their spectroscopic follow-up were only done on a redshift range of
+4.2.2). \nn{Their spectroscopic follow-up were done on a redshift range of
 $0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
 deeper. As such, these SNe~Ia} should also be a random sampling of the underlying
-SN population. The SNf data is particularly useful for studying SN property
-drift as it enable us to have a large SN~Ia sample at $z<0.1$.  The HST sample
-from Pantheon follows the same logic of having a search deeper than the
+SN population. The SNf sample is particularly useful for studying SN property
+drift as it enables us to have a large SN~Ia sample at $z<0.1$.
+
+\nn{Finally, we included t}he HST sample
+from Pantheon\nn{, that} follows the same logic of having a search deeper than the
 follow-up and we therefore kept it entirely \citep{strolger04}, see
-table~\ref{tab:sample}. \nn{We present the stretch distribution, redshift histogram
-and stretch cdf of these five surveys Fig.~\ref{fig:sample}}.
+Table~\ref{tab:sample}. \nn{We present the stretch distribution and redshift histogram of these five surveys in Fig.~\ref{fig:sample}}.
 
@@
     \centering
-    \caption{\nn{Source surveys and number of SNe~Ia used in our ''complete''
+    \caption{\nn{Source surveys and number of SNe~Ia used in our ``complete''
     sample. Conservative cuts are indicated in brackets.}}
@@
         HST & -- & 26\\[0.30em]
+        \hline\\[-0.8em]
         Total & -- & 569(422)\\[0.30em]
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75.pdf}
-    \caption{\nn{\textit{Main}: Stretches of each survey in full and empty markers for the
-    conservative and nominal cuts, respectively, as a function of redshift. \textit{Right}:
-    normed CDF of each survey's stretch distribution. \textit{Top}: combined redshift 
-    histogram of all the data, in dark gray for the conservative cuts;
-    in light gray for the nominal cuts.}}
+    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb.pdf}
+    \caption{\nn{\textit{Main}: SN stretches in each survey as a function of redshift, using solid (resp. empty) markers for the
+    conservative (resp. fiducial) selection. \textit{Top}: combined redshift 
+    histogram for all surveys, in dark (resp. light) gray for the conservative (resp. fiducial) selection cuts.}}
     \label{fig:sample}
@@
 %estimated to be 0.15 and 0.65, respectively."}
-% ADDIND: PS cut from pantheon close to 0.27 see Fig. 6 of Scolnic 2018
+% ADDIND: PS cut from Pantheon close to 0.27 see Fig. 6 of Scolnic 2018
 
@@
 the number of Gyr-old stars in the Universe, i.e. the stellar mass \nn{(M)}.
-Hence, if we denote $\delta(z)$ (resp. $\psi(z)$) the fraction of young (resp.
+Hence, if we denote $\delta(z)$ (resp. $\psi(z) = 1 - \delta(z)$) the fraction of young (resp.
 old) SNe~Ia in the Universe as a function of redshift, then their ratio
@@
 rate \nn{(SFR/M)} which goes as $(1+z)^{2.8}$ until $z\sim2$
-\citep[e.g.,][]{tasca2015}. Hence since \nn{$\delta(0.05) \sim \psi(0.05)$}
+\citep[e.g.,][]{tasca2015}. Since \nn{$\delta(0.05) \sim \psi(0.05)$}
 \citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
-\citep{mannucci2006,rodney2014}, \nn{and $\delta + \psi = 1$,} then we find in
-\cite{rigault2018}:
-\begin{align}
+\citep{mannucci2006,rodney2014}, we conclude in
+\cite{rigault2018} that
+%\begin{align}
+%    \label{eq:delta}
+%    \delta(z) &= \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
+%    \mathrm{and}\\
+%    \psi(z) &= \left( K \times (1+z)^{+2.8} +1 \right)^{-1}
+%\end{align}
+\begin{equation}
     \label{eq:delta}
-    \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
-    \mathrm{and}\\
-    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1}
-\end{align}
-where $K=0.87$. This modelisation is comparable to the evolution predicted by
+    \yc{\delta(z) = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}}
+\end{equation}
+with $K=0.87$. This model is comparable to the evolution predicted by
 \cite{childress2014} based on SN rates in galaxies depending on their quenching
@@
     \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$) as a function of
-        the local specific star formation rate (LsSFR) for SNf data used in this
-        analysis. The color corresponds to the probability for the SNe~Ia to be
-        above $\log(\mathrm{LsSFR}=-10.82)$, which defines $p_y$; see
+        the local specific star formation rate (LsSFR) for SNf SNe used in this
+        analysis. The color corresponds to the probability $p_y$ for the SNe~Ia to be
+        young, i.e. to have $\log\mathrm{LsSFR} \geq -10.82$; see
         \cite{rigault2018}. \textit{Right}: histogram of $p_y$-weighted
-        marginalization of the SN stretch. On both panel, we illustrate the
-        Base model fitted on the SNf data. The young and old population
+        distribution of the SN stretchs. On both panel, we illustrate the
+        Base model fitted on the SNf SNe. The young and old population
         contributions are shown in purple and yellow, respectively.}
@@
 
-\nn{In~\cite{rigault2018} we presented the relation between SNe stretches and
-LsSFR measurements, which trace progenitor age, from the SNf sample, and is
-shown in Fig.~\ref{fig:stretchlssfr}}. To model the evolution of the SN stretch
+To model the evolution of the SN stretch
 as a function of redshift, given our aforementioned model of the evolution of
 the fraction of young and old SNe~Ia with redshift, we need to model the SN
-stretch distribution for each age subsample. Given the structure of the
-stretch--LsSFR scatter plot shown in Fig.~\ref{fig:stretchlssfr}, we define our
-model as follows: for the young population the underlying stretch distribution
-is modeled as $\mathcal{N}(\mu_1, \sigma_1^2)$, i.e. a normal distribution
-centered on $\mu_1$ with a $\sigma_1$ width; the old population stretch
-distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
-\mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a gaussian mixture where one mode is the
-same as the young population one. \nn{When combining our prediction of the
-evolution of young SNe~Ia as a function of redshift (eq.~\ref{eq:delta}),
+stretch distribution for each age subsample.
+
+\nn{In~\cite{rigault2018}, we presented the relation between SNe stretches and
+LsSFR measurements, which trace progenitor age, from the SNf sample, and is
+shown in Fig.~\ref{fig:stretchlssfr}}. Given the structure of the
+stretch-LsSFR scatter plot, our model is defined as follows: 
+\begin{itemize}
+    \item for the young population the underlying stretch distribution
+    is modeled as a normal distribution $\mathcal{N}(\mu_1, \sigma_1^2)$; 
+    \item the old population stretch distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
+    \mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a Gaussian mixture where one mode is the
+    same as the young population one. \yc{explain meaning of $a$}
+\end{itemize}
+\nn{When combining our prediction of the
+evolution of young SNe~Ia as a function of redshift (Eq.~\ref{eq:delta}),
 and the stretch distribution of both young and old SNe~Ia, our model for the
-underlying distribution of stretch $X_1\left(z\right)$ as a function of
+underlying distribution of stretch $X_1(z)$ as a function of
 redshift is:}
@@
     \label{eq:stretchz}
-    \nn{X_1\left(z \right) =
-    } &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
-    (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
-    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
+    \nn{X_1(z) = {}}
+    \delta(z) &\times \mathcal{N}(\mu_1,\sigma_1^2) + {}\nonumber\\
+    (1 - \delta(z)) &\times \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
+    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right].
 \end{align}
 
-The estimation of the 5 free parameters
-($\theta\equiv{\mu_1,\mu_2,\sigma_1,\sigma_2,a}$) of the model is given by
-minimizing a pseudo-$\chi^2$ $\mathcal{L}$ such that:
-
+The \yc{maximum-likelihood estimate} of the 5 free parameters
+$\theta\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained by
+minimizing a pseudo-$\chi^2$ $\mathcal{L}$ defined as:
 \begin{equation}
     \label{eq:likelihood}
-    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
-    \mathrm{d}x^{i}_{1}, \nn{y^i}}\right),
+    \mathcal{L} = -2 \sum_i \ln \prob{x^{i}_{1}}{\theta;
+    \mathrm{d}x^{i}_{1}, \nn{y^i}},
 \end{equation}
-\nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
-    $y^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
-    \textsc{\texttt{SALT2.4}} stretch and the probability that the SN is young,
-    respectively. Depending on the data available, there are two ways to fit
-these 5 free parameters}.
-
-\subsubsection{\nn{Parameters estimation with LsSFR data}}
-\label{sec:modelpy}
-
-\nn{When using SNf data, we choose $y^i$ to be their given $p_y{}^i$, computed
-from their LsSFR data (see Fig.~\ref{fig:stretchlssfr}). \nn{It represents the
-probability to be above $\log(\mathrm{LsSFR} = -10.82)$.} Thus, we would have
-$\mathcal{P}$ from eq.~\ref{eq:likelihood} to be}
-
+with
 \begin{align}
     \label{eq:likelihoodsnf}
-    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}} =
-    \nn{p_y{}^i} \times &\mathcal{N}\left(\mu_1,
-    \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
-    \nonumber\\
-        (1-\nn{p_y{}^i}) \times \nn{\bigg[} a \times &
-        \mathcal{N}\left(\mu_1,
-        \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
-        \nonumber\\
-     (1-a) \times &\mathcal{N}\left(\mu_2,
- \sqrt{\sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) \nn{\bigg]}
+    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, y^i} =
+    y^i &\times \mathcal{N}\left(x^{i}_{1} \mid \mu_1, \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
+        (1-y^i) &\times \nn{\bigg[} 
+        a \times \mathcal{N}\left(x^{i}_{1} \mid \mu_1,
+        \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
+     &(1-a) \times \mathcal{N}\left(x^{i}_{1} \mid \mu_2,
+     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) \nn{\bigg]}
 \end{align}
-
-In practice, since ''$a$'' is defined between 0 and 1, we fit for $\alpha$ such
+where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
+$y^i$ are the \textsc{\texttt{SALT2.4}} stretch, its associated error 
+and the probability that the SN is young, respectively. 
+In practice, to ensure fraction $a$ is constrained between 0 and 1,
+we fit for $\alpha$ such
 that $a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.
-\mr{Results on fitting the SNf data with this model are shown
-table~\ref{tab:modelresults} and illustrated in Fig.~\ref{fig:stretchlssfr}.}
+
+Depending on \yc{whether $y^i$ can be 
+estimated directly from LsSFR measurements or not}, there are two ways to proceed.
+
+\subsubsection{\yc{With LsSFR measurements}}
+\label{sec:modelpy}
+
+\yc{For the SNfactory sample, we can readily set $y^i = p_y{}^i$, the
+probability to have $\log \mathrm{LsSFR} \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}),
+to minimize Eq.~\ref{eq:likelihood} with respect to~$\theta$.}
+\mr{Results on fitting the SNf SNe with this model are shown
+Table~\ref{tab:modelresults} and illustrated in Fig.~\ref{fig:stretchlssfr}.}
 
@@
     \caption{Best fit values of the parameters for the Base stretch distribution
-    model when applied to the SNf dataset only (114 SNe~Ia), the nominal 569
+    model when applied to the SNf dataset only (114 SNe~Ia), the reference 569
 SN~Ia sample or the conservative one (422).}
@@
         SNf & $0.41 \pm 0.08$  & $0.55 \pm 0.06$ & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$ & $0.48^{+0.08}_{-0.08}$ \\[0.15em]
-        All & $0.37 \pm 0.05$  & $0.61 \pm 0.04$ & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$ & $0.51^{+0.09}_{-0.10}$ \\[0.15em]
-        All(cons) & $0.38 \pm 0.05$  & $0.60 \pm 0.04$ & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\[0.15em]
+        \yc{Reference} & $0.37 \pm 0.05$  & $0.61 \pm 0.04$ & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$ & $0.51^{+0.09}_{-0.10}$ \\[0.15em]
+        \yc{Conservative} & $0.38 \pm 0.05$  & $0.60 \pm 0.04$ & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\[0.15em]
         \hline
@@
 
-\subsubsection{Without LsSFR data}
+\subsubsection{Without LsSFR measurements}
 \label{sec:modelnopy}
 
-Given eq.~\ref{eq:stretchz}, we can extend the analysis to non-SNfactory sample
-by fitting the free parameters of the model ($\theta\equiv{\mu_1, \mu_2,
-\sigma_1, \sigma_2, a}$) assuming the evolution of the fraction of young SNe~Ia
-as a function of redshift given by $\delta(z)$ and given, this time, $x_1^{i}$,
-$\mathrm{d}x^{i}_{1}$ and $z^{i}$, the stretch, error on the stretch and
-redshift of any given SN $i$. In practice, it means that we are still minimizing
-for the same eq.~\ref{eq:likelihood}, but \nn{using $y^i = \delta(z^i)$, which
-is equivalent to replacing $p_y{}^i$ with $\delta(z^i)$ in
-eq.~\ref{eq:likelihoodsnf}.}
-
-\mr{For the rest of the analysis, we will therefore fit for
-eq.~\ref{eq:likelihood} using \nn{$p_y{}^i$} --~the
+\yc{In lack of direct LsSFR measurements (i.e. $p_y^i$), we can extend the analysis to 
+non-SNfactory samples by using the redshift-evolution of the fraction $\delta(z)$ 
+of young SNe~Ia (Eq.~\ref{eq:delta}) as a proxy for the probability of a 
+SN to be young.} This still corresponds to minimizing Eq.~\ref{eq:likelihood} 
+with respect to the parameters 
+$\theta\equiv(\mu_1, \mu_2, \sigma_1, \sigma_2, a)$ 
+of the stretch distribution $X_1$ (Eq.~\ref{eq:stretchz}), 
+but this time assuming $y^i = \delta(z^i)$ for any given SN~$i$.
+
+\mr{For the rest of the analysis, we will therefore minimize
+Eq.~\ref{eq:likelihood} using \nn{$p_y{}^i$} --~the
 probability for the SN $i$ to be young~-- when available (i.e. for SNf
@@
 
-\mr{Results on fitting all the 569(422) SNe~Ia data (with conservative cuts)
-with this model are given table~\ref{tab:modelresults} and illustrated in
+\mr{Results on fitting this model on all the 569 (resp. 422) SNe from the 
+fiducial (resp. conservative) sample
+are given Table~\ref{tab:modelresults} and illustrated in
 Fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
-stretch per bin of redshift follows our redshift drift modeling. That is, when
+stretch per bin of redshift closely follows our redshift drift modeling; that is, when
 considering selection-bias free SN~Ia samples, SNe~Ia at higher redshift have on
@@
 redshift ($-0.17\pm 0.10$ at $z\sim0.05$). This is indeed what is expected if
-old environments favor low SN stretch \citep[e.g.]{howell2007} \textit{and} if
+old environments favor low SN stretches \citep[e.g.][]{howell2007} \emph{and} if
 the fraction of old SNe~Ia reduces as a function of redshift. See
@@
     of each redshift bins. Full and light markers are used when considering
-        the nominal or the conservative sample, respectively, and are colored by
-        their mean redshift. The orange horizontal line represents the mean
-        redshift of the nominal sample illustrating the expectation if the SN
-        stretch distribution is not drifting. Best fits of our Base drifting
+    the fiducial or the conservative samples, respectively. The orange horizontal line represents the mean
+    redshift of the reference sample illustrating the expectation if the SN
+    stretch distribution is not evolving with redshift. Best fits of our Base drifting
     model are shown as blue, dashed-blue and gray, when fitted on the
-    nominal, the conservative or the SNf dataset, respectively; all are similar.
+    fiducial sample, the conservative one or the SNf dataset only, respectively; all are \yc{compatible}.
 The light-blue illustrates the amplitude of the error of the best fit model when
-considering the nominal dataset.}
+considering the reference dataset.}
     \label{fig:modelall}
@@
 
-\subsection{Other modelings}
+\subsection{Alternative models}
 \label{sec:othermodel}
 
-In section~\nn{\ref{sec:basemodel}} we have modeled the underlying stretch
-distribution as a single gaussian for the young SNe Ia and a combination of two
-gaussians for the old SNe Ia population, i.e., the same as the young-population
+In section~\nn{\ref{sec:basemodel}}, we have modeled the underlying stretch
+distribution following \cite{rigault2018}, i.e. as a single Gaussian for the young SNe~Ia and a mixture of two
+Gaussians for the old SNe Ia population, the same as the young-population
 one plus another for the fast declining SNe~Ia that seem to only exist in old
-populations. This is our “Base” model.
-
-\cite{howell2007} used a simpler modelisation: a single gaussian for each age
-sub-group. To test \nn{different modeling choices}, we have implemented a suite
-of extra parametrisation that we also fit to the data following the procedure
+populations. This is our so-called ``Base'' model.
+
+However, to test \nn{different modeling choices}, we have implemented a suite
+of alternative parametrizations that we also adjust to the data following the procedure
 described section~\ref{sec:modelnopy}.
@@
 
-\nn{We thus consider a model ''Howell+drift'', with one single gaussian per age
-group as in \citet{howell2007}}. In addition, since we aim at probing the SN
-redshift drift, we are also fitting the ''Base'' model and \nn{this
-aforementioned extension} by fixing the population drift parameter $\delta(z)$
-to a free parameter constant $\delta(z)=f$.
-
-We are also considering alternative non-drifting models, and notably the one
+\cite{howell2007} used a simpler model with a normal distribution for each young 
+and old populations. \nn{We thus consider a ``Howell+drift'' model, with one 
+single Gaussian per age group}. \yc{one has to explain the ``drift'' part!!!}
+
+\yc{Alternatively, since we aim at probing the existence of a redshift drift, 
+we are also testing non-drifting models by restricting the ``Base'' and ``Howell'' models to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe; these models are labeled ``Base constant'' and ``Howell constant''.} \yc{fix labels!!!}
+
+We are also considering other non-drifting models, and notably the one
 developed for the Beam with Bias Correction \cite[BBC,][]{scolnic2016,
-kessler2017}, currently used by all recent SN Cosmological analyses
+kessler2017}, currently used in recent SN cosmological analyses
 \cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to account
 for Malmquist biases. The BBC formalism assumes sample-based (hence
-non-drifting) asymmetric Gaussian distributions ($\mathcal{N}\left(\mu,
-\sigma^{-}{}^2\, \mathrm{if}\,x_1<\mu\,\mathrm{else}\, \sigma^{+}\right)$). The
-idea behind the sample-based approach is twofold: (1) Malmquist biases are
+intrinsically non-drifting) asymmetric Gaussian stretch distributions: $\mathcal{N}\left(\mu,
+\sigma_{-}^2\;\text{if}\;x_1<\mu\;\text{else}\;\sigma_{+}^2\right)$. The
+idea behind this sample-based approach is twofold: (1) Malmquist biases are
 driven by survey properties and (2) because current surveys cover limited
-redshift ranges, doing so contains some potential redshift evolution information
-\citep{scolnic2016, scolnic2018a}. See further discussion concerning the BBC
-modeling in section~\ref{sec:bbc}. For the sake of the comparison, we are also
-running Gaussian and asymmetric Gaussian non-drifting models.
-
-The ability of these models to explain the data is detailed in
-section~\ref{sec:results}.
+redshift ranges, doing so covers some potential redshift evolution information
+\citep{scolnic2016, scolnic2018a}. See further discussion concerning the BBC in section~\ref{sec:bbc}.
+
+Finally, for the sake of completeness, we also
+consider simple ``Gaussian'' %$\mathcal{N}(\mu, \sigma)$ 
+and ``asymmetric'' Gaussian %$\mathcal{N}(\mu, \sigma_{\pm})$ 
+non-drifting models.
+
+% The ability of these models to explain the observations is detailed in
+% section~\ref{sec:results}.
 
@@
 \label{sec:results}
-We applied each model on both the conservative and non-conservative samples (cf.
-section \ref{sec:sample}). Because the models presented
-section~\ref{sec:modeling} have various degrees of freedom, we use the Akaike
-Information Criterion (AIC) \citep{burnham2004} to compare them. This estimator
+
+We applied each model on both the reference and conservative samples (cf.
+section~\ref{sec:sample}). Because the models presented
+in section~\ref{sec:modeling} have different degrees of freedom, we use the Akaike
+Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare them. This estimator
 penalizes extra degrees of freedom to avoid over-fitting the data. It is defined
@@
 \begin{equation}
-    \nn{\mathrm{AIC} = 2k - 2\ln L}
+    \mathrm{AIC} = \yc{\mathcal{L}_{min}} + 2k
 \end{equation}
-where $k$ is the number of free parameters and \nn{$-2\ln L = \mathcal{L}$} is
-the pseudo-$\chi^2$ defined eq.~\eqref{eq:likelihood}. The \nn{reference} model
-is the one with the smaller AIC and the probability for another model to be at
+where \yc{$\mathcal{L}_{min}$} is
+the minimum value of the pseudo-$\chi^2$ as defined Eq.~\eqref{eq:likelihood} and $k$ is the number of free parameters. The \nn{reference} model
+is the one with the smallest AIC, and the probability for another model to be at
 least as representative of the data as this one is given by:
@@
 \end{equation}
-Results are presented in table \ref{tab:comp} and are illustrated
+\yc{with $\Delta\mathrm{AIC} < 0$}. Results are presented in Table~\ref{tab:comp} and are illustrated
 Fig.~\ref{fig:mod_comp}.
@@
         section~\ref{sec:modeling}. The drift column indicates if the model
-        contains the age-drifting model ($\delta(z)$) or not (f, i.e.
-        non-drifting). The BBC modeling is sample based which somewhat contains
+        contains the age-drifting model ($\delta(z)$) or not (constant $f$, i.e.
+        non-drifting). The BBC is sample based which somewhat includes
         redshift dependency information. $\mathcal{L}$ is the pseudo-$\chi^{2}$
-        defined eq.~\ref{eq:likelihood} and Proba is the probability that a
+        defined Eq.~\ref{eq:likelihood} and Proba is the probability that a
         model is at least as good as the reference model to describe the data.
@@
     \begin{tabular}{c c c  |c c c c|  c c c c}\hline\hline\\[-0.8em]
-        & & & \multicolumn{4}{|c}{All SNe~Ia (569)} & \multicolumn{4}{|c}{All SNe~Ia (conservative; 422)} \\
+        & & & \multicolumn{4}{|c}{\yc{Fiducial sample (569 SNe)}} & \multicolumn{4}{|c}{\yc{Conservative sample (422 SNe)}} \\
         %\cline{2-6}
-        Name & drift & Free param &
+
+        Name & drift & $k$ &
         $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba & $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba \\[0.15em]
@@
         Howell & $\delta(z)$ & 4
-        & 1463.3 & 1471.3 & -4.6 & $1.0\times10^{-1}$
-        & 1088.2 & 1096.2 & -6.7 & $3.4\times10^{-2}$\\[0.15em]
+        & 1463.3 & 1471.3 & $-4.6$ & $1.0\times10^{-1}$
+        & 1088.2 & 1096.2 & $-6.7$ & $3.4\times10^{-2}$\\[0.15em]
 
         Asymmetric & -- & 3
-        & 1485.2 & 1491.2 & -24.5 & $4.7\times10^{-6}$
-        & 1101.3 & 1107.3 & -17.8 & $1.4\times10^{-4}$\\[0.15em]
+        & 1485.2 & 1491.2 & $-24.5$ & $4.7\times10^{-6}$
+        & 1101.3 & 1107.3 & $-17.8$ & $1.4\times10^{-4}$\\[0.15em]
 
         Howell & $f$ & 5
-        & 1484.2 & 1494.2 & -27.5 & $1.0\times10^{-6}$
-        & 1101.2 & 1111.2 & -21.7 & $1.9\times10^{-5}$\\[0.15em]
+        & 1484.2 & 1494.2 & $-27.5$ & $1.0\times10^{-6}$
+        & 1101.2 & 1111.2 & $-21.7$ & $1.9\times10^{-5}$\\[0.15em]
 
         Base & $f$ & 6
-        & 1484.2 & 1496.2 & -29.5 & $3.9\times10^{-7}$
-        & 1101.2 & 1113.2 & -23.7 & $7.1\times10^{-6}$\\[0.15em]
-
-        BBC-modeling & per sample & 3x5
-        & 1468.2 & 1498.2 & -31.5 & $1.5\times10^{-7}$
-        & 1083.6 & 1113.6 & -24.1 & $5.7\times10^{-6}$\\[0.15em]
+        & 1484.2 & 1496.2 & $-29.5$ & $3.9\times10^{-7}$
+        & 1101.2 & 1113.2 & $-23.7$ & $7.1\times10^{-6}$\\[0.15em]
+
+        BBC & per sample & 3x5
+        & 1468.2 & 1498.2 & $-31.5$ & $1.5\times10^{-7}$
+        & 1083.6 & 1113.6 & $-24.1$ & $5.7\times10^{-6}$\\[0.15em]
 
         Gaussian & -- & 2
-        & 1521.8 & 1525.8 & -59.1 & $1.5\times10^{-13}$
-        & 1142.6 & 1146.6 & -57.1 & $4.0\times10^{-13}$ \\\hline\hline
+        & 1521.8 & 1525.8 & $-59.1$ & $1.5\times10^{-13}$
+        & 1142.6 & 1146.6 & $-57.1$ & $4.0\times10^{-13}$ \\\hline\hline
     \end{tabular}
@@
     \includegraphics[width=\linewidth]{Article_figures/mod_comp.pdf}
-    \caption{$\Delta$AIC between the reference model and the other considered
-    models(see also table~\ref{tab:comp}). The full and open blue markers
+    \caption{$\Delta$AIC between the reference ``Base'' model and the other
+    models (see also Table~\ref{tab:comp}). The full and open blue markers
 correspond to model with and without redshift drift, respectively. Light markers
-show the results when this analysis is done on the conservative sample.
-Color-bands illustrate the validity of the models, from Acceptable to Excluded,
-see inside text. All non-drifting models are excluded.}
+show the results when this analysis is performed on the conservative sample 
+rather than the fiducial one. Color-bands illustrate the validity of the models, from 
+Acceptable ($\Delta\mathrm{AIC} > -5$) to Excluded ($\Delta\mathrm{AIC} < -20$),
+see text \yc{This is not explained in the text!}. \yc{According to the AIC}, all non-drifting models (open symbols) are excluded for the reference sample.}
     \label{fig:mod_comp}
@@
 
-\mr{We find that underlying stretch distributions with no redshift drift are all
-    excluded to describe the data as well as the Base redshift model. In fact,
-    the best non drifting model (Asymmetric) has $2\times10^{-4}\%$ of chance to
-    describe the data as well as Base. This is visible fig~\ref{fig:modelall}
-    where the mean stretches per bin of redshift of equal sample size suggest a
-    redshift evolution following our Base model --~fitted either on SNf, on the
-    nominal 569 SNe~Ia dataset or the conservative one~-- rather than having a
-constant mean value.}
-
-\mr{As just mentioned, the \nn{reference model is Base}, but we also implemented
-    a Base$-(\sigma_2)$ model, where both the low-stretch (mode 2 that only
-    exists in old environments) and high-stretch (mode 1 that exists in all
-    environments) normal distributions share the same width. It is marginally
-    better than the Base model where both distributions have their own
-    \nn{dispersion} ($\Delta \mathrm{AIC}<2$). Similarly, letting the
-    high-stretch distribution to vary between the young and the old population
-    (\nn{model we call} Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$) only
-    marginally improves the fit ($\Delta\mathcal{L}=3.6$ in comparison to Base)
-    at a cost of 2 more degrees of freedom, making this mode less likely. This
-    suggests that, indeed, the high-stretch mode (mode 1) exists in all kind of
-    age-populations. We also remark that considering the stretch distribution of
-    each age-population as independent normal distributions, as suggested by
-\cite{howell2007}, is \nn{significantly disfavored}.}
-
-\mr{Our results strongly suggest that the underlying distribution of SN~Ia
-    stretch evolves as a function of redshift. Our redshift drift modeling
-    presented section~\ref{sec:modeling} is a much better representation of the
-    pantheon data (see section~\ref{sec:sample}) than any non-drifting models,
-    including the sample-based BBC one. Considering or not the conservative
-sample has no effect on our conclusion as illustrated Fig.~\ref{fig:modelall}
-and presented table~\ref{tab:comp}.}
-
+\yc{We first note that the best model (with smallest AIC) is the so-called Base model, 
+implementing \cite{rigault2018} description of the stretch distribution and its 
+redshift evolution. This is therefore the reference model, both on the fiducial and 
+conservative samples.}
+
+\mr{\yc{Furthermore,} we find that underlying stretch distributions not including 
+    a redshift drift are all
+    excluded to describe the data as well as the Base model. In fact,
+    the best non drifting model (Asymmetric) has \yc{a very marginal chance 
+    ($p = 2\times10^{-6}$)} \yc{not same number as in table!!!} to
+    describe the data as well as the Base model.}
+    \yc{This result is just a quantitative assessment of qualitative facts clearly visible in 
+    Fig.~\ref{fig:modelall}: the mean SN stretch per bin of redshift strongly 
+    suggests a redshift evolution rather than having a constant value, and this evolution is compatible with the redshift drift expected from Eq.~\ref{eq:delta}.}
+
+\yc{Are the next 2 paragraphs (now commented out) still pertinent? The model does not appear in table nor 
+fig. Mickael, please rephrase and remove what is not needed anymore.}
+
+% \mr{As just mentioned, the \nn{reference model is Base}, but we also implemented
+%     a Base$-(\sigma_2)$ model, where both the low-stretch (mode 2 that only
+%     exists in old environments) and high-stretch (mode 1 that exists in all
+%     environments) normal distributions share the same width. It is marginally
+%     better than the Base model where both distributions have their own
+%     \nn{dispersion} ($\Delta \mathrm{AIC}<2$). Similarly, letting the
+%     high-stretch distribution to vary between the young and the old population
+%     (\nn{model we call} Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$) only
+%     marginally improves the fit ($\Delta\mathcal{L}=3.6$ in comparison to Base)
+%     at a cost of 2 more degrees of freedom, making this mode less likely. This
+%     suggests that, indeed, the high-stretch mode (mode 1) exists in all kind of
+%     age-populations. We also remark that considering the stretch distribution of
+%     each age-population as independent normal distributions, as suggested by
+% \cite{howell2007}, is \nn{significantly disfavored}.}
+
+% \mr{Our results strongly suggest that the underlying distribution of SN~Ia
+%     stretch evolves as a function of redshift. Our redshift drift modeling
+%     presented section~\ref{sec:modeling} is a much better representation of the
+%     Pantheon data (see section~\ref{sec:sample}) than any non-drifting models,
+%     including the sample-based BBC one. Considering or not the conservative
+% sample has no effect on our conclusion as illustrated Fig.~\ref{fig:modelall}
+% and presented Table~\ref{tab:comp}.}
+
+\yc{This paragraph is not clear: what is the impact, and its importance, on which case?
+Isn't it in contradiction with the conclusions of BBC section, where it is said that proper handling of stretch correction is of paramount importance for accurate cosmology???}
 \mr{To the best of our knowledge, the SN~Ia stretch redshift drift has never
@@
     Malmquist bias correction. Indeed, as long as the Phillip's relation
-    standardization parameter $\alpha$ is not redshift dependent (study behind
+    standardization parameter $\alpha$ is not redshift dependent (a study behind
     the scope of this paper, but see e.g. \citealt{scolnic2018a}), if a sample
@@
 
-\mr{The state-of-the-art tool for doing so is the BBC described in
-    \cite{scolnic2016} and \cite{kessler2017}, which is used in all recent SN
-    Cosmological analyses \citep{jones2018b, scolnic2018a, brout2019,
-    descosmopaper2019} including the direct measurement of H$_0$
-\citep{riess2016,riess2019}. As already discussed, it currently assumes an
-asymmetric stretch distribution per sample. \nn{A more detailed discussion on
-this modeling is presented section \ref{sec:bbc}.}}
-
 \section{\nn{BBC discussion}}
 \label{sec:bbc}
+
+\mr{The state-of-the-art tool for doing so \yc{what?} is the BBC described in
+    \cite{scolnic2016} and \cite{kessler2017}, which is used in recent SN
+    cosmological analyses \citep{jones2018b, scolnic2018a, brout2019,
+    descosmopaper2019} including the direct measurement of H$_0$
+\citep{riess2016,riess2019}.}
+
 \mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
     distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
-    modelisation. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the best of all
-    the non-drift model, but still $\Delta \mathcal{L}=-11.5$ worse than our
-    best model, it is penalized by requiring 15 free parameters, 3 per sample.
+    modelisation \yc{in our analysis}. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the smallest of all
+    non-drift models (but still $\Delta \mathcal{L}=-11.5$ worse than the
+    reference Base model), it is penalized by requiring 15 free parameters (3 for each of the 5 samples of the analysis).
     \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
-    highlighted that, because current surveys span limited redshift ranges, the
-    per sample approach somewhat accounts for unmodeled redshift drifts. We
-    stress here that, as measurements of modern surveys cover larger redshift
-    ranges to reduce calibration systematic  uncertainties, this is becoming
-    less true, notably for PS1, DES and, soon, LSST. Already with the current
+    highlighted that, because \yc{traditional} surveys span limited redshift ranges, the
+    per sample approach can somewhat account for \yc{implicit} redshift drifts. We
+    stress here that, as measurements of modern surveys cover \yc{increasingly} larger redshift
+    ranges to reduce calibration systematic  uncertainties, this approach is becoming
+    less \yc{valid}, notably for PS1, DES and, soon, LSST \yc{redshift ranges? Why not SNLS too?}. Already with the current
     surveys, the BBC sample-based technique is \nn{ruled out} to be an as good
@@
     detailed discussion of the consequence of this results for cosmology follows
-in section~\ref{sec:bbc}}
-
-\mr{We report in table~\ref{tab:bbc} the samples' $\sigma^-$, $\sigma^+$ and
-    $\mu^0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
+in section~\ref{sec:bbc}.}
+
+\mr{We report in Table~\ref{tab:bbc} the samples' $\sigma_{\pm}$ and
+    $\mu_0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
     SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{our} construction
     of \nn{a} selection-bias free sample, see
-    section~\ref{sec:sample}. If we were to use the \cite{scolnic2016} and
+    section~\ref{sec:sample}. \yc{WHY??? The developed BBC for \textit{incomplete} samples!} If we were to use \cite{scolnic2016} and
     \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
     the $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC
-    modeling would go from -32 to -47. \nn{It would seem that our minimization
-    of these parameters is closer to our Base model than theirs}. Furthermore,
+    modeling would go even deeper from $-32$ to $-47$. \nn{It would seem that our minimization
+    of these parameters is closer to our Base model than theirs}. \yc{???} \yc{Last},
     fixing the Base model parameters to those derived using SNf data only (top
-    row table.~\ref{tab:modelresults}) and using the BBC parameters from the
+    row Table~\ref{tab:modelresults}) and using the BBC parameters from the
     aforementioned literature for PS1, SDSS and SNLS, we find
-    $\Delta\mathrm{AIC}=-31.7$. Removing the SNf data reduces this difference to
-    $\Delta\mathrm{AIC}=-5$. Such a reduction is expected for two reasons.
+    $\Delta\mathrm{AIC}=-31.7$, but removing the SNf sample reduces this difference to
+    $\Delta\mathrm{AIC}=-5$ \yc{not clear: the SNf sample is removed or not???}. Such a \yc{dramatic apparent improvement} is expected for two reasons.
     First, it removes the nearby SNe~Ia, and thereby significantly reduces the
@@
     Second, the SNf data are better at discriminating the individual stretch
-    distributions by age since the probability for a SN to be young $p_y{}^i$ is
-    known while other samples rely on $\delta(z^i)$ which is the fraction
-    of young SNe~Ia at the SN redshift, i.e. a probabilistic proxy for
-    $p_y{}^i$. While in any case the Base-model is favored, the relative weight
+    distributions by age since the probability $y^i$ for a SN to be young is
+    known ($y^i = p_y{}^i$) while other samples rely on a probabilistic proxy for
+    $y^i$, namely the fraction $\delta(z^i)$ 
+    of young SNe~Ia at SN redshift. While in any case the Base model is favored, the relative weight
     of the SNf sample highlights the importance of adding non-targeted nearby
@@
     \begin{tabular}{c c c c c c c}\hline\hline\\[-0.8em]
-    Asymmetric & $\sigma^-$ & $\sigma^+$ & $\mu^0$ \\\hline\\[-0.8em]
+    Asymmetric & $\sigma_{-}$ & $\sigma_{+}$ & $\mu_0$ \\\hline\\[-0.8em]
     SNf &  1.34 $\pm$ 0.13 & 0.41 $\pm$ 0.10 & 0.68 $\pm$ 0.15\\[0.15em]
@@
     \caption{Underlying SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
-        distributions for the PS1 sample. The Green line(band) shows our best
+        distributions for the PS1 sample. The Green line (band) shows our best
         fitted asymmetric Gaussian distribution (its error) mimicking the BBC
-        approach. The dashed line shows the noe estimated by
+        approach. The dashed line shows the noe \yc{???} estimated by
     \cite{scolnic2018a}. The blue to red curves display our Base drifting
@@
 
-\mr{We illustrate Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
+\mr{We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
 underlying stretch distribution between the BBC modeling and our Base drifting
 model for the PS1 sample. Our model is bimodal and the relative
-amplitude of each mode depends on the fraction of old SNe~Ia in the sample
---~the highest the fraction of old SNe~Ia, the highest the amplitude of the
-low-stretch mode~-- and is therefore redshift dependent. As expected the two
+amplitude of each mode depends on the redshift-dependent fraction of old SNe~Ia in the sample: the higher the fraction of old SNe~Ia, the higher the amplitude of the
+low-stretch mode. As expected the two
 approaches strongly differ in modeling the negative part of the SN stretch
@@
 of the BBC analysis, a full study using our Base model given
-eq.~\ref{eq:stretchz} in place of the sample-based asymmetric modeling would be
+Eq.~\ref{eq:stretchz} in place of the sample-based asymmetric modeling would be
 necessary to have an exact understanding of the effect this inaccurate modeling
@@
 from \cite{scolnic2018a}. We see in this figure that the amplitude of the
-magnitude bias caused by an accurate modeling of the underlying stretch
+magnitude bias caused by an \yc{inaccurate} modeling of the underlying stretch
 distribution is of the order of few tens of millimag, which is the typical scale
@@
             expected magnitude difference between $\Lambda$CDM and $w$CDM
-            cosmologies ($w=-1\pm 10,\,5\,\mathrm{and}\,1\%$) or, in orange,
+            cosmologies ($w=-1\pm 10,\,5\,\mathrm{and}\,1\%$) or, in \yc{blue},
             more exotic $w_0,w_a$CDM examples. This figure illustrates that the
@@
         modeling of the underlying stretch distribution is comparable to the
-    observational effect of exotic form of dark energy.}
+    observational effect of exotic form of dark energy. \yc{Is this fig. still relevant?}}
     \label{fig:magdrift}
@@
 \label{sec:ccl}
-We have presented a study of the drift of the underlying SNe~Ia stretch
+
+We have presented a study of the drift of the underlying \yc{I still prefer ``intrinsic''} SNe~Ia stretch
 distribution as a function of redshift. We used SNe~Ia from magnitude-limited
-surveys from the pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
-well as HST to which we added SNfactory data from \cite{rigault2018} for the
+surveys from the Pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
+well as HST, to which we added SNfactory data from \cite{rigault2018} for the
 low-redshift bin. We only considered the supernovae that have been discovered
@@
 modeling. We also studied the stretch distribution modeling \nn{of the}
-current version of the BBC algorithm used for Malmquist bias correction by all
+current version of the BBC algorithm used for Malmquist bias correction by 
 recent Dark energy or Hubble constant cosmological analyses.
 
-Our conclusions are the following:
+Our conclusions are the following.
 \begin{enumerate}
@@
 
-    \item Given 1. we conclude that the SNe~Ia \nn{stretch} is indeed drifting, as
-        already suggested by \citep[e.g.][]{howell2007}.
+    \item Given 1., we conclude that the SNe~Ia \nn{stretch} is indeed drifting, as
+        already suggested by e.g.~\cite{howell2007}.
 
@@
     with: $a=0.50$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
-    $\sigma_2=0.56$, see table~\ref{tab:modelresults} and using the
+    $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}) and using the
     age-population drift modelization $\delta(z)$ defined in \cite{rigault2018}
@@
     \begin{align}
-        \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\nonumber
+        \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}.
     \end{align}
@@
 % Y
- Z
+% Z
 \end{thebibliography}
@@
 
-The fitted parameters are showed table \ref{tab:val}.
+The fitted parameters are shown in Table~\ref{tab:val}.
 
@@
 To test the coherence of this model, we also fitted it using all the data from
-the surveys indistinctly. These results are shown figure \ref{fig:model_all}.
+the surveys indistinctly. These results are shown Fig.~\ref{fig:model_all}.
 

### 2020-02-18 Nicolas.tex
Finished YC changes (at pdf/hist and BBC hist près)

@@
 redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
-We see there that the redshift limits we have selected roughly
-correspond to the peak of these histograms. \nn{As the peaks indicate that we
-are starting to miss the faintest SNe~Ia, it seems logical to have the redshift
-cuts fall near said peaks.}
+We see there that the redshift limits we have selected roughly \nn{but logically}
+correspond to the peak of these histograms. \nn{as the peaks indicate that we
+are starting to miss the faintest SNe~Ia}.
 
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55.pdf}
+    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-inferno.pdf}
     \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb.pdf}
+    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb-inferno.pdf}
     \caption{\nn{\textit{Main}: SN stretches in each survey as a function of redshift, using solid (resp. empty) markers for the
@@
 \begin{itemize}
-    \item for the young population the underlying stretch distribution
+    \item for the young \nn{(i.e. $\log \mathrm{LsSFR}>-10.82$)} population, the underlying stretch distribution
     is modeled as a normal distribution $\mathcal{N}(\mu_1, \sigma_1^2)$; 
-    \item the old population stretch distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
+    \item the old \nn{(i.e. $\log \mathrm{LsSFR}<-10.82$)} population stretch distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
     \mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a Gaussian mixture where one mode is the
-    same as the young population one. \yc{explain meaning of $a$}
+    same as the young population one\nn{, $a$ representing the relative importance
+    of the two Gaussians}.
 \end{itemize}
@@
     \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, y^i} =
-    y^i &\times \mathcal{N}\left(x^{i}_{1} \mid \mu_1, \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
+    y^i &\times \yc{\mathcal{N}\left(x^{i}_{1} \mid \mu_1, \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right)} + {} \nonumber\\
         (1-y^i) &\times \nn{\bigg[} 
-        a \times \mathcal{N}\left(x^{i}_{1} \mid \mu_1,
-        \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
-     &(1-a) \times \mathcal{N}\left(x^{i}_{1} \mid \mu_2,
-     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) \nn{\bigg]}
+        a \times \yc{\mathcal{N}\left(x^{i}_{1} \mid \mu_1,
+        \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right)} + {} \nonumber\\
+     &(1-a) \times \yc{\mathcal{N}\left(x^{i}_{1} \mid \mu_2,
+     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right)} \nn{\bigg]}
 \end{align}
@@
 and old populations. \nn{We thus consider a ``Howell+drift'' model, with one 
-single Gaussian per age group}. \yc{one has to explain the ``drift'' part!!!}
+single Gaussian per age group and the $\delta(z)$ drift from \cite{rigault2018}}. \uline{\yc{one has to explain the ``drift'' part!!!} \nn{done}}
 
 \yc{Alternatively, since we aim at probing the existence of a redshift drift, 
-we are also testing non-drifting models by restricting the ``Base'' and ``Howell'' models to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe; these models are labeled ``Base constant'' and ``Howell constant''.} \yc{fix labels!!!}
+we are also testing non-drifting models by restricting the ``Base'' and ``Howell'' models to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe; these models are labeled ``Base constant'' and ``Howell constant''.} \yc{\uline{fix labels!!!
+\nn{done}}}
 
@@
 \end{equation}
-\yc{with $\Delta\mathrm{AIC} < 0$}. Results are presented in Table~\ref{tab:comp} and are illustrated
-Fig.~\ref{fig:mod_comp}.
+\yc{with $\Delta\mathrm{AIC} < 0$}.
+Results are presented in Table~\ref{tab:comp} and are illustrated Fig.~\ref{fig:mod_comp}\nn{, where a $\Delta$AIC of -5 corresponds to a $z$-score of 1.4, -10 to a $z$-score of 2.5, and -20 to a $z$-score of 3.9}.
 
@@
 Acceptable ($\Delta\mathrm{AIC} > -5$) to Excluded ($\Delta\mathrm{AIC} < -20$),
-see text \yc{This is not explained in the text!}. \yc{According to the AIC}, all non-drifting models (open symbols) are excluded for the reference sample.}
+see text \yc{\uline{This is not explained in the text!
+\nn{Now done (conversion to z-score)}}}.
+\yc{According to the AIC}, all non-drifting models (open symbols) are excluded for the reference sample.}
     \label{fig:mod_comp}
@@
     the best non drifting model (Asymmetric) has \yc{a very marginal chance 
-    ($p = 2\times10^{-6}$)} \yc{not same number as in table!!!} to
+    ($p = 5\times10^{-6}$)} to
     describe the data as well as the Base model.}
@@
 
-\yc{Are the next 2 paragraphs (now commented out) still pertinent? The model does not appear in table nor 
-fig. Mickael, please rephrase and remove what is not needed anymore.}
+\textit{\yc{Are the next 2 paragraphs (now commented out) still pertinent? The model does not appear in table nor 
+fig. Mickael, please rephrase and remove what is not needed anymore.}}
 
@@
 
-\yc{This paragraph is not clear: what is the impact, and its importance, on which case?
-Isn't it in contradiction with the conclusions of BBC section, where it is said that proper handling of stretch correction is of paramount importance for accurate cosmology???}
+\textit{\yc{This paragraph is not clear: what is the impact, and its importance, on which case?
+Isn't it in contradiction with the conclusions of BBC section, where it is said that proper handling of stretch correction is of paramount importance for accurate cosmology???}}
 \mr{To the best of our knowledge, the SN~Ia stretch redshift drift has never
@@
 
-\mr{The state-of-the-art tool for doing so \yc{what?} is the BBC described in
+\mr{The state-of-the-art tool for doing so \textit{\yc{what?} \nn{mimic a selection-bias free sample}} is the BBC described in
     \cite{scolnic2016} and \cite{kessler2017}, which is used in recent SN
@@
     ranges to reduce calibration systematic uncertainties, this approach is becoming
-    less \yc{valid}, notably for PS1, DES and, soon, LSST \yc{redshift ranges? Why not SNLS too?}. Already with the current
+    less \yc{valid}, notably for PS1, DES and, soon, LSST \textit{\yc{redshift ranges? Why not SNLS too?} \nn{Mickaël probably meant to cite only 3 of them?}}. Already with the current
     surveys, the BBC sample-based technique is \nn{ruled out} to be an as good
-    representation of the data as our best model ($p=4\times 10^{-8}$). A more
+    representation of the data as our best model ($p=2\times 10^{-7}$). A more
     detailed discussion of the consequence of this results for cosmology follows
@@
     of \nn{a} selection-bias free sample, see
-    section~\ref{sec:sample}. \yc{WHY??? The developed BBC for \textit{incomplete} samples!} If we were to use \cite{scolnic2016} and
+    section~\ref{sec:sample}. \textit{\yc{WHY??? They developed BBC for \textbf{incomplete} samples!} \nn{Yes, but their final underlying distribution
+    is supposed to have been corrected for selection effects, because they were
+    taken into account and added to their incomplete sample. If we ultimately find
+    the same underlying distribution taking data without selection effects,
+    it might mean we're doing something that concurs with what they took into
+    account.}} If we were to use \cite{scolnic2016} and
     \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
@@
     modeling would go even deeper from $-32$ to $-47$. \nn{It would seem that our minimization
-    of these parameters is closer to our Base model than theirs}. \yc{???} \yc{Last},
+    of these parameters is closer to our Base model than theirs}. \textit{\yc{???} \nn{Is the question "what's the use of this information"? If so, good question.}} \yc{Last},
     fixing the Base model parameters to those derived using SNf data only (top
@@
     aforementioned literature for PS1, SDSS and SNLS, we find
-    $\Delta\mathrm{AIC}=-31.7$, but removing the SNf sample reduces this difference to
-    $\Delta\mathrm{AIC}=-5$ \yc{not clear: the SNf sample is removed or not???}. Such a \yc{dramatic apparent improvement} is expected for two reasons.
+    $\Delta\mathrm{AIC}=-31.7$, \nn{and then} removing the SNf sample reduces this difference to
+    $\Delta\mathrm{AIC}=-5$ \textit{\yc{not clear: the SNf sample is removed or not???} \nn{Yes.
+Tried to clarify with "and then".}}.
+Such a \yc{dramatic apparent improvement} is expected for two reasons.
     First, it removes the nearby SNe~Ia, and thereby significantly reduces the
@@
         fitted asymmetric Gaussian distribution (its error) mimicking the BBC
-        approach. The dashed line shows the noe \yc{???} estimated by
+        approach. The dashed line shows that estimated by
     \cite{scolnic2018a}. The blue to red curves display our Base drifting
@@
         modeling of the underlying stretch distribution is comparable to the
-    observational effect of exotic form of dark energy. \yc{Is this fig. still relevant?}}
+    observational effect of exotic form of dark energy. \textit{\yc{Is this fig. still relevant?} \nn{Only as a possible representation of the potential effect of this
+    difference in modeling. Why would it not anymore?}}}
     \label{fig:magdrift}
@@
 
-We have presented a study of the drift of the underlying \yc{I still prefer ``intrinsic''} SNe~Ia stretch
+We have presented a study of the drift of the underlying \textit{\yc{I still prefer ``intrinsic''}} SNe~Ia stretch
 distribution as a function of redshift. We used SNe~Ia from magnitude-limited

### 2020-03-26 Nicolas.tex
Greg corr

--- 2020-02-17 Nicolas.tex
@@
 \begin{document}
-\title{Redshift Evolution of the Type Ia Supernova Population Stretch
+\title{Redshift Evolution of the Underlying Type Ia Supernova Stretch
 Distribution}
@@
     \inst{2}
+    \and Y. Copin\inst{1}   
     \and R. Graziani\inst{2}
+    %HERE ALPHABETIC ORDER
     \and M. Briday\inst{1}
-    \and Y. Copin\inst{1}
     \and Y.-L. Kim\inst{1}
+    \and Saul Perlmutter\inst{3}
+    \and G. Smadja\inst{1}   
 }
@@
     1, Villeurbanne; CNRS/IN2P3, Institut de Physique des Deux Infinis, Lyon
-    \and Université Clermont Auvergne, CNRS/IN2P3, Laboratoire de
+    \and 
+    Université Clermont Auvergne, CNRS/IN2P3, Laboratoire de
     Physique de Clermont, F-63000 Clermont-Ferrand, France.
+    \and
+    Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720 
 }
@@
 
-\abstract{\mr{ABSTRACT NOT DONE YET.}
-Type Ia supernovae (SNe Ia) allow for the construction of the Hubble
-    diagram, giving us information about the Universe's expansion and its
-    fundamental components, one of which is dark energy. But systematic
-    uncertainties are now starting to be limiting in our ability to measure
-    those parameters. In particular, the physics of SNe Ia is still mostly
-    unknown, and is thought not to change in time/with the redshift.}
-    {In an attempt to reduce those uncertainties, we try to find an empirical
-    law describing SNe Ia's length of explosion (stretch) evolution with the
-redshift.}
-    {We started by getting a complete sample representing all of the stretch
-        distribution that Nature can give us, before using LsSFR measurements,
-        an age tracer which evolution with redshift is known, that has been
-        shown to have a strong correlation with the stretch. We compare their
-        AIC, an estimator of the relative quality of statistical models that
-    includes the number of free parameters, to determine which ones describe
-best the data.}
-    {Models with an evolution of the stretch with the redshift have a better
-    AIC than the ones without.}
-    {We find that implementing these models allows us to fit the data better
-    than models without stretch evolution.}
+\abstract{
+{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
+measure of survey statistic increase, the question of astrophysical systematic
+uncertainties rises, and notably that of the SN~Ia population redshift drift.}
+{In this paper, we study the redshift evolution of SN~Ia
+\textsc{\texttt{SALT2.4}} lightcurve stretches, which is a purely intrinsic SN
+property, to probe this drift.}
+{The SN stretch has been shown to strongly correlate with the SN's environment,
+and notably with stellar age tracers. Following the prediction of the fraction
+of young and old SNe~Ia in nature from \cite{rigault2018}, and assuming
+non-drifting underlying stretch distribution for each, we model the expected
+SN~Ia stretch distribution as a function of redshift. We test our prediction
+against data from the literature chosen to have negligible selection
+effects to ensure that any observed drift is indeed astrophysical and not
+artificial.}
+{We clearly demonstrate that the underlying SNe~Ia stretch distribution is
+evolving as a function of redshift, and that the young/old drifting model is a
+much better description of the data than any non-drifting model, including the
+sample-based asymmetric distributions used by the Beams with Bias Correction
+algorithm, used to correct Malmquist bias. Our favored underlying stretch
+model is bimodal: a high-stretch mode shared by both young and old environment,
+and a low-stretch mode only existing in old environments.}
+{The underlying SNe~Ia population is evolving as a function of redshift and the
+exact impact on cosmology remains to be studied. Yet, the astrophysical drift of
+the SN stretch distribution does affect current Malmquist bias corrections and
+thereby distances derived from SN affected by selection effects. We highlight
+that such a bias will increase with surveys covering larger redshift ranges,
+which is particularly important for LSST.}
+
+}
 
@@
 by cosmic microwave background (CMB) data measured by Planck assuming the
-standard $\L$CDM\nn{,} or when the SN luminosity is anchored at intermediate
+standard $\L$CDM, or when the SN luminosity is anchored at intermediate
 redshift by the baryon acoustic oscillation (BAO) scale
@@
 red giant branch technique in place of the Cepheids seem to favor an
-\mr{intermediate} value of H$_0$ \citep{freedman2019}, time delay measurements
-from strong lensing seem to also favor high H$_0$ values \citep{wong2019}.
+intermediate value of H$_0$ \citep{freedman2019,freedman2020}, time delay
+measurements from strong lensing seem to also favor high H$_0$ values
+\citep{wong2019}.
 
@@
 For the last decades, numerous analysis have studied the relation between SNe~Ia
-and host properties\nn{,} finding first that the standardized SNe~Ia magnitude
-significantly depends on the host stellar-mass, SNe~Ia from high-mass host
-\nn{being brighter on average}
-\cite[e.g.][]{kelly2010, sullivan2010, childress2013, betoule2014, rigault2018, kim19}.
-This mass-step correction is currently used in cosmological analyses
-\citep[e.g.][]{betoule2014, scolnic2018a}, including for deriving H$_0$
-\citep{riess2016, riess2019}. Yet, the underlying connection between the SN and
-the host remains unclear when using global properties such as the host stellar
-mass, which raises the question of the accuracy of the correction. More
-recently, studies have used the local SN environment to probe more direct
-connections between the SN and its environments \citep{rigault2013}\nn{,}
+and host properties, finding first that the standardized SNe~Ia magnitude
+significantly depends on the host stellar-mass, SNe~Ia from high-mass host being
+brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
+betoule2014, rigault2018, kim19}.  This mass-step correction is currently used
+in cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including
+for deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
+between the SN and the host remains unclear when using global properties such as
+the host stellar mass, which raises the question of the accuracy of the
+correction. More recently, studies have used the local SN environment to probe
+more direct connections between the SN and its environments \citep{rigault2013},
 showing that local age-tracers such as the local specific star formation rate or
 the local color are more strongly correlated with the standardized SN magnitude
-\citep{rigault2018, roman2018, kim18}, suggesting the age to be the driving parameter
+\citep{rigault2018, roman2018, kim18}, suggesting age as the driving parameter
 underlying the mass-step. If true, this would have a significant impact for
@@
 In this paper, we take a step aside to probe the validity of our modeling of the
-SN population that we claim to be constituted of two age-populations
-\citep{rigault2013,rigault2015,rigault2018}: one old and one young, the former
+SN population which we claim to be constituted of two age-populations
+\citep{rigault2013,rigault2015,rigault2018}: one old and one younger, the former
 having on average lower lightcurve stretches and being brighter after
-standardization. We indeed use the correlation between the SN age probed with
-local specific star formation rate in \cite{rigault2018} and the SN stretch to
-model the expected evolution of the underlying SN stretch distribution as a
-function of redshift. This modeling relies on three assumptions: (1) there
-\nn{are indeed} 2 populations of SNe~Ia; (2) the relative fraction of each of
-these populations as a function of redshift follows the age-modeling made in
-\citep{rigault2018} and (3) the underlying distribution of stretch for each
-age-sample can be modeled and do not significantly drift. This paper aims at
-testing this with datasets from the literature.
-
-% SERIES OF PAPER
-%This paper is part of a series of papers where we test the validity of the two
-%SNe Ia population models developed in \citep{rigault2018}. Briday et al. in
-%prep will present how using different environmental tracer is affecting our
-%ability to distinguish the two populations, and Rigault et al. in prep will
-%further study the connection with H$_0$. Here we analyze predictions concerning
-%the SNe Ia redshift drifts: (1) that the distribution of the SN~Ia stretch, a
-%purely intrinsic SN~Ia property, is evolving as a function of redshift and (2)
-%that this population-drift follows the relation that \citep{rigault2018}
-%derived assuming that one population is “young“ with a rate proportional to the
-%star formation rate in the Universe and the second is "old" with a rate
-%proportional to the stellar mass.
-
-The concept of the SNe~Ia age dichotomy aroused when the SN~Ia rate has first
-been studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008}
-have shown that the relative SNe~Ia rate in galaxies could only be explained if
-two populations existed, one young, following the host star formation activity,
-and one old following the host stellar mass (the so called “prompt and delayed“
-or “A+B“ model). In \cite{rigault2018} we used the specific Star Formation Rate
-at the SN location (Local sSFR or LsSFR) to classify which are the prompts
-(those with a high LsSFR) and which are the delayed (those with low LsSFR).
-Furthermore, since the first SNe~Ia host analysis, the SN stretch has been known
-to be strongly correlated with the SN host properties \citep{hamuy1996,
-hamuy2000}, which has been extensively confirmed since \citep[e.g.][]{neill2009,
-sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
-rigault2013, pan2014, kim19}. Following the “A+B“ model and the connection between SN
-stretch and host properties, \cite{howell2007} first discussed the potential
-redshift drift of the SN stretch distribution. In this paper we revisit this
-analysis with the most recent SNe~Ia dataset.
+standardization. We use the correlation between the SN age probed with local
+specific star formation rate in \cite{rigault2018} and the SN stretch to model
+the expected evolution of the underlying SN stretch distribution as a function
+of redshift. This modeling relies on three assumptions: (1) there are two
+populations of SNe~Ia; (2) the relative fraction of each of these populations as
+a function of redshift follows the age-modeling presented in \cite{rigault2018}
+and (3) the underlying distribution of stretch for each age-sample can be
+modeled and does not significantly drift. This paper aims at testing this with
+datasets from the literature.
+
+The concept of the SNe~Ia age dichotomy arose with study of the the SN~Ia rate.
+\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
+the relative SNe~Ia rate in galaxies could only be explained if two populations
+existed, one young, following the host star formation activity, and one old
+following the host stellar mass (the so called “prompt and delayed“ or “A+B“
+model). In \cite{rigault2018} we used the specific Star Formation Rate at the SN
+location (Local sSFR or LsSFR) to classify which are the younger (those with a
+high LsSFR) and which are the older (those with low LsSFR). Furthermore, since
+the first SNe~Ia host analyses, the SN stretch has been known to be strongly
+correlated with the SN host properties \citep{hamuy1996, hamuy2000}, which has
+been extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
+    lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
+pan2014, kim19}. Following the “A+B“ model and the connection between SN stretch
+and host properties, \cite{howell2007} first discussed the potential redshift
+drift of the SN stretch distribution. In this paper we revisit this analysis
+with the most recent SNe~Ia dataset.
 
@@
 analysis, which is based on the Pantheon dataset \citep{scolnic2018a}. We
-discuss the importance of obtaining a "complete" sample, i.e. representative of
+discuss the importance of obtaining a ``complete'' sample, i.e. representative of
 the true underlying SNe Ia distribution, and how we build one from the Pantheon
 sample.  We then present in section~\ref{sec:modeling} our modeling of the
-distribution of stretch as a function of redshift based stretch distribution of
-young and old SNe~Ia. Our results are presented in section~\ref{sec:results}
-where we test if the SN stretch does evolve as a function of redshift and if the
-age model is in good agreement with this evolution. Consequence for cosmology of
-our results are briefly discussed in section~\ref{sec:bbc} and we conclude in
-section~\ref{sec:ccl}.
-
-%Throughout the paper we use the Planck 2015 cosmology from the astropy library.
-
-\section{Complete Sample Construction}
-\label{sec:sample}
+distribution of stretch as a function of redshift based models for the stretch
+distribution of young and old SNe~Ia. Our results are presented in
+section~\ref{sec:results}, where we test whether the SN stretch does evolve as a
+function of redshift and if the age model is in good agreement with this
+evolution. Consequences for cosmology of our results are briefly discussed in
+section~\ref{sec:bbc} and we conclude in section~\ref{sec:ccl}.
+
+\section{Complete Sample Construction}\label{sec:sample}
 
@@
 stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
-with the lightcurve stretch (and color), the first SN~Ia that a
+with the lightcurve stretch (and color), the first SNe~Ia that a
 magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
 Consequently, if selection effects are not accounted for, one might confuse true
-population drift with selection effects, and conversely \nn{\citep[e.g.,][]{kim19}}.
-
-\nn{Assuming} sufficient (and random \yc{unbiased?}) spectroscopic follow-up for acquiring
-typing and host redshift, the selection effects of magnitude-limited surveys
-should be negligible bellow a given redshift where even the faintest \nn{SN~Ia}
-can be observed. In contrast, targeted surveys have highly complex selection
-functions and will be discarded from our analysis. Fortunately, modern SN
-cosmology samples such as the Pantheon dataset are now dominated by magnitude
-limited surveys.
+population drift with selection effects, and conversely.
+
+Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring typing
+and host redshift, the selection effects of magnitude-limited surveys should be
+negligible below a given redshift where even the faintest SNe~Ia can be observed.
+In contrast, targeted surveys have highly complex selection functions and will
+be discarded from our analysis. Fortunately, modern SN cosmology samples such as
+the Pantheon dataset are now dominated by magnitude limited surveys.
 
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
-    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$) lightcurve parameters
-        of SNe~Ia from the SDSS, PS1 and SNLS samples from the Pantheon catalog.
-        The individual SNe are shown as blue dots and a 2D histogram is
-        shown in gray to highlight point density. The ellipse
-        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the
-        corresponding standardized absolute magnitude using the $\alpha$ and
-        $\beta$ coefficients from \cite{scolnic2018a}. \nn{The lines represent
-        the $(x_1, c)$ evolution for $m = m_{lim}$ in Eq.~\ref{eq:m_lim},
-    for $z$ between $0.60$ and $z=1.00$ using SNLS's $m_{lim}$ of $24.8$~mag. The SNe
-    above these lines are those for which $m > m_{lim}$, meaning they cannot be
-    observed.}}
+    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
+        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples from
+        the Pantheon catalog.  The individual SNe are shown as blue dots and a
+        2D histogram is shown in gray to highlight point density. The ellipse
+        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the corresponding
+        standardized absolute magnitude using the $\alpha$ and $\beta$
+        coefficients from \cite{scolnic2018a}. The grey diagonal lines represent
+    the $(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and
+$z=1.70$ using SNLS's $m_{lim}$ of $24.8$~mag.}
     \label{fig:maglim}
@@
 We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
-distributions of SNe~Ia from \nn{the following surveys}: PanStarrs
+distributions of SNe~Ia from the following surveys: PanStarrs
 \citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
@@
 \citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
-c)$ plan with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates
-the full distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011}
-and \citet{campbell2013} for similar contours, the second using a more
-conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and
-$c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
-standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$ along the aforementioned ellipse
-given the standardization coefficient $\alpha=0.156$ and $\beta=3.14$ from
-\cite{scolnic2018a}: the faintest SN~Ia is that with $(x_1=-1.66, c=0.25)$
-and an absolute standardized magnitude at peak in Bessel-B band of
-$M^{t_0}_{min} = -18.31$~mag. Since one need to detect this object typically a week
-before and 10 days after peak to build a suitable lightcurve, the limiting
-\nn{standardized} absolute magnitude effectively is approximately
-$M_{min} = -18.00$~mag. Hence, given the \nn{\sout{$5\sigma$ point source detection}}
-magnitude limit $m_{lim}$ of a magnitude limited survey, one can derive the
-maximum redshift $z_{lim}$ above which the faintest SNe~Ia will be missed \nn{using the
-relation between apparent magnitude, redshift and absolute magnitude given by:}
-\begin{equation}\label{eq:m_lim}
-    \nn{\cancel{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\;\text{mag}}
-    < m_{lim}}}
-\end{equation}
-\begin{equation}\label{eq:m_lim}
-    \yc{\mu(z_{lim}) = m_{lim} - M_{min}} \quad\text{min vs. lim???}
-\end{equation}
+c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full distribution
+\citep{guy2007,betoule2014}; see also \citet{bazin2011} and \citet{campbell2013}
+for similar contours, the second using a more conservative $|c| \leq 0.2$ cut.
+Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is $M_0=-19.36$
+\citep{kessler2009,scolnic2014}, we can derive the absolute standardized
+magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$ along the
+aforementioned ellipse given the standardization coefficient $\alpha=0.156$ and
+$\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is that with
+$(x_1=-1.66, c=0.25)$ and an absolute standardized magnitude at peak in Bessel-B
+band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect this object
+typically a week before and 10 days after peak to build a suitable lightcurve,
+the limiting standardized absolute magnitude effectively is approximately
+$M_{lim} = -18.00$~mag. Hence, given the magnitude limit $m_{lim}$ of a
+magnitude limited survey, one can derive the maximum redshift $z_{lim}$ above
+which the faintest SNe~Ia will be missed using the relation between apparent
+magnitude, redshift and absolute magnitude $\mu(z_{lim}) = m_{lim} - M_{lim}$.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
+    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia  from
+        the SDSS, PS1 and SNLS dataset respectively (data from  Pantheon
+        \citealt{scolnic2018a}).  The colored parts represent the distribution
+        of SNe~Ia kept in our analysis for they are supposedly free from
+        selection bias (see section~\ref{sec:sample}). The dark (resp. light)
+    color responds to the conservative (resp. fiducial) selection cut.}
+    \label{fig:cuts}
+\end{figure}
 
@@
 filter that has a $5\sigma$ depth of
-24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT final release website.}}. 
-This converts to a $z_{lim}=0.60$, in perfect agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. \nn{Fig.~14 of \citealt{perrett2010}
-suggests a lower limit of $z_{lim}=0.55$; both will be considered as discussed below.}
+24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
+final release website.}}. This converts to a $z_{lim}=0.60$, in agreement with
+\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.  Fig.~14 of
+\citealt{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both will be
+considered, as discussed below.
 
@@
 
-In similar redshift range, SDSS has a limiting magnitude of
-22.5 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However,
-the SDSS surveys were more sensitive to limited spectroscopic resources.
-\cite{kessler2009} \nn{pointed out} that during year-1 of SDSS, SNe~Ia with
+In similar redshift range, SDSS has a limiting magnitude of 22.5
+\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
+SDSS surveys were more sensitive to limited spectroscopic resources.
+\cite{kessler2009} pointed out that during year-1 of SDSS, SNe~Ia with
 $r-mag<20.5$ were favored for spectroscopic follow up, corresponding to a
 redshift cut at $0.15$. For the rest of the SDSS survey, additional
-spectroscopic resources were used, and \cite{kessler2009} and
-\cite{dilday2008} show a relative completeness up to $z_{lim}=0.2$. Following
-these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit \sout{for
-the rest of \nn{this study}}.
-
-For the rest of the analysis, we will consider \nn{\sout{both
-cases,}} the \nn{aforementioned} redshift limits as well as more conservative cuts
-for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for
-PS1 and $z_{lim}=0.55$ for SNLS.
+spectroscopic resources were used, and \cite{kessler2009} and \cite{dilday2008}
+show a relative completeness up to $z_{lim}=0.2$. Following these analyses we
+will use $z_{lim}=0.2$ as the baseline SDSS redshift limit.
+
+For the rest of the analysis, we will consider the aforementioned redshift
+limits as well as more conservative cuts for each of these surveys, namely
+$z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS1 and $z_{lim}=0.55$ for SNLS.
 
@@
 redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
-We see there that the redshift limits we have selected roughly
-correspond to the peak of these histograms. \nn{As the peaks indicate that we
-are starting to miss the faintest SNe~Ia, it seems logical to have the redshift
-cuts fall near said peaks.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55.pdf}
-    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
-    SNLS from the Pantheon dataset. The colored part of the
-distribution shows which SNe~Ia has been kept in our analysis for they are
-supposedly free from selection bias (see section~\ref{sec:sample}). The
-dark (resp. light) color responds to the conservative (resp. fiducial) selection cut.}
-    \label{fig:cuts}
-\end{figure}
+As expected, the selected redshift limits roughly correspond to the peak of
+these histograms.
 
 In addition, we use the SNe~Ia from the Nearby Supernova Factory
-\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
+\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
 been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
-4.2.2). \nn{Their spectroscopic follow-up were done on a redshift range of
+4.2.2). Their spectroscopic follow-up was done over a redshift range of
 $0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
-deeper. As such, these SNe~Ia} should also be a random sampling of the underlying
-SN population. The SNf sample is particularly useful for studying SN property
+deeper. As such, these SNe~Ia should also be a random sampling of the underlying
+SN population. The SNfactory sample is particularly useful for studying SN property
 drift as it enables us to have a large SN~Ia sample at $z<0.1$.
 
-\nn{Finally, we included t}he HST sample
-from Pantheon\nn{, that} follows the same logic of having a search deeper than the
-follow-up and we therefore kept it entirely \citep{strolger04}, see
-Table~\ref{tab:sample}. \nn{We present the stretch distribution and redshift histogram of these five surveys in Fig.~\ref{fig:sample}}.
+Finally, we included the HST sample from Pantheon, that follows the same logic
+of having a search deeper than the follow-up and we therefore kept it entirely
+\citep{strolger04}, see Table~\ref{tab:sample}. We present the stretch
+distribution and redshift histogram of these five surveys in
+Fig.~\ref{fig:sample}.
 
@@
     \centering
-    \caption{\nn{Source surveys and number of SNe~Ia used in our ``complete''
-    sample. Conservative cuts are indicated in brackets.}}
+    \caption{Composition of the SNe~Ia dataset used in this analysis.
+    Conservative cuts are indicated in brackets.}
     \label{tab:sample}
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb.pdf}
-    \caption{\nn{\textit{Main}: SN stretches in each survey as a function of redshift, using solid (resp. empty) markers for the
-    conservative (resp. fiducial) selection. \textit{Top}: combined redshift 
-    histogram for all surveys, in dark (resp. light) gray for the conservative (resp. fiducial) selection cuts.}}
+    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb-cividis.pdf}
+    \caption{\textit{Lower}: \textsc{\texttt{SALT2.4}} lightcurve stretch as a
+        function of redshift (in log) for each survey considered in this
+        analysis (see legend).  Full (resp. empty) markers correspond to the
+        conservative (resp. fiducial) selection cuts. \textit{Top}: combined
+        redshift histogram in dark (resp. light) gray for the conservative
+    (resp. fiducial) selection cuts.}
     \label{fig:sample}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%
-%   LITERATURE        %
-%%%%%%%%%%%%%%%%%%%%%%%
-%\mri{COMPARE HERE WITH LITERATURE. E.G. FIG. 6 OF SCOLNIC 2018, PS zmax
-%AROUND0.27}
-
-%\mri{Perrett et al. 2010 | SNLS | zmax ="z $\sim$ 0.6"}
-
-%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, Fig. 10 suggest still
-%okat z=0.2}
-
-%\mri{Kessler 2009 "For the SDSS-II, the image-subtraction
-%pipeline efficiency (	subtr) is complete up to a redshift of z ∼ 0.2"}
-
-%\mri{Kesslet 2009 | "For the SDSS-II and SNLS samples, the cutoff redshifts are
-%estimated to be 0.15 and 0.65, respectively."}
-% ADDIND: PS cut from Pantheon close to 0.27 see Fig. 6 of Scolnic 2018
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%       INFO          %
-%%%%%%%%%%%%%%%%%%%%%%%
-%Fraction of random SDSS stretches being < mean_conservative =  0.299
-%Fraction of random PS1 stretches being < mean_conservative =  0.254
-%Fraction of random SNLS stretches being < mean_conservative =  0.417
-
-\section{Modeling the redshift drift}
-\label{sec:modeling}
-
-In~\cite{rigault2018} we presented a modeling of the evolution of the fraction
-of prompt and delayed SNe~Ia as a function of redshift following former work on
+\section{Modeling the redshift drift}\label{sec:modeling}
+
+In~\cite{rigault2018} we presented a model for the evolution of the fraction of
+younger and older SNe~Ia as a function of redshift following former work on
 rates and delay time distributions \citep[e.g.,][]{mannucci2005,
 scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
-In short, we assumed that the number of prompt SNe~Ia follows the star formation
-activity \nn{(SFR)} in the Universe while the number of delayed SNe~Ia follows
-the number of Gyr-old stars in the Universe, i.e. the stellar mass \nn{(M)}.
-Hence, if we denote $\delta(z)$ (resp. $\psi(z) = 1 - \delta(z)$) the fraction of young (resp.
-old) SNe~Ia in the Universe as a function of redshift, then their ratio
-$\delta/\psi$ is expected to follow the evolution of the specific star formation
-rate \nn{(SFR/M)} which goes as $(1+z)^{2.8}$ until $z\sim2$
-\citep[e.g.,][]{tasca2015}. Since \nn{$\delta(0.05) \sim \psi(0.05)$}
+In short, we assumed that the number of younger SNe~Ia follows the star
+formation activity (SFR) in the Universe while the number of older SNe~Ia
+follows the number of Gyr-old stars in the Universe, i.e. the stellar mass
+(M$^*$). Hence, if we denote $\delta(z)$ (resp. $\psi(z) = 1-\delta(z)$) the
+fraction of young (resp. old) SNe~Ia in the Universe as a function of redshift,
+then their ratio $\delta/\psi$ is expected to follow the evolution of the
+specific star formation rate (SFR/M$^*$) which goes as $(1+z)^{2.8}$ until
+$z\sim2$ \citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
 \citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
-\citep{mannucci2006,rodney2014}, we conclude in
-\cite{rigault2018} that
-%\begin{align}
-%    \label{eq:delta}
-%    \delta(z) &= \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
-%    \mathrm{and}\\
-%    \psi(z) &= \left( K \times (1+z)^{+2.8} +1 \right)^{-1}
-%\end{align}
+\citep{mannucci2006,rodney2014}, we concluded in \cite{rigault2018} that
 \begin{equation}
     \label{eq:delta}
-    \yc{\delta(z) = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}}
+    \delta(z) = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
 \end{equation}
@@
 
-\subsection{\nn{Base underlying distribution}}
+\subsection{“Base” underlying stretch distribution}
 \label{sec:basemodel}
@@
     \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
-    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$) as a function of
-        the local specific star formation rate (LsSFR) for SNf SNe used in this
-        analysis. The color corresponds to the probability $p_y$ for the SNe~Ia to be
-        young, i.e. to have $\log\mathrm{LsSFR} \geq -10.82$; see
-        \cite{rigault2018}. \textit{Right}: histogram of $p_y$-weighted
-        distribution of the SN stretchs. On both panel, we illustrate the
-        Base model fitted on the SNf SNe. The young and old population
-        contributions are shown in purple and yellow, respectively.}
+    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$)
+        as a function of the local specific star formation rate (LsSFR) for SNf
+        SNe used in this analysis. The color corresponds to the probability,
+        $p_y$, for the SNe~Ia to be young, i.e. to have $\log\mathrm{LsSFR} \geq
+        -10.82$; see \cite{rigault2018}. \textit{Right}: histogram of
+        $p_y$-weighted distribution of the SN stretchs. On both panels, we
+        illustrate the Base model fitted on the SNfactory SNe. The younger and
+    older population contributions are shown in purple and yellow,
+respectively.}
     \label{fig:stretchlssfr}
@@
 
-To model the evolution of the SN stretch
-as a function of redshift, given our aforementioned model of the evolution of
-the fraction of young and old SNe~Ia with redshift, we need to model the SN
-stretch distribution for each age subsample.
-
-\nn{In~\cite{rigault2018}, we presented the relation between SNe stretches and
-LsSFR measurements, which trace progenitor age, from the SNf sample, and is
-shown in Fig.~\ref{fig:stretchlssfr}}. Given the structure of the
-stretch-LsSFR scatter plot, our model is defined as follows: 
+To model the evolution of the SN stretch as a function of redshift, given our
+aforementioned model of the evolution of the fraction of younger and older
+SNe~Ia with redshift, we need to model the SN stretch distribution for each age
+subsample.
+
+In~\cite{rigault2018}, we presented the relation between SNe stretches and LsSFR
+measurements, which trace progenitor age, using the SNfactory sample. This
+relation is shown in Fig.~\ref{fig:stretchlssfr}, and given the structure of the
+stretch-LsSFR scatter plot, our model of the underlying SN~Ia stretch
+distribution is defined as follows:
 \begin{itemize}
-    \item for the young population the underlying stretch distribution
-    is modeled as a normal distribution $\mathcal{N}(\mu_1, \sigma_1^2)$; 
-    \item the old population stretch distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
-    \mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a Gaussian mixture where one mode is the
-    same as the young population one. \yc{explain meaning of $a$}
+    \item for the younger (i.e., $\log( \mathrm{LsSFR})>-10.82$) population, the
+        underlying stretch distribution is modeled as a normal distribution
+        $\mathcal{N}(\mu_1, \sigma_1^2)$; 
+    \item the old (i.e., $\log( \mathrm{LsSFR})<-10.82$) population stretch
+        distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) +
+        (1-a)\times \mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a Gaussian mixture
+        where one mode is the same as for the young population, $a$
+        representing the relative influence of the two Gaussians.
 \end{itemize}
-\nn{When combining our prediction of the
-evolution of young SNe~Ia as a function of redshift (Eq.~\ref{eq:delta}),
-and the stretch distribution of both young and old SNe~Ia, our model for the
-underlying distribution of stretch $X_1(z)$ as a function of
-redshift is:}
+When combining our prediction of the evolution of younger SNe~Ia as a function
+of redshift (Eq.~\ref{eq:delta}), and the stretch distribution of both younger
+and older SNe~Ia, our model for the underlying distribution of stretch,
+$X_1(z)$, as a function of redshift is:
 \begin{align}
     \label{eq:stretchz}
-    \nn{X_1(z) = {}}
+    X_1(z) = {}
     \delta(z) &\times \mathcal{N}(\mu_1,\sigma_1^2) + {}\nonumber\\
@@
 
-The \yc{maximum-likelihood estimate} of the 5 free parameters
+The maximum-likelihood estimate of the 5 free parameters
 $\theta\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained by
-minimizing a pseudo-$\chi^2$ $\mathcal{L}$ defined as:
+minimizing a pseudo-$\chi^2$ $\mathcal{X}^2$ defined as:
 \begin{equation}
     \label{eq:likelihood}
-    \mathcal{L} = -2 \sum_i \ln \prob{x^{i}_{1}}{\theta;
-    \mathrm{d}x^{i}_{1}, \nn{y^i}},
+    \mathcal{X}^2 = -2 \sum_i \ln \prob{x^{i}_{1}}{\theta;
+    \mathrm{d}x^{i}_{1}, y^i},
 \end{equation}
@@
     y^i &\times \mathcal{N}\left(x^{i}_{1} \mid \mu_1, \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
-        (1-y^i) &\times \nn{\bigg[} 
+        (1-y^i) &\times \bigg[
         a \times \mathcal{N}\left(x^{i}_{1} \mid \mu_1,
@@
      &(1-a) \times \mathcal{N}\left(x^{i}_{1} \mid \mu_2,
-     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) \nn{\bigg]}
+     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) \bigg]
 \end{align}
 where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
-$y^i$ are the \textsc{\texttt{SALT2.4}} stretch, its associated error 
-and the probability that the SN is young, respectively. 
-In practice, to ensure fraction $a$ is constrained between 0 and 1,
-we fit for $\alpha$ such
-that $a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.
-
-Depending on \yc{whether $y^i$ can be 
-estimated directly from LsSFR measurements or not}, there are two ways to proceed.
-
-\subsubsection{\yc{With LsSFR measurements}}
-\label{sec:modelpy}
-
-\yc{For the SNfactory sample, we can readily set $y^i = p_y{}^i$, the
-probability to have $\log \mathrm{LsSFR} \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}),
-to minimize Eq.~\ref{eq:likelihood} with respect to~$\theta$.}
-\mr{Results on fitting the SNf SNe with this model are shown
-Table~\ref{tab:modelresults} and illustrated in Fig.~\ref{fig:stretchlssfr}.}
+$y^i$ are the \textsc{\texttt{SALT2.4}} stretch, its associated error and the
+probability that the SN is young, respectively.  In practice, to ensure fraction
+$a$ is constrained between 0 and 1, we fit for $\alpha$ such that
+$a=\arctan(\alpha)/\pi+0.5$, which results in an asymmetric error on $a$.
+
+Depending on whether $y^i$ can be estimated directly from LsSFR measurements or
+not, there are two ways to proceed, which we now discuss.
+
+\subsubsection{With LsSFR measurements}\label{sec:modelpy}
+
+For the SNfactory sample, we can readily set $y^i = p_y{}^i$, the probability to
+have $\log \mathrm{LsSFR} \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
+minimize Eq.~\ref{eq:likelihood} with respect to~$\theta$.  Results on fitting
+the SNfactory SNe with this model are shown Table~\ref{tab:modelresults} and
+illustrated in Fig.~\ref{fig:stretchlssfr}.
 
@@
     \caption{Best fit values of the parameters for the Base stretch distribution
-    model when applied to the SNf dataset only (114 SNe~Ia), the reference 569
+    model when applied to the SNfactory dataset only (114 SNe~Ia), the reference 569
 SN~Ia sample or the conservative one (422).}
@@
         \hline\\[-0.8em]
-        SNf & $0.41 \pm 0.08$  & $0.55 \pm 0.06$ & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$ & $0.48^{+0.08}_{-0.08}$ \\[0.15em]
-        \yc{Reference} & $0.37 \pm 0.05$  & $0.61 \pm 0.04$ & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$ & $0.51^{+0.09}_{-0.10}$ \\[0.15em]
-        \yc{Conservative} & $0.38 \pm 0.05$  & $0.60 \pm 0.04$ & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\[0.15em]
+        SNfactory & $0.41 \pm 0.08$  & $0.55 \pm 0.06$ & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$ & $0.48^{+0.08}_{-0.08}$ \\[0.15em]
+        Fiducial & $0.37 \pm 0.05$  & $0.61 \pm 0.04$ & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$ & $0.51^{+0.09}_{-0.10}$ \\[0.15em]
+        Conservative & $0.38 \pm 0.05$  & $0.60 \pm 0.04$ & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\[0.15em]
         \hline
@@
 
-\subsubsection{Without LsSFR measurements}
-\label{sec:modelnopy}
-
-\yc{In lack of direct LsSFR measurements (i.e. $p_y^i$), we can extend the analysis to 
-non-SNfactory samples by using the redshift-evolution of the fraction $\delta(z)$ 
-of young SNe~Ia (Eq.~\ref{eq:delta}) as a proxy for the probability of a 
-SN to be young.} This still corresponds to minimizing Eq.~\ref{eq:likelihood} 
-with respect to the parameters 
-$\theta\equiv(\mu_1, \mu_2, \sigma_1, \sigma_2, a)$ 
-of the stretch distribution $X_1$ (Eq.~\ref{eq:stretchz}), 
-but this time assuming $y^i = \delta(z^i)$ for any given SN~$i$.
-
-\mr{For the rest of the analysis, we will therefore minimize
-Eq.~\ref{eq:likelihood} using \nn{$p_y{}^i$} --~the
-probability for the SN $i$ to be young~-- when available (i.e. for SNf
-dataset) and $\delta(z^{i})$ --~the expected fraction of young SNe~Ia at the
-SN redshift $z^{i}$~-- otherwise.}
-
-\mr{Results on fitting this model on all the 569 (resp. 422) SNe from the 
-fiducial (resp. conservative) sample
-are given Table~\ref{tab:modelresults} and illustrated in
-Fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
-stretch per bin of redshift closely follows our redshift drift modeling; that is, when
-considering selection-bias free SN~Ia samples, SNe~Ia at higher redshift have on
-average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
-redshift ($-0.17\pm 0.10$ at $z\sim0.05$). This is indeed what is expected if
-old environments favor low SN stretches \citep[e.g.][]{howell2007} \emph{and} if
-the fraction of old SNe~Ia reduces as a function of redshift. See
-section~\ref{sec:results} for a more quantitative discussion.}
+\subsubsection{Without LsSFR measurements}\label{sec:modelnopy}
+
+When lacking direct LsSFR measurements (i.e. $p_y^i$), we can extend the
+analysis to non-SNfactory samples by using the redshift-evolution of the
+fraction $\delta(z)$ of young SNe~Ia (Eq.~\ref{eq:delta}) as a proxy for the
+probability of a SN to be young. This still corresponds to minimizing
+Eq.~\ref{eq:likelihood} with respect to the parameters $\theta\equiv(\mu_1,
+\mu_2, \sigma_1, \sigma_2, a)$ of the stretch distribution $X_1$
+(Eq.~\ref{eq:stretchz}), but this time assuming $y^i = \delta(z^i)$ for any
+given SN~$i$.
+
+For the rest of the analysis, we will therefore minimize Eq.~\ref{eq:likelihood}
+using $p_y{}^i$ --~the probability for the SN $i$ to be young~-- when available
+(i.e. for SNfactory dataset), and $\delta(z^{i})$ --~the expected fraction of
+young SNe~Ia at the SN redshift $z^{i}$~-- otherwise.
+
+Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
+(resp. conservative) sample are given Table~\ref{tab:modelresults} and
+illustrated in Fig.~\ref{fig:modelall}. We see in this figure that the measured
+mean SN~Ia stretch per bin of redshift closely follows our redshift drift
+modeling; that is, when considering selection-bias-free SN~Ia samples, SNe~Ia at
+higher redshift have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$)
+than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$). This is indeed
+what is expected if old environments favor low SN stretches
+\citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
+a function of redshift. See section~\ref{sec:results} for a more quantitative
+discussion.
 
@@
         as a function of redshift. Markers show the mean stretch measured in
-    redshift bins of equal sample size indicated in light gray at the bottom
-    of each redshift bins. Full and light markers are used when considering
-    the fiducial or the conservative samples, respectively. The orange horizontal line represents the mean
-    redshift of the reference sample illustrating the expectation if the SN
-    stretch distribution is not evolving with redshift. Best fits of our Base drifting
-    model are shown as blue, dashed-blue and gray, when fitted on the
-    fiducial sample, the conservative one or the SNf dataset only, respectively; all are \yc{compatible}.
-The light-blue illustrates the amplitude of the error of the best fit model when
-considering the reference dataset.}
+        redshift bins of equal sample size, indicated in light gray at the
+        bottom of each redshift bins. Full and light markers are used when
+        considering the fiducial or the conservative samples, respectively. The
+        orange horizontal line represents the mean stretch of the reference
+        sample, illustrating the expectation if the SN stretch distribution is
+        not evolving with redshift. Best fits of our Base drifting model are
+        shown as blue, dashed-blue and gray, when fitted on the fiducial sample,
+        the conservative one or the SNfactory dataset only, respectively; all
+        are compatible. The light-blue band illustrates the amplitude of the
+    error of the best fit model when considering the reference dataset.}
     \label{fig:modelall}
@@
 
-\subsection{Alternative models}
-\label{sec:othermodel}
-
-In section~\nn{\ref{sec:basemodel}}, we have modeled the underlying stretch
-distribution following \cite{rigault2018}, i.e. as a single Gaussian for the young SNe~Ia and a mixture of two
-Gaussians for the old SNe Ia population, the same as the young-population
-one plus another for the fast declining SNe~Ia that seem to only exist in old
-populations. This is our so-called ``Base'' model.
-
-However, to test \nn{different modeling choices}, we have implemented a suite
-of alternative parametrizations that we also adjust to the data following the procedure
-described section~\ref{sec:modelnopy}.
-
-%Namely we consider:
-%\begin{itemize}
- %   \item “Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$”: Base model but where the high-stretch Gaussian of the old population is not set as that of the young population (+2 degrees of freedom);
- %   \item “Base$-(\sigma_2)$”: Base model but where $\sigma_2$ is fixed at the same value as $\sigma_1$ (-1 degree of freedom);
- %   \item “Howell”: one single Gaussian per age group as in \cite{howell2007}.
-%\end{itemize}
-
-\cite{howell2007} used a simpler model with a normal distribution for each young 
-and old populations. \nn{We thus consider a ``Howell+drift'' model, with one 
-single Gaussian per age group}. \yc{one has to explain the ``drift'' part!!!}
-
-\yc{Alternatively, since we aim at probing the existence of a redshift drift, 
-we are also testing non-drifting models by restricting the ``Base'' and ``Howell'' models to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe; these models are labeled ``Base constant'' and ``Howell constant''.} \yc{fix labels!!!}
-
-We are also considering other non-drifting models, and notably the one
-developed for the Beam with Bias Correction \cite[BBC,][]{scolnic2016,
-kessler2017}, currently used in recent SN cosmological analyses
-\cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to account
-for Malmquist biases. The BBC formalism assumes sample-based (hence
-intrinsically non-drifting) asymmetric Gaussian stretch distributions: $\mathcal{N}\left(\mu,
-\sigma_{-}^2\;\text{if}\;x_1<\mu\;\text{else}\;\sigma_{+}^2\right)$. The
-idea behind this sample-based approach is twofold: (1) Malmquist biases are
-driven by survey properties and (2) because current surveys cover limited
-redshift ranges, doing so covers some potential redshift evolution information
-\citep{scolnic2016, scolnic2018a}. See further discussion concerning the BBC in section~\ref{sec:bbc}.
-
-Finally, for the sake of completeness, we also
-consider simple ``Gaussian'' %$\mathcal{N}(\mu, \sigma)$ 
-and ``asymmetric'' Gaussian %$\mathcal{N}(\mu, \sigma_{\pm})$ 
-non-drifting models.
-
-% The ability of these models to explain the observations is detailed in
-% section~\ref{sec:results}.
-
-\section{Results}
-\label{sec:results}
+\subsection{Alternative models}\label{sec:othermodel}
+
+In section~\ref{sec:basemodel}, we have modeled the underlying stretch
+distribution following \cite{rigault2018}, i.e.\ as a single Gaussian for the
+young SNe~Ia and a mixture of two Gaussians for the old SNe Ia population, one
+the same as for the young-population plus another for the fast declining SNe~Ia
+that seem to only exist in old host galaxy stellar populations. This is our
+so-called ``Base'' model.
+
+However, to test different modeling choices, we have implemented a suite of
+alternative parametrizations that we also adjust to the data following the
+procedure described in section~\ref{sec:modelnopy}.
+
+\cite{howell2007} used a simpler model per age bin, assuming a normal
+distribution for both the young and the old population. We thus consider a
+``Howell+drift'' model, with one single Gaussian per age group and the
+$\delta(z)$  drift from \cite[][see Eq.~\ref{eq:delta}]{rigault2018}.
+
+Alternatively, since we aim at probing the existence of a redshift drift, we
+also test non-drifting models by restricting the ``Base'' and ``Howell'' models
+to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe;
+these models are labeled ``Base constant'' and ``Howell constant''.
+
+We also consider other non-drifting models, notably the one developed for Beams
+with Bias Correction \cite[BBC,][]{scolnic2016, kessler2017}, used in recent SN
+cosmological analyses \cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016,
+riess2019} to account for Malmquist biases. The BBC formalism assumes
+sample-based (hence intrinsically non-drifting) asymmetric Gaussian stretch
+distributions: $\mathcal{N}\left(\mu,
+\sigma_{-}^2\;\text{if}\;x_1<\mu,\;\text{else}\;\sigma_{+}^2\right)$. The idea
+behind this sample-based approach is twofold: (1) Malmquist biases are driven by
+survey properties and (2) because current surveys cover limited redshift ranges,
+doing so covers some potential redshift evolution information
+\citep{scolnic2016, scolnic2018a}. See further discussion concerning BBC in
+section~\ref{sec:bbc}.
+
+Finally, for the sake of completeness, we also consider simple ``Gaussian'' and
+``asymmetric'' Gaussian non-drifting models.
+
+\section{Results}\label{sec:results}
 
 We applied each model on both the reference and conservative samples (cf.
-section~\ref{sec:sample}). Because the models presented
-in section~\ref{sec:modeling} have different degrees of freedom, we use the Akaike
-Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare them. This estimator
-penalizes extra degrees of freedom to avoid over-fitting the data. It is defined
-as follow:
+section~\ref{sec:sample}). Because the models presented in
+section~\ref{sec:modeling} have different degrees of freedom, we use the Akaike
+Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare them. This
+estimator penalizes extra degrees of freedom to avoid over-fitting the data. It
+is defined as follow:
 \begin{equation}
-    \mathrm{AIC} = \yc{\mathcal{L}_{min}} + 2k
+    \mathrm{AIC} = \mathcal{X}^2_{min} + 2k
 \end{equation}
-where \yc{$\mathcal{L}_{min}$} is
-the minimum value of the pseudo-$\chi^2$ as defined Eq.~\eqref{eq:likelihood} and $k$ is the number of free parameters. The \nn{reference} model
-is the one with the smallest AIC, and the probability for another model to be at
-least as representative of the data as this one is given by:
-\begin{equation}
-    p(\mathrm{other} \geq \mathrm{ref}) =
-    \exp\left(\Delta\mathrm{AIC}/2\right)
-\end{equation}
-\yc{with $\Delta\mathrm{AIC} < 0$}. Results are presented in Table~\ref{tab:comp} and are illustrated
-Fig.~\ref{fig:mod_comp}.
+where $\mathcal{X}^2_{min}$ is the minimum value of the pseudo-$\chi^2$ as
+defined Eq.~\eqref{eq:likelihood} and $k$ is the number of free parameters. The
+reference model is the one with the smallest AIC. In comparison to this model,
+those with $\Delta\mathrm{AIC}<5$ are acceptable, those with
+$5<\Delta\mathrm{AIC}<10$ are unfavored and those with $\Delta\mathrm{AIC}>20$
+are excluded. This roughly corresponds to 2, 3 and 5 $\sigma$ limits for a
+Gaussian probability distribution.
 
@@
     \centering
-    \caption{Comparison of the relative ability for a model to describe the data
-        using AIC criterion. See the model definition
-        section~\ref{sec:modeling}. The drift column indicates if the model
-        contains the age-drifting model ($\delta(z)$) or not (constant $f$, i.e.
-        non-drifting). The BBC is sample based which somewhat includes
-        redshift dependency information. $\mathcal{L}$ is the pseudo-$\chi^{2}$
-        defined Eq.~\ref{eq:likelihood} and Proba is the probability that a
-        model is at least as good as the reference model to describe the data.
-    See also Fig.~\ref{fig:mod_comp}.}
+    \caption{Comparison of the relative ability for a model to describe the
+    data. For each considered model, we report if the model is drifting or not,
+its number of free parameters and, for both the fiducial and the conservative
+cuts, the pseudo-$\chi^2$, $\mathcal{X}^2$ (see Eq.~\ref{eq:likelihood}), the AIC
+and the AIC difference ($\Delta$AIC) between this model and the Base model used
+as reference for it has the lowest AIC.}
     \label{tab:comp}
-    \begin{tabular}{c c c  |c c c c|  c c c c}\hline\hline\\[-0.8em]
-        & & & \multicolumn{4}{|c}{\yc{Fiducial sample (569 SNe)}} & \multicolumn{4}{|c}{\yc{Conservative sample (422 SNe)}} \\
+    \begin{tabular}{c c c  |c c c |  c c c }\hline\hline\\[-0.8em]
+        & & & \multicolumn{3}{|c}{Fiducial sample (569 SNe)} & \multicolumn{3}{|c}{Conservative sample (422 SNe)} \\
         %\cline{2-6}
@@
         Name & drift & $k$ &
-        $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba & $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba \\[0.15em]
+        $\mathcal{X}^2$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ &  $\mathcal{X}^2$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$\\[0.15em]
         \hline\\[-0.8em]
@@
         Base & $\delta(z)$ & 5
-        & 1456.7 & 1466.7 & -- & --
-        & 1079.5 & 1089.5 & -- & --\\[0.15em]
-
-        Howell & $\delta(z)$ & 4
-        & 1463.3 & 1471.3 & $-4.6$ & $1.0\times10^{-1}$
-        & 1088.2 & 1096.2 & $-6.7$ & $3.4\times10^{-2}$\\[0.15em]
+        & 1456.7 & 1466.7 & -- 
+        & 1079.5 & 1089.5 & -- \\[0.15em]
+
+        Howell+drift & $\delta(z)$ & 4
+        & 1463.3 & 1471.3 & $-4.6$  %&$1.0\times10^{-1}$
+        & 1088.2 & 1096.2 & $-6.7$ 
+        %& $3.4\times10^{-2}$
+        \\[0.15em]
 
         Asymmetric & -- & 3
-        & 1485.2 & 1491.2 & $-24.5$ & $4.7\times10^{-6}$
-        & 1101.3 & 1107.3 & $-17.8$ & $1.4\times10^{-4}$\\[0.15em]
-
-        Howell & $f$ & 5
-        & 1484.2 & 1494.2 & $-27.5$ & $1.0\times10^{-6}$
-        & 1101.2 & 1111.2 & $-21.7$ & $1.9\times10^{-5}$\\[0.15em]
-
-        Base & $f$ & 6
-        & 1484.2 & 1496.2 & $-29.5$ & $3.9\times10^{-7}$
-        & 1101.2 & 1113.2 & $-23.7$ & $7.1\times10^{-6}$\\[0.15em]
+        & 1485.2 & 1491.2 & $-24.5$  %&$4.7\times10^{-6}$
+        & 1101.3 & 1107.3 & $-17.8$ 
+        %& $1.4\times10^{-4}$
+        \\[0.15em]
+
+        Howell:const & $f$ & 5
+        & 1484.2 & 1494.2 & $-27.5$  %&$1.0\times10^{-6}$
+        & 1101.2 & 1111.2 & $-21.7$ 
+        %& $1.9\times10^{-5}$
+        \\[0.15em]
+
+        Base:const & $f$ & 6
+        & 1484.2 & 1496.2 & $-29.5$  %&$3.9\times10^{-7}$
+        & 1101.2 & 1113.2 & $-23.7$ 
+        %& $7.1\times10^{-6}$
+        \\[0.15em]
 
         BBC & per sample & 3x5
-        & 1468.2 & 1498.2 & $-31.5$ & $1.5\times10^{-7}$
-        & 1083.6 & 1113.6 & $-24.1$ & $5.7\times10^{-6}$\\[0.15em]
+        & 1468.2 & 1498.2 & $-31.5$  %&$1.5\times10^{-7}$
+        & 1083.6 & 1113.6 & $-24.1$ 
+        %& $5.7\times10^{-6}$
+        \\[0.15em]
 
         Gaussian & -- & 2
-        & 1521.8 & 1525.8 & $-59.1$ & $1.5\times10^{-13}$
-        & 1142.6 & 1146.6 & $-57.1$ & $4.0\times10^{-13}$ \\\hline\hline
+        & 1521.8 & 1525.8 & $-59.1$  %&$1.5\times10^{-13}$
+        & 1142.6 & 1146.6 & $-57.1$ 
+        %& $4.0\times10^{-13}$ 
+        \\\hline\hline
     \end{tabular}
@@
     \includegraphics[width=\linewidth]{Article_figures/mod_comp.pdf}
-    \caption{$\Delta$AIC between the reference ``Base'' model and the other
-    models (see also Table~\ref{tab:comp}). The full and open blue markers
-correspond to model with and without redshift drift, respectively. Light markers
-show the results when this analysis is performed on the conservative sample 
-rather than the fiducial one. Color-bands illustrate the validity of the models, from 
-Acceptable ($\Delta\mathrm{AIC} > -5$) to Excluded ($\Delta\mathrm{AIC} < -20$),
-see text \yc{This is not explained in the text!}. \yc{According to the AIC}, all non-drifting models (open symbols) are excluded for the reference sample.}
+    \caption{$\Delta$AIC between ``Base'' model (reference) and other models
+        (see Table~\ref{tab:comp}). Full and open blue markers correspond to
+        models with and without redshift drift, respectively. Light markers show
+        the results when this analysis is performed on the conservative sample
+        rather than the fiducial one.  Color-bands illustrate the validity of
+        the models, from  Acceptable ($\Delta\mathrm{AIC} > -5$) to Excluded
+        ($\Delta\mathrm{AIC} < -20$); see text. According to the AIC, all
+    non-drifting models (open symbols) are excluded to be as good representation
+of the data in comparison to our Base (drifting) model.}
     \label{fig:mod_comp}
@@
 
-\yc{We first note that the best model (with smallest AIC) is the so-called Base model, 
-implementing \cite{rigault2018} description of the stretch distribution and its 
-redshift evolution. This is therefore the reference model, both on the fiducial and 
-conservative samples.}
-
-\mr{\yc{Furthermore,} we find that underlying stretch distributions not including 
-    a redshift drift are all
-    excluded to describe the data as well as the Base model. In fact,
-    the best non drifting model (Asymmetric) has \yc{a very marginal chance 
-    ($p = 2\times10^{-6}$)} \yc{not same number as in table!!!} to
-    describe the data as well as the Base model.}
-    \yc{This result is just a quantitative assessment of qualitative facts clearly visible in 
-    Fig.~\ref{fig:modelall}: the mean SN stretch per bin of redshift strongly 
-    suggests a redshift evolution rather than having a constant value, and this evolution is compatible with the redshift drift expected from Eq.~\ref{eq:delta}.}
-
-\yc{Are the next 2 paragraphs (now commented out) still pertinent? The model does not appear in table nor 
-fig. Mickael, please rephrase and remove what is not needed anymore.}
-
-% \mr{As just mentioned, the \nn{reference model is Base}, but we also implemented
-%     a Base$-(\sigma_2)$ model, where both the low-stretch (mode 2 that only
-%     exists in old environments) and high-stretch (mode 1 that exists in all
-%     environments) normal distributions share the same width. It is marginally
-%     better than the Base model where both distributions have their own
-%     \nn{dispersion} ($\Delta \mathrm{AIC}<2$). Similarly, letting the
-%     high-stretch distribution to vary between the young and the old population
-%     (\nn{model we call} Base$+(\mu_1^{\mathrm{O}}, \sigma_1^{\mathrm{O}})$) only
-%     marginally improves the fit ($\Delta\mathcal{L}=3.6$ in comparison to Base)
-%     at a cost of 2 more degrees of freedom, making this mode less likely. This
-%     suggests that, indeed, the high-stretch mode (mode 1) exists in all kind of
-%     age-populations. We also remark that considering the stretch distribution of
-%     each age-population as independent normal distributions, as suggested by
-% \cite{howell2007}, is \nn{significantly disfavored}.}
-
-% \mr{Our results strongly suggest that the underlying distribution of SN~Ia
-%     stretch evolves as a function of redshift. Our redshift drift modeling
-%     presented section~\ref{sec:modeling} is a much better representation of the
-%     Pantheon data (see section~\ref{sec:sample}) than any non-drifting models,
-%     including the sample-based BBC one. Considering or not the conservative
-% sample has no effect on our conclusion as illustrated Fig.~\ref{fig:modelall}
-% and presented Table~\ref{tab:comp}.}
-
-\yc{This paragraph is not clear: what is the impact, and its importance, on which case?
-Isn't it in contradiction with the conclusions of BBC section, where it is said that proper handling of stretch correction is of paramount importance for accurate cosmology???}
-\mr{To the best of our knowledge, the SN~Ia stretch redshift drift has never
-    been explicitly accounted for in cosmological analyses. Not doing so is a
-    second order issue for SN cosmology as it only affects the way one does
-    Malmquist bias correction. Indeed, as long as the Phillip's relation
-    standardization parameter $\alpha$ is not redshift dependent (a study behind
-    the scope of this paper, but see e.g. \citealt{scolnic2018a}), if a sample
-    is free from selection effects, the stretch-corrected SNe~Ia magnitudes used
-    for cosmology are blind to the underlying stretch distribution.  However,
-    since the first missed SNe~Ia are the faintest, which are also those with
-    the lowest stretch, once we include SNe~Ia affected by selection effects we
-    need to account for the fact that only the brightest SNe~Ia exist in the
-    low-stretch end of the sample after standardization. The modeling of the SN
-    stretch then enters the analysis to understand which SNe~Ia are missing
-given the selection functions to mimic a selection-bias free sample.}
-
-\section{\nn{BBC discussion}}
-\label{sec:bbc}
-
-\mr{The state-of-the-art tool for doing so \yc{what?} is the BBC described in
-    \cite{scolnic2016} and \cite{kessler2017}, which is used in recent SN
-    cosmological analyses \citep{jones2018b, scolnic2018a, brout2019,
-    descosmopaper2019} including the direct measurement of H$_0$
-\citep{riess2016,riess2019}.}
-
-\mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
+We first note that the best model (with smallest AIC) is the so-called Base
+model, implementing the \cite{rigault2018} description of the stretch
+distribution and its redshift evolution. This is therefore the reference model,
+both on the fiducial and conservative samples.
+
+Furthermore, we find that underlying stretch distributions not including a
+redshift drift are all excluded as description of the data relative to the Base
+model. In fact, the best non-drifting model (Asymmetric) has a very marginal
+chance ($p \equiv \exp\left(\Delta\mathrm{AIC}/2\right) = 5\times10^{-6}$) to
+describe the data as well as the Base model. This result is just a quantitative
+assessment of qualitative facts clearly visible in Fig.~\ref{fig:modelall}: the
+mean SN stretch per bin of redshift strongly suggests redshift evolution
+rather than a constant value, and this evolution is compatible with the
+redshift drift expected from Eq.~\ref{eq:delta}.
+
+For the sake of robustness, we also tested allowing the high-stretch mode of the
+old-population to differ from the young-population mode, hence adding two
+degrees of freedom. The corresponding best fit is not significantly better
+resulting in a $\Delta$AIC of $-0.4$, strengthening the fact that the young and
+old population indeed share the same underlying high-stretch mode. We also
+allowed the young population to have a low-stretch mode, finding its amplitude
+to be compatible with 0 ($<2\%$).
+
+\section{Discussion}\label{sec:discussion}
+
+To the best of our knowledge, the SN~Ia stretch redshift drift has never been
+explicitly accounted for in cosmological analyses. Not doing so is a second
+order issue for SN cosmology as it only affects the way one does Malmquist bias
+correction.  Indeed, as long as the Phillip's relation standardization parameter
+$\alpha$ is not redshift dependent (a study behind the scope of this paper, but
+see e.g. \citealt{scolnic2018a}), if a sample is free from selection effects,
+the stretch-corrected SNe~Ia magnitudes used for cosmology are blind to the
+underlying stretch distribution. Yet, modern surveys do have significant
+Malmquist bias at least for the upper half of their SN redshift distribution.
+Then, an ill-modeling of the underlying stretch distribution will bias the
+derived SN magnitude for these \citep[see e.g.,][]{rubin2015,rubin2016}. The
+fact that the currently-used stretch modeling for Malmquist bias correction in
+SN cosmology (BBC) is excluded to be an as good description of the data then our
+drifting model is worrisome and we further discussed in the next sections.
+
+\subsection{BBC formalism and underlying stretch distributions}\label{sec:bbc}
+
+The commonly used tool for doing Malmquist bias correction in SN cosmology is
+the BBC formalism described in \cite{scolnic2016} and \cite{kessler2017}, which
+is used in recent SN cosmological analyses \citep{jones2018b, scolnic2018a,
+brout2019, descosmopaper2019} including the direct measurement of $H_0$
+\citep{riess2016,riess2019}.
+
+Surprisingly, the BBC technique, which assumes an individual asymmetric
 distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
-    modelisation \yc{in our analysis}. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the smallest of all
-    non-drift models (but still $\Delta \mathcal{L}=-11.5$ worse than the
-    reference Base model), it is penalized by requiring 15 free parameters (3 for each of the 5 samples of the analysis).
-    \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
-    highlighted that, because \yc{traditional} surveys span limited redshift ranges, the
-    per sample approach can somewhat account for \yc{implicit} redshift drifts. We
-    stress here that, as measurements of modern surveys cover \yc{increasingly} larger redshift
-    ranges to reduce calibration systematic  uncertainties, this approach is becoming
-    less \yc{valid}, notably for PS1, DES and, soon, LSST \yc{redshift ranges? Why not SNLS too?}. Already with the current
-    surveys, the BBC sample-based technique is \nn{ruled out} to be an as good
-    representation of the data as our best model ($p=4\times 10^{-8}$). A more
-    detailed discussion of the consequence of this results for cosmology follows
-in section~\ref{sec:bbc}.}
-
-\mr{We report in Table~\ref{tab:bbc} the samples' $\sigma_{\pm}$ and
-    $\mu_0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
-    SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{our} construction
-    of \nn{a} selection-bias free sample, see
-    section~\ref{sec:sample}. \yc{WHY??? The developed BBC for \textit{incomplete} samples!} If we were to use \cite{scolnic2016} and
-    \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
-    the $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC
-    modeling would go even deeper from $-32$ to $-47$. \nn{It would seem that our minimization
-    of these parameters is closer to our Base model than theirs}. \yc{???} \yc{Last},
-    fixing the Base model parameters to those derived using SNf data only (top
-    row Table~\ref{tab:modelresults}) and using the BBC parameters from the
-    aforementioned literature for PS1, SDSS and SNLS, we find
-    $\Delta\mathrm{AIC}=-31.7$, but removing the SNf sample reduces this difference to
-    $\Delta\mathrm{AIC}=-5$ \yc{not clear: the SNf sample is removed or not???}. Such a \yc{dramatic apparent improvement} is expected for two reasons.
-    First, it removes the nearby SNe~Ia, and thereby significantly reduces the
-    redshift \nn{lever} arm to probe the impact of redshift drifts.
-    Second, the SNf data are better at discriminating the individual stretch
-    distributions by age since the probability $y^i$ for a SN to be young is
-    known ($y^i = p_y{}^i$) while other samples rely on a probabilistic proxy for
-    $y^i$, namely the fraction $\delta(z^i)$ 
-    of young SNe~Ia at SN redshift. While in any case the Base model is favored, the relative weight
-    of the SNf sample highlights the importance of adding non-targeted nearby
-surveys, such as the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,
-graham2019} to unambiguously confirm our \nn{model}.}
-
-%Using S18/SK16 values of \sigma_-, \sigma_+ and \mu_0, we find that SDSS, PS1 and SNLS's $\mathcal{L}$ to be 457.9, 397.6 and 254.5 respectively, and their AIC to be 464.0, 403.7 and 260.8 respectively, systematically higher than those we find in this study (see table \ref{tab:bbc_comp}).
+modelisations in our analysis (see section~\ref{sec:results}.) While its
+pseudo-$\chi^{2}$, $\mathcal{X}^2$, is the smallest of all non-drift models (but
+still $\Delta \mathcal{X}^2=-11.5$ worse than the reference Base model), it is
+also penalized for requiring 15 free parameters (3 for each of the 5 samples of
+the analysis). \citet[][section~2]{scolnic2016} and
+\citet[][section~5.4]{scolnic2018a} highlighted that, because traditional
+surveys span limited redshift ranges, the per sample approach can somewhat
+account for implicit redshift drifts. We stress here that, as measurements of
+modern surveys cover increasingly larger redshift ranges to reduce calibration
+systematic uncertainties, this approach is becoming less valid, notably for PS1,
+DES and, soon, LSST. Already with current surveys, the BBC sample-based
+technique is ruled out at representing the data as well as our best model;
+$\Delta\mathrm{AIC}<-20$, which could be interpreted as a probability $p=2\times
+10^{-7}$ of being an as good representation of the data as the Base model. Note
+that even considering the SNfactory dataset like any other (i.e.  ignoring the
+LsSFR information see section~\ref{sec:modeling}), reduces this difference to
+$\Delta\mathrm{AIC}<-10$, still strongly unfavoring the modeling currently used
+as part of the BBC Malmquist bias correction.
+
+We report in Table~\ref{tab:bbc} the samples' $\sigma_{\pm}$ and $\mu_0$ fitted
+on the supposedly selection-free samples using our fiducial cuts (see
+section~\ref{sec:sample}. We find our results in close agreement with
+\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
+derived these model parameters using the full BBC formalism, using numerous
+simulations to model the selection effects (see details e.g., section 3 of
+\citealt{kessler2017}). The agreement between our fit of the asymmetric
+Gaussians on the supposedly selection-free part of the samples and the results
+derived using the BBC formalism supports our approach to get a sample with
+negligible selection effects. If we were to use \cite{scolnic2016} and
+\cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively, the
+$\Delta\mathrm{AIC}$ between our Base drifting model and the BBC modeling would
+go even deeper from $-32$ to $-47$.
 
@@
     \centering
-    \caption{Best fit parameters for our sample-based asymmetric modeling of the
+    \caption{Best-fit parameters for our sample-based asymmetric modeling of the
     underlying stretch distribution.}
@@
     Asymmetric & $\sigma_{-}$    & $\sigma_{+}$    & $\mu_0$ \\\hline\\[-0.8em]
-    SNf &  1.34 $\pm$ 0.13 & 0.41 $\pm$ 0.10 & 0.68 $\pm$ 0.15\\[0.15em]
+    SNfactory  & 1.34 $\pm$ 0.13 & 0.41 $\pm$ 0.10 & 0.68 $\pm$ 0.15 \\[0.15em]
     SDSS       & 1.31 $\pm$ 0.11 & 0.42 $\pm$ 0.09 & 0.72 $\pm$ 0.13 \\[0.15em]
@@
     SNLS       & 1.41 $\pm$ 0.13 & 0.15 $\pm$ 0.13 & 1.22 $\pm$ 0.15 \\[0.15em]
-    HST &  0.76 $\pm$ 0.36 & 0.79 $\pm$ 0.35 & 0.11 $\pm$ 0.44 \\\hline\hline
+    HST        & 0.76 $\pm$ 0.36 & 0.79 $\pm$ 0.35 & 0.11 $\pm$ 0.44 \\\hline
+    \hline
     \end{tabular}
@@
 
+\subsection{First step toward quantification of the cosmological impact of
+inaccurate stretch distribution modeling.}
+
 \begin{figure}
     \centering
-    \includegraphics[width=\linewidth]{Article_figures/bbc_comp_PS1.pdf}
-    \caption{Underlying SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
-        distributions for the PS1 sample. The Green line (band) shows our best
-        fitted asymmetric Gaussian distribution (its error) mimicking the BBC
-        approach. The dashed line shows the noe \yc{???} estimated by
-    \cite{scolnic2018a}. The blue to red curves display our Base drifting
-stretch model estimated at each PS SN redshift; bluer for lower redshifts.}
+    \includegraphics[width=\linewidth]{Article_figures/bbc_comp_PS1_hist-nr.pdf}
+    \caption{Distribution of the PS1 SN~Ia \textsc{\texttt{SALT2.4}} stretch
+        ($x_1$) after the fiducial redshift limit cut (grey histogram). This
+        distribution is supposed to be a random draw from the underlying
+        stretch distribution. The green line shows the BBC model of this
+        underlying distribution (asymetric Gaussian). The full line (band) is
+        our best fit (its error); the dashed line shows \cite{scolnic2018a}
+        result. The black line (band) show our best fitted Base-modeling (its
+        error, see Table XX) that includes redshift drift. For illustration, we
+        show as colored (from blue to red with increasing redshifts) the
+    evolution of the underlying stretch distribution as a function of redshift
+for the redshift range covered by PS1 data.}
     \label{fig:bbc_pdf_ps1}
@@
 
-\mr{We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
+We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
 underlying stretch distribution between the BBC modeling and our Base drifting
-model for the PS1 sample. Our model is bimodal and the relative
-amplitude of each mode depends on the redshift-dependent fraction of old SNe~Ia in the sample: the higher the fraction of old SNe~Ia, the higher the amplitude of the
-low-stretch mode. As expected the two
-approaches strongly differ in modeling the negative part of the SN stretch
-distribution. The BBC asymmetric distribution goes in the middle of the bimodal
-distribution, largely over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
-largely under-estimating it at $x_1\sim-1.7$ for typical PS1 SN redshifts. This
-means that the bias-correction of the individual SN stretch
-$\overline{\delta}_{x_1}$ \citep{kessler2017} and consequently the magnitude
-bias correction $\mu_B$ is most likely inaccurate. However, given the complexity
-of the BBC analysis, a full study using our Base model given
-Eq.~\ref{eq:stretchz} in place of the sample-based asymmetric modeling would be
-necessary to have an exact understanding of the effect this inaccurate modeling
-has on cosmology.}
-
-\mr{Nonetheless, to have a rough estimation of the expected amplitude of the
-effect, we show Fig.~\ref{fig:magdrift} the difference of mean stretch
-correction $\alpha\times\left(\langle x_1 \rangle_{\mathrm{BBC}} - \langle x_1
-\rangle_{\mathrm{Base}}\right)$ as a function of redshift, using $\alpha=0.156$
-from \cite{scolnic2018a}. We see in this figure that the amplitude of the
-magnitude bias caused by an \yc{inaccurate} modeling of the underlying stretch
-distribution is of the order of few tens of millimag, which is the typical scale
-that exotic forms of dark energy could have on $\mu(z)$. In the era of modern
-cosmology where we aim at probing $w_0$ at a sub-percent and $w_a$ at a few
-percent \citep[e.g.,][]{lsstpaper}, it is therefore of paramount importance to
-further understand the exact modeling of the lightcurve parameters and
-especially their potential redshift drift; see also discussions about the impact
-of stretch and color distribution modeling for cosmology in \citealt{rubin2015}
-and \citealt{rubin2016}.}
+model for the PS1 sample. Our model is bimodal and the relative amplitude of
+each mode depends on the redshift-dependent fraction of old SNe~Ia in the
+sample: the higher the fraction of old SNe~Ia (lower redshift), the higher the
+amplitude of the low-stretch mode. This redshift dependency is shown as blue to
+red underlying distributions in Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges
+covered by PS1. The observed $x_1$ histogram corresponds to the sum of
+individual underlying SN-redshift distributions. As expected, the two approaches
+strongly differ in modeling the negative part of the SN stretch distribution.
+The BBC asymmetric distribution goes through the middle of the bimodal
+distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
+under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting
+model for typical PS1 SN redshifts. This means that the bias-correction of
+individual SN stretch, $\overline{\delta}_{x_1}$, \citep{kessler2017} and
+consequently the magnitude bias correction, $\mu_B$, is most likely inaccurate
+and this inaccuracy is redshift dependent .
+
+The amplitude of this bias for cosmology is beyond the scope of this paper given
+the complexity of the BBC analysis. It would require a full study using our Base
+model given Eq.~\ref{eq:stretchz} in place of the sample-based asymmetric
+modeling as part of the BBC simulations.
+
+Nonetheless, it is interesting to study the  expected mean stretch difference as
+a function of redshift when considering that the underlying SN stretch
+distribution follows either the BBC formalism or our Base drifting modeling.
+This is shown Fig.~\ref{fig:magdrift}, where, for clarity we convert this
+difference in $x_1$ to a difference in magnitude using $\alpha=0.156$
+\citep{scolnic2018a}.
+
+We see in this figure that the amplitude of modeling difference is a few tens of
+millimag, which corresponds to the expected bias affecting SNe~Ia impacted
+by sample selection effects. For comparison, the amplitude of this effect is
+similar in scale to the effect of exotic forms of dark energy could
+have on $\mu(z)$. Consequently, in the era of modern cosmology, where we aim at
+probing $w_0$ at a sub-percent and $w_a$ at a few percent
+\citep[e.g.,][]{lsstpaper}, it is therefore of paramount importance to further
+understand the exact modeling of the lightcurve parameters when using SNe~Ia
+affected by Malmquist bias.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{Article_figures/BBC_distmod_w0wa.pdf}
-    \caption{Difference of mean stretch standardization (in mag) between the BBC
-        algorithm used in \cite{scolnic2018a} and our Base drifting model.
-        \nn{The small colored points show the SDSS, PS1 and SNLS evolution,
-            while the orange markers represent their mean evolution in 30
-            regular redshift bins.} For comparison, we show in gray bands the
-            expected magnitude difference between $\Lambda$CDM and $w$CDM
-            cosmologies ($w=-1\pm 10,\,5\,\mathrm{and}\,1\%$) or, in \yc{blue},
-            more exotic $w_0,w_a$CDM examples. This figure illustrates that the
-            amplitude of the potential magnitude bias caused by inaccurate
-        modeling of the underlying stretch distribution is comparable to the
-    observational effect of exotic form of dark energy. \yc{Is this fig. still relevant?}}
+    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
+    \caption{Difference of mean stretch standardization (in mag, using
+        $\alpha=0.156$ from \citealt{scolnic2018a}) between the BBC algorithm
+        and our Base drifting model, as a function of redshift, considering the
+        SDSS, PS1 and SNLS datasets. The colored markers show the difference for
+        SDSS, PS1 and SNLS samples (see legend) considering the individual SN
+    redshifts (no redshift cut applied). The orange markers represent their mean
+evolution in 30 regularly-spaced redshift bins.}
     \label{fig:magdrift}
@@
 
-We have presented a study of the drift of the underlying \yc{I still prefer ``intrinsic''} SNe~Ia stretch
+We have presented a study of the drift of the underlying SNe~Ia stretch
 distribution as a function of redshift. We used SNe~Ia from magnitude-limited
@@
 low-redshift bin. We only considered the supernovae that have been discovered
-in redshift ranges of each surveys where selection effect are negligible. This
-way the observed SNe~Ia stretch are random sampling of the true underlying
+in redshift ranges of each surveys where selection effects are negligible. This
+way the observed SNe~Ia stretches are random sampling of the true underlying
 distribution. This corresponds to 569 SNe~Ia (422 when considering more
@@
 
-Following observations made in \cite{rigault2018} we introduced a redshift drift
-modeling \nn{which} depends on the expected fraction of young and old SNe~Ia as a
-function of redshift, each age-population having its own underlying stretch
-distribution. We have studied various suites of distribution\nn{, including} non-drifting
-modeling. We also studied the stretch distribution modeling \nn{of the}
-current version of the BBC algorithm used for Malmquist bias correction by 
-recent Dark energy or Hubble constant cosmological analyses.
-
-Our conclusions are the following.
+Following observations made in \cite{rigault2018}, we introduced a redshift
+drift modeling which depends on the expected fraction of young and old SNe~Ia as
+a function of redshift, each age-population having its own underlying stretch
+distribution. We have studied various suites of distributions, including
+non-drifting modeling. We also studied the stretch distribution modeling of the
+current version of the BBC Malmquist bias correction algorithm used by recent
+SN~Ia cosmological analyses.
+
+Our conclusions are the following:
 \begin{enumerate}
 
-    \item Non-drifting models are excluded to be as good descriptions of the
-        data \nn{as} our Base model. This model assumes that: (1) the young
-        population has an unimodal Gaussian stretch distribution while the old
-        population stretch one is bimodal, one mode being the young one; (2) the
-        evolution of the relative fraction of young and old SNe~Ia follows the
-        prediction made in \cite{rigault2018}.
-
-    \item Given 1., we conclude that the SNe~Ia \nn{stretch} is indeed drifting, as
-        already suggested by e.g.~\cite{howell2007}.
-
-    \item The BBC approach, which assumes each survey to have its own asymmetric
+    \item Non-drifting models are excluded as good descriptions of the data of
+        our Base model. This model assumes that: (1) the younger population has
+        a unimodal Gaussian stretch distribution while the older population
+        stretch one is bimodal, one mode being the same as the young one; (2)
+        the evolution of the relative fraction of younger and older SNe~Ia
+        follows the prediction made in \cite{rigault2018}.
+
+    \item Given 1., we conclude that the SNe~Ia stretch is indeed drifting, as
+        previously suggested by e.g.~\cite{howell2007}.
+
+    \item The BBC approach, which assumes each survey has its own asymmetric
         Gaussian distribution, is excluded to be an as good description of the
-        data as our drifting model ($p=4\times10^{-8}$). Hence, the BBC
-        sample-based approach does not accurately account for redshift drift \nn{contrary
-        to the suggestion of} \cite{scolnic2016} and \cite{scolnic2018a}.
-
-    \item An inaccurate underlying stretch distribution modeling is estimated to
-        bias the mean standardized SNe~Ia \nn{to} the order of a few percents, which
-        is comparable in scale to the observational signature of exotic forms of
-        dark energy.
-
-    \item We suggest to use the following stretch population modeling:
+        data as our drifting model. Hence, the BBC sample-based approach does
+        not accurately account for redshift drift, unlike what was suggested by
+        \cite{scolnic2016} and \cite{scolnic2018a}.
+
+    \item An inaccurate underlying stretch distribution model is estimated to
+        bias the mean standardized magnitude of SNe~Ia affected selection
+        effects to the order of a few percents. This is comparable in scale to
+        the observational signature of exotic forms of dark energy and is
+        therefore a significant source of systematic that should be carefully
+        studied.
+
+    \item We suggest the use of the following stretch population model:
         \begin{align}
             \label{eqconclusion:stretchz}
-            \nn{X_1\left(z \right) =
-            } &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
+            X_1\left(z \right) =
+             &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
             (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
@@
         \end{align}
-    with: $a=0.50$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
+    with: $a=0.51$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
     $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}) and using the
@@
 
-\bibitem[Aldering(2004)]{aldering2004} Aldering, G.\ 2004, APS April Meeting Abstracts 2004, V4.003
+\bibitem[Aldering et al.(2002)]{aldering2002} Aldering, G., Adam, G., Antilogus, P., et al.\ 2002, \procspie, 61
 
@@
 
+\bibitem[Freedman et al.(2020)]{freedman2020} Freedman, W.~L., Madore, B.~F., Hoyt, T., et al.\ 2020, arXiv e-prints, arXiv:2002.01550
+
 \bibitem[Frieman et al.(2008)]{frieman2008} Frieman, J.~A., Bassett, B., Becker, A., et al.\ 2008, \aj, 135, 338
@@
 % I
+\bibitem[Ivezi{\'c} et al.(2019)]{lsstpaper} Ivezi{\'c}, {\v{Z}}., Kahn, S.~M., Tyson, J.~A., et al.\ 2019, \apj, 873, 111
+
 % J
@@
 % Y
+
 % Z
@@
 \end{itemize}
-The ones without a stretch evolution are the same ones but with an "F" implying
+The ones without a stretch evolution are the same ones but with an ``F'' implying
 we set $\delta(z) = f = \mathrm{cst}$, and two others:
@@
      &  & \multicolumn{3}{|c}{Asym (this)} & \multicolumn{3}{|c}{Asym (SK18/S18)} & \multicolumn{3}{|c}{Base SNf} \\
-    N$_{\mathrm{SN}}$ & Survey & $\mathcal{L}$ & AIC & Free param & $\mathcal{L}$ & AIC & Free param & $\mathcal{L}$ & AIC & Free param \\\hline
+    N$_{\mathrm{SN}}$ & Survey & $\mathcal{X}^2$ & AIC & Free param & $\mathcal{X}^2$ & AIC & Free param & $\mathcal{X}^2$ & AIC & Free param \\\hline
     167 & SDSS & 445.0 & 451.2 & 3 & 457.9 & 464.0 & 3 & 452.9 & 452.9 & 0\\
@@
      & \multicolumn{3}{|c}{Asym (this)}  & \multicolumn{3}{|c}{Base SNf} \\
-    Survey & $\mathcal{L}$ & AIC & Free param & $\mathcal{L}$ & AIC & Free param \\\hline
+    Survey & $\mathcal{X}^2$ & AIC & Free param & $\mathcal{X}^2$ & AIC & Free param \\\hline
     SNf & 299.3 & 305.5 & 3 & 268.6 & 279.1 & 5\\
@@
     \end{tabular}
-\end{table*}
+\end{table*}we are also testing non-drifting models by restricting the ``Base'' and ``Howell'' models to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe; these models are labeled ``Base constant'' and ``Howell constant''.} \yc{\uline{fix labels!!!
+\nn{done}}}

### 2020-04-15 Nicolas.tex
Change on name, Yannick corr, soon Greg2+Saul+Nordin

@@
 
-\usepackage{color}
+\usepackage[dvipsnames]{xcolor}
 \newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
 \newcommand{\mri}[1]{{\textcolor{red}{#1}}}
-\newcommand{\nn}[1]{{\textcolor[rgb]{1, 0.27, 0}{#1}}}
-\newcommand{\yc}[1]{{\textcolor{blue}{#1}}}
+\newcommand{\nn}[1]{{\textcolor[rgb]{0.25, 0.50, 0}{#1}}}
+\newcommand{\yc}[1]{{\textcolor{BrickRed}{#1}}}
 
@@
     \thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
-    \inst{2}
-    \and Y. Copin\inst{1}   
+    \inst{2}\and\\
+    Y. Copin\inst{1}   
     \and R. Graziani\inst{2}
     %HERE ALPHABETIC ORDER
+    \and G. Aldering\inst{3}
     \and M. Briday\inst{1}
@@
 {In this paper, we study the redshift evolution of SN~Ia
-\textsc{\texttt{SALT2.4}} lightcurve stretches, which is a purely intrinsic SN
+\textsc{\texttt{SALT2.4}} lightcurve stretch\yc{\sout{es}}, \yc{\sout{which is}} a purely intrinsic SN
 property, to probe this drift.}
 {The SN stretch has been shown to strongly correlate with the SN's environment,
-and notably with stellar age tracers. Following the prediction of the fraction
-of young and old SNe~Ia in nature from \cite{rigault2018}, and assuming
-non-drifting underlying stretch distribution for each, we model the expected
+and notably with stellar age tracers. Following the \yc{evolution of} the fraction
+of young and old SNe~Ia \yc{\sout{in nature} as predicted by} \cite{rigault2018}, and assuming
+non-drifting \yc{$\to$constant/stationary?}  underlying stretch distribution for each \yc{population}, we model the expected
 SN~Ia stretch distribution as a function of redshift. We test our prediction
@@
 effects to ensure that any observed drift is indeed astrophysical and not
-artificial.}
-{We clearly demonstrate that the underlying SNe~Ia stretch distribution is
+\yc{\sout{artificial} observational}.}
+{We clearly demonstrate that the underlying SN\yc{e?}~Ia stretch distribution is
 evolving as a function of redshift, and that the young/old drifting model is a
@@
 sample-based asymmetric distributions used by the Beams with Bias Correction
-algorithm, used to correct Malmquist bias. Our favored underlying stretch
-model is bimodal: a high-stretch mode shared by both young and old environment,
-and a low-stretch mode only existing in old environments.}
-{The underlying SNe~Ia population is evolving as a function of redshift and the
-exact impact on cosmology remains to be studied. Yet, the astrophysical drift of
+algorithm used to correct Malmquist bias. Our favored underlying stretch
+model is bimodal: a high-stretch mode shared by both young and old environments,
+and a low-stretch mode \yc{\sout{only existing in} exclusive to} old environments.}
+{\sout{The underlying SNe~Ia population is evolving as a function of redshift \yc{This is purely a repetition of a previous sentence} \nn{\textit{see new sentence}} and the
+exact impact on cosmology remains to be studied.} \nn{The precise impact of the redshift evolution of the underlying SNe Ia population on cosmology remains to be studied}.
+Yet, the astrophysical drift of
 the SN stretch distribution does affect current Malmquist bias corrections and
@@
 \section{Introduction}
+
+\yc{YC: you should run a grammar corrector, e.g. \texttt{grammarly}. e.g. "SNe dichotomy" vs. "SN dichotomy", "analysis" vs. "analyses", etc.}
+
 Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
@@
 understand the properties of dark energy (DE) as it is the only tool able to
-precisely map the recent expansion rate $z<0.5$, when DE is driving it
+precisely map the recent expansion rate ($z<0.5$), when DE is driving it
 \citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
@@
 the SNe~Ia are anchored on Cepheids \citep[the SH0ES
-project][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
-by cosmic microwave background (CMB) data measured by Planck assuming the
-standard $\L$CDM, or when the SN luminosity is anchored at intermediate
+project,][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
+\yc{from} cosmic microwave background (CMB) data measured by Planck assuming the
+standard $\L$CDM \yc{ref?} \nn{textit{isn't it at the end of the sentence? Do you think it should be before the "or"?}}, or when the SN luminosity is anchored at intermediate
 redshift by the baryon acoustic oscillation (BAO) scale
-\citep{riess2019,reid2019,planck2016, feeney2019}. While using the tip of the
+\citep{riess2019,reid2019,planck2018, feeney2019}. While using the tip of the
 red giant branch technique in place of the Cepheids seem to favor an
@@
 matter-radiation decoupling moment caused by a (fine tuned) early dark energy
-\citep{poulin2019}.
+\citep{poulin2019}. \yc{I don't see the point of going into this debate...}
 
 Alternatively, systematic effects affecting one or several of the aforementioned
-analysis could also explain the tension.  In \cite{rigault2015} we suggested
-that SNe Ia from the Cepheid calibrator sample differs by construction from  the
-Hubble flow as it strongly favors young stellar population environments where
+analyses could also explain the tension. \cite{rigault2015} suggested
+that SNe Ia from the Cepheid calibrator sample differ by construction from the
+Hubble flow \yc{ones} as it \yc{they?} strongly favors \yc{SNe don't favor anything...} \nn{\textit{but the Cepheid sample does, right?}} young stellar population environments where
 one could find Cepheids. This selection effect would impact the derivation of
-H$_0$ if SNe~Ia from young and older environments differ in average magnitudes.
-
-For the last decades, numerous analysis have studied the relation between SNe~Ia
+H$_0$ if SNe~Ia from young and older environments differ in average \yc{standardized} magnitudes.
+
+For the last decades, numerous analyses have studied the relation between SNe~Ia
 and host properties, finding first that the standardized SNe~Ia magnitude
-significantly depends on the host stellar-mass, SNe~Ia from high-mass host being
+significantly depends on the host stellar mass, SNe~Ia from high-mass host being
 brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
@@
 for deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
-between the SN and the host remains unclear when using global properties such as
+between the SNe and their host remains unclear when using global properties such as
 the host stellar mass, which raises the question of the accuracy of the
@@
 more direct connections between the SN and its environments \citep{rigault2013},
-showing that local age-tracers such as the local specific star formation rate or
+showing that local age tracers such as the local specific star formation rate or
 the local color are more strongly correlated with the standardized SN magnitude
@@
 cosmology because the redshift drift magnitude correction to apply might
-strongly vary \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
-importance of local SN environmental study remains highly debated
-\cite[e.g.][]{jones2015,jones2019} and especially the impact it has on the
+strongly vary \yc{just the existence of a redshift-drift mag. correction is of importance here, even if it does not vary} \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
+importance of local SN environmental studies remains highly debated
+\cite[e.g.][]{jones2015,jones2019} and especially the impact it \yc{what does `it' refer to, the `importance'?} \nn{\textit{no, to the SN environment. Do you think it needs rephrasing?}} has on the
 derivation of H$_0$ \citep{jones2015,riess2016,riess2018,rose2019}.
@@
 having on average lower lightcurve stretches and being brighter after
-standardization. We use the correlation between the SN age probed with local
-specific star formation rate in \cite{rigault2018} and the SN stretch to model
+standardization. We use the correlation between the SN age, \yc{as} probed by local
+specific star formation rate \citep{rigault2018}, and the SN stretch to model
 the expected evolution of the underlying SN stretch distribution as a function
 of redshift. This modeling relies on three assumptions: (1) there are two
-populations of SNe~Ia; (2) the relative fraction of each of these populations as
-a function of redshift follows the age-modeling presented in \cite{rigault2018}
-and (3) the underlying distribution of stretch for each age-sample can be
-modeled and does not significantly drift. This paper aims at testing this with
+\yc{distinct} populations of SNe~Ia; (2) the relative fraction of each of these populations as
+a function of redshift follows the \yc{model} presented in \cite{rigault2018}
+and (3) the underlying distribution of stretch for each age sample \yc{is constant} \nn{\textit{what does it mean for a distribution to be constant?
+It seems weird that we say "does not significantly evolve" after saying it is constant, no?}} and does not significantly \yc{evolve with time}.
+This paper aims at testing this \yc{specific model} with
 datasets from the literature.
 
-The concept of the SNe~Ia age dichotomy arose with study of the the SN~Ia rate.
+\yc{Shouldn't this paragraph, introducing the concept of old and young populations and lsSFR, be moved \emph{before} the previous one?}
+The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate.
 \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
-the relative SNe~Ia rate in galaxies could only be explained if two populations
+the relative SNe~Ia rate in galaxies could be explained if two populations
 existed, one young, following the host star formation activity, and one old
-following the host stellar mass (the so called “prompt and delayed“ or “A+B“
+following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
 model). In \cite{rigault2018} we used the specific Star Formation Rate at the SN
-location (Local sSFR or LsSFR) to classify which are the younger (those with a
+location (Local sSFR or LsSFR \yc{this has already been used multiple times, define before}) to classify which are the younger (those with a
 high LsSFR) and which are the older (those with low LsSFR). Furthermore, since
 the first SNe~Ia host analyses, the SN stretch has been known to be strongly
-correlated with the SN host properties \citep{hamuy1996, hamuy2000}, which has
+correlated with the SN host properties \citep{hamuy1996, hamuy2000}, which \yc{what does this ``which'' refer to?} has
 been extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
     lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
-pan2014, kim19}. Following the “A+B“ model and the connection between SN stretch
+pan2014, kim19}. Following the ``A+B'' model and the connection between SN stretch
 and host properties, \cite{howell2007} first discussed the potential redshift
@@
 We present in section~\ref{sec:sample} the sample we are using for this
-analysis, which is based on the Pantheon dataset \citep{scolnic2018a}. We
+analysis, \yc{derived from} the Pantheon catalog \citep{scolnic2018a}. We
 discuss the importance of obtaining a ``complete'' sample, i.e. representative of
@@
 sample. We then present in section~\ref{sec:modeling} our modeling of the
-distribution of stretch as a function of redshift based models for the stretch
-distribution of young and old SNe~Ia. Our results are presented in
-section~\ref{sec:results}, where we test whether the SN stretch does evolve as a
-function of redshift and if the age model is in good agreement with this
+distribution of stretch as a function of redshift\nn{-}based models for the stretch
+distribution of young and old SNe~Ia \yc{what??? rephrase} \nn{\textit{what isn't clear? Does the hyphen help?}}. Our results are presented in
+section~\ref{sec:results}, where we test whether the SN stretch \yc{distribution} evolves as a
+function of redshift and if the \yc{aforementioned} age model is in good agreement with this
 evolution. Consequences for cosmology of our results are briefly discussed in
@@
 
-We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
+We base our analysis on the \yc{most recent comprehensive} SNe~Ia compilation, the Pantheon catalog from
 \cite{scolnic2018a}. A naive approach to test the SN stretch redshift drift
 would be to simply compare the observed SN stretch distributions in a few bins
-of redshift.  In practice, selection effects are affecting the observed SN
+of redshift.  In practice, \yc{however}, \yc{differential} selection effects are affecting the observed SN
 stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
@@
 magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
-Consequently, if selection effects are not accounted for, one might confuse true
-population drift with selection effects, and conversely.
+Consequently, if \yc{magnitude-related} selection effects are not accounted for, one might confuse true
+population drift with \yc{\sout{selection effects} survey properties}, and conversely.
 
@@
 and host redshift, the selection effects of magnitude-limited surveys should be
-negligible below a given redshift where even the faintest SNe~Ia can be observed.
+negligible below a given redshift \yc{at which} even the faintest SNe~Ia can be observed.
 In contrast, targeted surveys have highly complex selection functions and will
 be discarded from our analysis. Fortunately, modern SN cosmology samples such as
-the Pantheon dataset are now dominated by magnitude limited surveys.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
-    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
-        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples from
-        the Pantheon catalog.  The individual SNe are shown as blue dots and a
-        2D histogram is shown in gray to highlight point density. The ellipse
-        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the corresponding
-        standardized absolute magnitude using the $\alpha$ and $\beta$
-        coefficients from \cite{scolnic2018a}. The grey diagonal lines represent
-    the $(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and
-$z=1.70$ using SNLS's $m_{lim}$ of $24.8$~mag.}
-    \label{fig:maglim}
-\end{figure}
-
-We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
-distributions of SNe~Ia from the following surveys: PanStarrs
+the Pantheon \yc{one} are now dominated by magnitude-limited surveys.
+
+We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia from the following surveys: PanStarrs
 \citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
@@
 $\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is that with
-$(x_1=-1.66, c=0.25)$ and an absolute standardized magnitude at peak in Bessel-B
+$(x_1=-1.66, c=0.25)$ and an absolute standardized magnitude at peak in Bessel $B$
 band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect this object
 typically a week before and 10 days after peak to build a suitable lightcurve,
-the limiting standardized absolute magnitude effectively is approximately
+the \yc{effective} limiting standardized absolute magnitude is approximately
 $M_{lim} = -18.00$~mag. Hence, given the magnitude limit $m_{lim}$ of a
@@
 
+% YC: Figs and Tables *after* their 1st ref. in the text!
 \begin{figure}
     \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
-    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia  from
-        the SDSS, PS1 and SNLS dataset respectively (data from  Pantheon
-        \citealt{scolnic2018a}).  The colored parts represent the distribution
-        of SNe~Ia kept in our analysis for they are supposedly free from
-        selection bias (see section~\ref{sec:sample}). The dark (resp. light)
-    color responds to the conservative (resp. fiducial) selection cut.}
-    \label{fig:cuts}
+    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
+    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
+        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples of
+        the Pantheon catalog.  The individual SNe are shown as blue dots and a
+        2D histogram is shown in gray to highlight point density \yc{I don't think it's really needed}. The ellipse
+        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the corresponding
+        standardized absolute magnitude using the $\alpha$ and $\beta$
+        coefficients from \cite{scolnic2018a}. The grey diagonal lines represent
+    the $(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and
+$z=1.70$ using SNLS's $m_{lim}$ of $24.8$~mag.}
+    \label{fig:maglim}
 \end{figure}
 
-SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
-redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
-filter that has a $5\sigma$ depth of
+SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$; at these
+redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
+filter, that has a $5\sigma$ depth of
 24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
@@
 \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.  Fig.~14 of
-\citealt{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both will be
+\cite{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both \yc{limits} will be
 considered, as discussed below.
@@
 Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
-$5\sigma$ depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
-agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure could also
-suggest a more conservative $z_{lim}$ of 0.27.
-
-In similar redshift range, SDSS has a limiting magnitude of 22.5
+$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
+agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure \yc{is also suggestive of} a more conservative $z_{lim}=0.27$.
+
+In a similar redshift range, SDSS has a limiting magnitude of 22.5
 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
-SDSS surveys were more sensitive to limited spectroscopic resources.
-\cite{kessler2009} pointed out that during year-1 of SDSS, SNe~Ia with
-$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to a
+SDSS surveys were more sensitive to limited spectroscopic resources;
+\cite{kessler2009} pointed out that during \yc{first} year of SDSS, SNe~Ia with
+$r-mag<20.5$ were favored for spectroscopic follow-up, corresponding to a
 redshift cut at $0.15$. For the rest of the SDSS survey, additional
-spectroscopic resources were used, and \cite{kessler2009} and \cite{dilday2008}
-show a relative completeness up to $z_{lim}=0.2$. Following these analyses we
+spectroscopic resources were \yc{available}, and \cite{kessler2009} and \cite{dilday2008}
+show a relative completeness \yc{what is a \emph{relative completeness}???} up to $z_{lim}=0.2$. Following these analyses, we
 will use $z_{lim}=0.2$ as the baseline SDSS redshift limit.
@@
 For the rest of the analysis, we will consider the aforementioned redshift
-limits as well as more conservative cuts for each of these surveys, namely
-$z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS1 and $z_{lim}=0.55$ for SNLS.
+limits, as well as more conservative \yc{values} for each of these surveys, namely
+$z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS1 and $z_{lim}=0.55$ for SNLS. \yc{why do you remind the conservative values, and not the fiducial ones? Just refer to Table~1.}
 
 The complete sample selection is summarized in Table~\ref{tab:sample}, and the
-redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
+redshift distribution of these three surveys are \yc{is?} shown in Fig.~\ref{fig:cuts}.
 As expected, the selected redshift limits roughly correspond to the peak of
 these histograms.
-
-In addition, we use the SNe~Ia from the Nearby Supernova Factory
-\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
-been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
-4.2.2). Their spectroscopic follow-up was done over a redshift range of
-$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
-deeper. As such, these SNe~Ia should also be a random sampling of the underlying
-SN population. The SNfactory sample is particularly useful for studying SN property
-drift as it enables us to have a large SN~Ia sample at $z<0.1$.
-
-Finally, we included the HST sample from Pantheon, that follows the same logic
-of having a search deeper than the follow-up and we therefore kept it entirely
-\citep{strolger04}, see Table~\ref{tab:sample}. We present the stretch
-distribution and redshift histogram of these five surveys in
-Fig.~\ref{fig:sample}.
 
@@
     \begin{tabular}{l c c}
-    \hline\hline\\[-0.8em]
-        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
-        \hline\\[-0.8em]
-        SNf & -- & 114\\[0.30em]
-        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
-        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
-        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
-        HST & -- & 26\\[0.30em]
-        \hline\\[-0.8em]
-        Total & -- & 569(422)\\[0.30em]
+    \hline\hline%\\[-0.8em] % YC: don't, it's not your business to do that!
+        Survey & $z_{lim}$ & $N_{\mathrm{SN}}$ \\%[0.15em]
+        \hline%\\[-0.8em]
+        SNf & -- & 114\\%[0.30em]
+        SDSS & 0.20 (0.15) & 167 (82)\\%[0.30em]
+        PS1 & 0.30 (0.27) & 160 (122)\\%[0.30em]        
+        SNLS & 0.60 (0.55) & 102 (78)\\%[0.30em]
+        HST & -- & 26\\%[0.30em]
+        \hline%\\[-0.8em]
+        Total & -- & 569 (422)\\%[0.30em]
         \hline
@@
     \centering
+    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
+    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia from
+        the SDSS, PS1 and SNLS dataset respectively (data from Pantheon,
+        \citealt{scolnic2018a}).  The colored parts represent the distribution
+        of SNe~Ia kept in our analysis for they are supposedly free from
+        selection bias (see section~\ref{sec:sample}). The darker (resp. lighter)
+    color responds to the conservative (resp. fiducial) selection cut.}
+    \label{fig:cuts}
+\end{figure}
+
+In addition, we use the SNe~Ia from the Nearby Supernova Factory
+\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
+been discovered from non-targeted searches (114 SNe~Ia, see their sections~3 and~4.2.2; \mr{SNe~Ia time series are published in \citealt{Saunders2020}}). 
+The spectroscopic follow-up was done over a redshift range of
+$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the search was much
+deeper. As such, these SNe~Ia \yc{\sout{should also} are assumed to} be a random sampling of the underlying
+SN population. The SNfactory sample is particularly useful for studying SN property
+drift, as it enables us to have a large SN~Ia sample at $z<0.1$.
+
+Finally, we include the HST sample from Pantheon, that \yc{\sout{follows the same logic
+of having} similarly have} a search deeper than the follow-up and \yc{that} we therefore kept entirely
+\citep{strolger04}. %, see Table~\ref{tab:sample}.
+
+We present the stretch
+distribution and redshift histogram of these five surveys \yc{up to their respective $z_{lim}$} in
+Fig.~\ref{fig:sample}.
+
+\yc{You should also present/refer to Fig.5, showing the significant evolution of mean stretch as function of $z$ in complete samples. Next Sec. aims at modeling this evolution.}
+
+\begin{figure}
+    \centering
     \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb-cividis.pdf}
-    \caption{\textit{Lower}: \textsc{\texttt{SALT2.4}} lightcurve stretch as a
-        function of redshift (in log) for each survey considered in this
-        analysis (see legend).  Full (resp. empty) markers correspond to the
-        conservative (resp. fiducial) selection cuts. \textit{Top}: combined
+    \caption{\textit{Bottom:} \textsc{\texttt{SALT2.4}} lightcurve stretch as a
+        function of redshift for each survey considered in this
+        analysis (see legend).  Solid (resp. open) markers correspond to the
+        conservative (resp. fiducial) \yc{\sout{selection} redshift} cuts. \textit{Top:} combined
         redshift histogram in dark (resp. light) gray for the conservative
-    (resp. fiducial) selection cuts.}
+    (resp. fiducial) \yc{redshift} cuts.}
     \label{fig:sample}
@@
 
-In~\cite{rigault2018} we presented a model for the evolution of the fraction of
-younger and older SNe~Ia as a function of redshift following former work on
+\cite{rigault2018} presented a model for the evolution of the fraction of
+younger vs. older SNe~Ia as a function of redshift following former work on
 rates and delay time distributions \citep[e.g.,][]{mannucci2005,
 scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
-In short, we assumed that the number of younger SNe~Ia follows the star
-formation activity (SFR) in the Universe while the number of older SNe~Ia
+In short, it was assumed that the number of \yc{``young''} SNe~Ia follows the star
+formation activity (SFR \yc{R for activity? H for History?}) in the Universe, while the number of \yc{``old''} SNe~Ia
 follows the number of Gyr-old stars in the Universe, i.e. the stellar mass
@@
 fraction of young (resp. old) SNe~Ia in the Universe as a function of redshift,
-then their ratio $\delta/\psi$ is expected to follow the evolution of the
-specific star formation rate (SFR/M$^*$) which goes as $(1+z)^{2.8}$ until
+then the ratio $\delta/\psi$ is expected to follow the evolution of the
+specific star formation rate (SFR/M$^*$), which goes as $(1+z)^{2.8}$ until
 $z\sim2$ \citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
 \citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
-\citep{mannucci2006,rodney2014}, we concluded in \cite{rigault2018} that
+\citep{mannucci2006,rodney2014}, \cite{rigault2018} concluded that
 \begin{equation}
@@
 
-\subsection{“Base” underlying stretch distribution}
+\subsection{``Base'' underlying stretch distribution}
 \label{sec:basemodel}
+
+\yc{There is a generic confusion between ``SN stretch'' (individual SN, as in Eq.3/4) and something like ``SN stretch distribution'' (population, as in Eq.2). This is confusing...}
+
+To model the evolution of the SN stretch \yc{distribution} as a function of redshift, given our
+aforementioned model of the evolution of the fraction of younger and older
+SNe~Ia with \yc{\sout{redshift} cosmic time}, we need to model the SN stretch distribution for each age
+subsample. \yc{That paragraph sounds like a tautology...}
+
+\yc{I suggest we remove the use of ``we'' when referring to previous papers, like ``In R+18, we did...'' to be replaced by ``R+18 did...''}
+
+\cite{rigault2018} presented the relation between SN stretch and LsSFR
+measurement, \yc{a progenitor age tracer}, using the SNfactory sample. This
+relation is shown in Fig.~\ref{fig:stretchlssfr} \yc{for SNf SNe used in the current analysis}. Given the structure of the
+stretch-LsSFR scatter plot, our model of the underlying SN~Ia stretch
+distribution is defined as follows:
+\begin{itemize}
+\item for the younger population (i.e., $\log(\mathrm{LsSFR})>-10.82$), the
+\yc{\sout{underlying}} stretch distribution is modeled as a \yc{single} normal distribution
+$\mathcal{N}(\mu_1, \sigma_1^2)$; 
+\item the older population (i.e., $\log( \mathrm{LsSFR})<-10.82$) stretch
+distribution is modeled as a \yc{bimodal} Gaussian mixture $a\times \mathcal{N}(\mu_1, \sigma_1^2) +
+(1-a)\times \mathcal{N}(\mu_2, \sigma_2^2)$, 
+where one mode is the same as for the young population, $a$
+representing the relative influence of the two \yc{modes}.
+\end{itemize}
 
@@
         $p_y$, for the SNe~Ia to be young, i.e. to have $\log\mathrm{LsSFR} \geq
-        -10.82$; see \cite{rigault2018}. \textit{Right}: histogram of
-        $p_y$-weighted distribution of the SN stretchs. On both panels, we
-        illustrate the Base model fitted on the SNfactory SNe. The younger and
-    older population contributions are shown in purple and yellow,
-respectively.}
+        -10.82$ \citep[see][]{rigault2018}. \textit{Right}: $p_y$-weighted histogram of the SN stretches, \yc{as well as the adjusted Base model}; the younger and
+    older population contributions are shown in purple and yellow, respectively.}
     \label{fig:stretchlssfr}
@@
 
-To model the evolution of the SN stretch as a function of redshift, given our
-aforementioned model of the evolution of the fraction of younger and older
-SNe~Ia with redshift, we need to model the SN stretch distribution for each age
-subsample.
-
-In~\cite{rigault2018}, we presented the relation between SNe stretches and LsSFR
-measurements, which trace progenitor age, using the SNfactory sample. This
-relation is shown in Fig.~\ref{fig:stretchlssfr}, and given the structure of the
-stretch-LsSFR scatter plot, our model of the underlying SN~Ia stretch
-distribution is defined as follows:
-\begin{itemize}
-    \item for the younger (i.e., $\log( \mathrm{LsSFR})>-10.82$) population, the
-        underlying stretch distribution is modeled as a normal distribution
-        $\mathcal{N}(\mu_1, \sigma_1^2)$; 
-    \item the old (i.e., $\log( \mathrm{LsSFR})<-10.82$) population stretch
-        distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) +
-        (1-a)\times \mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a Gaussian mixture
-        where one mode is the same as for the young population, $a$
-        representing the relative influence of the two Gaussians.
-\end{itemize}
-When combining our prediction of the evolution of younger SNe~Ia as a function
-of redshift (Eq.~\ref{eq:delta}), and the stretch distribution of both younger
-and older SNe~Ia, our model for the underlying distribution of stretch,
-$X_1(z)$, as a function of redshift is:
-\begin{align}
+When combined to the previously described prediction of the evolution of younger SNe~Ia as a function
+of redshift (Eq.~\ref{eq:delta}), our model $X_1(z)$ for the underlying stretch distribution as a function of redshift is written
+\begin{multline}
 \label{eq:stretchz}
-    X_1(z) = {}
-    \delta(z) &\times \mathcal{N}(\mu_1,\sigma_1^2) + {}\nonumber\\
-    (1 - \delta(z)) &\times \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
+X_1(z) = 
+\delta(z) \times \mathcal{N}(\mu_1,\sigma_1^2) + {} \\
+(1 - \delta(z)) \times \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
 (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right].
-\end{align}
-
-The maximum-likelihood estimate of the 5 free parameters
-$\theta\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained by
-minimizing a pseudo-$\chi^2$ $\mathcal{X}^2$ defined as:
+\end{multline}
+\yc{In fact, not clear what this eq. does here: it is supposed to be superseded by Eq.3/4. Maybe it should be moved to Sec.3.1.2 when we use $\delta(z)$ as a proxy to $y^i$.}
+
+The maximum-likelihood estimate of the 5~free parameters
+$\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained by
+minimizing the \yc{following} pseudo-$\chi^2$: %$\mathcal{X}^2$ defined as:
 \begin{equation}
     \label{eq:likelihood}
-    \mathcal{X}^2 = -2 \sum_i \ln \prob{x^{i}_{1}}{\theta;
+    \mathcal{X}^2 = -2 \sum_i \ln \prob{x^{i}_{1}}{\vec{\theta};
     \mathrm{d}x^{i}_{1}, y^i},
@@
     \label{eq:likelihoodsnf}
-    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, y^i} =
+    \prob{x^{i}_{1}}{\vec{\theta}; \mathrm{d}x^{i}_{1}, y^i} =
     y^i &\times \mathcal{N}\left(x^{i}_{1} \mid \mu_1, \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
@@
      &(1-a) \times \mathcal{N}\left(x^{i}_{1} \mid \mu_2,
-     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) \bigg]
+     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) \bigg],
 \end{align}
@@
 $a$ is constrained between 0 and 1, we fit for $\alpha$ such that
-$a=\arctan(\alpha)/\pi+0.5$, which results in an asymmetric error on $a$.
+$a=\arctan(\alpha)/\pi+0.5$ \yc{\sout{, which results in an asymmetric error on $a$}}.
+\yc{Given the fact that 1.
+it's barely asymmetric ($a\sim 0.5$), 2.
+no one cares, it should be symmetrized (quadratically).
+Probably $\sigma$ errors are not symmetric either!}
 
@@
 have $\log \mathrm{LsSFR} \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
-minimize Eq.~\ref{eq:likelihood} with respect to~$\theta$.  Results on fitting
-the SNfactory SNe with this model are shown Table~\ref{tab:modelresults} and
+minimize Eq.~\ref{eq:likelihood} with respect to~$\vec{\theta}$.  Results on fitting
+the SNf SNe with this model are shown Table~\ref{tab:modelresults} and
 illustrated in Fig.~\ref{fig:stretchlssfr}.
@@
     \label{tab:modelresults}
-    \begin{tabular}{l c c c c c}
-    \hline\hline\\[-0.8em]
-        Sample & $\mu_1$  & $\sigma_1$ &$\mu_2$  & $\sigma_2$ & a \\[0.15em]
-        \hline\\[-0.8em]
-        SNfactory & $0.41 \pm 0.08$  & $0.55 \pm 0.06$ & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$ & $0.48^{+0.08}_{-0.08}$ \\[0.15em]
-        Fiducial & $0.37 \pm 0.05$  & $0.61 \pm 0.04$ & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$ & $0.51^{+0.09}_{-0.10}$ \\[0.15em]
-        Conservative & $0.38 \pm 0.05$  & $0.60 \pm 0.04$ & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\[0.15em]
+    \begin{tabular}{lccccc}
+    \hline\hline%\\[-0.8em]
+        Sample & $\mu_1$  & $\sigma_1$ &$\mu_2$  & $\sigma_2$ & a \\%[0.15em]
+        \hline%\\[-0.8em]
+        SNfactory & $0.41 \pm 0.08$  & $0.55 \pm 0.06$ & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$ & $0.48^{+0.08}_{-0.08}$ \\%[0.15em]
+        Fiducial & $0.37 \pm 0.05$  & $0.61 \pm 0.04$ & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$ & $0.51^{+0.09}_{-0.10}$ \\%[0.15em]
+        Conservative & $0.38 \pm 0.05$  & $0.60 \pm 0.04$ & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\%[0.15em]
         \hline
@@
 probability of a SN to be young. This still corresponds to minimizing
-Eq.~\ref{eq:likelihood} with respect to the parameters $\theta\equiv(\mu_1,
+Eq.~\ref{eq:likelihood} with respect to the parameters $\vec{\theta}\equiv(\mu_1,
 \mu_2, \sigma_1, \sigma_2, a)$ of the stretch distribution $X_1$
@@
 (resp. conservative) sample are given Table~\ref{tab:modelresults} and
-illustrated in Fig.~\ref{fig:modelall}. We see in this figure that the measured
+\yc{predicted redshift evolution of mean stretch} illustrated in Fig.~\ref{fig:modelall}.
+\yc{I'm confused: grey points in Fig.5 are model independent, aren't they?
+Only the curve is related to the model.} We see in this figure that the measured
 mean SN~Ia stretch per bin of redshift closely follows our redshift drift
@@
 higher redshift have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$)
-than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$). This is indeed
+than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$). \yc{This is model-independant, and a nice result by itself, and should be already mentioned right at the end of Sec.2.}  This is indeed
 what is expected if old environments favor low SN stretches
@@
 distribution following \cite{rigault2018}, i.e.\ as a single Gaussian for the
-young SNe~Ia and a mixture of two Gaussians for the old SNe Ia population, one
-the same as for the young-population plus another for the fast declining SNe~Ia
-that seem to only exist in old host galaxy stellar populations. This is our
-so-called ``Base'' model.
-
+``young'' SNe~Ia and a mixture of two Gaussians for the ``old'' SNe Ia population, one
+\yc{being} the same as for the young population, plus another one for the fast-declining SNe~Ia
+that seem to only exist in old \yc{\sout{host galaxy stellar populations} local environments}.
+This is our so-called ``Base'' model.
 However, to test different modeling choices, we have implemented a suite of
@@
 
-\cite{howell2007} used a simpler model per age bin, assuming a normal
-distribution for both the young and the old population. We thus consider a
+\cite{howell2007} used a simpler \yc{unimodal} model per age category, assuming a \yc{single} normal
+distribution for \yc{each of} the young and old populations. We thus consider a
 ``Howell+drift'' model, with one single Gaussian per age group and the
-$\delta(z)$  drift from \cite[][see Eq.~\ref{eq:delta}]{rigault2018}.
-
-Alternatively, since we aim at probing the existence of a redshift drift, we
-also test non-drifting models by restricting the ``Base'' and ``Howell'' models
-to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe;
-these models are labeled ``Base constant'' and ``Howell constant''.
-
-We also consider other non-drifting models, notably the one developed for Beams
+$\delta(z)$ drift from Eq.~\ref{eq:delta}.
+
+Alternatively, since we aim at probing the existence of an \yc{evolution with redshift}, we
+also test \yc{constant} models by restricting the ``Base'' and ``Howell'' models
+to use a supposedly redshift-independent fraction $\delta(z) \equiv f$ of young SNe;
+these models are \yc{hereafter} labeled ``Base+constant'' and ``Howell+constant''.
+
+We also consider another \yc{intrinsically} non-drifting model, the one developed for Beams
 with Bias Correction \cite[BBC,][]{scolnic2016, kessler2017}, used in recent SN
@@
 behind this sample-based approach is twofold: (1) Malmquist biases are driven by
-survey properties and (2) because current surveys cover limited redshift ranges,
-doing so covers some potential redshift evolution information
+survey properties and (2) \yc{since} current surveys cover limited redshift ranges,
+doing so \yc{what?} covers some potential redshift evolution information
 \citep{scolnic2016, scolnic2018a}. See further discussion concerning BBC in
@@
 
-Finally, for the sake of completeness, we also consider simple ``Gaussian'' and
-``asymmetric'' Gaussian non-drifting models.
+Finally, for the sake of completeness, we also consider simple the ``Gaussian'' and
+``Asymmetric'' Gaussian non-drifting models.
 
@@
 
-We applied each model on both the reference and conservative samples (cf.
-section~\ref{sec:sample}). Because the models presented in
-section~\ref{sec:modeling} have different degrees of freedom, we use the Akaike
-Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare them. This
-estimator penalizes extra degrees of freedom to avoid over-fitting the data. It
+We \yc{adjusted} each of the models described above on both the reference and conservative samples (cf.
+section~\ref{sec:sample}); \yc{results are gathered in Table~\ref{tab:comp}, and illustrated in Fig.~\ref{fig:mod_comp}.}
+
+\yc{Since} the \yc{various} models have different degrees of freedom, we use the Akaike
+Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare \yc{their ability to properly describe the observations}. This
+estimator penalizes extra degrees of freedom to avoid over-fitting the data, and
 is defined as follow:
@@
 where $\mathcal{X}^2_{min}$ is the minimum value of the pseudo-$\chi^2$ as
-defined Eq.~\eqref{eq:likelihood} and $k$ is the number of free parameters. The
-reference model is the one with the smallest AIC. In comparison to this model,
-those with $\Delta\mathrm{AIC}<5$ are acceptable, those with
-$5<\Delta\mathrm{AIC}<10$ are unfavored and those with $\Delta\mathrm{AIC}>20$
-are excluded. This roughly corresponds to 2, 3 and 5 $\sigma$ limits for a
+defined Eq.~\eqref{eq:likelihood}, and $k$ is the number of free parameters \yc{to be adjusted}. The
+reference model is the one with the smallest AIC; in comparison to this model,
+the models with $\Delta\mathrm{AIC}<5$ are \yc{coined} acceptable, the ones with
+$5<\Delta\mathrm{AIC}<20$ are unfavored, and those with $\Delta\mathrm{AIC}>20$
+are \yc{deemed} excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for a
 Gaussian probability distribution.
@@
     \label{tab:comp}
-    \begin{tabular}{c c c  |c c c |  c c c }\hline\hline\\[-0.8em]
-        & & & \multicolumn{3}{|c}{Fiducial sample (569 SNe)} & \multicolumn{3}{|c}{Conservative sample (422 SNe)} \\
+    \begin{tabular}{ccc|ccc|ccc}
+    \hline\hline%\\[-0.8em]
+        & & & \multicolumn{3}{c}{Fiducial sample (569 SNe)} & \multicolumn{3}{|c}{Conservative sample (422 SNe)} \\
         %\cline{2-6}
-
         Name & drift & $k$ &
-        $\mathcal{X}^2$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ &  $\mathcal{X}^2$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$\\[0.15em]
-        \hline\\[-0.8em]
+        $\mathcal{X}^2$ & AIC & $\Delta \mathrm{AIC}$ & $\mathcal{X}^2$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$\\%[0.15em]
+        \hline%\\[-0.8em]
 
@@
         & 1456.7 & 1466.7 & -- 
-        & 1079.5 & 1089.5 & -- \\[0.15em]
+        & 1079.5 & 1089.5 & -- \\%[0.15em]
 
@@
         %& $3.4\times10^{-2}$
-        \\[0.15em]
+        \\%[0.15em]
 
@@
         %& $1.4\times10^{-4}$
-        \\[0.15em]
-
-        Howell:const & $f$ & 5
+        \\%[0.15em]
+
+        Howell+const & $f$ & 5
         & 1484.2 & 1494.2 & $-27.5$  %&$1.0\times10^{-6}$
@@
         %& $1.9\times10^{-5}$
-        \\[0.15em]
-
-        Base:const & $f$ & 6
+        \\%[0.15em]
+
+        Base+const & $f$ & 6
         & 1484.2 & 1496.2 & $-29.5$  %&$3.9\times10^{-7}$
@@
         %& $7.1\times10^{-6}$
-        \\[0.15em]
-
-        BBC & per sample & 3x5
+        \\%[0.15em]
+
+        \yc{BBC-like?} & per sample & 3x5
         & 1468.2 & 1498.2 & $-31.5$  %&$1.5\times10^{-7}$
@@
         %& $5.7\times10^{-6}$
-        \\[0.15em]
+        \\%[0.15em]
 
@@
         %& $4.0\times10^{-13}$ 
-        \\\hline\hline
+        \\
+        \hline
     \end{tabular}
@@
         models with and without redshift drift, respectively. Light markers show
-        the results when this analysis is performed on the conservative sample
+        the results when the analysis is performed on the conservative sample
         rather than the fiducial one.  Color-bands illustrate the validity of
         the models, from  Acceptable ($\Delta\mathrm{AIC} > -5$) to Excluded
-        ($\Delta\mathrm{AIC} < -20$); see text. According to the AIC, all
-    non-drifting models (open symbols) are excluded to be as good representation
-of the data in comparison to our Base (drifting) model.}
+        ($\Delta\mathrm{AIC} < -20$), see text. According to the AIC, all
+    non-drifting models (open symbols) are excluded to be as good a representation
+of the data as the Base (drifting) model.}
     \label{fig:mod_comp}
@@
 distribution and its redshift evolution. This is therefore the reference model,
-both on the fiducial and conservative samples.
-
-Furthermore, we find that underlying stretch distributions not including a
-redshift drift are all excluded as description of the data relative to the Base
-model. In fact, the best non-drifting model (Asymmetric) has a very marginal
+both on the fiducial \yc{somewhere, you still use the ``ref. sample'' terminology!} and conservative samples.
+
+Furthermore, we find that \yc{\sout{underlying stretch distributions not including a
+redshift drift} redshift-independent stretch distributions} are all excluded as description of the data relative to the Base
+model. In fact, the best non-drifting model (the Asymmetric one) has a very marginal
 chance ($p \equiv \exp\left(\Delta\mathrm{AIC}/2\right) = 5\times10^{-6}$) to
@@
 assessment of qualitative facts clearly visible in Fig.~\ref{fig:modelall}: the
-mean SN stretch per bin of redshift strongly suggests redshift evolution
-rather than a constant value, and this evolution is compatible with the
-redshift drift expected from Eq.~\ref{eq:delta}.
-
-For the sake of robustness, we also tested allowing the high-stretch mode of the
-old-population to differ from the young-population mode, hence adding two
-degrees of freedom. The corresponding best fit is not significantly better
-resulting in a $\Delta$AIC of $-0.4$, strengthening the fact that the young and
-old population indeed share the same underlying high-stretch mode. We also
+mean SN stretch per bin of redshift strongly suggests \yc{a significant} redshift evolution
+rather than a constant value, and this evolution is \yc{\sout{compatible with the
+redshift drift expected from} well described by}  Eq.~\ref{eq:delta}.
+
+\mr{Surprisingly, the sample-based Gaussian asymmetric
+modeling used by current implementations of the BBC technique \citep{scolnic2016, kessler2017}, is one of the worst modelisations of our analysis (see section~\ref{sec:results}). While its
+pseudo-$\chi^{2}$ is the smallest of all \yc{redshift-independent} models (but still $\Delta\mathcal{X}^2 = -11.5$ worse than the reference Base model), it is
+\yc{strongly} penalized for requiring 15~free parameters (\yc{$\mu_0, \sigma_{\pm}$}~for each of the 5~samples of
+the analysis). 
+We report in Table~\ref{tab:bbc} the samples' $\mu_0$ and $\sigma_{\pm}$ adjusted
+on the \sout{supposedly} selection-free samples using our fiducial cuts (see
+section~\ref{sec:sample}. We find our results in close agreement with
+\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
+derived these model parameters using the full BBC formalism, using numerous
+simulations to model the selection effects (see details e.g., section~3 of
+\citealt{kessler2017}). The agreement between our fit of the asymmetric Gaussians on the \yc{supposedly} selection-free part of the samples and the results
+derived using the BBC formalism supports our approach to get a sample with negligible selection effects.} \yc{It's not very clear what you mean to convey in this sentence.}
+
+We further discuss the consequence of this result for cosmology in section~\ref{sec:discussion}.
+
+\begin{table}
+    \centering
+    \caption{Best-fit parameters for our sample-based asymmetric modeling of the
+    underlying stretch distribution.}
+    \label{tab:bbc}
+    \begin{tabular}{ccccccc}
+    \hline\hline%\\[-0.8em]
+    Asymmetric & $\sigma_{-}$    & $\sigma_{+}$    & $\mu_0$ \\
+    \hline%\\[-0.8em]
+    SNfactory  & 1.34 $\pm$ 0.13 & 0.41 $\pm$ 0.10 & 0.68 $\pm$ 0.15 \\%[0.15em]
+    SDSS       & 1.31 $\pm$ 0.11 & 0.42 $\pm$ 0.09 & 0.72 $\pm$ 0.13 \\%[0.15em]
+    PS1        & 1.01 $\pm$ 0.11 & 0.52 $\pm$ 0.12 & 0.38 $\pm$ 0.16 \\%[0.15em]
+    SNLS       & 1.41 $\pm$ 0.13 & 0.15 $\pm$ 0.13 & 1.22 $\pm$ 0.15 \\%[0.15em]
+    HST        & 0.76 $\pm$ 0.36 & 0.79 $\pm$ 0.35 & 0.11 $\pm$ 0.44 \\
+    \hline
+    \end{tabular}
+\end{table}
+
+For the sake of robustness \yc{???}, we also \yc{performed tests} allowing the high-stretch mode of the
+old population to differ from the young population mode, hence adding two
+degrees of freedom. The corresponding fit is not significantly better,
+with a $\Delta$AIC of $-0.4$; this strengthens the fact that the young and
+old populations indeed \yc{appear to} share the same underlying high-stretch mode. We also
 allowed the young population to have a low-stretch mode, finding its amplitude
-to be compatible with 0 ($<2\%$).
-
-\section{Discussion}\label{sec:discussion}
-
-To the best of our knowledge, the SN~Ia stretch redshift drift has never been
-explicitly accounted for in cosmological analyses. Not doing so is a second
-order issue for SN cosmology as it only affects the way one does Malmquist bias
-correction.  Indeed, as long as the Phillip's relation standardization parameter
-$\alpha$ is not redshift dependent (a study behind the scope of this paper, but
-see e.g. \citealt{scolnic2018a}), if a sample is free from selection effects,
-the stretch-corrected SNe~Ia magnitudes used for cosmology are blind to the
-underlying stretch distribution. Yet, modern surveys do have significant
-Malmquist bias at least for the upper half of their SN redshift distribution.
-Then, an ill-modeling of the underlying stretch distribution will bias the
-derived SN magnitude for these \citep[see e.g.,][]{rubin2015,rubin2016}. The
-fact that the currently-used stretch modeling for Malmquist bias correction in
-SN cosmology (BBC) is excluded to be an as good description of the data then our
-drifting model is worrisome and we further discussed in the next sections.
+to be compatible with~0 ($<2\% \pm \yc{???}$).
+
+\section{Discussion}
+\label{sec:discussion}
+
+To the best of our knowledge, the SN~Ia stretch redshift drift has never been explicitly accounted for in cosmological analyses, \mr{though bayesian hierarchy formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016} or Steve \citep{hinton2019} can easily allow it}.
+Not doing
+so is a second order issue for SN cosmology, as it only affects the way one does \yc{account for}
+Malmquist bias. \yc{I don't understand why you say it's a 2nd order effect if all samples are significantly Malmquist biased!?} Indeed, as long as Phillip's relation \yc{ref???}
+standardization parameter $\alpha$ is not \yc{intrinsically} redshift dependent (a study behind the
+scope of this paper, but see e.g. \citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
+cosmology are blind to the underlying stretch distribution \yc{\emph{for complete samples}}. \yc{However}, all modern surveys
+do have significant Malmquist bias at least for the upper half of their SN
+redshift distribution.  \yc{As a consequence}, an ill-modeling of the underlying stretch
+distribution will bias the derived SN magnitude for these \yc{these what?} \citep[see
+e.g.,][]{rubin2015,rubin2016}. 
+The fact that the currently-used \mr{sample-based asymmetric Gaussian distribution modeling} is excluded to be an as good description of the data then our drifting model is worrisome;  we further discuss this in the next sections.
+\yc{I don't see the point of this kind of sentence...}
 
 \subsection{BBC formalism and underlying stretch distributions}\label{sec:bbc}
+
+\yc{IMHO, you should take care of not converting this paper into an anti-BBC paper: just show the results, and let people make their opinion.
+Furthermore, I find this paragraph mostly a repetition of previous statements and results...}
 
@@
 brout2019, descosmopaper2019} including the direct measurement of $H_0$
-\citep{riess2016,riess2019}.
-
-Surprisingly, the BBC technique, which assumes an individual asymmetric
-distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
-modelisations in our analysis (see section~\ref{sec:results}.) While its
-pseudo-$\chi^{2}$, $\mathcal{X}^2$, is the smallest of all non-drift models (but
-still $\Delta \mathcal{X}^2=-11.5$ worse than the reference Base model), it is
-also penalized for requiring 15 free parameters (3 for each of the 5 samples of
-the analysis). \citet[][section~2]{scolnic2016} and
-\citet[][section~5.4]{scolnic2018a} highlighted that, because traditional
-surveys span limited redshift ranges, the per sample approach can somewhat
-account for implicit redshift drifts. We stress here that, as measurements of
-modern surveys cover increasingly larger redshift ranges to reduce calibration
-systematic uncertainties, this approach is becoming less valid, notably for PS1,
+\citep{riess2016,riess2019}. \yc{This all have already been said before, there's no point at repeating it IMHO. Maybe just a simpler introductory sentence without all the references?}
+
+\mr{Yet, the current underlying stretch modeling of the BBC formalism, a sample-based asymmetric Gaussian distribution, is excluded to be an as good representation of the data than our Base drifting modeling, see section~\ref{sec:results}.
+\yc{Again, unnecessary repetition.} Then, unlike what  \citet[][section~2]{scolnic2016} and
+\citet[][section~5.4]{scolnic2018a} suggested, the fact that traditional surveys span limited redshift ranges, the per-sample approach cannot correctly account for implicit redshift drifts.} \yc{I don't understand what they suggested and what is wrong...
+Rephrase.} We stress here that, as measurements of
+modern surveys \yc{try to} cover increasingly larger redshift ranges \yc{in order} to reduce calibration
+systematic uncertainties, this approach \yc{which one?} is becoming less valid, notably for PS1,
 DES and, soon, LSST. Already with current surveys, the BBC sample-based
@@
 $\Delta\mathrm{AIC}<-20$, which could be interpreted as a probability $p=2\times
-10^{-7}$ of being an as good representation of the data as the Base model. Note
-that even considering the SNfactory dataset like any other (i.e.  ignoring the
-LsSFR information see section~\ref{sec:modeling}), reduces this difference to
+10^{-7}$ of being an as good representation of the data as the Base model. \yc{Again, unnecessary repetition.} Note
+that even considering the SNfactory dataset like any other (i.e., ignoring the
+LsSFR measurements, see section~\ref{sec:modeling}), reduces this difference to
 $\Delta\mathrm{AIC}<-10$, still strongly unfavoring the modeling currently used
-as part of the BBC Malmquist bias correction.
-
-We report in Table~\ref{tab:bbc} the samples' $\sigma_{\pm}$ and $\mu_0$ fitted
-on the supposedly selection-free samples using our fiducial cuts (see
-section~\ref{sec:sample}. We find our results in close agreement with
-\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
-derived these model parameters using the full BBC formalism, using numerous
-simulations to model the selection effects (see details e.g., section 3 of
-\citealt{kessler2017}). The agreement between our fit of the asymmetric
-Gaussians on the supposedly selection-free part of the samples and the results
-derived using the BBC formalism supports our approach to get a sample with
-negligible selection effects. If we were to use \cite{scolnic2016} and
-\cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively, the
+as part of the \mr{sample-based formalism}.
+If we were to use \cite{scolnic2016} and
+\cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively \yc{which parameters?}, the
 $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC modeling would
 go even deeper from $-32$ to $-47$.
-
-\begin{table}
-    \centering
-    \caption{Best-fit parameters for our sample-based asymmetric modeling of the
-    underlying stretch distribution.}
-    \label{tab:bbc}
-    \begin{tabular}{c c c c c c c}\hline\hline\\[-0.8em]
-    Asymmetric & $\sigma_{-}$    & $\sigma_{+}$    & $\mu_0$ \\\hline\\[-0.8em]
-    SNfactory  & 1.34 $\pm$ 0.13 & 0.41 $\pm$ 0.10 & 0.68 $\pm$ 0.15 \\[0.15em]
-    SDSS       & 1.31 $\pm$ 0.11 & 0.42 $\pm$ 0.09 & 0.72 $\pm$ 0.13 \\[0.15em]
-    PS1        & 1.01 $\pm$ 0.11 & 0.52 $\pm$ 0.12 & 0.38 $\pm$ 0.16 \\[0.15em]
-    SNLS       & 1.41 $\pm$ 0.13 & 0.15 $\pm$ 0.13 & 1.22 $\pm$ 0.15 \\[0.15em]
-    HST        & 0.76 $\pm$ 0.36 & 0.79 $\pm$ 0.35 & 0.11 $\pm$ 0.44 \\\hline
-    \hline
-    \end{tabular}
-\end{table}
+\yc{But you would be using parameters derived specifically for incomplete samples on specially-constructed complete sample: what is the validity of the comparison???}
 
@@
         ($x_1$) after the fiducial redshift limit cut (grey histogram). This
-        distribution is supposed to be a random draw from the underlying
-        stretch distribution. The green line shows the BBC model of this
-        underlying distribution (asymetric Gaussian). The full line (band) is
-        our best fit (its error); the dashed line shows \cite{scolnic2018a}
-        result. The black line (band) show our best fitted Base-modeling (its
-        error, see Table XX) that includes redshift drift. For illustration, we
-        show as colored (from blue to red with increasing redshifts) the
-    evolution of the underlying stretch distribution as a function of redshift
-for the redshift range covered by PS1 data.}
+        distribution is supposed to be a random draw from the underlying stretch
+        distribution. The green line shows the BBC model of this underlying
+        distribution (asymmetric Gaussian). The full line (band) is our best fit
+        (its error); the dashed line shows \cite{scolnic2018a} result. The black
+        line (band) show our best fitted Base-modeling (its error, see Table
+        \ref{tab:modelresults}) that includes redshift drift. For illustration,
+        we show as colored (from blue to red with increasing redshifts) the
+        evolution of the underlying stretch distribution as a function of
+        redshift for the redshift range covered by PS1 data.}
     \label{fig:bbc_pdf_ps1}
@@
 model for the PS1 sample. Our model is bimodal and the relative amplitude of
-each mode depends on the redshift-dependent fraction of old SNe~Ia in the
-sample: the higher the fraction of old SNe~Ia (lower redshift), the higher the
-amplitude of the low-stretch mode. This redshift dependency is shown as blue to
+each mode depends on the redshift-dependent fraction of old \yc{vs.\ young} SNe~Ia in the
+sample: the higher the fraction of old SNe~Ia (at lower redshift), the higher the
+amplitude of the \yc{old-specific} low-stretch mode. This redshift dependency is shown as blue to
 red underlying distributions in Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges
 covered by PS1. The observed $x_1$ histogram corresponds to the sum of
-individual underlying SN-redshift distributions. As expected, the two approaches
+individual underlying SN-redshift distributions. \yc{Not clear: you mean the black curve -- a model -- or the histogram -- an observation?} As expected, the two approaches \yc{which ones?}
 strongly differ in modeling the negative part of the SN stretch distribution.
@@
 under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting
-model for typical PS1 SN redshifts. This means that the bias-correction of
-individual SN stretch, $\overline{\delta}_{x_1}$, \citep{kessler2017} and
+model for typical PS1 SN redshifts. \yc{This is definitely not obvious from the Fig. solely...} This means that the bias-correction of
+individual SN stretches, $\overline{\delta}_{x_1}$ \citep{kessler2017}, and
 consequently the magnitude bias correction, $\mu_B$, is most likely inaccurate
-and this inaccuracy is redshift dependent .
-
-The amplitude of this bias for cosmology is beyond the scope of this paper given
+and \sout{this inaccuracy is} redshift dependent.
+
+The amplitude of this \yc{magnitude} bias for cosmology is beyond the scope of this paper given
 the complexity of the BBC analysis. It would require a full study using our Base
-model given Eq.~\ref{eq:stretchz} in place of the sample-based asymmetric
+model (Eq.~\ref{eq:stretchz}) in place of the sample-based asymmetric
 modeling as part of the BBC simulations.
 
-Nonetheless, it is interesting to study the  expected mean stretch difference as
+Nonetheless, it is \yc{enlightening} to study the  expected mean stretch difference as
 a function of redshift when considering that the underlying SN stretch
-distribution follows either the BBC formalism or our Base drifting modeling.
+distribution follows either the BBC formalism or our Base drifting model.
 This is shown Fig.~\ref{fig:magdrift}, where, for clarity we convert this
-difference in $x_1$ to a difference in magnitude using $\alpha=0.156$
+difference in \yc{mean??? stretch} $x_1$ to a difference in magnitude using $\alpha=0.156$
 \citep{scolnic2018a}.
-
-We see in this figure that the amplitude of modeling difference is a few tens of
-millimag, which corresponds to the expected bias affecting SNe~Ia impacted
-by sample selection effects. For comparison, the amplitude of this effect is
-similar in scale to the effect of exotic forms of dark energy could
-have on $\mu(z)$. Consequently, in the era of modern cosmology, where we aim at
-probing $w_0$ at a sub-percent and $w_a$ at a few percent
-\citep[e.g.,][]{lsstpaper}, it is therefore of paramount importance to further
-understand the exact modeling of the lightcurve parameters when using SNe~Ia
-affected by Malmquist bias.
 
@@
 
+We see that the amplitude of the modeling difference is a few tens of
+millimag, which corresponds to the expected bias affecting SNe~Ia impacted
+by sample selection effects. For comparison, the amplitude of this effect is
+similar in scale to the effect exotic forms of dark energy could
+have on $\mu(z)$. Consequently, in the era of modern cosmology, where we aim at
+probing $w_0$ at a sub-percent \yc{level} and $w_a$ at the ten-percent \yc{precision}
+\citep[e.g.,][]{lsstpaper}, it is of paramount importance to further
+understand the exact modeling of the lightcurve parameters when using SNe~Ia
+affected by Malmquist bias.
+
 \section{Conclusion}
@@
 
-We have presented a study of the drift of the underlying SNe~Ia stretch
-distribution as a function of redshift. We used SNe~Ia from magnitude-limited
-surveys from the Pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
-well as HST, to which we added SNfactory data from \cite{rigault2018} for the
-low-redshift bin. We only considered the supernovae that have been discovered
-in redshift ranges of each surveys where selection effects are negligible. This
-way the observed SNe~Ia stretches are random sampling of the true underlying
-distribution. This corresponds to 569 SNe~Ia (422 when considering more
-conservative cuts).
-
-Following observations made in \cite{rigault2018}, we introduced a redshift
-drift modeling which depends on the expected fraction of young and old SNe~Ia as
-a function of redshift, each age-population having its own underlying stretch
-distribution. We have studied various suites of distributions, including
-non-drifting modeling. We also studied the stretch distribution modeling of the
-current version of the BBC Malmquist bias correction algorithm used by recent
-SN~Ia cosmological analyses.
-
-Our conclusions are the following:
+We have presented a study of the drift of the \yc{intrinsic} SNe~Ia stretch
+distribution as a function of redshift.
+We \yc{built a magnitude-limited SN~Ia sample} from the Pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to which we added HST and SNfactory data from \cite{rigault2018} for the
+\yc{high- and} low-redshift bins. We only considered the SNe that have been discovered
+in the redshift range of each survey where selection effects are negligible, so that the observed SNe~Ia stretches are random sampling of the true underlying
+distribution. This \yc{resulted in a} 569 SN~Ia \yc{fiducial sample} (422 SNe when more
+conservative cuts were considered).
+
+Following \yc{predictions} made in \cite{rigault2018}, we introduced a redshift
+drift model which depends on the expected fraction of ``young'' and ``old'' SNe~Ia as
+a function of redshift, each age population having its own \yc{intrinsic} stretch distribution.
+
+In addition to this ``base'' modeling, we have studied various distributions, including \yc{redshift independent} models; we also studied the prediction from the ``Beams
+with Bias Correction'', a Malmquist bias correction algorithm used by recent SN~Ia cosmological analyses, \yc{implemented as a} per-sample asymmetric Gaussian distribution.
+
+Our conclusions are the following: \yc{reread and rephrase...}
 \begin{enumerate}
-
     \item Non-drifting models are excluded as good descriptions of the data of
-        our Base model. This model assumes that: (1) the younger population has
-        a unimodal Gaussian stretch distribution while the older population
-        stretch one is bimodal, one mode being the same as the young one; (2)
+        our Base model. \yc{This does not mean anything...} This model \yc{which one?} assumes that: (1) the younger population has
+        a unimodal Gaussian stretch distribution, while the older population
+        stretch \yc{distribution} is bimodal, one mode being the same as the young one; (2)
         the evolution of the relative fraction of younger and older SNe~Ia
@@
 
-    \item Given 1., we conclude that the SNe~Ia stretch is indeed drifting, as
-        previously suggested by e.g.~\cite{howell2007}.
-
-    \item The BBC approach, which assumes each survey has its own asymmetric
-        Gaussian distribution, is excluded to be an as good description of the
-        data as our drifting model. Hence, the BBC sample-based approach does
-        not accurately account for redshift drift, unlike what was suggested by
+    \item Given 1., we conclude that the SNe~Ia stretch is indeed \yc{\sout{drifting} redshift dependent}, as previously suggested by e.g.~\cite{howell2007}.
+\mr{This result is largely independent of details on each age-population modeling.} \yc{Indeed, and this should be stated first!}
+
+    \item \yc{reread \& rephrase: ``assuming'' is not excluded!} \mr{Assuming survey-based asymmetric
+        Gaussian distribution, as done, e.g., in current implementation of the BBC, is excluded to be an as good description of the
+        data as our drifting model}. Hence, sample-based approach does not accurately account for redshift drift, unlike what was suggested by
         \cite{scolnic2016} and \cite{scolnic2018a}.
@@
         bias the mean standardized magnitude of SNe~Ia affected selection
-        effects to the order of a few percents. This is comparable in scale to
+        effects to the order of a few percents. \yc{???} This is comparable in scale to
         the observational signature of exotic forms of dark energy and is
@@
 
-    \item We suggest the use of the following stretch population model:
+    \item \mr{Given the current dataset}, we suggest the use of the following stretch population model \yc{as a function of redshift}:
         \begin{align}
@@
     $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}) and using the
-    age-population drift modelization $\delta(z)$ defined in \cite{rigault2018}
+    age-population drift modelization $\delta(z)$ defined in \cite{rigault2018} \yc{too many R+18 everywhere!}
     with $K=0.87$:
@@
 \end{enumerate}
+
+\mr{We considered in this paper a simple Gaussian mixture modeling, but additional data free from
+significant Malmquist bias would enable us to refine it.
+We highlight that samples at the low- and high-redshift ends of the Hubble diagram would be particularly helpful for this drifting analysis; fortunately this will soon be provided by the
+Zwicky Transient Facility \citep{bellm2019, graham2019} and Subaru SNe~Ia program (REF HERE ASK NICOLAS), respectively.}
 
@@
 % A
-\bibitem[Abbott et al.(2018)]{abbott2018} Abbott, T.~M.~C., Abdalla, F.~B., Alarcon, A., et al.\ 2018, \prd, 98, 043526
+%\bibitem[Abbott et al.(2018)]{abbott2018} Abbott, T.~M.~C., Abdalla, F.~B., Alarcon, A., et al.\ 2018, \prd, 98, 043526
 
@@
 
-\bibitem[Ata et al.(2017)]{ata2017} Ata, M., Kitaura, F.-S., Chuang, C.-H., et al.\ 2017, \mnras, 467, 3993
+%\bibitem[Ata et al.(2017)]{ata2017} Ata, M., Kitaura, F.-S., Chuang, C.-H., et al.\ 2017, \mnras, 467, 3993
 
@@
 
-\bibitem[Chabanier et al.(2019)]{chabanier2019} Chabanier, S., Millea, M., \& Palanque-Delabrouille, N.\ 2019, \mnras, 489, 2247
+%\bibitem[Chabanier et al.(2019)]{chabanier2019} Chabanier, S., Millea, M., \& Palanque-Delabrouille, N.\ 2019, \mnras, 489, 2247
 
@@
 
-\bibitem[Coles, \& Jones(1991)]{coles1991} Coles, P., \& Jones, B.\ 1991, \mnras, 248, 1
+%\bibitem[Coles, \& Jones(1991)]{coles1991} Coles, P., \& Jones, B.\ 1991, \mnras, 248, 1
 
@@
 \bibitem[Hamuy et al.(2000)]{hamuy2000} Hamuy, M., Trager, S.~C., Pinto, P.~A., et al.\ 2000, \aj, 120, 1479
+
+\bibitem[Hinton et al.(2019)]{hinton2019} Hinton, S.~R., Davis, T.~M., Kim, A.~G., et al.\ 2019, \apj, 876, 15
 
@@
 
-\bibitem[Planck Collaboration et al.(2016)]{planck2016} Planck Collaboration, Ade, P.~A.~R., Aghanim, N., et al.\ 2016, \aap, 594, A13
+%\bibitem[Planck Collaboration et al.(2016)]{planck2016} Planck Collaboration, Ade, P.~A.~R., Aghanim, N., et al.\ 2016, \aap, 594, A13
+
+\bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration, Aghanim, N., Akrami, Y., et al.\ 2018, arXiv e-prints, arXiv:1807.06209
 
 \bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T., et al.\ 2019, \prl, 122, 221301
-
-\bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration, Aghanim, N., Akrami, Y., et al.\ 2018, arXiv e-prints, arXiv:1807.06209
 
@@
 
-\bibitem[Scolnic et al.(2018)]{scolnic2018b} Scolnic, D.~M., Lochner, M., Gris, P., et al.\ 2018, arXiv e-prints, arXiv:1812.00516
+%\bibitem[Scolnic et al.(2018)]{scolnic2018b} Scolnic, D.~M., Lochner, M., Gris, P., et al.\ 2018, arXiv e-prints, arXiv:1812.00516
 
 \bibitem[Scolnic et al.(2019)]{scolnicastro2020} Scolnic, D., Perlmutter, S., Aldering, G., et al.\ 2019, Astro2020: Decadal Survey on Astronomy and Astrophysics, 2020, 270
+
+\bibitem[Shariff et al.(2016)]{shariff2016} Shariff, H., Jiao, X., Trotta, R., et al.\ 2016, \apj, 827, 1
 

### 2020-04-17 Nicolas.tex
Corrections after MR call

@@
 \newcommand{\nn}[1]{{\textcolor[rgb]{0.25, 0.50, 0}{#1}}}
+\newcommand{\nnn}[1]{{\textcolor[rgb]{0.40, 0.0, 0.20}{#1}}}
 \newcommand{\yc}[1]{{\textcolor{BrickRed}{#1}}}
@@
     \and M. Briday\inst{1}
+    \and J. Nordin
     \and Y.-L. Kim\inst{1}
@@
 \abstract{
-{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
-measure of survey statistic increase, the question of astrophysical systematic
-uncertainties rises, and notably that of the SN~Ia population redshift drift.}
+{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as survey statistics increase, the question of
+astrophysical systematic uncertainties rises,  notably that of
+the SN~Ia population redshift drift.}
 {In this paper, we study the redshift evolution of SN~Ia
-\textsc{\texttt{SALT2.4}} lightcurve stretch\yc{\sout{es}}, \yc{\sout{which is}} a purely intrinsic SN
-property, to probe this drift.}
-{The SN stretch has been shown to strongly correlate with the SN's environment,
-and notably with stellar age tracers. Following the \yc{evolution of} the fraction
-of young and old SNe~Ia \yc{\sout{in nature} as predicted by} \cite{rigault2018}, and assuming
-non-drifting \yc{$\to$constant/stationary?}  underlying stretch distribution for each \yc{population}, we model the expected
-SN~Ia stretch distribution as a function of redshift. We test our prediction
-against data from the literature chosen to have negligible selection
-effects to ensure that any observed drift is indeed astrophysical and not
-\yc{\sout{artificial} observational}.}
-{We clearly demonstrate that the underlying SN\yc{e?}~Ia stretch distribution is
-evolving as a function of redshift, and that the young/old drifting model is a
-much better description of the data than any non-drifting model, including the
-sample-based asymmetric distributions used by the Beams with Bias Correction
-algorithm used to correct Malmquist bias. Our favored underlying stretch
-model is bimodal: a high-stretch mode shared by both young and old environments,
-and a low-stretch mode \yc{\sout{only existing in} exclusive to} old environments.}
-{\sout{The underlying SNe~Ia population is evolving as a function of redshift \yc{This is purely a repetition of a previous sentence} \nn{\textit{see new sentence}} and the
-exact impact on cosmology remains to be studied.} \nn{The precise impact of the redshift evolution of the underlying SNe Ia population on cosmology remains to be studied}.
-Yet, the astrophysical drift of
-the SN stretch distribution does affect current Malmquist bias corrections and
-thereby distances derived from SN affected by selection effects. We highlight
-that such a bias will increase with surveys covering larger redshift ranges,
-which is particularly important for LSST.}
-
-}
+\textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN property, to
+probe this drift.}
+{The SN stretch has been shown to strongly correlate with the SN's environment, notably with stellar age tracers. Following the evolution of
+fraction of young and old SNe~Ia as predicted by \cite{rigault2018}, and
+assuming constant underlying stretch
+distribution for each population \mr{as a function of redshift}, model the expected SN~Ia stretch
+distribution as a function of redshift. We test our prediction against data
+from the literature chosen to have negligible selection effects to ensure
+that any observed drift is indeed astrophysical and not observational.}
+{We clearly demonstrate that the underlying SNe~Ia stretch distribution is
+evolving as a function of redshift, and that the young/old drifting model is
+a much better description of the data than any non-drifting model, including
+the sample-based asymmetric distributions used by the Beams with Bias
+Correction algorithm to correct Malmquist bias. Our favored underlying
+stretch model is bimodal: a high-stretch mode shared by both young and old
+environments, and a low-stretch mode exclusive to old environments.}
+\nn{The
+precise impact of the redshift evolution of the underlying SNe Ia population on
+cosmology remains to be studied}. Yet, the astrophysical drift of the SN stretch
+distribution does affect current Malmquist bias corrections and thereby
+distances derived from SN affected by selection effects. We highlight that such
+a bias will increase with surveys covering larger redshift ranges, which is
+particularly important for LSST.}
 
@@
 \section{Introduction}
-
-\yc{YC: you should run a grammar corrector, e.g. \texttt{grammarly}. e.g. "SNe dichotomy" vs. "SN dichotomy", "analysis" vs. "analyses", etc.}
 
@@
 project,][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
-\yc{from} cosmic microwave background (CMB) data measured by Planck assuming the
-standard $\L$CDM \yc{ref?} \nn{textit{isn't it at the end of the sentence? Do you think it should be before the "or"?}}, or when the SN luminosity is anchored at intermediate
-redshift by the baryon acoustic oscillation (BAO) scale
-\citep{riess2019,reid2019,planck2018, feeney2019}. While using the tip of the
+from cosmic microwave background (CMB) data measured by Planck assuming the
+standard $\L$CDM \citep{planck2018,riess2019,reid2019}, or when the SN luminosity is anchored
+at intermediate redshift by the baryon acoustic oscillation (BAO) scale \citep{feeney2019}. While using the tip of the
 red giant branch technique in place of the Cepheids seem to favor an
@@
 
-The H$_0$ tension has received a lot of attention as it could be a sign of new
+The H$_0$ tension has received a lot of attention, as it could be a sign of new
 fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
-tension when accounting for all other probes \citep{knox2019} and the current
-most promising scenario appears to be a burst of expansion at the
-matter-radiation decoupling moment caused by a (fine tuned) early dark energy
-\citep{poulin2019}. \yc{I don't see the point of going into this debate...}
-
+tension when accounting for all other probes \citep{knox2019}, but see e.g. \cite{poulin2019}.
 Alternatively, systematic effects affecting one or several of the aforementioned
-analyses could also explain the tension. \cite{rigault2015} suggested
-that SNe Ia from the Cepheid calibrator sample differ by construction from the
-Hubble flow \yc{ones} as it \yc{they?} strongly favors \yc{SNe don't favor anything...} \nn{\textit{but the Cepheid sample does, right?}} young stellar population environments where
-one could find Cepheids. This selection effect would impact the derivation of
-H$_0$ if SNe~Ia from young and older environments differ in average \yc{standardized} magnitudes.
+analyses could also explain the tension. \cite{rigault2015} suggested that SNe
+Ia from the Cepheid calibrator sample differ by construction from the Hubble
+flow \nn{sample} ones as \nn{the former} strongly
+favors young stellar population\nn{s}, while the latter not. This selection effect would impact the derivation of H$_0$ if
+SNe~Ia from young and older environments differ in average standardized
+magnitudes.
 
@@
 for deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
-between the SNe and their host remains unclear when using global properties such as
-the host stellar mass, which raises the question of the accuracy of the
+between the SNe and their host remains unclear when using global properties such
+as the host stellar mass, which raises the question of the accuracy of the
 correction. More recently, studies have used the local SN environment to probe
 more direct connections between the SN and its environments \citep{rigault2013},
-showing that local age tracers such as the local specific star formation rate or
-the local color are more strongly correlated with the standardized SN magnitude
-\citep{rigault2018, roman2018, kim18}, suggesting age as the driving parameter
-underlying the mass-step. If true, this would have a significant impact for
-cosmology because the redshift drift magnitude correction to apply might
-strongly vary \yc{just the existence of a redshift-drift mag. correction is of importance here, even if it does not vary} \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
+showing that local age tracers such as the Local specific Star Formation Rate
+\nn{(LsSFR)} or the local color are more strongly correlated with the standardized
+SN magnitude \citep{rigault2018, roman2018, kim18}, suggesting age as the
+driving parameter underlying the mass-step. If true, this would have a
+significant impact for cosmology, \mr{since the environmental correction to apply to SN standardization could strongly vary with redshift.} \citep{rigault2013, childress2014, scolnic2018a}. Yet, the
 importance of local SN environmental studies remains highly debated
-\cite[e.g.][]{jones2015,jones2019} and especially the impact it \yc{what does `it' refer to, the `importance'?} \nn{\textit{no, to the SN environment. Do you think it needs rephrasing?}} has on the
-derivation of H$_0$ \citep{jones2015,riess2016,riess2018,rose2019}.
+\cite[e.g.][]{jones2015, jones2019} and especially the impact \mr{of such an astrophysical bias} has on the derivation of
+H$_0$ \citep{jones2015, riess2016, riess2018, rose2019}.
+
+%\nn{We point that there has been sample-based approach meant to account for an unmodeled variation of host-related parameters thanks to multiple redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show aren't as effective as a direct modeling of such a drift}.
+
+The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate. 
+\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
+    shown that the relative SNe~Ia rate in galaxies could be explained if two
+    populations existed, one young, following the host star formation activity,
+    and one old following the host stellar mass (the so called ``prompt and
+    delayed'' or ``A+B'' model). In \cite{rigault2018} we used the LsSFR to classify which
+    are the younger (those with a high LsSFR) and which are the older (those
+    with low LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN
+    stretch has been known to be strongly correlated with the SN host properties
+    \citep{hamuy1996, hamuy2000}, correlation that has been extensively confirmed since \citep[e.g.][]{neill2009,
+sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
+rigault2013, pan2014, kim19}. Following the ``A+B'' model and the connection
+between SN stretch and host properties, \cite{howell2007} first discussed the
+potential redshift drift of the SN stretch distribution. In this paper we
+revisit this analysis with the most recent SNe~Ia dataset.
 
 In this paper, we take a step aside to probe the validity of our modeling of the
-SN population which we claim to be constituted of two age-populations
+SN population, which we claim to be constituted of two age-populations
 \citep{rigault2013,rigault2015,rigault2018}: one old and one younger, the former
 having on average lower lightcurve stretches and being brighter after
-standardization. We use the correlation between the SN age, \yc{as} probed by local
-specific star formation rate \citep{rigault2018}, and the SN stretch to model
+standardization. We use the correlation between the SN age, as probed by the LsSFR, and the SN stretch to model
 the expected evolution of the underlying SN stretch distribution as a function
 of redshift. This modeling relies on three assumptions: (1) there are two
-\yc{distinct} populations of SNe~Ia; (2) the relative fraction of each of these populations as
-a function of redshift follows the \yc{model} presented in \cite{rigault2018}
-and (3) the underlying distribution of stretch for each age sample \yc{is constant} \nn{\textit{what does it mean for a distribution to be constant?
-It seems weird that we say "does not significantly evolve" after saying it is constant, no?}} and does not significantly \yc{evolve with time}.
-This paper aims at testing this \yc{specific model} with
-datasets from the literature.
-
-\yc{Shouldn't this paragraph, introducing the concept of old and young populations and lsSFR, be moved \emph{before} the previous one?}
-The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate.
-\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
-the relative SNe~Ia rate in galaxies could be explained if two populations
-existed, one young, following the host star formation activity, and one old
-following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
-model). In \cite{rigault2018} we used the specific Star Formation Rate at the SN
-location (Local sSFR or LsSFR \yc{this has already been used multiple times, define before}) to classify which are the younger (those with a
-high LsSFR) and which are the older (those with low LsSFR). Furthermore, since
-the first SNe~Ia host analyses, the SN stretch has been known to be strongly
-correlated with the SN host properties \citep{hamuy1996, hamuy2000}, which \yc{what does this ``which'' refer to?} has
-been extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
-    lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
-pan2014, kim19}. Following the ``A+B'' model and the connection between SN stretch
-and host properties, \cite{howell2007} first discussed the potential redshift
-drift of the SN stretch distribution. In this paper we revisit this analysis
-with the most recent SNe~Ia dataset.
+distinct populations of SNe~Ia; (2) the relative fraction of each of these
+populations as a function of redshift follows the model presented in
+\cite{rigault2018} and (3) the underlying distribution of stretch for each age
+sample is constant. This paper aims at testing this specific model with datasets from the literature. We highlight
 
 We present in section~\ref{sec:sample} the sample we are using for this
-analysis, \yc{derived from} the Pantheon catalog \citep{scolnic2018a}. We
-discuss the importance of obtaining a ``complete'' sample, i.e. representative of
-the true underlying SNe Ia distribution, and how we build one from the Pantheon
-sample. We then present in section~\ref{sec:modeling} our modeling of the
-distribution of stretch as a function of redshift\nn{-}based models for the stretch
-distribution of young and old SNe~Ia \yc{what??? rephrase} \nn{\textit{what isn't clear? Does the hyphen help?}}. Our results are presented in
-section~\ref{sec:results}, where we test whether the SN stretch \yc{distribution} evolves as a
-function of redshift and if the \yc{aforementioned} age model is in good agreement with this
-evolution. Consequences for cosmology of our results are briefly discussed in
-section~\ref{sec:bbc} and we conclude in section~\ref{sec:ccl}.
+analysis, derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
+importance of obtaining a ``complete'' sample, i.e.\ representative of the true
+underlying SNe Ia distribution, and how we build one from the Pantheon sample.
+We then present in section~\ref{sec:modeling} our modeling of the distribution of stretch and our results are presented in
+section~\ref{sec:results}.
+In this section we test whether the SN stretch distribution evolves as a function of redshift and if the aforementioned age model is in good agreement with this evolution.
+We briefly discuss these results in the context of SN cosmology in section~\ref{sec:bbc} and we conclude in
+section~\ref{sec:ccl}.
 
@@
 
-We base our analysis on the \yc{most recent comprehensive} SNe~Ia compilation, the Pantheon catalog from
-\cite{scolnic2018a}. A naive approach to test the SN stretch redshift drift
-would be to simply compare the observed SN stretch distributions in a few bins
-of redshift.  In practice, \yc{however}, \yc{differential} selection effects are affecting the observed SN
-stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
-with the lightcurve stretch (and color), the first SNe~Ia that a
-magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
-Consequently, if \yc{magnitude-related} selection effects are not accounted for, one might confuse true
-population drift with \yc{\sout{selection effects} survey properties}, and conversely.
+We base our analysis on the most recent comprehensive SNe~Ia
+compilation, the Pantheon catalog from \cite{scolnic2018a}. A naive approach to
+test the SN stretch redshift drift would be to simply compare the observed SN
+stretch distributions in a few bins of redshift. In practice, however, \mr{differential} selection effects are affecting the observed SN stretch
+distributions. Indeed, because the observed SN~Ia magnitude correlates with the
+lightcurve stretch (and color), the first SNe~Ia that a magnitude-limited survey
+will miss are the lowest-stretch (and reddest) ones. Consequently, if
+magnitude-related selection effects are not accounted for, one might confuse
+true population drift with survey properties, and conversely.
 
@@
 and host redshift, the selection effects of magnitude-limited surveys should be
-negligible below a given redshift \yc{at which} even the faintest SNe~Ia can be observed.
-In contrast, targeted surveys have highly complex selection functions and will
-be discarded from our analysis. Fortunately, modern SN cosmology samples such as
-the Pantheon \yc{one} are now dominated by magnitude-limited surveys.
-
-We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia from the following surveys: PanStarrs
-\citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
-\citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
+negligible below a given redshift at which even the faintest SNe~Ia can be
+observed. In contrast, targeted surveys have highly complex selection functions
+and will be discarded from our analysis. Fortunately, modern SN cosmology
+samples such as the Pantheon one are now dominated by magnitude-limited
+surveys.
+
+We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia
+from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
+Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
 \citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
-c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full distribution
-\citep{guy2007,betoule2014}; see also \citet{bazin2011} and \citet{campbell2013}
-for similar contours, the second using a more conservative $|c| \leq 0.2$ cut.
-Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is $M_0=-19.36$
-\citep{kessler2009,scolnic2014}, we can derive the absolute standardized
-magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$ along the
-aforementioned ellipse given the standardization coefficient $\alpha=0.156$ and
-$\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is that with
-$(x_1=-1.66, c=0.25)$ and an absolute standardized magnitude at peak in Bessel $B$
-band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect this object
-typically a week before and 10 days after peak to build a suitable lightcurve,
-the \yc{effective} limiting standardized absolute magnitude is approximately
-$M_{lim} = -18.00$~mag. Hence, given the magnitude limit $m_{lim}$ of a
-magnitude limited survey, one can derive the maximum redshift $z_{lim}$ above
-which the faintest SNe~Ia will be missed using the relation between apparent
-magnitude, redshift and absolute magnitude $\mu(z_{lim}) = m_{lim} - M_{lim}$.
+c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
+distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
+\citet{campbell2013} for similar contours, the second using a more conservative
+$|c| \leq 0.2$ cut.  Assuming the SN absolute magnitude with $x_1=0$ and $c=0$
+is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
+standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
+along the aforementioned ellipse given the standardization coefficient
+$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is
+that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
+in Bessel $B$ band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect
+this object typically 5 days before and a week after peak to build a suitable lightcurve, the effective limiting standardized
+absolute magnitude is approximately $M_{lim} = -18.00$~mag. Hence, given the
+magnitude limit $m_{lim}$ of a magnitude limited survey, one can derive the
+maximum redshift $z_{lim}$ above which the faintest SNe~Ia will be missed using
+the relation between apparent magnitude, redshift and absolute magnitude
+$\mu(z_{lim}) = m_{lim} - M_{lim}$.
 
@@
         lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples of
-        the Pantheon catalog.  The individual SNe are shown as blue dots and a
-        2D histogram is shown in gray to highlight point density \yc{I don't think it's really needed}. The ellipse
-        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the corresponding
-        standardized absolute magnitude using the $\alpha$ and $\beta$
-        coefficients from \cite{scolnic2018a}. The grey diagonal lines represent
-    the $(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and
-$z=1.70$ using SNLS's $m_{lim}$ of $24.8$~mag.}
+        the Pantheon catalog. The individual SNe are shown as blue dots. The ellipse $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by
+the corresponding standardized absolute magnitude using the $\alpha$ and $\beta$
+coefficients from \cite{scolnic2018a}. The grey diagonal lines represent the
+$(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and $z=1.70$
+using SNLS's $m_{lim}$ of $24.8$~mag.}
     \label{fig:maglim}
@@
 \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.  Fig.~14 of
-\cite{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both \yc{limits} will be
+\cite{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both limits will be
 considered, as discussed below.
@@
 $5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
-agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure \yc{is also suggestive of} a more conservative $z_{lim}=0.27$.
+agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure is also
+suggestive of a more conservative $z_{lim}=0.27$.
 
@@
 SDSS surveys were more sensitive to limited spectroscopic resources;
-\cite{kessler2009} pointed out that during \yc{first} year of SDSS, SNe~Ia with
+\cite{kessler2009} pointed out that during the first year of SDSS, SNe~Ia with
 $r-mag<20.5$ were favored for spectroscopic follow-up, corresponding to a
 redshift cut at $0.15$. For the rest of the SDSS survey, additional
-spectroscopic resources were \yc{available}, and \cite{kessler2009} and \cite{dilday2008}
-show a relative completeness \yc{what is a \emph{relative completeness}???} up to $z_{lim}=0.2$. Following these analyses, we
-will use $z_{lim}=0.2$ as the baseline SDSS redshift limit.
-
-For the rest of the analysis, we will consider the aforementioned redshift
-limits, as well as more conservative \yc{values} for each of these surveys, namely
-$z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS1 and $z_{lim}=0.55$ for SNLS. \yc{why do you remind the conservative values, and not the fiducial ones? Just refer to Table~1.}
-
-The complete sample selection is summarized in Table~\ref{tab:sample}, and the
-redshift distribution of these three surveys are \yc{is?} shown in Fig.~\ref{fig:cuts}.
+spectroscopic resources were available, and \cite{kessler2009} and
+\cite{dilday2008} show a \mr{reasonable} completeness up to
+$z_{lim}=0.2$. Following these analyses, we will use $z_{lim}=0.2$ as the
+baseline SDSS redshift limit.
+
+The sample selection is summarized in Table~\ref{tab:sample}, and the redshift distribution of these three surveys is shown in Fig.~\ref{fig:cuts}.
 As expected, the selected redshift limits roughly correspond to the peak of
@@
 \citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
-been discovered from non-targeted searches (114 SNe~Ia, see their sections~3 and~4.2.2; \mr{SNe~Ia time series are published in \citealt{Saunders2020}}). 
-The spectroscopic follow-up was done over a redshift range of
-$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the search was much
-deeper. As such, these SNe~Ia \yc{\sout{should also} are assumed to} be a random sampling of the underlying
-SN population. The SNfactory sample is particularly useful for studying SN property
+been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
+and~4.2.2; \mr{SNe~Ia time series are published in \citealt{Saunders2020}, see also, Aldering in prep. and Ponder in prep.)}. The
+spectroscopic follow-up was done over a redshift range of $0.02 < z <
+0.09$~\citep[as in][]{rigault2018}, while the search was much deeper. As such,
+these SNe~Ia are assumed to be a random sampling of the underlying SN
+population. The SNfactory sample is particularly useful for studying SN property
 drift, as it enables us to have a large SN~Ia sample at $z<0.1$.
 
-Finally, we include the HST sample from Pantheon, that \yc{\sout{follows the same logic
-of having} similarly have} a search deeper than the follow-up and \yc{that} we therefore kept entirely
-\citep{strolger04}. %, see Table~\ref{tab:sample}.
-
-We present the stretch
-distribution and redshift histogram of these five surveys \yc{up to their respective $z_{lim}$} in
-Fig.~\ref{fig:sample}.
-
-\yc{You should also present/refer to Fig.5, showing the significant evolution of mean stretch as function of $z$ in complete samples. Next Sec. aims at modeling this evolution.}
+Finally, we include the HST sample from Pantheon, that similarly have a search
+deeper than the follow-up and that we therefore kept entirely
+\citep{strolger04}.
+
+We present the stretch distribution and redshift histogram of these five surveys
+up to their respective $z_{lim}$ in Fig.~\ref{fig:sample}.
+\mr{The evolution of the mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are split in redshift bins of regular sample size.
+We see that SNe~Ia at higher redshift have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch distribution is drifting.}
 
@@
     \caption{\textit{Bottom:} \textsc{\texttt{SALT2.4}} lightcurve stretch as a
-        function of redshift for each survey considered in this
-        analysis (see legend).  Solid (resp. open) markers correspond to the
-        conservative (resp. fiducial) \yc{\sout{selection} redshift} cuts. \textit{Top:} combined
-        redshift histogram in dark (resp. light) gray for the conservative
-    (resp. fiducial) \yc{redshift} cuts.}
+        function of redshift for each survey considered in this analysis (see
+        legend). Solid (resp. open) markers correspond to the conservative
+        (resp. fiducial) redshift cuts. \textit{Top:} combined redshift
+    histogram in dark (resp. light) gray for the conservative (resp. fiducial)
+redshift cuts.}
     \label{fig:sample}
@@
 \cite{rigault2018} presented a model for the evolution of the fraction of
-younger vs. older SNe~Ia as a function of redshift following former work on
+younger and older SNe~Ia as a function of redshift following former work on
 rates and delay time distributions \citep[e.g.,][]{mannucci2005,
 scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
-In short, it was assumed that the number of \yc{``young''} SNe~Ia follows the star
-formation activity (SFR \yc{R for activity? H for History?}) in the Universe, while the number of \yc{``old''} SNe~Ia
+In short, it was assumed that the number of ``young'' SNe~Ia follows the star
+formation \nn{rate} (SFR) in the Universe, while the number of ``old'' SNe~Ia
 follows the number of Gyr-old stars in the Universe, i.e. the stellar mass
@@
 fraction of young (resp. old) SNe~Ia in the Universe as a function of redshift,
-then the ratio $\delta/\psi$ is expected to follow the evolution of the
-specific star formation rate (SFR/M$^*$), which goes as $(1+z)^{2.8}$ until
-$z\sim2$ \citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
+then the ratio $\delta/\psi$ is expected to follow the evolution of the specific
+star formation rate (SFR/M$^*$), which goes as $(1+z)^{2.8}$ until $z\sim2$
+\citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
 \citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
@@
 
-\yc{There is a generic confusion between ``SN stretch'' (individual SN, as in Eq.3/4) and something like ``SN stretch distribution'' (population, as in Eq.2). This is confusing...}
-
-To model the evolution of the SN stretch \yc{distribution} as a function of redshift, given our
-aforementioned model of the evolution of the fraction of younger and older
-SNe~Ia with \yc{\sout{redshift} cosmic time}, we need to model the SN stretch distribution for each age
-subsample. \yc{That paragraph sounds like a tautology...}
-
-\yc{I suggest we remove the use of ``we'' when referring to previous papers, like ``In R+18, we did...'' to be replaced by ``R+18 did...''}
-
-\cite{rigault2018} presented the relation between SN stretch and LsSFR
-measurement, \yc{a progenitor age tracer}, using the SNfactory sample. This
-relation is shown in Fig.~\ref{fig:stretchlssfr} \yc{for SNf SNe used in the current analysis}. Given the structure of the
-stretch-LsSFR scatter plot, our model of the underlying SN~Ia stretch
-distribution is defined as follows:
-\begin{itemize}
-\item for the younger population (i.e., $\log(\mathrm{LsSFR})>-10.82$), the
-\yc{\sout{underlying}} stretch distribution is modeled as a \yc{single} normal distribution
-$\mathcal{N}(\mu_1, \sigma_1^2)$; 
-\item the older population (i.e., $\log( \mathrm{LsSFR})<-10.82$) stretch
-distribution is modeled as a \yc{bimodal} Gaussian mixture $a\times \mathcal{N}(\mu_1, \sigma_1^2) +
-(1-a)\times \mathcal{N}(\mu_2, \sigma_2^2)$, 
-where one mode is the same as for the young population, $a$
-representing the relative influence of the two \yc{modes}.
-\end{itemize}
+To model the evolution of the \nn{full} SN stretch distribution as a function of redshift, given our aforementioned model of the evolution of the fraction of
+younger and older SNe~Ia with cosmic time, we need to model the SN stretch distribution for each age
+subsample.
 
@@
     \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$)
-        as a function of the local specific star formation rate (LsSFR) for SNf
-        SNe used in this analysis. The color corresponds to the probability,
-        $p_y$, for the SNe~Ia to be young, i.e. to have $\log\mathrm{LsSFR} \geq
-        -10.82$ \citep[see][]{rigault2018}. \textit{Right}: $p_y$-weighted histogram of the SN stretches, \yc{as well as the adjusted Base model}; the younger and
-    older population contributions are shown in purple and yellow, respectively.}
+        as a function of the local specific star formation rate (LsSFR) for
+        SNfactory SNe used in this analysis. The color corresponds to the
+        probability, $p_y$, for the SNe~Ia to be young, i.e. to have
+        $\log\mathrm{LsSFR} \geq -10.82$ \citep[see][]{rigault2018}.
+        \textit{Right}: $p_y$-weighted histogram of the SN stretches, as well as
+        the adjusted Base model; the younger and older population contributions
+        are shown in purple and yellow, respectively.}
     \label{fig:stretchlssfr}
@@
 
-When combined to the previously described prediction of the evolution of younger SNe~Ia as a function
-of redshift (Eq.~\ref{eq:delta}), our model $X_1(z)$ for the underlying stretch distribution as a function of redshift is written
-\begin{multline}
-\label{eq:stretchz}
-X_1(z) = 
-\delta(z) \times \mathcal{N}(\mu_1,\sigma_1^2) + {} \\
-(1 - \delta(z)) \times \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
-(1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right].
+\cite{rigault2018} presented the relation between SN stretch and LsSFR
+measurement, a progenitor age tracer, using the SNfactory sample. This relation
+is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory SNe used in the
+current analysis. Given the structure of the stretch-LsSFR scatter plot, our
+model of the underlying SN~Ia stretch distribution is defined as follows:
+
+\begin{itemize}
+    \item for the younger population (i.e., $\log(\mathrm{LsSFR})\geq-10.82$),
+        the stretch distribution is modeled as a single normal distribution
+        $\mathcal{N}(\mu_1, \sigma_1{}^2)$; 
+    \item the older population (i.e., $\log( \mathrm{LsSFR})<-10.82$) stretch
+        distribution is modeled as a bimodal Gaussian mixture $a\times
+        \mathcal{N}(\mu_1, \sigma_1{}^2) + (1-a)\times \mathcal{N}(\mu_2,
+        \sigma_2{}^2)$, where one mode is the same as for the young population,
+        $a$ representing the relative influence of the two modes.
+\end{itemize}
+
+\mr{The stretch probability distribution function (pdf) of a given SN will be the linear combination of the stretch distributions of these two population weighted by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
+But generally, the fraction of young SNIa as a function of redshift is given by $\delta$(z) (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model of the underlying distribution of SNIa as a function of redshift $X_1(z)$ is given by}:
+\begin{multline}\label{eq:stretchz}
+    X_1(z) = \delta(z) \times \mathcal{N}(\mu_1,\sigma_1{}^2) + {} \\
+    (1-\delta(z)) \times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
+    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right].
 \end{multline}
-\yc{In fact, not clear what this eq. does here: it is supposed to be superseded by Eq.3/4. Maybe it should be moved to Sec.3.1.2 when we use $\delta(z)$ as a proxy to $y^i$.}
+
+\subsection{Base model applied to data}
+\label{sec:basemodelapplied}
+
+Given the probability $y^i$ that a given SN is young, and assuming our Base (see section~\ref{sec:basemodel}), the probability to measure a \textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
+\begin{align}\label{eq:likelihoodsnf}
+    \prob{x^i_1}{\vec{\theta}; \mathrm{d}x^i_1, y^i} =
+    y^i & \times
+    \mathcal{N}\left(x^i_1 \mid \mu_1, \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
+    \nonumber\\
+    (1-y^i) &\times \bigg[
+    a \times \mathcal{N}\left(x^i_1 \mid \mu_1,
+    \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
+    \nonumber\\
+    & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
+    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg],
+\end{align}
 
 The maximum-likelihood estimate of the 5~free parameters
-$\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained by
-minimizing the \yc{following} pseudo-$\chi^2$: %$\mathcal{X}^2$ defined as:
-\begin{equation}
-    \label{eq:likelihood}
-    \mathcal{X}^2 = -2 \sum_i \ln \prob{x^{i}_{1}}{\vec{\theta};
-    \mathrm{d}x^{i}_{1}, y^i},
+$\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained
+by minimizing the following:
+\begin{equation}\label{eq:likelihood}
+    -2\ln(L) = -2 \sum_i \ln \prob{x^{i}_{1}}{\vec{\theta};
+    \mathrm{d}x^{i}_{1}, y^i}.
 \end{equation}
-with
-\begin{align}
-    \label{eq:likelihoodsnf}
-    \prob{x^{i}_{1}}{\vec{\theta}; \mathrm{d}x^{i}_{1}, y^i} =
-    y^i &\times \mathcal{N}\left(x^{i}_{1} \mid \mu_1, \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
-        (1-y^i) &\times \bigg[
-        a \times \mathcal{N}\left(x^{i}_{1} \mid \mu_1,
-        \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
-     &(1-a) \times \mathcal{N}\left(x^{i}_{1} \mid \mu_2,
-     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) \bigg],
-\end{align}
-where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
-$y^i$ are the \textsc{\texttt{SALT2.4}} stretch, its associated error and the
-probability that the SN is young, respectively.  In practice, to ensure fraction
-$a$ is constrained between 0 and 1, we fit for $\alpha$ such that
-$a=\arctan(\alpha)/\pi+0.5$ \yc{\sout{, which results in an asymmetric error on $a$}}.
-\yc{Given the fact that 1.
-it's barely asymmetric ($a\sim 0.5$), 2.
-no one cares, it should be symmetrized (quadratically).
-Probably $\sigma$ errors are not symmetric either!}
-
-Depending on whether $y^i$ can be estimated directly from LsSFR measurements or
-not, there are two ways to proceed, which we now discuss.
+
+Depending on whether $y^i$ can be estimated directly from LsSFR measurements or not, there are two ways to proceed, which we now discuss.
 
@@
 
-For the SNfactory sample, we can readily set $y^i = p_y{}^i$, the probability to
+For the SNfactory sample, we can readily set $y^i = p_y^i$, the probability to
 have $\log \mathrm{LsSFR} \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
-minimize Eq.~\ref{eq:likelihood} with respect to~$\vec{\theta}$.  Results on fitting
-the SNf SNe with this model are shown Table~\ref{tab:modelresults} and
+minimize Eq.~\ref{eq:likelihood} with respect to~$\vec{\theta}$. Results on
+fitting the SNf SNe with this model are shown Table~\ref{tab:modelresults} and
 illustrated in Fig.~\ref{fig:stretchlssfr}.
@@
     \caption{Best fit values of the parameters for the Base stretch distribution
-    model when applied to the SNfactory dataset only (114 SNe~Ia), the reference 569
-SN~Ia sample or the conservative one (422).}
+    model when applied to the SNfactory dataset only (114 SNe~Ia), the fiducial
+569 SN~Ia sample or the conservative one (422).}
     \label{tab:modelresults}
     \begin{tabular}{lccccc}
-    \hline\hline%\\[-0.8em]
-        Sample & $\mu_1$  & $\sigma_1$ &$\mu_2$  & $\sigma_2$ & a \\%[0.15em]
-        \hline%\\[-0.8em]
-        SNfactory & $0.41 \pm 0.08$  & $0.55 \pm 0.06$ & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$ & $0.48^{+0.08}_{-0.08}$ \\%[0.15em]
-        Fiducial & $0.37 \pm 0.05$  & $0.61 \pm 0.04$ & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$ & $0.51^{+0.09}_{-0.10}$ \\%[0.15em]
-        Conservative & $0.38 \pm 0.05$  & $0.60 \pm 0.04$ & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$ & $0.47^{+0.09}_{-0.08}$ \\%[0.15em]
+    \hline\hline
+        Sample & $\mu_1$  & $\sigma_1$
+               & $\mu_2$ & $\sigma_2$
+               & a \\
+        \hline
+        SNfactory & $0.41 \pm 0.08$ & $0.55 \pm 0.06$
+                  & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$
+                  & $0.48^{+0.08}_{-0.08}$ \\
+        Fiducial & $0.37 \pm 0.05$  & $0.61 \pm 0.04$
+                 & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$
+                 & $0.51^{+0.09}_{-0.10}$ \\
+        Conservative & $0.38 \pm 0.05$  & $0.60 \pm 0.04$
+                     & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$
+                     & $0.47^{+0.09}_{-0.08}$ \\
         \hline
@@
 probability of a SN to be young. This still corresponds to minimizing
-Eq.~\ref{eq:likelihood} with respect to the parameters $\vec{\theta}\equiv(\mu_1,
-\mu_2, \sigma_1, \sigma_2, a)$ of the stretch distribution $X_1$
-(Eq.~\ref{eq:stretchz}), but this time assuming $y^i = \delta(z^i)$ for any
-given SN~$i$.
+Eq.~\ref{eq:likelihood} with respect to the parameters
+$\vec{\theta}\equiv(\mu_1, \mu_2, \sigma_1, \sigma_2, a)$ of the stretch
+distribution $X_1$ (Eq.~\ref{eq:stretchz}), but this time assuming $y^i =
+\delta(z^i)$ for any given SN~$i$.
 
 For the rest of the analysis, we will therefore minimize Eq.~\ref{eq:likelihood}
-using $p_y{}^i$ --~the probability for the SN $i$ to be young~-- when available
-(i.e. for SNfactory dataset), and $\delta(z^{i})$ --~the expected fraction of
-young SNe~Ia at the SN redshift $z^{i}$~-- otherwise.
+using $p_y^i$ --~the probability for the SN $i$ to be young~-- when available
+(i.e. for SNfactory dataset), and $\delta(z^i)$ --~the expected fraction of
+young SNe~Ia at the SN redshift $z^i$~-- otherwise.
 
 Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
-(resp. conservative) sample are given Table~\ref{tab:modelresults} and
-\yc{predicted redshift evolution of mean stretch} illustrated in Fig.~\ref{fig:modelall}.
-\yc{I'm confused: grey points in Fig.5 are model independent, aren't they?
-Only the curve is related to the model.} We see in this figure that the measured
-mean SN~Ia stretch per bin of redshift closely follows our redshift drift
-modeling; that is, when considering selection-bias-free SN~Ia samples, SNe~Ia at
-higher redshift have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$)
-than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$). \yc{This is model-independant, and a nice result by itself, and should be already mentioned right at the end of Sec.2.}  This is indeed
-what is expected if old environments favor low SN stretches
+(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
+predicted redshift evolution of mean stretch (mean of Eq.~\ref{eq:stretchz})
+illustrated as a blue band in Fig.~\ref{fig:modelall}. 
+We see in this figure that the measured mean SN~Ia stretch per redshift bins of equal sample size closely follows our redshift drift modeling.
+This is indeed what is expected if old environments favor low SN stretches
 \citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
@@
         considering the fiducial or the conservative samples, respectively. The
-        orange horizontal line represents the mean stretch of the reference
+        orange horizontal line represents the mean stretch of the fiducial
         sample, illustrating the expectation if the SN stretch distribution is
@@
         are compatible. The light-blue band illustrates the amplitude of the
-    error of the best fit model when considering the reference dataset.}
+    error of the best fit model when considering the fiducial dataset.}
     \label{fig:modelall}
@@
 distribution following \cite{rigault2018}, i.e.\ as a single Gaussian for the
-``young'' SNe~Ia and a mixture of two Gaussians for the ``old'' SNe Ia population, one
-\yc{being} the same as for the young population, plus another one for the fast-declining SNe~Ia
-that seem to only exist in old \yc{\sout{host galaxy stellar populations} local environments}.
-This is our so-called ``Base'' model.
-However, to test different modeling choices, we have implemented a suite of
-alternative parametrizations that we also adjust to the data following the
-procedure described in section~\ref{sec:modelnopy}.
-
-\cite{howell2007} used a simpler \yc{unimodal} model per age category, assuming a \yc{single} normal
-distribution for \yc{each of} the young and old populations. We thus consider a
-``Howell+drift'' model, with one single Gaussian per age group and the
-$\delta(z)$ drift from Eq.~\ref{eq:delta}.
-
-Alternatively, since we aim at probing the existence of an \yc{evolution with redshift}, we
-also test \yc{constant} models by restricting the ``Base'' and ``Howell'' models
-to use a supposedly redshift-independent fraction $\delta(z) \equiv f$ of young SNe;
-these models are \yc{hereafter} labeled ``Base+constant'' and ``Howell+constant''.
-
-We also consider another \yc{intrinsically} non-drifting model, the one developed for Beams
-with Bias Correction \cite[BBC,][]{scolnic2016, kessler2017}, used in recent SN
-cosmological analyses \cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016,
-riess2019} to account for Malmquist biases. The BBC formalism assumes
+``young'' SNe~Ia and a mixture of two Gaussians for the ``old'' SNe Ia
+population, one being the same as for the young population, plus another one for
+the fast-declining SNe~Ia that seem to only exist in old local environments.
+This is our so-called ``Base'' model. However, to test different modeling
+choices, we have implemented a suite of alternative parametrizations that we
+also adjust to the data following the procedure described in
+section~\ref{sec:modelnopy}.
+
+\cite{howell2007} used a simpler unimodal model per age category, assuming a
+single normal distribution for each of the young and old populations.  We thus
+consider a ``Howell+drift'' model, with one single Gaussian per age group and
+the $\delta(z)$ drift from Eq.~\ref{eq:delta}.
+
+Alternatively, since we aim at probing the existence of an evolution with
+redshift, we also test constant models by restricting the ``Base'' and
+``Howell'' models to use a supposedly redshift-independent fraction $\delta(z)
+\equiv f$ of young SNe; these models are hereafter labeled ``Base+constant'' and
+``Howell+constant''.
+
+We also consider another intrinsically non-drifting model, the one developed for
+Beams with Bias Correction \cite[BBC,][]{scolnic2016, kessler2017}, used in
+recent SN cosmological analyses \cite[e.g.][]{scolnic2018a, descosmopaper2019,
+riess2016, riess2019} to account for Malmquist biases. The BBC formalism assumes
 sample-based (hence intrinsically non-drifting) asymmetric Gaussian stretch
 distributions: $\mathcal{N}\left(\mu,
-\sigma_{-}^2\;\text{if}\;x_1<\mu,\;\text{else}\;\sigma_{+}^2\right)$. The idea
+\sigma_-{}^2\;\text{if}\;x_1<\mu,\;\text{else}\;\sigma_+{}^2\right)$. The idea
 behind this sample-based approach is twofold: (1) Malmquist biases are driven by
-survey properties and (2) \yc{since} current surveys cover limited redshift ranges,
-doing so \yc{what?} covers some potential redshift evolution information
-\citep{scolnic2016, scolnic2018a}. See further discussion concerning BBC in
-section~\ref{sec:bbc}.
-
-Finally, for the sake of completeness, we also consider simple the ``Gaussian'' and
-``Asymmetric'' Gaussian non-drifting models.
+survey properties and (2) since current surveys cover limited redshift ranges, having a sample-based approach covers
+some potential redshift evolution information \citep{scolnic2016, scolnic2018a}.
+See further discussion concerning BBC in section~\ref{sec:bbc}.
+
+Finally, for the sake of completeness, we also consider simple the ``Gaussian''
+and ``Asymmetric'' Gaussian non-drifting models.
 
@@
 
-We \yc{adjusted} each of the models described above on both the reference and conservative samples (cf.
-section~\ref{sec:sample}); \yc{results are gathered in Table~\ref{tab:comp}, and illustrated in Fig.~\ref{fig:mod_comp}.}
-
-\yc{Since} the \yc{various} models have different degrees of freedom, we use the Akaike
-Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare \yc{their ability to properly describe the observations}. This
-estimator penalizes extra degrees of freedom to avoid over-fitting the data, and
-is defined as follow:
+We adjusted each of the models described above on both the fiducial and
+conservative samples (cf. section~\ref{sec:sample}); results are gathered
+in Table~\ref{tab:comp}, and illustrated in Fig.~\ref{fig:mod_comp}.
+
+Since the various models have different degrees of freedom, we use the Akaike
+Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare their ability
+to properly describe the observations. This estimator penalizes extra degrees of
+freedom to avoid over-fitting the data, and is defined as follow:
 \begin{equation}
-    \mathrm{AIC} = \mathcal{X}^2_{min} + 2k
+    \mathrm{AIC} = -2\ln(L) + 2k
 \end{equation}
-where $\mathcal{X}^2_{min}$ is the minimum value of the pseudo-$\chi^2$ as
-defined Eq.~\eqref{eq:likelihood}, and $k$ is the number of free parameters \yc{to be adjusted}. The
-reference model is the one with the smallest AIC; in comparison to this model,
-the models with $\Delta\mathrm{AIC}<5$ are \yc{coined} acceptable, the ones with
-$5<\Delta\mathrm{AIC}<20$ are unfavored, and those with $\Delta\mathrm{AIC}>20$
-are \yc{deemed} excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for a
-Gaussian probability distribution.
+where $-2\ln(L)$ \mr{is derived by minimizing} Eq.~\eqref{eq:likelihood}, and $k$ is the number of free parameters to be
+adjusted. The reference model is the one with the smallest AIC; in comparison to
+this model, the models with $\Delta\mathrm{AIC}<5$ are coined acceptable,
+the ones with $5<\Delta\mathrm{AIC}<20$ are unfavored, and those with
+$\Delta\mathrm{AIC}>20$ are deemed excluded. This roughly corresponds to 2,
+3 and 5~$\sigma$ limits for a Gaussian probability distribution.
 
@@
     \centering
-\caption{Comparison of the relative ability for a model to describe the
-data. For each considered model, we report if the model is drifting or not,
-its number of free parameters and, for both the fiducial and the conservative
-cuts, the pseudo-$\chi^2$, $\mathcal{X}^2$ (see Eq.~\ref{eq:likelihood}), the AIC
-and the AIC difference ($\Delta$AIC) between this model and the Base model used
-as reference for it has the lowest AIC.}
+    \caption{Comparison of the relative ability of each model to describe the
+        data. For each considered model, we report if the model is drifting or
+        not, its number of free parameters and, for both the fiducial and the
+        conservative cuts, $-2\ln(L)$ (see
+    Eq.~\ref{eq:likelihood}), the AIC and the AIC difference ($\Delta$AIC)
+between this model and the Base model used as reference for it has the lowest
+AIC.}
     \label{tab:comp}
     \begin{tabular}{ccc|ccc|ccc}
-    \hline\hline%\\[-0.8em]
-        & & & \multicolumn{3}{c}{Fiducial sample (569 SNe)} & \multicolumn{3}{|c}{Conservative sample (422 SNe)} \\
-        %\cline{2-6}
+    \hline\hline
+        & & & \multicolumn{3}{c}{Fiducial sample (569 SNe)}
+            & \multicolumn{3}{|c}{Conservative sample (422 SNe)} \\
         Name & drift & $k$ &
-        $\mathcal{X}^2$ & AIC & $\Delta \mathrm{AIC}$ & $\mathcal{X}^2$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$\\%[0.15em]
-        \hline%\\[-0.8em]
+        $-2\ln(L)$ & AIC & $\Delta$AIC & $-2\ln(L)$ & AIC & $\Delta$ AIC\\
+        \hline
 
@@
         & 1456.7 & 1466.7 & -- 
-        & 1079.5 & 1089.5 & -- \\%[0.15em]
+        & 1079.5 & 1089.5 & -- \\
 
         Howell+drift & $\delta(z)$ & 4
-        & 1463.3 & 1471.3 & $-4.6$  %&$1.0\times10^{-1}$
+        & 1463.3 & 1471.3 & $-4.6$
         & 1088.2 & 1096.2 & $-6.7$ 
-        %& $3.4\times10^{-2}$
-        \\%[0.15em]
+        \\
 
         Asymmetric & -- & 3
-        & 1485.2 & 1491.2 & $-24.5$  %&$4.7\times10^{-6}$
+        & 1485.2 & 1491.2 & $-24.5$
         & 1101.3 & 1107.3 & $-17.8$ 
-        %& $1.4\times10^{-4}$
-        \\%[0.15em]
+        \\
 
         Howell+const & $f$ & 5
-        & 1484.2 & 1494.2 & $-27.5$  %&$1.0\times10^{-6}$
+        & 1484.2 & 1494.2 & $-27.5$
         & 1101.2 & 1111.2 & $-21.7$ 
-        %& $1.9\times10^{-5}$
-        \\%[0.15em]
+        \\
 
         Base+const & $f$ & 6
-        & 1484.2 & 1496.2 & $-29.5$  %&$3.9\times10^{-7}$
+        & 1484.2 & 1496.2 & $-29.5$
         & 1101.2 & 1113.2 & $-23.7$ 
-        %& $7.1\times10^{-6}$
-        \\%[0.15em]
-
-        \yc{BBC-like?} & per sample & 3x5
-        & 1468.2 & 1498.2 & $-31.5$  %&$1.5\times10^{-7}$
+        \\
+
+        Per Sample Asym. & per sample & 3$\times$5
+        & 1468.2 & 1498.2 & $-31.5$
         & 1083.6 & 1113.6 & $-24.1$ 
-        %& $5.7\times10^{-6}$
-        \\%[0.15em]
+        \\
 
         Gaussian & -- & 2
-        & 1521.8 & 1525.8 & $-59.1$  %&$1.5\times10^{-13}$
+        & 1521.8 & 1525.8 & $-59.1$
         & 1142.6 & 1146.6 & $-57.1$ 
-        %& $4.0\times10^{-13}$ 
         \\
@@
 model, implementing the \cite{rigault2018} description of the stretch
-distribution and its redshift evolution. This is therefore the reference model,
-both on the fiducial \yc{somewhere, you still use the ``ref. sample'' terminology!} and conservative samples.
-
-Furthermore, we find that \yc{\sout{underlying stretch distributions not including a
-redshift drift} redshift-independent stretch distributions} are all excluded as description of the data relative to the Base
-model. In fact, the best non-drifting model (the Asymmetric one) has a very marginal
+distribution and its redshift evolution. This is therefore the reference model, both on the fiducial and conservative samples. 
+% $\chi^2=-2\ln(L)-\ln(2\pi\sigma)$ ~ 900 so 1.6 per DoF.
+
+Furthermore, we find that redshift-independent stretch distributions are all
+excluded as \nn{suitable} descriptions of the data relative to the Base model.
+In fact, the best non-drifting model (the Asymmetric one) has a very marginal
 chance ($p \equiv \exp\left(\Delta\mathrm{AIC}/2\right) = 5\times10^{-6}$) to
@@
 assessment of qualitative facts clearly visible in Fig.~\ref{fig:modelall}: the
-mean SN stretch per bin of redshift strongly suggests \yc{a significant} redshift evolution
-rather than a constant value, and this evolution is \yc{\sout{compatible with the
-redshift drift expected from} well described by}  Eq.~\ref{eq:delta}.
-
-\mr{Surprisingly, the sample-based Gaussian asymmetric
-modeling used by current implementations of the BBC technique \citep{scolnic2016, kessler2017}, is one of the worst modelisations of our analysis (see section~\ref{sec:results}). While its
-pseudo-$\chi^{2}$ is the smallest of all \yc{redshift-independent} models (but still $\Delta\mathcal{X}^2 = -11.5$ worse than the reference Base model), it is
-\yc{strongly} penalized for requiring 15~free parameters (\yc{$\mu_0, \sigma_{\pm}$}~for each of the 5~samples of
-the analysis). 
-We report in Table~\ref{tab:bbc} the samples' $\mu_0$ and $\sigma_{\pm}$ adjusted
-on the \sout{supposedly} selection-free samples using our fiducial cuts (see
-section~\ref{sec:sample}. We find our results in close agreement with
-\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
-derived these model parameters using the full BBC formalism, using numerous
-simulations to model the selection effects (see details e.g., section~3 of
-\citealt{kessler2017}). The agreement between our fit of the asymmetric Gaussians on the \yc{supposedly} selection-free part of the samples and the results
-derived using the BBC formalism supports our approach to get a sample with negligible selection effects.} \yc{It's not very clear what you mean to convey in this sentence.}
-
-We further discuss the consequence of this result for cosmology in section~\ref{sec:discussion}.
+mean SN stretch per bin of redshift strongly suggests a significant redshift
+evolution rather than a constant value, and this evolution is well described by
+Eq.~\ref{eq:delta}.
+
+Surprisingly, the sample-based Gaussian asymmetric modeling used by current
+implementations of the BBC technique \citep{scolnic2016, kessler2017}, is one of
+the worst modelisations of our analysis (see section~\ref{sec:results}). While
+its $-2\ln(L)$ is the smallest of all redshift-independent models (but still $-11.5$ worse than the reference Base model), it is
+strongly penalized for requiring 15~free parameters ($\mu_0, \sigma_{\pm}$~for
+each of the 5~samples of the analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be
+interpreted as a probability $p=2\times 10^{-7}$ of being an as good representation of the data as the Base model.
+We report in Table~\ref{tab:bbc} the
+samples' $\mu_0$ and $\sigma_{\pm}$ adjusted on the \nn{nominally}
+selection-free samples using our fiducial cuts (see section~\ref{sec:sample}).
+We find our results in close agreement with \cite{scolnic2016} for SNLS and SDSS
+and with \cite{scolnic2018a} for PS1, who derived these model parameters using
+the full BBC formalism, using numerous simulations to model the selection
+effects (see details e.g., section~3 of \citealt{kessler2017}). The agreement
+between our fit of the asymmetric Gaussians on the supposedly selection-free
+part of the samples and the results derived using the BBC formalism supports our
+approach to get a sample with negligible selection effects. 
+If we were to use \cite{scolnic2016} and \cite{scolnic2018a}
+best fit values \nn{of the $\mu_0, \sigma_{\pm}$ asymmetric parameters} for SNLS, SDSS and PS1, respectively,
+the $\Delta$AIC between our Base drifting model and the BBC
+modeling would go even deeper from $-32$ to $-47$.
+We further discuss the consequence of this result for cosmology in
+section~\ref{sec:discussion}.
 
@@
     \begin{tabular}{ccccccc}
-    \hline\hline%\\[-0.8em]
+    \hline\hline
     Asymmetric & $\sigma_{-}$    & $\sigma_{+}$    & $\mu_0$ \\
-    \hline%\\[-0.8em]
-    SNfactory  & 1.34 $\pm$ 0.13 & 0.41 $\pm$ 0.10 & 0.68 $\pm$ 0.15 \\%[0.15em]
-    SDSS       & 1.31 $\pm$ 0.11 & 0.42 $\pm$ 0.09 & 0.72 $\pm$ 0.13 \\%[0.15em]
-    PS1        & 1.01 $\pm$ 0.11 & 0.52 $\pm$ 0.12 & 0.38 $\pm$ 0.16 \\%[0.15em]
-    SNLS       & 1.41 $\pm$ 0.13 & 0.15 $\pm$ 0.13 & 1.22 $\pm$ 0.15 \\%[0.15em]
+    \hline
+    SNfactory  & 1.34 $\pm$ 0.13 & 0.41 $\pm$ 0.10 & 0.68 $\pm$ 0.15 \\
+    SDSS       & 1.31 $\pm$ 0.11 & 0.42 $\pm$ 0.09 & 0.72 $\pm$ 0.13 \\
+    PS1        & 1.01 $\pm$ 0.11 & 0.52 $\pm$ 0.12 & 0.38 $\pm$ 0.16 \\
+    SNLS       & 1.41 $\pm$ 0.13 & 0.15 $\pm$ 0.13 & 1.22 $\pm$ 0.15 \\
     HST        & 0.76 $\pm$ 0.36 & 0.79 $\pm$ 0.35 & 0.11 $\pm$ 0.44 \\
@@
 
-For the sake of robustness \yc{???}, we also \yc{performed tests} allowing the high-stretch mode of the
-old population to differ from the young population mode, hence adding two
-degrees of freedom. The corresponding fit is not significantly better,
-with a $\Delta$AIC of $-0.4$; this strengthens the fact that the young and
-old populations indeed \yc{appear to} share the same underlying high-stretch mode. We also
-allowed the young population to have a low-stretch mode, finding its amplitude
-to be compatible with~0 ($<2\% \pm \yc{???}$).
-
-\section{Discussion}
-\label{sec:discussion}
-
-To the best of our knowledge, the SN~Ia stretch redshift drift has never been explicitly accounted for in cosmological analyses, \mr{though bayesian hierarchy formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016} or Steve \citep{hinton2019} can easily allow it}.
-Not doing
-so is a second order issue for SN cosmology, as it only affects the way one does \yc{account for}
-Malmquist bias. \yc{I don't understand why you say it's a 2nd order effect if all samples are significantly Malmquist biased!?} Indeed, as long as Phillip's relation \yc{ref???}
-standardization parameter $\alpha$ is not \yc{intrinsically} redshift dependent (a study behind the
-scope of this paper, but see e.g. \citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
-cosmology are blind to the underlying stretch distribution \yc{\emph{for complete samples}}. \yc{However}, all modern surveys
-do have significant Malmquist bias at least for the upper half of their SN
-redshift distribution.  \yc{As a consequence}, an ill-modeling of the underlying stretch
-distribution will bias the derived SN magnitude for these \yc{these what?} \citep[see
-e.g.,][]{rubin2015,rubin2016}. 
-The fact that the currently-used \mr{sample-based asymmetric Gaussian distribution modeling} is excluded to be an as good description of the data then our drifting model is worrisome;  we further discuss this in the next sections.
-\yc{I don't see the point of this kind of sentence...}
-
-\subsection{BBC formalism and underlying stretch distributions}\label{sec:bbc}
-
-\yc{IMHO, you should take care of not converting this paper into an anti-BBC paper: just show the results, and let people make their opinion.
-Furthermore, I find this paragraph mostly a repetition of previous statements and results...}
-
-The commonly used tool for doing Malmquist bias correction in SN cosmology is
-the BBC formalism described in \cite{scolnic2016} and \cite{kessler2017}, which
-is used in recent SN cosmological analyses \citep{jones2018b, scolnic2018a,
-brout2019, descosmopaper2019} including the direct measurement of $H_0$
-\citep{riess2016,riess2019}. \yc{This all have already been said before, there's no point at repeating it IMHO. Maybe just a simpler introductory sentence without all the references?}
-
-\mr{Yet, the current underlying stretch modeling of the BBC formalism, a sample-based asymmetric Gaussian distribution, is excluded to be an as good representation of the data than our Base drifting modeling, see section~\ref{sec:results}.
-\yc{Again, unnecessary repetition.} Then, unlike what  \citet[][section~2]{scolnic2016} and
-\citet[][section~5.4]{scolnic2018a} suggested, the fact that traditional surveys span limited redshift ranges, the per-sample approach cannot correctly account for implicit redshift drifts.} \yc{I don't understand what they suggested and what is wrong...
-Rephrase.} We stress here that, as measurements of
-modern surveys \yc{try to} cover increasingly larger redshift ranges \yc{in order} to reduce calibration
-systematic uncertainties, this approach \yc{which one?} is becoming less valid, notably for PS1,
-DES and, soon, LSST. Already with current surveys, the BBC sample-based
-technique is ruled out at representing the data as well as our best model;
-$\Delta\mathrm{AIC}<-20$, which could be interpreted as a probability $p=2\times
-10^{-7}$ of being an as good representation of the data as the Base model. \yc{Again, unnecessary repetition.} Note
-that even considering the SNfactory dataset like any other (i.e., ignoring the
-LsSFR measurements, see section~\ref{sec:modeling}), reduces this difference to
-$\Delta\mathrm{AIC}<-10$, still strongly unfavoring the modeling currently used
-as part of the \mr{sample-based formalism}.
-If we were to use \cite{scolnic2016} and
-\cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively \yc{which parameters?}, the
-$\Delta\mathrm{AIC}$ between our Base drifting model and the BBC modeling would
-go even deeper from $-32$ to $-47$.
-\yc{But you would be using parameters derived specifically for incomplete samples on specially-constructed complete sample: what is the validity of the comparison???}
-
-\subsection{First step toward quantification of the cosmological impact of
-inaccurate stretch distribution modeling.}
+We also performed tests allowing
+the high-stretch mode of the old population to differ from the young population mode, hence adding two degrees of freedom. The corresponding fit is not
+significantly better, with a $\Delta$AIC of $-0.4$. 
+this strengthens the fact
+that the young and old populations indeed appear to share the same underlying
+high-stretch mode. 
+\mr{Furthermore, one might wonder whether a low-stretch mode might also exist in the young-population, see Fig.~\ref{fig:stretchlssfr}.
+We tested that by allowing this population to also be bi-modal, finding the amplitude of this low-stretch mode to be compatible with~0 in this young population ($<2\%$).
+More generally, this rises the question of inaccurate tracing of age by a given environmental tracer (here the LsSFR).
+A dedicated analysis be presented in Briday et al.
+in prep.}
+
+\mr{Finally, considering the SNfactory dataset like any other
+(i.e., ignoring the LsSFR measurements, see section~\ref{sec:modeling}),
+reduces the significance the results presented in this section, as expected.
+Yet, non-drifting models remain strongly disfavored, and for instance, the best fitted sample-based gaussian asymmetric modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the data than our Base drifting modeling.}
+
+\section{Discussion}\label{sec:discussion}
+
+To the best of our knowledge, a SN~Ia stretch redshift drift \mr{modeling} has never been explicitly \mr{used} in cosmological analyses, though bayesian hierarchy
+formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016} or Steve
+\citep{hinton2019} can easily allow it; see e.g., section 1.3 and 2.5 of \cite{rubin2015}.
+Not doing so is a second order issue for
+SN cosmology, as it only affects the way one accounts for Malmquist bias. Indeed, as long as Phillips' relation \citep{phillips1993} standardization parameter
+$\alpha$ is not redshift dependent (a study behind the scope of
+this paper, but see e.g. \citealt{scolnic2018a}), the stretch-corrected SNe~Ia
+magnitudes used for cosmology are blind to the underlying stretch distribution for complete samples. 
+However, surveys usually do have
+significant Malmquist bias at least for the upper half of their SN redshift distribution.
+As a consequence, an ill-modeling of the underlying stretch distribution will bias the SN magnitude derived from such surveys.
+
+\subsection{Issue of with the per Sample Asymmetric modeling}
+\label{sec:bbc}
+
+Commonly used Malmquist bias correction techniques, such as the BBC-formalism, assume per Sample Asymmetric Gaussian functions for modeling the underlying stretch and color distributions.
+Yet, as shown in Section~\ref{sec:results}, such a sample-based distribution is excluded as being as good as our drifting model.
+Then, unlike what \citet[][section~2]{scolnic2016} and
+\citet[][section~5.4]{scolnic2018a} suggested, i.e. that
+traditional surveys span limited redshift ranges and therefore that the per-sample approach accounts for implicit redshift drifts, might not be correct.
+We stress here that, as measurements of modern surveys try to cover increasingly larger redshift ranges in order to reduce calibration systematic uncertainties, this sample-based approach  is becoming less valid, notably for PS1, DES and, soon, LSST.
+
+We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
+underlying stretch distribution between the per-sample asymmetric modeling and our Base drifting
+model for the PS1 sample. Our model is bimodal and the relative amplitude of
+each mode depends on the redshift-dependent fraction of old and young SNe~Ia in
+the sample: the higher the fraction of old SNe~Ia (at lower redshift), the
+higher the amplitude of the old-specific low-stretch mode. This redshift
+dependency is shown as blue to red underlying distributions in
+Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges covered by PS1. The observed
+$x_1$ histogram follows our modeling defined as the sum of individual underlying SN-redshift distributions.
+As expected, the two modeling approaches differ mostly in the negative part of the SN stretch distribution.
+The asymmetric Gaussian distribution goes through the middle of the
+bimodal distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
+under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN redshifts.
+\mr{This means that the SN bias-corrected standardized magnitude estimated at redshift affected by selection effect would be biased by ill-modeling of the true underlying stretch distribution.}
 
@@
         distribution (asymmetric Gaussian). The full line (band) is our best fit
-        (its error); the dashed line shows \cite{scolnic2018a} result. The black
-        line (band) show our best fitted Base-modeling (its error, see Table
-        \ref{tab:modelresults}) that includes redshift drift. For illustration,
-        we show as colored (from blue to red with increasing redshifts) the
-        evolution of the underlying stretch distribution as a function of
+        (its error); the dashed line shows the \cite{scolnic2018a} result. The
+        black line (band) show our best fitted Base-modeling (its error, see
+        Table \ref{tab:modelresults}) that includes redshift drift. For
+    illustration, we show as colored (from blue to red with increasing
+redshifts) the evolution of the underlying stretch distribution as a function of
 redshift for the redshift range covered by PS1 data.}
@@
 
-We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
-underlying stretch distribution between the BBC modeling and our Base drifting
-model for the PS1 sample. Our model is bimodal and the relative amplitude of
-each mode depends on the redshift-dependent fraction of old \yc{vs.\ young} SNe~Ia in the
-sample: the higher the fraction of old SNe~Ia (at lower redshift), the higher the
-amplitude of the \yc{old-specific} low-stretch mode. This redshift dependency is shown as blue to
-red underlying distributions in Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges
-covered by PS1. The observed $x_1$ histogram corresponds to the sum of
-individual underlying SN-redshift distributions. \yc{Not clear: you mean the black curve -- a model -- or the histogram -- an observation?} As expected, the two approaches \yc{which ones?}
-strongly differ in modeling the negative part of the SN stretch distribution.
-The BBC asymmetric distribution goes through the middle of the bimodal
-distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
-under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting
-model for typical PS1 SN redshifts. \yc{This is definitely not obvious from the Fig. solely...} This means that the bias-correction of
-individual SN stretches, $\overline{\delta}_{x_1}$ \citep{kessler2017}, and
-consequently the magnitude bias correction, $\mu_B$, is most likely inaccurate
-and \sout{this inaccuracy is} redshift dependent.
-
-The amplitude of this \yc{magnitude} bias for cosmology is beyond the scope of this paper given
-the complexity of the BBC analysis. It would require a full study using our Base
-model (Eq.~\ref{eq:stretchz}) in place of the sample-based asymmetric
-modeling as part of the BBC simulations.
-
-Nonetheless, it is \yc{enlightening} to study the  expected mean stretch difference as
+The amplitude of this magnitude bias for cosmology is beyond the scope of this paper given the complexity of the BBC analysis. It would require a full
+study using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
+asymmetric modeling as part of the BBC simulations.
+
+Nonetheless, it is enlightening to study the expected mean stretch difference as
 a function of redshift when considering that the underlying SN stretch
-distribution follows either the BBC formalism or our Base drifting model.
-This is shown Fig.~\ref{fig:magdrift}, where, for clarity we convert this
-difference in \yc{mean??? stretch} $x_1$ to a difference in magnitude using $\alpha=0.156$
+distribution follows either the per-sample Asymmetric Gaussian modeling or our Base drifting model. This
+is shown Fig.~\ref{fig:magdrift}, where, for clarity we convert this difference
+in mean stretch $x_1$ to a difference in magnitude using $\alpha=0.156$
 \citep{scolnic2018a}.
+\mr{We strengthen here that this does not corresponds to the actual impact on cosmological distances derived from SNeIa depending on which underlying stretch distribution modeling as it does not account for the amplitude of the Malmquist bias correction as a function of redshift and surveys.}
 
@@
     \caption{Difference of mean stretch standardization (in mag, using
-        $\alpha=0.156$ from \citealt{scolnic2018a}) between the BBC algorithm
+        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample Gaussian asymmetric modeling
         and our Base drifting model, as a function of redshift, considering the
@@
     redshifts (no redshift cut applied). The orange markers represent their mean
-evolution in 30 regularly-spaced redshift bins.}
+    evolution in 30 regularly-spaced redshift bins. 
+    \mr{This must be multiplied by the amplitude of the Malmquist bias correction to derive the SN distance modulus bias affecting cosmology, see text.}}
     \label{fig:magdrift}
@@
 
-We see that the amplitude of the modeling difference is a few tens of
-millimag, which corresponds to the expected bias affecting SNe~Ia impacted
-by sample selection effects. For comparison, the amplitude of this effect is
-similar in scale to the effect exotic forms of dark energy could
-have on $\mu(z)$. Consequently, in the era of modern cosmology, where we aim at
-probing $w_0$ at a sub-percent \yc{level} and $w_a$ at the ten-percent \yc{precision}
-\citep[e.g.,][]{lsstpaper}, it is of paramount importance to further
-understand the exact modeling of the lightcurve parameters when using SNe~Ia
-affected by Malmquist bias.
-
-\section{Conclusion}
-\label{sec:ccl}
-
-We have presented a study of the drift of the \yc{intrinsic} SNe~Ia stretch
-distribution as a function of redshift.
-We \yc{built a magnitude-limited SN~Ia sample} from the Pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to which we added HST and SNfactory data from \cite{rigault2018} for the
-\yc{high- and} low-redshift bins. We only considered the SNe that have been discovered
-in the redshift range of each survey where selection effects are negligible, so that the observed SNe~Ia stretches are random sampling of the true underlying
-distribution. This \yc{resulted in a} 569 SN~Ia \yc{fiducial sample} (422 SNe when more
+We see that the amplitude of the modeling difference is a few tens of millimag. In the era of modern cosmology, where we aim at probing $w_0$ at a
+sub-percent level and $w_a$ at the ten-percent precision
+\citep[e.g.,][]{lsstpaper}, we stress that correctly modeling of potential SN redshift drift should be further studied and care should be made when using samples affected by selection effects.
+
+\section{Conclusion}\label{sec:ccl}
+
+We have presented a study of the drift of the \mr{underlying} SNe~Ia stretch distribution as a function of redshift. We
+built a magnitude-limited SN~Ia sample from the Pantheon dataset
+\citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to which we added HST and SNfactory
+data from \cite{rigault2018} for the high- and low-redshift bins. We only
+considered the SNe that have been discovered in the redshift range of each
+survey where selection effects are negligible, so that the observed SNe~Ia
+stretches are random sampling of the true underlying distribution. This
+resulted in a 569 SN~Ia fiducial sample (422 SNe when more
 conservative cuts were considered).
 
-Following \yc{predictions} made in \cite{rigault2018}, we introduced a redshift
-drift model which depends on the expected fraction of ``young'' and ``old'' SNe~Ia as
-a function of redshift, each age population having its own \yc{intrinsic} stretch distribution.
-
-In addition to this ``base'' modeling, we have studied various distributions, including \yc{redshift independent} models; we also studied the prediction from the ``Beams
-with Bias Correction'', a Malmquist bias correction algorithm used by recent SN~Ia cosmological analyses, \yc{implemented as a} per-sample asymmetric Gaussian distribution.
-
-Our conclusions are the following: \yc{reread and rephrase...}
+Following predictions made in \cite{rigault2018}, we introduced a redshift drift
+model which depends on the expected fraction of ``young'' and ``old'' SNe~Ia as
+a function of redshift, each age population having its own underlying
+stretch distribution.
+
+In addition to this ``base'' modeling, we have studied various distributions,
+including redshift independent models; we also studied the prediction from the
+``Beams with Bias Correction'', a Malmquist bias correction algorithm used by
+\nn{several} recent SN~Ia cosmological analyses, implemented as a per-sample
+asymmetric Gaussian distribution.
+
+Our conclusions are the following:
 \begin{enumerate}
-    \item Non-drifting models are excluded as good descriptions of the data of
-        our Base model. \yc{This does not mean anything...} This model \yc{which one?} assumes that: (1) the younger population has
-        a unimodal Gaussian stretch distribution, while the older population
-        stretch \yc{distribution} is bimodal, one mode being the same as the young one; (2)
-        the evolution of the relative fraction of younger and older SNe~Ia
+
+    \item The underlying SN~Ia stretch distribution is         redshift dependent, as previously suggested by e.g.~\cite{howell2007}. This
+        result is largely independent of details on each age-population
+        modeling. \yc{\textit{Indeed, and this should be stated first!}}
+
+    \item Non-drifting models are indeed excluded as \nn{suitable} descriptions of the
+        data \nn{relative to} our Base model. This model assumes
+        that: (1) the younger population has a unimodal Gaussian stretch
+        distribution, while the older population stretch distribution
+        is bimodal, one mode being the same as the young one; (2) the
+        evolution of the relative fraction of younger and older SNe~Ia
         follows the prediction made in \cite{rigault2018}.
 
-    \item Given 1., we conclude that the SNe~Ia stretch is indeed \yc{\sout{drifting} redshift dependent}, as previously suggested by e.g.~\cite{howell2007}.
-\mr{This result is largely independent of details on each age-population modeling.} \yc{Indeed, and this should be stated first!}
-
-    \item \yc{reread \& rephrase: ``assuming'' is not excluded!} \mr{Assuming survey-based asymmetric
-        Gaussian distribution, as done, e.g., in current implementation of the BBC, is excluded to be an as good description of the
-        data as our drifting model}. Hence, sample-based approach does not accurately account for redshift drift, unlike what was suggested by
-        \cite{scolnic2016} and \cite{scolnic2018a}.
+    \item 
+        \mr{Using} survey-based asymmetric Gaussian distributions, as  done, e.g., in the current implementation of the BBC, is excluded to be good descriptions of the data \nn{relative to} our drifting model.
+Hence, the sample-based approach does not accurately
+        account for redshift drift.
 
     \item An inaccurate underlying stretch distribution model is estimated to
-        bias the mean standardized magnitude of SNe~Ia affected selection
-        effects to the order of a few percents. \yc{???} This is comparable in scale to
-        the observational signature of exotic forms of dark energy and is
-        therefore a significant source of systematic that should be carefully
-        studied.
-
-    \item \mr{Given the current dataset}, we suggest the use of the following stretch population model \yc{as a function of redshift}:
-        \begin{align}
-            \label{eqconclusion:stretchz}
+        bias the mean standardized magnitude of SNe~Ia affected \nn{by}
+        selection effects at the order of a few percents. This is
+        comparable in scale to the observational signature of exotic forms of
+        dark energy and is therefore a significant source of systematic that
+        should be carefully studied.
+
+    \item Given the current dataset, we suggest the use of the following stretch
+        population model as a function of redshift:
+        \begin{align}\label{eqconclusion:stretchz}
             X_1\left(z \right) =
@@
         $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}) and using the
-    age-population drift modelization $\delta(z)$ defined in \cite{rigault2018} \yc{too many R+18 everywhere!}
-    with $K=0.87$:
+        age-population drift modelization $\delta(z)$ defined in with $K=0.87$:
         \begin{align}
@@
 
-\mr{We considered in this paper a simple Gaussian mixture modeling, but additional data free from
-significant Malmquist bias would enable us to refine it.
-We highlight that samples at the low- and high-redshift ends of the Hubble diagram would be particularly helpful for this drifting analysis; fortunately this will soon be provided by the
-Zwicky Transient Facility \citep{bellm2019, graham2019} and Subaru SNe~Ia program (REF HERE ASK NICOLAS), respectively.}
+\nn{In this paper, we considered} a simple Gaussian mixture modeling, but
+additional data free from significant Malmquist bias would enable us to refine
+it. We note that samples at the low- and high-redshift
+ends of the Hubble diagram would be particularly helpful for this drifting
+analysis; fortunately this will soon be provided by the Zwicky Transient Facility \citep{bellm2019, graham2019} and Subaru and SeeChange SNe~Ia programs, respectively.
 
@@
 \bibitem[Perrett et al.(2010)]{perrett2010} Perrett, K., Balam, D., Sullivan, M., et al.\ 2010, \aj, 140, 518
+
+\bibitem[Phillips(1993)]{phillips1993} Phillips, M.~M.\ 1993, \apjl, 413, L105
 

### 2020-04-23 Nicolas.tex
Corrections after everyone's questions. Greg corr and comm tbd

@@
     \and M. Briday\inst{1}
-    \and J. Nordin
+    \and J. Nordin\inst{4}
     \and Y.-L. Kim\inst{1}
     \and Saul Perlmutter\inst{3}
-    \and G. Smadja\inst{1}   
+    % Gerard ne souhaite pas etre co-auteur.
 }
@@
     1 Cyclotron Road, Berkeley, CA, 94720 
+    \and
+    Institut fur Physik, Humboldt-Universität zu Berlin, Newtonstr. 15, 12489 Berlin
 }
@@
 \cite{rigault2018} and (3) the underlying distribution of stretch for each age
-sample is constant. This paper aims at testing this specific model with datasets from the literature. We highlight
+sample is constant.
+This paper aims at testing this specific model with datasets from the literature.
+\mr{Note that the progenitor age as traced by the LsSFR seems to capture physical feature intrinsic to the progenitor and/or explosion mechanism that the stretch alone is not capturing \citep{nordin2018}.}
 
@@
 \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.  Fig.~14 of
-\cite{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both limits will be
+\cite[][\mr{see their section~5}]{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both limits will be
 considered, as discussed below.
@@
 SDSS surveys were more sensitive to limited spectroscopic resources;
-\cite{kessler2009} pointed out that during the first year of SDSS, SNe~Ia with
+\cite[][\mr{see their section~2}]{kessler2009} pointed out that during the first year of SDSS, SNe~Ia with
 $r-mag<20.5$ were favored for spectroscopic follow-up, corresponding to a
@@
 these histograms.
+\mr{We show in section~\ref{sec:results} that this sample selection indeed provide a subset of SNe~Ia with insignificant selection effects when compared to state-of-the-art Malmquist correction techniques.}
 
@@
 We tested that by allowing this population to also be bi-modal, finding the amplitude of this low-stretch mode to be compatible with~0 in this young population ($<2\%$).
-More generally, this rises the question of inaccurate tracing of age by a given environmental tracer (here the LsSFR).
+More generally, this raises the question of inaccurate tracing of age by a given environmental tracer (here the LsSFR).
 A dedicated analysis be presented in Briday et al.
@@
 
-\subsection{Issue of with the per Sample Asymmetric modeling}
-\label{sec:bbc}
-
 Commonly used Malmquist bias correction techniques, such as the BBC-formalism, assume per Sample Asymmetric Gaussian functions for modeling the underlying stretch and color distributions.
@@
 \citet[][section~5.4]{scolnic2018a} suggested, i.e. that
-traditional surveys span limited redshift ranges and therefore that the per-sample approach accounts for implicit redshift drifts, might not be correct.
+traditional surveys span limited redshift ranges and that therefore the per-sample approach accounts for implicit redshift drifts, \sout{might not be correct} \nn{a direct modeling of the redshift drift is more appropriate than a sample-based approach}.
 We stress here that, as measurements of modern surveys try to cover increasingly larger redshift ranges in order to reduce calibration systematic uncertainties, this sample-based approach is becoming less valid, notably for PS1, DES and, soon, LSST.
@@
 under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN redshifts.
-\mr{This means that the SN bias-corrected standardized magnitude estimated at redshift affected by selection effect would be biased by ill-modeling of the true underlying stretch distribution.}
+\mr{This means that the SN bias-corrected standardized magnitude estimated at redshift affected by selection effect would be biased by an ill-modeling of the true underlying stretch distribution.}
 
@@
 \citep{scolnic2018a}.
-\mr{We strengthen here that this does not corresponds to the actual impact on cosmological distances derived from SNeIa depending on which underlying stretch distribution modeling as it does not account for the amplitude of the Malmquist bias correction as a function of redshift and surveys.}
+\mr{We stress here that this does not correspond to the actual impact on cosmological distances derived from SNeIa as this simple stretch average comparison does not account for the amplitude of the Malmquist bias correction to be applied as a function of redshift and surveys.}
 
@@
         result is largely independent of details on each age-population
-        modeling. \yc{\textit{Indeed, and this should be stated first!}}
+        modeling.
 
@@
         \mr{Using} survey-based asymmetric Gaussian distributions, as  done, e.g., in the current implementation of the BBC, is excluded to be good descriptions of the data \nn{relative to} our drifting model.
-Hence, the sample-based approach does not accurately
-        account for redshift drift.
+Hence, the sample-based approach does not accurately account for redshift drift and even less so as survey span larger redshift ranges.
+\mr{We warn that, even if extra degrees of freedom might be acceptable SN cosmological studies, extrapolating the SN property distributions from the volume-limited part of a survey to the magnitude-limited one affected by sample selection would still be inaccurate since the distribution would drift.}
 
@@
         $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}) and using the
-        age-population drift modelization $\delta(z)$ defined in with $K=0.87$:
+        age-population drift modelization $\delta(z)$ with $K=0.87$:
         \begin{align}
@@
     (grant agreement n 759194 - USNAC).
+    This work was supported in part by the Director, Office of Science, Office of High Energy Physics of the U.S. Department of Energy under Contract No. DE-AC025CH11231.
     % How has received that support for this project ? 
@@
 
+\bibitem[Nordin et al.(2018)]{nordin2018} Nordin, J., Aldering, G., Antilogus, P., et al.\ 2018, \aap, 614, A71
+
 % O
@@
 \bibitem[Sako et al.(2008)]{sako2008} Sako, M., Bassett, B., Becker, A., et al.\ 2008, \aj, 135, 348
+
+\bibitem[Saunders et al.(2020)]{saunders2020} Saunders, C., Aldering, G., Antilogus, P., et al.\ 2020, VizieR Online Data Catalog, J/ApJ/869/167
 

### 2020-04-23 Nicolas.tex
Greg_corr_03 added

@@
     \and
-    Institut fur Physik, Humboldt-Universität zu Berlin, Newtonstr. 15, 12489 Berlin
+    Institut fur Physik, Humboldt-Universität zu Berlin, Newtonstr. 15,
+    12489 Berlin
 }
@@
 \abstract{
-{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as survey statistics increase, the question of
-astrophysical systematic uncertainties rises,  notably that of
-the SN~Ia population redshift drift.}
-{In this paper, we study the redshift evolution of SN~Ia
-\textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN property, to
-probe this drift.}
-{The SN stretch has been shown to strongly correlate with the SN's environment, notably with stellar age tracers. Following the evolution of
-fraction of young and old SNe~Ia as predicted by \cite{rigault2018}, and
-assuming constant underlying stretch
-distribution for each population \mr{as a function of redshift}, model the expected SN~Ia stretch
-distribution as a function of redshift. We test our prediction against data
-from the literature chosen to have negligible selection effects to ensure
-that any observed drift is indeed astrophysical and not observational.}
-{We clearly demonstrate that the underlying SNe~Ia stretch distribution is
-evolving as a function of redshift, and that the young/old drifting model is
-a much better description of the data than any non-drifting model, including
-the sample-based asymmetric distributions used by the Beams with Bias
-Correction algorithm to correct Malmquist bias. Our favored underlying
+The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
+survey statistics increase, the question of astrophysical systematic
+uncertainties rises, notably that of the SN~Ia population redshift drift. In
+this paper, we study the redshift evolution of SN~Ia \textsc{\texttt{SALT2.4}}
+lightcurve stretch, a purely intrinsic SN property, to probe this drift. The SN
+stretch has been shown to strongly correlate with the SN's environment, notably
+with stellar age tracers. Following the evolution of fraction of young and old
+SNe~Ia as predicted by \cite{rigault2018}, and assuming constant underlying
+stretch distribution for each population as a function of redshift, model the
+expected SN~Ia stretch distribution as a function of redshift. We test our
+prediction against data from the literature chosen to have negligible selection
+effects to ensure that any observed drift is indeed astrophysical and not
+observational. We clearly demonstrate that the underlying SNe~Ia stretch
+distribution is evolving as a function of redshift, and that the young/old
+drifting model is a much better description of the data than any non-drifting
+model, including the sample-based asymmetric distributions used by the Beams
+with Bias Correction algorithm to correct Malmquist bias. Our favored underlying
 stretch model is bimodal: a high-stretch mode shared by both young and old
-environments, and a low-stretch mode exclusive to old environments.}
-\nn{The
-precise impact of the redshift evolution of the underlying SNe Ia population on
-cosmology remains to be studied}. Yet, the astrophysical drift of the SN stretch
+environments, and a low-stretch mode exclusive to old environments. The precise
+impact of the redshift evolution of the underlying SNe Ia population on
+cosmology remains to be studied. Yet, the astrophysical drift of the SN stretch
 distribution does affect current Malmquist bias corrections and thereby
@@
 a bias will increase with surveys covering larger redshift ranges, which is
-particularly important for LSST.}
+particularly important for LSST.
+}
 
@@
 \citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
-the SNe~Ia are anchored on Cepheids \citep[the SH0ES
-project,][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
-from cosmic microwave background (CMB) data measured by Planck assuming the
-standard $\L$CDM \citep{planck2018,riess2019,reid2019}, or when the SN luminosity is anchored
-at intermediate redshift by the baryon acoustic oscillation (BAO) scale \citep{feeney2019}. While using the tip of the
-red giant branch technique in place of the Cepheids seem to favor an
-intermediate value of H$_0$ \citep{freedman2019,freedman2020}, time delay
-measurements from strong lensing seem to also favor high H$_0$ values
-\citep{wong2019}.
+the SNe~Ia are anchored on Cepheids \citep[the SH0ES project,][]{riess2009,
+riess2016} is $\sim5\sigma$ higher than what is predicted from cosmic microwave
+background (CMB) data measured by Planck assuming the standard $\L$CDM
+\citep{planck2018, riess2019, reid2019}, or when the SN luminosity is anchored
+at intermediate redshift by the baryon acoustic oscillation (BAO) scale
+\citep{feeney2019}. While using the tip of the red giant branch technique in
+place of the Cepheids seem to favor an intermediate value of H$_0$
+\citep{freedman2019, freedman2020}, time delay measurements from strong lensing
+seem to also favor high H$_0$ values \citep{wong2019}.
 
@@
 fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
-tension when accounting for all other probes \citep{knox2019}, but see e.g. \cite{poulin2019}.
-Alternatively, systematic effects affecting one or several of the aforementioned
-analyses could also explain the tension. \cite{rigault2015} suggested that SNe
-Ia from the Cepheid calibrator sample differ by construction from the Hubble
-flow \nn{sample} ones as \nn{the former} strongly
-favors young stellar population\nn{s}, while the latter not. This selection effect would impact the derivation of H$_0$ if
-SNe~Ia from young and older environments differ in average standardized
-magnitudes.
+tension when accounting for all other probes \citep{knox2019}, but see e.g.
+\cite{poulin2019}. Alternatively, systematic effects affecting one or several
+of the aforementioned analyses could also explain the tension.
+\cite{rigault2015} suggested that SNe Ia from the Cepheid calibrator sample
+differ by construction from the Hubble flow sample ones as the former
+strongly favors young stellar populations, while the latter not. This
+selection effect would impact the derivation of H$_0$ if SNe~Ia from young and
+older environments differ in average standardized magnitudes.
 
@@
 brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
-betoule2014, rigault2018, kim19}.  This mass-step correction is currently used
-in cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including
-for deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
+betoule2014, rigault2018, kim19}. This mass-step correction is currently used in
+cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including for
+deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
 between the SNe and their host remains unclear when using global properties such
@@
 showing that local age tracers such as the Local specific Star Formation Rate
-\nn{(LsSFR)} or the local color are more strongly correlated with the standardized
-SN magnitude \citep{rigault2018, roman2018, kim18}, suggesting age as the
-driving parameter underlying the mass-step. If true, this would have a
-significant impact for cosmology, \mr{since the environmental correction to apply to SN standardization could strongly vary with redshift.} \citep{rigault2013, childress2014, scolnic2018a}. Yet, the
-importance of local SN environmental studies remains highly debated
-\cite[e.g.][]{jones2015, jones2019} and especially the impact \mr{of such an astrophysical bias} has on the derivation of
+(LsSFR) or the local color are more strongly correlated with the standardized SN
+magnitude \citep{rigault2018, roman2018, kim18}, suggesting age as the driving
+parameter underlying the mass-step. If true, this would have a significant
+impact for cosmology, since the environmental correction to apply to SN
+standardization could strongly vary with redshift. \citep{rigault2013,
+childress2014, scolnic2018a}. Yet, the importance of local SN environmental
+studies remains highly debated \cite[e.g.][]{jones2015, jones2019} and
+especially the impact of such an astrophysical bias has on the derivation of
 H$_0$ \citep{jones2015, riess2016, riess2018, rose2019}.
@@
 The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate.
-\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
-    shown that the relative SNe~Ia rate in galaxies could be explained if two
-    populations existed, one young, following the host star formation activity,
-    and one old following the host stellar mass (the so called ``prompt and
-    delayed'' or ``A+B'' model). In \cite{rigault2018} we used the LsSFR to classify which
-    are the younger (those with a high LsSFR) and which are the older (those
-    with low LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN
-    stretch has been known to be strongly correlated with the SN host properties
-    \citep{hamuy1996, hamuy2000}, correlation that has been extensively confirmed since \citep[e.g.][]{neill2009,
-sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
-rigault2013, pan2014, kim19}. Following the ``A+B'' model and the connection
-between SN stretch and host properties, \cite{howell2007} first discussed the
-potential redshift drift of the SN stretch distribution. In this paper we
-revisit this analysis with the most recent SNe~Ia dataset.
+\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
+the relative SNe~Ia rate in galaxies could be explained if two populations
+existed, one young, following the host star formation activity, and one old
+following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
+model). In \cite{rigault2018} we used the LsSFR to classify which are the
+younger (those with a high LsSFR) and which are the older (those with low
+LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN stretch has
+been known to be strongly correlated with the SN host properties
+\citep{hamuy1996, hamuy2000}, correlation that has been extensively confirmed
+since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
+gupta2011, dandrea2011, childress2013, rigault2013, pan2014, kim19}. Following
+the ``A+B'' model and the connection between SN stretch and host properties,
+\cite{howell2007} first discussed the potential redshift drift of the SN stretch
+distribution. In this paper we revisit this analysis with the most recent SNe~Ia
+dataset.
 
@@
 having on average lower lightcurve stretches and being brighter after
-standardization. We use the correlation between the SN age, as probed by the LsSFR, and the SN stretch to model
-the expected evolution of the underlying SN stretch distribution as a function
-of redshift. This modeling relies on three assumptions: (1) there are two
-distinct populations of SNe~Ia; (2) the relative fraction of each of these
-populations as a function of redshift follows the model presented in
-\cite{rigault2018} and (3) the underlying distribution of stretch for each age
-sample is constant.
-This paper aims at testing this specific model with datasets from the literature.
-\mr{Note that the progenitor age as traced by the LsSFR seems to capture physical feature intrinsic to the progenitor and/or explosion mechanism that the stretch alone is not capturing \citep{nordin2018}.}
-
-We present in section~\ref{sec:sample} the sample we are using for this
+standardization. We use the correlation between the SN age, as probed by the
+LsSFR, and the SN stretch to model the expected evolution of the underlying SN
+stretch distribution as a function of redshift. This modeling relies on three
+assumptions: (1) there are two distinct populations of SNe~Ia; (2) the relative
+fraction of each of these populations as a function of redshift follows the
+model presented in \cite{rigault2018} and (3) the underlying distribution of
+stretch for each age sample is constant. This paper aims at testing this
+specific model with datasets from the literature. Note that the progenitor age
+as traced by the LsSFR seems to capture physical feature intrinsic to the
+progenitor and/or explosion mechanism that the stretch alone is not capturing
+\citep{nordin2018}.
+
+We present in Section~\ref{sec:sample} the sample we are using for this
 analysis, derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
@@
 underlying SNe Ia distribution, and how we build one from the Pantheon sample.
-We then present in section~\ref{sec:modeling} our modeling of the distribution of stretch and our results are presented in
-section~\ref{sec:results}.
-In this section we test whether the SN stretch distribution evolves as a function of redshift and if the aforementioned age model is in good agreement with this evolution.
-We briefly discuss these results in the context of SN cosmology in section~\ref{sec:bbc} and we conclude in
-section~\ref{sec:ccl}.
+We then present in Section~\ref{sec:modeling} our modeling of the distribution
+of stretch and our results are presented in Section~\ref{sec:results}. In this
+section we test whether the SN stretch distribution evolves as a function of
+redshift and if the aforementioned age model is in good agreement with this
+evolution. We briefly discuss these results in the context of SN cosmology in
+Section~\ref{sec:discussion} and we conclude in Section~\ref{sec:ccl}.
 
@@
 
-We base our analysis on the most recent comprehensive SNe~Ia
-compilation, the Pantheon catalog from \cite{scolnic2018a}. A naive approach to
-test the SN stretch redshift drift would be to simply compare the observed SN
-stretch distributions in a few bins of redshift. In practice, however, \mr{differential} selection effects are affecting the observed SN stretch
-distributions. Indeed, because the observed SN~Ia magnitude correlates with the
-lightcurve stretch (and color), the first SNe~Ia that a magnitude-limited survey
-will miss are the lowest-stretch (and reddest) ones. Consequently, if
-magnitude-related selection effects are not accounted for, one might confuse
-true population drift with survey properties, and conversely.
+We base our analysis on the most recent comprehensive SNe~Ia compilation, the
+Pantheon catalog from \cite{scolnic2018a}. A naive approach to test the SN
+stretch redshift drift would be to simply compare the observed SN stretch
+distributions in a few bins of redshift. In practice, however, differential
+selection effects are affecting the observed SN stretch distributions. Indeed,
+because the observed SN~Ia magnitude correlates with the lightcurve stretch (and
+color), the first SNe~Ia that a magnitude-limited survey will miss are the
+lowest-stretch (and reddest) ones. Consequently, if magnitude-related selection
+effects are not accounted for, one might confuse true population drift with
+survey properties, and conversely.
 
@@
 and will be discarded from our analysis. Fortunately, modern SN cosmology
-samples such as the Pantheon one are now dominated by magnitude-limited
-surveys.
+samples such as the Pantheon one are now dominated by magnitude-limited surveys.
 
@@
 in Bessel $B$ band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect
-this object typically 5 days before and a week after peak to build a suitable lightcurve, the effective limiting standardized
-absolute magnitude is approximately $M_{lim} = -18.00$~mag. Hence, given the
-magnitude limit $m_{lim}$ of a magnitude limited survey, one can derive the
-maximum redshift $z_{lim}$ above which the faintest SNe~Ia will be missed using
-the relation between apparent magnitude, redshift and absolute magnitude
-$\mu(z_{lim}) = m_{lim} - M_{lim}$.
-
-% YC: Figs and Tables *after* their 1st ref. in the text!
+this object typically 5 days before and a week after peak to build a suitable
+lightcurve, the effective limiting standardized absolute magnitude is
+approximately $M_{lim} = -18.00$~mag. Hence, given the magnitude limit $m_{lim}$
+of a magnitude limited survey, one can derive the maximum redshift $z_{lim}$
+above which the faintest SNe~Ia will be missed using the relation between
+apparent magnitude, redshift and absolute magnitude $\mu(z_{lim}) = m_{lim} -
+M_{lim}$.
+
 \begin{figure}
@@
         lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples of
-        the Pantheon catalog. The individual SNe are shown as blue dots. The ellipse $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by
-the corresponding standardized absolute magnitude using the $\alpha$ and $\beta$
-coefficients from \cite{scolnic2018a}. The grey diagonal lines represent the
-$(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and $z=1.70$
-using SNLS's $m_{lim}$ of $24.8$~mag.}
+        the Pantheon catalog. The individual SNe are shown as blue dots. The
+        ellipse $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the
+        corresponding standardized absolute magnitude using the $\alpha$ and
+        $\beta$ coefficients from \cite{scolnic2018a}. The grey diagonal lines
+        represent the $(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between
+        $0.50$ and $z=1.70$ using SNLS's $m_{lim}$ of $24.8$~mag.}
     \label{fig:maglim}
@@
 
-SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$; at these
+SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
 redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
@@
 \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
-\cite[][\mr{see their section~5}]{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both limits will be
-considered, as discussed below.
-
-Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
+\cite[][see their Section~5]{perrett2010} suggests a lower limit of
+$z_{lim}=0.55$; both limits will be considered, as discussed below.
+
+Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
 $5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
@@
 In a similar redshift range, SDSS has a limiting magnitude of 22.5
-\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
-SDSS surveys were more sensitive to limited spectroscopic resources;
-\cite[][\mr{see their section~2}]{kessler2009} pointed out that during the first year of SDSS, SNe~Ia with
-$r-mag<20.5$ were favored for spectroscopic follow-up, corresponding to a
-redshift cut at $0.15$. For the rest of the SDSS survey, additional
+\citep{dilday2008, sako2008}, which would lead to a $z_{lim}=0.24$. However, the
+SDSS surveys were more sensitive to limited spectroscopic resources; \cite[][see
+their Section~2]{kessler2009} pointed out that during the first year of SDSS,
+SNe~Ia with $r-mag<20.5$ were favored for spectroscopic follow-up, corresponding
+to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
 spectroscopic resources were available, and \cite{kessler2009} and
-\cite{dilday2008} show a \mr{reasonable} completeness up to
-$z_{lim}=0.2$. Following these analyses, we will use $z_{lim}=0.2$ as the
-baseline SDSS redshift limit.
-
-The sample selection is summarized in Table~\ref{tab:sample}, and the redshift distribution of these three surveys is shown in Fig.~\ref{fig:cuts}.
-As expected, the selected redshift limits roughly correspond to the peak of
-these histograms.
-\mr{We show in section~\ref{sec:results} that this sample selection indeed provide a subset of SNe~Ia with insignificant selection effects when compared to state-of-the-art Malmquist correction techniques.}
+\cite{dilday2008} show a reasonable completeness up to $z_{lim}=0.2$. Following
+these analyses, we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit.
+
+The sample selection is summarized in Table~\ref{tab:sample}, and the redshift
+distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
+expected, the selected redshift limits roughly correspond to the peak of these
+histograms. We show in Section~\ref{sec:results} that this sample selection
+indeed provides a subset of SNe~Ia with insignificant selection effects when
+compared to state-of-the-art Malmquist correction techniques.
 
@@
     \begin{tabular}{l c c}
-    \hline\hline%\\[-0.8em] % YC: don't, it's not your business to do that!
-        Survey & $z_{lim}$ & $N_{\mathrm{SN}}$ \\%[0.15em]
-        \hline%\\[-0.8em]
-        SNf & -- & 114\\%[0.30em]
-        SDSS & 0.20 (0.15) & 167 (82)\\%[0.30em]
-        PS1 & 0.30 (0.27) & 160 (122)\\%[0.30em]        
-        SNLS & 0.60 (0.55) & 102 (78)\\%[0.30em]
-        HST & -- & 26\\%[0.30em]
-        \hline%\\[-0.8em]
-        Total & -- & 569 (422)\\%[0.30em]
+        \hline\hline
+        Survey & $z_{lim}$ & $N_{\mathrm{SN}}$ \\
+        \hline
+        SNf   & --          & 114 \\
+        SDSS  & 0.20 (0.15) & 167 (82)\\
+        PS1   & 0.30 (0.27) & 160 (122)\\
+        SNLS  & 0.60 (0.55) & 102 (78)\\
+        HST   & --          & 26 \\
+        \hline
+        Total & --          & 569 (422) \\
         \hline
@@
         of SNe~Ia kept in our analysis for they are supposedly free from
-        selection bias (see section~\ref{sec:sample}). The darker (resp. lighter)
+        selection bias (see Section~\ref{sec:sample}). The darker (resp. lighter)
         color responds to the conservative (resp. fiducial) selection cut.}
@@
 been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
-and~4.2.2; \mr{SNe~Ia time series are published in \citealt{Saunders2020}, see also, Aldering in prep. and Ponder in prep.)}. The
-spectroscopic follow-up was done over a redshift range of $0.02 < z <
-0.09$~\citep[as in][]{rigault2018}, while the search was much deeper. As such,
-these SNe~Ia are assumed to be a random sampling of the underlying SN
-population. The SNfactory sample is particularly useful for studying SN property
-drift, as it enables us to have a large SN~Ia sample at $z<0.1$.
+and~4.2.2; SNe~Ia time series are published in \citealt{Saunders2020}
+\nn{\textit{Greg: for now you may need to cite SNEMO and SUGAR to remove this
+``?''}}, see also Aldering in prep. and Ponder in prep.). The spectroscopic
+follow-up was done over a redshift range of $0.02 < z < 0.09$~\citep[as
+in][]{rigault2018}, while the search was much deeper. As such, these SNe~Ia are
+assumed to be a random sampling of the underlying SN population. The SNfactory
+sample is particularly useful for studying SN property drift, as it enables us
+to have a large SN~Ia sample at $z<0.1$.
 
@@
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{lim}$ in Fig.~\ref{fig:sample}.
-\mr{The evolution of the mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are split in redshift bins of regular sample size.
-We see that SNe~Ia at higher redshift have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch distribution is drifting.}
+up to their respective $z_{lim}$ in Fig.~\ref{fig:sample}. The evolution of
+the mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are
+split in redshift bins of regular sample size. We see that SNe~Ia at higher
+redshift have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than
+those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the
+underlying stretch distribution is drifting.
 
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb-cividis.pdf}
+
     \caption{\textit{Bottom:} \textsc{\texttt{SALT2.4}} lightcurve stretch as a
         function of redshift for each survey considered in this analysis (see
-        legend). Solid (resp. open) markers correspond to the conservative
-        (resp. fiducial) redshift cuts. \textit{Top:} combined redshift
-    histogram in dark (resp. light) gray for the conservative (resp. fiducial)
-redshift cuts.}
+        legend). Solid (resp.\ open) markers correspond to the conservative
+        (resp.\ fiducial) redshift cuts. \textit{Top:} combined redshift
+        histogram in dark (resp.\ light) gray for the conservative (resp.
+        \ fiducial) redshift cuts.}
+
     \label{fig:sample}
@@
 In short, it was assumed that the number of ``young'' SNe~Ia follows the star
-formation \nn{rate} (SFR) in the Universe, while the number of ``old'' SNe~Ia
+formation rate (SFR) in the Universe, while the number of ``old'' SNe~Ia
 follows the number of Gyr-old stars in the Universe, i.e. the stellar mass
@@
 
-To model the evolution of the \nn{full} SN stretch distribution as a function of redshift, given our aforementioned model of the evolution of the fraction of
-younger and older SNe~Ia with cosmic time, we need to model the SN stretch distribution for each age
-subsample.
+To model the evolution of the full SN stretch distribution as a function of
+redshift, given our aforementioned model of the evolution of the fraction of
+younger and older SNe~Ia with cosmic time, we need to model the SN stretch
+distribution for each age subsample.
+
+\cite{rigault2018} presented the relation between SN stretch and LsSFR
+measurement, a progenitor age tracer, using the SNfactory sample. This relation
+is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory SNe used in the
+current analysis. Given the structure of the stretch-LsSFR scatter plot, our
+model of the underlying SN~Ia stretch distribution is defined as follows:
 
@@
 
-\cite{rigault2018} presented the relation between SN stretch and LsSFR
-measurement, a progenitor age tracer, using the SNfactory sample. This relation
-is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory SNe used in the
-current analysis. Given the structure of the stretch-LsSFR scatter plot, our
-model of the underlying SN~Ia stretch distribution is defined as follows:
-
 \begin{itemize}
@@
 
-\mr{The stretch probability distribution function (pdf) of a given SN will be the linear combination of the stretch distributions of these two population weighted by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
-But generally, the fraction of young SNIa as a function of redshift is given by $\delta$(z) (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model of the underlying distribution of SNIa as a function of redshift $X_1(z)$ is given by}:
+The stretch probability distribution function (pdf) of a given SN will be the
+linear combination of the stretch distributions of these two population weighted
+by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
+But generally, the fraction of young SNe~Ia as a function of redshift is given
+by $\delta$(z) (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
+of the underlying distribution of SNe~Ia as a function of redshift $X_1(z)$ is
+given by:
 \begin{multline}\label{eq:stretchz}
@@
 
-\subsection{Base model applied to data}
-\label{sec:basemodelapplied}
-
-Given the probability $y^i$ that a given SN is young, and assuming our Base (see section~\ref{sec:basemodel}), the probability to measure a \textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
+\subsection{Base model applied to data}\label{sec:basemodelapplied}
+
+Given the probability $y^i$ that a given SN is young, and assuming our Base (see
+Section~\ref{sec:basemodel}), the probability to measure a
+\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
 \begin{align}\label{eq:likelihoodsnf}
@@
     & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
-    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg],
+    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg]
 \end{align}
@@
 \begin{equation}\label{eq:likelihood}
-    -2\ln(L) = -2 \sum_i \ln \prob{x^{i}_{1}}{\vec{\theta};
-    \mathrm{d}x^{i}_{1}, y^i}.
+    -2\ln(L) = -2 \sum_i \ln \prob{x_1^i}{\vec{\theta};
+    \mathrm{d}x_1^i, y^i}.
 \end{equation}
 
-Depending on whether $y^i$ can be estimated directly from LsSFR measurements or not, there are two ways to proceed, which we now discuss.
+Depending on whether $y^i$ can be estimated directly from LsSFR measurements or
+not, there are two ways to proceed, which we now discuss.
 
@@
 predicted redshift evolution of mean stretch (mean of Eq.~\ref{eq:stretchz})
-illustrated as a blue band in Fig.~\ref{fig:modelall}. 
-We see in this figure that the measured mean SN~Ia stretch per redshift bins of equal sample size closely follows our redshift drift modeling.
-This is indeed what is expected if old environments favor low SN stretches
-\citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
-a function of redshift. See section~\ref{sec:results} for a more quantitative
-discussion.
+illustrated as a blue band in Fig.~\ref{fig:modelall}. We see in this figure
+that the measured mean SN~Ia stretch per redshift bins of equal sample size
+closely follows our redshift drift modeling. This is indeed what is expected if
+old environments favor low SN stretches \citep[e.g.][]{howell2007} \emph{and} if
+the fraction of old SNe~Ia declines as a function of redshift. See
+Section~\ref{sec:results} for a more quantitative discussion.
 
@@
 
-In section~\ref{sec:basemodel}, we have modeled the underlying stretch
+In Section~\ref{sec:basemodel}, we have modeled the underlying stretch
 distribution following \cite{rigault2018}, i.e.\ as a single Gaussian for the
@@
 also adjust to the data following the procedure described in
-section~\ref{sec:modelnopy}.
+Section~\ref{sec:modelnopy}.
 
@@
 behind this sample-based approach is twofold: (1) Malmquist biases are driven by
-survey properties and (2) since current surveys cover limited redshift ranges, having a sample-based approach covers
-some potential redshift evolution information \citep{scolnic2016, scolnic2018a}.
-See further discussion concerning BBC in section~\ref{sec:bbc}.
+survey properties and (2) since current surveys cover limited redshift ranges,
+having a sample-based approach covers some potential redshift evolution
+information \citep{scolnic2016, scolnic2018a}. See further discussion
+concerning BBC in Section~\ref{sec:discussion}.
 
@@
 We adjusted each of the models described above on both the fiducial and
-conservative samples (cf. section~\ref{sec:sample}); results are gathered
+conservative samples (cf. Section~\ref{sec:sample}); results are gathered
 in Table~\ref{tab:comp}, and illustrated in Fig.~\ref{fig:mod_comp}.
-
-Since the various models have different degrees of freedom, we use the Akaike
-Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare their ability
-to properly describe the observations. This estimator penalizes extra degrees of
-freedom to avoid over-fitting the data, and is defined as follow:
-\begin{equation}
-    \mathrm{AIC} = -2\ln(L) + 2k
-\end{equation}
-where $-2\ln(L)$ \mr{is derived by minimizing} Eq.~\eqref{eq:likelihood}, and $k$ is the number of free parameters to be
-adjusted. The reference model is the one with the smallest AIC; in comparison to
-this model, the models with $\Delta\mathrm{AIC}<5$ are coined acceptable,
-the ones with $5<\Delta\mathrm{AIC}<20$ are unfavored, and those with
-$\Delta\mathrm{AIC}>20$ are deemed excluded. This roughly corresponds to 2,
-3 and 5~$\sigma$ limits for a Gaussian probability distribution.
 
@@
         not, its number of free parameters and, for both the fiducial and the
-        conservative cuts, $-2\ln(L)$ (see
-    Eq.~\ref{eq:likelihood}), the AIC and the AIC difference ($\Delta$AIC)
-between this model and the Base model used as reference for it has the lowest
-AIC.}
+        conservative cuts, $-2\ln(L)$ (see Eq.~\ref{eq:likelihood}), the AIC and
+        the AIC difference ($\Delta$AIC) between this model and the Base model
+        used as reference for it has the lowest AIC.}
     \label{tab:comp}
@@
         ($\Delta\mathrm{AIC} < -20$), see text. According to the AIC, all
-    non-drifting models (open symbols) are excluded to be as good a representation
-of the data as the Base (drifting) model.}
+        non-drifting models (open symbols) are excluded to be as good a
+        representation of the data as the Base (drifting) model.}
     \label{fig:mod_comp}
@@
 
+Since the various models have different degrees of freedom, we use the Akaike
+Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare their ability
+to properly describe the observations. This estimator penalizes extra degrees of
+freedom to avoid over-fitting the data, and is defined as follow:
+\begin{equation}
+    \mathrm{AIC} = -2\ln(L) + 2k
+\end{equation}
+where $-2\ln(L)$ is derived by minimizing Eq.~\eqref{eq:likelihood}, and $k$ is
+the number of free parameters to be adjusted. The reference model is the one
+with the smallest AIC; in comparison to this model, the models with
+$\Delta\mathrm{AIC}<5$ are coined acceptable, the ones with
+$5<\Delta\mathrm{AIC}<20$ are unfavored, and those with $\Delta\mathrm{AIC}>20$
+are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
+a Gaussian probability distribution.
+
 We first note that the best model (with smallest AIC) is the so-called Base
 model, implementing the \cite{rigault2018} description of the stretch
-distribution and its redshift evolution. This is therefore the reference model, both on the fiducial and conservative samples. 
+distribution and its redshift evolution. This is therefore the reference model,
+both on the fiducial and conservative samples.
+
 % $\chi^2=-2\ln(L)-\ln(2\pi\sigma)$ ~ 900 so 1.6 per DoF.
@@
 Furthermore, we find that redshift-independent stretch distributions are all
-excluded as \nn{suitable} descriptions of the data relative to the Base model.
-In fact, the best non-drifting model (the Asymmetric one) has a very marginal
+excluded as suitable descriptions of the data relative to the Base model.  In
+fact, the best non-drifting model (the Asymmetric one) has a very marginal
 chance ($p \equiv \exp\left(\Delta\mathrm{AIC}/2\right) = 5\times10^{-6}$) to
@@
 implementations of the BBC technique \citep{scolnic2016, kessler2017}, is one of
-the worst modelisations of our analysis (see section~\ref{sec:results}). While
-its $-2\ln(L)$ is the smallest of all redshift-independent models (but still $-11.5$ worse than the reference Base model), it is
-strongly penalized for requiring 15~free parameters ($\mu_0, \sigma_{\pm}$~for
-each of the 5~samples of the analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be
-interpreted as a probability $p=2\times 10^{-7}$ of being an as good representation of the data as the Base model.
-We report in Table~\ref{tab:bbc} the
-samples' $\mu_0$ and $\sigma_{\pm}$ adjusted on the \nn{nominally}
-selection-free samples using our fiducial cuts (see section~\ref{sec:sample}).
-We find our results in close agreement with \cite{scolnic2016} for SNLS and SDSS
-and with \cite{scolnic2018a} for PS1, who derived these model parameters using
-the full BBC formalism, using numerous simulations to model the selection
-effects (see details e.g., section~3 of \citealt{kessler2017}). The agreement
-between our fit of the asymmetric Gaussians on the supposedly selection-free
-part of the samples and the results derived using the BBC formalism supports our
-approach to get a sample with negligible selection effects. 
-If we were to use \cite{scolnic2016} and \cite{scolnic2018a}
-best fit values \nn{of the $\mu_0, \sigma_{\pm}$ asymmetric parameters} for SNLS, SDSS and PS1, respectively,
-the $\Delta$AIC between our Base drifting model and the BBC
-modeling would go even deeper from $-32$ to $-47$.
-We further discuss the consequence of this result for cosmology in
-section~\ref{sec:discussion}.
+the worst modelisations of our analysis (see Section~\ref{sec:results}). While
+its $-2\ln(L)$ is the smallest of all redshift-independent models (but still
+$-11.5$ worse than the reference Base model), it is strongly penalized for
+requiring 15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the 5~samples of
+the analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted
+as a probability $p=2\times 10^{-7}$ of being an as good representation of the
+data as the Base model. We report in Table~\ref{tab:bbc} the samples' $\mu_0$
+and $\sigma_{\pm}$ adjusted on the nominally selection-free samples using our
+fiducial cuts (see Section~\ref{sec:sample}). We find our results in close
+agreement with \cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a}
+for PS1, who derived these model parameters using the full BBC formalism, using
+numerous simulations to model the selection effects (see details e.g., Section~3
+of \citealt{kessler2017}). The agreement between our fit of the asymmetric
+Gaussians on the supposedly selection-free part of the samples and the results
+derived using the BBC formalism supports our approach to get a sample with
+negligible selection effects. If we were to use \cite{scolnic2016} and
+\cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
+parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
+Base drifting model and the BBC modeling would go even deeper from $-32$ to
+$-47$. We further discuss the consequence of this result for cosmology in
+Section~\ref{sec:discussion}.
 
@@
 
-We also performed tests allowing
-the high-stretch mode of the old population to differ from the young population mode, hence adding two degrees of freedom. The corresponding fit is not
-significantly better, with a $\Delta$AIC of $-0.4$. 
-this strengthens the fact
-that the young and old populations indeed appear to share the same underlying
-high-stretch mode. 
-\mr{Furthermore, one might wonder whether a low-stretch mode might also exist in the young-population, see Fig.~\ref{fig:stretchlssfr}.
-We tested that by allowing this population to also be bi-modal, finding the amplitude of this low-stretch mode to be compatible with~0 in this young population ($<2\%$).
-More generally, this raises the question of inaccurate tracing of age by a given environmental tracer (here the LsSFR).
-A dedicated analysis be presented in Briday et al.
-in prep.}
-
-\mr{Finally, considering the SNfactory dataset like any other
-(i.e., ignoring the LsSFR measurements, see section~\ref{sec:modeling}),
-reduces the significance the results presented in this section, as expected.
-Yet, non-drifting models remain strongly disfavored, and for instance, the best fitted sample-based gaussian asymmetric modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the data than our Base drifting modeling.}
+We also performed tests allowing the high-stretch mode of the old population to
+differ from the young population mode, hence adding two degrees of freedom. The
+corresponding fit is not significantly better, with a $\Delta$AIC of $-0.4$.
+this strengthens the fact that the young and old populations indeed appear to
+share the same underlying high-stretch mode.  Furthermore, one might wonder
+whether a low-stretch mode might also exist in the young-population, see
+Fig.~\ref{fig:stretchlssfr}. We tested that by allowing this population to also
+be bi-modal, finding the amplitude of this low-stretch mode to be compatible
+with~0 in this young population ($<2\%$). More generally, this raises the
+question of inaccurate tracing of age by a given environmental tracer (here the
+LsSFR). A dedicated analysis will be presented in Briday et al.\ in prep.
+
+Finally, \sout{considering the SNfactory dataset like any other (i.e.,} ignoring
+the LsSFR measurements \nn{-- available only for the SNfactory dataset}, see
+Section~\ref{sec:modeling} -- \sout{),} reduces the significance of the results
+presented in this section, as expected. Yet, non-drifting models remain strongly
+disfavored, and for instance, the best fitted sample-based Gaussian asymmetric
+modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the data than
+our Base drifting modeling.
 
@@
 
-To the best of our knowledge, a SN~Ia stretch redshift drift \mr{modeling} has never been explicitly \mr{used} in cosmological analyses, though bayesian hierarchy
+To the best of our knowledge, a SN~Ia stretch redshift drift modeling has never
+been explicitly used in cosmological analyses, though bayesian hierarchy
 formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016} or Steve
-\citep{hinton2019} can easily allow it; see e.g., section 1.3 and 2.5 of \cite{rubin2015}.
-Not doing so is a second order issue for
-SN cosmology, as it only affects the way one accounts for Malmquist bias. Indeed, as long as Phillips' relation \citep{phillips1993} standardization parameter
-$\alpha$ is not redshift dependent (a study behind the scope of
-this paper, but see e.g. \citealt{scolnic2018a}), the stretch-corrected SNe~Ia
-magnitudes used for cosmology are blind to the underlying stretch distribution for complete samples. 
-However, surveys usually do have
-significant Malmquist bias at least for the upper half of their SN redshift distribution.
-As a consequence, an ill-modeling of the underlying stretch distribution will bias the SN magnitude derived from such surveys.
-
-Commonly used Malmquist bias correction techniques, such as the BBC-formalism, assume per Sample Asymmetric Gaussian functions for modeling the underlying stretch and color distributions.
-Yet, as shown in Section~\ref{sec:results}, such a sample-based distribution is excluded as being as good as our drifting model.
-Then, unlike what \citet[][section~2]{scolnic2016} and
-\citet[][section~5.4]{scolnic2018a} suggested, i.e. that
-traditional surveys span limited redshift ranges and that therefore the per-sample approach accounts for implicit redshift drifts, \sout{might not be correct} \nn{a direct modeling of the redshift drift is more appropriate than a sample-based approach}.
-We stress here that, as measurements of modern surveys try to cover increasingly larger redshift ranges in order to reduce calibration systematic uncertainties, this sample-based approach is becoming less valid, notably for PS1, DES and, soon, LSST.
+\citep{hinton2019} can easily allow it; see e.g., section 1.3 and 2.5 of
+\cite{rubin2015}. Not doing so is a second order issue for SN cosmology, as it
+only affects the way one accounts for Malmquist bias. Indeed, as long as
+Phillips' relation \citep{phillips1993} standardization parameter $\alpha$ is
+not redshift dependent (a study behind the scope of this paper, but see e.g.
+\citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
+cosmology are blind to the underlying stretch distribution for complete samples.
+However, surveys usually do have significant Malmquist bias at least for the
+upper half of their SN redshift distribution. As a consequence, an ill-modeling
+of the underlying stretch distribution will bias the SN magnitude derived from
+such surveys.
+
+Commonly used Malmquist bias correction techniques, such as the BBC-formalism,
+assume per Sample Asymmetric Gaussian functions for modeling the underlying
+stretch and color distributions. Yet, as shown in Section~\ref{sec:results},
+such a sample-based distribution is excluded as being as good as our drifting
+model. Then, unlike what \citet[][Section~2]{scolnic2016} and
+\citet[][Section~5.4]{scolnic2018a} suggested, i.e. that traditional surveys
+span limited redshift ranges and that therefore the per-sample approach accounts
+for implicit redshift drifts, a direct modeling of the redshift drift is more
+appropriate than a sample-based approach. We stress here that, as measurements
+of modern surveys try to cover increasingly larger redshift ranges in order to
+reduce calibration systematic uncertainties, this sample-based approach is
+becoming less valid, notably for PS1, DES and, soon, LSST.
 
 We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
-underlying stretch distribution between the per-sample asymmetric modeling and our Base drifting
-model for the PS1 sample. Our model is bimodal and the relative amplitude of
-each mode depends on the redshift-dependent fraction of old and young SNe~Ia in
-the sample: the higher the fraction of old SNe~Ia (at lower redshift), the
-higher the amplitude of the old-specific low-stretch mode. This redshift
-dependency is shown as blue to red underlying distributions in
+underlying stretch distribution between the per-sample asymmetric modeling and
+our Base drifting model for the PS1 sample. Our model is bimodal and the
+relative amplitude of each mode depends on the redshift-dependent fraction of
+old and young SNe~Ia in the sample: the higher the fraction of old SNe~Ia (at
+lower redshift), the higher the amplitude of the old-specific low-stretch mode.
+This redshift dependency is shown as blue to red underlying distributions in
 Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges covered by PS1. The observed
-$x_1$ histogram follows our modeling defined as the sum of individual underlying SN-redshift distributions.
-As expected, the two modeling approaches differ mostly in the negative part of the SN stretch distribution.
-The asymmetric Gaussian distribution goes through the middle of the
-bimodal distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
-under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN redshifts.
-\mr{This means that the SN bias-corrected standardized magnitude estimated at redshift affected by selection effect would be biased by an ill-modeling of the true underlying stretch distribution.}
+$x_1$ histogram follows our modeling defined as the sum of individual underlying
+SN-redshift distributions. As expected, the two modeling approaches differ
+mostly in the negative part of the SN stretch distribution. The asymmetric
+Gaussian distribution goes through the middle of the bimodal distribution,
+over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and under-estimating it at
+$x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN
+redshifts. This means that the SN bias-corrected standardized magnitude
+estimated at redshifts affected by selection effect would be biased by an
+ill-modeling of the true underlying stretch distribution.
 
@@
         illustration, we show as colored (from blue to red with increasing
-redshifts) the evolution of the underlying stretch distribution as a function of
-redshift for the redshift range covered by PS1 data.}
+        redshifts) the evolution of the underlying stretch distribution as a
+        function of redshift for the redshift range covered by PS1 data.}
     \label{fig:bbc_pdf_ps1}
@@
 
-The amplitude of this magnitude bias for cosmology is beyond the scope of this paper given the complexity of the BBC analysis. It would require a full
-study using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
+The amplitude of this magnitude bias for cosmology is beyond the scope of this
+paper given the complexity of the BBC analysis. It would require a full study
+using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
 asymmetric modeling as part of the BBC simulations.
@@
 a function of redshift when considering that the underlying SN stretch
-distribution follows either the per-sample Asymmetric Gaussian modeling or our Base drifting model. This
-is shown Fig.~\ref{fig:magdrift}, where, for clarity we convert this difference
-in mean stretch $x_1$ to a difference in magnitude using $\alpha=0.156$
-\citep{scolnic2018a}.
-\mr{We stress here that this does not correspond to the actual impact on cosmological distances derived from SNeIa as this simple stretch average comparison does not account for the amplitude of the Malmquist bias correction to be applied as a function of redshift and surveys.}
+distribution follows either the per-sample Asymmetric Gaussian modeling or our
+Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
+we convert this difference in mean stretch $x_1$ to a difference in magnitude
+using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
+correspond to the actual impact on cosmological distances derived from SNe~Ia as
+this simple stretch average comparison does not account for the amplitude of the
+Malmquist bias correction to be applied as a function of redshift and surveys.
 
@@
     \caption{Difference of mean stretch standardization (in mag, using
-        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample Gaussian asymmetric modeling
-        and our Base drifting model, as a function of redshift, considering the
-        SDSS, PS1 and SNLS datasets. The colored markers show the difference for
-        SDSS, PS1 and SNLS samples (see legend) considering the individual SN
-    redshifts (no redshift cut applied). The orange markers represent their mean
-    evolution in 30 regularly-spaced redshift bins. 
-    \mr{This must be multiplied by the amplitude of the Malmquist bias correction to derive the SN distance modulus bias affecting cosmology, see text.}}
+        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
+        Gaussian asymmetric modeling and our Base drifting model, as a function
+        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
+        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
+        considering the individual SN redshifts (no redshift cut applied). The
+        orange markers represent their mean evolution in 30 regularly-spaced
+        redshift bins. This must be multiplied by the amplitude of the Malmquist
+        bias correction to derive the SN distance modulus bias affecting
+        cosmology, see text.}
     \label{fig:magdrift}
@@
 
-We see that the amplitude of the modeling difference is a few tens of millimag. In the era of modern cosmology, where we aim at probing $w_0$ at a
-sub-percent level and $w_a$ at the ten-percent precision
-\citep[e.g.,][]{lsstpaper}, we stress that correctly modeling of potential SN redshift drift should be further studied and care should be made when using samples affected by selection effects.
+We see that the amplitude of the modeling difference is a few tens of millimag.
+In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
+level and $w_a$ at the ten-percent precision \citep[e.g.,][]{lsstpaper}, we
+stress that correctly modeling of potential SN redshift drift should be further
+studied and care should be made when using samples affected by selection
+effects.
 
@@
 
-We have presented a study of the drift of the \mr{underlying} SNe~Ia stretch distribution as a function of redshift. We
-built a magnitude-limited SN~Ia sample from the Pantheon dataset
-\citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to which we added HST and SNfactory
-data from \cite{rigault2018} for the high- and low-redshift bins. We only
-considered the SNe that have been discovered in the redshift range of each
-survey where selection effects are negligible, so that the observed SNe~Ia
-stretches are random sampling of the true underlying distribution. This
-resulted in a 569 SN~Ia fiducial sample (422 SNe when more
+We have presented a study of the drift of the underlying SNe~Ia stretch
+distribution as a function of redshift. We built a magnitude-limited SN~Ia
+sample from the Pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to
+which we added HST and SNfactory data from \cite{rigault2018} for the high- and
+low-redshift bins. We only considered the SNe that have been discovered in the
+redshift range of each survey where selection effects are negligible, so that
+the observed SNe~Ia stretches are random sampling of the true underlying
+distribution. This resulted in a 569 SN~Ia fiducial sample (422 SNe when more
 conservative cuts were considered).
@@
 ``Beams with Bias Correction'', a Malmquist bias correction algorithm used by
-\nn{several} recent SN~Ia cosmological analyses, implemented as a per-sample
+several recent SN~Ia cosmological analyses, implemented as a per-sample
 asymmetric Gaussian distribution.
@@
 
-    \item The underlying SN~Ia stretch distribution is         redshift dependent, as previously suggested by e.g.~\cite{howell2007}. This
-        result is largely independent of details on each age-population
-        modeling.
-
-    \item Non-drifting models are indeed excluded as \nn{suitable} descriptions of the
-        data \nn{relative to} our Base model. This model assumes
-        that: (1) the younger population has a unimodal Gaussian stretch
-        distribution, while the older population stretch distribution
-        is bimodal, one mode being the same as the young one; (2) the
-        evolution of the relative fraction of younger and older SNe~Ia
-        follows the prediction made in \cite{rigault2018}.
-
-    \item 
-        \mr{Using} survey-based asymmetric Gaussian distributions, as  done, e.g., in the current implementation of the BBC, is excluded to be good descriptions of the data \nn{relative to} our drifting model.
-Hence, the sample-based approach does not accurately account for redshift drift and even less so as survey span larger redshift ranges.
-\mr{We warn that, even if extra degrees of freedom might be acceptable SN cosmological studies, extrapolating the SN property distributions from the volume-limited part of a survey to the magnitude-limited one affected by sample selection would still be inaccurate since the distribution would drift.}
+    \item The underlying SN~Ia stretch distribution is redshift dependent, as
+        previously suggested by e.g.~\cite{howell2007}. This result is largely
+        independent of details on each age-population modeling.
+
+    \item Non-drifting models are indeed excluded as suitable descriptions of
+        the data relative to our Base model. This model assumes that: (1) the
+        younger population has a unimodal Gaussian stretch distribution, while
+        the older population stretch distribution is bimodal, one mode being the
+        same as the young one; (2) the evolution of the relative fraction of
+        younger and older SNe~Ia follows the prediction made in
+        \cite{rigault2018}.
+
+    \item Using survey-based asymmetric Gaussian distributions, as  done, e.g.,
+        in the current implementation of the BBC, is excluded to be good
+        descriptions of the data relative to our drifting model. Hence, the
+        sample-based approach does not accurately account for redshift drift and
+        even less so as survey span larger redshift ranges. We warn that, even
+        if extra degrees of freedom might be acceptable \nn{given the large
+        number of SNe~Ia in} cosmological studies, extrapolating the SN property
+        distributions from the volume-limited part of a survey to the
+        magnitude-limited one affected by sample selection would still be
+        inaccurate since the distribution would drift.
 
     \item An inaccurate underlying stretch distribution model is estimated to
-        bias the mean standardized magnitude of SNe~Ia affected \nn{by}
+        bias the mean standardized magnitude of SNe~Ia affected by
         selection effects at the order of a few percents. This is
@@
 
-\nn{In this paper, we considered} a simple Gaussian mixture modeling, but
-additional data free from significant Malmquist bias would enable us to refine
-it. We note that samples at the low- and high-redshift
-ends of the Hubble diagram would be particularly helpful for this drifting
-analysis; fortunately this will soon be provided by the Zwicky Transient Facility \citep{bellm2019, graham2019} and Subaru and SeeChange SNe~Ia programs, respectively.
+In this paper, we considered a simple Gaussian mixture modeling, but additional
+data free from significant Malmquist bias would enable us to refine it. We note
+that samples at the low- and high-redshift ends of the Hubble diagram would be
+particularly helpful for this drifting analysis; fortunately this will soon be
+provided by the Zwicky Transient Facility \citep{bellm2019, graham2019} and
+Subaru and SeeChange SNe~Ia programs, respectively.
 
@@
     (grant agreement n 759194 - USNAC).
-    This work was supported in part by the Director, Office of Science, Office of High Energy Physics of the U.S. Department of Energy under Contract No. DE-AC025CH11231.
-    % How has received that support for this project ? 
-    %Support has been provided by the Institut Universitaire de France, the CNES, and the region Auvergne-Rhone-Alpes.
+    This work was supported in part by the Director, Office of Science, Office
+    of High Energy Physics of the U.S.  Department of Energy under Contract No.
+    DE-AC025CH11231.
 \end{acknowledgements}
@@
 
-%For clarity with the next models, we named it 3G2M2S$_{\text{SNf}}$ for it has a total of 3 gaussians but with only 2 means and 2 standard deviations, and has been fitted on SNf data only.
+%For clarity with the next models, we named it 3G2M2S$_{\text{SNf}}$ for it has a total of 3 Gaussians but with only 2 means and 2 standard deviations, and has been fitted on SNf data only.
 
@@
     \item 2G2M2S, model taken from HOWELL 2009 where we added $\delta(z)$;
-    \item 3G3M3S, with three independent gaussians.
+    \item 3G3M3S, with three independent Gaussians.
 \end{itemize}

### 2020-04-23 Nicolas.tex
Missing commit of Nora changes before MR calll

--- 2020-04-17 Nicolas.tex
@@
 \abstract{
-{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as survey statistics increase, the question of
-astrophysical systematic uncertainties rises,  notably that of
+{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
+\nn{\sout{measure of}} survey statistic\nn{s} increase, the question of
+astrophysical systematic uncertainties rises, \nn{\sout{and}} notably that of
 the SN~Ia population redshift drift.}
@@
 probe this drift.}
-{The SN stretch has been shown to strongly correlate with the SN's environment, notably with stellar age tracers. Following the evolution of
+{The SN stretch has been shown to strongly correlate with the SN's environment,
+\nn{\sout{and}} notably with stellar age tracers. Following the evolution of
 fraction of young and old SNe~Ia as predicted by \cite{rigault2018}, and
-assuming constant underlying stretch
-distribution for each population \mr{as a function of redshift}, model the expected SN~Ia stretch
+assuming non-drifting \yc{$\to$constant/stationary?} underlying stretch
+distribution for each population, model the expected SN~Ia stretch
 distribution as a function of redshift. We test our prediction against data
@@
 environments, and a low-stretch mode exclusive to old environments.}
-\nn{The
+{\sout{The underlying SNe~Ia population is evolving as a function of redshift
+    \yc{This is purely a repetition of a previous sentence} \nn{\textit{see new
+sentence}} and the exact impact on cosmology remains to be studied.} \nn{The
 precise impact of the redshift evolution of the underlying SNe Ia population on
@@
 particularly important for LSST.}
+}
 
@@
 \section{Introduction}
+
+\yc{\textit{YC: you should run a grammar corrector, e.g. \texttt{grammarly}.
+e.g. "SNe dichotomy" vs. "SN dichotomy", "analysis" vs. "analyses", etc.}}
+\nn{\textit{By the way, as we capitalize ``Fig.'' before a reference, shouldn't
+we capitalize ``Section'' before reference?}}
 
@@
 project,][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
-from cosmic microwave background (CMB) data measured by Planck assuming the
-standard $\L$CDM \citep{planck2018,riess2019,reid2019}, or when the SN luminosity is anchored
-at intermediate redshift by the baryon acoustic oscillation (BAO) scale \citep{feeney2019}. While using the tip of the
+\yc{from} cosmic microwave background (CMB) data measured by Planck assuming the
+standard $\L$CDM \yc{ref?} \nn{\textit{isn't it at the end of the sentence? Do
+you think it should be before the "or"?}}, or when the SN luminosity is anchored
+at intermediate redshift by the baryon acoustic oscillation (BAO) scale
+\citep{riess2019,reid2019,planck2018, feeney2019}. While using the tip of the
 red giant branch technique in place of the Cepheids seem to favor an
@@
 fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
-tension when accounting for all other probes \citep{knox2019}, but see e.g. \cite{poulin2019}.
+tension when accounting for all other probes \citep{knox2019} and \nn{\sout{the
+current most promising}} \nn{one currently possible} scenario appears to be a
+burst of expansion at the matter-radiation decoupling moment caused by a (fine
+tuned) early dark energy \citep{poulin2019}. \yc{\textit{I don't see the point
+of going into this debate...}}
+
 Alternatively, systematic effects affecting one or several of the aforementioned
@@
 Ia from the Cepheid calibrator sample differ by construction from the Hubble
-flow \nn{sample} ones as \nn{the former} strongly
-favors young stellar population\nn{s}, while the latter not. This selection effect would impact the derivation of H$_0$ if
+flow \nn{sample} ones as \nn{\sout{it \yc{they?}}} \nn{the former} strongly
+favors young stellar population\nn{s} \nn{\sout{environments where one could
+find Cepheids}}. This selection effect would impact the derivation of H$_0$ if
 SNe~Ia from young and older environments differ in average standardized
@@
 driving parameter underlying the mass-step. If true, this would have a
-significant impact for cosmology, \mr{since the environmental correction to apply to SN standardization could strongly vary with redshift.} \citep{rigault2013, childress2014, scolnic2018a}. Yet, the
+significant impact for cosmology\nn{, not only because there would be a redshift
+drift magnitude correction to apply, but also bevause it might strongly
+vary} \yc{\textit{just the existence of a redshift-drift mag.\ correction is
+of importance here, even if it does not vary}} \nn{\textit{now taken
+into account}} \citep{rigault2013, childress2014, scolnic2018a}. Yet, the
 importance of local SN environmental studies remains highly debated
-\cite[e.g.][]{jones2015, jones2019} and especially the impact \mr{of such an astrophysical bias} has on the derivation of
+\cite[e.g.][]{jones2015, jones2019} and especially the impact it
+\yc{\textit{what does `it' refer to, the `importance'?}} \nn{\textit{no, to the
+SN environment. Do you think it needs rephrasing?}} has on the derivation of
 H$_0$ \citep{jones2015, riess2016, riess2018, rose2019}.
 
-%\nn{We point that there has been sample-based approach meant to account for an unmodeled variation of host-related parameters thanks to multiple redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show aren't as effective as a direct modeling of such a drift}.
-
-The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate. 
-\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
+\yc{\textit{Shouldn't this paragraph, introducing the concept of old and young
+populations and lsSFR, be moved \emph{\nn{here}}?}} \nn{\textit{I agree. Now
+done.}}
+
+\nn{The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia
+    rate. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
     shown that the relative SNe~Ia rate in galaxies could be explained if two
@@
     and one old following the host stellar mass (the so called ``prompt and
-    delayed'' or ``A+B'' model). In \cite{rigault2018} we used the LsSFR to classify which
+    delayed'' or ``A+B'' model). In \cite{rigault2018} we used the \nnn{LsSFR
+    (}specific Star Formation Rate at the SN location\nnn{)} to classify which
     are the younger (those with a high LsSFR) and which are the older (those
@@
     stretch has been known to be strongly correlated with the SN host properties
-    \citep{hamuy1996, hamuy2000}, correlation that has been extensively confirmed since \citep[e.g.][]{neill2009,
+    \citep{hamuy1996, hamuy2000}, \nnn{correlation that} \nnn{\sout{which}}
+    \yc{\textit{what does this ``which'' refer to?}} \nnn{\textit{is that
+    better?}} has been extensively confirmed since \citep[e.g.][]{neill2009,
 sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
@@
 potential redshift drift of the SN stretch distribution. In this paper we
-revisit this analysis with the most recent SNe~Ia dataset.
+revisit this analysis with the most recent SNe~Ia dataset.}
 
@@
 having on average lower lightcurve stretches and being brighter after
-standardization. We use the correlation between the SN age, as probed by the LsSFR, and the SN stretch to model
+standardization. We use the correlation between the SN age, as probed by local
+specific star formation rate \citep{rigault2018}, and the SN stretch to model
 the expected evolution of the underlying SN stretch distribution as a function
@@
 \cite{rigault2018} and (3) the underlying distribution of stretch for each age
-sample is constant. This paper aims at testing this specific model with datasets from the literature. We highlight
+sample \sout{\yc{is constant}} \nn{is a fixed model} and does not significantly
+evolve with time. This paper aims at testing this specific model with datasets
+from the literature.
 
@@
 underlying SNe Ia distribution, and how we build one from the Pantheon sample.
-We then present in section~\ref{sec:modeling} our modeling of the distribution of stretch and our results are presented in
-section~\ref{sec:results}.
-In this section we test whether the SN stretch distribution evolves as a function of redshift and if the aforementioned age model is in good agreement with this evolution.
-We briefly discuss these results in the context of SN cosmology in section~\ref{sec:bbc} and we conclude in
+We then present in section~\ref{sec:modeling} our modeling of the distribution
+of stretch\nn{, using \sout{as a function of}} redshift\nn{-}based models for
+the stretch distribution of young and old SNe~Ia \yc{\textit{what??? rephrase}}
+\nn{\textit{Is that better?}}. Our results are presented in
+section~\ref{sec:results}, where we test whether the SN stretch distribution
+evolves as a function of redshift and if the aforementioned age model is in good
+agreement with this evolution. Consequences for cosmology of our results are
+briefly discussed in section~\ref{sec:bbc} and we conclude in
 section~\ref{sec:ccl}.
@@
 
-We base our analysis on the most recent comprehensive SNe~Ia
+We base our analysis on the most recent \yc{comprehensive}\nn{\textit{What does
+it mean? Are there SNe~Ia compilation that are not comprehensive?}} SNe~Ia
 compilation, the Pantheon catalog from \cite{scolnic2018a}. A naive approach to
 test the SN stretch redshift drift would be to simply compare the observed SN
-stretch distributions in a few bins of redshift. In practice, however, \mr{differential} selection effects are affecting the observed SN stretch
+stretch distributions in a few bins of redshift. In practice, however,
+\yc{differential} \nn{\textit{Do you mean ``different'', or that it is based on
+        differences (the definition of ``differential'' as given by my search
+engine)?}} selection effects are affecting the observed SN stretch
 distributions. Indeed, because the observed SN~Ia magnitude correlates with the
@@
 in Bessel $B$ band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect
-this object typically 5 days before and a week after peak to build a suitable lightcurve, the effective limiting standardized
+this object typically \nn{\sout{a week} 5 days before and \sout{10 days} a week}
+after peak to build a suitable lightcurve, the effective limiting standardized
 absolute magnitude is approximately $M_{lim} = -18.00$~mag. Hence, given the
@@
         lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples of
-        the Pantheon catalog. The individual SNe are shown as blue dots. The ellipse $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by
+        the Pantheon catalog. The individual SNe are shown as blue dots
+        \nn{\sout{and a 2D histogram is shown in gray to highlight point
+        density}}. \yc{\textit{I don't think it's really needed}} \nn{\textit{Ok,
+changed. Better?}}. The ellipse $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by
 the corresponding standardized absolute magnitude using the $\alpha$ and $\beta$
@@
 spectroscopic resources were available, and \cite{kessler2009} and
-\cite{dilday2008} show a \mr{reasonable} completeness up to
+\cite{dilday2008} show a relative completeness \yc{\textit{what is a
+\emph{relative completeness}???}} \nn{\textit{In Dilday 2008, they give a median
+redshift of <z> = .22 for spectroscopically confirmed SNe~Ia}} up to
 $z_{lim}=0.2$. Following these analyses, we will use $z_{lim}=0.2$ as the
@@
 
-The sample selection is summarized in Table~\ref{tab:sample}, and the redshift distribution of these three surveys is shown in Fig.~\ref{fig:cuts}.
+\nn{\sout{For the rest of the analysis, we will consider the aforementioned redshift
+limits, as well as more conservative values for each of these surveys, namely
+$z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS1 and $z_{lim}=0.55$ for SNLS.}}
+\yc{why do you remind the conservative values, and not the fiducial ones? Just
+refer to Table~1.}
+
+The complete sample selection is summarized in Table~\ref{tab:sample}, and the
+redshift distribution of these three surveys is shown in Fig.~\ref{fig:cuts}.
 As expected, the selected redshift limits roughly correspond to the peak of
@@
 been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
-and~4.2.2; \mr{SNe~Ia time series are published in \citealt{Saunders2020}, see also, Aldering in prep. and Ponder in prep.)}. The
+and~4.2.2; \mr{SNe~Ia time series are published in \citealt{Saunders2020}}
+\nn{should add Aldering in prep. and Pander in prep. as suggested by Greg}). The
 spectroscopic follow-up was done over a redshift range of $0.02 < z <
@@
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{lim}$ in Fig.~\ref{fig:sample}.
-\mr{The evolution of the mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are split in redshift bins of regular sample size.
-We see that SNe~Ia at higher redshift have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch distribution is drifting.}
+up to their respective $z_{lim}$ in Fig.~\ref{fig:sample}. \nn{The combined
+samples' mean stretches in bins of redshift is also presented
+in Fig.~\ref{fig:modelall}, where we see that when considering
+selection-bias-free SN~Ia samples, SNe~Ia at higher redshift have on average
+larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift
+($-0.17\pm 0.10$ at $z\sim0.05$). That is the evolution that we attempt to model
+in the next section.} \nnn{\textit{Last part added following Yannick's comment
+about the fact that this result is model-independent; see last paragraph before
+section 3.2}}
+
+\yc{\textit{You should also present/refer to Fig.5, showing the significant
+evolution of mean stretch as function of $z$ in complete samples. Next Sec. aims
+at modeling this evolution.}} \nn{\textit{Done.}}
 
@@
 
-\subsection{``Base'' underlying stretch distribution}
-\label{sec:basemodel}
-
-To model the evolution of the \nn{full} SN stretch distribution as a function of redshift, given our aforementioned model of the evolution of the fraction of
-younger and older SNe~Ia with cosmic time, we need to model the SN stretch distribution for each age
-subsample.
+\subsection{``Base'' underlying stretch distribution}\label{sec:basemodel}
+
+\yc{\textit{There is a generic confusion between ``SN stretch'' (individual SN,
+as in Eq.3/4) and something like ``SN stretch distribution'' (population, as in
+Eq.2).  This is confusing...}}
+
+To model the evolution of the \nn{global} SN stretch distribution as a function
+of redshift, given our aforementioned model of the evolution of the fraction of
+younger and older SNe~Ia with \yc{\sout{redshift} cosmic time} \nn{\textit{Is it
+        really common to call $z$ the ``cosmic time''? What does this
+modification bring to the understanding of this evolution? We might make the
+confusion between the ``age'' of a SNe, and the ``time'' of its explosion given
+by the redshift...}}, we need to model the SN stretch distribution for each age
+subsample. \yc{\textit{That paragraph sounds like a tautology...}}
+\nn{\textit{Added ``global'' so that we see its meaning}}
+
+\yc{\textit{I suggest we remove the use of ``we'' when referring to previous papers,
+like ``In R+18, we did...'' to be replaced by ``R+18 did...''}} \nn{\textit{That
+has been a recurring question. This paper is part of a series of paper, so it
+seemed ok, but for that I'll trust the common way, though I feel like lying by
+making it sound impartial when we actually use ``our'' results. People should
+know that we take our previous work as a baseline.}}
+
+\cite{rigault2018} presented the relation between SN stretch and LsSFR
+measurement, a progenitor age tracer, using the SNfactory sample. This relation
+is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory SNe used in the
+current analysis. Given the structure of the stretch-LsSFR scatter plot, our
+model of the underlying SN~Ia stretch distribution is defined as follows:
+
+\begin{itemize}
+    \item for the younger population (i.e., $\log(\mathrm{LsSFR})\geq-10.82$),
+        the stretch distribution is modeled as a single normal distribution
+        $\mathcal{N}(\mu_1, \sigma_1{}^2)$; 
+    \item the older population (i.e., $\log( \mathrm{LsSFR})<-10.82$) stretch
+        distribution is modeled as a bimodal Gaussian mixture $a\times
+        \mathcal{N}(\mu_1, \sigma_1{}^2) + (1-a)\times \mathcal{N}(\mu_2,
+        \sigma_2{}^2)$, where one mode is the same as for the young population,
+        $a$ representing the relative influence of the two modes.
+\end{itemize}
 
@@
 
-\cite{rigault2018} presented the relation between SN stretch and LsSFR
-measurement, a progenitor age tracer, using the SNfactory sample. This relation
-is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory SNe used in the
-current analysis. Given the structure of the stretch-LsSFR scatter plot, our
-model of the underlying SN~Ia stretch distribution is defined as follows:
-
-\begin{itemize}
-    \item for the younger population (i.e., $\log(\mathrm{LsSFR})\geq-10.82$),
-        the stretch distribution is modeled as a single normal distribution
-        $\mathcal{N}(\mu_1, \sigma_1{}^2)$; 
-    \item the older population (i.e., $\log( \mathrm{LsSFR})<-10.82$) stretch
-        distribution is modeled as a bimodal Gaussian mixture $a\times
-        \mathcal{N}(\mu_1, \sigma_1{}^2) + (1-a)\times \mathcal{N}(\mu_2,
-        \sigma_2{}^2)$, where one mode is the same as for the young population,
-        $a$ representing the relative influence of the two modes.
-\end{itemize}
-
-\mr{The stretch probability distribution function (pdf) of a given SN will be the linear combination of the stretch distributions of these two population weighted by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
-But generally, the fraction of young SNIa as a function of redshift is given by $\delta$(z) (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model of the underlying distribution of SNIa as a function of redshift $X_1(z)$ is given by}:
+When combined to the previously described prediction of the evolution of younger
+SNe~Ia as a function of redshift (Eq.~\ref{eq:delta}), our model $X_1(z)$ for
+the underlying stretch distribution as a function of redshift is written
 \begin{multline}\label{eq:stretchz}
@@
 
-\subsection{Base model applied to data}
-\label{sec:basemodelapplied}
-
-Given the probability $y^i$ that a given SN is young, and assuming our Base (see section~\ref{sec:basemodel}), the probability to measure a \textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
+\yc{\textit{In fact, not clear what this eq. does here: it is supposed to be
+superseded by Eq.3/4. Maybe it should be moved to Sec.3.1.2 when we use
+$\delta(z)$ as a proxy to $y^i$.}} 
+\nn{\textit{It should not! The evolution as a function of $z$ \underline{doesn't
+depend on how we determine the parameters}. \textbf{This} is what people should
+use for the ``Base'' model, and \textbf{then} there a two ways to get the
+parameters. If you think there is a clearer way for people to understand this,
+please tell us. All 3 eq. (2), (3) and (4) are of the same importance and are
+used in the .py we wrote.}}
+
+The maximum-likelihood estimate of the 5~free parameters
+$\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained
+by minimizing the following pseudo-$\chi^2$:
+\begin{equation}\label{eq:likelihood}
+    -2\ln(L) = -2 \sum_i \ln \prob{x^{i}_{1}}{\vec{\theta};
+    \mathrm{d}x^{i}_{1}, y^i},
+\end{equation}
+with
 \begin{align}\label{eq:likelihoodsnf}
@@
 
-The maximum-likelihood estimate of the 5~free parameters
-$\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained
-by minimizing the following:
-\begin{equation}\label{eq:likelihood}
-    -2\ln(L) = -2 \sum_i \ln \prob{x^{i}_{1}}{\vec{\theta};
-    \mathrm{d}x^{i}_{1}, y^i}.
-\end{equation}
-
-Depending on whether $y^i$ can be estimated directly from LsSFR measurements or not, there are two ways to proceed, which we now discuss.
+where $i$ is the index of the SN~Ia, $x^i_1$, $\mathrm{d}x^i_1$ and $y^i$ are
+the \textsc{\texttt{SALT2.4}} stretch, its associated error and the probability
+that the SN is young, respectively. In practice, to ensure fraction $a$ is
+constrained between 0 and 1, we fit for $\alpha$ such that
+$a=\arctan(\alpha)/\pi+0.5$ \yc{\sout{, which results in an asymmetric error on
+$a$}}. \yc{\textit{Given the fact that 1. it's barely asymmetric ($a\sim 0.5$),
+2. no one cares, it should be symmetrized (quadratically). Probably $\sigma$
+errors are not symmetric either!}} \nn{\textit{Mickaël, do you have arguments to
+maintain an asymmetric definition of $a$?}}
+
+Depending on whether $y^i$ can be estimated directly from LsSFR measurements or
+not, there are two ways to proceed, which we now discuss.
 
@@
 predicted redshift evolution of mean stretch (mean of Eq.~\ref{eq:stretchz})
-illustrated as a blue band in Fig.~\ref{fig:modelall}. 
-We see in this figure that the measured mean SN~Ia stretch per redshift bins of equal sample size closely follows our redshift drift modeling.
-This is indeed what is expected if old environments favor low SN stretches
+illustrated in Fig.~\ref{fig:modelall}. \yc{\textit{I'm confused: grey points in
+Fig.5 are model independent, aren't they? Only the curve is related to the
+model}}. \nn{\textit{Yes. What is confusing?}} We see in this figure that the
+measured mean SN~Ia stretch per bin of redshift\nn{, defined so that each
+contain an equal number of SNe of the fiducial sample,} closely follows our
+redshift drift modeling\nn{\sout{; that is, when considering
+selection-bias-free SN~Ia samples, SNe~Ia at higher redshift have on
+average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at
+lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$)}, and namely that SNe~Ia at
+higher redshift have a larger average stretch ($0.34\pm 0.10$ at
+$z\sim0.65$) than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$)}.
+\yc{\textit{This is model-independant, and a nice result by itself, and should
+be already mentioned right at the end of Sec.2}}. \nn{\textit{I agree. I dumbly
+copied that right before Sec.3, and reworded that part for a reminder}}. This is
+indeed what is expected if old environments favor low SN stretches
 \citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
@@
 behind this sample-based approach is twofold: (1) Malmquist biases are driven by
-survey properties and (2) since current surveys cover limited redshift ranges, having a sample-based approach covers
+survey properties and (2) since current surveys cover limited redshift ranges,
+doing so \yc{\textit{what?}} \nn{\textit{having a sample-based approach}} covers
 some potential redshift evolution information \citep{scolnic2016, scolnic2018a}.
@@
 \begin{equation}
-    \mathrm{AIC} = -2\ln(L) + 2k
+    \mathrm{AIC} = -2\ln(L)_{min} + 2k
 \end{equation}
-where $-2\ln(L)$ \mr{is derived by minimizing} Eq.~\eqref{eq:likelihood}, and $k$ is the number of free parameters to be
+where $-2\ln(L)_{min}$ is the minimum value of the pseudo-$\chi^2$ as defined
+Eq.~\eqref{eq:likelihood}, and $k$ is the number of free parameters to be
 adjusted. The reference model is the one with the smallest AIC; in comparison to
@@
         not, its number of free parameters and, for both the fiducial and the
-        conservative cuts, $-2\ln(L)$ (see
+        conservative cuts, the pseudo-$\chi^2$, $-2\ln(L)$ (see
     Eq.~\ref{eq:likelihood}), the AIC and the AIC difference ($\Delta$AIC)
@@
 
-        Per Sample Asym. & per sample & 3$\times$5
+        BBC-like & per sample & 3$\times$5
         & 1468.2 & 1498.2 & $-31.5$
@@
 model, implementing the \cite{rigault2018} description of the stretch
-distribution and its redshift evolution. This is therefore the reference model, both on the fiducial and conservative samples. 
-% $\chi^2=-2\ln(L)-\ln(2\pi\sigma)$ ~ 900 so 1.6 per DoF.
+distribution and its redshift evolution. This is therefore the reference model,
+both on the fiducial \yc{\textit{somewhere, you still use the ``ref. sample''
+terminology!}} \nn{\textit{corrected, thanks.}} and conservative samples.
 
@@
 the worst modelisations of our analysis (see section~\ref{sec:results}). While
-its $-2\ln(L)$ is the smallest of all redshift-independent models (but still $-11.5$ worse than the reference Base model), it is
+its pseudo-$\chi^{2}$ is the smallest of all redshift-independent models (but
+still $\Delta\chi^2 = -11.5$ worse than the reference Base model), it is
 strongly penalized for requiring 15~free parameters ($\mu_0, \sigma_{\pm}$~for
-each of the 5~samples of the analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be
-interpreted as a probability $p=2\times 10^{-7}$ of being an as good representation of the data as the Base model.
-We report in Table~\ref{tab:bbc} the
+each of the 5~samples of the analysis). We report in Table~\ref{tab:bbc} the
 samples' $\mu_0$ and $\sigma_{\pm}$ adjusted on the \nn{nominally}
@@
 part of the samples and the results derived using the BBC formalism supports our
-approach to get a sample with negligible selection effects. 
-If we were to use \cite{scolnic2016} and \cite{scolnic2018a}
-best fit values \nn{of the $\mu_0, \sigma_{\pm}$ asymmetric parameters} for SNLS, SDSS and PS1, respectively,
-the $\Delta$AIC between our Base drifting model and the BBC
-modeling would go even deeper from $-32$ to $-47$.
+approach to get a sample with negligible selection effects. \yc{\textit{It's not
+very clear what you mean to convey in this sentence.}} \nn{\textit{Basically,
+because we get the same results but with a different approach, theirs being
+complicated and supposedly good for selection effects, we think it can show that
+ours (using the limiting magnitude blablabla) is at least as good.}}
+
 We further discuss the consequence of this result for cosmology in
@@
 
-We also performed tests allowing
-the high-stretch mode of the old population to differ from the young population mode, hence adding two degrees of freedom. The corresponding fit is not
-significantly better, with a $\Delta$AIC of $-0.4$. 
-this strengthens the fact
+For the sake of robustness \yc{???}\nn{???}, we also performed tests allowing
+the high-stretch mode of the old population to differ from the young population
+mode, hence adding two degrees of freedom. The corresponding fit is not
+significantly better, with a $\Delta$AIC of $-0.4$; this strengthens the fact
 that the young and old populations indeed appear to share the same underlying
-high-stretch mode. 
-\mr{Furthermore, one might wonder whether a low-stretch mode might also exist in the young-population, see Fig.~\ref{fig:stretchlssfr}.
-We tested that by allowing this population to also be bi-modal, finding the amplitude of this low-stretch mode to be compatible with~0 in this young population ($<2\%$).
-More generally, this rises the question of inaccurate tracing of age by a given environmental tracer (here the LsSFR).
-A dedicated analysis be presented in Briday et al.
-in prep.}
-
-\mr{Finally, considering the SNfactory dataset like any other
+high-stretch mode. We also allowed the young population to have a low-stretch
+mode \nn{\textit{Well I didn't! But I strongly think we should, as well as
+Martin, Jakob and now Greg too. We've let the 3 Gaussians to have
+independant $\mu$s and $\sigma$s, but never the young having the low-stretch,
+unless you did at some point. If so, it should be adressed earlier and compared,
+because Fig.~\ref{fig:stretchlssfr} invites us to think it does.}}, finding its
+amplitude to be compatible with~0 ($<2\% \pm \yc{???}$).
+
+\section{Discussion}\label{sec:discussion}
+
+To the best of our knowledge, the SN~Ia stretch redshift drift has never been
+explicitly accounted for in cosmological analyses, though bayesian hierarchy
+formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016} or Steve
+\citep{hinton2019} can easily allow it. Not doing so is a second order issue for
+SN cosmology, as it only affects the way one accounts for Malmquist bias.
+\yc{\textit{I don't understand why you say it's a 2nd order effect if all
+samples are significantly Malmquist biased!?}}. \nn{\textit{Yes, but is
+it taken into account?}} Indeed, as long as \nn{\sout{Phillip} Tripp
+\textit{right?}}'s relation \yc{\textit{ref?}} standardization parameter
+$\alpha$ is not intrinsically redshift dependent (a study behind the scope of
+this paper, but see e.g. \citealt{scolnic2018a}), the stretch-corrected SNe~Ia
+magnitudes used for cosmology are blind to the underlying stretch distribution
+\yc{for complete samples} \nn{\textit{Why is that? Even for sample not
+``complete'', if $\alpha$ doesn't depend on $z$, $\alpha*x_1$ won't see the
+underlying distribution, right?}}. However, all modern surveys do have
+significant Malmquist bias at least for the upper half of their SN redshift
+distribution. As a consequence, an ill-modeling of the underlying stretch
+distribution will bias the derived SN magnitude for these \yc{\textit{these
+what?}} \nn{\textit{These surveys. Reword?}} \citep[see e.g.,][]{rubin2015,
+rubin2016}. \sout{The fact that the currently-used sample-based asymmetric Gaussian
+distribution modeling is excluded \nn{as being as good as our drifting model is
+worrisome and is further discussed} this in the next sections.} \yc{\textit{I
+don't see the point of this kind of sentence...}} \nn{\textit{I guess that if
+the most common tool is biased, it has to be adressed, and this serves as a
+transition, but I think we could just go into the subsection directly.}}
+
+\subsection{BBC formalism and underlying stretch distributions}\label{sec:bbc}
+
+\yc{\textit{IMHO, you should take care of not converting this paper into an
+anti-BBC paper: just show the results, and let people make their opinion.
+Furthermore, I find this paragraph mostly a repetition of previous statements
+and results...}} \nn{\textit{Probably.}} \bigbreak
+
+\sout{The commonly used tool for doing Malmquist bias correction in SN cosmology
+is the BBC formalism described in} \cite{scolnic2016}\sout{ and}
+\cite{kessler2017}\sout{, which is used in recent SN cosmological analyses}
+\citep{jones2018b, scolnic2018a, brout2019, descosmopaper2019}
+\sout{including the direct measurement of $H_0$} \citep{riess2016,
+riess2019}. \yc{\textit{This all have already been said before, there's
+no point at repeating it IMHO. Maybe just a simpler introductory
+sentence without all the references?}} \nn{\textit{The first part yes, but
+not the second. See proposed paragraph below.}} \bigbreak
+
+\nn{The BBC formalism previously mentioned is used in some recent SN
+cosmological analyses \citep{jones2018b, scolnic2018a, brout2019,
+descosmopaper2019} including the direct measurement of $H_0$
+\citep{riess2016, riess2019}. Yet, as shown in Section~\ref{sec:results},
+their sample-based asymmetric Gaussian distribution modeling is excluded as
+being as good as our drifting model.}
+
+\sout{Yet, the current underlying stretch modeling of the BBC formalism, a
+sample-based asymmetric Gaussian distribution, is excluded to be an as good
+representation of the data than our Base drifting modeling, see
+section~\ref{sec:results}.} \yc{\textit{Again, unnecessary repetition.}}
+\nn{\textit{Here, I agree. I moved this part in the previous paragraph.}} Then,
+unlike what \citet[][section~2]{scolnic2016} and
+\citet[][section~5.4]{scolnic2018a} suggested, \nn{\sout{the fact that
+traditional surveys span limited redshift ranges, the per-sample approach
+cannot correctly account for implicit redshift drifts} namely that a per-sample
+approach might be able to account for unmodeled redshift drift thanks to the
+limited redshift span of traditional surveys, the results indicate that it is
+far from being as useful as directly modeling the redshift drift}. \yc{\textit{I
+don't understand what they suggested and what is wrong...  Rephrase.}}
+\nn{\textit{Done.}} We stress here that, as measurements of modern surveys
+try to cover increasingly larger redshift ranges in order to
+reduce calibration systematic uncertainties, this \nn{sample-based} approach
+\yc{\textit{which one?}} \nn{\textit{Done.}}
+is becoming less valid, notably for PS1, DES and, soon, LSST. Already with
+current surveys, the BBC sample-based technique is ruled out at representing
+the data as well as our best model: $\Delta\mathrm{AIC}<-20$, which could be
+interpreted as a probability $p=2\times 10^{-7}$ of being an as good
+representation of the data as the Base model. \yc{\textit{Again, unnecessary
+repetition.}} \nn{\textit{I think that with the rewording I done, it's ok. I
+don't find anything else to put after ``a probability of ...''.}}
+Note that even considering the SNfactory dataset like any other
 (i.e., ignoring the LsSFR measurements, see section~\ref{sec:modeling}),
-reduces the significance the results presented in this section, as expected.
-Yet, non-drifting models remain strongly disfavored, and for instance, the best fitted sample-based gaussian asymmetric modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the data than our Base drifting modeling.}
-
-\section{Discussion}\label{sec:discussion}
-
-To the best of our knowledge, a SN~Ia stretch redshift drift \mr{modeling} has never been explicitly \mr{used} in cosmological analyses, though bayesian hierarchy
-formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016} or Steve
-\citep{hinton2019} can easily allow it; see e.g., section 1.3 and 2.5 of \cite{rubin2015}.
-Not doing so is a second order issue for
-SN cosmology, as it only affects the way one accounts for Malmquist bias. Indeed, as long as Phillips' relation \citep{phillips1993} standardization parameter
-$\alpha$ is not redshift dependent (a study behind the scope of
-this paper, but see e.g. \citealt{scolnic2018a}), the stretch-corrected SNe~Ia
-magnitudes used for cosmology are blind to the underlying stretch distribution for complete samples. 
-However, surveys usually do have
-significant Malmquist bias at least for the upper half of their SN redshift distribution.
-As a consequence, an ill-modeling of the underlying stretch distribution will bias the SN magnitude derived from such surveys.
-
-\subsection{Issue of with the per Sample Asymmetric modeling}
-\label{sec:bbc}
-
-Commonly used Malmquist bias correction techniques, such as the BBC-formalism, assume per Sample Asymmetric Gaussian functions for modeling the underlying stretch and color distributions.
-Yet, as shown in Section~\ref{sec:results}, such a sample-based distribution is excluded as being as good as our drifting model.
-Then, unlike what \citet[][section~2]{scolnic2016} and
-\citet[][section~5.4]{scolnic2018a} suggested, i.e. that
-traditional surveys span limited redshift ranges and therefore that the per-sample approach accounts for implicit redshift drifts, might not be correct.
-We stress here that, as measurements of modern surveys try to cover increasingly larger redshift ranges in order to reduce calibration systematic uncertainties, this sample-based approach  is becoming less valid, notably for PS1, DES and, soon, LSST.
-
-We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
-underlying stretch distribution between the per-sample asymmetric modeling and our Base drifting
-model for the PS1 sample. Our model is bimodal and the relative amplitude of
-each mode depends on the redshift-dependent fraction of old and young SNe~Ia in
-the sample: the higher the fraction of old SNe~Ia (at lower redshift), the
-higher the amplitude of the old-specific low-stretch mode. This redshift
-dependency is shown as blue to red underlying distributions in
-Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges covered by PS1. The observed
-$x_1$ histogram follows our modeling defined as the sum of individual underlying SN-redshift distributions.
-As expected, the two modeling approaches differ mostly in the negative part of the SN stretch distribution.
-The asymmetric Gaussian distribution goes through the middle of the
-bimodal distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
-under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN redshifts.
-\mr{This means that the SN bias-corrected standardized magnitude estimated at redshift affected by selection effect would be biased by ill-modeling of the true underlying stretch distribution.}
+reduces this difference to $\Delta\mathrm{AIC}<-10$, still strongly
+disfavoring \nn{\sout{the modeling currently used as part of the \mr{sample-based
+formalism}} this modeling}.
+If we were to use \cite{scolnic2016} and \cite{scolnic2018a}
+best fit values \nn{of the $\mu_0, \sigma_{\pm}$ asymmetric parameters}
+for SNLS, SDSS and PS1, respectively \yc{\textit{which parameters?}}
+\nn{\textit{Now precised.}},
+the $\Delta$AIC between our Base drifting model and the BBC
+modeling would go even deeper from $-32$ to $-47$. \yc{\textit{But you would be
+using parameters derived specifically for incomplete samples on
+specially-constructed complete sample: what is the validity of the
+comparison???}} \nn{\textit{Well their parameters are made for incomplete
+samples, but \textbf{taking into account} the selection effects; we \textbf{got
+rid of them}, so technically, the underlying parameters should both represent
+those of the underlying distribution.}}
+
+\subsection{First step toward quantification of the cosmological impact of
+inaccurate stretch distribution modeling.}
 
@@
 
-The amplitude of this magnitude bias for cosmology is beyond the scope of this paper given the complexity of the BBC analysis. It would require a full
+We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
+underlying stretch distribution between the BBC modeling and our Base drifting
+model for the PS1 sample. Our model is bimodal and the relative amplitude of
+each mode depends on the redshift-dependent fraction of old and young SNe~Ia in
+the sample: the higher the fraction of old SNe~Ia (at lower redshift), the
+higher the amplitude of the old-specific low-stretch mode. This redshift
+dependency is shown as blue to red underlying distributions in
+Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges covered by PS1. The observed
+$x_1$ histogram\nn{, in grey,} corresponds to the sum of individual underlying
+SN-redshift distributions\nn{, represented by the black line.} \yc{\textit{Not
+clear: you mean the black curve -- a model -- or the histogram -- an
+observation?}} \nn{\textit{Is it clearer?}} As expected, the two approaches
+\yc{\textit{which ones?}} \nn{\textit{the BBC modeling and our Base
+drifting model, as introduec in the paragraph. Do you think we should repeat it?
+I feel that we've be talking about BBC vs.\ Base for so long it would be too
+muchr.}} strongly differ in modeling the negative part of the SN stretch
+distribution. The BBC asymmetric distribution goes through the middle of the
+bimodal distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
+under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting model
+for typical PS1 SN redshifts. \yc{\textit{This is definitely not obvious from the Fig.
+solely...}} \nn{\textit{I agree that both distributions are quite close to the
+histogram. Our black line may be closer to the bins at low redshift, that the
+green line is way too high at -0.9, but ours is way too low at =0.5 or so.}}
+This means that the bias-correction of individual SN stretches,
+$\overline{\delta}_{x_1}$ \citep{kessler2017}, and consequently the
+magnitude bias correction, $\mu_B$, is most likely inaccurate and redshift
+dependent.
+
+The amplitude of this magnitude bias for cosmology is beyond the scope of
+this paper given the complexity of the BBC analysis. It would require a full
 study using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
@@
 a function of redshift when considering that the underlying SN stretch
-distribution follows either the per-sample Asymmetric Gaussian modeling or our Base drifting model. This
+distribution follows either the BBC formalism or our Base drifting model. This
 is shown Fig.~\ref{fig:magdrift}, where, for clarity we convert this difference
@@
 \citep{scolnic2018a}.
-\mr{We strengthen here that this does not corresponds to the actual impact on cosmological distances derived from SNeIa depending on which underlying stretch distribution modeling as it does not account for the amplitude of the Malmquist bias correction as a function of redshift and surveys.}
 
@@
     \caption{Difference of mean stretch standardization (in mag, using
-        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample Gaussian asymmetric modeling
+        $\alpha=0.156$ from \citealt{scolnic2018a}) between the BBC algorithm
         and our Base drifting model, as a function of redshift, considering the
@@
     redshifts (no redshift cut applied). The orange markers represent their mean
-    evolution in 30 regularly-spaced redshift bins. 
-    \mr{This must be multiplied by the amplitude of the Malmquist bias correction to derive the SN distance modulus bias affecting cosmology, see text.}}
+evolution in 30 regularly-spaced redshift bins.}
     \label{fig:magdrift}
@@
 
-We see that the amplitude of the modeling difference is a few tens of millimag. In the era of modern cosmology, where we aim at probing $w_0$ at a
+We see that the amplitude of the modeling difference is a few tens of millimag,
+which corresponds to the expected bias affecting SNe~Ia impacted by sample
+selection effects. For comparison, the amplitude of this effect is similar in
+scale to the effect exotic forms of dark energy could have on $\mu(z)$.
+Consequently, in the era of modern cosmology, where we aim at probing $w_0$ at a
 sub-percent level and $w_a$ at the ten-percent precision
-\citep[e.g.,][]{lsstpaper}, we stress that correctly modeling of potential SN redshift drift should be further studied and care should be made when using samples affected by selection effects.
+\citep[e.g.,][]{lsstpaper}, it is of paramount importance to further understand
+the exact modeling of the lightcurve parameters when using SNe~Ia affected by
+Malmquist bias.
 
@@
 
-We have presented a study of the drift of the \mr{underlying} SNe~Ia stretch distribution as a function of redshift. We
+We have presented a study of the drift of the \yc{intrinsic} \nn{\textit{still
+up to debate I think. We've been using ``underlying'' for the most part of
+the paper...}} SNe~Ia stretch distribution as a function of redshift. We
 built a magnitude-limited SN~Ia sample from the Pantheon dataset
@@
 model which depends on the expected fraction of ``young'' and ``old'' SNe~Ia as
-a function of redshift, each age population having its own underlying
+a function of redshift, each age population having its own \yc{intrinsic}
 stretch distribution.
@@
 
-    \item The underlying SN~Ia stretch distribution is         redshift dependent, as previously suggested by e.g.~\cite{howell2007}. This
-        result is largely independent of details on each age-population
-        modeling. \yc{\textit{Indeed, and this should be stated first!}}
-
-    \item Non-drifting models are indeed excluded as \nn{suitable} descriptions of the
-        data \nn{relative to} our Base model. This model assumes
+    \item Non-drifting models are excluded as \nn{suitable} descriptions of the
+        data \nn{relative to} our Base model \yc{\textit{This does not mean
+        anything...}} \nn{\textit{Indeed. Better?}} \nn{\sout{This model},
+        which} \yc{\textit{which one?}} \nn{\textit{Now clearer?}} assumes
         that: (1) the younger population has a unimodal Gaussian stretch
@@
 
-    \item 
-        \mr{Using} survey-based asymmetric Gaussian distributions, as  done, e.g., in the current implementation of the BBC, is excluded to be good descriptions of the data \nn{relative to} our drifting model.
-Hence, the sample-based approach does not accurately
-        account for redshift drift.
+    \item Given 1., we conclude that the SNe~Ia stretch is indeed redshift
+        dependent, as previously suggested by e.g.~\cite{howell2007}. This
+        result is largely independent of details on each age-population
+        modeling. \yc{\textit{Indeed, and this should be stated first!}}
+
+    \item \yc{\textit{reread \& rephrase: ``assuming'' is not excluded!}}
+        \nn{Models} assuming survey-based asymmetric Gaussian distributions, as
+        done, e.g., in the current implementation of the BBC, \nn{\sout{is} are}
+        excluded to be good descriptions of the data \nn{relative to} our
+        drifting model. Hence, sample-based approach does not accurately
+        account for redshift drift, unlike what was suggested by
+        \cite{scolnic2016} and \cite{scolnic2018a}.
 
@@
         $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}) and using the
-        age-population drift modelization $\delta(z)$ defined in with $K=0.87$:
+        age-population drift modelization $\delta(z)$ defined in
+        \cite{rigault2018} \yc{\textit{too many R+18 everywhere!}} with $K=0.87$:
         \begin{align}
@@
 additional data free from significant Malmquist bias would enable us to refine
-it. We note that samples at the low- and high-redshift
+it. We \nn{\sout{highlight} note} that samples at the low- and high-redshift
 ends of the Hubble diagram would be particularly helpful for this drifting
-analysis; fortunately this will soon be provided by the Zwicky Transient Facility \citep{bellm2019, graham2019} and Subaru and SeeChange SNe~Ia programs, respectively.
+analysis; fortunately this will soon be provided by the Zwicky Transient
+Facility \citep{bellm2019, graham2019} and Subaru SNe~Ia program (REF HERE ASK
+NICOLAS), respectively. \nn{\textit{Suggestion of adding a reference to the
+upcoming See Chang paper with $\approx$ 25 SNe at z > 1.2}}
 
@@
 \bibitem[Perrett et al.(2010)]{perrett2010} Perrett, K., Balam, D., Sullivan, M., et al.\ 2010, \aj, 140, 518
-
-\bibitem[Phillips(1993)]{phillips1993} Phillips, M.~M.\ 1993, \apjl, 413, L105
 

### 2020-04-23 Nicolas.tex
Back to latest. Comms included, seem to be ok except mean bias in orange

--- 2020-04-23 Nicolas.tex
@@
         bias correction to derive the SN distance modulus bias affecting
-        cosmology, see text.}
+        cosmology, see text. \nn{\textit{Apparently, there is some issue with
+        the mean of all surveys combined. See Greg's comments.}}}
     \label{fig:magdrift}
@@
 % A
-%\bibitem[Abbott et al.(2018)]{abbott2018} Abbott, T.~M.~C., Abdalla, F.~B., Alarcon, A., et al.\ 2018, \prd, 98, 043526
-
-\bibitem[Abbott et al.(2019)]{descosmopaper2019} Abbott, T.~M.~C., Allam, S., Andersen, P., et al.\ 2019, \apjl, 872, L30
-
-\bibitem[Aldering et al.(2002)]{aldering2002} Aldering, G., Adam, G., Antilogus, P., et al.\ 2002, \procspie, 61
-
-\bibitem[Astier et al.(2006)]{astier2006} Astier, P., Guy, J., Regnault, N., et al.\ 2006, \aap, 447, 31
-
-%\bibitem[Ata et al.(2017)]{ata2017} Ata, M., Kitaura, F.-S., Chuang, C.-H., et al.\ 2017, \mnras, 467, 3993
-
-\bibitem[Aubourg et al.(2008)]{aubourg2008} Aubourg, {\'E}.,
-  Tojeiro, R., Jimenez, R., et al.\ 2008, \aap, 492, 631
+
+%\bibitem[Abbott et al.(2018)]{abbott2018} Abbott, T.~M.~C., Abdalla, F.~B.,
+%Alarcon, A., et al.\ 2018, \prd, 98, 043526
+
+\bibitem[Abbott et al.(2019)]{descosmopaper2019} Abbott, T.~M.~C., Allam, S.,
+Andersen, P., et al.\ 2019, \apjl, 872, L30
+
+\bibitem[Aldering et al.(2002)]{aldering2002} Aldering, G., Adam, G., Antilogus,
+P., et al.\ 2002, \procspie, 61
+
+\bibitem[Astier et al.(2006)]{astier2006} Astier, P., Guy, J., Regnault, N., et
+al.\ 2006, \aap, 447, 31
+
+%\bibitem[Ata et al.(2017)]{ata2017} Ata, M., Kitaura, F.-S., Chuang, C.-H., et
+%al.\ 2017, \mnras, 467, 3993
+
+\bibitem[Aubourg et al.(2008)]{aubourg2008} Aubourg, {\'E}., Tojeiro, R.,
+Jimenez, R., et al.\ 2008, \aap, 492, 631
 
 % B
-\bibitem[Bazin et al.(2011)]{bazin2011} Bazin, G., Ruhlmann-Kleider, V., Palanque-Delabrouille, N., et al.\ 2011, \aap, 534, A43
-
-\bibitem[Bellm et al.(2019)]{bellm2019} Bellm, E.~C., Kulkarni, S.~R., Graham, M.~J., et al.\ 2019, \pasp, 131, 018002
-
-\bibitem[Betoule et al.(2014)]{betoule2014} Betoule, M., Kessler, R., Guy, J., et al.\ 2014, \aap, 568, A22
-
-\bibitem[Brout et al.(2019)]{brout2019} Brout, D., Scolnic, D., Kessler, R., et al.\ 2019, \apj, 874, 150
+
+\bibitem[Bazin et al.(2011)]{bazin2011} Bazin, G., Ruhlmann-Kleider, V.,
+Palanque-Delabrouille, N., et al.\ 2011, \aap, 534, A43
+
+\bibitem[Bellm et al.(2019)]{bellm2019} Bellm, E.~C., Kulkarni, S.~R., Graham,
+M.~J., et al.\ 2019, \pasp, 131, 018002
+
+\bibitem[Betoule et al.(2014)]{betoule2014} Betoule, M., Kessler, R., Guy, J.,
+et al.\ 2014, \aap, 568, A22
+
+\bibitem[Brout et al.(2019)]{brout2019} Brout, D., Scolnic, D., Kessler, R., et
+al.\ 2019, \apj, 874, 150
 
@@
 % C
-\bibitem[Campbell et al.(2013)]{campbell2013} Campbell, H., D'Andrea, C.~B., Nichol, R.~C., et al.\ 2013, \apj, 763, 88
-
-%\bibitem[Chabanier et al.(2019)]{chabanier2019} Chabanier, S., Millea, M., \& Palanque-Delabrouille, N.\ 2019, \mnras, 489, 2247
-
-\bibitem[Childress et al.(2013)]{childress2013} Childress, M., Aldering, G., Antilogus, P., et al.\ 2013, \apj, 770, 108
-
-\bibitem[Childress et al.(2014)]{childress2014} Childress, M.~J., Wolf, C., \& Zahid, H.~J.\ 2014, \mnras, 445, 1898
-
-%\bibitem[Coles, \& Jones(1991)]{coles1991} Coles, P., \& Jones, B.\ 1991, \mnras, 248, 1
-
-%\bibitem[Conley et al.(2011)]{conley2011} Conley, A., Guy, J., Sullivan, M., et al.\ 2011, \apjs, 192, 1
+
+\bibitem[Campbell et al.(2013)]{campbell2013} Campbell, H., D'Andrea, C.~B.,
+Nichol, R.~C., et al.\ 2013, \apj, 763, 88
+
+%\bibitem[Chabanier et al.(2019)]{chabanier2019} Chabanier, S., Millea, M., \&
+%Palanque-Delabrouille, N.\ 2019, \mnras, 489, 2247
+
+\bibitem[Childress et al.(2013)]{childress2013} Childress, M., Aldering, G.,
+Antilogus, P., et al.\ 2013, \apj, 770, 108
+
+\bibitem[Childress et al.(2014)]{childress2014} Childress, M.~J., Wolf, C., \&
+Zahid, H.~J.\ 2014, \mnras, 445, 1898
+
+%\bibitem[Coles, \& Jones(1991)]{coles1991} Coles, P., \& Jones, B.\ 1991,
+%\mnras, 248, 1
+
+%\bibitem[Conley et al.(2011)]{conley2011} Conley, A., Guy, J., Sullivan, M., et
+%al.\ 2011, \apjs, 192, 1
 
 % D
-\bibitem[D'Andrea et al.(2011)]{dandrea2011} D'Andrea, C.~B., Gupta, R.~R., Sako, M., et al.\ 2011, \apj, 743, 172
-
-\bibitem[Dilday et al.(2008)]{dilday2008} Dilday, B., Kessler, R., Frieman, J.~A., et al.\ 2008, \apj, 682, 262
+
+\bibitem[D'Andrea et al.(2011)]{dandrea2011} D'Andrea, C.~B., Gupta, R.~R.,
+Sako, M., et al.\ 2011, \apj, 743, 172
+
+\bibitem[Dilday et al.(2008)]{dilday2008} Dilday, B., Kessler, R., Frieman,
+J.~A., et al.\ 2008, \apj, 682, 262
 
@@
 % F
-\bibitem[Feeney et al.(2019)]{feeney2019} Feeney, S.~M., Peiris, H.~V., Williamson, A.~R., et al.\ 2019, \prl, 122, 061105
-
-\bibitem[Freedman et al.(2019)]{freedman2019} Freedman, W.~L., Madore, B.~F., Hatt, D., et al.\ 2019, \apj, 882, 34
-
-\bibitem[Freedman et al.(2020)]{freedman2020} Freedman, W.~L., Madore, B.~F., Hoyt, T., et al.\ 2020, arXiv e-prints, arXiv:2002.01550
-
-\bibitem[Frieman et al.(2008)]{frieman2008} Frieman, J.~A., Bassett, B., Becker, A., et al.\ 2008, \aj, 135, 338
+
+\bibitem[Feeney et al.(2019)]{feeney2019} Feeney, S.~M., Peiris, H.~V.,
+Williamson, A.~R., et al.\ 2019, \prl, 122, 061105
+
+\bibitem[Freedman et al.(2019)]{freedman2019} Freedman, W.~L., Madore, B.~F.,
+Hatt, D., et al.\ 2019, \apj, 882, 34
+
+\bibitem[Freedman et al.(2020)]{freedman2020} Freedman, W.~L., Madore, B.~F.,
+Hoyt, T., et al.\ 2020, arXiv e-prints, arXiv:2002.01550
+
+\bibitem[Frieman et al.(2008)]{frieman2008} Frieman, J.~A., Bassett, B., Becker,
+A., et al.\ 2008, \aj, 135, 338
 
 % G
-\bibitem[Graham et al.(2019)]{graham2019} Graham, M.~J., Kulkarni, S.~R., Bellm, E.~C., et al.\ 2019, \pasp, 131, 078001
-
-\bibitem[Gupta et al.(2011)]{gupta2011} Gupta, R.~R., D'Andrea, C.~B., Sako, M., et al.\ 2011, \apj, 740, 92
-
-\bibitem[Guy et al.(2007)]{guy2007} Guy, J., Astier, P., Baumont, S., et al.\ 2007, \aap, 466, 11
-
-%\bibitem[Guy et al.(2010)]{guy2010} Guy, J., Sullivan, M., Conley, A., et al.\ 2010, \aap, 523, A7
-
-%\bibitem[Graziani et al.(2019)]{graziani2019} Graziani, R., Courtois, H.~M., Lavaux, G., et al.\ 2019, \mnras, 488, 5438
+
+\bibitem[Graham et al.(2019)]{graham2019} Graham, M.~J., Kulkarni, S.~R., Bellm,
+E.~C., et al.\ 2019, \pasp, 131, 078001
+
+\bibitem[Gupta et al.(2011)]{gupta2011} Gupta, R.~R., D'Andrea, C.~B., Sako, M.,
+et al.\ 2011, \apj, 740, 92
+
+\bibitem[Guy et al.(2007)]{guy2007} Guy, J., Astier, P., Baumont, S., et al.\
+2007, \aap, 466, 11
+
+%\bibitem[Guy et al.(2010)]{guy2010} Guy, J., Sullivan, M., Conley, A., et al.\
+%2010, \aap, 523, A7
+
+%\bibitem[Graziani et al.(2019)]{graziani2019} Graziani, R., Courtois, H.~M.,
+%Lavaux, G., et al.\ 2019, \mnras, 488, 5438
 
 % H
-\bibitem[Hamuy et al.(1996)]{hamuy1996} Hamuy, M., Phillips, M.~M., Suntzeff, N.~B., et al.\ 1996, \aj, 112, 2391
-
-\bibitem[Hamuy et al.(2000)]{hamuy2000} Hamuy, M., Trager, S.~C., Pinto, P.~A., et al.\ 2000, \aj, 120, 1479
-
-\bibitem[Hinton et al.(2019)]{hinton2019} Hinton, S.~R., Davis, T.~M., Kim, A.~G., et al.\ 2019, \apj, 876, 15
-
-\bibitem[Howell et al.(2007)]{howell2007} Howell, D.~A., Sullivan, M., Conley, A., et al.\ 2007, \apjl, 667, L37
-
-%\bibitem[Howell et al.(2009)]{howell2009} Howell, D.~A., Sullivan, M., Brown, E.~F., et al.\ 2009, \apj, 691, 661
+
+\bibitem[Hamuy et al.(1996)]{hamuy1996} Hamuy, M., Phillips, M.~M., Suntzeff,
+N.~B., et al.\ 1996, \aj, 112, 2391
+
+\bibitem[Hamuy et al.(2000)]{hamuy2000} Hamuy, M., Trager, S.~C., Pinto, P.~A.,
+et al.\ 2000, \aj, 120, 1479
+
+\bibitem[Hinton et al.(2019)]{hinton2019} Hinton, S.~R., Davis, T.~M., Kim,
+A.~G., et al.\ 2019, \apj, 876, 15
+
+\bibitem[Howell et al.(2007)]{howell2007} Howell, D.~A., Sullivan, M., Conley,
+A., et al.\ 2007, \apjl, 667, L37
+
+%\bibitem[Howell et al.(2009)]{howell2009} Howell, D.~A., Sullivan, M., Brown,
+%E.~F., et al.\ 2009, \apj, 691, 661
 
 % I
-\bibitem[Ivezi{\'c} et al.(2019)]{lsstpaper} Ivezi{\'c}, {\v{Z}}., Kahn, S.~M., Tyson, J.~A., et al.\ 2019, \apj, 873, 111
+
+\bibitem[Ivezi{\'c} et al.(2019)]{lsstpaper} Ivezi{\'c}, {\v{Z}}., Kahn, S.~M.,
+Tyson, J.~A., et al.\ 2019, \apj, 873, 111
 
 % J
-\bibitem[Jones et al.(2015)]{jones2015} Jones, D.~O., Riess, A.~G., \& Scolnic, D.~M.\ 2015, \apj, 812, 3
-1
-
-\bibitem[Jones et al.(2018)]{jones2018} Jones, D.~O., Riess, A.~G., Scolnic, D.~M., et al.\ 2018, \apj, 867, 108
-
-\bibitem[Jones et al.(2018)b]{jones2018b} Jones, D.~O., Scolnic, D.~M., Riess, A.~G., et al.\ 2018, \apj, 857, 51
-
-\bibitem[Jones et al.(2019)]{jones2019} Jones, D.~O., Scolnic, D.~M., Foley, R.~J., et al.\ 2019, \apj, 881, 19
+
+\bibitem[Jones et al.(2015)]{jones2015} Jones, D.~O., Riess, A.~G., \& Scolnic,
+D.~M.\ 2015, \apj, 812, 3 1
+
+\bibitem[Jones et al.(2018)]{jones2018} Jones, D.~O., Riess, A.~G., Scolnic,
+D.~M., et al.\ 2018, \apj, 867, 108
+
+\bibitem[Jones et al.(2018)b]{jones2018b} Jones, D.~O., Scolnic, D.~M., Riess,
+A.~G., et al.\ 2018, \apj, 857, 51
+
+\bibitem[Jones et al.(2019)]{jones2019} Jones, D.~O., Scolnic, D.~M., Foley,
+R.~J., et al.\ 2019, \apj, 881, 19
 
 % K
-\bibitem[Kelly et al.(2010)]{kelly2010} Kelly, P.~L., Hicken, M., Burke, D.~L., et al.\ 2010, \apj, 715, 743
-
-\bibitem[Kessler et al.(2009)]{kessler2009} Kessler, R., Becker, A.~C., Cinabro, D., et al.\ 2009, \apjs, 185, 32
-
-\bibitem[Kessler \& Scolnic(2017)]{kessler2017} Kessler, R., \& Scolnic, D.\ 2017, \apj, 836, 56
-
-\bibitem[Kim et al.(2018)]{kim18} Kim, Y.-L., Smith, M., Sullivan, M., et al.\ 2018, \apj, 854, 24
-
-\bibitem[Kim et al.(2019)]{kim19} Kim, Y.-L., Kang, Y., \& Lee, Y.-W.\ 2019, Journal of Korean Astronomical Society, 52, 181
-
-\bibitem[Knox \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv e-prints, arXiv:1908.03663
+
+\bibitem[Kelly et al.(2010)]{kelly2010} Kelly, P.~L., Hicken, M., Burke, D.~L.,
+et al.\ 2010, \apj, 715, 743
+
+\bibitem[Kessler et al.(2009)]{kessler2009} Kessler, R., Becker, A.~C., Cinabro,
+D., et al.\ 2009, \apjs, 185, 32
+
+\bibitem[Kessler \& Scolnic(2017)]{kessler2017} Kessler, R., \& Scolnic, D.\
+2017, \apj, 836, 56
+
+\bibitem[Kim et al.(2018)]{kim18} Kim, Y.-L., Smith, M., Sullivan, M., et al.\
+2018, \apj, 854, 24
+
+\bibitem[Kim et al.(2019)]{kim19} Kim, Y.-L., Kang, Y., \& Lee, Y.-W.\ 2019,
+Journal of Korean Astronomical Society, 52, 181
+
+\bibitem[Knox \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv
+e-prints, arXiv:1908.03663
 
 % L
-\bibitem[Lampeitl et al.(2010)]{lampeitl2010} Lampeitl, H., Smith, M., Nichol, R.~C., et al.\ 2010, \apj, 722, 566
+
+\bibitem[Lampeitl et al.(2010)]{lampeitl2010} Lampeitl, H., Smith, M., Nichol,
+R.~C., et al.\ 2010, \apj, 722, 566
 
 % M
-\bibitem[Mannucci et al.(2005)]{mannucci2005} Mannucci, F.,
-  Della Valle, M., Panagia, N., et al.\ 2005, \aap, 433, 807 
-\bibitem[Mannucci et al.(2006)]{mannucci2006} Mannucci, F.,
-  Della Valle, M., \& Panagia, N.\ 2006, \mnras, 370, 773
-
-\bibitem[Maoz et al.(2014)]{maozmannucci2014} Maoz, D., Mannucci,
-  F., \& Nelemans, G.\ 2014, \araa, 52, 107
+
+\bibitem[Mannucci et al.(2005)]{mannucci2005} Mannucci, F., Della Valle, M.,
+Panagia, N., et al.\ 2005, \aap, 433, 807
+
+\bibitem[Mannucci et al.(2006)]{mannucci2006} Mannucci, F., Della Valle, M., \&
+Panagia, N.\ 2006, \mnras, 370, 773
+
+\bibitem[Maoz et al.(2014)]{maozmannucci2014} Maoz, D., Mannucci, F., \&
+Nelemans, G.\ 2014, \araa, 52, 107
 
 % N
-\bibitem[Neill et al.(2006)]{neill2006} Neill, J.~D., Sullivan, M., Balam, D., et al.\ 2006, \aj, 132, 1126
-
-\bibitem[Neill et al.(2009)]{neill2009} Neill, J.~D., Sullivan, M., Howell, D.~A., et al.\ 2009, \apj, 707, 1449
-
-\bibitem[Nordin et al.(2018)]{nordin2018} Nordin, J., Aldering, G., Antilogus, P., et al.\ 2018, \aap, 614, A71
+
+\bibitem[Neill et al.(2006)]{neill2006} Neill, J.~D., Sullivan, M., Balam, D.,
+et al.\ 2006, \aj, 132, 1126
+
+\bibitem[Neill et al.(2009)]{neill2009} Neill, J.~D., Sullivan, M., Howell,
+D.~A., et al.\ 2009, \apj, 707, 1449
+
+\bibitem[Nordin et al.(2018)]{nordin2018} Nordin, J., Aldering, G., Antilogus,
+P., et al.\ 2018, \aap, 614, A71
 
@@
 % P
-\bibitem[Pan et al.(2014)]{pan2014} Pan, Y.-C., Sullivan, M., Maguire, K., et al.\ 2014, \mnras, 438, 1391
-
-\bibitem[Perlmutter et al.(1999)]{perlmutter1999} Perlmutter, S., Aldering, G., Goldhaber, G., et al.\ 1999, \apj, 517, 565
-
-\bibitem[Perrett et al.(2010)]{perrett2010} Perrett, K., Balam, D., Sullivan, M., et al.\ 2010, \aj, 140, 518
+
+\bibitem[Pan et al.(2014)]{pan2014} Pan, Y.-C., Sullivan, M., Maguire, K., et
+al.\ 2014, \mnras, 438, 1391
+
+\bibitem[Perlmutter et al.(1999)]{perlmutter1999} Perlmutter, S., Aldering, G.,
+Goldhaber, G., et al.\ 1999, \apj, 517, 565
+
+\bibitem[Perrett et al.(2010)]{perrett2010} Perrett, K., Balam, D., Sullivan,
+M., et al.\ 2010, \aj, 140, 518
 
@@
 
-%\bibitem[Planck Collaboration et al.(2016)]{planck2016} Planck Collaboration, Ade, P.~A.~R., Aghanim, N., et al.\ 2016, \aap, 594, A13
-
-\bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration, Aghanim, N., Akrami, Y., et al.\ 2018, arXiv e-prints, arXiv:1807.06209
-
-\bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T., et al.\ 2019, \prl, 122, 221301
+%\bibitem[Planck Collaboration et al.(2016)]{planck2016} Planck Collaboration,
+%Ade, P.~A.~R., Aghanim, N., et al.\ 2016, \aap, 594, A13
+
+\bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration,
+Aghanim, N., Akrami, Y., et al.\ 2018, arXiv e-prints, arXiv:1807.06209
+
+\bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T.,
+et al.\ 2019, \prl, 122, 221301
 
@@
 % R
-\bibitem[Reid et al.(2019)]{reid2019} Reid, M.~J., Pesce, D.~W., \& Riess, A.~G.\ 2019, arXiv e-prints, arXiv:1908.05625
-
-\bibitem[Rest et al.(2014)]{rest2014} Rest, A., Scolnic, D., Foley, R.~J., et al.\ 2014, \apj, 795, 44
-
-\bibitem[Riess et al.(1998)]{riess1998} Riess, A.~G., Filippenko, A.~V., Challis, P., et al.\ 1998, \aj, 116, 1009
-
-\bibitem[Riess et al.(2009)]{riess2009} Riess, A.~G., Macri, L., Casertano, S., et al.\ 2009, \apj, 699, 539
-
-\bibitem[Riess et al.(2016)]{riess2016} Riess, A.~G., Macri, L.~M., Hoffmann, S.~L., et al.\ 2016, \apj, 826, 56
-
-\bibitem[Riess et al.(2018)]{riess2018} Riess, A.~G., Casertano, S., Yuan, W., et al.\ 2018, \apj, 861, 126
-
-\bibitem[Riess et al.(2019)]{riess2019} Riess, A.~G., Casertano, S., Yuan, W., et al.\ 2019, \apj, 876, 85
-
-\bibitem[{Rigault {et~al.}(2013)}]{rigault2013}
-Rigault, M., Copin, Y., Aldering, G., {et~al.} 2013, \aap, 560, A66
-
-\bibitem[Rigault et al.(2015)]{rigault2015} Rigault, M., Aldering, G., Kowalski, M., et al.\ 2015, \apj, 802, 20
-
-\bibitem[Rigault et al.(2018)]{rigault2018} Rigault, M.,
-  Brinnel, V., Aldering, G., et al.\ 2018, arXiv:1806.03849
-
-\bibitem[Rodney et al.(2014)]{rodney2014} Rodney, S.~A.,
-  Riess, A.~G., Strolger, L.-G., et al.\ 2014, \aj, 148, 13
-
-\bibitem[Roman et al.(2018)]{roman2018} Roman, M., Hardin, D., Betoule, M., et al.\ 2018, \aap, 615, A68
-
-\bibitem[Rose et al.(2019)]{rose2019} Rose, B.~M., Garnavich, P.~M., \& Berg, M.~A.\ 2019, \apj, 874, 32
-
-\bibitem[Rubin et al.(2015)]{rubin2015} Rubin, D., Aldering, G., Barbary, K., et al.\ 2015, \apj, 813, 137
-
-\bibitem[Rubin \& Hayden(2016)]{rubin2016} Rubin, D., \& Hayden, B.\ 2016, \apjl, 833, L30
+
+\bibitem[Reid et al.(2019)]{reid2019} Reid, M.~J., Pesce, D.~W., \& Riess,
+A.~G.\ 2019, arXiv e-prints, arXiv:1908.05625
+
+\bibitem[Rest et al.(2014)]{rest2014} Rest, A., Scolnic, D., Foley, R.~J., et
+al.\ 2014, \apj, 795, 44
+
+\bibitem[Riess et al.(1998)]{riess1998} Riess, A.~G., Filippenko, A.~V.,
+Challis, P., et al.\ 1998, \aj, 116, 1009
+
+\bibitem[Riess et al.(2009)]{riess2009} Riess, A.~G., Macri, L., Casertano, S.,
+et al.\ 2009, \apj, 699, 539
+
+\bibitem[Riess et al.(2016)]{riess2016} Riess, A.~G., Macri, L.~M., Hoffmann,
+S.~L., et al.\ 2016, \apj, 826, 56
+
+\bibitem[Riess et al.(2018)]{riess2018} Riess, A.~G., Casertano, S., Yuan, W.,
+et al.\ 2018, \apj, 861, 126
+
+\bibitem[Riess et al.(2019)]{riess2019} Riess, A.~G., Casertano, S., Yuan, W.,
+et al.\ 2019, \apj, 876, 85
+
+\bibitem[{Rigault {et~al.}(2013)}]{rigault2013} Rigault, M., Copin, Y.,
+Aldering, G., {et~al.} 2013, \aap, 560, A66
+
+\bibitem[Rigault et al.(2015)]{rigault2015} Rigault, M., Aldering, G., Kowalski,
+M., et al.\ 2015, \apj, 802, 20
+
+\bibitem[Rigault et al.(2018)]{rigault2018} Rigault, M., Brinnel, V., Aldering,
+G., et al.\ 2018, arXiv:1806.03849
+
+\bibitem[Rodney et al.(2014)]{rodney2014} Rodney, S.~A., Riess, A.~G., Strolger,
+L.-G., et al.\ 2014, \aj, 148, 13
+
+\bibitem[Roman et al.(2018)]{roman2018} Roman, M., Hardin, D., Betoule, M., et
+al.\ 2018, \aap, 615, A68
+
+\bibitem[Rose et al.(2019)]{rose2019} Rose, B.~M., Garnavich, P.~M., \& Berg,
+M.~A.\ 2019, \apj, 874, 32
+
+\bibitem[Rubin et al.(2015)]{rubin2015} Rubin, D., Aldering, G., Barbary, K., et
+al.\ 2015, \apj, 813, 137
+
+\bibitem[Rubin \& Hayden(2016)]{rubin2016} Rubin, D., \& Hayden, B.\ 2016,
+\apjl, 833, L30
 
 % S
-\bibitem[Sako et al.(2008)]{sako2008} Sako, M., Bassett, B., Becker, A., et al.\ 2008, \aj, 135, 348
-
-\bibitem[Saunders et al.(2020)]{saunders2020} Saunders, C., Aldering, G., Antilogus, P., et al.\ 2020, VizieR Online Data Catalog, J/ApJ/869/167
-
-\bibitem[Scannapieco \& Bildsten(2005)]{scannapieco2005} Scannapieco, E., \& Bildsten, L.\ 2005, \apjl, 629, L85
-
-\bibitem[Scolnic et al.(2014)]{scolnic2014} Scolnic, D., Rest, A., Riess, A., et al.\ 2014, \apj, 795, 45
-
-\bibitem[Scolnic \& Kessler(2016)]{scolnic2016} Scolnic, D., \& Kessler, R.\ 2016, \apjl, 822, L35
-
-\bibitem[Scolnic et al.(2018)]{scolnic2018a} Scolnic, D.~M., Jones, D.~O., Rest, A., et al.\ 2018a, \apj, 859, 101
-
-%\bibitem[Scolnic et al.(2018)]{scolnic2018b} Scolnic, D.~M., Lochner, M., Gris, P., et al.\ 2018, arXiv e-prints, arXiv:1812.00516
-
-\bibitem[Scolnic et al.(2019)]{scolnicastro2020} Scolnic, D., Perlmutter, S., Aldering, G., et al.\ 2019, Astro2020: Decadal Survey on Astronomy and Astrophysics, 2020, 270
-
-\bibitem[Shariff et al.(2016)]{shariff2016} Shariff, H., Jiao, X., Trotta, R., et al.\ 2016, \apj, 827, 1
-
-\bibitem[Strolger et al.(2004)]{strolger04} Strolger, L.-G., Riess, A.~G., Dahlen, T., et al.\ 2004, \apj, 613, 200
-
-\bibitem[Sullivan et al.(2006)]{sullivan2006} Sullivan, M., Le  Borgne, D., Pritchet, C.~J., et al.\ 2006, \apj, 648, 868
-
-\bibitem[Sullivan et al.(2010)]{sullivan2010} Sullivan, M., Conley, A., Howell, D.~A., et al.\ 2010, \mnras, 406, 782
+
+\bibitem[Sako et al.(2008)]{sako2008} Sako, M., Bassett, B., Becker, A., et al.\
+2008, \aj, 135, 348
+
+\bibitem[Saunders et al.(2020)]{saunders2020} Saunders, C., Aldering, G.,
+Antilogus, P., et al.\ 2020, VizieR Online Data Catalog, J/ApJ/869/167
+
+\bibitem[Scannapieco \& Bildsten(2005)]{scannapieco2005} Scannapieco, E., \&
+Bildsten, L.\ 2005, \apjl, 629, L85
+
+\bibitem[Scolnic et al.(2014)]{scolnic2014} Scolnic, D., Rest, A., Riess, A., et
+al.\ 2014, \apj, 795, 45
+
+\bibitem[Scolnic \& Kessler(2016)]{scolnic2016} Scolnic, D., \& Kessler, R.\
+2016, \apjl, 822, L35
+
+\bibitem[Scolnic et al.(2018)]{scolnic2018a} Scolnic, D.~M., Jones, D.~O., Rest,
+A., et al.\ 2018a, \apj, 859, 101
+
+%\bibitem[Scolnic et al.(2018)]{scolnic2018b} Scolnic, D.~M., Lochner, M., Gris,
+%P., et al.\ 2018, arXiv e-prints, arXiv:1812.00516
+
+\bibitem[Scolnic et al.(2019)]{scolnicastro2020} Scolnic, D., Perlmutter, S.,
+Aldering, G., et al.\ 2019, Astro2020: Decadal Survey on Astronomy and
+Astrophysics, 2020, 270
+
+\bibitem[Shariff et al.(2016)]{shariff2016} Shariff, H., Jiao, X., Trotta, R.,
+et al.\ 2016, \apj, 827, 1
+
+\bibitem[Strolger et al.(2004)]{strolger04} Strolger, L.-G., Riess, A.~G.,
+Dahlen, T., et al.\ 2004, \apj, 613, 200
+
+\bibitem[Sullivan et al.(2006)]{sullivan2006} Sullivan, M., Le  Borgne, D.,
+Pritchet, C.~J., et al.\ 2006, \apj, 648, 868
+
+\bibitem[Sullivan et al.(2010)]{sullivan2010} Sullivan, M., Conley, A., Howell,
+D.~A., et al.\ 2010, \mnras, 406, 782
 
 % T
-\bibitem[Tasca et al.(2015)]{tasca2015} Tasca, L.~A.~M., Le F{\`e}vre, O., Hathi, N.~P., et al.\ 2015, \aap, 581, A54
+
+\bibitem[Tasca et al.(2015)]{tasca2015} Tasca, L.~A.~M., Le F{\`e}vre, O.,
+Hathi, N.~P., et al.\ 2015, \aap, 581, A54
 
@@
 % W
-\bibitem[Wiseman et al.(2020)]{wiseman2020} Wiseman, P., Smith, M., Childress, M., et al.\ 2020, arXiv e-prints, arXiv:2001.02640
-
-\bibitem[Wong et al.(2019)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F., et al.\ 2019, arXiv e-prints, arXiv:1907.04869
+
+\bibitem[Wiseman et al.(2020)]{wiseman2020} Wiseman, P., Smith, M., Childress,
+M., et al.\ 2020, arXiv e-prints, arXiv:2001.02640
+
+\bibitem[Wong et al.(2019)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F.,
+et al.\ 2019, arXiv e-prints, arXiv:1907.04869
 
@@
 % Y
-
 % Z
@@
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  BACKUP 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-%For clarity with the next models, we named it 3G2M2S$_{\text{SNf}}$ for it has a total of 3 Gaussians but with only 2 means and 2 standard deviations, and has been fitted on SNf data only.
-
-%We implemented and compared 10 models in total, 4 of which have an
-%evolution with the redshift from $\delta(z)$, and 6 don't ($\delta(z) = f =\mathrm{cst}$). The ones with and evolution are:
-\begin{itemize}
-    \item 3G2M2S, the one we described (but fitted on all the data);
-    \item 3G2M1S, where this time $\sigma_1 \equiv \sigma_2$;
-    \item 2G2M2S, model taken from HOWELL 2009 where we added $\delta(z)$;
-    \item 3G3M3S, with three independent Gaussians.
-\end{itemize}
-The ones without a stretch evolution are the same ones but with an ``F'' implying
-we set $\delta(z) = f = \mathrm{cst}$, and two others:
-\begin{itemize}
-    \item 1G1M1S, where there is no distinction between old and young SNe;
-    \item 1G1M2S, taken from KESLLER 2017 and used in recent cosmological
-        analysis SCOLNIC 2018. It's an asymetric model where
-\end{itemize}
-\begin{align}
- p(x_1^i, \mathrm{d} x_1^i | \mu, \sigma_-, \sigma_+) = 
-    \begin{cases}
-        \mathcal{N} \left(\mu, \sqrt{\sigma_-{}^2 + \mathrm{d} x_1^i{}^{2}}\right) (x_1^i) & \text{if
-        } x_1^i\geq \mu\\
-        \mathcal{N} \left(\mu, \sqrt{\sigma_+{}^2 + \mathrm{d} x_1^i{}^{2}}\right) (x_1^i), &
-        \text{else}
-    \end{cases}
-\end{align}
-
-The fitted parameters are shown in Table~\ref{tab:val}.
-
-\begin{table*}[htbp!]
-    \centering
-    \caption{Values of the best-it parameters for all the models.
-In red are the non-coherent ones which coincides with the models without an evolution of the fraction of young SNe Ia with the redshift.}
-    \label{tab:val}
-    \begin{tabular}{c c c c c c c}\hline\hline
-
-        Model & $a$ & $f$ & $\mu_1$ & $\sigma_1$ & $\mu_2$ &
-        $\sigma_2$ \\\hline
-
-        3G2M2S$_{\mathrm{SNf}}$ & $0.48 \pm 0.07$ & none & $0.39 \pm 0.07$ &
-        $0.56 \pm 0.05$ & $-1.5 \pm 0.1$ & $0.52 \pm 0.09$ \\
-
-        3G2M2S & $0.48 \pm 0.17$ & none & $0.36 \pm 0.08$ & $0.61 \pm 0.05$ &
-        $-1.3 \pm 0.2$ & $0.60 \pm 0.12$ \\
-
-        3G2M2SF & \textcolor{red}{$0.1 \pm 0.6$} & \textcolor{red}{$0.2 \pm
-        0.6$} & $-0.9 \pm 0.7$ & $0.7 \pm 0.3$ & $0.5 \pm 0.2$ & $0.6 \pm 0.1$
-        \\
-
-        3G2M1S & $0.47 \pm 0.07$ & none & $0.35 \pm 0.04$ & $0.61 \pm 0.03$ &
-        $-1.25 \pm 0.10$ & $\sigma_1$ \\
-
-        3G2M1SF & \textcolor{red}{$0.2 \pm 0.9$} & $0.7 \pm 0.3$ & $0.36 \pm
-        0.04$ & $0.60 \pm 0.03$ & $-1.23 \pm 0.10$ & $\sigma_1$ \\
-
-        2G2M2S & none & none & $0.49 \pm 0.04$ & $0.54 \pm 0.03$ & $-0.72 \pm
-        0.08$ & $0.83 \pm 0.07$ \\
-
-        2G2M2SF & none & $0.3 \pm 0.2$ & $-0.9 \pm 0.6$ & $0.7 \pm 0.2$ & $0.5
-        \pm 0.2$ & $0.56 \pm 0.09$ \\\hline
-
-    \end{tabular} \bigbreak
-
-\begin{tabular}{c c c}\hline\hline
-
-    Model & $\mu$ & $\sigma$ \\\hline
-
-    1G1M1S & $0.01 \pm 0.04$ & $0.90 \pm 0.03$ \\\hline
-
-\end{tabular} \bigbreak
-
-\begin{tabular}{c c c c c}\hline\hline
-
-    Model & $\mu$ & $\sigma_-$ & $\sigma_+$ \\\hline
-
-    1G1M2S & $0.16617 \pm 0.00004$ & $1.07 \pm 0.04$ & $0.69 \pm 0.03$ \\\hline
-
-\end{tabular} \bigbreak
-
-\begin{tabular}{c c c c c c c c c}\hline\hline
-
-    Model & $a$ & $f$ & $\mu_1$ & $\sigma_1$ & $\mu_2$ & $\sigma_2$ &
-    $\mu_3$ & $\sigma_3$ \\\hline
-
-    3G3M3S & $0.14 \pm 0.08$ & none & $0.51 \pm 0.06$ & $0.54 \pm 0.04$ & $-1.9
-    \pm 0.2$ & $0.29 \pm 0.11$ & $-0.55 \pm 0.12$ & $0.67 \pm 0.15$ \\
-
-    3G3M3SF & $0.2 \pm 0.2 $ & $0.10 \pm 0.04 $ & $-1.7 \pm 0.2$ & $0.4 \pm 0.1$
-            & $0.9 \pm 0.1$ & $0.3 \pm 0.2$ & $0.0 \pm 0.2$ & $0.7 \pm 0.1$
-            \\\hline
-
-\end{tabular} \bigbreak
-\end{table*}
-
-To test the coherence of this model, we also fitted it using all the data from
-the surveys indistinctly. These results are shown Fig.~\ref{fig:model_all}.
-
-\begin{table*}
-    \centering
-    \caption{BBC comparison}
-    \label{tab:bbc_comp}
-    \begin{tabular}{c c |c c c |c c c |c c c}\hline\hline\\[-1.05em]
-     &  & \multicolumn{3}{|c}{Asym (this)} & \multicolumn{3}{|c}{Asym (SK18/S18)} & \multicolumn{3}{|c}{Base SNf} \\
-    N$_{\mathrm{SN}}$ & Survey & $\mathcal{X}^2$ & AIC & Free param & $\mathcal{X}^2$ & AIC & Free param & $\mathcal{X}^2$ & AIC & Free param \\\hline
-    167 & SDSS & 445.0 & 451.2 & 3 & 457.9 & 464.0 & 3 & 452.9 & 452.9 & 0\\
-    160 & PS1 & 397.4 & 403.6 & 3 & 397.6 & 403.7 & 3 & 405.2 & 405.2 & 0\\
-    102 & SNLS & 252.7 & 259.0 & 3 & 254.5 & 260.8 & 3 & 269.5 & 269.5 & 0\\
-    114 & SNf & 299.3 & 305.5 & 3 & -- & -- & -- & 268.6 & 279.1 & 5\\
-    543 & Total & 1394.5 & 1419.1 & 12 & -- & -- & -- & 1396.0 & 1406.2 & 5\\
-    429 & Total-SNf & 1095.2 & 1113.6 & 9 & 1110.0 & 1128.4 & 9 & 1127.5 & 1127.5 & 0\\\hline\hline
-    \end{tabular}
-\end{table*}
-
-\begin{table*}
-    \centering
-    \caption{BBC comparison}
-    \label{tab:bbc_comp}
-    \begin{tabular}{c |c c c |c c c}\hline\hline\\[-1.05em]
-     & \multicolumn{3}{|c}{Asym (this)}  & \multicolumn{3}{|c}{Base SNf} \\
-    Survey & $\mathcal{X}^2$ & AIC & Free param & $\mathcal{X}^2$ & AIC & Free param \\\hline
-    SNf & 299.3 & 305.5 & 3 & 268.6 & 279.1 & 5\\
-    SDSS & 445.0 & 451.2 & 3 & 452.9 & 452.9 & 0\\
-    PS1 & 397.4 & 403.6 & 3 & 405.2 & 405.2 & 0\\
-    SNLS & 252.7 & 259.0 & 3  & 269.5 & 269.5 & 0\\
-    Total & 1394.5 & 1419.1 & 12 & 1396.0 & 1406.2 & 5\\\hline\hline
-    \end{tabular}
-\end{table*}we are also testing non-drifting models by restricting the ``Base'' and ``Howell'' models to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe; these models are labeled ``Base constant'' and ``Howell constant''.} \yc{\uline{fix labels!!!
-\nn{done}}}

### 2020-05-13 Nicolas.tex
48h to go

@@
 \newcommand{\yc}[1]{{\textcolor{BrickRed}{#1}}}
+\newcommand{\agk}[1]{\textcolor{purple}{AK: #1}}
+
+\newcommand{\orcid}[1]{
+\href{https://orcid.org/#1}{
+\includegraphics[height=11pt]{General_figures/ORCIDiD_icon128x128.png}}}
 
@@
 
-\author{N. Nicolas
-    \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
-    \inst{1}
-    \and M. Rigault
-    \thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
-    \inst{2}\and\\
-    Y. Copin\inst{1}   
-    \and R. Graziani\inst{2}
+\author{
+    N. Nicolas \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution} \inst{1} 
+    \and M. Rigault \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
+    \orcid{0000-0002-8121-2560}
+    \and\\ Y. Copin \inst{1}
+    \and R. Graziani \inst{2}
     %HERE ALPHABETIC ORDER
@@
     \and J. Nordin\inst{4}
+    \and A.~G. Kim\inst{3}
     \and Y.-L. Kim\inst{1}
@@
 
-\abstract{
-The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
-survey statistics increase, the question of astrophysical systematic
-uncertainties rises, notably that of the SN~Ia population redshift drift. In
-this paper, we study the redshift evolution of SN~Ia \textsc{\texttt{SALT2.4}}
-lightcurve stretch, a purely intrinsic SN property, to probe this drift. The SN
-stretch has been shown to strongly correlate with the SN's environment, notably
-with stellar age tracers. Following the evolution of fraction of young and old
-SNe~Ia as predicted by \cite{rigault2018}, and assuming constant underlying
-stretch distribution for each population as a function of redshift, model the
-expected SN~Ia stretch distribution as a function of redshift. We test our
-prediction against data from the literature chosen to have negligible selection
-effects to ensure that any observed drift is indeed astrophysical and not
+\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
+    unknown, and as survey statistics increase, the question of astrophysical
+    systematic uncertainties rises, notably that of the SN~Ia population
+    redshift drift. In this paper, we study the redshift evolution of SN~Ia
+    \textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN
+    property, to probe this drift. The SN stretch has been shown to strongly
+    correlate with the SN's environment, notably with stellar age tracers. We
+    model the underlying stretch distribution as a function of reshift, using
+    the evolution of fraction of young and old SNe~Ia as predicted by
+    \cite{rigault2018}, and assuming constant underlying stretch distribution
+    for each age population made of Gaussian-mixtures.  We test our prediction
+    against data from the literature chosen to have negligible selection effects
+    to ensure that any observed drift is indeed astrophysical and not
     observational. We clearly demonstrate that the underlying SNe~Ia stretch
     distribution is evolving as a function of redshift, and that the young/old
-drifting model is a much better description of the data than any non-drifting
-model, including the sample-based asymmetric distributions used by the Beams
-with Bias Correction algorithm to correct Malmquist bias. Our favored underlying
-stretch model is bimodal: a high-stretch mode shared by both young and old
-environments, and a low-stretch mode exclusive to old environments. The precise
-impact of the redshift evolution of the underlying SNe Ia population on
-cosmology remains to be studied. Yet, the astrophysical drift of the SN stretch
+    drifting model is a much better description of the data than any
+    non-drifting model, including the sample-based asymmetric distributions
+    usually used to correct Malmquist bias. Our favored underlying stretch model
+    is bimodal: a high-stretch mode shared by both young and old environments,
+    and a low-stretch mode exclusive to old environments. The precise impact of
+    the redshift evolution of the underlying SNe Ia population on cosmology
+    remains to be studied. Yet, the astrophysical drift of the SN stretch
     distribution does affect current Malmquist bias corrections and thereby
-distances derived from SN affected by selection effects. We highlight that such
-a bias will increase with surveys covering larger redshift ranges, which is
-particularly important for LSST.
-}
+    distances derived from SN affected by selection effects. We highlight that
+    such a bias will increase with surveys covering larger redshift ranges,
+    which is particularly important for LSST.}
 
@@
 
-%\nn{We point that there has been sample-based approach meant to account for an unmodeled variation of host-related parameters thanks to multiple redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show aren't as effective as a direct modeling of such a drift}.
-
-The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate.
+%\nn{We point that there has been sample-based approach meant to account for an
+%unmodeled variation of host-related parameters thanks to multiple
+%redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show
+%aren't as effective as a direct modeling of such a drift}.
+
+The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
 \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
@@
     \caption{Composition of the SNe~Ia dataset used in this analysis.
-    Conservative cuts are indicated in brackets.}
+    Conservative cuts are indicated in parentheses.}
     \label{tab:sample}
@@
 been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
-and~4.2.2; SNe~Ia time series are published in \citealt{Saunders2020}
-\nn{\textit{Greg: for now you may need to cite SNEMO and SUGAR to remove this
-``?''}}, see also Aldering in prep. and Ponder in prep.). The spectroscopic
-follow-up was done over a redshift range of $0.02 < z < 0.09$~\citep[as
-in][]{rigault2018}, while the search was much deeper. As such, these SNe~Ia are
-assumed to be a random sampling of the underlying SN population. The SNfactory
-sample is particularly useful for studying SN property drift, as it enables us
-to have a large SN~Ia sample at $z<0.1$.
+and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
+\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
+range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
+much deeper. \mr{These 114 SNfactory SNe~Ia are thus part of the volume limited
+part of the survey (Aldering in prep.).} As such, these SNe~Ia are assumed to be
+a random sampling of the underlying SN population. The SNfactory sample is
+particularly useful for studying SN property drift, as it enables us to have a
+large SN~Ia sample at $z<0.1$.
 
@@
 But generally, the fraction of young SNe~Ia as a function of redshift is given
-by $\delta$(z) (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
+by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
 of the underlying distribution of SNe~Ia as a function of redshift $X_1(z)$ is
@@
 
-Given the probability $y^i$ that a given SN is young, and assuming our Base (see
-Section~\ref{sec:basemodel}), the probability to measure a
+Given the probability $y^i$ that a given SN is young, and assuming our Base
+\nn{model} (see Section~\ref{sec:basemodel}), the probability to measure a
 \textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
@@
 
-For the SNfactory sample, we can readily set $y^i = p_y^i$, the probability to
-have $\log \mathrm{LsSFR} \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
-minimize Eq.~\ref{eq:likelihood} with respect to~$\vec{\theta}$. Results on
-fitting the SNf SNe with this model are shown Table~\ref{tab:modelresults} and
-illustrated in Fig.~\ref{fig:stretchlssfr}.
+\mr{Following \cite{rigault2018}, we set $p_y^i$ as one minus the cumulative
+distribution function of the $\log(\mathrm{LsSFR})$ at
+$-10.82\,\mathrm{dex}$. This is defining the probability that a SN projenitor
+is classified as young (see further discussion about this assumption in Briday
+et al. in~prep). Since the drift is negligible within the SNfactory dataset
+($\Delta\delta(z)<3\%$ from the highest to the lowest redshift of the dataset),
+one can readily use $y^i = p_y^i$ to minimize Eq.~\ref{eq:likelihood} with
+respect to $\vec{\theta}$. Results on fitting the SNfactory SNe with this model
+are shown Table~\ref{tab:modelresults} and illustrated in
+Fig.~\ref{fig:stretchlssfr}.}
 
@@
 
+%\agk{You don't define $p^i_y$. The language is is sloppy. What is relevant is
+%the opposite, the probability that a galaxy with low mass gets the measured
+%LsSFR. These are in general different. The solution you use is only correct
+%for a model in which the rates for low and high-LsSFR galaxies is the same.
+%The probability in general should include a prior on $l$.
+
+%\begin{align}
+%p(x^i_{o1},l_o^i | \mu_1,\mu_2,\sigma_1,\sigma_2,a) & = \int dl^i p(x^i_{o1},l_o^i | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
+%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
+% & = \int dl^i p(x^i_{o1} | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
+% p(l_o^i | li) 
+%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
+%& = \int_{-\infty}^{-10.82} dl^i [a\mathcal{N}(\mu_1,\sigma_1{}^2) +
+%    (1-a)\mathcal{N}(\mu_2,\sigma_2{}^2] 
+% p(l_o^i | l^i) 
+%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
+%&+\int_{-10.82}^{\infty} dl^i\mathcal{N}(\mu_1,\sigma_1{}^2)
+% p(l_o^i | l^i) 
+%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber
+%\end{align}
+%Note that you use rates without LsSFR measurements, so you are doing something
+%inconsistent here. There should be a $\delta{z}$ term here. However, since the
+%rates are the same at $z=0.05$, maybe this doesn't matter so much.
+%}
+
 \subsubsection{Without LsSFR measurements}\label{sec:modelnopy}
@@
 (resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
-predicted redshift evolution of mean stretch (mean of Eq.~\ref{eq:stretchz})
-illustrated as a blue band in Fig.~\ref{fig:modelall}. We see in this figure
-that the measured mean SN~Ia stretch per redshift bins of equal sample size
-closely follows our redshift drift modeling. This is indeed what is expected if
-old environments favor low SN stretches \citep[e.g.][]{howell2007} \emph{and} if
-the fraction of old SNe~Ia declines as a function of redshift. See
-Section~\ref{sec:results} for a more quantitative discussion.
+predicted redshift evolution of mean stretch \mr{(expected $x_1$ given the
+distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
+Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances}.
+We see in this figure that the measured mean SN~Ia stretch per redshift bins of
+equal sample size closely follows our redshift drift modeling. This is indeed
+what is expected if old environments favor low SN stretches
+\citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
+a function of redshift. See Section~\ref{sec:results} for a more quantitative
+discussion.
 
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
+    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf-hist_on.pdf}
     \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
@@
         are compatible. The light-blue band illustrates the amplitude of the
-    error of the best fit model when considering the fiducial dataset.}
+        error (incl. covariance) of the best fit model when considering the
+        fiducial dataset. \mr{The redshift distributions of each sample are
+        shown as light-histograms, see lower left legend.}}
     \label{fig:modelall}
@@
 having a sample-based approach covers some potential redshift evolution
-information \citep{scolnic2016, scolnic2018a}. See further discussion
-concerning BBC in Section~\ref{sec:discussion}.
-
-Finally, for the sake of completeness, we also consider simple the ``Gaussian''
-and ``Asymmetric'' Gaussian non-drifting models.
+information \citep{scolnic2016, scolnic2018a}. See further discussion concerning
+BBC in Section~\ref{sec:discussion}.
+
+Finally, for the sake of completeness, we also consider Gaussian
+and Asymmetric Gaussian non-drifting models.
 
@@
 are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
-a Gaussian probability distribution.
-
-We first note that the best model (with smallest AIC) is the so-called Base
-model, implementing the \cite{rigault2018} description of the stretch
-distribution and its redshift evolution. This is therefore the reference model,
-both on the fiducial and conservative samples.
+a Gaussian probability distribution. \mr{However, we highlight that the best of
+the compared models might not be the true underlying model.}
+
+The best model (with smallest AIC) is the so-called Base model and thus is out
+the reference model. This is true both on the fiducial and conservative samples.
 
@@
 Surprisingly, the sample-based Gaussian asymmetric modeling used by current
-implementations of the BBC technique \citep{scolnic2016, kessler2017}, is one of
-the worst modelisations of our analysis (see Section~\ref{sec:results}). While
-its $-2\ln(L)$ is the smallest of all redshift-independent models (but still
-$-11.5$ worse than the reference Base model), it is strongly penalized for
-requiring 15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the 5~samples of
-the analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted
-as a probability $p=2\times 10^{-7}$ of being an as good representation of the
-data as the Base model. We report in Table~\ref{tab:bbc} the samples' $\mu_0$
-and $\sigma_{\pm}$ adjusted on the nominally selection-free samples using our
-fiducial cuts (see Section~\ref{sec:sample}). We find our results in close
-agreement with \cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a}
-for PS1, who derived these model parameters using the full BBC formalism, using
-numerous simulations to model the selection effects (see details e.g., Section~3
-of \citealt{kessler2017}). The agreement between our fit of the asymmetric
+implementations of the BBC technique \citep{scolnic2016, kessler2017} has one of
+the highest AIC value in our analysis (see Section~\ref{sec:results}). While its
+$-2\ln(L)$ is the smallest of all redshift-independent models (but still $-11.5$
+worse than the reference Base model), it is strongly penalized for requiring
+15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the 5~samples of the
+analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted as a
+probability $p=2\times 10^{-7}$ of being an as good representation of the data
+as the Base model.
+\mr{Remark that the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
+k\ln(n)$, with $n$ is the number of datapoint) might be better suited than AIC
+when comparin models to the sample-based one since it explicitely accounts for
+the fact that each sample is fitted separately; the sample-based model BIC is
+the sum of the BIC per sample. Doing so we find $\Delta\mathrm{BIC}=-48$, also
+excluding the sample-base asymmetric Gaussian model as an as good description of
+the data then the Base-model.}
+
+%\agk{I think it does worse not because of the more parameters but that it has 5
+%complexity penalties whereas the base model only has 1.  AIC doesn't account
+%for the fact that each sample is fit independently.  It seems a fairer metric
+%would be the BIC, where you adjust $n$ for each sample.  This would recover
+%$\sim 16$ in the IC for this model.}
+
+We report in Table~\ref{tab:bbc} the samples' $\mu_0$ and $\sigma_{\pm}$
+adjusted on the nominally selection-free samples using our fiducial cuts (see
+Section~\ref{sec:sample}). We find our results in close agreement with
+\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
+derived these model parameters using the full BBC formalism, using numerous
+simulations to model the selection effects (see details e.g., Section~3 of
+\citealt{kessler2017}). The agreement between our fit of the asymmetric
 Gaussians on the supposedly selection-free part of the samples and the results
@@
 
-Finally, \sout{considering the SNfactory dataset like any other (i.e.,} ignoring
-the LsSFR measurements \nn{-- available only for the SNfactory dataset}, see
-Section~\ref{sec:modeling} -- \sout{),} reduces the significance of the results
-presented in this section, as expected. Yet, non-drifting models remain strongly
-disfavored, and for instance, the best fitted sample-based Gaussian asymmetric
-modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the data than
-our Base drifting modeling.
+Finally, ignoring the LsSFR measurements --~available only for the SNfactory
+dataset, see Section~\ref{sec:modeling}~-- reduces the significance of the
+results presented in this section, as expected. Yet, non-drifting models remain
+strongly disfavored, and for instance, the best fitted sample-based Gaussian
+asymmetric modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the
+data than our Base drifting modeling.
 
@@
 redshifts. This means that the SN bias-corrected standardized magnitude
-estimated at redshifts affected by selection effect would be biased by an
+estimated at a redshift affected by selection effect would be biased by an
 ill-modeling of the true underlying stretch distribution.
@@
 asymmetric modeling as part of the BBC simulations.
-
-Nonetheless, it is enlightening to study the expected mean stretch difference as
-a function of redshift when considering that the underlying SN stretch
-distribution follows either the per-sample Asymmetric Gaussian modeling or our
-Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
-we convert this difference in mean stretch $x_1$ to a difference in magnitude
-using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
-correspond to the actual impact on cosmological distances derived from SNe~Ia as
-this simple stretch average comparison does not account for the amplitude of the
-Malmquist bias correction to be applied as a function of redshift and surveys.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
-    \caption{Difference of mean stretch standardization (in mag, using
-        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
-        Gaussian asymmetric modeling and our Base drifting model, as a function
-        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
-        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
-        considering the individual SN redshifts (no redshift cut applied). The
-        orange markers represent their mean evolution in 30 regularly-spaced
-        redshift bins. This must be multiplied by the amplitude of the Malmquist
-        bias correction to derive the SN distance modulus bias affecting
-        cosmology, see text. \nn{\textit{Apparently, there is some issue with
-        the mean of all surveys combined. See Greg's comments.}}}
-    \label{fig:magdrift}
-\end{figure}
-
-We see that the amplitude of the modeling difference is a few tens of millimag.
+\mr{However, we already highlight that even if a non-drifting sample-based model
+could provide comparable result in the volume-limited part of the various
+samples, these models would differ when extrapolating at higher redshift,
+precisely where the  underlying will matter for correcting Malmquist biases.}
+
+%Nonetheless, it is enlightening to study the expected mean stretch difference
+%as a function of redshift when considering that the underlying SN stretch
+%distribution follows either the per-sample Asymmetric Gaussian modeling or our
+%Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
+%we convert this difference in mean stretch $x_1$ to a difference in magnitude
+%using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
+%correspond to the actual impact on cosmological distances derived from SNe~Ia
+%as this simple stretch average comparison does not account for the amplitude of
+%the Malmquist bias correction to be applied as a function of redshift and
+%surveys.
+
+%\agk{ The AG modeling contains Malmquist bias. To be fair you should plot Base
+%with the survey-specific Malmquist bias. There is something strange going on at
+%$z<0.2$, can you add SNfactory to this plot? I think you fill find that the
+%systematic difference is driven by differences in the $x_1$ distribution of
+%SNf.}
+
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
+%    \caption{Difference of mean stretch standardization (in mag, using
+%        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
+%        Gaussian asymmetric modeling and our Base drifting model, as a function
+%        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
+%        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
+%        considering the individual SN redshifts (no redshift cut applied). The
+%        orange markers represent their mean evolution in 30 regularly-spaced
+%        redshift bins. This must be multiplied by the amplitude of the Malmquist
+%        bias correction to derive the SN distance modulus bias affecting
+%        cosmology, see text. \nn{\textit{Apparently, there is some issue with
+%        the mean of all surveys combined. See Greg's comments.}}}
+%    \label{fig:magdrift}
+%\end{figure}
+
+%We see that the amplitude of the modeling difference is a few tens of millimag.
 In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
@@
         younger and older SNe~Ia follows the prediction made in
-        \cite{rigault2018}.
+        \cite{rigault2018}. This second result strongly supports the existence
+        of both young and old SN~Ia populations, in agreement with rate studies.
 
@@
         even less so as survey span larger redshift ranges. We warn that, even
-        if extra degrees of freedom might be acceptable \nn{given the large
-        number of SNe~Ia in} cosmological studies, extrapolating the SN property
+        if extra degrees of freedom might be acceptable given the large
+        number of SNe~Ia in cosmological studies, extrapolating the SN property
         distributions from the volume-limited part of a survey to the
@@
 
-    \item An inaccurate underlying stretch distribution model is estimated to
-        bias the mean standardized magnitude of SNe~Ia affected by
-        selection effects at the order of a few percents. This is
-        comparable in scale to the observational signature of exotic forms of
-        dark energy and is therefore a significant source of systematic that
-        should be carefully studied.
+    %\item An inaccurate underlying stretch distribution model is estimated to
+    %    bias the mean standardized magnitude of SNe~Ia affected by
+    %    selection effects at the order of a few percents. This is
+    %    comparable in scale to the observational signature of %exotic forms of
+    %    dark energy and is therefore a significant source of systematic that
+    %    should be carefully studied.
 
@@
 
+\bibitem[Aldering et al.(2020)]{aldering2020} Aldering, G., Antilogus, P.,
+Aragon, C., et al.\ 2020, Research Notes of the American Astronomical Society,
+4, 63
+
 \bibitem[Astier et al.(2006)]{astier2006} Astier, P., Guy, J., Regnault, N., et

### 2020-05-13 Nicolas.tex
YC add

@@
     N. Nicolas \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution} \inst{1} 
-    \and M. Rigault \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
+    \and \mr{M. Rigault} \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
     \orcid{0000-0002-8121-2560}
-    \and\\ Y. Copin \inst{1}
+    \and\\ \yc{Y. Copin} \inst{1}
+    \orcid{0000-0002-5317-7518}
     \and R. Graziani \inst{2}
@@
     \textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN
-    property, to probe this drift. The SN stretch has been shown to strongly
-    correlate with the SN's environment, notably with stellar age tracers. We
-    model the underlying stretch distribution as a function of reshift, using
+    property, to probe this drift. \yc{\emph{isn't a 'drift' \textbf{exactly} a redshift evolution?}} The SN stretch has been shown to strongly
+    correlate with the SN environment, notably with stellar age tracers. We
+    model the underlying stretch distribution as a function of redshift, using
     the evolution of fraction of young and old SNe~Ia as predicted by
     \cite{rigault2018}, and assuming constant underlying stretch distribution
-    for each age population made of Gaussian-mixtures.  We test our prediction
-    against data from the literature chosen to have negligible selection effects
-    to ensure that any observed drift is indeed astrophysical and not
-    observational. We clearly demonstrate that the underlying SNe~Ia stretch
+    for each age population made of Gaussian mixtures.  We test our prediction
+    against \yc{published samples} chosen to have negligible \yc{magnitude} selection effects,
+    \yc{so} that any observed \yc{change} is indeed \yc{of astrophysical and not
+    observational origin}. We clearly demonstrate that the underlying SN~Ia stretch
     distribution is evolving as a function of redshift, and that the young/old
     drifting model is a much better description of the data than any
-    non-drifting model, including the sample-based asymmetric distributions
-    usually used to correct Malmquist bias. Our favored underlying stretch model
-    is bimodal: a high-stretch mode shared by both young and old environments,
+    \yc{\sout{non-drifting} constant} model, including the sample-based asymmetric distributions
+    usually used to correct Malmquist bias. \yc{The} favored underlying stretch model
+    is \yc{the bimodal one introduced in \cite{rigault2018}}: a high-stretch mode shared by both young and old environments,
     and a low-stretch mode exclusive to old environments. The precise impact of
-    the redshift evolution of the underlying SNe Ia population on cosmology
+    the redshift evolution of the SN Ia population \yc{intrinsic properties} on cosmology
     remains to be studied. Yet, the astrophysical drift of the SN stretch
@@
     distances derived from SN affected by selection effects. We highlight that
-    such a bias will increase with surveys covering larger redshift ranges,
+    such a bias will increase with surveys covering \yc{increasingly} larger redshift ranges,
     which is particularly important for LSST.}
@@
 \maketitle
+
+\yc{\emph{IMHO too many occurrences of `drift' (53!); this could be replaced by '[redshift] evolution' to be opposed to 'constant'.}}
 
@@
 range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
-much deeper. \mr{These 114 SNfactory SNe~Ia are thus part of the volume limited
-part of the survey (Aldering in prep.).} As such, these SNe~Ia are assumed to be
+much deeper. \mr{These 114 SNfactory SNe~Ia are thus \yc{in} the volume limited
+part of the survey (Aldering \yc{et al.}, in prep.)}, \yc{and are therefore} assumed to be
 a random sampling of the underlying SN population. The SNfactory sample is
@@
 
-\mr{Following \cite{rigault2018}, we set $p_y^i$ as one minus the cumulative
+\mr{Following \cite{rigault2018}, we \yc{define?} $p_y^i$ as one minus the cumulative
 distribution function of the $\log(\mathrm{LsSFR})$ at
-$-10.82\,\mathrm{dex}$. This is defining the probability that a SN projenitor
-is classified as young (see further discussion about this assumption in Briday
+$-10.82\,\mathrm{dex}$. This is defining the probability that a SN progenitor
+is classified as young (see further discussion about this assumption \yc{\emph{which assumption, if it's purely your own definition of young vs. old?}} in Briday
 et al. in~prep). Since the drift is negligible within the SNfactory dataset
@@
 are shown Table~\ref{tab:modelresults} and illustrated in
-Fig.~\ref{fig:stretchlssfr}.}
+Fig.~\ref{fig:stretchlssfr}.} \yc{\emph{Now I'm confused why one does not have \textbf{exactly} $y^i = p_y^i$ when LsSFR measurement is available, but you seem to say it's an assumption.}}
 
@@
         fiducial dataset. \mr{The redshift distributions of each sample are
-        shown as light-histograms, see lower left legend.}}
+        shown as light-histograms, see lower left legend.} \yc{\emph{This is upper panel of Fig.3? Why is it again split in subsamples, while it was not deemed necessary in dedicated Fig.3?}}}
     \label{fig:modelall}
@@
 
-Finally, for the sake of completeness, we also consider Gaussian
-and Asymmetric Gaussian non-drifting models.
+Finally, for the sake of completeness, we also consider \yc{constant} Gaussian
+and Asymmetric Gaussian \yc{\sout{non-drifting}} models.
 
@@
 a Gaussian probability distribution. \mr{However, we highlight that the best of
-the compared models might not be the true underlying model.}
+the compared models might not be the true underlying model.} \yc{\emph{why this sentence? Is there any \textbf{true} intrinsic model?}}
 
@@
 as the Base model.
-\mr{Remark that the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
-k\ln(n)$, with $n$ is the number of datapoint) might be better suited than AIC
-when comparin models to the sample-based one since it explicitely accounts for
-the fact that each sample is fitted separately; the sample-based model BIC is
-the sum of the BIC per sample. Doing so we find $\Delta\mathrm{BIC}=-48$, also
-excluding the sample-base asymmetric Gaussian model as an as good description of
-the data then the Base-model.}
+
+\yc{Remark that, when comparing models adjusted on individual subsamples rather than globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
+k\ln(n)$, with $n$ the number of data points) might be better suited than AIC, since it explicitly accounts for
+the fact that each subsample is fitted separately: the sample-based model BIC is
+rightfully the sum of the BIC for each sample. Doing so, we find $\Delta\mathrm{BIC}=-48$, 
+again \sout{excluding} \emph{refuting} the sample-based asymmetric Gaussian model as being as pertinent as the Base model.}
 
@@
 To the best of our knowledge, a SN~Ia stretch redshift drift modeling has never
-been explicitly used in cosmological analyses, though bayesian hierarchy
+been explicitly used in cosmological analyses, though Bayesian hierarchy
 formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016} or Steve
@@
 could provide comparable result in the volume-limited part of the various
-samples, these models would differ when extrapolating at higher redshift,
-precisely where the  underlying will matter for correcting Malmquist biases.}
+samples, these models would differ when extrapolating at higher redshifts,
+precisely where the  underlying \yc{\emph{missing word?}} will matter for correcting Malmquist biases.}
 
@@
 several recent SN~Ia cosmological analyses, implemented as a per-sample
-asymmetric Gaussian distribution.
+asymmetric Gaussian distribution. \yc{\emph{of what?}}
 
@@
 \begin{enumerate}
-
-    \item The underlying SN~Ia stretch distribution is redshift dependent, as
+    \item The underlying SN~Ia stretch distribution is \yc{significantly} redshift dependent, as
         previously suggested by e.g.~\cite{howell2007}. This result is largely
@@
 
-    \item Non-drifting models are indeed excluded as suitable descriptions of
+    \item \yc{\sout{Non-drifting} Redshift-independent} models are indeed excluded as suitable descriptions of
         the data relative to our Base model. This model assumes that: (1) the
@@
         \cite{rigault2018}. This second result strongly supports the existence
-        of both young and old SN~Ia populations, in agreement with rate studies.
+        of both young and old SN~Ia populations, in agreement with rate studies. \yc{\emph{ref???}}
 
@@
         in the current implementation of the BBC, is excluded to be good
-        descriptions of the data relative to our drifting model. Hence, the
+        descriptions \yc{\emph{rephrase: "using is excluded"?}} of the data relative to our drifting model. Hence, the
         sample-based approach does not accurately account for redshift drift and
-        even less so as survey span larger redshift ranges. We warn that, even
+        even less so as survey span \yc{increasingly} larger redshift ranges. We \yc{\sout{warn} stipulate} that, even
         if extra degrees of freedom might be acceptable given the large
         number of SNe~Ia in cosmological studies, extrapolating the SN property
-        distributions from the volume-limited part of a survey to the
-        magnitude-limited one affected by sample selection would still be
-        inaccurate since the distribution would drift.
+        distributions from the volume-limited part of a survey to its
+        \yc{Malmquist-biased} magnitude-limited one \yc{\sout{affected by sample selection}} would still be
+        inaccurate \yc{\sout{since the distribution would drift} because of the redshift evolution}.
 
@@
         population model as a function of redshift:
-        \begin{align}\label{eqconclusion:stretchz}
+        \begin{align*}
+        \label{eqconclusion:stretchz}
             X_1\left(z \right) =
@@
             (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
-        \end{align}
-        with: $a=0.51$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
-        $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}) and using the
-        age-population drift modelization $\delta(z)$ with $K=0.87$:
-        \begin{align}
-            \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}.
-        \end{align}
+            \tag{\ref{eq:stretchz}}
+        \end{align*}
+        with $a=0.51$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
+        $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}), and using the
+        age-population drift model         \begin{align*}
+            \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
+            \tag{\ref{eq:delta}}
+        \end{align*}
+        with $K=0.87$.
 \end{enumerate}
@@
 In this paper, we considered a simple Gaussian mixture modeling, but additional
-data free from significant Malmquist bias would enable us to refine it. We note
+data free from significant Malmquist bias would enable us to refine it \yc{as necessary}. We note
 that samples at the low- and high-redshift ends of the Hubble diagram would be
@@
 provided by the Zwicky Transient Facility \citep{bellm2019, graham2019} and
-Subaru and SeeChange SNe~Ia programs, respectively.
+Subaru and SeeChange SNe~Ia programs, respectively. \yc{\emph{cannot say who's who in 'A and B and C'.}}
 

### 2020-05-13 Nicolas.tex
NN changes

@@
 \usepackage[dvipsnames]{xcolor}
-\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
-\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
-\newcommand{\nn}[1]{{\textcolor[rgb]{0.25, 0.50, 0}{#1}}}
-\newcommand{\nnn}[1]{{\textcolor[rgb]{0.40, 0.0, 0.20}{#1}}}
-\newcommand{\yc}[1]{{\textcolor{BrickRed}{#1}}}
+\newcommand{\mr}[1]{\textcolor[rgb]{0.60,0.10,0.6}{#1}}
+\newcommand{\mri}[1]{\textcolor{red}{#1}}
+\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
+\newcommand{\nnn}[1]{\nn{\textit{#1}}}
+\newcommand{\yc}[1]{\textcolor{BrickRed}{#1}}
 \newcommand{\agk}[1]{\textcolor{purple}{AK: #1}}
@@
     N. Nicolas \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution} \inst{1} 
-    \and \mr{M. Rigault} \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
+    \and M. Rigault \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
     \orcid{0000-0002-8121-2560}
-    \and\\ \yc{Y. Copin} \inst{1}
+    \and\\ Y. Copin \inst{1}
     \orcid{0000-0002-5317-7518}
@@
     \textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN
-    property, to probe this drift. \yc{\emph{isn't a 'drift' \textbf{exactly} a redshift evolution?}} The SN stretch has been shown to strongly
-    correlate with the SN environment, notably with stellar age tracers. We
-    model the underlying stretch distribution as a function of redshift, using
-    the evolution of fraction of young and old SNe~Ia as predicted by
+    property, to probe this drift. \yc{\emph{isn't a 'drift' \textbf{exactly} a
+    redshift evolution?}} The SN stretch has been shown to strongly correlate
+    with the SN environment, notably with stellar age tracers. We model the
+    underlying stretch distribution as a function of redshift, using the
+    evolution of fraction of young and old SNe~Ia as predicted by
     \cite{rigault2018}, and assuming constant underlying stretch distribution
     for each age population made of Gaussian mixtures. We test our prediction
-    against \yc{published samples} chosen to have negligible \yc{magnitude} selection effects,
-    \yc{so} that any observed \yc{change} is indeed \yc{of astrophysical and not
-    observational origin}. We clearly demonstrate that the underlying SN~Ia stretch
-    distribution is evolving as a function of redshift, and that the young/old
-    drifting model is a much better description of the data than any
-    \yc{\sout{non-drifting} constant} model, including the sample-based asymmetric distributions
-    usually used to correct Malmquist bias. \yc{The} favored underlying stretch model
-    is \yc{the bimodal one introduced in \cite{rigault2018}}: a high-stretch mode shared by both young and old environments,
-    and a low-stretch mode exclusive to old environments. The precise impact of
-    the redshift evolution of the SN Ia population \yc{intrinsic properties} on cosmology
-    remains to be studied. Yet, the astrophysical drift of the SN stretch
+    against published samples chosen to have negligible magnitude selection
+    effects, so that any observed change is indeed of astrophysical and not
+    observational origin. We clearly demonstrate that the underlying SN~Ia
+    stretch distribution is evolving as a function of redshift, and that the
+    young/old drifting model is a much better description of the data than any
+    \yc{\sout{non-drifting} constant} model, including the sample-based
+    asymmetric distributions usually used to correct Malmquist bias. The favored
+    underlying stretch model is \yc{the bimodal one introduced in
+    \cite{rigault2018}} \nnn{to my knowledge, R+18 doesn't introduce a
+    stretch model, but rather discusses the differences between the stretch
+    distributions for the young and old pop, never saying anything about stretch
+    modes}: a high-stretch mode shared by both young and old environments, and a
+    low-stretch mode exclusive to old environments. The precise impact of the
+    redshift evolution of the SN Ia population intrinsic properties on
+    cosmology remains to be studied. Yet, the astrophysical drift of the SN stretch
     distribution does affect current Malmquist bias corrections and thereby
-    distances derived from SN affected by selection effects. We highlight that
-    such a bias will increase with surveys covering \yc{increasingly} larger redshift ranges,
-    which is particularly important for LSST.}
+    distances derived from SN affected by selection effects. We highlight that such
+    a bias will increase with surveys covering increasingly larger redshift
+    ranges, which is particularly important for LSST.}
 
@@
 
-\yc{\emph{IMHO too many occurrences of `drift' (53!); this could be replaced by '[redshift] evolution' to be opposed to 'constant'.}}
+\yc{\emph{IMHO too many occurrences of `drift' (53!); this could be replaced by
+'[redshift] evolution' to be opposed to 'constant'.}}
 
@@
 range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
-much deeper. \mr{These 114 SNfactory SNe~Ia are thus \yc{in} the volume limited
-part of the survey (Aldering \yc{et al.}, in prep.)}, \yc{and are therefore} assumed to be
-a random sampling of the underlying SN population. The SNfactory sample is
+much deeper. These 114 SNfactory SNe~Ia are thus in the volume limited part
+of the survey (Aldering et al., in prep.), and are therefore assumed to be a
+random sampling of the underlying SN population. The SNfactory sample is
 particularly useful for studying SN property drift, as it enables us to have a
@@
 given by:
-\begin{multline}\label{eq:stretchz}
-    X_1(z) = \delta(z) \times \mathcal{N}(\mu_1,\sigma_1{}^2) + {} \\
-    (1-\delta(z)) \times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
-    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right].
-\end{multline}
+\begin{align}\label{eq:stretchz}
+    X_1(z) = \delta(z)&\times \mathcal{N}(\mu_1,\sigma_1{}^2) + \nonumber \\
+    (1-\delta(z))&\times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
+    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right]
+\end{align}
 
@@
 Given the probability $y^i$ that a given SN is young, and assuming our Base
-\nn{model} (see Section~\ref{sec:basemodel}), the probability to measure a
+model (see Section~\ref{sec:basemodel}), the probability to measure a
 \textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
@@
 
-\mr{Following \cite{rigault2018}, we \yc{define?} $p_y^i$ as one minus the cumulative
-distribution function of the $\log(\mathrm{LsSFR})$ at
-$-10.82\,\mathrm{dex}$. This is defining the probability that a SN progenitor
-is classified as young (see further discussion about this assumption \yc{\emph{which assumption, if it's purely your own definition of young vs. old?}} in Briday
-et al. in~prep). Since the drift is negligible within the SNfactory dataset
-($\Delta\delta(z)<3\%$ from the highest to the lowest redshift of the dataset),
-one can readily use $y^i = p_y^i$ to minimize Eq.~\ref{eq:likelihood} with
-respect to $\vec{\theta}$. Results on fitting the SNfactory SNe with this model
-are shown Table~\ref{tab:modelresults} and illustrated in
-Fig.~\ref{fig:stretchlssfr}.} \yc{\emph{Now I'm confused why one does not have \textbf{exactly} $y^i = p_y^i$ when LsSFR measurement is available, but you seem to say it's an assumption.}}
+Following \cite{rigault2018}, we define $p_y^i$ as one minus the cumulative
+distribution function of the $\log(\mathrm{LsSFR})$ at $-10.82\,\mathrm{dex}$.
+This is defining the probability that a SN progenitor is classified as
+young\nn{, thus assuming the LsSFR to be a perfect age tracer} (see further
+discussion about this assumption \yc{\emph{which assumption, if it's purely your
+own definition of young vs.\ old?}} in Briday et al.\ in~prep). \nn{This
+definition should be weighted by the $\delta(z)$ probability for a SN to be
+young, since the two effects play their roles, but} the drift is negligible
+within the SNfactory dataset ($\Delta\delta(z)<3\%$ from the highest to the
+lowest redshift of the dataset)\nn{; thus} one can readily use $y^i = p_y^i$ to
+minimize Eq.~\ref{eq:likelihood} with respect to $\vec{\theta}$. Results on
+fitting the SNfactory SNe with this model are shown Table~\ref{tab:modelresults}
+and illustrated in Fig.~\ref{fig:stretchlssfr}. \yc{\emph{Now I'm confused why
+one does not have \textbf{exactly} $y^i = p_y^i$ when LsSFR measurement is
+available, but you seem to say it's an assumption.}}
 
@@
 
-%\agk{You don't define $p^i_y$. The language is is sloppy. What is relevant is
-%the opposite, the probability that a galaxy with low mass gets the measured
-%LsSFR. These are in general different. The solution you use is only correct
-%for a model in which the rates for low and high-LsSFR galaxies is the same.
-%The probability in general should include a prior on $l$.
-
-%\begin{align}
-%p(x^i_{o1},l_o^i | \mu_1,\mu_2,\sigma_1,\sigma_2,a) & = \int dl^i p(x^i_{o1},l_o^i | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
-%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
-% & = \int dl^i p(x^i_{o1} | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
-% p(l_o^i | li) 
-%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
-%& = \int_{-\infty}^{-10.82} dl^i [a\mathcal{N}(\mu_1,\sigma_1{}^2) +
-%    (1-a)\mathcal{N}(\mu_2,\sigma_2{}^2] 
-% p(l_o^i | l^i) 
-%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
-%&+\int_{-10.82}^{\infty} dl^i\mathcal{N}(\mu_1,\sigma_1{}^2)
-% p(l_o^i | l^i) 
-%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber
-%\end{align}
-%Note that you use rates without LsSFR measurements, so you are doing something
-%inconsistent here. There should be a $\delta{z}$ term here. However, since the
-%rates are the same at $z=0.05$, maybe this doesn't matter so much.
-%}
-
 \subsubsection{Without LsSFR measurements}\label{sec:modelnopy}
@@
 (resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
-predicted redshift evolution of mean stretch \mr{(expected $x_1$ given the
+predicted redshift evolution of mean stretch (expected $x_1$ given the
 distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
-Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances}.
+Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances.
 We see in this figure that the measured mean SN~Ia stretch per redshift bins of
@@
         error (incl. covariance) of the best fit model when considering the
-        fiducial dataset. \mr{The redshift distributions of each sample are
-        shown as light-histograms, see lower left legend.} \yc{\emph{This is upper panel of Fig.3? Why is it again split in subsamples, while it was not deemed necessary in dedicated Fig.3?}}}
+        fiducial dataset. The redshift distributions of each sample are
+        shown as light-histograms, see lower left legend. \yc{\emph{This is upper panel of Fig.3? Why is it again split in subsamples, while it was not deemed necessary in dedicated Fig.3?}}}
     \label{fig:modelall}
@@
 are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
-a Gaussian probability distribution. \mr{However, we highlight that the best of
-the compared models might not be the true underlying model.} \yc{\emph{why this sentence? Is there any \textbf{true} intrinsic model?}}
+a Gaussian probability distribution. However, we highlight that the best of the
+compared models \nn{is not meant to} be the true underlying model. \nn{It only
+means that it is the best at describing the observations while minimizing the
+number of free parameters}. \yc{\emph{why this sentence? Is there any
+\textbf{true} intrinsic model?}}
 
@@
 
-\yc{Remark that, when comparing models adjusted on individual subsamples rather than globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
-k\ln(n)$, with $n$ the number of data points) might be better suited than AIC, since it explicitly accounts for
-the fact that each subsample is fitted separately: the sample-based model BIC is
-rightfully the sum of the BIC for each sample. Doing so, we find $\Delta\mathrm{BIC}=-48$, 
-again \sout{excluding} \emph{refuting} the sample-based asymmetric Gaussian model as being as pertinent as the Base model.}
-
-%\agk{I think it does worse not because of the more parameters but that it has 5
-%complexity penalties whereas the base model only has 1.  AIC doesn't account
-%for the fact that each sample is fit independently.  It seems a fairer metric
-%would be the BIC, where you adjust $n$ for each sample.  This would recover
-%$\sim 16$ in the IC for this model.}
+Remark that, when comparing models adjusted on individual subsamples rather than
+globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
+k\ln(n)$, with $n$ the number of data points) might be better suited than AIC,
+since it explicitly accounts for the fact that each subsample is fitted
+separately: the sample-based model BIC is rightfully the sum of the BIC for each
+sample. Doing so, we find $\Delta\mathrm{BIC}=-48$, again refuting the
+sample-based asymmetric Gaussian model as being as pertinent as the Base model.
 
@@
 using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
-asymmetric modeling as part of the BBC simulations.
-\mr{However, we already highlight that even if a non-drifting sample-based model
-could provide comparable result in the volume-limited part of the various
-samples, these models would differ when extrapolating at higher redshifts,
-precisely where the  underlying \yc{\emph{missing word?}} will matter for correcting Malmquist biases.}
-
-%Nonetheless, it is enlightening to study the expected mean stretch difference
-%as a function of redshift when considering that the underlying SN stretch
-%distribution follows either the per-sample Asymmetric Gaussian modeling or our
-%Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
-%we convert this difference in mean stretch $x_1$ to a difference in magnitude
-%using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
-%correspond to the actual impact on cosmological distances derived from SNe~Ia
-%as this simple stretch average comparison does not account for the amplitude of
-%the Malmquist bias correction to be applied as a function of redshift and
-%surveys.
-
-%\agk{ The AG modeling contains Malmquist bias. To be fair you should plot Base
-%with the survey-specific Malmquist bias. There is something strange going on at
-%$z<0.2$, can you add SNfactory to this plot? I think you fill find that the
-%systematic difference is driven by differences in the $x_1$ distribution of
-%SNf.}
-
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
-%    \caption{Difference of mean stretch standardization (in mag, using
-%        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
-%        Gaussian asymmetric modeling and our Base drifting model, as a function
-%        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
-%        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
-%        considering the individual SN redshifts (no redshift cut applied). The
-%        orange markers represent their mean evolution in 30 regularly-spaced
-%        redshift bins. This must be multiplied by the amplitude of the Malmquist
-%        bias correction to derive the SN distance modulus bias affecting
-%        cosmology, see text. \nn{\textit{Apparently, there is some issue with
-%        the mean of all surveys combined. See Greg's comments.}}}
-%    \label{fig:magdrift}
-%\end{figure}
-
-%We see that the amplitude of the modeling difference is a few tens of millimag.
+asymmetric modeling as part of the BBC simulations. However, we already
+highlight that even if a non-drifting sample-based model could provide
+comparable result in the volume-limited part of the various samples, these
+models would differ when extrapolating at higher redshifts, precisely where the
+underlying distribution will matter for correcting Malmquist biases.
+
 In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
@@
 several recent SN~Ia cosmological analyses, implemented as a per-sample
-asymmetric Gaussian distribution. \yc{\emph{of what?}}
+asymmetric Gaussian \nn{stretch} distribution.
 
@@
 \begin{enumerate}
-    \item The underlying SN~Ia stretch distribution is \yc{significantly} redshift dependent, as
-        previously suggested by e.g.~\cite{howell2007}. This result is largely
-        independent of details on each age-population modeling.
-
-    \item \yc{\sout{Non-drifting} Redshift-independent} models are indeed excluded as suitable descriptions of
-        the data relative to our Base model. This model assumes that: (1) the
-        younger population has a unimodal Gaussian stretch distribution, while
-        the older population stretch distribution is bimodal, one mode being the
-        same as the young one; (2) the evolution of the relative fraction of
-        younger and older SNe~Ia follows the prediction made in
-        \cite{rigault2018}. This second result strongly supports the existence
-        of both young and old SN~Ia populations, in agreement with rate studies. \yc{\emph{ref???}}
-
-    \item Using survey-based asymmetric Gaussian distributions, as done, e.g.,
-        in the current implementation of the BBC, is excluded to be good
-        descriptions \yc{\emph{rephrase: "using is excluded"?}} of the data relative to our drifting model. Hence, the
-        sample-based approach does not accurately account for redshift drift and
-        even less so as survey span \yc{increasingly} larger redshift ranges. We \yc{\sout{warn} stipulate} that, even
-        if extra degrees of freedom might be acceptable given the large
-        number of SNe~Ia in cosmological studies, extrapolating the SN property
-        distributions from the volume-limited part of a survey to its
-        \yc{Malmquist-biased} magnitude-limited one \yc{\sout{affected by sample selection}} would still be
-        inaccurate \yc{\sout{since the distribution would drift} because of the redshift evolution}.
-
-    %\item An inaccurate underlying stretch distribution model is estimated to
-    %    bias the mean standardized magnitude of SNe~Ia affected by
-    %    selection effects at the order of a few percents. This is
-    %    comparable in scale to the observational signature of %exotic forms of
-    %    dark energy and is therefore a significant source of systematic that
-    %    should be carefully studied.
+    \item The underlying SN~Ia stretch distribution is significantly redshift
+        dependent, as previously suggested by e.g.~\cite{howell2007}. This
+        result is largely independent of details on each age-population
+        modeling.
+
+    \item Redshift-independent models are indeed excluded as suitable
+        descriptions of the data relative to our Base model. This model assumes
+        that: (1) the younger population has a unimodal Gaussian stretch
+        distribution, while the older population stretch distribution is
+        bimodal, one mode being the same as the young one; (2) the evolution of
+        the relative fraction of younger and older SNe~Ia follows the prediction
+        made in \cite{rigault2018}. This second result strongly supports the
+        existence of both young and old SN~Ia populations, in agreement with
+        rate studies. \yc{\emph{ref???}}
+
+    \item \nn{Models} using survey-based asymmetric Gaussian distributions, as
+        done, e.g., in the current implementation of the BBC, are excluded to be
+        good descriptions of the data relative to our drifting model. Hence,
+        the sample-based approach does not accurately account for redshift drift
+        and even less so as survey span increasingly larger redshift ranges. We
+        stipulate that, even if extra degrees of freedom might be acceptable
+        given the large number of SNe~Ia in cosmological studies, extrapolating
+        the SN property distributions from the volume-limited part of a survey
+        to its Malmquist-biased magnitude-limited one  would still be inaccurate
+        because of the redshift evolution.
 
@@
             X_1\left(z \right) =
-             &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
-            (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
-            (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
+            \delta(z)&\times\mathcal{N}(\mu_1,\sigma_1{}^2)\,+\nonumber\\
+            (1-\delta(z))&\times \left[a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
+            (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2)\right]
             \tag{\ref{eq:stretchz}}
@@
 In this paper, we considered a simple Gaussian mixture modeling, but additional
-data free from significant Malmquist bias would enable us to refine it \yc{as necessary}. We note
-that samples at the low- and high-redshift ends of the Hubble diagram would be
-particularly helpful for this drifting analysis; fortunately this will soon be
-provided by the Zwicky Transient Facility \citep{bellm2019, graham2019} and
-Subaru and SeeChange SNe~Ia programs, respectively. \yc{\emph{cannot say who's who in 'A and B and C'.}}
+data free from significant Malmquist bias would enable us to refine it as
+necessary. We note that samples at the low- and high-redshift ends of the Hubble
+diagram would be particularly helpful for this drifting analysis; fortunately
+this will soon be provided by the Zwicky Transient Facility \citep{bellm2019,
+graham2019} \nn{on one hand}, and Subaru and SeeChange SNe~Ia programs \nn{on
+the other hand}, respectively. \nnn{need to check which is which}
 

### 2020-05-13 Nicolas.tex
Just ref and last sentence
Last of AGK

--- 2020-05-13 Nicolas.tex
@@
 
-\newcommand{\orcid}[1]{
-\href{https://orcid.org/#1}{
-\includegraphics[height=11pt]{General_figures/ORCIDiD_icon128x128.png}}}
-
 \def\l{\lambda}\def\L{\Lambda}
@@
 
-\author{
-    N. Nicolas \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution} \inst{1} 
-    \and M. Rigault \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
-    \orcid{0000-0002-8121-2560}
-    \and\\ Y. Copin \inst{1}
-    \and R. Graziani \inst{2}
+\author{N. Nicolas
+    \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
+    \inst{1}
+    \and M. Rigault
+    \thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
+    \inst{2}\and\\
+    Y. Copin\inst{1}   
+    \and R. Graziani\inst{2}
     %HERE ALPHABETIC ORDER
@@
     \and J. Nordin\inst{4}
-    \and A.~G. Kim\inst{3}
     \and Y.-L. Kim\inst{1}
@@
 
-\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
-    unknown, and as survey statistics increase, the question of astrophysical
-    systematic uncertainties rises, notably that of the SN~Ia population
-    redshift drift. In this paper, we study the redshift evolution of SN~Ia
-    \textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN
-    property, to probe this drift. The SN stretch has been shown to strongly
-    correlate with the SN's environment, notably with stellar age tracers. We
-    model the underlying stretch distribution as a function of reshift, using
-    the evolution of fraction of young and old SNe~Ia as predicted by
-    \cite{rigault2018}, and assuming constant underlying stretch distribution
-    for each age population made of Gaussian-mixtures.  We test our prediction
-    against data from the literature chosen to have negligible selection effects
-    to ensure that any observed drift is indeed astrophysical and not
+\abstract{
+The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
+survey statistics increase, the question of astrophysical systematic
+uncertainties rises, notably that of the SN~Ia population redshift drift. In
+this paper, we study the redshift evolution of SN~Ia \textsc{\texttt{SALT2.4}}
+lightcurve stretch, a purely intrinsic SN property, to probe this drift. The SN
+stretch has been shown to strongly correlate with the SN's environment, notably
+with stellar age tracers. Following the evolution of fraction of young and old
+SNe~Ia as predicted by \cite{rigault2018}, and assuming constant underlying
+stretch distribution for each population as a function of redshift, model the
+expected SN~Ia stretch distribution as a function of redshift. We test our
+prediction against data from the literature chosen to have negligible selection
+effects to ensure that any observed drift is indeed astrophysical and not
 observational. We clearly demonstrate that the underlying SNe~Ia stretch
 distribution is evolving as a function of redshift, and that the young/old
-    drifting model is a much better description of the data than any
-    non-drifting model, including the sample-based asymmetric distributions
-    usually used to correct Malmquist bias. Our favored underlying stretch model
-    is bimodal: a high-stretch mode shared by both young and old environments,
-    and a low-stretch mode exclusive to old environments. The precise impact of
-    the redshift evolution of the underlying SNe Ia population on cosmology
-    remains to be studied. Yet, the astrophysical drift of the SN stretch
+drifting model is a much better description of the data than any non-drifting
+model, including the sample-based asymmetric distributions used by the Beams
+with Bias Correction algorithm to correct Malmquist bias. Our favored underlying
+stretch model is bimodal: a high-stretch mode shared by both young and old
+environments, and a low-stretch mode exclusive to old environments. The precise
+impact of the redshift evolution of the underlying SNe Ia population on
+cosmology remains to be studied. Yet, the astrophysical drift of the SN stretch
 distribution does affect current Malmquist bias corrections and thereby
-    distances derived from SN affected by selection effects. We highlight that
-    such a bias will increase with surveys covering larger redshift ranges,
-    which is particularly important for LSST.}
+distances derived from SN affected by selection effects. We highlight that such
+a bias will increase with surveys covering larger redshift ranges, which is
+particularly important for LSST.
+}
 
@@
 
-%\nn{We point that there has been sample-based approach meant to account for an
-%unmodeled variation of host-related parameters thanks to multiple
-%redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show
-%aren't as effective as a direct modeling of such a drift}.
-
-The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
+%\nn{We point that there has been sample-based approach meant to account for an unmodeled variation of host-related parameters thanks to multiple redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show aren't as effective as a direct modeling of such a drift}.
+
+The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate.
+\agk{This grammatical error appears a lot in this article.  In compound
+nouns the first noun is typically singular, so this should
+be SN~Ia age dichotomy.}
 \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
@@
     \caption{Composition of the SNe~Ia dataset used in this analysis.
-    Conservative cuts are indicated in parentheses.}
+    Conservative cuts are indicated in brackets.
+    \agk{parentheses not brackets}}
     \label{tab:sample}
@@
 been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
-and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
-\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
-range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
-much deeper. \mr{These 114 SNfactory SNe~Ia are thus part of the volume limited
-part of the survey (Aldering in prep.).} As such, these SNe~Ia are assumed to be
-a random sampling of the underlying SN population. The SNfactory sample is
-particularly useful for studying SN property drift, as it enables us to have a
-large SN~Ia sample at $z<0.1$.
+and~4.2.2; SNe~Ia time series are published in \citealt{Saunders2020}
+\nn{\textit{Greg: for now you may need to cite SNEMO and SUGAR to remove this
+``?''}}, see also Aldering in prep. and Ponder in prep.). The spectroscopic
+follow-up was done over a redshift range of $0.02 < z < 0.09$~\citep[as
+in][]{rigault2018}, while the search was much deeper. \mr{These 114 SNfactory SNeIa 
+thus are part of the volume limited part of the survey (Aldering in prep.).}
+As such, these SNe~Ia are
+assumed to be a random sampling of the underlying SN population. The SNfactory
+sample is particularly useful for studying SN property drift, as it enables us
+to have a large SN~Ia sample at $z<0.1$.
 
@@
 underlying stretch distribution is drifting.
+\agk{how can this be asserted when you just showed that none of the surveys
+are complete at $z \sim 0.65$!?} \mr{MR: Alex, There is HST too, see top of Fig. 3. | in the latest bin 68 SNLS, 26 HST}
 
@@
 But generally, the fraction of young SNe~Ia as a function of redshift is given
-by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
+by $\delta$(z) \agk{Convention is that letters that are variables
+are italicized.  $z$, $H_0$.}
+(see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
 of the underlying distribution of SNe~Ia as a function of redshift $X_1(z)$ is
@@
 
-Given the probability $y^i$ that a given SN is young, and assuming our Base
-\nn{model} (see Section~\ref{sec:basemodel}), the probability to measure a
+Given the probability $y^i$ that a given SN is young, and assuming our Base (see
+Section~\ref{sec:basemodel}), the probability to measure a
 \textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
@@
 
-\mr{Following \cite{rigault2018}, we set $p_y^i$ as one minus the cumulative
-distribution function of the $\log(\mathrm{LsSFR})$ at
-$-10.82\,\mathrm{dex}$. This is defining the probability that a SN projenitor
-is classified as young (see further discussion about this assumption in Briday
-et al. in~prep). Since the drift is negligible within the SNfactory dataset
-($\Delta\delta(z)<3\%$ from the highest to the lowest redshift of the dataset),
-one can readily use $y^i = p_y^i$ to minimize Eq.~\ref{eq:likelihood} with
-respect to $\vec{\theta}$. Results on fitting the SNfactory SNe with this model
-are shown Table~\ref{tab:modelresults} and illustrated in
-Fig.~\ref{fig:stretchlssfr}.}
+For the SNfactory sample, we can readily set $y^i = p_y^i$, the probability to
+have $\log \mathrm{LsSFR} \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
+minimize Eq.~\ref{eq:likelihood} with respect to~$\vec{\theta}$. Results on
+fitting the SNf SNe with this model are shown Table~\ref{tab:modelresults} and
+illustrated in Fig.~\ref{fig:stretchlssfr}.
 
@@
 \end{table*}
-
-%\agk{You don't define $p^i_y$. The language is is sloppy. What is relevant is
-%the opposite, the probability that a galaxy with low mass gets the measured
-%LsSFR. These are in general different. The solution you use is only correct
-%for a model in which the rates for low and high-LsSFR galaxies is the same.
-%The probability in general should include a prior on $l$.
-
-%\begin{align}
-%p(x^i_{o1},l_o^i | \mu_1,\mu_2,\sigma_1,\sigma_2,a) & = \int dl^i p(x^i_{o1},l_o^i | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
-%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
-% & = \int dl^i p(x^i_{o1} | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
-% p(l_o^i | li) 
-%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
-%& = \int_{-\infty}^{-10.82} dl^i [a\mathcal{N}(\mu_1,\sigma_1{}^2) +
-%    (1-a)\mathcal{N}(\mu_2,\sigma_2{}^2] 
-% p(l_o^i | l^i) 
-%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
-%&+\int_{-10.82}^{\infty} dl^i\mathcal{N}(\mu_1,\sigma_1{}^2)
-% p(l_o^i | l^i) 
-%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber
-%\end{align}
-%Note that you use rates without LsSFR measurements, so you are doing something
-%inconsistent here. There should be a $\delta{z}$ term here. However, since the
-%rates are the same at $z=0.05$, maybe this doesn't matter so much.
-%}
+\agk{You don't define $p^i_y$.  The language is
+is sloppy.  What is relevant is the opposite, the probability
+that a galaxy with low mass gets the measured LsSFR.
+These are in general different.
+The solution you use is only correct for a model in which the
+rates for low and high-LsSFR galaxies is the same.
+The probability in general should include a prior
+on $l$.
+\begin{align}
+p(x^i_{o1},l_o^i | \mu_1,\mu_2,\sigma_1,\sigma_2,a) & = \int dl^i p(x^i_{o1},l_o^i | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
+p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
+ & = \int dl^i p(x^i_{o1} | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
+ p(l_o^i | li) 
+p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
+& = \int_{-\infty}^{-10.82} dl^i [a\mathcal{N}(\mu_1,\sigma_1{}^2) +
+    (1-a)\mathcal{N}(\mu_2,\sigma_2{}^2] 
+ p(l_o^i | l^i) 
+p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
+&+\int_{-10.82}^{\infty} dl^i\mathcal{N}(\mu_1,\sigma_1{}^2)
+ p(l_o^i | l^i) 
+p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber
+\end{align}
+Note that you use rates without LsSFR measurements,
+so you are doing something inconsistent here.
+There should be a $\delta{z}$ term here.
+However, since the rates are the same at $z=0.05$,
+maybe this doesn't matter so much.
+}
 
@@
 (resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
-predicted redshift evolution of mean stretch \mr{(expected $x_1$ given the
-distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
-Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances}.
-We see in this figure that the measured mean SN~Ia stretch per redshift bins of
-equal sample size closely follows our redshift drift modeling. This is indeed
-what is expected if old environments favor low SN stretches
-\citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
-a function of redshift. See Section~\ref{sec:results} for a more quantitative
-discussion.
+predicted redshift evolution of mean stretch (mean of Eq.~\ref{eq:stretchz}
+\agk{This is not the mean of Eq.~4.  It is the expected
+value of $x_1$ given the distribution of Eq.~4.
+Are the uncertainties in the parameters (and their
+covariances) taken into account in this expected
+value and the shaded blue band?  If not, they be.})
+illustrated as a blue band in Fig.~\ref{fig:modelall}. We see in this figure
+that the measured mean SN~Ia stretch per redshift bins of equal sample size
+closely follows our redshift drift modeling. This is indeed what is expected if
+old environments favor low SN stretches \citep[e.g.][]{howell2007} \emph{and} if
+the fraction of old SNe~Ia declines as a function of redshift. See
+Section~\ref{sec:results} for a more quantitative discussion.
 
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf-hist_on.pdf}
+    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
     \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
@@
         are compatible. The light-blue band illustrates the amplitude of the
-        error (incl. covariance) of the best fit model when considering the
-        fiducial dataset. \mr{The redshift distributions of each sample are
-        shown as light-histograms, see lower left legend.}}
+    error of the best fit model when considering the fiducial dataset. \agk{Confirming this band includes parameter covariance.}}
     \label{fig:modelall}
@@
 having a sample-based approach covers some potential redshift evolution
-information \citep{scolnic2016, scolnic2018a}. See further discussion concerning
-BBC in Section~\ref{sec:discussion}.
-
-Finally, for the sake of completeness, we also consider Gaussian
-and Asymmetric Gaussian non-drifting models.
+information \citep{scolnic2016, scolnic2018a}. See further discussion
+concerning BBC in Section~\ref{sec:discussion}.
+
+Finally, for the sake of completeness, we also consider simple \agk{?} the ``Gaussian''
+and ``Asymmetric'' Gaussian non-drifting models.
 
@@
 are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
-a Gaussian probability distribution. \mr{However, we highlight that the best of
-the compared models might not be the true underlying model.}
-
-The best model (with smallest AIC) is the so-called Base model and thus is out
-the reference model. This is true both on the fiducial and conservative samples.
+a Gaussian probability distribution.
+
+We first note that the best model (with smallest AIC) is the so-called Base
+model, implementing the \cite{rigault2018} description of the stretch
+distribution and its redshift evolution. This is therefore the reference model,
+both on the fiducial and conservative samples.
 
@@
 Surprisingly, the sample-based Gaussian asymmetric modeling used by current
-implementations of the BBC technique \citep{scolnic2016, kessler2017} has one of
-the highest AIC value in our analysis (see Section~\ref{sec:results}). While its
-$-2\ln(L)$ is the smallest of all redshift-independent models (but still $-11.5$
-worse than the reference Base model), it is strongly penalized for requiring
-15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the 5~samples of the
-analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted as a
-probability $p=2\times 10^{-7}$ of being an as good representation of the data
-as the Base model.
-\mr{Remark that the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
-k\ln(n)$, with $n$ is the number of datapoint) might be better suited than AIC
-when comparin models to the sample-based one since it explicitely accounts for
-the fact that each sample is fitted separately; the sample-based model BIC is
-the sum of the BIC per sample. Doing so we find $\Delta\mathrm{BIC}=-48$, also
-excluding the sample-base asymmetric Gaussian model as an as good description of
-the data then the Base-model.}
-
-%\agk{I think it does worse not because of the more parameters but that it has 5
-%complexity penalties whereas the base model only has 1.  AIC doesn't account
-%for the fact that each sample is fit independently.  It seems a fairer metric
-%would be the BIC, where you adjust $n$ for each sample.  This would recover
-%$\sim 16$ in the IC for this model.}
-
-We report in Table~\ref{tab:bbc} the samples' $\mu_0$ and $\sigma_{\pm}$
-adjusted on the nominally selection-free samples using our fiducial cuts (see
-Section~\ref{sec:sample}). We find our results in close agreement with
-\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
-derived these model parameters using the full BBC formalism, using numerous
-simulations to model the selection effects (see details e.g., Section~3 of
-\citealt{kessler2017}). The agreement between our fit of the asymmetric
+implementations of the BBC technique \citep{scolnic2016, kessler2017}, is one of
+the worst
+\agk{the word ``worst'' is inappropriate for the most highly AIC excluded
+model}
+modelisations of our analysis (see Section~\ref{sec:results}). While
+its $-2\ln(L)$ is the smallest of all redshift-independent models (but still
+$-11.5$ worse than the reference Base model), it is strongly penalized for
+requiring 15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the 5~samples of
+the analysis).
+\agk{I think it does worse not because of the more parameters but that it has 5 complexity penalties
+whereas the base model only has 1.  AIC doesn't account for the fact that each sample is fit independently.  It seems a fairer
+metric would be the BIC, where you adjust $n$ for each sample.  This would recover $\sim 16$ in the
+IC for this model.}
+Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted
+as a probability $p=2\times 10^{-7}$ of being an as good representation of the
+data as the Base model. We report in Table~\ref{tab:bbc} the samples' $\mu_0$
+and $\sigma_{\pm}$ adjusted on the nominally selection-free samples using our
+fiducial cuts (see Section~\ref{sec:sample}). We find our results in close
+agreement with \cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a}
+for PS1, who derived these model parameters using the full BBC formalism, using
+numerous simulations to model the selection effects (see details e.g., Section~3
+of \citealt{kessler2017}). The agreement between our fit of the asymmetric
 Gaussians on the supposedly selection-free part of the samples and the results
@@
 
-Finally, ignoring the LsSFR measurements --~available only for the SNfactory
-dataset, see Section~\ref{sec:modeling}~-- reduces the significance of the
-results presented in this section, as expected. Yet, non-drifting models remain
-strongly disfavored, and for instance, the best fitted sample-based Gaussian
-asymmetric modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the
-data than our Base drifting modeling.
+Finally, ignoring
+the LsSFR measurements --~available only for the SNfactory dataset, see
+Section~\ref{sec:modeling}~-- reduces the significance of the results
+presented in this section, as expected. Yet, non-drifting models remain strongly
+disfavored, and for instance, the best fitted sample-based Gaussian asymmetric
+modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the data than
+our Base drifting modeling.
 
@@
 redshifts. This means that the SN bias-corrected standardized magnitude
-estimated at a redshift affected by selection effect would be biased by an
+estimated at redshifts affected by selection effect would be biased by an
 ill-modeling of the true underlying stretch distribution.
@@
 asymmetric modeling as part of the BBC simulations.
-\mr{However, we already highlight that even if a non-drifting sample-based model
-could provide comparable result in the volume-limited part of the various
-samples, these models would differ when extrapolating at higher redshift,
-precisely where the  underlying will matter for correcting Malmquist biases.}
-
-%Nonetheless, it is enlightening to study the expected mean stretch difference
-%as a function of redshift when considering that the underlying SN stretch
-%distribution follows either the per-sample Asymmetric Gaussian modeling or our
-%Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
-%we convert this difference in mean stretch $x_1$ to a difference in magnitude
-%using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
-%correspond to the actual impact on cosmological distances derived from SNe~Ia
-%as this simple stretch average comparison does not account for the amplitude of
-%the Malmquist bias correction to be applied as a function of redshift and
-%surveys.
-
-%\agk{ The AG modeling contains Malmquist bias. To be fair you should plot Base
-%with the survey-specific Malmquist bias. There is something strange going on at
-%$z<0.2$, can you add SNfactory to this plot? I think you fill find that the
-%systematic difference is driven by differences in the $x_1$ distribution of
-%SNf.}
-
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
-%    \caption{Difference of mean stretch standardization (in mag, using
-%        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
-%        Gaussian asymmetric modeling and our Base drifting model, as a function
-%        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
-%        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
-%        considering the individual SN redshifts (no redshift cut applied). The
-%        orange markers represent their mean evolution in 30 regularly-spaced
-%        redshift bins. This must be multiplied by the amplitude of the Malmquist
-%        bias correction to derive the SN distance modulus bias affecting
-%        cosmology, see text. \nn{\textit{Apparently, there is some issue with
-%        the mean of all surveys combined. See Greg's comments.}}}
-%    \label{fig:magdrift}
-%\end{figure}
-
-%We see that the amplitude of the modeling difference is a few tens of millimag.
+
+Nonetheless, it is enlightening to study the expected mean stretch difference as
+a function of redshift when considering that the underlying SN stretch
+distribution follows either the per-sample Asymmetric Gaussian modeling or our
+Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
+we convert this difference in mean stretch $x_1$ to a difference in magnitude
+using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
+correspond to the actual impact on cosmological distances derived from SNe~Ia as
+this simple stretch average comparison does not account for the amplitude of the
+Malmquist bias correction to be applied as a function of redshift and surveys.
+\agk{
+The AG modeling contains Malmquist bias.  To be fair you
+should plot Base with the survey-specific Malmquist bias.
+There is something strange going on at $z<0.2$,
+can you add SNfactory to this plot?  I think you
+fill find that the systematic difference
+is driven by differences in the $x_1$ distribution
+of SNf.}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
+    \caption{Difference of mean stretch standardization (in mag, using
+        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
+        Gaussian asymmetric modeling and our Base drifting model, as a function
+        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
+        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
+        considering the individual SN redshifts (no redshift cut applied). The
+        orange markers represent their mean evolution in 30 regularly-spaced
+        redshift bins. This must be multiplied by the amplitude of the Malmquist
+        bias correction to derive the SN distance modulus bias affecting
+        cosmology, see text. \nn{\textit{Apparently, there is some issue with
+        the mean of all surveys combined. See Greg's comments.}}}
+    \label{fig:magdrift}
+\end{figure}
+
+We see that the amplitude of the modeling difference is a few tens of millimag.
 In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
@@
         younger and older SNe~Ia follows the prediction made in
-        \cite{rigault2018}. This second result strongly supports the existence
-        of both young and old SN~Ia populations, in agreement with rate studies.
+        \cite{rigault2018}.
 
@@
         even less so as survey span larger redshift ranges. We warn that, even
-        if extra degrees of freedom might be acceptable given the large
-        number of SNe~Ia in cosmological studies, extrapolating the SN property
+        if extra degrees of freedom might be acceptable \nn{given the large
+        number of SNe~Ia in} cosmological studies, extrapolating the SN property
         distributions from the volume-limited part of a survey to the
@@
 
-    %\item An inaccurate underlying stretch distribution model is estimated to
-    %    bias the mean standardized magnitude of SNe~Ia affected by
-    %    selection effects at the order of a few percents. This is
-    %    comparable in scale to the observational signature of %exotic forms of
-    %    dark energy and is therefore a significant source of systematic that
-    %    should be carefully studied.
+    \item An inaccurate underlying stretch distribution model is estimated to
+        bias the mean standardized magnitude of SNe~Ia affected by
+        selection effects at the order of a few percents. This is
+        comparable in scale to the observational signature of exotic forms of
+        dark energy and is therefore a significant source of systematic that
+        should be carefully studied.
 
@@
 
-\bibitem[Aldering et al.(2020)]{aldering2020} Aldering, G., Antilogus, P.,
-Aragon, C., et al.\ 2020, Research Notes of the American Astronomical Society,
-4, 63
-
 \bibitem[Astier et al.(2006)]{astier2006} Astier, P., Guy, J., Regnault, N., et

### 2020-05-13 Nicolas.tex
MR answered before 48h to go

@@
 stretch has been shown to strongly correlate with the SN's environment, notably
-with stellar age tracers. Following the evolution of fraction of young and old
-SNe~Ia as predicted by \cite{rigault2018}, and assuming constant underlying
-stretch distribution for each population as a function of redshift, model the
-expected SN~Ia stretch distribution as a function of redshift. We test our
-prediction against data from the literature chosen to have negligible selection
+with stellar age tracers.
+We model the underlying stretch distribution as a function of reshift, using the evolution of fraction of young and old SNe~Ia as predicted by \cite{rigault2018}, and assuming constant underlying stretch distribution for each age population made of Gaussian-mixtures.
+We test our prediction against data from the literature chosen to have negligible selection
 effects to ensure that any observed drift is indeed astrophysical and not
@@
 drifting model is a much better description of the data than any non-drifting
-model, including the sample-based asymmetric distributions used by the Beams
-with Bias Correction algorithm to correct Malmquist bias. Our favored underlying
-stretch model is bimodal: a high-stretch mode shared by both young and old
+model, including the sample-based asymmetric distributions usually used to correct Malmquist bias. Our favored underlying stretch model is bimodal: a high-stretch mode shared by both young and old
 environments, and a low-stretch mode exclusive to old environments. The precise
@@
 
-The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate.
-\agk{This grammatical error appears a lot in this article.  In compound
-nouns the first noun is typically singular, so this should
-be SN~Ia age dichotomy.}
+The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
 \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
@@
     \caption{Composition of the SNe~Ia dataset used in this analysis.
-    Conservative cuts are indicated in brackets.
-    \agk{parentheses not brackets}}
+    Conservative cuts are indicated in parentheses.}
     \label{tab:sample}
@@
 been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
-and~4.2.2; SNe~Ia time series are published in \citealt{Saunders2020}
-\nn{\textit{Greg: for now you may need to cite SNEMO and SUGAR to remove this
-``?''}}, see also Aldering in prep. and Ponder in prep.). The spectroscopic
+and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also Aldering in prep. and Ponder in prep.). The spectroscopic
 follow-up was done over a redshift range of $0.02 < z < 0.09$~\citep[as
@@
 underlying stretch distribution is drifting.
-\agk{how can this be asserted when you just showed that none of the surveys
-are complete at $z \sim 0.65$!?} \mr{MR: Alex, There is HST too, see top of Fig. 3. | in the latest bin 68 SNLS, 26 HST}
 
@@
 But generally, the fraction of young SNe~Ia as a function of redshift is given
-by $\delta$(z) \agk{Convention is that letters that are variables
-are italicized.  $z$, $H_0$.}
-(see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
+by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
 of the underlying distribution of SNe~Ia as a function of redshift $X_1(z)$ is
@@
 
-For the SNfactory sample, we can readily set $y^i = p_y^i$, the probability to
-have $\log \mathrm{LsSFR} \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
-minimize Eq.~\ref{eq:likelihood} with respect to~$\vec{\theta}$. Results on
+For the SNfactory sample, we can readily set $y^i = p_y^i$, with the $p_y^i$ the probability that $\log \mathrm{LsSFR} \geq -10.82$ given the measurement errors (see Fig.~\ref{fig:stretchlssfr}), to
+minimize Eq.~\ref{eq:likelihood} with respect to~$\vec{\theta}$.
+\mr{This assumes that the LsSFR is a perfect age tracers (see Briday et al.
+in~prep) and that the redshift drift within the SNfactory sample is negligible such that the likelihood and the posterior distribution of $p_y$ are equal  ($\Delta\delta(z)<3\%$ from the highest to the lowest redshift of the dataset).}
+Results on
 fitting the SNf SNe with this model are shown Table~\ref{tab:modelresults} and
@@
 \end{table*}
-\agk{You don't define $p^i_y$.  The language is
-is sloppy.  What is relevant is the opposite, the probability
-that a galaxy with low mass gets the measured LsSFR.
-These are in general different.
-The solution you use is only correct for a model in which the
-rates for low and high-LsSFR galaxies is the same.
-The probability in general should include a prior
-on $l$.
-\begin{align}
-p(x^i_{o1},l_o^i | \mu_1,\mu_2,\sigma_1,\sigma_2,a) & = \int dl^i p(x^i_{o1},l_o^i | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
-p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
- & = \int dl^i p(x^i_{o1} | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
- p(l_o^i | li) 
-p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
-& = \int_{-\infty}^{-10.82} dl^i [a\mathcal{N}(\mu_1,\sigma_1{}^2) +
-    (1-a)\mathcal{N}(\mu_2,\sigma_2{}^2] 
- p(l_o^i | l^i) 
-p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
-&+\int_{-10.82}^{\infty} dl^i\mathcal{N}(\mu_1,\sigma_1{}^2)
- p(l_o^i | l^i) 
-p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber
-\end{align}
-Note that you use rates without LsSFR measurements,
-so you are doing something inconsistent here.
+%\agk{You don't define $p^i_y$.  The language is
+%is sloppy.  What is relevant is the opposite, the probability
+%that a galaxy with low mass gets the measured LsSFR.
+%These are in general different.
+%The solution you use is only correct for a model in which the
+%rates for low and high-LsSFR galaxies is the same.
+%The probability in general should include a prior
+%on $l$.
+%\begin{align}
+%p(x^i_{o1},l_o^i | \mu_1,\mu_2,\sigma_1,\sigma_2,a) & = \int dl^i p(x^i_{o1},l_o^i | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
+%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
+% & = \int dl^i p(x^i_{o1} | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
+% p(l_o^i | li) 
+%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
+%& = \int_{-\infty}^{-10.82} dl^i [a\mathcal{N}(\mu_1,\sigma_1{}^2) +
+%    (1-a)\mathcal{N}(\mu_2,\sigma_2{}^2] 
+% p(l_o^i | l^i) 
+%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
+%&+\int_{-10.82}^{\infty} dl^i\mathcal{N}(\mu_1,\sigma_1{}^2)
+% p(l_o^i | l^i) 
+%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber
+%\end{align}
+%Note that you use rates without LsSFR measurements, so you are doing something inconsistent here.
 There should be a $\delta{z}$ term here.
-However, since the rates are the same at $z=0.05$,
-maybe this doesn't matter so much.
-}
+However, since the rates are the same at $z=0.05$, maybe this doesn't matter so much.
+%}
 
@@
 Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
-(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
-predicted redshift evolution of mean stretch (mean of Eq.~\ref{eq:stretchz}
-\agk{This is not the mean of Eq.~4.  It is the expected
-value of $x_1$ given the distribution of Eq.~4.
-Are the uncertainties in the parameters (and their
-covariances) taken into account in this expected
-value and the shaded blue band?  If not, they be.})
-illustrated as a blue band in Fig.~\ref{fig:modelall}. We see in this figure
+(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the predicted redshift evolution of mean stretch \mr{(expected $x_1$ given the distribution of Eq.~\ref{eq:likelihood})
+illustrated as a blue band in Fig.~\ref{fig:modelall} accounting for parameters errors (and their covariances)}. 
+We see in this figure
 that the measured mean SN~Ia stretch per redshift bins of equal sample size
@@
         the conservative one or the SNfactory dataset only, respectively; all
-        are compatible. The light-blue band illustrates the amplitude of the
-    error of the best fit model when considering the fiducial dataset. \agk{Confirming this band includes parameter covariance.}}
+        are compatible.
+The light-blue band illustrates the amplitude of the error (incl.
+covariance) of the best fit model when considering the fiducial dataset.
+\mr{The redshift distributions of each sample are shown as light-histograms, see lower left legend.}}
     \label{fig:modelall}
@@
 
-Finally, for the sake of completeness, we also consider simple \agk{?} the ``Gaussian''
-and ``Asymmetric'' Gaussian non-drifting models.
+Finally, for the sake of completeness, we also consider Gaussian
+and Asymmetric Gaussian non-drifting models.
 
@@
 $5<\Delta\mathrm{AIC}<20$ are unfavored, and those with $\Delta\mathrm{AIC}>20$
-are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
-a Gaussian probability distribution.
-
-We first note that the best model (with smallest AIC) is the so-called Base
-model, implementing the \cite{rigault2018} description of the stretch
-distribution and its redshift evolution. This is therefore the reference model,
-both on the fiducial and conservative samples.
+are deemed excluded.
+This roughly corresponds to 2, 3 and 5~$\sigma$ limits for a Gaussian probability distribution.
+\mr{We however highlight that this best of the compared model might however not be the true underlying model.}
+
+The best model (with smallest AIC) is the so-called Base
+model and thus is out the reference model. This is true both on the fiducial and conservative samples.
 
@@
 
-Surprisingly, the sample-based Gaussian asymmetric modeling used by current
-implementations of the BBC technique \citep{scolnic2016, kessler2017}, is one of
-the worst
-\agk{the word ``worst'' is inappropriate for the most highly AIC excluded
-model}
-modelisations of our analysis (see Section~\ref{sec:results}). While
+Surprisingly, the sample-based Gaussian asymmetric modeling used by current implementations of the BBC technique \citep{scolnic2016, kessler2017} has one of the highest AIC value
+in our analysis (see Section~\ref{sec:results}). While
 its $-2\ln(L)$ is the smallest of all redshift-independent models (but still
@@
 the analysis).
-\agk{I think it does worse not because of the more parameters but that it has 5 complexity penalties
-whereas the base model only has 1.  AIC doesn't account for the fact that each sample is fit independently.  It seems a fairer
-metric would be the BIC, where you adjust $n$ for each sample.  This would recover $\sim 16$ in the
-IC for this model.}
+%\agk{I think it does worse not because of the more parameters but that it has 5 complexity penalties
+%whereas the base model only has 1.  AIC doesn't account for the fact that each sample is fit independently.  It seems a fairer
+%metric would be the BIC, where you adjust $n$ for each sample.  This would recover $\sim 16$ in the
+%IC for this model.}
 Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted
@@
 asymmetric modeling as part of the BBC simulations. 
-
-Nonetheless, it is enlightening to study the expected mean stretch difference as
-a function of redshift when considering that the underlying SN stretch
-distribution follows either the per-sample Asymmetric Gaussian modeling or our
-Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
-we convert this difference in mean stretch $x_1$ to a difference in magnitude
-using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
-correspond to the actual impact on cosmological distances derived from SNe~Ia as
-this simple stretch average comparison does not account for the amplitude of the
-Malmquist bias correction to be applied as a function of redshift and surveys.
-\agk{
-The AG modeling contains Malmquist bias.  To be fair you
-should plot Base with the survey-specific Malmquist bias.
-There is something strange going on at $z<0.2$,
-can you add SNfactory to this plot?  I think you
-fill find that the systematic difference
-is driven by differences in the $x_1$ distribution
-of SNf.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
-    \caption{Difference of mean stretch standardization (in mag, using
-        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
-        Gaussian asymmetric modeling and our Base drifting model, as a function
-        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
-        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
-        considering the individual SN redshifts (no redshift cut applied). The
-        orange markers represent their mean evolution in 30 regularly-spaced
-        redshift bins. This must be multiplied by the amplitude of the Malmquist
-        bias correction to derive the SN distance modulus bias affecting
-        cosmology, see text. \nn{\textit{Apparently, there is some issue with
-        the mean of all surveys combined. See Greg's comments.}}}
-    \label{fig:magdrift}
-\end{figure}
-
-We see that the amplitude of the modeling difference is a few tens of millimag.
-In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
-level and $w_a$ at the ten-percent precision \citep[e.g.,][]{lsstpaper}, we
-stress that correctly modeling of potential SN redshift drift should be further
-studied and care should be made when using samples affected by selection
-effects.
+\mr{However, we already highlight that even if the a non-drifting sample-based model could provide comparable result in the volume-limited part of the various samples, these models would differ when extrapolating at higher redshift, precisely where the  underlying will matter for correcting Malmquist biases.}
+
+%Nonetheless, it is enlightening to study the expected mean stretch difference as a function of redshift when considering that the underlying SN stretch distribution follows either the per-sample Asymmetric Gaussian modeling or our Base drifting model.
+This is shown Fig.~\ref{fig:magdrift}, where, for clarity we convert this difference in mean stretch $x_1$ to a difference in magnitude using $\alpha=0.156$ \citep{scolnic2018a}.
+We stress here that this does not correspond to the actual impact on cosmological distances derived from SNe~Ia as this simple stretch average comparison does not account for the amplitude of the Malmquist bias correction to be applied as a function of redshift and surveys.
+%\agk{
+%The AG modeling contains Malmquist bias.  To be fair you
+%should plot Base with the survey-specific Malmquist bias.
+%There is something strange going on at $z<0.2$,
+%can you add SNfactory to this plot?  I think you
+%fill find that the systematic difference
+%is driven by differences in the $x_1$ distribution
+%of SNf.}
+
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
+%    \caption{Difference of mean stretch standardization (in mag, using
+%        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
+%        Gaussian asymmetric modeling and our Base drifting model, as a function
+%        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
+%        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
+%        considering the individual SN redshifts (no redshift cut applied). The
+%        orange markers represent their mean evolution in 30 regularly-spaced
+%        redshift bins. This must be multiplied by the amplitude of the Malmquist
+%        bias correction to derive the SN distance modulus bias affecting
+%        cosmology, see text. \nn{\textit{Apparently, there is some issue with
+%        the mean of all surveys combined. See Greg's comments.}}}
+%    \label{fig:magdrift}
+%\end{figure}
+
+%We see that the amplitude of the modeling difference is a few tens of millimag.
+In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent level and $w_a$ at the ten-percent precision \citep[e.g.,][]{lsstpaper}, we stress that correctly modeling of potential SN redshift drift should be further studied and care should be made when using samples affected by selection effects.
 
@@
         younger and older SNe~Ia follows the prediction made in
-        \cite{rigault2018}.
+        \cite{rigault2018}. This second result strongly supports the existence of both young and old SN~Ia populations, in agreement with rate studies.
 

### 2020-05-16 Nicolas.tex
Should go Monday

--- 2020-05-13 Nicolas.tex
@@
-\documentclass[]{aa} %bibnumber
-
-%\usepackage{hyperref}
+\documentclass[]{aa}
+
 \usepackage{txfonts}
@@
 \usepackage[colorlinks,breaklinks]{hyperref}
-\usepackage[normalem]{ulem}
-\usepackage{cancel}
 \hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}
@@
 \usepackage[dvipsnames]{xcolor}
-\newcommand{\mr}[1]{\textcolor[rgb]{0.60,0.10,0.6}{#1}}
-\newcommand{\mri}[1]{\textcolor{red}{#1}}
-\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
-\newcommand{\nnn}[1]{\nn{\textit{#1}}}
-\newcommand{\yc}[1]{\textcolor{BrickRed}{#1}}
-\newcommand{\agk}[1]{\textcolor{purple}{AK: #1}}
 
@@
     \and R. Graziani \inst{2}
-    %HERE ALPHABETIC ORDER
     \and G. Aldering\inst{3}
@@
     \and J. Nordin\inst{4}
-    \and A.~G. Kim\inst{3}
     \and Y.-L. Kim\inst{1}
+    \orcid{0000-0002-1031-0796}
     \and Saul Perlmutter\inst{3}
-    % Gerard ne souhaite pas etre co-auteur.
 }
@@
     systematic uncertainties rises, notably that of the SN~Ia population
-    redshift drift. In this paper, we study the redshift evolution of SN~Ia
-    \textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN
-    property, to probe this drift. \yc{\emph{isn't a 'drift' \textbf{exactly} a
-    redshift evolution?}} The SN stretch has been shown to strongly correlate
-    with the SN environment, notably with stellar age tracers. We model the
-    underlying stretch distribution as a function of redshift, using the
-    evolution of fraction of young and old SNe~Ia as predicted by
-    \cite{rigault2018}, and assuming constant underlying stretch distribution
-    for each age population made of Gaussian mixtures. We test our prediction
-    against published samples chosen to have negligible magnitude selection
-    effects, so that any observed change is indeed of astrophysical and not
-    observational origin. We clearly demonstrate that the underlying SN~Ia
-    stretch distribution is evolving as a function of redshift, and that the
-    young/old drifting model is a much better description of the data than any
-    \yc{\sout{non-drifting} constant} model, including the sample-based
-    asymmetric distributions usually used to correct Malmquist bias. The favored
-    underlying stretch model is \yc{the bimodal one introduced in
-    \cite{rigault2018}} \nnn{to my knowledge, R+18 doesn't introduce a
-    stretch model, but rather discusses the differences between the stretch
-    distributions for the young and old pop, never saying anything about stretch
-    modes}: a high-stretch mode shared by both young and old environments, and a
-    low-stretch mode exclusive to old environments. The precise impact of the
-    redshift evolution of the SN Ia population intrinsic properties on
-    cosmology remains to be studied. Yet, the astrophysical drift of the SN stretch
-    distribution does affect current Malmquist bias corrections and thereby
-    distances derived from SN affected by selection effects. We highlight that such
-    a bias will increase with surveys covering increasingly larger redshift
-    ranges, which is particularly important for LSST.}
+    evolution. In this paper, we study the dependence with redshift of the SN~Ia
+    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
+    probe its potential redshift drift. The SN stretch has been shown to
+    strongly correlate with the SN environment, notably with stellar age
+    tracers. We model the underlying stretch distribution as a function of
+    redshift, using the evolution of fraction of young and old SNe~Ia as
+    predicted by \cite{rigault2018}, and assuming constant underlying stretch
+    distribution for each age population made of Gaussian mixtures. We test our
+    prediction against published samples chosen to have negligible magnitude
+    selection effects, so that any observed change is indeed of astrophysical
+    and not observational origin. We clearly demonstrate that the underlying
+    SN~Ia stretch distribution is evolving as a function of redshift, and that
+    the young/old drifting model is a much better description of the data than
+    any time-constant model, including the sample-based asymmetric distributions
+    usually used to correct Malmquist bias. The favored underlying stretch model
+    is the bimodal one derived from \cite{rigault2018}: a high-stretch mode
+    shared by both young and old environments, and a low-stretch mode exclusive
+    to old environments. The precise impact of the redshift evolution of the SN
+    Ia population intrinsic properties on cosmology remains to be studied. Yet,
+    the astrophysical drift of the SN stretch distribution does affect current
+    Malmquist bias corrections and thereby distances derived from SN affected by
+    selection effects. We highlight that such a bias will increase with surveys
+    covering increasingly larger redshift ranges, which is particularly
+    important for LSST.}
 
@@
 \maketitle
-
-\yc{\emph{IMHO too many occurrences of `drift' (53!); this could be replaced by
-'[redshift] evolution' to be opposed to 'constant'.}}
 
@@
 \citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
-Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
-\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
+Hubble Constant ($H_0$), provided one can calibrate their absolute magnitude
+\citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ derived when
 the SNe~Ia are anchored on Cepheids \citep[the SH0ES project,][]{riess2009,
@@
 \citep{feeney2019}. While using the tip of the red giant branch technique in
-place of the Cepheids seem to favor an intermediate value of H$_0$
+place of the Cepheids seem to favor an intermediate value of $H_0$
 \citep{freedman2019, freedman2020}, time delay measurements from strong lensing
-seem to also favor high H$_0$ values \citep{wong2019}.
-
-The H$_0$ tension has received a lot of attention, as it could be a sign of new
-fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
+seem to also favor high $H_0$ values \citep{wong2019}.
+
+The $H_0$ tension has received a lot of attention, as it could be a sign of new
+fundamental physics. Yet, no simple solution is able to accommodate this $H_0$
 tension when accounting for all other probes \citep{knox2019}, but see e.g.
@@
 strongly favors young stellar populations, while the latter not. This
-selection effect would impact the derivation of H$_0$ if SNe~Ia from young and
+selection effect would impact the derivation of $H_0$ if SNe~Ia from young and
 older environments differ in average standardized magnitudes.
@@
 cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including for
-deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
+deriving $H_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
 between the SNe and their host remains unclear when using global properties such
@@
 especially the impact of such an astrophysical bias has on the derivation of
-H$_0$ \citep{jones2015, riess2016, riess2018, rose2019}.
-
-%\nn{We point that there has been sample-based approach meant to account for an
-%unmodeled variation of host-related parameters thanks to multiple
-%redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show
-%aren't as effective as a direct modeling of such a drift}.
+$H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.
 
@@
 that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
-in Bessel $B$ band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect
+in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one ought to detect
 this object typically 5 days before and a week after peak to build a suitable
 lightcurve, the effective limiting standardized absolute magnitude is
-approximately $M_{lim} = -18.00$~mag. Hence, given the magnitude limit $m_{lim}$
-of a magnitude limited survey, one can derive the maximum redshift $z_{lim}$
+approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit $m_{\lim}$
+of a magnitude limited survey, one can derive the maximum redshift $z_{\lim}$
 above which the faintest SNe~Ia will be missed using the relation between
-apparent magnitude, redshift and absolute magnitude $\mu(z_{lim}) = m_{lim} -
-M_{lim}$.
+apparent magnitude, redshift and absolute magnitude $\mu(z_{\lim}) = m_{\lim} -
+M_{\lim}$.
 
@@
         $\beta$ coefficients from \cite{scolnic2018a}. The grey diagonal lines
-        represent the $(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between
-        $0.50$ and $z=1.70$ using SNLS's $m_{lim}$ of $24.8$~mag.}
+        represent the $(x_1, c)$ evolution for $m = m_{\lim}$, for $z$ between
+        $0.50$ and $z=1.70$ using SNLS's $m_{\lim}$ of $24.8$~mag.}
     \label{fig:maglim}
@@
 24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
-final release website.}}. This converts to a $z_{lim}=0.60$, in agreement with
+final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
 \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
 \cite[][see their Section~5]{perrett2010} suggests a lower limit of
-$z_{lim}=0.55$; both limits will be considered, as discussed below.
+$z_{\lim}=0.55$; both limits will be considered, as discussed below.
 
 Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
-$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
+$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.30$ in
 agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure is also
-suggestive of a more conservative $z_{lim}=0.27$.
+suggestive of a more conservative $z_{\lim}=0.27$.
 
 In a similar redshift range, SDSS has a limiting magnitude of 22.5
-\citep{dilday2008, sako2008}, which would lead to a $z_{lim}=0.24$. However, the
+\citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However, the
 SDSS surveys were more sensitive to limited spectroscopic resources; \cite[][see
 their Section~2]{kessler2009} pointed out that during the first year of SDSS,
-SNe~Ia with $r-mag<20.5$ were favored for spectroscopic follow-up, corresponding
+SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
 to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
 spectroscopic resources were available, and \cite{kessler2009} and
-\cite{dilday2008} show a reasonable completeness up to $z_{lim}=0.2$. Following
-these analyses, we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit.
+\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
+these analyses, we will use $z_{\lim}=0.2$ as the baseline SDSS redshift limit.
 
@@
         \hline\hline
-        Survey & $z_{lim}$ & $N_{\mathrm{SN}}$ \\
+        Survey & $z_{\lim}$ & $N_{\mathrm{SN}}$ \\
         \hline
@@
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{lim}$ in Fig.~\ref{fig:sample}. The evolution of
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. The evolution of
 the mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb-cividis.pdf}
-
+    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_stac_75-lb-cividis.pdf}
     \caption{\textit{Bottom:} \textsc{\texttt{SALT2.4}} lightcurve stretch as a
@@
         legend). Solid (resp.\ open) markers correspond to the conservative
-        (resp.\ fiducial) redshift cuts. \textit{Top:} combined redshift
-        histogram in dark (resp.\ light) gray for the conservative (resp.
+        (resp.\ fiducial) redshift cuts. \textit{Top:} stacked redshift
+        histograms in dark (resp.\ light) colors for the conservative (resp.
         \ fiducial) redshift cuts.}
-
     \label{fig:sample}
@@
 
-Following \cite{rigault2018}, we define $p_y^i$ as one minus the cumulative
-distribution function of the $\log(\mathrm{LsSFR})$ at $-10.82\,\mathrm{dex}$.
-This is defining the probability that a SN progenitor is classified as
-young\nn{, thus assuming the LsSFR to be a perfect age tracer} (see further
-discussion about this assumption \yc{\emph{which assumption, if it's purely your
-own definition of young vs.\ old?}} in Briday et al.\ in~prep). \nn{This
-definition should be weighted by the $\delta(z)$ probability for a SN to be
-young, since the two effects play their roles, but} the drift is negligible
-within the SNfactory dataset ($\Delta\delta(z)<3\%$ from the highest to the
-lowest redshift of the dataset)\nn{; thus} one can readily use $y^i = p_y^i$ to
+For the SNfactory sample, we can readily set $y^i = p^i_y$, the probability to
+have $\log(\textrm{LsSFR}) \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
 minimize Eq.~\ref{eq:likelihood} with respect to $\vec{\theta}$. Results on
-fitting the SNfactory SNe with this model are shown Table~\ref{tab:modelresults}
-and illustrated in Fig.~\ref{fig:stretchlssfr}. \yc{\emph{Now I'm confused why
-one does not have \textbf{exactly} $y^i = p_y^i$ when LsSFR measurement is
-available, but you seem to say it's an assumption.}}
+fitting the SNf SNe with this model are shown Table~\ref{tab:modelresults} and
+illustrated in Fig.~\ref{fig:modelall}.
 
@@
                & $\mu_2$ & $\sigma_2$
-               & a \\
+               & $a$ \\
         \hline
-        SNfactory & $0.41 \pm 0.08$ & $0.55 \pm 0.06$
+        SNfactory & $ 0.41 \pm 0.08$ & $0.55 \pm 0.06$
                   & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$
-                  & $0.48^{+0.08}_{-0.08}$ \\
-        Fiducial & $0.37 \pm 0.05$  & $0.61 \pm 0.04$
+                  & $ 0.48 \pm 0.08$ \\
+        Fiducial & $ 0.37 \pm 0.05$ & $0.61 \pm 0.04$
                  & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$
-                 & $0.51^{+0.09}_{-0.10}$ \\
-        Conservative & $0.38 \pm 0.05$  & $0.60 \pm 0.04$
+                 & $ 0.51 \pm 0.09$ \\
+        Conservative & $ 0.38 \pm 0.05$ & $0.60 \pm 0.04$
                      & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$
-                     & $0.47^{+0.09}_{-0.08}$ \\
+                     & $ 0.47 \pm 0.09$ \\
         \hline
@@
         error (incl. covariance) of the best fit model when considering the
-        fiducial dataset. The redshift distributions of each sample are
-        shown as light-histograms, see lower left legend. \yc{\emph{This is upper panel of Fig.3? Why is it again split in subsamples, while it was not deemed necessary in dedicated Fig.3?}}}
+        fiducial dataset. 
+        %The redshift distributions of each sample are
+        %shown as light-histograms, see lower left legend.
+        }
     \label{fig:modelall}
@@
 
-Finally, for the sake of completeness, we also consider \yc{constant} Gaussian
-and Asymmetric Gaussian \yc{\sout{non-drifting}} models.
+Finally, for the sake of completeness, we also consider redshift-independent
+pure and asymmetric Gaussian models.
 
@@
         Name & drift & $k$ &
-        $-2\ln(L)$ & AIC & $\Delta$AIC & $-2\ln(L)$ & AIC & $\Delta$ AIC\\
+        $-2\ln(L)$ & AIC & $\Delta$AIC & $-2\ln(L)$ & AIC & $\Delta$AIC\\
         \hline
@@
 are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
-a Gaussian probability distribution. However, we highlight that the best of the
-compared models \nn{is not meant to} be the true underlying model. \nn{It only
-means that it is the best at describing the observations while minimizing the
-number of free parameters}. \yc{\emph{why this sentence? Is there any
-\textbf{true} intrinsic model?}}
-
-The best model (with smallest AIC) is the so-called Base model and thus is out
-the reference model. This is true both on the fiducial and conservative samples.
-
-% $\chi^2=-2\ln(L)-\ln(2\pi\sigma)$ ~ 900 so 1.6 per DoF.
+a Gaussian probability distribution.
+
+The best model (with smallest AIC) is the so-called Base model and thus is our
+reference model; this is true both on the fiducial and conservative samples.
+The Base model also has the smallest $-2\ln(L)$, making it the most likely
+model even ignoring the over-fitting issue accounted for by the AIC formalism.
 
@@
 several recent SN~Ia cosmological analyses, implemented as a per-sample
-asymmetric Gaussian \nn{stretch} distribution.
+asymmetric Gaussian stretch distribution.
 
@@
         existence of both young and old SN~Ia populations, in agreement with
-        rate studies. \yc{\emph{ref???}}
-
-    \item \nn{Models} using survey-based asymmetric Gaussian distributions, as
-        done, e.g., in the current implementation of the BBC, are excluded to be
-        good descriptions of the data relative to our drifting model. Hence,
-        the sample-based approach does not accurately account for redshift drift
-        and even less so as survey span increasingly larger redshift ranges. We
+        rate studies \cite{mannucci2005, scannapieco2005, sullivan2006,
+        aubourg2008}.
+
+    \item Models using survey-based asymmetric Gaussian distributions, as done,
+        e.g., in the current implementation of the BBC, are excluded to be good
+        descriptions of the data relative to our drifting model. Hence, the
+        sample-based approach does not accurately account for redshift drift and
+        even less so as survey span increasingly larger redshift ranges. We
         stipulate that, even if extra degrees of freedom might be acceptable
@@
 diagram would be particularly helpful for this drifting analysis; fortunately
-this will soon be provided by the Zwicky Transient Facility \citep{bellm2019,
-graham2019} \nn{on one hand}, and Subaru and SeeChange SNe~Ia programs \nn{on
-the other hand}, respectively. \nnn{need to check which is which}
+this will soon be provided by the Zwicky Transient Facility
+\citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
+programs (high-$z$), respectively.
 
@@
 
-\bibliographystyle{aa} % style aa.bst
+\bibliographystyle{aa}
 \begin{thebibliography}{} 
 % A
-
-%\bibitem[Abbott et al.(2018)]{abbott2018} Abbott, T.~M.~C., Abdalla, F.~B.,
-%Alarcon, A., et al.\ 2018, \prd, 98, 043526
 
@@
 
-%\bibitem[Ata et al.(2017)]{ata2017} Ata, M., Kitaura, F.-S., Chuang, C.-H., et
-%al.\ 2017, \mnras, 467, 3993
-
 \bibitem[Aubourg et al.(2008)]{aubourg2008} Aubourg, {\'E}., Tojeiro, R.,
@@
 
-%\bibitem[Chabanier et al.(2019)]{chabanier2019} Chabanier, S., Millea, M., \&
-%Palanque-Delabrouille, N.\ 2019, \mnras, 489, 2247
-
 \bibitem[Childress et al.(2013)]{childress2013} Childress, M., Aldering, G.,
@@
 
-%\bibitem[Coles, \& Jones(1991)]{coles1991} Coles, P., \& Jones, B.\ 1991,
-%\mnras, 248, 1
-
-%\bibitem[Conley et al.(2011)]{conley2011} Conley, A., Guy, J., Sullivan, M., et
-%al.\ 2011, \apjs, 192, 1
-
 % D
@@
 
-%\bibitem[Guy et al.(2010)]{guy2010} Guy, J., Sullivan, M., Conley, A., et al.\
-%2010, \aap, 523, A7
-
-%\bibitem[Graziani et al.(2019)]{graziani2019} Graziani, R., Courtois, H.~M.,
-%Lavaux, G., et al.\ 2019, \mnras, 488, 5438
-
 % H
@@
 
-%\bibitem[Howell et al.(2009)]{howell2009} Howell, D.~A., Sullivan, M., Brown,
-%E.~F., et al.\ 2009, \apj, 691, 661
-
 % I
@@
 
-%\bibitem[Planck Collaboration et al.(2016)]{planck2016} Planck Collaboration,
-%Ade, P.~A.~R., Aghanim, N., et al.\ 2016, \aap, 594, A13
-
 \bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration,
@@
 A., et al.\ 2018a, \apj, 859, 101
-
-%\bibitem[Scolnic et al.(2018)]{scolnic2018b} Scolnic, D.~M., Lochner, M., Gris,
-%P., et al.\ 2018, arXiv e-prints, arXiv:1812.00516
 

### 2020-05-19 Nicolas.tex
This is it! 1 column for referee

@@
-\documentclass[]{aa}
+\documentclass[referee]{aa}
 
@@
 \hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}
-
-\usepackage[dvipsnames]{xcolor}
 
@@
 
-\date{Received 2 November 1992 / Accepted 7 January 1993}
+\date{Submitted to A\&A the 19th of May 2020}
 
@@
         redshift bins of equal sample size, indicated in light gray at the
-        bottom of each redshift bins. Full and light markers are used when
+        bottom of each redshift bin. Full and light markers are used when
         considering the fiducial or the conservative samples, respectively. The
@@
         error (incl. covariance) of the best fit model when considering the
-        fiducial dataset. 
-        %The redshift distributions of each sample are
-        %shown as light-histograms, see lower left legend.
-        }
+        fiducial dataset.}
     \label{fig:modelall}
@@
 
-        Per Sample Asym. & per sample & 3$\times$5
+        Per sample Asym. & per sample & 3$\times$5
         & 1468.2 & 1498.2 & $-31.5$
@@
 Commonly used Malmquist bias correction techniques, such as the BBC-formalism,
-assume per Sample Asymmetric Gaussian functions for modeling the underlying
+assume per sample asymmetric Gaussian functions for modeling the underlying
 stretch and color distributions. Yet, as shown in Section~\ref{sec:results},
@@
         distribution is supposed to be a random draw from the underlying stretch
-        distribution. The green line shows the BBC model of this underlying
+        distribution. The green lines show the BBC model of this underlying
         distribution (asymmetric Gaussian). The full line (band) is our best fit
         (its error); the dashed line shows the \cite{scolnic2018a} result. The
-        black line (band) show our best fitted Base-modeling (its error, see
+        black line (band) shows our best fitted Base-modeling (its error, see
         Table \ref{tab:modelresults}) that includes redshift drift. For
@@
 level and $w_a$ at the ten-percent precision \citep[e.g.,][]{lsstpaper}, we
-stress that correctly modeling of potential SN redshift drift should be further
-studied and care should be made when using samples affected by selection
+stress that correct modeling of potential SN redshift drift should be further
+studied and care should be taken when using samples affected by selection
 effects.
@@
 In addition to this ``base'' modeling, we have studied various distributions,
-including redshift independent models; we also studied the prediction from the
-``Beams with Bias Correction'', a Malmquist bias correction algorithm used by
-several recent SN~Ia cosmological analyses, implemented as a per-sample
-asymmetric Gaussian stretch distribution.
+including redshift independent models; we also studied the prediction from a per-sample asymmetric Gaussian stretch distribution used, for instance, by the Beams with Bias Correction Malmquist bias correction algorithm \citep{scolnic2016, kessler2017}.
 

### 2020-06-17 referee_answer.tex
Started referee answer

\documentclass[11pt,a4paper]{paper}

\usepackage{preambule}
\fancyhead[L]{\scriptsize \textsc{Nora Nicolas}}

\begin{document}

\title{Referee answer}

\section{Of the surveys' spectroscopic follow-up}
\begin{enumerate}

    \item SNLS's detection efficiency $\e \approx 0$ for $i \approx
        \SI{24.8}{mag}$\smallbreak

        \textcolor{red}{A limiting magnitude of $m_{\lim} = \SI{23.5}{mag}
        \Rightarrow z_{\lim} = 0.36$, which would lead to only 26/236 SNe
    instead of 102/236 with our current cut}

    \item HST may have a follow-up efficiency that we should take into account
        like we did for SDSS;

    \item Misunderstanding about the 20\% of SNf's SNe that had selection
        effects:\smallbreak

        \textcolor{dgreen}{The 80\% of SNf's SNe that had no selection
effects are the 114 SNe that are in our sample}.
\end{enumerate}

\section{Of the $x_1$ bias that doesn't appear on $m$}
The referee insists on the fact that ``Biases in x1 are expected as a function
of redshift simply from survey modeling/selection effects'', showing the
following figure from \textsc{Kessler \& Scolnic 2017} to point
out that biases in $x_1$ become apparent much sooner that biases on $m$.

\begin{figure}[htbp!]
    \centering
    \begin{subfigure}
        \centering
        \includegraphics[width=.3\linewidth]{Answer_figures/Fig_biasCor_mb.eps}
    \end{subfigure}
    \begin{subfigure}
        \centering
        \includegraphics[width=.3\linewidth]{Answer_figures/Fig_biasCor_x1.eps}
    \end{subfigure}
    \begin{subfigure}
        \centering
        \includegraphics[width=.3\linewidth]{Answer_figures/Fig_biasCor_c.eps}
    \end{subfigure}
    \captionsetup{justification=centering}
    \caption{Bias corrections $\mkern
        1.5mu\overline{\mkern-1.5mu\delta\mkern-1.5mu}\mkern 1.5mu_{m_B}$,
        $\alpha \overline{\delta}_{x_1}$, and $\beta\overline{\delta}_{c}$ are
        shown as a function of redshift. The pre-factors $\alpha, \beta$ are
        used to show the bias in distance-modulus magnitudes. The parameter
    selection ranges are shown on each panel.}
    \label{fig:KS17}
\end{figure}

\begin{wrapfigure}[10]{R}{.4\linewidth}
    \vspace*{-20pt}
    \centering
    \includegraphics[width=\linewidth]{Answer_figures/hist_surveys2_btw_cividis.pdf}
    \captionsetup{justification=centering}
    \caption{\small Redshift histograms of SNe~Ia from the SDSS and PS1 datasets
    respectively}
    \label{fig:hists}
\end{wrapfigure}
Yet, from the same figure, biases in $c$ appear at $z=0$. Here we want to
convince the referee that if we don't find any sign of color bias in our sample,
we may consider little to no bias on $x_1$. We thus studied the $x_1$ and $c$
distributions of the end of SDSS and the start of PS1, for $0.10 < z < 0.20$. In
this redshift range, the SDSS cut dataset contains the most questionable SNe~Ia,
for the SNe between $0.15 < z < 0.20$ are between our conservative and fiducial
cuts, due to limited spectroscopic resources; the PS1 dataset is however quite
robust for these SNe are far from both the conservative and fiducial cuts; see
Fig. \ref{fig:hists}. \bigbreak

The results are shown Fig. \ref{fig:distrib}

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=\linewidth]{Answer_figures/both-cut_SDSS_PS1-010-020.pdf}
    \captionsetup{justification=centering}
    \caption{$x_1$ and $c$ distributions of the SDSS and PS1 samples for $0.10 <
    z < 0.20$. A Kolmogorov-Smirnov test doesn't show any indication that the
samples are not taken from the same distribution.}
\label{fig:distrib}
\end{figure}

\end{document}

### 2020-06-17 Nicolas.tex
2 columns for submit
Started referee answer

@@
-\documentclass[referee]{aa}
+\documentclass[]{aa}
 
@@
 In addition to this ``base'' modeling, we have studied various distributions,
-including redshift independent models; we also studied the prediction from a per-sample asymmetric Gaussian stretch distribution used, for instance, by the Beams with Bias Correction Malmquist bias correction algorithm \citep{scolnic2016, kessler2017}.
+including redshift independent models; we also studied the prediction from a
+per-sample asymmetric Gaussian stretch distribution used, for instance, by the
+Beams with Bias Correction Malmquist bias correction algorithm
+\citep{scolnic2016, kessler2017}.
 

### 2020-06-17 referee_answer.tex
Added means

@@
         \centering
-        \includegraphics[width=.3\linewidth]{Answer_figures/Fig_biasCor_mb.eps}
+        \includegraphics[width=.20\linewidth]{Answer_figures/Fig_biasCor_mb.eps}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=.3\linewidth]{Answer_figures/Fig_biasCor_x1.eps}
+        \includegraphics[width=.20\linewidth]{Answer_figures/Fig_biasCor_x1.eps}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=.3\linewidth]{Answer_figures/Fig_biasCor_c.eps}
+        \includegraphics[width=.20\linewidth]{Answer_figures/Fig_biasCor_c.eps}
     \end{subfigure}
@@
 
-\begin{wrapfigure}[10]{R}{.4\linewidth}
+\begin{wrapfigure}[13]{R}{.4\linewidth}
     \vspace*{-20pt}
@@
 
-The results are shown Fig. \ref{fig:distrib}
+The results are shown Fig. \ref{fig:distrib}. We represented the normed
+histograms along with their error bars, and found that they don't differ much.
+To ascertain this idea, we used a Kolmogorov-Smirnov test that gave us a p-value
+of $0.265$ for the color, and $0.137$ for the stretch; while it doesn't tell us
+that both the SDSS and PS1 parts come from the same distribution, as would be
+expected if we were free from selection effects (so that the samples were random
+draws from what Nature could give us), it is not excluded.
 
@@
     \captionsetup{justification=centering}
-    \caption{$x_1$ and $c$ distributions of the SDSS and PS1 samples for $0.10 <
-    z < 0.20$. A Kolmogorov-Smirnov test doesn't show any indication that the
-samples are not taken from the same distribution.}
+    \caption{$x_1$ and $c$ normed histograms of the SDSS and PS1 samples for
+        $0.10 < z < 0.20$ and their errorbars. On the left, we find in orange
+        lines the Base model distributions for $z = 0.10, 0.15, 0.20$ from top
+        to bottom. The p-values are the results from a Kolmogorov-Smirnov test;
+    they don't show any indication that the samples are not taken from the same
+distribution, for both $c$ and $x_1$.}
 \label{fig:distrib}
@@
 
+Moreover, we tried to see if a particular trend of color evolution was visible,
+reproducting the Fig. 3 from the original N+20 paper. We find that the averages
+are compatible with a constant mean color, see Fig. \ref{fig:means}
+
+\begin{figure}[htbp!]
+    \centering
+    \begin{subfigure}
+        \centering
+        \includegraphics[width=.45\linewidth]{Answer_figures/mean_stretchs-nHST.pdf}
+        %\captionsetup{justification=centering}
+        %\caption{Mean stretch of the complete sample}
+        %\label{fig:mean_x}
+    \end{subfigure}
+    \begin{subfigure}
+        \centering
+        \includegraphics[width=.45\linewidth]{Answer_figures/mean_colors-nHST.pdf}
+        %\captionsetup{justification=centering}
+        %\caption{Mean color of the complete sample}
+        %\label{fig:mean_c}
+    \end{subfigure}
+    \centering
+    \captionsetup{justification=centering}
+
+    \caption{Mean stretch and color of the complete sample in the same 6 bins of
+        equal sample size. The open marker show the means for the complete
+        sample from which we removed HST data. For the color, we might think
+        that there is an evolution for the 5 first bins, but the 6th one breaks
+    it, even without HST data, while for the stretch, with our without HST data
+we see the same evolution}
+
+    \label{fig:means}
+\end{figure}
+
 \end{document}

### 2020-06-18 referee_answer.tex
Complete means, ideas for 1 & 3

@@
     \centering
-    \begin{subfigure}
+    \begin{subfigure}[t]{.20\linewidth}
         \centering
-        \includegraphics[width=.20\linewidth]{Answer_figures/Fig_biasCor_mb.eps}
+        \includegraphics[width=\linewidth]{Answer_figures/Fig_biasCor_mb.eps}
     \end{subfigure}
-    \begin{subfigure}
+    \begin{subfigure}[t]{.20\linewidth}
         \centering
-        \includegraphics[width=.20\linewidth]{Answer_figures/Fig_biasCor_x1.eps}
+        \includegraphics[width=\linewidth]{Answer_figures/Fig_biasCor_x1.eps}
     \end{subfigure}
-    \begin{subfigure}
+    \begin{subfigure}[t]{.20\linewidth}
         \centering
-        \includegraphics[width=.20\linewidth]{Answer_figures/Fig_biasCor_c.eps}
+        \includegraphics[width=\linewidth]{Answer_figures/Fig_biasCor_c.eps}
     \end{subfigure}
@@
 Moreover, we tried to see if a particular trend of color evolution was visible,
-reproducting the Fig. 3 from the original N+20 paper. We find that the averages
-are compatible with a constant mean color, see Fig. \ref{fig:means}
+reproducting the Fig. 3 from the original N+20 paper plotting the color instead
+of the stretch without any modeling (Fig. \ref{fig:mean_c}), and recalled what
+it looked like for the stretch (Fig. \ref{fig:mean_x}). For additional
+information, we represented:
+\begin{itemize}
+    \item The full, uncut dataset in dark grey;
+    \item The fiducial sample in light grey;
+    \item The SNe above the fiducial cut in green (the SNe that we removed from
+        the full sample to obtain the fiducial ones);
+    \item The conservative sample il transparent light grey;
+    \item And the fiducial sample without HST data.
+\end{itemize}
+The last one serves to show that even without HST data, the trend of the color
+to go up in the last bin is not governed by HST only. We find that the averages
+are compatible with a constant mean color, and that the fiducial and
+conservative samples are quite close in terms of mean values while removing the
+low-color trend that is expected from selection effects. Concerning the stretch,
+the fact that the full, fiducial and conservative samples are close are an
+indicator that the selection effect have maginal impact on mean $x_1$ in
+comparison to the drift.
 
@@
     \centering
-    \begin{subfigure}
+    \begin{subfigure}[t]{0.70\linewidth}
         \centering
-        \includegraphics[width=.45\linewidth]{Answer_figures/mean_stretchs-nHST.pdf}
-        %\captionsetup{justification=centering}
-        %\caption{Mean stretch of the complete sample}
-        %\label{fig:mean_x}
+        \includegraphics[width=\linewidth]{Answer_figures/mean_stretchs-nHST_full-sup.pdf}
+        \captionsetup{justification=centering}
+        \caption{Mean stretch of the complete sample}
+        \label{fig:mean_x}
     \end{subfigure}
-    \begin{subfigure}
+    \begin{subfigure}[t]{0.70\linewidth}
         \centering
-        \includegraphics[width=.45\linewidth]{Answer_figures/mean_colors-nHST.pdf}
-        %\captionsetup{justification=centering}
-        %\caption{Mean color of the complete sample}
-        %\label{fig:mean_c}
+        \includegraphics[width=\linewidth]{Answer_figures/mean_colors-nHST_full-sup.pdf}
+        \captionsetup{justification=centering}
+        \caption{Mean color of the complete sample}
+        \label{fig:mean_c}
     \end{subfigure}
@@
     \caption{Mean stretch and color of the complete sample in the same 6 bins of
-        equal sample size. The open marker show the means for the complete
-        sample from which we removed HST data. For the color, we might think
-        that there is an evolution for the 5 first bins, but the 6th one breaks
-    it, even without HST data, while for the stretch, with our without HST data
-we see the same evolution}
+        equal sample size (based on the fiducial sample). The dark grey markers
+        represent the full, uncut sample; the light grey ones the fiducial
+        sample; the SNe above the fiducial cut are in green; in transparent
+        light grey is the conservative sample and the open marker shows the
+        means for the complete sample from which we removed HST data. For the
+        color, we might think that there is an evolution for the 5 first bins,
+        but the 6th one breaks it, even without HST data, while for the stretch,
+    with our without HST data we see the same evolution}
 
@@
 
+\section{Of the $x_1$ drift of simulated surveys}
+Finally, we wanted to bringto the attention of the referee the fact that a full
+SNANA analysis will soon be ongoing and should be the subject of a dedicated
+paper, that could then give more in depth insight concerning the simulations
+thoughts that were shared with us.
 \end{document}

### 2020-06-24 referee_answer.tex
Update

@@
 \title{Referee answer}
+First, thank you for your interest in our paper. We are glad you made those
+comments, both the general and precise ones. With this answer, we hope we will
+be able to further complete the discussion about selection effects and their
+extent in our sample. \bigbreak
+
+The main point that we want to bring to the attention of the referee is that a
+full SNANA analysis in collaboration with Daniel \textsc{Scolnic} with the
+proposed results of \textsc{Nicolas} 2020 has started, and should be a whole
+paper in itself. We are eager to work with such a powerful simulation tool to
+have an even better insight concerning those effects and how they affect
+cosmology. \bigbreak
+
+Concerning the precise points brought up to our attention, here are some
+elements of answers:
 
@@
 \section{Of the $x_1$ bias that doesn't appear on $m$}
-The referee insists on the fact that ``Biases in x1 are expected as a function
-of redshift simply from survey modeling/selection effects'', showing the
-following figure from \textsc{Kessler \& Scolnic 2017} to point
-out that biases in $x_1$ become apparent much sooner that biases on $m$.
 
-\begin{figure}[htbp!]
+\begin{wrapfigure}[11]{R}{.5\linewidth}
     \centering
-    \begin{subfigure}[t]{.20\linewidth}
+    \vspace{-20pt}
+    \begin{subfigure}[t]{.30\linewidth}
         \centering
@@
     \end{subfigure}
-    \begin{subfigure}[t]{.20\linewidth}
+    \begin{subfigure}[t]{.30\linewidth}
         \centering
@@
     \end{subfigure}
-    \captionsetup{justification=centering}
+    \captionsetup{justification=centering, size=scriptsize}
     \caption{Bias corrections $\mkern
@@
     \label{fig:KS17}
-\end{figure}
+\end{wrapfigure}
 
-\begin{wrapfigure}[13]{R}{.4\linewidth}
-    \vspace*{-20pt}
-    \centering
-    \includegraphics[width=\linewidth]{Answer_figures/hist_surveys2_btw_cividis.pdf}
-    \captionsetup{justification=centering}
-    \caption{\small Redshift histograms of SNe~Ia from the SDSS and PS1 datasets
-    respectively}
-    \label{fig:hists}
-\end{wrapfigure}
+Then, the referee insists on the fact that ``Biases in x1 are expected as a
+function of redshift simply from survey modeling/selection effects'', showing
+the following figure from \textsc{Kessler \& Scolnic 2017} to point out that
+biases in $x_1$ become apparent much sooner that biases on $m$.
+
 Yet, from the same figure, biases in $c$ appear at $z=0$. Here we want to
@@
 
+\begin{wrapfigure}[10]{R}{.4\linewidth}
+    \vspace*{-28pt}
+    \centering
+    \includegraphics[width=\linewidth]{Answer_figures/hist_surveys2_btw_cividis.pdf}
+    \captionsetup{justification=centering, size=scriptsize}
+    \caption{Redshift histograms of SNe~Ia from the SDSS and PS1 datasets
+    respectively}
+    \label{fig:hists}
+\end{wrapfigure}
+
 The results are shown Fig. \ref{fig:distrib}. We represented the normed
@@
 expected if we were free from selection effects (so that the samples were random
-draws from what Nature could give us), it is not excluded.
+draws from what Nature could give us), it is not excluded. This particuliar
+figure will appear upon re-submission of the paper for it adresses pertinent
+questions.
+
+\vspace{30pt}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{Answer_figures/both-cut_SDSS_PS1-010-020.pdf}
-    \captionsetup{justification=centering}
+    \includegraphics[width=.9\linewidth]{Answer_figures/both-cut_SDSS_PS1-010-020.pdf}
+    \captionsetup{justification=centering, font=small}
     \caption{$x_1$ and $c$ normed histograms of the SDSS and PS1 samples for
@@
 reproducting the Fig. 3 from the original N+20 paper plotting the color instead
-of the stretch without any modeling (Fig. \ref{fig:mean_c}), and recalled what
-it looked like for the stretch (Fig. \ref{fig:mean_x}). For additional
+of the stretch without any modeling (Fig. \ref{fig:means}, right), and recalled what
+it looked like for the stretch (Fig. \ref{fig:means}, left). For additional
 information, we represented:
@@
     \centering
-    \begin{subfigure}[t]{0.70\linewidth}
+    \begin{subfigure}[t]{0.49\linewidth}
         \centering
         \includegraphics[width=\linewidth]{Answer_figures/mean_stretchs-nHST_full-sup.pdf}
-        \captionsetup{justification=centering}
-        \caption{Mean stretch of the complete sample}
-        \label{fig:mean_x}
     \end{subfigure}
-    \begin{subfigure}[t]{0.70\linewidth}
+    \begin{subfigure}[t]{0.49\linewidth}
         \centering
         \includegraphics[width=\linewidth]{Answer_figures/mean_colors-nHST_full-sup.pdf}
-        \captionsetup{justification=centering}
-        \caption{Mean color of the complete sample}
-        \label{fig:mean_c}
     \end{subfigure}
     \centering
-    \captionsetup{justification=centering}
-
+    \captionsetup{justification=centering, size=small}
     \caption{Mean stretch and color of the complete sample in the same 6 bins of
@@
     with our without HST data we see the same evolution}
-
     \label{fig:means}
@@
 thoughts that were shared with us.
+
 \end{document}

### 2020-06-26 Nicolas.tex
Included color test in main article

@@
 \section{Complete Sample Construction}\label{sec:sample}
+\subsection{Applying redshift cuts}\label{ssec:cuts}
 
@@
     \label{fig:sample}
+\end{figure}
+
+\subsection{Verifying completeness}\label{ssec:verify}
+
+In order to ensure that the effects we model are only due to the supposed drift,
+we need to check for possible Malmquist bias left in the sample. Because color
+is most sensitive to selection effects, we studied both the stretch and color of
+the end of SDSS and the start of PS1, for $0.10 < z < 0.20$. In this redshift
+range, because the SDSS SNe between $0.15 < z < 0.20$ are between our
+conservative and fiducial cuts, these SNe are the most prone to have selection
+effects. However, the PS1 dataset is quite robust because these SNe are far from
+both the conservative and fiducial cuts; see Fig. \ref{fig:cuts}.
+
+The results are shown Fig. \ref{fig:distrib}. We represented the normed
+histograms along with their error bars, and found that they don't differ much.
+To ascertain this idea, we used a Kolmogorov-Smirnov test that gave us a p-value
+of $0.265$ for the color, and $0.137$ for the stretch; while it doesn't tell us
+that both the SDSS and PS1 parts come from the same distribution, as would be
+expected if we were free from selection effects (so that the samples were random
+draws from what Nature could give us), it is not excluded. This particuliar
+figure will appear upon re-submission of the paper for it adresses pertinent
+questions.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=17cm]{Article_figures/both-cut_SDSS_PS1-010-020.pdf}
+    \caption{$x_1$ and $c$ normed histograms of the SDSS and PS1 samples for
+        $0.10 < z < 0.20$ and their errorbars. On the left, we find in orange
+        lines the Base model distributions for $z = 0.10, 0.15, 0.20$ from top
+        to bottom. The p-values are the results from a Kolmogorov-Smirnov test;
+    they don't show any indication that the samples are not taken from the same
+distribution, for both $c$ and $x_1$.}
+    \label{fig:distrib}
+\end{figure*}
+
+Moreover, we tried to see if a particular trend of color evolution was visible,
+reproducting Fig. \ref{fig:sample} plotting the color instead of the stretch
+without any modeling (Fig. \ref{fig:means}, top), and recalled what it looked
+like for the stretch (Fig. \ref{fig:means}, bottom). For additional information,
+we represented:
+
+\begin{itemize}
+    \item The full, uncut dataset in dark grey;
+    \item The fiducial sample in light grey;
+    \item The SNe above the fiducial cut in green (the SNe that we removed from
+        the full sample to obtain the fiducial ones);
+    \item The conservative sample il transparent light grey;
+    \item And the fiducial sample without HST data.
+\end{itemize}
+
+The last one serves to show that even without HST data, the trend of the color
+to go up in the last bin is not governed by HST only. We find that the averages
+are compatible with a constant mean color, and that the fiducial and
+conservative samples are quite close in terms of mean values while removing the
+low-color trend that is expected from selection effects. Concerning the stretch,
+the fact that the full, fiducial and conservative samples are close are an
+indicator that the selection effect have maginal impact on mean $x_1$ in
+comparison to the drift.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{Article_figures/mean_both-nHST_full-sup.pdf}
+    \caption{Mean stretch and color of the complete sample in the same 6 bins of
+        equal sample size (based on the fiducial sample). The dark grey markers
+        represent the full, uncut sample; the light grey ones the fiducial
+        sample; the SNe above the fiducial cut are in green; in transparent
+        light grey is the conservative sample and the open marker shows the
+        means for the complete sample from which we removed HST data. For the
+        color, we might think that there is an evolution for the 5 first bins,
+    but the 6th one breaks it, even without HST data, while for the stretch,
+with our without HST data we see the same evolution}
+    \label{fig:means}
 \end{figure}

### 2020-06-26 referee_answer.tex
Minor tweaks in referee answer

@@
 reproducting the Fig. 3 from the original N+20 paper plotting the color instead
-of the stretch without any modeling (Fig. \ref{fig:means}, right), and recalled what
-it looked like for the stretch (Fig. \ref{fig:means}, left). For additional
+of the stretch without any modeling (Fig. \ref{fig:means}, right), and recalled
+what it looked like for the stretch (Fig. \ref{fig:means}, left). For additional
 information, we represented:
@@
 \section{Of the $x_1$ drift of simulated surveys}
-Finally, we wanted to bringto the attention of the referee the fact that a full
-SNANA analysis will soon be ongoing and should be the subject of a dedicated
-paper, that could then give more in depth insight concerning the simulations
+A full SNANA analysis will soon be ongoing and should be the subject of a
+dedicated paper, giving more indepth insight concerning the simulations
 thoughts that were shared with us.

### 2020-06-30 referee_answer.tex
Update

@@
 \section{Of the $x_1$ drift of simulated surveys}
-A full SNANA analysis will soon be ongoing and should be the subject of a
-dedicated paper, giving more indepth insight concerning the simulations
-thoughts that were shared with us.
+We recall here that a full SNANA analysis will soon be ongoing and should be the
+subject of a dedicated paper, giving more indepth insight concerning the
+simulations thoughts that were shared with us. \bigbreak
+
+The referee pointed out that simulated samples without drift parameters do
+recover a stretch evolution, saying that the $x_1$ drift could be entirely
+explained by selection effects. We agree that both our supposed drift and
+selection effects go the same way, so it seems natural to see this trend in
+other studies. However, concerning the Fig. 6 of \textsc{Mosher et al.} 2014
+which uses SDSS data we see that this trend actually goes up, according to
+selection effects, only after $z > 0.2$, which corresponds to our redshift cut
+for that survey. We feel that the plots the referee showed us are mostly
+examples that simulated samples include both drift and selection effects, the
+two of them stacking at the end of the redshift range. \bigbreak
+
+Concerning the inclusion of the Per sample Asym model without using simulations
+beforehand, we want to point out that the idea is actually to compare models
+prescribing a drift parameter with those that don't, using pertinent models
+(Howell, Per sample Asym for instance). The Per sample Asym model is used in
+current cosmological studies and was said by their authors to be able to account
+for unmodeled redshift drifts thanks to the limited span of the surveys: we
+also wanted to test this affirmation. It would seem that this model is relevant
+for this study. \bigbreak
+
+\section{General comments}
+The parameters comparison in the referee's comments are indeed not agreeing with
+each other. There may be two reasons for that:
+\begin{enumerate}
+
+    \item Even if we work on the same surveys, the datasets we use are really
+        different from \textsc{Scolnic \& Kessler} 2016 because they worked on
+        their corrected full sample while we used a subset of that through the
+        redshift cuts; differences in the computed parameters are expected in
+        that matter.
+
+    \item Moreover, the parameters we say are in really good agreement are those
+        of PS1 from \textsc{Scolnic \& Kessler} \underline{2018}, in which they
+        presented updated results that are in much better agreement with ours;
+        we suspect that the same work done for SNLS and SDSS would go the same
+        way.
+
+\end{enumerate}
 

### 2020-06-30 Nicolas.tex
Update

--- 2020-06-17 Nicolas.tex
@@
 
+\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
 \newcommand{\prob}[2]{\mathcal{P}\left( #1 \mid #2\right)}
@@
     \and Saul Perlmutter\inst{3}
+    \and M. Smith\inst{4}
+    \orcid{0000-0002-3321-1432}
 }
@@
     12489 Berlin
+    \and
+    University of Southampton: Southampton, GB
 }
@@
 \section{Complete Sample Construction}\label{sec:sample}
+\subsection{\nn{Applying redshift cuts}}\label{ssec:cuts}
 
@@
 \end{figure}
+
+\subsection{\nn{Verifying completeness}}\label{ssec:verify}
+
+\nn{In order to ensure that the effects we model are only due to the supposed drift,
+we need to check for possible Malmquist bias left in the sample. Because color
+is most sensitive to selection effects, we studied both the stretch and color of
+the end of SDSS and the start of PS1, for $0.10 < z < 0.20$. In this redshift
+range, because the SDSS SNe between $0.15 < z < 0.20$ are between our
+conservative and fiducial cuts, these SNe are the most prone to have selection
+effects. However, the PS1 dataset is quite robust because these SNe are far from
+both the conservative and fiducial cuts; see Fig. \ref{fig:cuts}.}
+
+\nn{The results are shown Fig. \ref{fig:distrib}. We represented the normed
+histograms along with their error bars, and found that they don't differ much.
+To ascertain this idea, we used a Kolmogorov-Smirnov test that gave us a p-value
+of $0.265$ for the color, and $0.137$ for the stretch; while it doesn't tell us
+that both the SDSS and PS1 parts come from the same distribution, as would be
+expected if we were free from selection effects (so that the samples were random
+draws from what Nature could give us), it is not excluded. }
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=17cm]{Article_figures/both-cut_SDSS_PS1-010-020.pdf}
+    \caption{\nn{$x_1$ and $c$ normed histograms of the SDSS and PS1 samples for
+        $0.10 < z < 0.20$ and their errorbars. On the left, we find in orange
+        lines the Base model distributions for $z = 0.10, 0.15, 0.20$ from top
+        to bottom. The p-values are the results from a Kolmogorov-Smirnov test;
+    they don't show any indication that the samples are not taken from the same
+distribution, for both $c$ and $x_1$.}}
+    \label{fig:distrib}
+\end{figure*}
 

### 2020-08-18 Nicolas.tex
New image

@@
 \newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
+\newcommand{\mr}[1]{\textcolor[rgb]{0.5, 0.20, 0.3}{#1}}
 \newcommand{\prob}[2]{\mathcal{P}\left( #1 \mid #2\right)}
@@
 
-\subsection{\nn{Verifying completeness}}\label{ssec:verify}
-
-\nn{In order to ensure that the effects we model are only due to the supposed drift,
-we need to check for possible Malmquist bias left in the sample. Because color
-is most sensitive to selection effects, we studied both the stretch and color of
-the end of SDSS and the start of PS1, for $0.10 < z < 0.20$. In this redshift
-range, because the SDSS SNe between $0.15 < z < 0.20$ are between our
-conservative and fiducial cuts, these SNe are the most prone to have selection
-effects. However, the PS1 dataset is quite robust because these SNe are far from
-both the conservative and fiducial cuts; see Fig. \ref{fig:cuts}.}
-
-\nn{The results are shown Fig. \ref{fig:distrib}. We represented the normed
-histograms along with their error bars, and found that they don't differ much.
-To ascertain this idea, we used a Kolmogorov-Smirnov test that gave us a p-value
-of $0.265$ for the color, and $0.137$ for the stretch; while it doesn't tell us
-that both the SDSS and PS1 parts come from the same distribution, as would be
-expected if we were free from selection effects (so that the samples were random
-draws from what Nature could give us), it is not excluded. }
-
-\begin{figure*}
+\subsection{\mr{Testing the construction of a volume limited sample}}\label{ssec:verify}
+
+\mr{In this analysis, we have built a volume limited sample from a set of
+magnitude limited ones, using a simple redshift limit.  A ongoing follow up
+analysis will make use of the SNANA package \citep{SNANA} to account for
+more subtle selection functions. However, the strength of the current
+approach is that it is independent of potential inaccuracy in the modeling
+of the complex selection functions of the aforementioned dataset. Yet, it is
+necessary to test if the sample we have created could show sign of leftover
+sample selections.}
+
+\mr{To do so, we compare the stretch and color distributions of the SNe~Ia in
+our sample between datasets of overlapping redshifts. If indeed we have built
+a volume limited sample, the distributions of SNe~Ia parameters at a given
+redshift should not depend of their origin surveys. When considering a
+redshift range, this has to be narrow enough such that any drift would be
+negligible. The two samples that overlap the most in redshift are PS1 and
+SDSS in the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}).
+This corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and
+thus potentially the most affected by potential selection effect leftover
+(see the corresponding discussion in section~\ref{sec:sample}). In that same
+redshift range, PS1 has 52 SNe~Ia, which are at in the lowest redshift bin
+and thus unlikely to have any selection function issues. We present the
+stretch and color distributions of both samples in the redshift range $0.10 <
+z < 0.20$ in Fig.~\ref{fig:distrib}, where we also provide the p-values of
+Kolmogorov-Smirnov tests. The high p-values both in color and stretch
+distributions ($>10\%$) strengthen the visual impression that there is no
+reason to believe that the SDSS and PS1 SNe~Ia in their overlapping redshift
+range are not drawn from the same underlying distributions. A similar
+analysis is doable for PS1 and SNLS in their redshift range $0.20 < z <
+0.30$ where the same conclusion could be drawn, i.e., that there is no sign
+of discrepancy between in the stretch and color distributions between the
+low- and high-end of the SNLS and PS1 datasets, respectively. Nonetheless,
+the small size of the SNLS dataset at $z < 0.30$ (12 SNe~Ia vs. 90 for PS1)
+limits the interest of the test as only strong deviation would be
+noticeable. Extending the redshift range to $0.20 < z < 0.40$ (though we have no
+PS1 data above 0.3) enables to get 31 SNe~Ia from SNLS in total, the stretch
+and color p-values remains high, at 34\% and 4\% respectively.}
+
+\begin{figure}
     \centering
-    \includegraphics[width=17cm]{Article_figures/both-cut_SDSS_PS1-010-020.pdf}
+    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_PS1-010-020.pdf}
     \caption{\nn{$x_1$ and $c$ normed histograms of the SDSS and PS1 samples for
@@
     \label{fig:distrib}
-\end{figure*}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SNLS_PS1-020-030.pdf}
+    \caption{\mr{Same as Fig.~\ref{fig:distrib} for the PS1 and SNLS data in
+    their overlapping redshift range $0.20 < z < 0.30$.}}
+
+    \label{fig:distrib-ps-snls}
+\end{figure}
+
+\mr{We highlight that the SNe~Ia color is more prone to selection effects than
+stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of
+\cite{kessler2017}. Hence, since the color shows no sign of leftover selection
+effects, it further supports our claim that our simple redshift-based selection
+criteria is sufficient to test the redshift drift of the SN~Ia stretch
+distribution.}
+
+\nn{In order to ensure that the effects we model are only due to the supposed
+drift, we need to check for possible Malmquist bias left in the sample.
+Because color is most sensitive to selection effects, we studied both the
+stretch and color of the end of SDSS and the start of PS1, for $0.10 < z <
+0.20$. In this redshift range, }
+
+\nn{The results are shown Fig. \ref{fig:distrib}. We represented the normed
+histograms along with their error bars, and found that they don't differ much.
+To ascertain this idea, we used a Kolmogorov-Smirnov test that gave us a p-value
+of $0.265$ for the color, and $0.137$ for the stretch; while it doesn't tell us
+that both the SDSS and PS1 parts come from the same distribution, as would be
+expected if we were free from selection effects (so that the samples were random
+draws from what Nature could give us), it is not excluded.}
 
@@
 
+\bibitem[Kessler et al.(2009)]{SNANA} Kessler, R., Bernstein, J.~P., Cinabro,
+D., et al.\ 2009, \pasp, 121, 1028
+
 \bibitem[Kessler \& Scolnic(2017)]{kessler2017} Kessler, R., \& Scolnic, D.\

### 2020-09-17 Nicolas.tex
Including MR modifications. Referee answer on GD. YC's comments to be in

@@
     \and M. Briday\inst{1}
-    \and J. Nordin\inst{4}
     \and Y.-L. Kim\inst{1}
     \orcid{0000-0002-1031-0796}
+    \and J. Nordin\inst{4}
     \and Saul Perlmutter\inst{3}
@@
     tracers. We model the underlying stretch distribution as a function of
-    redshift, using the evolution of fraction of young and old SNe~Ia as
+    redshift, using the evolution of the fraction of young and old SNe~Ia as
     predicted by \cite{rigault2018}, and assuming constant underlying stretch
@@
 \citet{campbell2013} for similar contours, the second using a more conservative
-$|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and $c=0$
-is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
+$|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is
+$M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
 standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
@@
 lightcurve, the effective limiting standardized absolute magnitude is
-approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit $m_{\lim}$
-of a magnitude limited survey, one can derive the maximum redshift $z_{\lim}$
-above which the faintest SNe~Ia will be missed using the relation between
-apparent magnitude, redshift and absolute magnitude $\mu(z_{\lim}) = m_{\lim} -
-M_{\lim}$.
+approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit
+$m_{\lim}$ of a magnitude limited survey, one can derive the maximum redshift
+$z_{\lim}$ above which the faintest SNe~Ia will be missed using the relation
+between apparent magnitude, redshift and absolute magnitude $\mu(z_{\lim}) =
+m_{\lim} - M_{\lim}$.
 
@@
 In a similar redshift range, SDSS has a limiting magnitude of 22.5
-\citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However, the
-SDSS surveys were more sensitive to limited spectroscopic resources; \cite[][see
-their Section~2]{kessler2009} pointed out that during the first year of SDSS,
-SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
-to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
-spectroscopic resources were available, and \cite{kessler2009} and
+\citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However,
+the SDSS surveys were more sensitive to limited spectroscopic resources;
+\cite[][see their Section~2]{kessler2009} pointed out that during the first year
+of SDSS, SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up,
+corresponding to a redshift cut at $0.15$. For the rest of the SDSS survey,
+additional spectroscopic resources were available, and \cite{kessler2009} and
 \cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
@@
 range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
-much deeper. These 114 SNfactory SNe~Ia are thus in the volume limited part
-of the survey (Aldering et al., in prep.), and are therefore assumed to be a
-random sampling of the underlying SN population. The SNfactory sample is
-particularly useful for studying SN property drift, as it enables us to have a
-large SN~Ia sample at $z<0.1$.
+much deeper. These 114 SNfactory SNe~Ia are thus in the volume limited part of
+the survey (Aldering et al., in prep.), and are therefore assumed to be a random
+sampling of the underlying SN population. The SNfactory sample is particularly
+useful for studying SN property drift, as it enables us to have a large SN~Ia
+sample at $z<0.1$.
 
@@
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. The evolution of
-the mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are
-split in redshift bins of regular sample size. We see that SNe~Ia at higher
-redshift have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than
-those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the
-underlying stretch distribution is drifting.
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. The evolution of the
+mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are split
+in redshift bins of regular sample size. We see that SNe~Ia at higher redshift
+have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at
+lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
+stretch distribution is drifting.
 
@@
 
-\subsection{\mr{Testing the construction of a volume limited sample}}\label{ssec:verify}
+\subsection{\mr{Testing the construction of a volume limited sample}}
+\label{ssec:verify}
 
 \mr{In this analysis, we have built a volume limited sample from a set of
-magnitude limited ones, using a simple redshift limit.  A ongoing follow up
+    magnitude limited ones, using a simple redshift limit. An ongoing follow up
     analysis will make use of the SNANA package \citep{SNANA} to account for
@@
 \mr{To do so, we compare the stretch and color distributions of the SNe~Ia in
-our sample between datasets of overlapping redshifts. If indeed we have built
-a volume limited sample, the distributions of SNe~Ia parameters at a given
-redshift should not depend of their origin surveys. When considering a
-redshift range, this has to be narrow enough such that any drift would be
+    our sample between datasets of overlapping redshifts. If indeed we have
+    built a volume limited sample, the distributions of any SNe~Ia parameter at
+    a given redshift should not depend of their origin surveys. When considering
+    a redshift range, this has to be narrow enough such that any drift would be
     negligible. The two samples that overlap the most in redshift are PS1 and
@@
     redshift range, PS1 has 52 SNe~Ia, which are at in the lowest redshift bin
-and thus unlikely to have any selection function issues. We present the
-stretch and color distributions of both samples in the redshift range $0.10 <
-z < 0.20$ in Fig.~\ref{fig:distrib}, where we also provide the p-values of
-Kolmogorov-Smirnov tests. The high p-values both in color and stretch
-distributions ($>10\%$) strengthen the visual impression that there is no
-reason to believe that the SDSS and PS1 SNe~Ia in their overlapping redshift
-range are not drawn from the same underlying distributions. A similar
-analysis is doable for PS1 and SNLS in their redshift range $0.20 < z <
-0.30$ where the same conclusion could be drawn, i.e., that there is no sign
-of discrepancy between in the stretch and color distributions between the
-low- and high-end of the SNLS and PS1 datasets, respectively. Nonetheless,
-the small size of the SNLS dataset at $z < 0.30$ (12 SNe~Ia vs. 90 for PS1)
-limits the interest of the test as only strong deviation would be
-noticeable. Extending the redshift range to $0.20 < z < 0.40$ (though we have no
-PS1 data above 0.3) enables to get 31 SNe~Ia from SNLS in total, the stretch
-and color p-values remains high, at 34\% and 4\% respectively.}
+    and thus unlikely to have any selection issue. The Fig.~\ref{fig:distrib}
+    present the stretch and color distribution of both, together with p-values
+    of Kolmogorov-Smirnov tests that test if the $x_1$ and $c$ of the SDSS and
+    PS1 samples at $0.10<z<0.20$ arise from the same underlying distribution.
+    Those p-values are high ($>10\%$), and if one cannot conclude from this test
+    that they SNe~Ia are from the same underlying distribution, one has no
+    ground to claim they are not. This is in agreement with the visual
+    impression from Fig.  ~\ref{fig:distrib}.  A similar analysis is doable for
+    PS1 and SNLS in their redshift range $0.20 < z < 0.30$ where the same
+    conclusion could be drawn, i.e., that there is no sign of discrepancy
+    between in the stretch and color distributions between the low- and high-end
+    of the SNLS and PS1 datasets, respectively. Nonetheless, the small size of
+    the SNLS dataset at $z < 0.30$ (12 SNe~Ia vs. 90 for PS1) limits the
+    interest of the test as only strong deviation would be noticeable. Extending
+the redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
+enables to get 31 SNe~Ia from SNLS in total, the stretch and color p-values
+remains high, at 34\% and 4\% respectively.}
 
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_PS1-010-020.pdf}
-    \caption{\nn{$x_1$ and $c$ normed histograms of the SDSS and PS1 samples for
-        $0.10 < z < 0.20$ and their errorbars. On the left, we find in orange
-        lines the Base model distributions for $z = 0.10, 0.15, 0.20$ from top
-        to bottom. The p-values are the results from a Kolmogorov-Smirnov test;
-    they don't show any indication that the samples are not taken from the same
-distribution, for both $c$ and $x_1$.}}
+    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
+    \caption{\mr{$x_1$ and $c$ distribution histograms of different surveys
+    overlapping in redshift. \emph{Facing up} SDSS and PS1 within the redshift
+    range $0.10 < z < 0.20$ ; \emph{facing down} PS1 and SNLS within the
+    redshift range $0.20<z<0.3$. Error bars show the Poisson noise. Our stretch
+    "Base" modeling is illustrated in orange at the mean redshift of the
+    redshift ranges, $0.15$ and $0.25$, respectively. P-values of the
+    Kolmogorov-Smirnov tests are indicated on the top (bottom) of each panel
+    showing no sign that the SDSS and PS1 (PS1 and SNLS) $x_1$ and $c$
+    distributions are not drawn from the same underlying distributions.}}
     \label{fig:distrib}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SNLS_PS1-020-030.pdf}
-    \caption{\mr{Same as Fig.~\ref{fig:distrib} for the PS1 and SNLS data in
-    their overlapping redshift range $0.20 < z < 0.30$.}}
-
-    \label{fig:distrib-ps-snls}
-\end{figure}
-
 \mr{We highlight that the SNe~Ia color is more prone to selection effects than
     stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of
-\cite{kessler2017}. Hence, since the color shows no sign of leftover selection
-effects, it further supports our claim that our simple redshift-based selection
-criteria is sufficient to test the redshift drift of the SN~Ia stretch
-distribution.}
-
-\nn{In order to ensure that the effects we model are only due to the supposed
-drift, we need to check for possible Malmquist bias left in the sample.
-Because color is most sensitive to selection effects, we studied both the
-stretch and color of the end of SDSS and the start of PS1, for $0.10 < z <
-0.20$. In this redshift range, }
-
-\nn{The results are shown Fig. \ref{fig:distrib}. We represented the normed
-histograms along with their error bars, and found that they don't differ much.
-To ascertain this idea, we used a Kolmogorov-Smirnov test that gave us a p-value
-of $0.265$ for the color, and $0.137$ for the stretch; while it doesn't tell us
-that both the SDSS and PS1 parts come from the same distribution, as would be
-expected if we were free from selection effects (so that the samples were random
-draws from what Nature could give us), it is not excluded.}
+    \cite{kessler2017}. Hence, since the color shows no sign of leftover
+    selection effects, it further supports our claim that our simple
+    redshift-based selection criteria is sufficient to test the redshift drift
+of the SN~Ia stretch distribution.}
 
@@
 In short, it was assumed that the number of ``young'' SNe~Ia follows the star
-formation rate (SFR) in the Universe, while the number of ``old'' SNe~Ia
-follows the number of Gyr-old stars in the Universe, i.e. the stellar mass
-(M$^*$). Hence, if we denote $\delta(z)$ (resp. $\psi(z) = 1-\delta(z)$) the
-fraction of young (resp. old) SNe~Ia in the Universe as a function of redshift,
-then the ratio $\delta/\psi$ is expected to follow the evolution of the specific
-star formation rate (SFR/M$^*$), which goes as $(1+z)^{2.8}$ until $z\sim2$
+formation rate (SFR) in the Universe, while the number of ``old'' SNe~Ia follows
+the number of Gyr-old stars in the Universe, i.e. the stellar mass (M$^*$).
+Hence, if we denote $\delta(z)$ (resp. $\psi(z) = 1-\delta(z)$) the fraction of
+young (resp. old) SNe~Ia in the Universe as a function of redshift, then the
+ratio $\delta/\psi$ is expected to follow the evolution of the specific star
+formation rate (SFR/M$^*$), which goes as $(1+z)^{2.8}$ until $z\sim2$
 \citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
-\citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
-\citep{mannucci2006,rodney2014}, \cite{rigault2018} concluded that
+\citep{rigault2013, rigault2018, wiseman2020}, in agreement with rate
+expectations \citep{mannucci2006, rodney2014}, \cite{rigault2018} concluded that
+
 \begin{equation}
@@
 
-We also consider another intrinsically non-drifting model, the one developed for
-Beams with Bias Correction \cite[BBC,][]{scolnic2016, kessler2017}, used in
-recent SN cosmological analyses \cite[e.g.][]{scolnic2018a, descosmopaper2019,
-riess2016, riess2019} to account for Malmquist biases. The BBC formalism assumes
-sample-based (hence intrinsically non-drifting) asymmetric Gaussian stretch
-distributions: $\mathcal{N}\left(\mu,
-\sigma_-{}^2\;\text{if}\;x_1<\mu,\;\text{else}\;\sigma_+{}^2\right)$. The idea
-behind this sample-based approach is twofold: (1) Malmquist biases are driven by
-survey properties and (2) since current surveys cover limited redshift ranges,
-having a sample-based approach covers some potential redshift evolution
-information \citep{scolnic2016, scolnic2018a}. See further discussion concerning
-BBC in Section~\ref{sec:discussion}.
+We also consider another intrinsically non-drifting model, \mr{the functional
+    form} developed for Beams with Bias Correction \cite[BBC,][]{scolnic2016,
+    kessler2017}, used in recent SN cosmological analyses
+    \cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to
+    account for Malmquist biases. The BBC formalism assumes sample-based (hence
+    intrinsically non-drifting) asymmetric Gaussian stretch distributions:
+    $\mathcal{N}\left(\mu,
+    \sigma_-{}^2\;\text{if}\;x_1<\mu,\;\text{else}\;\sigma_+{}^2\right)$. The
+    idea behind this sample-based approach is twofold: (1) Malmquist biases are
+    driven by survey properties and (2) since current surveys cover limited
+    redshift ranges, having a sample-based approach covers some potential
+    redshift evolution information \citep{scolnic2016, scolnic2018a}. See
+    further discussion concerning BBC in Section~\ref{sec:discussion}.
 
@@
     DE-AC025CH11231.
+    This project is partly financially supported by Region Rhone-Alpes-Auvergne
 \end{acknowledgements}
@@
 \end{document}
+end{document}

### 2020-10-14 Nicolas.tex
Integrated YC comments

@@
 \usepackage{txfonts}
+\usepackage{ulem}
 \usepackage{amsmath, mathtools}
@@
     distribution for each age population made of Gaussian mixtures. We test our
-    prediction against published samples chosen to have negligible magnitude
-    selection effects, so that any observed change is indeed of astrophysical
-    and not observational origin. We clearly demonstrate that the underlying
-    SN~Ia stretch distribution is evolving as a function of redshift, and that
-    the young/old drifting model is a much better description of the data than
-    any time-constant model, including the sample-based asymmetric distributions
-    usually used to correct Malmquist bias. The favored underlying stretch model
-    is the bimodal one derived from \cite{rigault2018}: a high-stretch mode
-    shared by both young and old environments, and a low-stretch mode exclusive
-    to old environments. The precise impact of the redshift evolution of the SN
-    Ia population intrinsic properties on cosmology remains to be studied. Yet,
-    the astrophysical drift of the SN stretch distribution does affect current
-    Malmquist bias corrections and thereby distances derived from SN affected by
-    selection effects. We highlight that such a bias will increase with surveys
-    covering increasingly larger redshift ranges, which is particularly
-    important for LSST.}
+    prediction against published samples \nn{\sout{chosen} cut} to have
+    negligible magnitude selection effects, so that any observed change is
+    indeed of astrophysical and not observational origin. \nn{An ongoing
+    analysis with the SNANA package will test this assumption in order to refine
+    the study.} We \nn{\sout{clearly}} demonstrate that the underlying SN~Ia stretch
+    distribution is evolving as a function of redshift, and that the young/old
+    drifting model is a much better description of the data than any time-constant
+    model, including the sample-based asymmetric distributions usually used to
+    correct Malmquist bias. The favored underlying stretch model is the bimodal one
+    derived from \cite{rigault2018}: a high-stretch mode shared by both young and
+    old environments, and a low-stretch mode exclusive to old environments. The
+    precise impact of the redshift evolution of the SN Ia population intrinsic
+    properties on cosmology remains to be studied. Yet, the astrophysical drift of
+    the SN stretch distribution does affect current Malmquist bias corrections and
+    thereby distances derived from SN affected by selection effects. We highlight
+    that such a bias will increase with surveys covering increasingly larger
+    redshift ranges, which is particularly important for LSST.}
 
@@
 \section{Complete Sample Construction}\label{sec:sample}
-\subsection{\nn{Applying redshift cuts}}\label{ssec:cuts}
+\subsection{Applying redshift cuts}\label{ssec:cuts}
 
@@
 24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
-final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
-\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
-\cite[][see their Section~5]{perrett2010} suggests a lower limit of
-$z_{\lim}=0.55$; both limits will be considered, as discussed below.
+final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement
+with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. \nn{However,
+because this simple approach might easily not be enough to alleviate all
+selection effects, we also take into account another, more restrictive,
+redshift limit:} Fig.~14 of \cite[][see their Section~5]{perrett2010} suggests a
+lower limit of $z_{\lim}=0.55$; both limits will be considered, as discussed
+below.
 
 Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
-$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.30$ in
+$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.31$ in
 agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure is also
-suggestive of a more conservative $z_{\lim}=0.27$.
+suggestive of a more conservative $z_{\lim}=0.27$, \nn{similarly considered in
+the following analysis as a first step to evaluate the robustness of our
+sample construction.}
 
@@
 \cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
-these analyses, we will use $z_{\lim}=0.2$ as the baseline SDSS redshift limit.
+these analyses, we will use $z_{\lim}=0.20$ as the baseline SDSS redshift limit,
+\nn{and the first 0.15 cut to compare the results of our modeling on a more
+restrictive sample.}
 
@@
 expected, the selected redshift limits roughly correspond to the peak of these
-histograms. We show in Section~\ref{sec:results} that this sample selection
-indeed provides a subset of SNe~Ia with insignificant selection effects when
+histograms. \nn{We test in Section~\ref{ssec:verify} whether we can consider
+this sample to be free of selection effects. The results exposed} in
+Section~\ref{sec:results} further supports that this sample selection indeed
+provides a subset of SNe~Ia with insignificant selection effects when
 compared to state-of-the-art Malmquist correction techniques.
@@
         SDSS & 0.20 (0.15) & 167 (82)\\
-        PS1   & 0.30 (0.27) & 160 (122)\\
+        PS1 & 0.31 (0.27) & 160 (122)\\
         SNLS & 0.60 (0.55) & 102 (78)\\
@@
 range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
-much deeper. These 114 SNfactory SNe~Ia are thus in the volume limited part of
-the survey (Aldering et al., in prep.), and are therefore assumed to be a random
-sampling of the underlying SN population. The SNfactory sample is particularly
-useful for studying SN property drift, as it enables us to have a large SN~Ia
-sample at $z<0.1$.
+much deeper. These 114 SNfactory SNe~Ia \nn{(80\% of the 141 SNe presented in
+the previously-cited paper)} are thus in the volume-limited part of the
+survey (Aldering et al., in prep.), and are therefore assumed to be a random
+sampling of the underlying SN population. The SNfactory sample is
+particularly useful for studying SN property drift, as it enables us to have
+a large SN~Ia sample at $z<0.1$.
 
@@
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. The evolution of the
-mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are split
-in redshift bins of regular sample size. We see that SNe~Ia at higher redshift
-have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at
-lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
-stretch distribution is drifting.
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}, \nn{where the
+disappearance of the low-stretch population (typically $x_1 < -1$) is what we
+test in this paper}. The evolution of the mean stretch is also shown in
+Fig.~\ref{fig:modelall}, where the data are split in redshift bins of regular
+sample size. We see that SNe~Ia at higher redshift have on average larger
+stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift ($-0.17\pm
+0.10$ at $z\sim0.05$), suggesting that the underlying stretch distribution is
+drifting.
 
@@
 
-\subsection{\mr{Testing the construction of a volume limited sample}}
-\label{ssec:verify}
-
-\mr{In this analysis, we have built a volume limited sample from a set of
-    magnitude limited ones, using a simple redshift limit. An ongoing follow up
-    analysis will make use of the SNANA package \citep{SNANA} to account for
-    more subtle selection functions. However, the strength of the current
-    approach is that it is independent of potential inaccuracy in the modeling
-    of the complex selection functions of the aforementioned dataset. Yet, it is
-    necessary to test if the sample we have created could show sign of leftover
-sample selections.}
-
-\mr{To do so, we compare the stretch and color distributions of the SNe~Ia in
-    our sample between datasets of overlapping redshifts. If indeed we have
-    built a volume limited sample, the distributions of any SNe~Ia parameter at
-    a given redshift should not depend of their origin surveys. When considering
-    a redshift range, this has to be narrow enough such that any drift would be
-    negligible. The two samples that overlap the most in redshift are PS1 and
-    SDSS in the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}).
-    This corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and
-    thus potentially the most affected by potential selection effect leftover
-    (see the corresponding discussion in section~\ref{sec:sample}). In that same
-    redshift range, PS1 has 52 SNe~Ia, which are at in the lowest redshift bin
-    and thus unlikely to have any selection issue. The Fig.~\ref{fig:distrib}
-    present the stretch and color distribution of both, together with p-values
-    of Kolmogorov-Smirnov tests that test if the $x_1$ and $c$ of the SDSS and
-    PS1 samples at $0.10<z<0.20$ arise from the same underlying distribution.
-    Those p-values are high ($>10\%$), and if one cannot conclude from this test
-    that they SNe~Ia are from the same underlying distribution, one has no
-    ground to claim they are not. This is in agreement with the visual
-    impression from Fig.  ~\ref{fig:distrib}.  A similar analysis is doable for
-    PS1 and SNLS in their redshift range $0.20 < z < 0.30$ where the same
-    conclusion could be drawn, i.e., that there is no sign of discrepancy
-    between in the stretch and color distributions between the low- and high-end
-    of the SNLS and PS1 datasets, respectively. Nonetheless, the small size of
-    the SNLS dataset at $z < 0.30$ (12 SNe~Ia vs. 90 for PS1) limits the
-    interest of the test as only strong deviation would be noticeable. Extending
-the redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
-enables to get 31 SNe~Ia from SNLS in total, the stretch and color p-values
-remains high, at 34\% and 4\% respectively.}
+\subsection{Testing the construction of a volume-limited sample}\label{ssec:verify}
+
+In this analysis, we have built a volume-limited sample from a set of magnitude
+limited ones, using a simple redshift limit. \nn{Even though it has the strength
+to be independent of potential inaccuracy in the modeling of the complex
+selection functions of the aforementioned dataset, it is suboptimal.
+That fact lead to the use of conservative cuts as first indicators that there is
+a drift that can't be explained by selection effects alone and a}n ongoing
+follow up analysis will make use of the SNANA package \citep{SNANA} to account
+for more subtle selection functions; \nn{however}, it is still necessary to test
+if the sample we have created could show sign of leftover sample selection.
+
+To do so, we compare the stretch and color distributions of the SNe~Ia in our
+sample between datasets of overlapping redshifts. If indeed we have built a
+volume-limited sample, the distributions of any SNe~Ia parameter at a given
+redshift should not depend of their survey or origin. When considering a
+redshift range, this has to be narrow enough such that any drift would be
+negligible. The two samples that overlap the most in redshift are PS1 and SDSS
+in the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
+corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and thus the
+most affected by potential selection effect leftover (see the corresponding
+discussion in section~\ref{sec:sample}). In that same redshift range, PS1 has 52
+SNe~Ia, which are at in the lowest redshift bin and thus unlikely to have any
+selection issue. The Fig.~\ref{fig:distrib} present the stretch and color
+distribution of both, together with p-values of Kolmogorov-Smirnov tests that
+\nn{probe whether} the $x_1$ and $c$ of the SDSS and PS1 samples at $0.10 < z <
+0.20$ arise from the same underlying distribution. Those p-values are high
+($>10\%$), and if one cannot conclude from this test that these SNe~Ia are from
+the same underlying distribution, one has no ground to claim they are not. This
+is in agreement with the visual impression from Fig.~\ref{fig:distrib}. A
+similar analysis \nn{was done} for PS1 and SNLS in the redshift range $0.20 < z
+< 0.31$ where the same conclusion \nn{can} be drawn, i.e., there is no sign of
+discrepancy in the stretch and color distributions between the low- and high-end
+of the SNLS and PS1 datasets, respectively. Nonetheless, the small size of the
+SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the interest of the
+test as only strong deviation would be noticeable. Extending the redshift range
+to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3) enables to get 31
+SNe~Ia from SNLS in total, the stretch and color p-values remains high, at 34\%
+and 4\% respectively.
 
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
-    \caption{\mr{$x_1$ and $c$ distribution histograms of different surveys
-    overlapping in redshift. \emph{Facing up} SDSS and PS1 within the redshift
-    range $0.10 < z < 0.20$ ; \emph{facing down} PS1 and SNLS within the
-    redshift range $0.20<z<0.3$. Error bars show the Poisson noise. Our stretch
-    "Base" modeling is illustrated in orange at the mean redshift of the
-    redshift ranges, $0.15$ and $0.25$, respectively. P-values of the
-    Kolmogorov-Smirnov tests are indicated on the top (bottom) of each panel
-    showing no sign that the SDSS and PS1 (PS1 and SNLS) $x_1$ and $c$
-    distributions are not drawn from the same underlying distributions.}}
+    \caption{$x_1$ and $c$ distribution histograms of different surveys
+        overlapping in redshift. \textit{Facing up}: SDSS and PS1 within the
+        redshift range $0.10 < z < 0.20$; \textit{facing down}: PS1 and SNLS
+        within the redshift range $0.20 < z < 0.31$. Error bars show the Poisson
+        noise. Our stretch ``Base'' modeling is illustrated in orange at the mean
+        redshift of the redshift ranges, $0.15$ and $0.25$, respectively.
+        P-values of the Kolmogorov-Smirnov tests are indicated on the top
+        (bottom) of each panel showing no sign that the SDSS and PS1 (PS1 and
+        SNLS) $x_1$ and $c$ distributions are not drawn from the same underlying
+        distributions.}
     \label{fig:distrib}
@@
 
-\mr{We highlight that the SNe~Ia color is more prone to selection effects than
+We highlight that the SNe~Ia color is more prone to selection effects than
 stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of
-    \cite{kessler2017}. Hence, since the color shows no sign of leftover
-    selection effects, it further supports our claim that our simple
-    redshift-based selection criteria is sufficient to test the redshift drift
-of the SN~Ia stretch distribution.}
+\cite{kessler2017}. Hence, since the color shows no sign of leftover selection
+effects, it further supports our claim that our simple redshift-based selection
+criteria is sufficient to test the \nn{existence of the} redshift drift of the
+SN~Ia stretch distribution.
 
@@
 what is expected if old environments favor low SN stretches
-\citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
+\citep[e.g.][]{howell2007} \textit{and} if the fraction of old SNe~Ia declines as
 a function of redshift. See Section~\ref{sec:results} for a more quantitative
@@
 
-We also consider another intrinsically non-drifting model, \mr{the functional
-    form} developed for Beams with Bias Correction \cite[BBC,][]{scolnic2016,
+We also consider another intrinsically non-drifting model, the functional form
+developed for Beams with Bias Correction \cite[BBC,][]{scolnic2016,
 kessler2017}, used in recent SN cosmological analyses
-    \cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to
-    account for Malmquist biases. The BBC formalism assumes sample-based (hence
+\cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to account
+for Malmquist biases. The BBC formalism assumes sample-based (hence
 intrinsically non-drifting) asymmetric Gaussian stretch distributions:
-    $\mathcal{N}\left(\mu,
-    \sigma_-{}^2\;\text{if}\;x_1<\mu,\;\text{else}\;\sigma_+{}^2\right)$. The
-    idea behind this sample-based approach is twofold: (1) Malmquist biases are
-    driven by survey properties and (2) since current surveys cover limited
-    redshift ranges, having a sample-based approach covers some potential
-    redshift evolution information \citep{scolnic2016, scolnic2018a}. See
-    further discussion concerning BBC in Section~\ref{sec:discussion}.
+$\mathcal{N}\left(\mu, \sigma_-{}^2\; \text{if} \;x_1<\mu,\; \text{else}
+\;\sigma_+{}^2\right)$. The idea behind this sample-based approach is twofold:
+(1) Malmquist biases are driven by survey properties and (2) since current
+surveys cover limited redshift ranges, having a sample-based approach covers
+some potential redshift evolution information \citep{scolnic2016, scolnic2018a}.
+See further discussion concerning BBC in Section~\ref{sec:discussion}.
 
@@
 
-We report in Table~\ref{tab:bbc} the samples' $\mu_0$ and $\sigma_{\pm}$
-adjusted on the nominally selection-free samples using our fiducial cuts (see
-Section~\ref{sec:sample}). We find our results in close agreement with
-\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
-derived these model parameters using the full BBC formalism, using numerous
-simulations to model the selection effects (see details e.g., Section~3 of
-\citealt{kessler2017}). The agreement between our fit of the asymmetric
-Gaussians on the supposedly selection-free part of the samples and the results
-derived using the BBC formalism supports our approach to get a sample with
-negligible selection effects. If we were to use \cite{scolnic2016} and
-\cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
-parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
-Base drifting model and the BBC modeling would go even deeper from $-32$ to
-$-47$. We further discuss the consequence of this result for cosmology in
-Section~\ref{sec:discussion}.
+We report in Table~\ref{tab:bbc} \nn{our determination of} the samples' $\mu_0$
+and $\sigma_{\pm}$ adjusted on the nominally selection-free samples using our
+fiducial cuts (see Section~\ref{sec:sample}) \nn{when implementing an asymmetric
+Gaussian fit}. We find our results in close agreement with \cite{scolnic2016}
+for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who derived these model
+parameters using the full BBC formalism, using numerous simulations to model the
+selection effects (see details e.g., Section~3 of \citealt{kessler2017}). The
+agreement between our fit of the asymmetric Gaussians on the supposedly
+selection-free part of the samples and the results derived using the BBC
+formalism supports our approach to get a sample with negligible selection
+effects. If we were to use \cite{scolnic2016} and \cite{scolnic2018a} best fit
+values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for SNLS, SDSS and
+PS1, respectively, the $\Delta$AIC between our Base drifting model and the BBC
+modeling would go even deeper from $-32$ to $-47$. We further discuss the
+consequence of this result for cosmology in Section~\ref{sec:discussion}.
 
@@
     \item The underlying SN~Ia stretch distribution is significantly redshift
-        dependent, as previously suggested by e.g.~\cite{howell2007}. This
-        result is largely independent of details on each age-population
-        modeling.
+        dependent, as previously suggested by e.g.~\cite{howell2007} \nn{in a
+        way that selection effects alone cannot explain}. This result is largely
+        independent of details on each age-population modeling.
 
@@
 \citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
-programs (high-$z$), respectively.
+programs (high-$z$), respectively. \nn{Moreover, a more detailed analysis using
+the SNANA framework is currently in progress in order to account for more subtle
+selection functions and refine the dataset.}
 
@@
     DE-AC025CH11231.
-    This project is partly financially supported by Region Rhone-Alpes-Auvergne
+    This project is partly financially supported by Région Rhône-Alpes-Auvergne.
 \end{acknowledgements}
@@
 \end{document}
-end{document}

### 2020-10-14 Nicolas.tex
MS comments

@@
 
-\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
+%\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
+\newcommand{\nn}[1]{\textcolor[rgb]{0.4, 0.8, 0}{#1}}
 \newcommand{\mr}[1]{\textcolor[rgb]{0.5, 0.20, 0.3}{#1}}
@@
     \and Saul Perlmutter\inst{3}
-    \and M. Smith\inst{4}
+    \and M. Smith\inst{2,5}
     \orcid{0000-0002-3321-1432}
@@
 these analyses, we will use $z_{\lim}=0.20$ as the baseline SDSS redshift limit,
-\nn{and the first 0.15 cut to compare the results of our modeling on a more
+\nn{and $z_{\lim}=0.15$ to compare the results of our modeling on a more
 restrictive sample.}
@@
 histograms. \nn{We test in Section~\ref{ssec:verify} whether we can consider
-this sample to be free of selection effects. The results exposed} in
-Section~\ref{sec:results} further supports that this sample selection indeed
+this sample to be free of selection effects. The consistency of the results for the two samples, discussed} in
+Section~\ref{sec:results}, further supports that this sample selection indeed
 provides a subset of SNe~Ia with insignificant selection effects when
@@
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}, \nn{where the
-disappearance of the low-stretch population (typically $x_1 < -1$) is what we
-test in this paper}. The evolution of the mean stretch is also shown in
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \nn{Proportionally few low-stretch SNe (typically $x_1 < -1$) are observed at $z>0.4$}. The evolution of the mean stretch is also shown in
 Fig.~\ref{fig:modelall}, where the data are split in redshift bins of regular
@@
 In this analysis, we have built a volume-limited sample from a set of magnitude
-limited ones, using a simple redshift limit. \nn{Even though it has the strength
-to be independent of potential inaccuracy in the modeling of the complex
-selection functions of the aforementioned dataset, it is suboptimal.
-That fact lead to the use of conservative cuts as first indicators that there is
-a drift that can't be explained by selection effects alone and a}n ongoing
-follow up analysis will make use of the SNANA package \citep{SNANA} to account
+limited ones, using a simple redshift limit. \nn{The advantage of this approach is that our results are robust to any potential inaccuracy in the modeling of the complex
+selection functions for the individual datasets.
+However, while we will find consistent results when considering a sample derived from conservative cuts (Section~\ref{sec:results}), there remains the possibility that our sample is not fully volume-limited and that unaccounted for selection effects remain.
+To fully account for this possibility, a} follow up analysis will make use of the SNANA package \citep{SNANA} to account
 for more subtle selection functions; \nn{however}, it is still necessary to test
 if the sample we have created could show sign of leftover sample selection.
+%While our conservative comparision sample is That fact lead to the use of conservative cuts as first indicators that there isa drift that can't be explained by selection effects alone and a}n ongoing
 
@@
 the same underlying distribution, one has no ground to claim they are not. This
-is in agreement with the visual impression from Fig.~\ref{fig:distrib}. A
-similar analysis \nn{was done} for PS1 and SNLS in the redshift range $0.20 < z
+is in agreement with the visual impression from Fig.~\ref{fig:distrib}. \nn{We perform a}
+similar analysis for PS1 and SNLS in the redshift range $0.20 < z
 < 0.31$ where the same conclusion \nn{can} be drawn, i.e., there is no sign of
@@
 \citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
-programs (high-$z$), respectively. \nn{Moreover, a more detailed analysis using
-the SNANA framework is currently in progress in order to account for more subtle
-selection functions and refine the dataset.}
+programs (high-$z$), respectively. \nn{To fully break the degeneracy between selection effects and population evolution, an analysis using
+the SNANA framework is underway.
+This analysis will allow for subtle selection effects to be considered, and by removing the need for a redshift cut, increase the redshift leverage and statistical significance of our results.}
 

### 2020-10-22 Nicolas.tex
YC comments

@@
 \newcommand{\mr}[1]{\textcolor[rgb]{0.5, 0.20, 0.3}{#1}}
+\newcommand{\yc}[1]{\textcolor[RGB]{217, 22, 102}{#1}}
 \newcommand{\prob}[2]{\mathcal{P}\left( #1 \mid #2\right)}
@@
     prediction against published samples \nn{\sout{chosen} cut} to have
-    negligible magnitude selection effects, so that any observed change is
+    \yc{\sout{negligible} marginal} magnitude selection effects, so that any observed change is
     indeed of astrophysical and not observational origin. \nn{An ongoing
     analysis with the SNANA package will test this assumption in order to refine
-    the study.} We \nn{\sout{clearly}} demonstrate that the underlying SN~Ia stretch
+    the study.} \yc{YC: I don't think this should be in the abstract, rather emphasize it's a preliminary study.} \yc{\sout{We \nn{\sout{clearly}} demonstrate} There are indications} that the underlying SN~Ia stretch
     distribution is evolving as a function of redshift, and that the young/old
-    drifting model is a much better description of the data than any time-constant
+    drifting model is a \yc{\sout{much}} better description of the data than any time-constant
     model, including the sample-based asymmetric distributions usually used to
@@
 
+\yc{YC: you should introduce the principle of a fiducial sample, and a more conservative one, smaller and therefore less statistically constraining, but inherently less prone to selection effects.
+If the redshift drift is still significant in the conservative sample, it would be even more significant in a carefully-tailored selection-free sample: that's the goal of this preliminary paper.}
+
 SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
@@
 selection effects, we also take into account another, more restrictive,
-redshift limit:} Fig.~14 of \cite[][see their Section~5]{perrett2010} suggests a
+redshift limit:} Fig.~14 of \citet[][see their Section~5]{perrett2010} suggests a
 lower limit of $z_{\lim}=0.55$; both limits will be considered, as discussed
@@
 the SDSS surveys were more sensitive to limited spectroscopic resources;
-\cite[][see their Section~2]{kessler2009} pointed out that during the first year
+\citet[][see their Section~2]{kessler2009} pointed out that during the first year
 of SDSS, SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up,
@@
 histograms. \nn{We test in Section~\ref{ssec:verify} whether we can consider
-this sample to be free of selection effects. The consistency of the results for the two samples, discussed} in
+this sample to be free of \yc{(or at least marginally affected by)} selection effects. The consistency of the results for the two samples, discussed} in
 Section~\ref{sec:results}, further supports that this sample selection indeed
@@
 particularly useful for studying SN property drift, as it enables us to have
-a large SN~Ia sample at $z<0.1$.
+a large \yc{complete} SN~Ia sample at $z<0.1$.
 
@@
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \nn{Proportionally few low-stretch SNe (typically $x_1 < -1$) are observed at $z>0.4$}. The evolution of the mean stretch is also shown in
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}.
+\nn{Proportionally few low-stretch SNe (typically $x_1 < -1$) are observed at $z>0.4$}.
+\yc{YC: this rather looks like $z\sim 0.2-0.3$, as in Fig.~\ref{fig:modelall}.} The evolution of the mean stretch is also shown in
 Fig.~\ref{fig:modelall}, where the data are split in redshift bins of regular
@@
 0.10$ at $z\sim0.05$), suggesting that the underlying stretch distribution is
-drifting.
+drifting. \yc{YC: I prefer ``evolving with redshift''.}
 
@@
         histograms in dark (resp.\ light) colors for the conservative (resp.
-        \ fiducial) redshift cuts.}
+        \ fiducial) redshift cuts. \yc{YC: you should add a running mean $\pm$ stddev to show the evolution at $z \sim 0.2-0.3$.}}
     \label{fig:sample}
@@
 selection functions for the individual datasets.
-However, while we will find consistent results when considering a sample derived from conservative cuts (Section~\ref{sec:results}), there remains the possibility that our sample is not fully volume-limited and that unaccounted for selection effects remain.
+\yc{YC: once again, I don't think it's a fair statement.
+You should rather emphasize that this drastic approach is more robust at the price of being statistically sub-optimal, but is enough to demonstrate the key finding: a redshift evolution of stretch compatible with R+18 model.} However, while we will find consistent results when considering a sample derived from conservative cuts (Section~\ref{sec:results}), there remains the possibility that our sample is not fully volume-limited and that unaccounted for selection effects remain.
 To fully account for this possibility, a} follow up analysis will make use of the SNANA package \citep{SNANA} to account
@@
 if the sample we have created could show sign of leftover sample selection.
-%While our conservative comparision sample is That fact lead to the use of conservative cuts as first indicators that there isa drift that can't be explained by selection effects alone and a}n ongoing
+%While our conservative comparison sample is That fact lead to the use of conservative cuts as first indicators that there isa drift that can't be explained by selection effects alone and a}n ongoing 
+\yc{YC: this paragraph to be rephrased IMHO.}
 
@@
 volume-limited sample, the distributions of any SNe~Ia parameter at a given
-redshift should not depend of their survey or origin. When considering a
+redshift should not depend of their survey or origin. \yc{YC: survey/origin, what's the difference?} When considering a
 redshift range, this has to be narrow enough such that any drift would be
-negligible. The two samples that overlap the most in redshift are PS1 and SDSS
+negligible. \yc{YC: why do you say that here, since you didn't fine-tuned the overlaping $z$-range?} The two samples that overlap the most in redshift are PS1 and SDSS
 in the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
@@
 most affected by potential selection effect leftover (see the corresponding
-discussion in section~\ref{sec:sample}). In that same redshift range, PS1 has 52
+discussion in section~\ref{sec:sample} \yc{YC: we are in Sec.~\ref{sec:sample}!}). In that same redshift range, PS1 has 52
 SNe~Ia, which are at in the lowest redshift bin and thus unlikely to have any
-selection issue. The Fig.~\ref{fig:distrib} present the stretch and color
-distribution of both, together with p-values of Kolmogorov-Smirnov tests that
-\nn{probe whether} the $x_1$ and $c$ of the SDSS and PS1 samples at $0.10 < z <
-0.20$ arise from the same underlying distribution. Those p-values are high
+selection issue. \yc{\sout{The}} Fig.~\ref{fig:distrib} present\yc{s} the stretch and color
+distribution of both \yc{YC: what?}, together with $p$-values of Kolmogorov-Smirnov tests that
+\yc{\sout{\nn{probe whether}}} the $x_1$ and $c$ of the SDSS and PS1 samples at $0.10 < z <
+0.20$ arise from the same underlying distribution. These $p$-values are high
 ($>10\%$), and if one cannot conclude from this test that these SNe~Ia are from
-the same underlying distribution, one has no ground to claim they are not. This
+the same underlying distribution, one has no ground to claim they are not.
+\yc{YC: I suggest ``These $p$-values are too high ($>10\%$) to exclude that both samples originate from the same underlying population.''} This
 is in agreement with the visual impression from Fig.~\ref{fig:distrib}. \nn{We perform a}
@@
 to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3) enables to get 31
-SNe~Ia from SNLS in total, the stretch and color p-values remains high, at 34\%
+SNe~Ia from SNLS in total, the stretch and color \yc{KS-test} $p$-values remains \yc{\sout{high} inconclusive}, at 34\%
 and 4\% respectively.
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
-    \caption{$x_1$ and $c$ distribution histograms of different surveys
+    \caption{$x_1$ \yc{(left)} and $c$ \yc{(right)} distribution histograms of different surveys
         overlapping in redshift. \textit{Facing up}: SDSS and PS1 within the
@@
         within the redshift range $0.20 < z < 0.31$. Error bars show the Poisson
-        noise. Our stretch ``Base'' modeling is illustrated in orange at the mean
+        noise \yc{YC: not visible enough}. Our stretch ``Base'' model\yc{\sout{ing}} is illustrated in orange at the mean
         redshift of the redshift ranges, $0.15$ and $0.25$, respectively.
-        P-values of the Kolmogorov-Smirnov tests are indicated on the top
-        (bottom) of each panel showing no sign that the SDSS and PS1 (PS1 and
+        \yc{Kolmogorov-Smirnov test $p$-values} are indicated on the top
+        (\yc{resp.} bottom) of each panel showing no sign that the SDSS and PS1 (\yc{resp.} PS1 and
         SNLS) $x_1$ and $c$ distributions are not drawn from the same underlying
@@
 
-We highlight that the SNe~Ia color is more prone to selection effects than
+We \yc{furthermore} highlight that the SNe~Ia color is more prone to selection effects than
 stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of
-\cite{kessler2017}. Hence, since the color shows no sign of leftover selection
-effects, it further supports our claim that our simple redshift-based selection
-criteria is sufficient to test the \nn{existence of the} redshift drift of the
-SN~Ia stretch distribution.
+\cite{kessler2017}. Hence, since the color \yc{distribution} shows no \yc{\sout{sign} significant hint} of \yc{\sout{leftover}} selection
+effect, it further supports our claim that our simple redshift-based selection
+criteria is sufficient to \yc{\sout{test the \nn{existence of the} redshift drift of the
+SN~Ia stretch distribution} build an unbiased SN~Ia sample at all redshifts required to test the redshift evolution of the stretch distribution}.
 
@@
 fiducial cuts (see Section~\ref{sec:sample}) \nn{when implementing an asymmetric
-Gaussian fit}. We find our results in close agreement with \cite{scolnic2016}
+Gaussian fit}. \yc{YC: ???} We find our results in close agreement with \cite{scolnic2016}
 for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who derived these model
@@
 
-In this paper, we considered a simple Gaussian mixture modeling, but additional
+In this paper, we considered a simple \yc{two-population} Gaussian mixture modeling, but additional
 data free from significant Malmquist bias would enable us to refine it as
@@
 \citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
-programs (high-$z$), respectively. \nn{To fully break the degeneracy between selection effects and population evolution, an analysis using
+programs (high-$z$), respectively.
+\nn{To \yc{\sout{fully break the degeneracy between} improve the disentangling of} \yc{observational} selection effects and \yc{intrinsic} population evolution, an analysis using
 the SNANA framework is underway.
-This analysis will allow for subtle selection effects to be considered, and by removing the need for a redshift cut, increase the redshift leverage and statistical significance of our results.}
+It will allow for subtle selection effects to be \yc{\sout{considered} accounted for}, and by removing the need for a \yc{naive} redshift cut, increase the redshift leverage and statistical significance of our results.}
 
@@
     This project has received funding from the European Research Council (ERC)
-    under the European Union's Horizon 2020 research and innovation programme
-    (grant agreement n 759194 - USNAC).
+    under the European Union's Horizon 2020 Research and Innovation program
+    (grant agreement no 759194 - USNAC).
     This work was supported in part by the Director, Office of Science, Office
@@
 
+\yc{YC: check again all arXiv references.}
+
 \bibliographystyle{aa}
@@
 \end{thebibliography}
-\end{document}
+\end{document}end{document}

### 2020-11-09 Nicolas.tex
MS and YC comments implmented

@@
     distribution for each age population made of Gaussian mixtures. We test our
-    prediction against published samples \nn{\sout{chosen} cut} to have
-    \yc{\sout{negligible} marginal} magnitude selection effects, so that any observed change is
-    indeed of astrophysical and not observational origin. \nn{An ongoing
-    analysis with the SNANA package will test this assumption in order to refine
-    the study.} \yc{YC: I don't think this should be in the abstract, rather emphasize it's a preliminary study.} \yc{\sout{We \nn{\sout{clearly}} demonstrate} There are indications} that the underlying SN~Ia stretch
-    distribution is evolving as a function of redshift, and that the young/old
-    drifting model is a \yc{\sout{much}} better description of the data than any time-constant
-    model, including the sample-based asymmetric distributions usually used to
-    correct Malmquist bias. The favored underlying stretch model is the bimodal one
-    derived from \cite{rigault2018}: a high-stretch mode shared by both young and
-    old environments, and a low-stretch mode exclusive to old environments. The
+    prediction against published samples cut to have marginal magnitude
+    selection effects, so that any observed change is indeed of astrophysical
+    and not observational origin. \nn{In this preliminary study,} there are
+    indications that the underlying SN~Ia stretch distribution is evolving as a
+    function of redshift, and that the young/old drifting model is a better
+    description of the data than any time-constant model, including the
+    sample-based asymmetric distributions usually used to correct Malmquist
+    bias. The favored underlying stretch model is the bimodal one derived from
+    \cite{rigault2018}: a high-stretch mode shared by both young and old
+    environments, and a low-stretch mode exclusive to old environments. The
     precise impact of the redshift evolution of the SN Ia population intrinsic
-    properties on cosmology remains to be studied. Yet, the astrophysical drift of
-    the SN stretch distribution does affect current Malmquist bias corrections and
-    thereby distances derived from SN affected by selection effects. We highlight
-    that such a bias will increase with surveys covering increasingly larger
-    redshift ranges, which is particularly important for LSST.}
+    properties on cosmology remains to be studied. Yet, the astrophysical drift
+    of the SN stretch distribution does affect current Malmquist bias
+    corrections and thereby distances derived from SN affected by selection
+effects. We highlight that such a bias will increase with surveys covering
+increasingly larger redshift ranges, which is particularly important for LSST.}
 
@@
 
-\yc{YC: you should introduce the principle of a fiducial sample, and a more conservative one, smaller and therefore less statistically constraining, but inherently less prone to selection effects.
-If the redshift drift is still significant in the conservative sample, it would be even more significant in a carefully-tailored selection-free sample: that's the goal of this preliminary paper.}
+\nn{We will thus consider a set of cuts that will define a first fiducial
+    sample, taking the farthest limits possible; but as this solution might be
+    suboptimal to create a complete sample, we also consider another set of cuts
+    to define a conservative sample, smaller and therefore less statistically
+    constraining, but inherently less prone to selection effects. If the
+    redshift drift is still significant in the conservative sample, it would be
+    even more significant in a carefully-tailored selection-free sample: the
+goal of this preliminary paper is to make a first implementation of this
+modeling before refining it on a better sample.}
 
@@
 histograms. \nn{We test in Section~\ref{ssec:verify} whether we can consider
-this sample to be free of \yc{(or at least marginally affected by)} selection effects. The consistency of the results for the two samples, discussed} in
+this sample to be free of (or at least marginally affected by) selection
+effects. The consistency of the results for the two samples, discussed} in
 Section~\ref{sec:results}, further supports that this sample selection indeed
-provides a subset of SNe~Ia with insignificant selection effects when
-compared to state-of-the-art Malmquist correction techniques.
+provides a subset of SNe~Ia with insignificant selection effects when compared
+to state-of-the-art Malmquist correction techniques.
 
@@
 particularly useful for studying SN property drift, as it enables us to have
-a large \yc{complete} SN~Ia sample at $z<0.1$.
+a large complete SN~Ia sample at $z<0.1$.
 
@@
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}.
-\nn{Proportionally few low-stretch SNe (typically $x_1 < -1$) are observed at $z>0.4$}.
-\yc{YC: this rather looks like $z\sim 0.2-0.3$, as in Fig.~\ref{fig:modelall}.} The evolution of the mean stretch is also shown in
-Fig.~\ref{fig:modelall}, where the data are split in redshift bins of regular
-sample size. We see that SNe~Ia at higher redshift have on average larger
-stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift ($-0.17\pm
-0.10$ at $z\sim0.05$), suggesting that the underlying stretch distribution is
-drifting. \yc{YC: I prefer ``evolving with redshift''.}
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \nn{Proportionally
+few low-stretch SNe (typically $x_1 < -1$) are observed at $z>0.2$}. The
+evolution of the mean stretch is also shown in Fig.~\ref{fig:modelall}, where
+the data are split in redshift bins of regular sample size. We see that SNe~Ia
+at higher redshift have on average larger stretch ($0.34 \pm 0.10$ at
+$z\sim0.65$) than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$),
+suggesting that the underlying stretch distribution is evolving with redshift.
 
@@
         (resp.\ fiducial) redshift cuts. \textit{Top:} stacked redshift
-        histograms in dark (resp.\ light) colors for the conservative (resp.
-        \ fiducial) redshift cuts. \yc{YC: you should add a running mean $\pm$ stddev to show the evolution at $z \sim 0.2-0.3$.}}
+    histograms in dark (resp.\ light) colors for the conservative (resp.  \
+fiducial) redshift cuts. \yc{YC: you should add a running mean $\pm$ stddev to
+show the evolution at $z \sim 0.2-0.3$.}}
     \label{fig:sample}
@@
 In this analysis, we have built a volume-limited sample from a set of magnitude
-limited ones, using a simple redshift limit. \nn{The advantage of this approach is that our results are robust to any potential inaccuracy in the modeling of the complex
-selection functions for the individual datasets.
-\yc{YC: once again, I don't think it's a fair statement.
-You should rather emphasize that this drastic approach is more robust at the price of being statistically sub-optimal, but is enough to demonstrate the key finding: a redshift evolution of stretch compatible with R+18 model.} However, while we will find consistent results when considering a sample derived from conservative cuts (Section~\ref{sec:results}), there remains the possibility that our sample is not fully volume-limited and that unaccounted for selection effects remain.
-To fully account for this possibility, a} follow up analysis will make use of the SNANA package \citep{SNANA} to account
-for more subtle selection functions; \nn{however}, it is still necessary to test
-if the sample we have created could show sign of leftover sample selection.
-%While our conservative comparison sample is That fact lead to the use of conservative cuts as first indicators that there isa drift that can't be explained by selection effects alone and a}n ongoing 
-\yc{YC: this paragraph to be rephrased IMHO.}
+limited ones, using a simple redshift limit. \nn{This drastic approach is robust
+at the price of being statistically sub-optimal, but is enough to
+demonstrate our key finding: a redshift evolution of stretch compatible with
+the \cite{rigault2018} model.} However, while we will find consistent results
+when considering a sample derived from conservative cuts
+(Section~\ref{sec:results}), there remains the possibility that our sample is
+not fully volume-limited and that unaccounted for selection effects remain. To
+fully account for this possibility, a follow up analysis will make use of the
+SNANA package \citep{SNANA} to account for more subtle selection functions;
+\nn{yet, it appears to still be necessary} to test if the sample we have created
+could show sign of leftover sample selection.
 
@@
 volume-limited sample, the distributions of any SNe~Ia parameter at a given
-redshift should not depend of their survey or origin. \yc{YC: survey/origin, what's the difference?} When considering a
+redshift should not depend of their survey of origin. When considering a
 redshift range, this has to be narrow enough such that any drift would be
-negligible. \yc{YC: why do you say that here, since you didn't fine-tuned the overlaping $z$-range?} The two samples that overlap the most in redshift are PS1 and SDSS
-in the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
+negligible. \yc{YC: \textit{why do you say that here, since you didn't fine-tune
+the overlaping $z$-range?}} \nn{\textit{Because we still took that into
+account?}} The two samples that overlap the most in redshift are PS1 and SDSS in
+the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
 corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and thus the
 most affected by potential selection effect leftover (see the corresponding
-discussion in section~\ref{sec:sample} \yc{YC: we are in Sec.~\ref{sec:sample}!}). In that same redshift range, PS1 has 52
-SNe~Ia, which are at in the lowest redshift bin and thus unlikely to have any
-selection issue. \yc{\sout{The}} Fig.~\ref{fig:distrib} present\yc{s} the stretch and color
-distribution of both \yc{YC: what?}, together with $p$-values of Kolmogorov-Smirnov tests that
-\yc{\sout{\nn{probe whether}}} the $x_1$ and $c$ of the SDSS and PS1 samples at $0.10 < z <
-0.20$ arise from the same underlying distribution. These $p$-values are high
-($>10\%$), and if one cannot conclude from this test that these SNe~Ia are from
-the same underlying distribution, one has no ground to claim they are not.
-\yc{YC: I suggest ``These $p$-values are too high ($>10\%$) to exclude that both samples originate from the same underlying population.''} This
-is in agreement with the visual impression from Fig.~\ref{fig:distrib}. \nn{We perform a}
-similar analysis for PS1 and SNLS in the redshift range $0.20 < z
-< 0.31$ where the same conclusion \nn{can} be drawn, i.e., there is no sign of
-discrepancy in the stretch and color distributions between the low- and high-end
-of the SNLS and PS1 datasets, respectively. Nonetheless, the small size of the
-SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the interest of the
-test as only strong deviation would be noticeable. Extending the redshift range
-to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3) enables to get 31
-SNe~Ia from SNLS in total, the stretch and color \yc{KS-test} $p$-values remains \yc{\sout{high} inconclusive}, at 34\%
-and 4\% respectively.
+discussion in section~\ref{ssec:cuts}). In that same redshift range, PS1 has 52
+SNe~Ia, which are in the lowest redshift bins and thus unlikely to have any
+selection issue. Fig.~\ref{fig:distrib} presents the stretch and color
+distribution of both \nn{these surveys}, together with $p$-values of
+Kolmogorov-Smirnov tests that \yc{\sout{\nn{probe whether}}}\nn{\textit{I don't
+see anything else to put here without repeating ``test''}} the $x_1$ and $c$ of
+the SDSS and PS1 samples at $0.10 < z < 0.20$ arise from the same underlying
+distribution. These $p$-values are high ($>10\%$), and if one cannot conclude
+from this test that these SNe~Ia are from the same underlying distribution, one
+has no ground to claim they are not. \yc{YC: I suggest ``These $p$-values are
+too high ($>10\%$) to exclude that both samples originate from the same
+underlying population.''} This is in agreement with the visual impression from
+Fig.~\ref{fig:distrib}. We perform a similar analysis for PS1 and SNLS in the
+redshift range $0.20 < z < 0.31$ where the same conclusion can be drawn, i.e.,
+there is no sign of discrepancy in the stretch and color distributions between
+the low- and high-end of the SNLS and PS1 datasets, respectively. Nonetheless,
+the small size of the SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1)
+limits the interest of the test as only strong deviation would be noticeable.
+Extending the redshift range to $0.20 < z < 0.40$ (though we have no PS1 data
+above 0.3) enables to get 31 SNe~Ia from SNLS in total, yet the stretch and
+color KS-test $p$-values remains inconclusive, at 34\% and 4\% respectively.
 
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
-    \caption{$x_1$ \yc{(left)} and $c$ \yc{(right)} distribution histograms of different surveys
-        overlapping in redshift. \textit{Facing up}: SDSS and PS1 within the
-        redshift range $0.10 < z < 0.20$; \textit{facing down}: PS1 and SNLS
+    \caption{$x_1$ (left) and $c$ (right) distribution histograms of different
+        surveys overlapping in redshift. \textit{Facing up}: SDSS and PS1 within
+        the redshift range $0.10 < z < 0.20$; \textit{facing down}: PS1 and SNLS
         within the redshift range $0.20 < z < 0.31$. Error bars show the Poisson
-        noise \yc{YC: not visible enough}. Our stretch ``Base'' model\yc{\sout{ing}} is illustrated in orange at the mean
-        redshift of the redshift ranges, $0.15$ and $0.25$, respectively.
-        \yc{Kolmogorov-Smirnov test $p$-values} are indicated on the top
-        (\yc{resp.} bottom) of each panel showing no sign that the SDSS and PS1 (\yc{resp.} PS1 and
-        SNLS) $x_1$ and $c$ distributions are not drawn from the same underlying
-        distributions.}
+        noise \yc{YC: not visible enough}. Our stretch ``Base'' model is
+        illustrated in orange at the mean redshift of the redshift ranges,
+        $0.15$ and $0.25$, respectively.  Kolmogorov-Smirnov test $p$-values are
+    indicated on the top (resp. bottom) of each panel showing no sign that the
+SDSS and PS1 (resp. PS1 and SNLS) $x_1$ and $c$ distributions are not drawn from
+the same underlying distributions.}
     \label{fig:distrib}
@@
 
-We \yc{furthermore} highlight that the SNe~Ia color is more prone to selection effects than
-stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of
-\cite{kessler2017}. Hence, since the color \yc{distribution} shows no \yc{\sout{sign} significant hint} of \yc{\sout{leftover}} selection
-effect, it further supports our claim that our simple redshift-based selection
-criteria is sufficient to \yc{\sout{test the \nn{existence of the} redshift drift of the
-SN~Ia stretch distribution} build an unbiased SN~Ia sample at all redshifts required to test the redshift evolution of the stretch distribution}.
+We furthermore highlight that the SNe~Ia color is more prone to selection
+effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
+Fig.~3 of \cite{kessler2017}. Hence, since the color distribution shows no
+significant hint of selection effect, it further supports our claim that our
+simple redshift-based selection criteria is sufficient to \nn{build an unbiased
+SNe~Ia sample at all redshifts required to test the redshift evolution of the
+stretch distribution}.
 
@@
 We report in Table~\ref{tab:bbc} \nn{our determination of} the samples' $\mu_0$
-and $\sigma_{\pm}$ adjusted on the nominally selection-free samples using our
-fiducial cuts (see Section~\ref{sec:sample}) \nn{when implementing an asymmetric
-Gaussian fit}. \yc{YC: ???} We find our results in close agreement with \cite{scolnic2016}
-for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who derived these model
-parameters using the full BBC formalism, using numerous simulations to model the
-selection effects (see details e.g., Section~3 of \citealt{kessler2017}). The
-agreement between our fit of the asymmetric Gaussians on the supposedly
-selection-free part of the samples and the results derived using the BBC
-formalism supports our approach to get a sample with negligible selection
-effects. If we were to use \cite{scolnic2016} and \cite{scolnic2018a} best fit
-values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for SNLS, SDSS and
-PS1, respectively, the $\Delta$AIC between our Base drifting model and the BBC
-modeling would go even deeper from $-32$ to $-47$. We further discuss the
-consequence of this result for cosmology in Section~\ref{sec:discussion}.
+and $\sigma_{\pm}$ \nn{when implementing an asymmetric Gaussian model}, adjusted
+on the nominally selection-free samples using our fiducial cuts (see
+Section~\ref{sec:sample}). We find our results in close agreement with
+\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
+derived these model parameters using the full BBC formalism, using numerous
+simulations to model the selection effects (see details e.g., Section~3 of
+\citealt{kessler2017}). The agreement between our fit of the asymmetric
+Gaussians on the supposedly selection-free part of the samples and the results
+derived using the BBC formalism supports our approach to get a sample with
+negligible selection effects. If we were to use \cite{scolnic2016} and
+\cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
+parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
+Base drifting model and the BBC modeling would go even deeper from $-32$ to
+$-47$. We further discuss the consequence of this result for cosmology in
+Section~\ref{sec:discussion}.
 
@@
 
-In this paper, we considered a simple \yc{two-population} Gaussian mixture modeling, but additional
-data free from significant Malmquist bias would enable us to refine it as
-necessary. We note that samples at the low- and high-redshift ends of the Hubble
-diagram would be particularly helpful for this drifting analysis; fortunately
-this will soon be provided by the Zwicky Transient Facility
+In this paper, we considered a simple two-population Gaussian mixture modeling,
+but additional data free from significant Malmquist bias would enable us to
+refine it as necessary. We note that samples at the low- and high-redshift ends
+of the Hubble diagram would be particularly helpful for this drifting analysis;
+fortunately this will soon be provided by the Zwicky Transient Facility
 \citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
-programs (high-$z$), respectively.
-\nn{To \yc{\sout{fully break the degeneracy between} improve the disentangling of} \yc{observational} selection effects and \yc{intrinsic} population evolution, an analysis using
-the SNANA framework is underway.
-It will allow for subtle selection effects to be \yc{\sout{considered} accounted for}, and by removing the need for a \yc{naive} redshift cut, increase the redshift leverage and statistical significance of our results.}
+programs (high-$z$), respectively. \nn{To improve the disentangling of
+observational} selection effects and \nn{intrinsic} population evolution, an
+analysis using the SNANA framework is underway. It will allow for subtle
+selection effects to be \nn{accounted for}, and by removing the need for a
+\nn{naive} redshift cut, increase the redshift leverage and statistical
+significance of our results.
 
@@
 \end{thebibliography}
-\end{document}end{document}
+\end{document}

### 2020-11-13 Nicolas.tex
MR comments to be included.

@@
     selection effects, so that any observed change is indeed of astrophysical
-    and not observational origin. \nn{In this preliminary study,} there are
+    and not observational origin. \nn{In this \mr{first} study,} there are
     indications that the underlying SN~Ia stretch distribution is evolving as a
@@
 negligible below a given redshift at which even the faintest SNe~Ia can be
-observed. In contrast, targeted surveys have highly complex selection functions
+observed. \mr{In practice such spectroscopic following has not always been the norm as discussed below.} In contrast, targeted surveys have highly complex selection functions
 and will be discarded from our analysis. Fortunately, modern SN cosmology
@@
 \nn{We will thus consider a set of cuts that will define a first fiducial
-    sample, taking the farthest limits possible; but as this solution might be
-    suboptimal to create a complete sample, we also consider another set of cuts
-    to define a conservative sample, smaller and therefore less statistically
-    constraining, but inherently less prone to selection effects. If the
-    redshift drift is still significant in the conservative sample, it would be
-    even more significant in a carefully-tailored selection-free sample: the
-goal of this preliminary paper is to make a first implementation of this
-modeling before refining it on a better sample.}
+    sample, taking the farthest limits possible\mr{.
+But} as this solution might be \mr{over simplistic} to create a complete sample \mr{, e.g., for it ignores follow-up efficiency}, we also consider another set of cuts
+    to define a conservative sample.
+\mr{This latter is} smaller and therefore \mr{has} less statistically constraining, but inherently less prone to selection effects.
+If the redshift drift is still significant in the conservative sample, it would be
+    even more significant in a carefully-tailored selection-free sample\mr{. T}he goal of this \mr{study} is to \mr{build} \mr{the} first implementation of \mr{a SN properties drift}
+    modeling before refining it on a \mr{more complex, yet more accurate, follow-up analysis accounting at the same time for precise selection modeling using of the SNANA package \citep{SNANA}.}}
 
@@
 final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement
-with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. \nn{However,
-because this simple approach might easily not be enough to alleviate all
-selection effects, we also take into account another, more restrictive,
-redshift limit:} Fig.~14 of \citet[][see their Section~5]{perrett2010} suggests a
-lower limit of $z_{\lim}=0.55$; both limits will be considered, as discussed
-below.
+with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. 
+Fig.~14 of \citet[][see their Section~5]{perrett2010} suggests however a
+lower limit of $z_{\lim}=0.55$. \mr{We will therefore use $z=0.60$ and $z=0.55$ as redshift limits for the fiducial and conservative cuts, respectively.}
 
@@
 $5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.31$ in
-agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure is also
-suggestive of a more conservative $z_{\lim}=0.27$, \nn{similarly considered in
-the following analysis as a first step to evaluate the robustness of our
-sample construction.}
+agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. \mr{If we were to be conservative, this figure would also
+suggest of a more stringent $z_{\lim}=0.27$ cut ; hence, we will use $0.31$ and $0.27$ for our fiducial and conservative cuts, respectively.}
 
@@
 \citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However,
-the SDSS surveys were more sensitive to limited spectroscopic resources;
-\citet[][see their Section~2]{kessler2009} pointed out that during the first year
-of SDSS, SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up,
+the SDSS surveys were more sensitive to limited spectroscopic resources as discussed before.
+\citet[][Section~2]{kessler2009} pointed out that during the first year of SDSS, SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up,
 corresponding to a redshift cut at $0.15$. For the rest of the SDSS survey,
@@
 \cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
-these analyses, we will use $z_{\lim}=0.20$ as the baseline SDSS redshift limit,
-\nn{and $z_{\lim}=0.15$ to compare the results of our modeling on a more
-restrictive sample.}
+these analyses, \mr{we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for out fiducial and conservative cuts, respectively.}
 
@@
 distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
-expected, the selected redshift limits roughly correspond to the peak of these
-histograms. \nn{We test in Section~\ref{ssec:verify} whether we can consider
-this sample to be free of (or at least marginally affected by) selection
-effects. The consistency of the results for the two samples, discussed} in
-Section~\ref{sec:results}, further supports that this sample selection indeed
-provides a subset of SNe~Ia with insignificant selection effects when compared
-to state-of-the-art Malmquist correction techniques.
+expected, the selected redshift limits \mr{are roughly located slightly before the peak of these histograms}.
+\nn{We test in Section~\ref{ssec:verify} \mr{the validity of our intent to build volume limited surveys from (mostly) magnitude limited ones with the aforementioned redshift limits.}}
+%We also highlight} that the consistency of the results between the fiducial and the conservative samples, discussed} in Section~\ref{sec:results}, further supports that this sample selection indeed provides a subset of SNe~Ia with marginal leftover selection effects when compared to state-of-the-art Malmquist correction techniques.
 
@@
 and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
-\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
-range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
-much deeper. These 114 SNfactory SNe~Ia \nn{(80\% of the 141 SNe presented in
-the previously-cited paper)} are thus in the volume-limited part of the
+\citealt{aldering2020}). \mr{For this dataset,} the spectroscopic follow-up was done over a redshift
+range of $0.02 < z < 0.09$, while the search was
+much deeper. These 114 SNfactory SNe~Ia 
+%\nn{(80\% of the 141 SNe presented in
+%the previously-cited paper)}  %%% NO NEED IT ADDS CONFUSION even though the referee were confused
+are thus in the volume-limited part of the
 survey (Aldering et al., in prep.), and are therefore assumed to be a random
@@
 \citep{strolger04}.
+\mr{At these redshift, the supernovae typing is challenging but the target classification were robust enough to include them within the cosmological analysis \citep{scolnic2018a} ; we do not add further quality cuts.}
 
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \nn{Proportionally
-few low-stretch SNe (typically $x_1 < -1$) are observed at $z>0.2$}. The
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \mr{We observe that the apparent fraction of low-stretch SNe (typically $x_1 < -1$) seems to decrease as a function of redshift.} The
 evolution of the mean stretch is also shown in Fig.~\ref{fig:modelall}, where
@@
 fiducial) redshift cuts. \yc{YC: you should add a running mean $\pm$ stddev to
-show the evolution at $z \sim 0.2-0.3$.}}
+show the evolution at $z \sim 0.2-0.3$.} \mr{Agreed if it is not too busy.}}
     \label{fig:sample}
@@
 
-In this analysis, we have built a volume-limited sample from a set of magnitude
-limited ones, using a simple redshift limit. \nn{This drastic approach is robust
-at the price of being statistically sub-optimal, but is enough to
+\mr{In section~\ref{ssec:cuts}}, we have built a volume-limited sample from a set of magnitude
+limited ones, using a simple redshift limit. \nn{This \mr{simplistic} approach is statistically sub-optimal, but is enough to
 demonstrate our key finding: a redshift evolution of stretch compatible with
-the \cite{rigault2018} model.} However, while we will find consistent results
-when considering a sample derived from conservative cuts
-(Section~\ref{sec:results}), there remains the possibility that our sample is
-not fully volume-limited and that unaccounted for selection effects remain. To
-fully account for this possibility, a follow up analysis will make use of the
-SNANA package \citep{SNANA} to account for more subtle selection functions;
-\nn{yet, it appears to still be necessary} to test if the sample we have created
-could show sign of leftover sample selection.
-
-To do so, we compare the stretch and color distributions of the SNe~Ia in our
+the \cite{rigault2018} model.} \mr{However, there remains the possibility that \mr{complex selection function related to follow-up efficiencies below our fiducial (or even conservative cuts)} might significant affect our sample, making it not fully volume-limited, which could, in turn, bias our conclusion on the astrophysical drift of the SNe~Ia population.}
+
+\mr{To test for the existence of potentially significant leftover selection biases in our sample,} we compare the stretch and color distributions of the SNe~Ia in our
 sample between datasets of overlapping redshifts. If indeed we have built a
@@
 negligible. \yc{YC: \textit{why do you say that here, since you didn't fine-tune
-the overlaping $z$-range?}} \nn{\textit{Because we still took that into
-account?}} The two samples that overlap the most in redshift are PS1 and SDSS in
+the overlaping $z$-range?}} \nn{\textit{NN: Because we still took that into
+account?}} \mr{\textit{MR:Because if there is a strong drift in the redshift range, and if the redshift distribution of the sample differ, their respective property distribution should differ.}}
+
+The two samples that overlap the most in redshift are PS1 and SDSS in
 the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
@@
 distribution of both \nn{these surveys}, together with $p$-values of
-Kolmogorov-Smirnov tests that \yc{\sout{\nn{probe whether}}}\nn{\textit{I don't
-see anything else to put here without repeating ``test''}} the $x_1$ and $c$ of
-the SDSS and PS1 samples at $0.10 < z < 0.20$ arise from the same underlying
-distribution. These $p$-values are high ($>10\%$), and if one cannot conclude
-from this test that these SNe~Ia are from the same underlying distribution, one
-has no ground to claim they are not. \yc{YC: I suggest ``These $p$-values are
+Kolmogorov-Smirnov tests (KS-test). \yc{These $p$-values are
 too high ($>10\%$) to exclude that both samples originate from the same
-underlying population.''} This is in agreement with the visual impression from
-Fig.~\ref{fig:distrib}. We perform a similar analysis for PS1 and SNLS in the
-redshift range $0.20 < z < 0.31$ where the same conclusion can be drawn, i.e.,
-there is no sign of discrepancy in the stretch and color distributions between
-the low- and high-end of the SNLS and PS1 datasets, respectively. Nonetheless,
+underlying population,} in agreement with the visual impression from
+Fig.~\ref{fig:distrib}.
+%The $p$-values are high ($>10\%$), and if one cannot conclude from a KS-test that indeed these SNe~Ia are from the same underlying distribution, there is no ground to claim they are not.
+
+We perform a similar analysis for PS1 and SNLS in the
+redshift range $0.20 < z < 0.31$ where the same conclusion can be drawn, i.e., there is no sign of discrepancy in the stretch and color distributions between the low- and high-end of the SNLS and PS1 datasets, respectively.
+Nonetheless,
 the small size of the SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1)
@@
 
-We furthermore highlight that the SNe~Ia color is more prone to selection
+We highlight that the SNe~Ia color is more prone to selection
 effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
 Fig.~3 of \cite{kessler2017}. Hence, since the color distribution shows no
-significant hint of selection effect, it further supports our claim that our
-simple redshift-based selection criteria is sufficient to \nn{build an unbiased
-SNe~Ia sample at all redshifts required to test the redshift evolution of the
+significant hint of leftover selection effect, it further supports our claim that our simple redshift-based selection criteria is sufficient to \nn{build the SNe~Ia sample required to test the redshift evolution of the
 stretch distribution}.
@@
 
-We have presented a study of the drift of the underlying SNe~Ia stretch
+We have presented a \mr{first} study of the drift of the underlying SNe~Ia stretch
 distribution as a function of redshift. We built a magnitude-limited SN~Ia
@@
 \citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
-programs (high-$z$), respectively. \nn{To improve the disentangling of
-observational} selection effects and \nn{intrinsic} population evolution, an
-analysis using the SNANA framework is underway. It will allow for subtle
-selection effects to be \nn{accounted for}, and by removing the need for a
-\nn{naive} redshift cut, increase the redshift leverage and statistical
-significance of our results.
+programs (high-$z$), respectively.
+
+\mr{The next step of the analysis will be to incorporate our modeling within the SNANA framework \citep{SNANA} both to more accurately account for selection functions and to test the impact of our modeling on the derivation of cosmological parameters ; this study is underway.}
 

### 2020-11-24 Nicolas.tex
All ready for submission, waiting co-authors

--- 2020-11-09 Nicolas.tex
@@
     distribution for each age population made of Gaussian mixtures. We test our
-    prediction against published samples cut to have marginal magnitude
+    prediction against published samples \textbf{cut} to have marginal magnitude
     selection effects, so that any observed change is indeed of astrophysical
-    and not observational origin. \nn{In this preliminary study,} there are
-    indications that the underlying SN~Ia stretch distribution is evolving as a
+    and not observational origin. \textbf{In this first study, there are
+    indications} that the underlying SN~Ia stretch distribution is evolving as a
     function of redshift, and that the young/old drifting model is a better
@@
 negligible below a given redshift at which even the faintest SNe~Ia can be
-observed. In contrast, targeted surveys have highly complex selection functions
-and will be discarded from our analysis. Fortunately, modern SN cosmology
-samples such as the Pantheon one are now dominated by magnitude-limited surveys.
+observed. \textbf{In practice such spectroscopic following has not always been
+the norm as discussed below.} In contrast, targeted surveys have highly
+complex selection functions and will be discarded from our analysis.
+Fortunately, modern SN cosmology samples such as the Pantheon one are now
+dominated by magnitude-limited surveys.
 
@@
 
-\nn{We will thus consider a set of cuts that will define a first fiducial
-    sample, taking the farthest limits possible; but as this solution might be
-    suboptimal to create a complete sample, we also consider another set of cuts
-    to define a conservative sample, smaller and therefore less statistically
-    constraining, but inherently less prone to selection effects. If the
+\textbf{We will thus consider a set of cuts that will define a first fiducial
+sample, taking the farthest limits possible. But as this solution might be
+over simplistic to create a complete sample, e.g., for it ignores follow-up
+efficiency, we also consider another set of cuts to define a conservative
+sample. This latter is smaller and therefore less statistically
+constraining, but also inherently less prone to selection effects. If the
 redshift drift is still significant in the conservative sample, it would be
-    even more significant in a carefully-tailored selection-free sample: the
-goal of this preliminary paper is to make a first implementation of this
-modeling before refining it on a better sample.}
+even more significant in a carefully-tailored selection-free sample. The
+goal of this study is to build the first implementation of a SN Ia
+properties drift modeling before refining it on a more complex, yet more
+accurate, follow-up analysis accounting at the same time for precise selection
+modeling using of the SNANA package \citep{SNANA}.}
 
@@
 24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
-final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement
-with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. \nn{However,
-because this simple approach might easily not be enough to alleviate all
-selection effects, we also take into account another, more restrictive,
-redshift limit:} Fig.~14 of \citet[][see their Section~5]{perrett2010} suggests a
-lower limit of $z_{\lim}=0.55$; both limits will be considered, as discussed
-below.
+final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
+\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
+\citet[][see their Section~5]{perrett2010} suggests however a lower limit of
+$z_{\lim}=0.55$. \textbf{We will therefore use $z=0.60$ and $z=0.55$ as redshift
+limits for the fiducial and conservative cuts, respectively.}
 
@@
 $5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.31$ in
-agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure is also
-suggestive of a more conservative $z_{\lim}=0.27$, \nn{similarly considered in
-the following analysis as a first step to evaluate the robustness of our
-sample construction.}
+agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. \textbf{If we were to be
+conservative, this figure would also suggest of a more stringent
+$z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
+conservative cuts, respectively.}
 
@@
 \citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However,
-the SDSS surveys were more sensitive to limited spectroscopic resources;
-\citet[][see their Section~2]{kessler2009} pointed out that during the first year
-of SDSS, SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up,
-corresponding to a redshift cut at $0.15$. For the rest of the SDSS survey,
-additional spectroscopic resources were available, and \cite{kessler2009} and
-\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
-these analyses, we will use $z_{\lim}=0.20$ as the baseline SDSS redshift limit,
-\nn{and $z_{\lim}=0.15$ to compare the results of our modeling on a more
-restrictive sample.}
+the SDSS surveys were more sensitive to limited spectroscopic resources
+\textbf{as discussed before}. \citet[][Section~2]{kessler2009} pointed out that
+during the first year of SDSS, SNe~Ia with $r<20.5$~mag were favored for
+spectroscopic follow-up, corresponding to a redshift cut at $0.15$. For the rest
+of the SDSS survey, additional spectroscopic resources were available, and
+\cite{kessler2009} and \cite{dilday2008} show a reasonable completeness up to
+$z_{\lim}=0.2$. Following these analyses, \textbf{we will use $z_{\lim}=0.20$
+and $z_{\lim}=0.15$ for out fiducial and conservative cuts, respectively.}
 
@@
 distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
-expected, the selected redshift limits roughly correspond to the peak of these
-histograms. \nn{We test in Section~\ref{ssec:verify} whether we can consider
-this sample to be free of (or at least marginally affected by) selection
-effects. The consistency of the results for the two samples, discussed} in
-Section~\ref{sec:results}, further supports that this sample selection indeed
-provides a subset of SNe~Ia with insignificant selection effects when compared
-to state-of-the-art Malmquist correction techniques.
+expected, the selected redshift limits \textbf{are roughly located slightly
+before the peak of these histograms}. \textbf{We test in
+Section~\ref{ssec:verify} the validity of our intent to build volume-limited
+surveys from (mostly) magnitude-limited ones with the aforementioned redshift
+limits.}
 
@@
 and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
-\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
-range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
-much deeper. These 114 SNfactory SNe~Ia \nn{(80\% of the 141 SNe presented in
-the previously-cited paper)} are thus in the volume-limited part of the
+\citealt{aldering2020}). \textbf{For this dataset,} the spectroscopic follow-up
+was done over a redshift range of $0.02 < z < 0.09$, while the search was much
+deeper. These 114 SNfactory SNe~Ia are thus in the volume-limited part of the
 survey (Aldering et al., in prep.), and are therefore assumed to be a random
-sampling of the underlying SN population. The SNfactory sample is
-particularly useful for studying SN property drift, as it enables us to have
-a large complete SN~Ia sample at $z<0.1$.
+sampling of the underlying SN population. The SNfactory sample is particularly
+useful for studying SN property drift, as it enables us to have a large complete
+SN~Ia sample at $z<0.1$.
 
@@
 deeper than the follow-up and that we therefore kept entirely
-\citep{strolger04}.
+\citep{strolger04}. \textbf{At these redshift, the supernovae typing is
+challenging but the target classification were robust enough to include them
+within the cosmological analysis \citep{scolnic2018a}; we do not add further
+quality cuts.}
 
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \nn{Proportionally
-few low-stretch SNe (typically $x_1 < -1$) are observed at $z>0.2$}. The
-evolution of the mean stretch is also shown in Fig.~\ref{fig:modelall}, where
-the data are split in redshift bins of regular sample size. We see that SNe~Ia
-at higher redshift have on average larger stretch ($0.34 \pm 0.10$ at
-$z\sim0.65$) than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$),
-suggesting that the underlying stretch distribution is evolving with redshift.
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe
+that the apparent fraction of low-stretch SNe (typically $x_1 < -1$) seems to
+decrease as a function of redshift.} The evolution of the mean stretch is also
+shown in Fig.~\ref{fig:modelall}, where the data are split in redshift bins of
+regular sample size. We see that SNe~Ia at higher redshift have on average
+larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift
+($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch
+distribution is \textbf{evolving with redshift}.
 
@@
     histograms in dark (resp.\ light) colors for the conservative (resp.  \
-fiducial) redshift cuts. \yc{YC: you should add a running mean $\pm$ stddev to
-show the evolution at $z \sim 0.2-0.3$.}}
+fiducial) redshift cuts.}
     \label{fig:sample}
@@
 
-In this analysis, we have built a volume-limited sample from a set of magnitude
-limited ones, using a simple redshift limit. \nn{This drastic approach is robust
-at the price of being statistically sub-optimal, but is enough to
+\textbf{In section~\ref{ssec:cuts}, we have built a volume-limited sample from a
+set of magnitude-limited ones, using a simple redshift limit. This
+simplistic approach is statistically sub-optimal, but is enough to
 demonstrate our key finding: a redshift evolution of stretch compatible with
-the \cite{rigault2018} model.} However, while we will find consistent results
-when considering a sample derived from conservative cuts
-(Section~\ref{sec:results}), there remains the possibility that our sample is
-not fully volume-limited and that unaccounted for selection effects remain. To
-fully account for this possibility, a follow up analysis will make use of the
-SNANA package \citep{SNANA} to account for more subtle selection functions;
-\nn{yet, it appears to still be necessary} to test if the sample we have created
-could show sign of leftover sample selection.
-
-To do so, we compare the stretch and color distributions of the SNe~Ia in our
+the \cite{rigault2018} model. However, there remains the possibility that
+complex selection function related to follow-up efficiencies below our fiducial
+(or even conservative cuts) might significantly affect our sample, making it not
+fully volume-limited, which could, in turn, bias our conclusion on the
+astrophysical drift of the SNe~Ia population.}
+
+\textbf{To test for the existence of potentially significant leftover selection biases in our sample, we compare the stretch and color distributions of the SNe~Ia in our
 sample between datasets of overlapping redshifts. If indeed we have built a
@@
 redshift range, this has to be narrow enough such that any drift would be
-negligible. \yc{YC: \textit{why do you say that here, since you didn't fine-tune
-the overlaping $z$-range?}} \nn{\textit{Because we still took that into
-account?}} The two samples that overlap the most in redshift are PS1 and SDSS in
-the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
-corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and thus the
-most affected by potential selection effect leftover (see the corresponding
-discussion in section~\ref{ssec:cuts}). In that same redshift range, PS1 has 52
-SNe~Ia, which are in the lowest redshift bins and thus unlikely to have any
-selection issue. Fig.~\ref{fig:distrib} presents the stretch and color
-distribution of both \nn{these surveys}, together with $p$-values of
-Kolmogorov-Smirnov tests that \yc{\sout{\nn{probe whether}}}\nn{\textit{I don't
-see anything else to put here without repeating ``test''}} the $x_1$ and $c$ of
-the SDSS and PS1 samples at $0.10 < z < 0.20$ arise from the same underlying
-distribution. These $p$-values are high ($>10\%$), and if one cannot conclude
-from this test that these SNe~Ia are from the same underlying distribution, one
-has no ground to claim they are not. \yc{YC: I suggest ``These $p$-values are
-too high ($>10\%$) to exclude that both samples originate from the same
-underlying population.''} This is in agreement with the visual impression from
-Fig.~\ref{fig:distrib}. We perform a similar analysis for PS1 and SNLS in the
-redshift range $0.20 < z < 0.31$ where the same conclusion can be drawn, i.e.,
-there is no sign of discrepancy in the stretch and color distributions between
-the low- and high-end of the SNLS and PS1 datasets, respectively. Nonetheless,
-the small size of the SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1)
-limits the interest of the test as only strong deviation would be noticeable.
-Extending the redshift range to $0.20 < z < 0.40$ (though we have no PS1 data
-above 0.3) enables to get 31 SNe~Ia from SNLS in total, yet the stretch and
-color KS-test $p$-values remains inconclusive, at 34\% and 4\% respectively.
+negligible.}
+
+%\yc{YC: \textit{why do you say that here, since you didn't fine-tune
+%the overlaping $z$-range?}} \nn{\textit{NN: Because we still took that into
+%account?}} \mr{\textit{MR:Because if there is a strong drift in the redshift
+%range, and if the redshift distribution of the sample differ, their respective
+%property distribution should differ.}}
+
+\textbf{The two samples that overlap the most in redshift are PS1 and SDSS in the
+redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This corresponds to
+the 146 SNe~Ia from SDSS at the highest redshift end and thus the most affected
+by potential selection effect leftover (see the corresponding discussion in
+section~\ref{ssec:cuts}). In that same redshift range, PS1 has 52 SNe~Ia, which
+are in the lowest redshift bins and thus unlikely to have any selection issue.
+Fig.~\ref{fig:distrib} presents the stretch and color distribution of both these
+surveys, together with $p$-values of Kolmogorov-Smirnov tests (KS-test). These
+$p$-values are too high ($>10\%$) to exclude that both samples originate from
+the same underlying population, in agreement with the visual impression from
+Fig.~\ref{fig:distrib}.}
+
+%The $p$-values are high ($>10\%$), and if one cannot conclude from a KS-test
+%that indeed these SNe~Ia are from the same underlying distribution, there is no
+%ground to claim they are not.
+
+\textbf{We perform a similar analysis for PS1 and SNLS in the redshift range $0.20 < z <
+0.31$ where the same conclusion can be drawn, i.e., there is no sign of
+discrepancy in the stretch and color distributions between the low- and high-end
+of the SNLS and PS1 datasets, respectively. Nonetheless, the small size of the
+SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the interest of the
+test as only strong deviation would be noticeable.  Extending the redshift range
+to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3) enables to get 31
+SNe~Ia from SNLS in total, yet the stretch and color KS-test $p$-values remains
+inconclusive, at 34\% and 4\% respectively.}
 
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
-    \caption{$x_1$ (left) and $c$ (right) distribution histograms of different
-        surveys overlapping in redshift. \textit{Facing up}: SDSS and PS1 within
-        the redshift range $0.10 < z < 0.20$; \textit{facing down}: PS1 and SNLS
-        within the redshift range $0.20 < z < 0.31$. Error bars show the Poisson
-        noise \yc{YC: not visible enough}. Our stretch ``Base'' model is
+    \caption{\textbf{$x_1$ (left) and $c$ (right) distribution histograms of
+            different surveys overlapping in redshift. \textit{Facing up}: SDSS
+            and PS1 within the redshift range $0.10 < z < 0.20$; \textit{facing
+            down}: PS1 and SNLS within the redshift range $0.20 < z < 0.31$.
+            Error bars show the Poisson noise. Our stretch ``Base'' model is
             illustrated in orange at the mean redshift of the redshift ranges,
-        $0.15$ and $0.25$, respectively.  Kolmogorov-Smirnov test $p$-values are
-    indicated on the top (resp. bottom) of each panel showing no sign that the
-SDSS and PS1 (resp. PS1 and SNLS) $x_1$ and $c$ distributions are not drawn from
-the same underlying distributions.}
+            $0.15$ and $0.25$, respectively.  Kolmogorov-Smirnov test $p$-values
+            are indicated on the top (resp.  bottom) of each panel showing no
+            sign that the SDSS and PS1 (resp. PS1 and SNLS) $x_1$ and $c$
+            distributions are not drawn from the same underlying
+    distributions.}}
     \label{fig:distrib}
@@
 
-We furthermore highlight that the SNe~Ia color is more prone to selection
-effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
-Fig.~3 of \cite{kessler2017}. Hence, since the color distribution shows no
-significant hint of selection effect, it further supports our claim that our
-simple redshift-based selection criteria is sufficient to \nn{build an unbiased
-SNe~Ia sample at all redshifts required to test the redshift evolution of the
-stretch distribution}.
+\textbf{We highlight that the SNe~Ia color is more prone to selection effects
+than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3
+of \cite{kessler2017}. Hence, since the color distribution shows no
+significant hint of leftover selection effect, it further supports our claim
+that our simple redshift-based selection criteria is sufficient to build the
+SNe~Ia sample required to test the redshift evolution of the stretch
+distribution.}
 
@@
 
-We report in Table~\ref{tab:bbc} \nn{our determination of} the samples' $\mu_0$
-and $\sigma_{\pm}$ \nn{when implementing an asymmetric Gaussian model}, adjusted
-on the nominally selection-free samples using our fiducial cuts (see
-Section~\ref{sec:sample}). We find our results in close agreement with
+We report in Table~\ref{tab:bbc} \textbf{our determination of} the samples'
+$\mu_0$ and $\sigma_{\pm}$ \textbf{when implementing an asymmetric Gaussian
+model}, adjusted on the nominally selection-free samples using our fiducial cuts
+(see Section~\ref{sec:sample}). We find our results in close agreement with
 \cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
@@
 corresponding fit is not significantly better, with a $\Delta$AIC of $-0.4$.
-this strengthens the fact that the young and old populations indeed appear to
+This strengthens the fact that the young and old populations indeed appear to
 share the same underlying high-stretch mode. Furthermore, one might wonder
@@
 
-We have presented a study of the drift of the underlying SNe~Ia stretch
-distribution as a function of redshift. We built a magnitude-limited SN~Ia
-sample from the Pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to
-which we added HST and SNfactory data from \cite{rigault2018} for the high- and
-low-redshift bins. We only considered the SNe that have been discovered in the
-redshift range of each survey where selection effects are negligible, so that
-the observed SNe~Ia stretches are random sampling of the true underlying
-distribution. This resulted in a 569 SN~Ia fiducial sample (422 SNe when more
-conservative cuts were considered).
+We have presented a \textbf{first} study of the drift of the underlying SNe~Ia
+stretch distribution as a function of redshift. We built a magnitude-limited
+SN~Ia sample from the Pantheon dataset \citep[][SDSS, PS1 and
+SNLS]{scolnic2018a}, to which we added HST and SNfactory data from
+\cite{rigault2018} for the high- and low-redshift bins. We only considered the
+SNe that have been discovered in the redshift range of each survey where
+selection effects are negligible, so that the observed SNe~Ia stretches are
+random sampling of the true underlying distribution. This resulted in a 569
+SN~Ia fiducial sample (422 SNe when more conservative cuts were considered).
 
@@
     \item The underlying SN~Ia stretch distribution is significantly redshift
-        dependent, as previously suggested by e.g.~\cite{howell2007} \nn{in a
-        way that selection effects alone cannot explain}. This result is largely
-        independent of details on each age-population modeling.
+        dependent, as previously suggested by e.g.~\cite{howell2007} \textbf{in
+        a way that selection effects alone cannot explain}. This result is
+        largely independent of details on each age-population modeling.
 
@@
 
-In this paper, we considered a simple two-population Gaussian mixture modeling,
-but additional data free from significant Malmquist bias would enable us to
-refine it as necessary. We note that samples at the low- and high-redshift ends
-of the Hubble diagram would be particularly helpful for this drifting analysis;
-fortunately this will soon be provided by the Zwicky Transient Facility
-\citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
-programs (high-$z$), respectively. \nn{To improve the disentangling of
-observational} selection effects and \nn{intrinsic} population evolution, an
-analysis using the SNANA framework is underway. It will allow for subtle
-selection effects to be \nn{accounted for}, and by removing the need for a
-\nn{naive} redshift cut, increase the redshift leverage and statistical
-significance of our results.
+In this paper, we considered a simple \textbf{two-population} Gaussian mixture
+modeling, but additional data free from significant Malmquist bias would enable
+us to refine it as necessary. We note that samples at the low- and high-redshift
+ends of the Hubble diagram would be particularly helpful for this drifting
+analysis; fortunately this will soon be provided by the Zwicky Transient
+Facility \citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange
+SNe~Ia programs (high-$z$), respectively.
+
+\textbf{The next step of the analysis will be to incorporate our modeling within
+the SNANA framework \citep{SNANA} both to more accurately account for selection
+functions and to test the impact of our modeling on the derivation of
+cosmological parameters; this study is underway.}
 
@@
 
-\yc{YC: check again all arXiv references.}
-
 \bibliographystyle{aa}
@@
 \bibitem[Freedman et al.(2020)]{freedman2020} Freedman, W.~L., Madore, B.~F.,
-Hoyt, T., et al.\ 2020, arXiv e-prints, arXiv:2002.01550
+Hoyt, T., et al.\ 2020, \apj, 891, 57. doi:10.3847/1538-4357/ab7339
 
@@
 
-\bibitem[Knox \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv
-e-prints, arXiv:1908.03663
+\bibitem[Knox \& Millea(2020)]{knox2019} Knox, L. \& Millea, M.\ 2020, \prd,
+101, 043533. doi:10.1103/PhysRevD.101.043533
 
@@
 
-\bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration,
-Aghanim, N., Akrami, Y., et al.\ 2018, arXiv e-prints, arXiv:1807.06209
+\bibitem[Planck Collaboration et al.(2020)]{planck2018} Planck
+Collaboration, Aghanim, N., Akrami, Y., et al.\ 2020, \aap, 641, A6.
+doi:10.1051/0004-6361/201833910
 
@@
 \bibitem[Reid et al.(2019)]{reid2019} Reid, M.~J., Pesce, D.~W., \& Riess,
-A.~G.\ 2019, arXiv e-prints, arXiv:1908.05625
+A.~G.\ 2019, \apjl, 886, L27. doi:10.3847/2041-8213/ab552d
 
@@
 \bibitem[Wiseman et al.(2020)]{wiseman2020} Wiseman, P., Smith, M., Childress,
-M., et al.\ 2020, arXiv e-prints, arXiv:2001.02640
-
-\bibitem[Wong et al.(2019)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F.,
-et al.\ 2019, arXiv e-prints, arXiv:1907.04869
+M., et al.\ 2020, \mnras, 495, 4040. doi:10.1093/mnras/staa1302
+
+\bibitem[Wong et al.(2020)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F.,
+et al.\ 2020, \mnras, 498, 1420. doi:10.1093/mnras/stz3094
 

### 2020-12-16 Nicolas.tex
Greg comments included and treated.

@@
 %\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
-\newcommand{\nn}[1]{\textcolor[rgb]{0.4, 0.8, 0}{#1}}
+\newcommand{\nn}[1]{\textcolor[rgb]{1.0, 0.6, 0.0}{#1}}
 \newcommand{\mr}[1]{\textcolor[rgb]{0.5, 0.20, 0.3}{#1}}
@@
 \author{
-    N. Nicolas \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution} \inst{1} 
-    \and M. Rigault \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
+    N. Nicolas \thanks{n.nicolas@ip2i.in2p3.fr, equal contribution} \inst{1} 
+    \and M. Rigault \thanks{m.rigault@ip2i.in2p3.fr, equal contribution} \inst{2}
     \orcid{0000-0002-8121-2560}
@@
 observed. \textbf{In practice such spectroscopic following has not always been
-the norm as discussed below.} In contrast, targeted surveys have highly
-complex selection functions and will be discarded from our analysis.
-Fortunately, modern SN cosmology samples such as the Pantheon one are now
-dominated by magnitude-limited surveys.
+the norm as discussed below.} In contrast, targeted surveys have highly complex
+selection functions and will be discarded from our analysis.  Modern SN
+cosmology samples such as the Pantheon one are now dominated by
+magnitude-limited surveys \nn{which have easier to model selection functions}.
 
@@
 accurate, follow-up analysis accounting at the same time for precise selection
-modeling using of the SNANA package \citep{SNANA}.}
+modeling using, e.g., the SNANA package \citep{SNANA}.}
 
@@
 Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
-$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.31$ in
+$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields $z_{\lim}=0.31$, in
 agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. \textbf{If we were to be
@@
 In a similar redshift range, SDSS has a limiting magnitude of 22.5
-\citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However,
-the SDSS surveys were more sensitive to limited spectroscopic resources
-\textbf{as discussed before}. \citet[][Section~2]{kessler2009} pointed out that
-during the first year of SDSS, SNe~Ia with $r<20.5$~mag were favored for
-spectroscopic follow-up, corresponding to a redshift cut at $0.15$. For the rest
-of the SDSS survey, additional spectroscopic resources were available, and
-\cite{kessler2009} and \cite{dilday2008} show a reasonable completeness up to
-$z_{\lim}=0.2$. Following these analyses, \textbf{we will use $z_{\lim}=0.20$
-and $z_{\lim}=0.15$ for out fiducial and conservative cuts, respectively.}
+\citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
+SDSS surveys \nn{had to contend with limited spectroscopic resources. As
+discussed in} \citet[][Section~2]{kessler2009}, during the first year of SDSS,
+SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
+to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
+spectroscopic resources were available, and \cite{kessler2009} and
+\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
+these analyses, \textbf{we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
+fiducial and conservative cuts, respectively.}
 
@@
 expected, the selected redshift limits \textbf{are roughly located slightly
-before the peak of these histograms}. \textbf{We test in
-Section~\ref{ssec:verify} the validity of our intent to build volume-limited
-surveys from (mostly) magnitude-limited ones with the aforementioned redshift
-limits.}
+    before the peak of these histograms}. \nn{In Section 2.2 we validate that
+    these redshift limits are effective for constructing nearly volume-limited
+subsamples from samples that were initially more-closely magnitude limited in
+their search or spectroscopic follow-up.}
 
@@
     \caption{Composition of the SNe~Ia dataset used in this analysis.
-    Conservative cuts are indicated in parentheses.}
+    Conservative cuts are indicated in parentheses. \nn{The SNf limit is set by
+\cite{rigault2018}, see text.}}
     \label{tab:sample}
@@
         \hline
-        SNf & -- & 114 \\
+        SNf & 0.08 & 114 \\
         SDSS & 0.20 (0.15) & 167 (82)\\
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
-    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia from
-        the SDSS, PS1 and SNLS dataset respectively (data from Pantheon,
-        \citealt{scolnic2018a}). The colored parts represent the distribution
-        of SNe~Ia kept in our analysis for they are supposedly free from
-        selection bias (see Section~\ref{sec:sample}). The darker (resp. lighter)
-        color responds to the conservative (resp. fiducial) selection cut.}
+    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia from the
+        SDSS, PS1 and SNLS dataset respectively (data from Pantheon,
+        \citealt{scolnic2018a}). The colored parts represent the distribution of
+        SNe~Ia kept in our analysis for they are supposedly free from selection
+    bias (see Section~\ref{sec:sample}). The darker (resp. lighter) color
+responds to the conservative (resp. fiducial) selection cut.}
     \label{fig:cuts}
@@
 and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
-\citealt{aldering2020}). \textbf{For this dataset,} the spectroscopic follow-up
-was done over a redshift range of $0.02 < z < 0.09$, while the search was much
-deeper. These 114 SNfactory SNe~Ia are thus in the volume-limited part of the
-survey (Aldering et al., in prep.), and are therefore assumed to be a random
+\citealt{aldering2020}). \nn{For this dataset, spectroscopic screening was done
+for candidates with $r \lesssim 19.5$. Then, redshift cuts were applied when
+selecting which SN Ia to follow, giving a redshift range of 0.02 < z < 0.09 ;
+further reduced to $<0.08$ in \cite{rigault2018} for extracting local host
+properties.} These 114 SNfactory SNe~Ia are thus in the volume-limited part of
+the survey (Aldering et al., in prep.), and are therefore assumed to be a random
 sampling of the underlying SN population. The SNfactory sample is particularly
@@
 
-\textbf{To test for the existence of potentially significant leftover selection biases in our sample, we compare the stretch and color distributions of the SNe~Ia in our
-sample between datasets of overlapping redshifts. If indeed we have built a
-volume-limited sample, the distributions of any SNe~Ia parameter at a given
-redshift should not depend of their survey of origin. When considering a
-redshift range, this has to be narrow enough such that any drift would be
-negligible.}
-
-%\yc{YC: \textit{why do you say that here, since you didn't fine-tune
-%the overlaping $z$-range?}} \nn{\textit{NN: Because we still took that into
-%account?}} \mr{\textit{MR:Because if there is a strong drift in the redshift
-%range, and if the redshift distribution of the sample differ, their respective
-%property distribution should differ.}}
-
-\textbf{The two samples that overlap the most in redshift are PS1 and SDSS in the
-redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This corresponds to
-the 146 SNe~Ia from SDSS at the highest redshift end and thus the most affected
-by potential selection effect leftover (see the corresponding discussion in
-section~\ref{ssec:cuts}). In that same redshift range, PS1 has 52 SNe~Ia, which
-are in the lowest redshift bins and thus unlikely to have any selection issue.
-Fig.~\ref{fig:distrib} presents the stretch and color distribution of both these
-surveys, together with $p$-values of Kolmogorov-Smirnov tests (KS-test). These
-$p$-values are too high ($>10\%$) to exclude that both samples originate from
-the same underlying population, in agreement with the visual impression from
-Fig.~\ref{fig:distrib}.}
-
-%The $p$-values are high ($>10\%$), and if one cannot conclude from a KS-test
-%that indeed these SNe~Ia are from the same underlying distribution, there is no
-%ground to claim they are not.
-
-\textbf{We perform a similar analysis for PS1 and SNLS in the redshift range $0.20 < z <
-0.31$ where the same conclusion can be drawn, i.e., there is no sign of
-discrepancy in the stretch and color distributions between the low- and high-end
-of the SNLS and PS1 datasets, respectively. Nonetheless, the small size of the
-SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the interest of the
-test as only strong deviation would be noticeable.  Extending the redshift range
-to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3) enables to get 31
-SNe~Ia from SNLS in total, yet the stretch and color KS-test $p$-values remains
-inconclusive, at 34\% and 4\% respectively.}
+\textbf{To test for the existence of potentially significant leftover selection
+biases in our sample, we compare the stretch and color distributions of the
+SNe~Ia in our sample between datasets of overlapping redshifts. If indeed we
+have built a volume-limited sample, the distributions of any SNe~Ia
+parameter at a given redshift should not depend on their survey of origin. When
+considering a redshift range, this has to be narrow enough such that any drift
+would be negligible.}
+
+\textbf{The two samples that overlap the most in redshift are PS1 and SDSS in
+the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
+corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and thus
+the most \nn{likely to be} affected by selection effects \nn{not eliminated
+by our redshift cut} (see the corresponding discussion in
+section~\ref{ssec:cuts}). In that same redshift range, PS1 has 52 SNe~Ia,
+which are in the lowest redshift bins and thus unlikely to have any
+selection issue.  \nn{To identify potential inconsistency between PS1 and
+SDSS, Fig.~\ref{fig:distrib} presents} the stretch and color
+distribution of both these surveys, together with $p$-values of
+Kolmogorov-Smirnov (KS) tests. \nn{These $p$-values are too high
+($>10\%$) and do not indicate that these subsamples are inconsistent}, in
+agreement with the visual impression from Fig.~\ref{fig:distrib}.}
+
+\textbf{We perform a similar analysis for PS1 and SNLS in the redshift range
+$0.20 < z < 0.31$ where the same conclusion can be drawn, i.e., there is no
+sign of discrepancy in the stretch and color distributions between the low-
+and high-end of the SNLS and PS1 datasets, respectively. Nonetheless, the
+small size of the SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1)
+limits the \nn{sensitivity} of the test, as only strong deviation would be
+noticeable. Extending the redshift range to $0.20 < z < 0.40$ (though we
+have no PS1 data above 0.3) enables to get 31 SNe~Ia from SNLS in total, yet
+\nn{the stretch p-values remains high at 34\% showing no sign of inconsistency}}
 
@@
 \end{equation}
-with $K=0.87$. This model is comparable to the evolution predicted by
-\cite{childress2014} based on SN rates in galaxies depending on their quenching
-time as a function of their stellar mass.
+with $K=0.87$. This model is comparable to the evolution \nn{subsequently}
+predicted by \cite{childress2014} based on SN rates in galaxies depending on
+their quenching time as a function of their stellar mass.
 
@@
 \end{align}
+
+\nn{This is our Base model.}
 
@@
 analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted as a
-probability $p=2\times 10^{-7}$ of being an as good representation of the data
+probability $p=2\times 10^{-7}$ of being as good a representation of the data
 as the Base model.
 
-Remark that, when comparing models adjusted on individual subsamples rather than
-globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
+\nn{Note} that, when comparing models adjusted on individual subsamples rather
+than globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
 k\ln(n)$, with $n$ the number of data points) might be better suited than AIC,
@@
 separately: the sample-based model BIC is rightfully the sum of the BIC for each
-sample. Doing so, we find $\Delta\mathrm{BIC}=-48$, again refuting the
-sample-based asymmetric Gaussian model as being as pertinent as the Base model.
+sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the sample-based
+asymmetric Gaussian model as being as pertinent as the Base model.
 
@@
 Gaussians on the supposedly selection-free part of the samples and the results
-derived using the BBC formalism supports our approach to get a sample with
-negligible selection effects. If we were to use \cite{scolnic2016} and
+derived using the BBC formalism supports our approach to \nn{construct} a sample
+with negligible selection effects. If we were to use \cite{scolnic2016} and
 \cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
@@
 corresponding fit is not significantly better, with a $\Delta$AIC of $-0.4$.
-This strengthens the fact that the young and old populations indeed appear to
-share the same underlying high-stretch mode. Furthermore, one might wonder
-whether a low-stretch mode might also exist in the young-population, see
-Fig.~\ref{fig:stretchlssfr}. We tested that by allowing this population to also
-be bi-modal, finding the amplitude of this low-stretch mode to be compatible
-with~0 in this young population ($<2\%$). More generally, this raises the
-question of inaccurate tracing of age by a given environmental tracer (here the
-LsSFR). A dedicated analysis will be presented in Briday et al.\ in prep.
+This \nn{reinforces our assumption} that the young and old populations indeed
+appear to share the same underlying high-stretch mode. Furthermore, one might
+wonder whether a low-stretch mode might also exist in the young-population, see
+Fig.~\ref{fig:stretchlssfr}. We \nn{test for this} by allowing this population
+to also be bi-modal, finding the amplitude of \nn{such a} low-stretch mode to be
+compatible with~0 in this young population ($<2\%$). More generally, this raises
+the question of\nn{how well a given environmental tracer (here LsSFR) traces the
+age}.  A dedicated analysis will be presented in Briday et al.\ in prep.
 
@@
 \cite{rubin2015}. Not doing so is a second order issue for SN cosmology, as it
-only affects the way one accounts for Malmquist bias. Indeed, as long as
-Phillips' relation \citep{phillips1993} standardization parameter $\alpha$ is
-not redshift dependent (a study behind the scope of this paper, but see e.g.
+only affects the way one accounts for Malmquist bias. Indeed, as long as the
+Phillips relation \citep{phillips1993} standardization parameter $\alpha$ is not
+redshift dependent (a study \nn{beyond} the scope of this paper, but see e.g.
 \citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
 cosmology are blind to the underlying stretch distribution for complete samples.
-However, surveys usually do have significant Malmquist bias at least for the
-upper half of their SN redshift distribution. As a consequence, an ill-modeling
-of the underlying stretch distribution will bias the SN magnitude derived from
-such surveys.
+However, surveys usually do have significant Malmquist bias for the upper half
+of their SN redshift distribution. As a consequence, an ill-modeling of the
+underlying stretch distribution will bias the SN magnitudes derived from such
+surveys.
 
@@
 for implicit redshift drifts, a direct modeling of the redshift drift is more
-appropriate than a sample-based approach. We stress here that, as measurements
+appropriate than a sample-based approach. We \nn{add} here that, as measurements
 of modern surveys try to cover increasingly larger redshift ranges in order to
-reduce calibration systematic uncertainties, this sample-based approach is
-becoming less valid, notably for PS1, DES and, soon, LSST.
+reduce calibration systematic uncertainties, this sample-based approach
+\nn{becomes} less valid, notably for PS1, DES and, soon, LSST.
 
@@
 Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges covered by PS1. The observed
-$x_1$ histogram follows our modeling defined as the sum of individual underlying
-SN-redshift distributions. As expected, the two modeling approaches differ
-mostly in the negative part of the SN stretch distribution. The asymmetric
-Gaussian distribution goes through the middle of the bimodal distribution,
-over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and under-estimating it at
-$x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN
-redshifts. This means that the SN bias-corrected standardized magnitude
-estimated at a redshift affected by selection effect would be biased by an
-ill-modeling of the true underlying stretch distribution.
+$x_1$ histogram follows \nn{the model we defined using} the sum of individual
+underlying SN-redshift distributions. As expected, the two modeling approaches
+differ mostly in the negative part of the SN stretch distribution. The
+asymmetric Gaussian distribution goes through the middle of the bimodal
+distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
+under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting model
+for typical PS1 SN redshifts. This means that the SN bias-corrected standardized
+magnitude estimated at a redshift affected by selection effects would be biased
+by an ill-modeling of the true underlying stretch distribution.
 
@@
 
-The amplitude of this magnitude bias for cosmology is beyond the scope of this
-paper given the complexity of the BBC analysis. It would require a full study
-using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
-asymmetric modeling as part of the BBC simulations. However, we already
-highlight that even if a non-drifting sample-based model could provide
+\nn{Assessing t}he amplitude of this magnitude bias for cosmology is beyond the
+scope of this paper given the complexity of the BBC analysis. It would require a
+full study using our Base model (Eq.~\ref{eq:stretchz}) in place of the
+sample-based asymmetric modeling as part of the BBC simulations. However, we
+already highlight that even if a non-drifting sample-based model could provide
 comparable result in the volume-limited part of the various samples, these
-models would differ when extrapolating at higher redshifts, precisely where the
+models would differ when extrapolating to higher redshifts, precisely where the
 underlying distribution will matter for correcting Malmquist biases.
 
-In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
-level and $w_a$ at the ten-percent precision \citep[e.g.,][]{lsstpaper}, we
-stress that correct modeling of potential SN redshift drift should be further
-studied and care should be taken when using samples affected by selection
-effects.
+In the era of modern cosmology, where we aim \nn{to measure} $w_0$ at a
+sub-percent level and $w_a$ \nn{with} ten-percent precision
+\citep[e.g.,][]{lsstpaper}, we stress that correct modeling of potential SN
+redshift drift should be further studied and care should be taken when using
+samples affected by selection effects.
 
@@
 
-We have presented a \textbf{first} study of the drift of the underlying SNe~Ia
+We have presented \nn{an initial} study of the drift of the underlying SNe~Ia
 stretch distribution as a function of redshift. We built a magnitude-limited
@@
 selection effects are negligible, so that the observed SNe~Ia stretches are
-random sampling of the true underlying distribution. This resulted in a 569
+a random sampling of the true underlying distribution. This resulted in a 569
 SN~Ia fiducial sample (422 SNe when more conservative cuts were considered).
@@
 model which depends on the expected fraction of ``young'' and ``old'' SNe~Ia as
-a function of redshift, each age population having its own underlying
+a function of redshift, \nn{with} each age population having its own underlying
 stretch distribution.
@@
     \item The underlying SN~Ia stretch distribution is significantly redshift
-        dependent, as previously suggested by e.g.~\cite{howell2007} \textbf{in
+        dependent, as previously suggested by e.g.~\cite{howell2007}, \textbf{in
         a way that selection effects alone cannot explain}. This result is
-        largely independent of details on each age-population modeling.
-
-    \item Redshift-independent models are indeed excluded as suitable
-        descriptions of the data relative to our Base model. This model assumes
-        that: (1) the younger population has a unimodal Gaussian stretch
+        largely independent of details on each age-population model.
+
+    \item Redshift-independent models are \nn{quantitatively} excluded as
+        suitable descriptions of the data relative to our Base model. This model
+        assumes that: (1) the younger population has a unimodal Gaussian stretch
         distribution, while the older population stretch distribution is
@@
     \item Models using survey-based asymmetric Gaussian distributions, as done,
-        e.g., in the current implementation of the BBC, are excluded to be good
-        descriptions of the data relative to our drifting model. Hence, the
+        e.g., in the current implementation of BBC, are excluded \nn{as a} good
+        description of the data relative to our drifting model. Hence, the
         sample-based approach does not accurately account for redshift drift and
-        even less so as survey span increasingly larger redshift ranges. We
-        stipulate that, even if extra degrees of freedom might be acceptable
-        given the large number of SNe~Ia in cosmological studies, extrapolating
-        the SN property distributions from the volume-limited part of a survey
-        to its Malmquist-biased magnitude-limited one would still be inaccurate
-        because of the redshift evolution.
+        even less so as survey span increasingly larger redshift ranges.  \nn{As
+        a result, if} extra degrees of freedom might be acceptable given the
+        large number of SNe~Ia in cosmological studies, extrapolating the SN
+        property distributions from the volume-limited part of a survey to its
+        Malmquist-biased magnitude-limited one would still be inaccurate because
+        of the redshift evolution.
 

### 2020-12-17 Nicolas.tex
Changed fig. 6 to mean + caption, changed affiliation

@@
     N. Nicolas \thanks{n.nicolas@ip2i.in2p3.fr, equal contribution} \inst{1} 
-    \and M. Rigault \thanks{m.rigault@ip2i.in2p3.fr, equal contribution} \inst{2}
+    \and M. Rigault \thanks{m.rigault@ip2i.in2p3.fr, equal contribution} \inst{1}
     \orcid{0000-0002-8121-2560}
@@
     \and Saul Perlmutter\inst{3}
-    \and M. Smith\inst{2,5}
+    \and M. Smith\inst{1,5}
     \orcid{0000-0002-3321-1432}
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
+    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf-mean.pdf}
     \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
@@
         redshift bins of equal sample size, indicated in light gray at the
-        bottom of each redshift bin. Full and light markers are used when
-        considering the fiducial or the conservative samples, respectively. The
-        orange horizontal line represents the mean stretch of the fiducial
-        sample, illustrating the expectation if the SN stretch distribution is
-        not evolving with redshift. Best fits of our Base drifting model are
-        shown as blue, dashed-blue and gray, when fitted on the fiducial sample,
-        the conservative one or the SNfactory dataset only, respectively; all
-        are compatible. The light-blue band illustrates the amplitude of the
-        error (incl. covariance) of the best fit model when considering the
-        fiducial dataset.}
+        bottom of each redshift bin. \nn{Individual $x_1$ errors, subdominant cause
+        of the total scatter, have not been accounted for when measuring the
+    $x_1$ means.} Full and light markers are used when considering the
+        fiducial or the conservative samples, respectively. The orange
+        horizontal line represents the mean stretch of the fiducial sample,
+        illustrating the expectation if the SN stretch distribution is not
+        evolving with redshift. Best fits of our Base drifting model are shown
+        as blue, dashed-blue and gray, when fitted on the fiducial sample, the
+        conservative one or the SNfactory dataset only, respectively; all are
+        compatible. The light-blue band illustrates the amplitude of the error
+    (incl. covariance) of the best fit model when considering the fiducial
+dataset.}
     \label{fig:modelall}

### 2021-01-06 Nicolas.tex
Changes following Greg discussion, hoping for last round of comments

--- 2020-12-16 Nicolas.tex
@@
 
-%\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
 \newcommand{\nn}[1]{\textcolor[rgb]{1.0, 0.6, 0.0}{#1}}
@@
     redshift, using the evolution of the fraction of young and old SNe~Ia as
-    predicted by \cite{rigault2018}, and assuming constant underlying stretch
+    predicted by \cite{rigault2020}, and assuming constant underlying stretch
     distribution for each age population made of Gaussian mixtures. We test our
@@
     bias. The favored underlying stretch model is the bimodal one derived from
-    \cite{rigault2018}: a high-stretch mode shared by both young and old
+    \cite{rigault2020}: a high-stretch mode shared by both young and old
     environments, and a low-stretch mode exclusive to old environments. The
@@
 brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
-betoule2014, rigault2018, kim19}. This mass-step correction is currently used in
+betoule2014, rigault2020, kim19}. This mass-step correction is currently used in
 cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including for
@@
 (LsSFR) or the local color are more strongly correlated with the standardized SN
-magnitude \citep{rigault2018, roman2018, kim18}, suggesting age as the driving
+magnitude \citep{rigault2020, roman2018, kim18}, suggesting age as the driving
 parameter underlying the mass-step. If true, this would have a significant
@@
 following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
-model). In \cite{rigault2018} we used the LsSFR to classify which are the
+model). In \cite{rigault2020} we used the LsSFR to classify which are the
 younger (those with a high LsSFR) and which are the older (those with low
@@
 SN population, which we claim to be constituted of two age-populations
-\citep{rigault2013,rigault2015,rigault2018}: one old and one younger, the former
+\citep{rigault2013,rigault2015,rigault2020}: one old and one younger, the former
 having on average lower lightcurve stretches and being brighter after
@@
 fraction of each of these populations as a function of redshift follows the
-model presented in \cite{rigault2018} and (3) the underlying distribution of
+model presented in \cite{rigault2020} and (3) the underlying distribution of
 stretch for each age sample is constant. This paper aims at testing this
@@
 cosmology samples such as the Pantheon one are now dominated by
-magnitude-limited surveys \nn{which have easier to model selection functions}.
+magnitude-limited surveys \textbf{which have easier to model selection functions}.
 
@@
 \citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
-SDSS surveys \nn{had to contend with limited spectroscopic resources. As
+SDSS surveys \textbf{had to contend with limited spectroscopic resources. As
 discussed in} \citet[][Section~2]{kessler2009}, during the first year of SDSS,
@@
 expected, the selected redshift limits \textbf{are roughly located slightly
-    before the peak of these histograms}. \nn{In Section 2.2 we validate that
-    these redshift limits are effective for constructing nearly volume-limited
-subsamples from samples that were initially more-closely magnitude limited in
-their search or spectroscopic follow-up.}
+before the peak of these histograms. In Section 2.2 we validate that these
+redshift limits are effective for constructing nearly volume-limited subsamples
+from samples that were initially more closely magnitude-limited in their search
+or spectroscopic follow-up.}
 
@@
     \caption{Composition of the SNe~Ia dataset used in this analysis.
-    Conservative cuts are indicated in parentheses. \nn{The SNf limit is set by
-\cite{rigault2018}, see text.}}
+    Conservative cuts are indicated in parentheses. \textbf{The SNf limit is set by
+\cite{rigault2020}, see text.}}
     \label{tab:sample}
@@
 In addition, we use the SNe~Ia from the Nearby Supernova Factory
-\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
+\citep[SNfactory,][]{aldering2002} published in \cite{rigault2020} and that have
 been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
 and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
-\citealt{aldering2020}). \nn{For this dataset, spectroscopic screening was done
-for candidates with $r \lesssim 19.5$. Then, redshift cuts were applied when
-selecting which SN Ia to follow, giving a redshift range of 0.02 < z < 0.09 ;
-further reduced to $<0.08$ in \cite{rigault2018} for extracting local host
-properties.} These 114 SNfactory SNe~Ia are thus in the volume-limited part of
-the survey (Aldering et al., in prep.), and are therefore assumed to be a random
-sampling of the underlying SN population. The SNfactory sample is particularly
-useful for studying SN property drift, as it enables us to have a large complete
-SN~Ia sample at $z<0.1$.
-
-Finally, we include the HST sample from Pantheon, that similarly have a search
-deeper than the follow-up and that we therefore kept entirely
-\citep{strolger04}. \textbf{At these redshift, the supernovae typing is
-challenging but the target classification were robust enough to include them
-within the cosmological analysis \citep{scolnic2018a}; we do not add further
-quality cuts.}
+\citealt{aldering2020}). \textbf{For this dataset, spectroscopic screening was
+    done for candidates with $r \lesssim 19.5$. Then, redshift cuts were applied
+    when selecting which SN Ia to follow, giving a redshift range of 0.02 < z <
+    0.09 ; further reduced to $<0.08$ in \cite{rigault2020} for extracting local
+host properties.} These 114 SNfactory SNe~Ia are thus in the volume-limited part
+of the survey (Aldering et al., in prep.), and are therefore assumed to be a
+random sampling of the underlying SN population. The SNfactory sample is
+particularly useful for studying SN property drift, as it enables us to have a
+large complete SN~Ia sample at $z<0.1$.
+
+Finally, we include the HST sample from Pantheon \citep{strolger04}.
+\textbf{These high-redshift SNe are of great interest as they provide the
+    greatest leverage for testing evolution. While at these redshift the
+    supernovae typing is challenging, the target classification was robust
+enough to include them within the cosmological analysis \citep{scolnic2018a} and
+we do not add further quality cuts. Section~\ref{sec:results} highlights that
+our results are not dependent on the inclusion of this dataset.}
 
@@
 demonstrate our key finding: a redshift evolution of stretch compatible with
-the \cite{rigault2018} model. However, there remains the possibility that
+the \cite{rigault2020} model. However, there remains the possibility that
 complex selection function related to follow-up efficiencies below our fiducial
@@
 corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and thus
-the most \nn{likely to be} affected by selection effects \nn{not eliminated
-by our redshift cut} (see the corresponding discussion in
-section~\ref{ssec:cuts}). In that same redshift range, PS1 has 52 SNe~Ia,
-which are in the lowest redshift bins and thus unlikely to have any
-selection issue.  \nn{To identify potential inconsistency between PS1 and
-SDSS, Fig.~\ref{fig:distrib} presents} the stretch and color
-distribution of both these surveys, together with $p$-values of
-Kolmogorov-Smirnov (KS) tests. \nn{These $p$-values are too high
-($>10\%$) and do not indicate that these subsamples are inconsistent}, in
-agreement with the visual impression from Fig.~\ref{fig:distrib}.}
+the most likely to be affected by selection effects not eliminated by our
+redshift cut (see the corresponding discussion in section~\ref{ssec:cuts}).
+In that same redshift range, PS1 has 52 SNe~Ia, which are in the lowest
+redshift bins and thus unlikely to have any selection issue. To identify
+potential inconsistency between PS1 and SDSS, Fig.~\ref{fig:distrib}
+presents the stretch and color distribution of both these surveys, together
+with $p$-values of Kolmogorov-Smirnov (KS) tests. These $p$-values are too
+high ($>10\%$) and do not indicate that these subsamples are inconsistent,
+in agreement with the visual impression from Fig.~\ref{fig:distrib}.}
 
@@
 small size of the SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1)
-limits the \nn{sensitivity} of the test, as only strong deviation would be
+limits the sensitivity of the test, as only strong deviation would be
 noticeable. Extending the redshift range to $0.20 < z < 0.40$ (though we
 have no PS1 data above 0.3) enables to get 31 SNe~Ia from SNLS in total, yet
-\nn{the stretch p-values remains high at 34\% showing no sign of inconsistency}}
+the stretch p-values remains high at 34\% showing no sign of inconsistency}
 
@@
 
-\cite{rigault2018} presented a model for the evolution of the fraction of
+\cite{rigault2020} presented a model for the evolution of the fraction of
 younger and older SNe~Ia as a function of redshift following former work on
@@
 \citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
-\citep{rigault2013, rigault2018, wiseman2020}, in agreement with rate
-expectations \citep{mannucci2006, rodney2014}, \cite{rigault2018} concluded that
+\citep{rigault2013, rigault2020, wiseman2020}, in agreement with rate
+expectations \citep{mannucci2006, rodney2014}, \cite{rigault2020} concluded that
 
@@
 \end{equation}
-with $K=0.87$. This model is comparable to the evolution \nn{subsequently}
+with $K=0.87$. This model is comparable to the evolution \textbf{subsequently}
 predicted by \cite{childress2014} based on SN rates in galaxies depending on
@@
 
-\cite{rigault2018} presented the relation between SN stretch and LsSFR
+\cite{rigault2020} presented the relation between SN stretch and LsSFR
 measurement, a progenitor age tracer, using the SNfactory sample. This relation
@@
         probability, $p_y$, for the SNe~Ia to be young, i.e. to have
-        $\log\mathrm{LsSFR} \geq -10.82$ \citep[see][]{rigault2018}.
+        $\log\mathrm{LsSFR} \geq -10.82$ \citep[see][]{rigault2020}.
         \textit{Right}: $p_y$-weighted histogram of the SN stretches, as well as
@@
 
-\nn{This is our Base model.}
+\textbf{This is our Base model.}
 
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
+    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf-mean.pdf}
     \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
-        as a function of redshift. Markers show the mean stretch measured in
-        redshift bins of equal sample size, indicated in light gray at the
-        bottom of each redshift bin. Full and light markers are used when
-        considering the fiducial or the conservative samples, respectively. The
-        orange horizontal line represents the mean stretch of the fiducial
-        sample, illustrating the expectation if the SN stretch distribution is
-        not evolving with redshift. Best fits of our Base drifting model are
-        shown as blue, dashed-blue and gray, when fitted on the fiducial sample,
-        the conservative one or the SNfactory dataset only, respectively; all
-        are compatible. The light-blue band illustrates the amplitude of the
-        error (incl. covariance) of the best fit model when considering the
-        fiducial dataset.}
+        as a function of redshift. Markers show the \textbf{stretch plain mean
+        (with error estimated from scatter)} measured in redshift bins of equal
+        sample size, indicated in light gray at the bottom of each redshift bin.
+        Full and light markers are used when considering the fiducial or the
+        conservative samples, respectively. The orange horizontal line
+        represents the mean stretch \textbf{for the non-evolving Gaussian model
+        (last line of Table~\ref{tab:comp}) on the fiducial sample}. Best
+        fits of our Base drifting model are shown as blue, dashed-blue and
+        gray, when fitted on the fiducial sample, the conservative one or
+        the SNfactory dataset only, respectively; all are compatible. The
+        light-blue band illustrates the amplitude of the error (incl.
+        covariance) of the best fit model when considering the fiducial
+    dataset.}
     \label{fig:modelall}
@@
 In Section~\ref{sec:basemodel}, we have modeled the underlying stretch
-distribution following \cite{rigault2018}, i.e.\ as a single Gaussian for the
+distribution following \cite{rigault2020}, i.e.\ as a single Gaussian for the
 ``young'' SNe~Ia and a mixture of two Gaussians for the ``old'' SNe Ia
@@
 
-\nn{Note} that, when comparing models adjusted on individual subsamples rather
-than globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
-k\ln(n)$, with $n$ the number of data points) might be better suited than AIC,
-since it explicitly accounts for the fact that each subsample is fitted
-separately: the sample-based model BIC is rightfully the sum of the BIC for each
-sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the sample-based
-asymmetric Gaussian model as being as pertinent as the Base model.
+\textbf{Note} that, when comparing models adjusted on individual subsamples
+rather than globally, the Bayesian Information Criterion ($\mathrm{BIC} =
+-2\ln(L) + k\ln(n)$, with $n$ the number of data points) might be better suited
+than AIC, since it explicitly accounts for the fact that each subsample is
+fitted separately: the sample-based model BIC is rightfully the sum of the BIC
+for each sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the
+sample-based asymmetric Gaussian model as being as pertinent as the Base model.
+
+\textbf{In order to ensure that our results are not driven by the poorly-modeled
+    HST sub-sample, we recompute $\Delta$AIC for each model excluding this
+    dataset; we find (as listed in Table~\ref{tab:comp}) -3.4, -24.5, -27.2,
+    -29.2, -26.9, -60.0 and -5.4, -17.4, -20.4, -22.4, -19.7, -57.9 for the
+    fiducial and conservative samples, respectively. The consistency of these
+    values with those in Table~\ref{tab:comp} show that the HST sub-sample does
+not drive our conclusions.}
 
@@
 Gaussians on the supposedly selection-free part of the samples and the results
-derived using the BBC formalism supports our approach to \nn{construct} a sample
-with negligible selection effects. If we were to use \cite{scolnic2016} and
-\cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
+derived using the BBC formalism supports our approach to \textbf{construct} a
+sample with negligible selection effects. If we were to use \cite{scolnic2016}
+and \cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
 parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
@@
 corresponding fit is not significantly better, with a $\Delta$AIC of $-0.4$.
-This \nn{reinforces our assumption} that the young and old populations indeed
-appear to share the same underlying high-stretch mode. Furthermore, one might
-wonder whether a low-stretch mode might also exist in the young-population, see
-Fig.~\ref{fig:stretchlssfr}. We \nn{test for this} by allowing this population
-to also be bi-modal, finding the amplitude of \nn{such a} low-stretch mode to be
-compatible with~0 in this young population ($<2\%$). More generally, this raises
-the question of\nn{how well a given environmental tracer (here LsSFR) traces the
-age}.  A dedicated analysis will be presented in Briday et al.\ in prep.
+This \textbf{reinforces our assumption} that the young and old populations
+indeed appear to share the same underlying high-stretch mode. Furthermore, one
+might wonder whether a low-stretch mode might also exist in the
+young-population, see Fig.~\ref{fig:stretchlssfr}. We \textbf{test for this} by
+allowing this population to also be bi-modal, finding the amplitude of
+\textbf{such a} low-stretch mode to be compatible with~0 in this young
+population ($<2\%$). More generally, this raises the question of \textbf{how
+well a given environmental tracer (here LsSFR) traces the age}. A dedicated
+analysis will be presented in Briday et al.\ in prep.
 
@@
 Phillips relation \citep{phillips1993} standardization parameter $\alpha$ is not
-redshift dependent (a study \nn{beyond} the scope of this paper, but see e.g.
-\citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
+redshift dependent (a study \textbf{beyond} the scope of this paper, but see
+e.g. \citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
 cosmology are blind to the underlying stretch distribution for complete samples.
@@
 for implicit redshift drifts, a direct modeling of the redshift drift is more
-appropriate than a sample-based approach. We \nn{add} here that, as measurements
-of modern surveys try to cover increasingly larger redshift ranges in order to
-reduce calibration systematic uncertainties, this sample-based approach
-\nn{becomes} less valid, notably for PS1, DES and, soon, LSST.
+appropriate than a sample-based approach. We \textbf{add} here that, as
+measurements of modern surveys try to cover increasingly larger redshift ranges
+in order to reduce calibration systematic uncertainties, this sample-based
+approach \textbf{becomes} less valid, notably for PS1, DES and, soon, LSST.
 
@@
 Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges covered by PS1. The observed
-$x_1$ histogram follows \nn{the model we defined using} the sum of individual
-underlying SN-redshift distributions. As expected, the two modeling approaches
-differ mostly in the negative part of the SN stretch distribution. The
-asymmetric Gaussian distribution goes through the middle of the bimodal
+$x_1$ histogram follows \textbf{the model we defined using} the sum of
+individual underlying SN-redshift distributions. As expected, the two modeling
+approaches differ mostly in the negative part of the SN stretch distribution.
+The asymmetric Gaussian distribution goes through the middle of the bimodal
 distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
@@
 
-\nn{Assessing t}he amplitude of this magnitude bias for cosmology is beyond the
-scope of this paper given the complexity of the BBC analysis. It would require a
-full study using our Base model (Eq.~\ref{eq:stretchz}) in place of the
-sample-based asymmetric modeling as part of the BBC simulations. However, we
+\textbf{Assessing t}he amplitude of this magnitude bias for cosmology is beyond
+the scope of this paper given the complexity of the BBC analysis. It would
+require a full study using our Base model (Eq.~\ref{eq:stretchz}) in place of
+the sample-based asymmetric modeling as part of the BBC simulations. However, we
 already highlight that even if a non-drifting sample-based model could provide
@@
 
-In the era of modern cosmology, where we aim \nn{to measure} $w_0$ at a
-sub-percent level and $w_a$ \nn{with} ten-percent precision
+In the era of modern cosmology, where we aim \textbf{to measure} $w_0$ at a
+sub-percent level and $w_a$ \textbf{with} ten-percent precision
 \citep[e.g.,][]{lsstpaper}, we stress that correct modeling of potential SN
@@
 
-We have presented \nn{an initial} study of the drift of the underlying SNe~Ia
+We have presented \textbf{an initial} study of the drift of the underlying SNe~Ia
 stretch distribution as a function of redshift. We built a magnitude-limited
@@
 SNLS]{scolnic2018a}, to which we added HST and SNfactory data from
-\cite{rigault2018} for the high- and low-redshift bins. We only considered the
+\cite{rigault2020} for the high- and low-redshift bins. We only considered the
 SNe that have been discovered in the redshift range of each survey where
@@
 
-Following predictions made in \cite{rigault2018}, we introduced a redshift drift
+Following predictions made in \cite{rigault2020}, we introduced a redshift drift
 model which depends on the expected fraction of ``young'' and ``old'' SNe~Ia as
-a function of redshift, \nn{with} each age population having its own underlying
-stretch distribution.
+a function of redshift, \textbf{with} each age population having its own
+underlying stretch distribution.
 
@@
 
-    \item Redshift-independent models are \nn{quantitatively} excluded as
+    \item Redshift-independent models are \textbf{quantitatively} excluded as
         suitable descriptions of the data relative to our Base model. This model
@@
         the relative fraction of younger and older SNe~Ia follows the prediction
-        made in \cite{rigault2018}. This second result strongly supports the
+        made in \cite{rigault2020}. This second result strongly supports the
         existence of both young and old SN~Ia populations, in agreement with
@@
     \item Models using survey-based asymmetric Gaussian distributions, as done,
-        e.g., in the current implementation of BBC, are excluded \nn{as a} good
-        description of the data relative to our drifting model. Hence, the
+        e.g., in the current implementation of BBC, are excluded \textbf{as a}
+        good description of the data relative to our drifting model. Hence, the
         sample-based approach does not accurately account for redshift drift and
-        even less so as survey span increasingly larger redshift ranges.  \nn{As
-        a result, if} extra degrees of freedom might be acceptable given the
-        large number of SNe~Ia in cosmological studies, extrapolating the SN
-        property distributions from the volume-limited part of a survey to its
-        Malmquist-biased magnitude-limited one would still be inaccurate because
-        of the redshift evolution.
+        even less so as survey span increasingly larger redshift ranges.
+        \textbf{As a result, if} extra degrees of freedom might be acceptable
+        given the large number of SNe~Ia in cosmological studies, extrapolating
+        the SN property distributions from the volume-limited part of a survey
+        to its Malmquist-biased magnitude-limited one would still be inaccurate
+        because of the redshift evolution.
 
@@
 
-\bibitem[Rigault et al.(2018)]{rigault2018} Rigault, M., Brinnel, V., Aldering,
-G., et al.\ 2018, arXiv:1806.03849
+\bibitem[Rigault et al.(2020)]{rigault2020} Rigault, M., Brinnel, V., Aldering,
+G., et al.\ 2029, \aap, 644, A176
 

### 2021-01-06 Nicolas.tex
YC comments

@@
 \textbf{We will thus consider a set of cuts that will define a first fiducial
-sample, taking the farthest limits possible. But as this solution might be
+sample, taking the farthest limits possible.
+\yc{YC: This sounds too strong and ill-defined.
+I suggest ``taking limits as initially suggested by completeness analysis''.} \yc{However,} as this solution might be
 over simplistic to create a complete sample, e.g., for it ignores follow-up
-efficiency, we also consider another set of cuts to define a conservative
+efficiency, we also consider another set of cuts to define a \yc{so-called} conservative
 sample. This latter is smaller and therefore less statistically
@@
 redshift drift is still significant in the conservative sample, it would be
-even more significant in a carefully-tailored selection-free sample. The
-goal of this study is to build the first implementation of a SN Ia
-properties drift modeling before refining it on a more complex, yet more
-accurate, follow-up analysis accounting at the same time for precise selection
+even more significant \yc{meaningful? (to avoid repetition)} in a carefully-tailored selection-free sample. The
+goal of this study is to build \yc{test?} the first implementation of a SN Ia
+properties drift modeling \yc{YC: modeling or model?} before refining it on a more complex, yet more
+accurate, follow-up analysis accounting at the same time \yc{YC: not clear, "same time" as what?} for precise selection
 modeling using, e.g., the SNANA package \citep{SNANA}.}
@@
 \citealt{aldering2020}). \textbf{For this dataset, spectroscopic screening was
-    done for candidates with $r \lesssim 19.5$. Then, redshift cuts were applied
-    when selecting which SN Ia to follow, giving a redshift range of 0.02 < z <
-    0.09 ; further reduced to $<0.08$ in \cite{rigault2020} for extracting local
+    done for candidates with $r \lesssim 19.5$; redshift cuts were \yc{then} applied
+    when selecting which SN Ia to follow, \yc{resulting in} a redshift range of $0.02 < z <
+    0.09$, further reduced to $<0.08$ in \cite{rigault2020} for extracting local
 host properties.} These 114 SNfactory SNe~Ia are thus in the volume-limited part
@@
 \textbf{These high-redshift SNe are of great interest as they provide the
-    greatest leverage for testing evolution. While at these redshift the
+    greatest leverage for testing evolution. While at these redshifts the
     supernovae typing is challenging, the target classification was robust
 enough to include them within the cosmological analysis \citep{scolnic2018a} and
-we do not add further quality cuts. Section~\ref{sec:results} highlights that
+we do not add further quality cuts. Section~\ref{sec:results} highlights that, \yc{while compatible with it,}
 our results are not dependent on the inclusion of this dataset.}
@@
 up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe
-that the apparent fraction of low-stretch SNe (typically $x_1 < -1$) seems to
+that the apparent \yc{YC: why "apparent" here?
+It's supposed to be intrinsic now that we're working on a complete sample.
+Too much "observe/apparent/seem".} fraction of low-stretch SNe (typically $x_1 < -1$) seems to
 decrease as a function of redshift.} The evolution of the mean stretch is also
@@
 
-\textbf{In section~\ref{ssec:cuts}, we have built a volume-limited sample from a
-set of magnitude-limited ones, using a simple redshift limit. This
-simplistic approach is statistically sub-optimal, but is enough to
+\textbf{In section~\ref{ssec:cuts}, we have built volume-limited sample\yc{s} from a
+set of magnitude-limited ones, using simple redshift \yc{cuts}. This
+simplistic approach is statistically sub-optimal, but \yc{should be} enough to
 demonstrate our key finding: a redshift evolution of stretch compatible with
-the \cite{rigault2020} model. However, there remains the possibility that
+\yc{\sout{the}} \cite{rigault2020} model. However, \yc{the possibility remains} that \yc{a}
 complex selection function related to follow-up efficiencies below our fiducial
-(or even conservative cuts) might significantly affect our sample, making it not
-fully volume-limited, which could, in turn, bias our conclusion on the
-astrophysical drift of the SNe~Ia population.}
-
-\textbf{To test for the existence of potentially significant leftover selection
+(or even conservative) cuts \yc{could still}  affect our sample, making it not
+fully volume-limited; \yc{this would}, in turn, bias our conclusion on the
+astrophysical drift of the SNe~Ia population. \yc{We now look at this possibility.}}
+
+\textbf{To test for the existence of \yc{potential} leftover selection
 biases in our sample, we compare the stretch and color distributions of the
-SNe~Ia in our sample between datasets of overlapping redshifts. If indeed we
-have built a volume-limited sample, the distributions of any SNe~Ia
-parameter at a given redshift should not depend on their survey of origin. When
-considering a redshift range, this has to be narrow enough such that any drift
+SNe~Ia \yc{\sout{in our sample} originating from different} datasets \yc{over} overlapping redshift \yc{ranges}: \yc{these distributions should be similar.
+Note that the} redshift range has to be narrow enough so that any drift
 would be negligible.}
@@
 the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
-corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and thus
-the most likely to be affected by selection effects not eliminated by our
-redshift cut (see the corresponding discussion in section~\ref{ssec:cuts}).
-In that same redshift range, PS1 has 52 SNe~Ia, which are in the lowest
+corresponds to the 146 SNe~Ia from SDSS at the \yc{high} redshift end and thus
+the most likely to be affected by \yc{residual} selection effects (see the corresponding discussion in section~\ref{ssec:cuts}).
+\yc{Over} that same redshift range, PS1 has 52 SNe~Ia, which are in the lowest
 redshift bins and thus unlikely to have any selection issue. To identify
-potential inconsistency between PS1 and SDSS, Fig.~\ref{fig:distrib}
-presents the stretch and color distribution of both these surveys, together
-with $p$-values of Kolmogorov-Smirnov (KS) tests. These $p$-values are too
-high ($>10\%$) and do not indicate that these subsamples are inconsistent,
+potential inconsistency between PS1 and SDSS \yc{subsamples}, Fig.~\ref{fig:distrib}
+\yc{(upper panels) compares} the stretch and color distribution of both these surveys. \yc{The associated Kolmogorov-Smirnov (KS) similarity test $p$-values} are too
+high ($>10\%$) \yc{to conclude on any inconsistency},
 in agreement with the visual impression from Fig.~\ref{fig:distrib}.}
 
-\textbf{We perform a similar analysis for PS1 and SNLS in the redshift range
-$0.20 < z < 0.31$ where the same conclusion can be drawn, i.e., there is no
+\textbf{We perform a similar analysis for PS1 and SNLS \yc{over} the redshift range
+$0.20 < z < 0.31$ \yc{(Fig.~\ref{fig:distrib}, lower panels)}, where the same conclusion can be drawn: there is no \yc{substantial}
 sign of discrepancy in the stretch and color distributions between the low-
@@
 small size of the SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1)
-limits the sensitivity of the test, as only strong deviation would be
+limits the sensitivity of the test, \yc{and} only strong deviation would be
 noticeable. Extending the redshift range to $0.20 < z < 0.40$ (though we
-have no PS1 data above 0.3) enables to get 31 SNe~Ia from SNLS in total, yet
-the stretch p-values remains high at 34\% showing no sign of inconsistency}
+have no PS1 data above 0.3) \yc{allows to increase the SNLS subsample to 31}, yet
+the stretch $p$-values remains high (34\%) showing no sign of inconsistency.}
 
@@
             distributions are not drawn from the same underlying
-    distributions.}}
+    distributions. \yc{YC: I still urge you to make errorbars and text more visible!}}}
     \label{fig:distrib}
@@
 
-\textbf{We highlight that the SNe~Ia color is more prone to selection effects
+\textbf{We \yc{finally} highlight that the SNe~Ia color is more prone to selection effects
 than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3
-of \cite{kessler2017}. Hence, since the color distribution shows no
+of \cite{kessler2017}. Hence, since the \yc{comparison of} color distributions shows no
 significant hint of leftover selection effect, it further supports our claim
-that our simple redshift-based selection criteria is sufficient to build the
+that our simple redshift-based selection criteria is sufficient to build \yc{a complete}
 SNe~Ia sample required to test the redshift evolution of the stretch
@@
 
-\subsection{Base model applied to data}\label{sec:basemodelapplied}
+\subsection{\yc{Comparison} to data}\label{sec:basemodelapplied}
 
@@
     -29.2, -26.9, -60.0 and -5.4, -17.4, -20.4, -22.4, -19.7, -57.9 for the
-    fiducial and conservative samples, respectively. The consistency of these
+    fiducial and conservative samples, respectively.
+\yc{YC: IMHO, it would be better to add these values in 2 new dedicated columns.
+Or are these precise values really needed?
+Just say something like "does not change $\Delta$AIC by more than few tenths".} The consistency of these
     values with those in Table~\ref{tab:comp} show that the HST sub-sample does

### 2021-01-12 Nicolas.tex
YC comments included

--- 2021-01-06 Nicolas.tex
@@
     predicted by \cite{rigault2020}, and assuming constant underlying stretch
-    distribution for each age population made of Gaussian mixtures. We test our
-    prediction against published samples \textbf{cut} to have marginal magnitude
-    selection effects, so that any observed change is indeed of astrophysical
-    and not observational origin. \textbf{In this first study, there are
-    indications} that the underlying SN~Ia stretch distribution is evolving as a
-    function of redshift, and that the young/old drifting model is a better
-    description of the data than any time-constant model, including the
-    sample-based asymmetric distributions usually used to correct Malmquist
+    distribution for each age population \textbf{consisting} of Gaussian
+    mixtures. We test our prediction against published samples \textbf{cut} to
+    have marginal magnitude selection effects, so that any observed change is
+    indeed of astrophysical and not observational origin. \textbf{In this first
+    study, there are indications} that the underlying SN~Ia stretch distribution
+    is evolving as a function of redshift, and that the young/old drifting model
+    is a better description of the data than any time-constant model, including
+    the sample-based asymmetric distributions usually used to correct Malmquist
     bias. The favored underlying stretch model is the bimodal one derived from
-    \cite{rigault2020}: a high-stretch mode shared by both young and old
-    environments, and a low-stretch mode exclusive to old environments. The
-    precise impact of the redshift evolution of the SN Ia population intrinsic
-    properties on cosmology remains to be studied. Yet, the astrophysical drift
-    of the SN stretch distribution does affect current Malmquist bias
-    corrections and thereby distances derived from SN affected by selection
-effects. We highlight that such a bias will increase with surveys covering
-increasingly larger redshift ranges, which is particularly important for LSST.}
+    \cite{rigault2020}, \textbf{composed of} a high-stretch mode shared by both
+    young and old environments, and a low-stretch mode exclusive to old
+    environments. The precise impact of the redshift evolution of the SN Ia
+    population intrinsic properties on cosmology remains to be studied. Yet, the
+    astrophysical drift of the SN stretch distribution does affect current
+    Malmquist bias corrections and thereby distances derived from SN affected by
+selection effects. We highlight that such a bias will increase with surveys
+covering increasingly larger redshift ranges, which is particularly important
+for LSST.}
 
@@
 the norm as discussed below.} In contrast, targeted surveys have highly complex
-selection functions and will be discarded from our analysis.  Modern SN
-cosmology samples such as the Pantheon one are now dominated by
-magnitude-limited surveys \textbf{which have easier to model selection functions}.
+selection functions and will be discarded from our analysis. Modern SN cosmology
+samples such as the Pantheon one are now dominated by magnitude-limited surveys
+\textbf{which have easier to model selection functions}.
 
@@
 \textbf{We will thus consider a set of cuts that will define a first fiducial
-sample, taking the farthest limits possible. But as this solution might be
-over simplistic to create a complete sample, e.g., for it ignores follow-up
-efficiency, we also consider another set of cuts to define a conservative
-sample. This latter is smaller and therefore less statistically
+    sample, taking the limits as initially suggested by the previous
+    completeness analysis. \textbf{However,} as this solution might be over
+    simplistic to create a complete sample, e.g., for it ignores follow-up
+    efficiency, we also consider another set of cuts to define a so-called
+    conservative sample. This latter is smaller and therefore less statistically
     constraining, but also inherently less prone to selection effects. If the
     redshift drift is still significant in the conservative sample, it would be
-even more significant in a carefully-tailored selection-free sample. The
-goal of this study is to build the first implementation of a SN Ia
-properties drift modeling before refining it on a more complex, yet more
-accurate, follow-up analysis accounting at the same time for precise selection
-modeling using, e.g., the SNANA package \citep{SNANA}.}
+    even more meaningful in a carefully-tailored selection-free sample. The goal
+    of this study is to build the first implementation of a SN Ia properties
+    drift modeling before refining it on a more complex, yet more accurate,
+follow-up analysis also accounting for precise selection modeling using, e.g.,
+the SNANA package \citep{SNANA}.}
 
@@
 \citealt{aldering2020}). \textbf{For this dataset, spectroscopic screening was
-    done for candidates with $r \lesssim 19.5$. Then, redshift cuts were applied
-    when selecting which SN Ia to follow, giving a redshift range of 0.02 < z <
-    0.09 ; further reduced to $<0.08$ in \cite{rigault2020} for extracting local
-host properties.} These 114 SNfactory SNe~Ia are thus in the volume-limited part
-of the survey (Aldering et al., in prep.), and are therefore assumed to be a
-random sampling of the underlying SN population. The SNfactory sample is
-particularly useful for studying SN property drift, as it enables us to have a
-large complete SN~Ia sample at $z<0.1$.
+    done for candidates with $r \lesssim 19.5$; redshift cuts were then applied
+    when selecting which SN Ia to follow, resulting in a redshift range of $0.02
+    < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} for extracting
+local host properties.} These 114 SNfactory SNe~Ia are thus in the
+volume-limited part of the survey (Aldering et al., in prep.), and are therefore
+assumed to be a random sampling of the underlying SN population. The SNfactory
+sample is particularly useful for studying SN property drift, as it enables us
+to have a large complete SN~Ia sample at $z<0.1$.
 
@@
 \textbf{These high-redshift SNe are of great interest as they provide the
-    greatest leverage for testing evolution. While at these redshift the
+    greatest leverage for testing evolution. While at these redshifts the
     supernovae typing is challenging, the target classification was robust
-enough to include them within the cosmological analysis \citep{scolnic2018a} and
-we do not add further quality cuts. Section~\ref{sec:results} highlights that
-our results are not dependent on the inclusion of this dataset.}
+    enough to include them within the cosmological analysis \citep{scolnic2018a}
+    and we do not add further quality cuts. Section~\ref{sec:results} highlights
+    that, while compatible with it, our results are not dependent on the
+    inclusion of this dataset.}
 
@@
 up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe
-that the apparent fraction of low-stretch SNe (typically $x_1 < -1$) seems to
-decrease as a function of redshift.} The evolution of the mean stretch is also
-shown in Fig.~\ref{fig:modelall}, where the data are split in redshift bins of
-regular sample size. We see that SNe~Ia at higher redshift have on average
-larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift
-($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch
-distribution is \textbf{evolving with redshift}.
+that the fraction of low-stretch SNe (typically $x_1 < -1$) seems to decrease as
+a function of redshift; this is confirmed in Fig.~\ref{fig:modelall} in which}
+the evolution of the mean stretch is shown, \textbf{with} the data split in
+redshift bins of regular sample size. We see that SNe~Ia at higher redshift have
+on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
+redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
+stretch distribution is \textbf{evolving with redshift}.
 
@@
 
-\textbf{In section~\ref{ssec:cuts}, we have built a volume-limited sample from a
-set of magnitude-limited ones, using a simple redshift limit. This
-simplistic approach is statistically sub-optimal, but is enough to
-demonstrate our key finding: a redshift evolution of stretch compatible with
-the \cite{rigault2020} model. However, there remains the possibility that
-complex selection function related to follow-up efficiencies below our fiducial
-(or even conservative cuts) might significantly affect our sample, making it not
-fully volume-limited, which could, in turn, bias our conclusion on the
-astrophysical drift of the SNe~Ia population.}
-
-\textbf{To test for the existence of potentially significant leftover selection
-biases in our sample, we compare the stretch and color distributions of the
-SNe~Ia in our sample between datasets of overlapping redshifts. If indeed we
-have built a volume-limited sample, the distributions of any SNe~Ia
-parameter at a given redshift should not depend on their survey of origin. When
-considering a redshift range, this has to be narrow enough such that any drift
-would be negligible.}
+\textbf{In section~\ref{ssec:cuts}, we have built volume-limited samples from a
+    set of magnitude-limited ones, using simple redshift cuts. This simplistic
+    approach is statistically sub-optimal, but should be enough to demonstrate
+    our key finding: a redshift evolution of stretch compatible with
+    \cite{rigault2020} model. However, the possibility remains that a complex
+    selection function related to follow-up efficiencies below our fiducial (or
+    even conservative) cuts could still affect our sample, making it not fully
+volume-limited; this would, in turn, bias our conclusion on the astrophysical
+drift of the SNe~Ia population. We now look at this possibility.}
+
+\textbf{To test for the existence of potential leftover selection biases in our
+    sample, we compare the stretch and color distributions of the SNe~Ia
+    originating from different datasets over overlapping redshift ranges: these
+    distributions should be similar. Note that the redshift range has to be
+    narrow enough so that any drift would be negligible.}
 
@@
     the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
-corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and thus
-the most likely to be affected by selection effects not eliminated by our
-redshift cut (see the corresponding discussion in section~\ref{ssec:cuts}).
-In that same redshift range, PS1 has 52 SNe~Ia, which are in the lowest
-redshift bins and thus unlikely to have any selection issue. To identify
-potential inconsistency between PS1 and SDSS, Fig.~\ref{fig:distrib}
-presents the stretch and color distribution of both these surveys, together
-with $p$-values of Kolmogorov-Smirnov (KS) tests. These $p$-values are too
-high ($>10\%$) and do not indicate that these subsamples are inconsistent,
-in agreement with the visual impression from Fig.~\ref{fig:distrib}.}
-
-\textbf{We perform a similar analysis for PS1 and SNLS in the redshift range
-$0.20 < z < 0.31$ where the same conclusion can be drawn, i.e., there is no
-sign of discrepancy in the stretch and color distributions between the low-
-and high-end of the SNLS and PS1 datasets, respectively. Nonetheless, the
-small size of the SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1)
-limits the sensitivity of the test, as only strong deviation would be
-noticeable. Extending the redshift range to $0.20 < z < 0.40$ (though we
-have no PS1 data above 0.3) enables to get 31 SNe~Ia from SNLS in total, yet
-the stretch p-values remains high at 34\% showing no sign of inconsistency}
+    corresponds to the 146 SNe~Ia from SDSS at the high redshift end and thus
+    the most likely to be affected by residual selection effects (see the
+    corresponding discussion in section~\ref{ssec:cuts}). Over that same
+    redshift range, PS1 has 52 SNe~Ia, which are in the lowest redshift bins and
+    thus unlikely to have any selection issue. To identify potential
+    inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib}
+    (upper panels) compares the stretch and color distribution of both these
+    surveys. The associated Kolmogorov-Smirnov (KS) similarity test $p$-values
+are too high ($>10\%$) to conclude on any inconsistency, in agreement with the
+visual impression from Fig.~\ref{fig:distrib}.}
+
+\textbf{We perform a similar analysis for PS1 and SNLS over the redshift range
+    $0.20 < z < 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same
+    conclusion can be drawn: there is no substantial sign of discrepancy in the
+    stretch and color distributions between the low- and high-end of the SNLS
+    and PS1 datasets, respectively. Nonetheless, the small size of the SNLS
+    dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of
+    the test, and only strong deviation would be noticeable. Extending the
+redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
+allows to increase the SNLS subsample to 31, yet the stretch $p$-values remains
+high (34\%) showing no sign of inconsistency.}
 
@@
             distributions are not drawn from the same underlying
-    distributions.}}
+    distributions}}
     \label{fig:distrib}
@@
 
-\textbf{We highlight that the SNe~Ia color is more prone to selection effects
-than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3
-of \cite{kessler2017}. Hence, since the color distribution shows no
-significant hint of leftover selection effect, it further supports our claim
-that our simple redshift-based selection criteria is sufficient to build the
-SNe~Ia sample required to test the redshift evolution of the stretch
-distribution.}
+\textbf{We finally highlight that the SNe~Ia color is more prone to selection
+    effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
+    Fig.~3 of \cite{kessler2017}. Hence, since the comparison of color
+    distributions shows no significant hint of leftover selection effect, it
+    further supports our claim that our simple redshift-based selection criteria
+    is sufficient to build a complete SNe~Ia sample required to test the
+    redshift evolution of the stretch distribution.}
 
@@
 
-\subsection{Base model applied to data}\label{sec:basemodelapplied}
+\subsection{Comparison to data}\label{sec:basemodelapplied}
 
@@
     HST sub-sample, we recompute $\Delta$AIC for each model excluding this
-    dataset; we find (as listed in Table~\ref{tab:comp}) -3.4, -24.5, -27.2,
-    -29.2, -26.9, -60.0 and -5.4, -17.4, -20.4, -22.4, -19.7, -57.9 for the
-    fiducial and conservative samples, respectively. The consistency of these
-    values with those in Table~\ref{tab:comp} show that the HST sub-sample does
-not drive our conclusions.}
+    dataset; we find that it does not change $\Delta$AIC by more than few
+    tenths. The consistency of these values with those in Table~\ref{tab:comp}
+show that the HST sub-sample does not drive our conclusions.}
 

### 2021-01-12 Nicolas.tex
Greg changes, diff to see them

@@
 
-\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
-    unknown, and as survey statistics increase, the question of astrophysical
-    systematic uncertainties rises, notably that of the SN~Ia population
+\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
+    and as survey statistics increase, the question of astrophysical
+    systematic uncertainties rises, notably that of SN~Ia population
     evolution. In this paper, we study the dependence with redshift of the SN~Ia
@@
     have marginal magnitude selection effects, so that any observed change is
-    indeed of astrophysical and not observational origin. \textbf{In this first
+    indeed of astrophysical and not observational in origin. \textbf{In this first
     study, there are indications} that the underlying SN~Ia stretch distribution
@@
     is a better description of the data than any time-constant model, including
-    the sample-based asymmetric distributions usually used to correct Malmquist
+    the sample-based asymmetric distributions often used to correct Malmquist
     bias. The favored underlying stretch model is the bimodal one derived from
@@
     Malmquist bias corrections and thereby distances derived from SN affected by
-selection effects. We highlight that such a bias will increase with surveys
+observational selection effects. We highlight that such a bias will increase with surveys
 covering increasingly larger redshift ranges, which is particularly important
@@
 Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
-have enabled the discovery of the acceleration of the Universe's expansion
+enabled the discovery of the acceleration of the Universe's expansion
 \citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
@@
 \citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ derived when
-the SNe~Ia are anchored on Cepheids \citep[the SH0ES project,][]{riess2009,
+the SNe~Ia are anchored to Cepheids \citep[the SH0ES project,][]{riess2009,
 riess2016} is $\sim5\sigma$ higher than what is predicted from cosmic microwave
@@
 \cite{poulin2019}. Alternatively, systematic effects affecting one or several
-of the aforementioned analyses could also explain the tension.
+of the aforementioned analyses might explain at least some of this tension.
 \cite{rigault2015} suggested that SNe Ia from the Cepheid calibrator sample
-differ by construction from the Hubble flow sample ones as the former
-strongly favors young stellar populations, while the latter not. This
+differ by construction from those in the Hubble flow sample, as the former
+strongly favors young stellar populations, while the latter does not. This
 selection effect would impact the derivation of $H_0$ if SNe~Ia from young and
@@
 
-For the last decades, numerous analyses have studied the relation between SNe~Ia
-and host properties, finding first that the standardized SNe~Ia magnitude
-significantly depends on the host stellar mass, SNe~Ia from high-mass host being
+%For the last decades, numerous analyses have studied 
+The relation between SNe~Ia
+and their host galaxy properties has been studied extensively. The first key finding
+was that that the standardized SNe~Ia magnitudes
+significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass host 
+galaxies being
 brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
@@
 deriving $H_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
-between the SNe and their host remains unclear when using global properties such
-as the host stellar mass, which raises the question of the accuracy of the
-correction. More recently, studies have used the local SN environment to probe
-more direct connections between the SN and its environments \citep{rigault2013},
+between the SNe and their host galaxies remains unclear, particularly when using global properties such
+as the host stellar mass. This in turn raises the question of the accuracy of such
+corrections since host galaxy global properties evolve with redshift. 
+More recently, studies have used the local SN host galaxy environment to probe
+more direct connections between the SNe and their host galaxy environments \citep{rigault2013},
 showing that local age tracers such as the Local specific Star Formation Rate
 (LsSFR) or the local color are more strongly correlated with the standardized SN
-magnitude \citep{rigault2020, roman2018, kim18}, suggesting age as the driving
+magnitude \citep{rigault2020, roman2018, kim18}. Further reinforcing this
+connection is the identification of SN~Ia spectral features that are 
+correlated with LsSFR \citep{nordin2018}. These results suggest age as the driving
 parameter underlying the mass-step. If true, this would have a significant
@@
 LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN stretch has
-been known to be strongly correlated with the SN host properties
-\citep{hamuy1996, hamuy2000}, correlation that has been extensively confirmed
+been known to be strongly correlated with the SN host galaxy properties
+\citep{hamuy1996, hamuy2000}. This correlation that has been extensively confirmed
 since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
 gupta2011, dandrea2011, childress2013, rigault2013, pan2014, kim19}. Following
-the ``A+B'' model and the connection between SN stretch and host properties,
+the ``A+B'' model and the connection between SN stretch and host galaxy properties,
 \cite{howell2007} first discussed the potential redshift drift of the SN stretch
-distribution. In this paper we revisit this analysis with the most recent SNe~Ia
-dataset.
+distribution. In this paper we revisit this question using the most recent SNe~Ia
+datasets.
 
@@
 stretch for each age sample is constant. This paper aims at testing this
-specific model with datasets from the literature. Note that the progenitor age
-as traced by the LsSFR seems to capture physical feature intrinsic to the
-progenitor and/or explosion mechanism that the stretch alone is not capturing
-\citep{nordin2018}.
+specific model with datasets from the literature. 
+% Note that the progenitor age
+% as traced by the LsSFR seems to capture physical feature intrinsic to the
+%progenitor and/or explosion mechanism that the stretch alone is not capturing
+% \citep{nordin2018}.
 
@@
 \section{Complete Sample Construction}\label{sec:sample}
+
+The ideal SN~Ia sample for studying this question would be a very deep, large-area,
+volume-limited sample. This would capture the true underlying stretch distribution
+function, and we would then study how it evolves with redshift. No such sample exists, 
+so first we must construct subsamples that are as near to volume-limited as possible 
+from existing high-redshift SN~Ia samples.
+
 \subsection{Applying redshift cuts}\label{ssec:cuts}
@@
 distributions in a few bins of redshift. In practice, however, differential
-selection effects are affecting the observed SN stretch distributions. Indeed,
+observational selection effects will affect the observed SN stretch distributions. Indeed,
 because the observed SN~Ia magnitude correlates with the lightcurve stretch (and
 color), the first SNe~Ia that a magnitude-limited survey will miss are the
-lowest-stretch (and reddest) ones. Consequently, if magnitude-related selection
+lowest-stretch (and reddest) ones. Consequently, if magnitude-related observational selection
 effects are not accounted for, one might confuse true population drift with
@@
 
-Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring typing
-and host redshift, the selection effects of magnitude-limited surveys should be
-negligible below a given redshift at which even the faintest SNe~Ia can be
-observed. \textbf{In practice such spectroscopic following has not always been
-the norm as discussed below.} In contrast, targeted surveys have highly complex
-selection functions and will be discarded from our analysis. Modern SN cosmology
-samples such as the Pantheon one are now dominated by magnitude-limited surveys
-\textbf{which have easier to model selection functions}.
+Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring type SN types
+and host galaxy redshifts, the observation selection effects of magnitude-limited surveys should be
+negligible below a given redshift at which even the faintest normal SNe~Ia can be
+observed. \textbf{Aiding in the construction of nearly volume-limited 
+subsamples is the fact that the SN~Ia population trails off towards fainter SNe~Ia.
+A complication is that complete spectroscopic follow-up has not always been
+the norm, as discussed below.}
+In contrast, targeted surveys have highly complex observational
+selection functions and so are discarded from our analysis. High-redshift SN cosmology
+samples, such as Pantheon, are predominately from magnitude-limited surveys from
+which volume-limited SN~Ia subsamples can be constructed.
+%\textbf{which have easier to model observational selection functions}.
 
@@
 c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
-distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
+parent distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
 \citet{campbell2013} for similar contours, the second using a more conservative
@@
     sample, taking the limits as initially suggested by the previous
-    completeness analysis. \textbf{However,} as this solution might be over
-    simplistic to create a complete sample, e.g., for it ignores follow-up
-    efficiency, we also consider another set of cuts to define a so-called
-    conservative sample. This latter is smaller and therefore less statistically
-    constraining, but also inherently less prone to selection effects. If the
+    completeness analysis. \textbf{However,} as this solution might be an overly
+    simplified way to create a complete sample, e.g., because it ignores spectroscopic follow-up
+    in efficiency for redshifts below $z_{lim}$, we also consider another set of cuts to define a so-called
+    ``conservative'' sample. The latter is smaller and therefore less statistically
+    constraining, but also even less prone to observational selection effects. If the
     redshift drift is still significant in the conservative sample, it would be
-    even more meaningful in a carefully-tailored selection-free sample. The goal
-    of this study is to build the first implementation of a SN Ia properties
-    drift modeling before refining it on a more complex, yet more accurate,
-follow-up analysis also accounting for precise selection modeling using, e.g.,
+    even more meaningful in a carefully-tailored selection-free sample. These
+    samples are adequate for the goal
+    of this study, which is to develop a first implementation of a model for drift in
+    SN Ia properties. If fruitful, the sample selection can later be refined, if needed, with a
+    more detailed
+    %drift modeling before refining it on a more complex, yet more accurate,
+%follow-up analysis also accounting for precise 
+model of the observational selection, e.g., using
 the SNANA package \citep{SNANA}.}
@@
 redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
-filter, that has a $5\sigma$ depth of
+filter, which has a $5\sigma$ depth of
 24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
@@
 $z_{\lim}=0.55$. \textbf{We will therefore use $z=0.60$ and $z=0.55$ as redshift
-limits for the fiducial and conservative cuts, respectively.}
+limits for the fiducial and conservative samples, respectively, for SNLS.}
 
@@
 $z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
-conservative cuts, respectively.}
+conservative samples, respectively, for PS1.}
 
@@
 spectroscopic resources were available, and \cite{kessler2009} and
-\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
-these analyses, \textbf{we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
-fiducial and conservative cuts, respectively.}
+\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Based on this,
+\textbf{we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
+fiducial and conservative samples, respectively, for SDSS.}
 
@@
         \citealt{scolnic2018a}). The colored parts represent the distribution of
-        SNe~Ia kept in our analysis for they are supposedly free from selection
+        SNe~Ia kept in our analysis for they are supposedly free from observational selection
     bias (see Section~\ref{sec:sample}). The darker (resp. lighter) color
@@
     enough to include them within the cosmological analysis \citep{scolnic2018a}
-    and we do not add further quality cuts. Section~\ref{sec:results} highlights
+    and we do not impose further cuts. Section~\ref{sec:results} highlights
     that, while compatible with it, our results are not dependent on the
@@
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe
-that the fraction of low-stretch SNe (typically $x_1 < -1$) seems to decrease as
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe here
+that the fraction of low-stretch SNe (typically $x_1 < -1$) appears to decrease as
 a function of redshift; this is confirmed in Fig.~\ref{fig:modelall} in which}
@@
 \textbf{In section~\ref{ssec:cuts}, we have built volume-limited samples from a
-    set of magnitude-limited ones, using simple redshift cuts. This simplistic
-    approach is statistically sub-optimal, but should be enough to demonstrate
-    our key finding: a redshift evolution of stretch compatible with
+    set of magnitude-limited ones, using simple redshift cuts. This simplified
+    approach is statistically sub-optimal, but should be enough to test
+    our key question: whether redshift evolution of stretch is compatible with the
     \cite{rigault2020} model. However, the possibility remains that a complex
-    selection function related to follow-up efficiencies below our fiducial (or
-    even conservative) cuts could still affect our sample, making it not fully
+    observational selection function related to spectroscopic follow-up efficiencies below our fiducial (or
+    even conservative) redshift cuts could still affect our sample, making it not fully
 volume-limited; this would, in turn, bias our conclusion on the astrophysical
@@
 
-\textbf{To test for the existence of potential leftover selection biases in our
+\textbf{To test for the existence of potential leftover observational selection biases in our
     sample, we compare the stretch and color distributions of the SNe~Ia
-    originating from different datasets over overlapping redshift ranges: these
-    distributions should be similar. Note that the redshift range has to be
+    originating from different datasets having overlapping redshift ranges: these
+    distributions should be similar if they reflect the underlying parent population. 
+    Note that the redshift range has to be
     narrow enough so that any drift would be negligible.}
@@
 \textbf{The two samples that overlap the most in redshift are PS1 and SDSS in
-    the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
-    corresponds to the 146 SNe~Ia from SDSS at the high redshift end and thus
-    the most likely to be affected by residual selection effects (see the
+    the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping subsample
+    consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus
+    the most likely to be affected by residual observational selection effects (see the
     corresponding discussion in section~\ref{ssec:cuts}). Over that same
     redshift range, PS1 has 52 SNe~Ia, which are in the lowest redshift bins and
-    thus unlikely to have any selection issue. To identify potential
+    thus unlikely to have any observational selection issue. To identify potential
     inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib}
     (upper panels) compares the stretch and color distribution of both these
-    surveys. The associated Kolmogorov-Smirnov (KS) similarity test $p$-values
-are too high ($>10\%$) to conclude on any inconsistency, in agreement with the
+    surveys. The resulting Kolmogorov-Smirnov (KS) similarity test $p$-values
+($p >10\%$) do not support any inconsistency, in agreement with the
 visual impression from Fig.~\ref{fig:distrib}.}
@@
     conclusion can be drawn: there is no substantial sign of discrepancy in the
-    stretch and color distributions between the low- and high-end of the SNLS
-    and PS1 datasets, respectively. Nonetheless, the small size of the SNLS
+    stretch and color distributions between the low- and high-end of our fiducial SNLS
+    and PS1 samples, respectively. Nonetheless, the small size of the SNLS
     dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of
-    the test, and only strong deviation would be noticeable. Extending the
+    this test, and only a strong deviation would be noticeable. Extending the
 redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
@@
 
-\textbf{We finally highlight that the SNe~Ia color is more prone to selection
+\textbf{We finally highlight that the SNe~Ia color is more prone to observational selection
     effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
     Fig.~3 of \cite{kessler2017}. Hence, since the comparison of color
-    distributions shows no significant hint of leftover selection effect, it
+    distributions shows no significant hint of leftover observational selection effect, it
     further supports our claim that our simple redshift-based selection criteria
-    is sufficient to build a complete SNe~Ia sample required to test the
+    are sufficient to build the complete SNe~Ia samples required to test the
     redshift evolution of the stretch distribution.}
@@
 \end{align}
-
-\textbf{This is our Base model.}
+\textbf{This is our Base drifting model.}
 
@@
 For the rest of the analysis, we will therefore minimize Eq.~\ref{eq:likelihood}
-using $p_y^i$ --~the probability for the SN $i$ to be young~-- when available
-(i.e. for SNfactory dataset), and $\delta(z^i)$ --~the expected fraction of
-young SNe~Ia at the SN redshift $z^i$~-- otherwise.
+using $p_y^i$ --- the probability for the SN $i$ to be young --- when available
+(i.e. for SNfactory dataset), and $\delta(z^i)$ --- the expected fraction of
+young SNe~Ia at the SN redshift $z^i$ --- otherwise.
 
@@
 
-\textbf{In order to ensure that our results are not driven by the poorly-modeled
-    HST sub-sample, we recompute $\Delta$AIC for each model excluding this
+\textbf{In order to ensure that our results are not driven by the incompletely-modeled
+    HST subsample, we recompute $\Delta$AIC for each model excluding this
     dataset; we find that it does not change $\Delta$AIC by more than few
     tenths. The consistency of these values with those in Table~\ref{tab:comp}
-show that the HST sub-sample does not drive our conclusions.}
-
-We report in Table~\ref{tab:bbc} \textbf{our determination of} the samples'
-$\mu_0$ and $\sigma_{\pm}$ \textbf{when implementing an asymmetric Gaussian
+show that the HST subsample does not drive our conclusions.}
+
+We report in Table~\ref{tab:bbc} \textbf{our determination of}
+$\mu_0$ and $\sigma_{\pm}$ \textbf{for each sample when implementing an asymmetric Gaussian
 model}, adjusted on the nominally selection-free samples using our fiducial cuts
@@
 derived these model parameters using the full BBC formalism, using numerous
-simulations to model the selection effects (see details e.g., Section~3 of
+simulations to model the observational selection effects (see details e.g., Section~3 of
 \citealt{kessler2017}). The agreement between our fit of the asymmetric
 Gaussians on the supposedly selection-free part of the samples and the results
-derived using the BBC formalism supports our approach to \textbf{construct} a
-sample with negligible selection effects. If we were to use \cite{scolnic2016}
+derived using the BBC formalism supports our approach to \textbf{constructing} a
+sample with negligible observational selection effects. If we were to use \cite{scolnic2016}
 and \cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
@@
 might wonder whether a low-stretch mode might also exist in the
-young-population, see Fig.~\ref{fig:stretchlssfr}. We \textbf{test for this} by
-allowing this population to also be bi-modal, finding the amplitude of
-\textbf{such a} low-stretch mode to be compatible with~0 in this young
-population ($<2\%$). More generally, this raises the question of \textbf{how
-well a given environmental tracer (here LsSFR) traces the age}. A dedicated
-analysis will be presented in Briday et al.\ in prep.
-
-Finally, ignoring the LsSFR measurements --~available only for the SNfactory
-dataset, see Section~\ref{sec:modeling}~-- reduces the significance of the
-results presented in this section, as expected. Yet, non-drifting models remain
-strongly disfavored, and for instance, the best fitted sample-based Gaussian
-asymmetric modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the
-data than our Base drifting modeling.
+young-population (see Fig.~\ref{fig:stretchlssfr}). We \textbf{test for this} by
+allowing this population to also be bimodal, finding the amplitude of
+\textbf{such a} low-stretch mode to be compatible with~0 ($<2\%$) in this young
+population. More generally, this raises the question of \textbf{how
+well a given environmental tracer (here LsSFR) traces the age}. An
+analysis dedicated to this question will be presented in Briday et al.\ in prep.
+
+Finally, ignoring the LsSFR measurements --- available only for the SNfactory
+dataset (see Section~\ref{sec:modeling}) --- reduces the significance of the
+results presented in this section, as expected. Even so, non-drifting models remain
+strongly disfavored. For instance, the best fitting sample-based Gaussian
+asymmetric model is still $\Delta\mathrm{AIC}<-10$ less representative of the
+data than our Base drifting model.
 
@@
 formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016} or Steve
-\citep{hinton2019} can easily allow it; see e.g., section 1.3 and 2.5 of
-\cite{rubin2015}. Not doing so is a second order issue for SN cosmology, as it
+\citep{hinton2019} can easily allow it (see e.g., sections~1.3 and 2.5 of
+\cite{rubin2015}). Not doing so is a second order issue for SN cosmology, as it
 only affects the way one accounts for Malmquist bias. Indeed, as long as the
@@
 However, surveys usually do have significant Malmquist bias for the upper half
-of their SN redshift distribution. As a consequence, an ill-modeling of the
+of their SN redshift distribution. As a consequence, mismodeling of the
 underlying stretch distribution will bias the SN magnitudes derived from such
@@
 Commonly used Malmquist bias correction techniques, such as the BBC-formalism,
-assume per sample asymmetric Gaussian functions for modeling the underlying
+assume per sample asymmetric Gaussian functions to model the underlying
 stretch and color distributions. Yet, as shown in Section~\ref{sec:results},
-such a sample-based distribution is excluded as being as good as our drifting
-model. Then, unlike what \citet[][Section~2]{scolnic2016} and
-\citet[][Section~5.4]{scolnic2018a} suggested, i.e. that traditional surveys
-span limited redshift ranges and that therefore the per-sample approach accounts
-for implicit redshift drifts, a direct modeling of the redshift drift is more
+such a sample-based distribution is excluded in comparison to our drifting
+model. Then, contrary to what \citet[][Section~2]{scolnic2016} and
+\citet[][Section~5.4]{scolnic2018a} have suggested --- i.e., that traditional surveys
+span sufficiently limited redshift ranges such that the per-sample approach accounts
+for implicit redshift drifts --- a direct modeling of the redshift drift is more
 appropriate than a sample-based approach. We \textbf{add} here that, as
@@
 lower redshift), the higher the amplitude of the old-specific low-stretch mode.
-This redshift dependency is shown as blue to red underlying distributions in
-Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges covered by PS1. The observed
+This redshift dependency of the underlying stretch distributions is shown running from blue to red in
+Fig.~\ref{fig:bbc_pdf_ps1} for the redshift range covered by PS1. The observed
 $x_1$ histogram follows \textbf{the model we defined using} the sum of
@@
 for typical PS1 SN redshifts. This means that the SN bias-corrected standardized
-magnitude estimated at a redshift affected by selection effects would be biased
-by an ill-modeling of the true underlying stretch distribution.
+magnitude estimated at a redshift suffering from by observational selection effects would be biased
+by an mismodeling of the true underlying stretch distribution.
 
@@
 redshift drift should be further studied and care should be taken when using
-samples affected by selection effects.
+samples affected by observational selection effects.
 
@@
 We have presented \textbf{an initial} study of the drift of the underlying SNe~Ia
-stretch distribution as a function of redshift. We built a magnitude-limited
-SN~Ia sample from the Pantheon dataset \citep[][SDSS, PS1 and
+stretch distribution as a function of redshift. We built effectively volume-limited
+SN~Ia subsamples from the Pantheon dataset \citep[][SDSS, PS1 and
 SNLS]{scolnic2018a}, to which we added HST and SNfactory data from
@@
 SNe that have been discovered in the redshift range of each survey where
-selection effects are negligible, so that the observed SNe~Ia stretches are
-a random sampling of the true underlying distribution. This resulted in a 569
-SN~Ia fiducial sample (422 SNe when more conservative cuts were considered).
+observational selection effects are negligible, so that the observed SNe~Ia stretches are
+a random sampling of the true underlying distribution. This resulted in a fiducial sample 
+of 569 SNe~Ia (422 SNe when more conservative cuts were applied).
 
@@
 
-In addition to this ``base'' modeling, we have studied various distributions,
+In addition to this ``base'' model, we have studied various distributions,
 including redshift independent models; we also studied the prediction from a
@@
         dependent, as previously suggested by e.g.~\cite{howell2007}, \textbf{in
-        a way that selection effects alone cannot explain}. This result is
+        a way that observational selection effects alone cannot explain}. This result is
         largely independent of details on each age-population model.
@@
         the relative fraction of younger and older SNe~Ia follows the prediction
-        made in \cite{rigault2020}. This second result strongly supports the
+        made in \cite{rigault2020}. This second result further supports the
         existence of both young and old SN~Ia populations, in agreement with
@@
 
-    \item Models using survey-based asymmetric Gaussian distributions, as done,
-        e.g., in the current implementation of BBC, are excluded \textbf{as a}
+    \item Models using survey-based asymmetric Gaussian distributions, e.g., as employed
+        in the current implementation of BBC, are excluded \textbf{as a}
         good description of the data relative to our drifting model. Hence, the
-        sample-based approach does not accurately account for redshift drift and
-        even less so as survey span increasingly larger redshift ranges.
-        \textbf{As a result, if} extra degrees of freedom might be acceptable
+        sample-based approach does not accurately account for redshift drift, a problem
+        that will be exacerbated as surveys span increasingly larger redshift ranges.
+        \textbf{As a result, even if the necessary} extra degrees of freedom might be acceptable
         given the large number of SNe~Ia in cosmological studies, extrapolating
         the SN property distributions from the volume-limited part of a survey
-        to its Malmquist-biased magnitude-limited one would still be inaccurate
+        to its Malmquist-biased magnitude-limited part would still be inaccurate
         because of the redshift evolution.
@@
 In this paper, we considered a simple \textbf{two-population} Gaussian mixture
-modeling, but additional data free from significant Malmquist bias would enable
-us to refine it as necessary. We note that samples at the low- and high-redshift
+modeling. Additional data free from significant Malmquist bias would enable
+us to refine it, as necessary. We note that samples at the low- and high-redshift
 ends of the Hubble diagram would be particularly helpful for this drifting
@@
 
-\textbf{The next step of the analysis will be to incorporate our modeling within
-the SNANA framework \citep{SNANA} both to more accurately account for selection
-functions and to test the impact of our modeling on the derivation of
+\textbf{The next step in this line of analysis will incorporate our model into
+the SNANA framework \citep{SNANA}, both to more accurately account for observational selection
+functions and to test the impact of our model on the derivation of
 cosmological parameters; this study is underway.}
@@
 \bibitem[Strolger et al.(2004)]{strolger04} Strolger, L.-G., Riess, A.~G.,
-Dahlen, T., et al.\ 2004, \apj, 613, 200
 

### 2021-01-15 Nicolas.tex
MR relecture and MB corrections. Bold to add.

@@
 Distribution}
+
+\titlerunning{Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution}
+\authorrunning{N.~Nicolas et al.}
 
@@
 fundamental physics. Yet, no simple solution is able to accommodate this $H_0$
-tension when accounting for all other probes \citep{knox2019}, but see e.g.
-\cite{poulin2019}. Alternatively, systematic effects affecting one or several
+tension when accounting for all other probes \citep{knox2019}. Alternatively, systematic effects affecting one or several
 of the aforementioned analyses might explain at least some of this tension.
@@
 and their host galaxy properties has been studied extensively. The first key finding
-was that that the standardized SNe~Ia magnitudes
+was that the standardized SNe~Ia magnitudes
 significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass host 
@@
 More recently, studies have used the local SN host galaxy environment to probe
-more direct connections between the SNe and their host galaxy environments \citep{rigault2013},
+more direct connections between the SNe and their host galaxy environment \citep{rigault2013},
 showing that local age tracers such as the Local specific Star Formation Rate
 (LsSFR) or the local color are more strongly correlated with the standardized SN
-magnitude \citep{rigault2020, roman2018, kim18}. Further reinforcing this
+magnitude \citep{roman2018, kim18,rigault2020}. Further reinforcing this
 connection is the identification of SN~Ia spectral features that are 
@@
 following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
-model). In \cite{rigault2020} we used the LsSFR to classify which are the
+model). \cite{rigault2020} used the LsSFR to classify which are the
 younger (those with a high LsSFR) and which are the older (those with low
@@
 
-In this paper, we take a step aside to probe the validity of our modeling of the
+In this paper, we take a step aside \mr{from the cosmological analyses} to probe the validity of our modeling of the
 SN population, which we claim to be constituted of two age-populations
@@
 of stretch and our results are presented in Section~\ref{sec:results}. In this
-section we test whether the SN stretch distribution evolves as a function of
+section, we test whether the SN stretch distribution evolves as a function of
 redshift and if the aforementioned age model is in good agreement with this
@@
 
-Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring type SN types
+Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring SN types
 and host galaxy redshifts, the observation selection effects of magnitude-limited surveys should be
@@
     completeness analysis. \textbf{However,} as this solution might be an overly
-    simplified way to create a complete sample, e.g., because it ignores spectroscopic follow-up
-    in efficiency for redshifts below $z_{lim}$, we also consider another set of cuts to define a so-called
+    simplified way to create a complete sample (e.g., because it ignores spectroscopic follow-up
+    in efficiency for redshifts below $z_{lim}$), we also consider another set of cuts to define a so-called
     ``conservative'' sample. The latter is smaller and therefore less statistically
@@
 expected, the selected redshift limits \textbf{are roughly located slightly
-before the peak of these histograms. In Section 2.2 we validate that these
+before the peak of these histograms. In Section \ref{ssec:verify} we validate that these
 redshift limits are effective for constructing nearly volume-limited subsamples
@@
 that the fraction of low-stretch SNe (typically $x_1 < -1$) appears to decrease as
-a function of redshift; this is confirmed in Fig.~\ref{fig:modelall} in which}
+a function of redshift; this is confirmed in Fig.~\ref{fig:modelall}, in which}
 the evolution of the mean stretch is shown, \textbf{with} the data split in
@@
 by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
-of the underlying distribution of SNe~Ia as a function of redshift $X_1(z)$ is
+of the underlying \textbf{stretch} distribution of SNe~Ia as a function of redshift $X_1(z)$ is
 given by:
@@
 \delta(z^i)$ for any given SN~$i$. 
-
+\\
+\\
 For the rest of the analysis, we will therefore minimize Eq.~\ref{eq:likelihood}
@@
 
-\bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T.,
-et al.\ 2019, \prl, 122, 221301
+%\bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T., et al.\ 2019, \prl, 122, 221301
 
@@
 
-\bibitem[{Rigault {et~al.}(2013)}]{rigault2013} Rigault, M., Copin, Y.,
+\bibitem[{Rigault et al.(2013)}]{rigault2013} Rigault, M., Copin, Y.,
 Aldering, G., {et~al.} 2013, \aap, 560, A66

### 2021-01-18 Nicolas.tex
One column for referee!

@@
 
-\titlerunning{Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution}
+\titlerunning{Redshift Evolution of the Underlying Type Ia Supernova Stretch
+Distribution}
 \authorrunning{N.~Nicolas et al.}
@@
 
-\institute{Université de Lyon, F-69622, Lyon, France; Université de Lyon
-    1, Villeurbanne; CNRS/IN2P3, Institut de Physique des Deux Infinis, Lyon
+\institute{Univ Lyon, Univ Claude Bernard Lyon 1, CNRS, IP2I Lyon / IN2P3, IMR
+    5822, F-69622, Villeurbanne, France
     \and 
@@
 
-\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
-    and as survey statistics increase, the question of astrophysical
-    systematic uncertainties rises, notably that of SN~Ia population
-    evolution. In this paper, we study the dependence with redshift of the SN~Ia
-    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
-    probe its potential redshift drift. The SN stretch has been shown to
-    strongly correlate with the SN environment, notably with stellar age
-    tracers. We model the underlying stretch distribution as a function of
+\abstract{The \textbf{detailed} nature of type Ia supernovae (SNe~Ia) remains
+    \textbf{uncertain}, and as survey statistics increase, the question of
+    astrophysical systematic uncertainties rises, notably that of SN~Ia
+    population evolution. In this paper, we study the dependence with redshift
+    of the SN~Ia \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN
+    property, to probe its potential redshift drift. The SN stretch has been
+    shown to strongly correlate with the SN environment, notably with stellar
+    age tracers. We model the underlying stretch distribution as a function of
     redshift, using the evolution of the fraction of young and old SNe~Ia as
@@
     have marginal magnitude selection effects, so that any observed change is
-    indeed of astrophysical and not observational in origin. \textbf{In this first
-    study, there are indications} that the underlying SN~Ia stretch distribution
-    is evolving as a function of redshift, and that the young/old drifting model
-    is a better description of the data than any time-constant model, including
-    the sample-based asymmetric distributions often used to correct Malmquist
-    bias. The favored underlying stretch model is the bimodal one derived from
-    \cite{rigault2020}, \textbf{composed of} a high-stretch mode shared by both
-    young and old environments, and a low-stretch mode exclusive to old
-    environments. The precise impact of the redshift evolution of the SN Ia
-    population intrinsic properties on cosmology remains to be studied. Yet, the
-    astrophysical drift of the SN stretch distribution does affect current
-    Malmquist bias corrections and thereby distances derived from SN affected by
-observational selection effects. We highlight that such a bias will increase with surveys
-covering increasingly larger redshift ranges, which is particularly important
-for LSST.}
+    indeed of astrophysical and not observational \textbf{in} origin. \textbf{In
+    this first study, there are indications} that the underlying SN~Ia stretch
+    distribution is evolving as a function of redshift, and that the young/old
+    drifting model is a better description of the data than any time-constant
+    model, including the sample-based asymmetric distributions \textbf{often}
+    used to correct Malmquist bias. The favored underlying stretch model is the
+    bimodal one derived from \cite{rigault2020}, \textbf{composed of} a
+    high-stretch mode shared by both young and old environments, and a
+    low-stretch mode exclusive to old environments. The precise impact of the
+    redshift evolution of the SN Ia population intrinsic properties on cosmology
+    remains to be studied. Yet, the astrophysical drift of the SN stretch
+    distribution does affect current Malmquist bias corrections and thereby
+distances derived from SN affected by \textbf{observational} selection effects.
+We highlight that such a bias will increase with surveys covering increasingly
+larger redshift ranges, which is particularly important for LSST.}
 
@@
 \citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ derived when
-the SNe~Ia are anchored to Cepheids \citep[the SH0ES project,][]{riess2009,
-riess2016} is $\sim5\sigma$ higher than what is predicted from cosmic microwave
-background (CMB) data measured by Planck assuming the standard $\L$CDM
-\citep{planck2018, riess2019, reid2019}, or when the SN luminosity is anchored
-at intermediate redshift by the baryon acoustic oscillation (BAO) scale
-\citep{feeney2019}. While using the tip of the red giant branch technique in
-place of the Cepheids seem to favor an intermediate value of $H_0$
-\citep{freedman2019, freedman2020}, time delay measurements from strong lensing
-seem to also favor high $H_0$ values \citep{wong2019}.
+the SNe~Ia are anchored \textbf{to} Cepheids \citep[the SH0ES
+project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
+from cosmic microwave background (CMB) data measured by Planck assuming the
+standard $\L$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
+luminosity is anchored at intermediate redshift by the baryon acoustic
+oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
+branch technique in place of the Cepheids seem to favor an intermediate value of
+$H_0$ \citep{freedman2019, freedman2020}, time delay measurements from strong
+lensing seem to also favor high $H_0$ values \citep{wong2019}.
 
@@
 fundamental physics. Yet, no simple solution is able to accommodate this $H_0$
-tension when accounting for all other probes \citep{knox2019}. Alternatively, systematic effects affecting one or several
-of the aforementioned analyses might explain at least some of this tension.
-\cite{rigault2015} suggested that SNe Ia from the Cepheid calibrator sample
-differ by construction from those in the Hubble flow sample, as the former
-strongly favors young stellar populations, while the latter does not. This
-selection effect would impact the derivation of $H_0$ if SNe~Ia from young and
-older environments differ in average standardized magnitudes.
-
-%For the last decades, numerous analyses have studied 
-The relation between SNe~Ia
-and their host galaxy properties has been studied extensively. The first key finding
-was that the standardized SNe~Ia magnitudes
-significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass host 
-galaxies being
-brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
-betoule2014, rigault2020, kim19}. This mass-step correction is currently used in
-cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including for
-deriving $H_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
-between the SNe and their host galaxies remains unclear, particularly when using global properties such
-as the host stellar mass. This in turn raises the question of the accuracy of such
-corrections since host galaxy global properties evolve with redshift. 
-More recently, studies have used the local SN host galaxy environment to probe
-more direct connections between the SNe and their host galaxy environment \citep{rigault2013},
+tension when accounting for all other probes \citep{knox2019}. Alternatively,
+systematic effects affecting one or several of the aforementioned analyses
+\textbf{might} explain at least some of this tension. \cite{rigault2015}
+suggested that SNe Ia from the Cepheid calibrator sample differ by construction
+from \textbf{those in the Hubble flow sample,} as the former strongly favors
+young stellar populations, while the latter \textbf{does} not. This selection
+effect would impact the derivation of $H_0$ if SNe~Ia from young and older
+environments differ in average standardized magnitudes.
+
+\textbf{The relation between SNe~Ia and their host galaxy properties has been
+studied extensively. The first key finding was} that the standardized SNe~Ia
+magnitudes significantly depend on the host \textbf{galaxy} stellar mass, with
+SNe~Ia from high-mass host \textbf{galaxies} being brighter on average
+\cite[e.g.][]{kelly2010, sullivan2010, childress2013, betoule2014, rigault2020,
+kim19}. This mass-step correction is currently used in cosmological analyses
+\citep[e.g.][]{betoule2014, scolnic2018a}, including for deriving $H_0$
+\citep{riess2016, riess2019}. Yet, the underlying connection between the SNe and
+their host \textbf{galaxies} remains unclear\textbf{, particularly} when using
+global properties such as the host stellar mass\textbf{. This in turn} raises
+the question of the accuracy of \textbf{such corrections since host galaxy
+global properties evolve with redshift}. More recently, studies have used the
+local SN \textbf{host galaxy} environment to probe more direct connections
+between the \textbf{SNe and their host galaxy environments} \citep{rigault2013},
 showing that local age tracers such as the Local specific Star Formation Rate
 (LsSFR) or the local color are more strongly correlated with the standardized SN
-magnitude \citep{roman2018, kim18,rigault2020}. Further reinforcing this
-connection is the identification of SN~Ia spectral features that are 
-correlated with LsSFR \citep{nordin2018}. These results suggest age as the driving
+magnitude \citep{roman2018, kim18,rigault2020}. \textbf{Further reinforcing this
+connection is the identification of SN~Ia spectral features that are correlated
+with LsSFR \citep{nordin2018}. These results suggest} age as the driving
 parameter underlying the mass-step. If true, this would have a significant
@@
 following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
-model). \cite{rigault2020} used the LsSFR to classify which are the
-younger (those with a high LsSFR) and which are the older (those with low
-LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN stretch has
-been known to be strongly correlated with the SN host galaxy properties
-\citep{hamuy1996, hamuy2000}. This correlation that has been extensively confirmed
-since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
+model). \cite{rigault2020} used the LsSFR to classify which are the younger
+(those with a high LsSFR) and which are the older (those with low LsSFR).
+Furthermore, since the first SNe~Ia host analyses, the SN stretch has been known
+to be strongly correlated with the SN host \textbf{galaxy} properties
+\citep{hamuy1996, hamuy2000}. This correlation that has been extensively
+confirmed since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
 gupta2011, dandrea2011, childress2013, rigault2013, pan2014, kim19}. Following
-the ``A+B'' model and the connection between SN stretch and host galaxy properties,
-\cite{howell2007} first discussed the potential redshift drift of the SN stretch
-distribution. In this paper we revisit this question using the most recent SNe~Ia
-datasets.
-
-In this paper, we take a step aside \mr{from the cosmological analyses} to probe the validity of our modeling of the
-SN population, which we claim to be constituted of two age-populations
-\citep{rigault2013,rigault2015,rigault2020}: one old and one younger, the former
-having on average lower lightcurve stretches and being brighter after
-standardization. We use the correlation between the SN age, as probed by the
-LsSFR, and the SN stretch to model the expected evolution of the underlying SN
-stretch distribution as a function of redshift. This modeling relies on three
-assumptions: (1) there are two distinct populations of SNe~Ia; (2) the relative
-fraction of each of these populations as a function of redshift follows the
-model presented in \cite{rigault2020} and (3) the underlying distribution of
-stretch for each age sample is constant. This paper aims at testing this
-specific model with datasets from the literature. 
-% Note that the progenitor age
-% as traced by the LsSFR seems to capture physical feature intrinsic to the
-%progenitor and/or explosion mechanism that the stretch alone is not capturing
-% \citep{nordin2018}.
+the ``A+B'' model and the connection between SN stretch and host galaxy
+properties, \cite{howell2007} first discussed the potential redshift drift of
+the SN stretch distribution. In this paper we revisit this \textbf{question
+using} the most recent SNe~Ia \textbf{datasets}.
+
+In this paper, we take a step aside \textbf{from the cosmological analyses} to
+probe the validity of our modeling of the SN population, which we claim to be
+constituted of two age-populations \citep{rigault2013,rigault2015,rigault2020}:
+one old and one younger, the former having on average lower lightcurve stretches
+and being brighter after standardization. We use the correlation between the SN
+age, as probed by the LsSFR, and the SN stretch to model the expected evolution
+of the underlying SN stretch distribution as a function of redshift. This
+modeling relies on three assumptions: (1) there are two distinct populations of
+SNe~Ia; (2) the relative fraction of each of these populations as a function of
+redshift follows the model presented in \cite{rigault2020} and (3) the
+underlying distribution of stretch for each age sample is constant. This paper
+aims at testing this specific model with datasets from the literature.
 
@@
 
-The ideal SN~Ia sample for studying this question would be a very deep, large-area,
-volume-limited sample. This would capture the true underlying stretch distribution
-function, and we would then study how it evolves with redshift. No such sample exists, 
-so first we must construct subsamples that are as near to volume-limited as possible 
-from existing high-redshift SN~Ia samples.
+\textbf{The ideal SN~Ia sample for studying this question would be a very deep,
+    large-area, volume-limited sample. This would capture the true underlying
+    stretch distribution function, and we would then study how it evolves with
+    redshift. No such sample exists, so first we must construct subsamples that
+are as near to volume-limited as possible from existing high-redshift SN~Ia
+samples.}
 
@@
 distributions in a few bins of redshift. In practice, however, differential
-observational selection effects will affect the observed SN stretch distributions. Indeed,
-because the observed SN~Ia magnitude correlates with the lightcurve stretch (and
-color), the first SNe~Ia that a magnitude-limited survey will miss are the
-lowest-stretch (and reddest) ones. Consequently, if magnitude-related observational selection
-effects are not accounted for, one might confuse true population drift with
-survey properties, and conversely.
-
-Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring SN types
-and host galaxy redshifts, the observation selection effects of magnitude-limited surveys should be
-negligible below a given redshift at which even the faintest normal SNe~Ia can be
-observed. \textbf{Aiding in the construction of nearly volume-limited 
-subsamples is the fact that the SN~Ia population trails off towards fainter SNe~Ia.
-A complication is that complete spectroscopic follow-up has not always been
-the norm, as discussed below.}
-In contrast, targeted surveys have highly complex observational
-selection functions and so are discarded from our analysis. High-redshift SN cosmology
-samples, such as Pantheon, are predominately from magnitude-limited surveys from
-which volume-limited SN~Ia subsamples can be constructed.
-%\textbf{which have easier to model observational selection functions}.
+\textbf{observational} selection effects \textbf{will affect} the observed SN
+stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
+with the lightcurve stretch (and color), the first SNe~Ia that a
+magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
+Consequently, if magnitude-related \textbf{observational} selection effects are
+not accounted for, one might confuse true population drift with survey
+properties, and conversely.
+
+Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring
+\textbf{SN types and host galaxy redshifts}, the \textbf{observation} selection
+effects of magnitude-limited surveys should be negligible below a given redshift
+at which even the faintest \textbf{normal} SNe~Ia can be observed.
+\textbf{Aiding in the construction of nearly volume-limited subsamples is the
+    fact that the SN~Ia population trails off towards fainter SNe~Ia. A
+    complication is that complete spectroscopic follow-up has not always been
+    the norm, as discussed below.} In contrast, targeted surveys have highly
+    complex \textbf{observational} selection functions and \textbf{so are}
+    discarded from our analysis. \textbf{High-redshift SN cosmology samples,
+    such as Pantheon, are predominately from magnitude-limited surveys from
+which volume-limited SN~Ia subsamples can be constructed.}
 
@@
 c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
-parent distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
-\citet{campbell2013} for similar contours, the second using a more conservative
-$|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is
-$M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
-standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
-along the aforementioned ellipse given the standardization coefficient
-$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is
-that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
-in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one ought to detect
-this object typically 5 days before and a week after peak to build a suitable
-lightcurve, the effective limiting standardized absolute magnitude is
-approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit
-$m_{\lim}$ of a magnitude limited survey, one can derive the maximum redshift
-$z_{\lim}$ above which the faintest SNe~Ia will be missed using the relation
-between apparent magnitude, redshift and absolute magnitude $\mu(z_{\lim}) =
-m_{\lim} - M_{\lim}$.
+\textbf{parent} distribution \citep{guy2007, betoule2014}; see also
+\citet{bazin2011} and \citet{campbell2013} for similar contours, the second
+using a more conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude
+with $x_1=0$ and $c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can
+derive the absolute standardized magnitude at maximum of light $M = M_0 - \alpha
+x_1 + \beta c$ along the aforementioned ellipse given the standardization
+coefficient $\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the
+faintest SN~Ia is that with $(x_1=-1.65, c=0.25)$ and an absolute standardized
+magnitude at peak in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one
+ought to detect this object typically 5 days before and a week after peak to
+build a suitable lightcurve, the effective limiting standardized absolute
+magnitude is approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude
+limit $m_{\lim}$ of a magnitude limited survey, one can derive the maximum
+redshift $z_{\lim}$ above which the faintest SNe~Ia will be missed using the
+relation between apparent magnitude, redshift and absolute magnitude
+$\mu(z_{\lim}) = m_{\lim} - M_{\lim}$.
 
@@
     sample, taking the limits as initially suggested by the previous
-    completeness analysis. \textbf{However,} as this solution might be an overly
-    simplified way to create a complete sample (e.g., because it ignores spectroscopic follow-up
-    in efficiency for redshifts below $z_{lim}$), we also consider another set of cuts to define a so-called
-    ``conservative'' sample. The latter is smaller and therefore less statistically
-    constraining, but also even less prone to observational selection effects. If the
-    redshift drift is still significant in the conservative sample, it would be
-    even more meaningful in a carefully-tailored selection-free sample. These
-    samples are adequate for the goal
-    of this study, which is to develop a first implementation of a model for drift in
-    SN Ia properties. If fruitful, the sample selection can later be refined, if needed, with a
-    more detailed
-    %drift modeling before refining it on a more complex, yet more accurate,
-%follow-up analysis also accounting for precise 
-model of the observational selection, e.g., using
-the SNANA package \citep{SNANA}.}
+    completeness analysis. However, as this solution might be an overly
+    simplified way to create a complete sample, e.g., because it ignores
+    spectroscopic follow-up in efficiency for redshifts below $z_{lim}$, we
+    also consider another set of cuts to define a so-called ``conservative''
+    sample. The latter is smaller and therefore less statistically constraining,
+    but also even less prone to observational selection effects. If the redshift
+    drift is still significant in the conservative sample, it would be even more
+    meaningful in a carefully-tailored selection-free sample. These samples are
+    adequate for the goal of this study, which is to develop a first
+    implementation of a model for drift in SN Ia properties. If fruitful, the
+    sample selection can later be refined, if needed, with a more detailed model
+of the observational selection, e.g., using the SNANA package \citep{SNANA}.}
 
@@
 redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
-filter, which has a $5\sigma$ depth of
+filter, \textbf{which} has a $5\sigma$ depth of
 24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
@@
 spectroscopic resources were available, and \cite{kessler2009} and
-\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Based on this,
-\textbf{we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
+\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$.
+\textbf{Based on this, we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
 fiducial and conservative samples, respectively, for SDSS.}
@@
 expected, the selected redshift limits \textbf{are roughly located slightly
-before the peak of these histograms. In Section \ref{ssec:verify} we validate that these
-redshift limits are effective for constructing nearly volume-limited subsamples
-from samples that were initially more closely magnitude-limited in their search
-or spectroscopic follow-up.}
+    before the peak of these histograms. In Section \ref{ssec:verify} we
+validate that these redshift limits are effective for constructing nearly
+volume-limited subsamples from samples that were initially more closely
+magnitude-limited in their search or spectroscopic follow-up.}
 
@@
     \caption{Composition of the SNe~Ia dataset used in this analysis.
-    Conservative cuts are indicated in parentheses. \textbf{The SNf limit is set by
-\cite{rigault2020}, see text.}}
+    Conservative cuts are indicated in parentheses. \textbf{The SNf limit is set
+by \cite{rigault2020}, see text.}}
     \label{tab:sample}
@@
         \citealt{scolnic2018a}). The colored parts represent the distribution of
-        SNe~Ia kept in our analysis for they are supposedly free from observational selection
-    bias (see Section~\ref{sec:sample}). The darker (resp. lighter) color
-responds to the conservative (resp. fiducial) selection cut.}
+        SNe~Ia kept in our analysis for they are supposedly free from
+    \textbf{observational} selection bias (see Section~\ref{sec:sample}). The
+darker (resp. lighter) color responds to the conservative (resp. fiducial)
+selection cut.}
     \label{fig:cuts}
@@
     set of magnitude-limited ones, using simple redshift cuts. This simplified
-    approach is statistically sub-optimal, but should be enough to test
-    our key question: whether redshift evolution of stretch is compatible with the
+    approach is statistically sub-optimal, but should be enough to test our key
+    question: whether redshift evolution of stretch is compatible with the
     \cite{rigault2020} model. However, the possibility remains that a complex
-    observational selection function related to spectroscopic follow-up efficiencies below our fiducial (or
-    even conservative) redshift cuts could still affect our sample, making it not fully
-volume-limited; this would, in turn, bias our conclusion on the astrophysical
-drift of the SNe~Ia population. We now look at this possibility.}
-
-\textbf{To test for the existence of potential leftover observational selection biases in our
-    sample, we compare the stretch and color distributions of the SNe~Ia
-    originating from different datasets having overlapping redshift ranges: these
-    distributions should be similar if they reflect the underlying parent population. 
-    Note that the redshift range has to be
-    narrow enough so that any drift would be negligible.}
+    observational selection function related to spectroscopic follow-up
+    efficiencies below our fiducial (or even conservative) redshift cuts could
+still affect our sample, making it not fully volume-limited; this would, in
+turn, bias our conclusion on the astrophysical drift of the SNe~Ia population.
+We now look at this possibility.}
+
+\textbf{To test for the existence of potential leftover observational selection
+    biases in our sample, we compare the stretch and color distributions of the
+    SNe~Ia originating from different datasets having overlapping redshift
+    ranges: these distributions should be similar if they reflect the underlying
+    parent population. Note that the redshift range has to be narrow enough so
+that any drift would be negligible.}
 
 \textbf{The two samples that overlap the most in redshift are PS1 and SDSS in
-    the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping subsample
-    consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus
-    the most likely to be affected by residual observational selection effects (see the
-    corresponding discussion in section~\ref{ssec:cuts}). Over that same
-    redshift range, PS1 has 52 SNe~Ia, which are in the lowest redshift bins and
-    thus unlikely to have any observational selection issue. To identify potential
-    inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib}
-    (upper panels) compares the stretch and color distribution of both these
-    surveys. The resulting Kolmogorov-Smirnov (KS) similarity test $p$-values
-($p >10\%$) do not support any inconsistency, in agreement with the
-visual impression from Fig.~\ref{fig:distrib}.}
+    the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
+    overlapping subsample consists of the 146 SNe~Ia at the high-redshift end of
+    SDSS and thus the most likely to be affected by residual observational
+    selection effects (see the corresponding discussion in
+    section~\ref{ssec:cuts}). Over that same redshift range, PS1 has 52 SNe~Ia,
+    which are in the lowest redshift bins and thus unlikely to have any
+    observational selection issue. To identify potential inconsistency between
+    the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib} (upper panels) compares
+    the stretch and color distribution of both these surveys. The resulting
+Kolmogorov-Smirnov (KS) similarity test $p$-values ($p >10\%$) do not support
+any inconsistency, in agreement with the visual impression from
+Fig.~\ref{fig:distrib}.}
 
@@
     conclusion can be drawn: there is no substantial sign of discrepancy in the
-    stretch and color distributions between the low- and high-end of our fiducial SNLS
-    and PS1 samples, respectively. Nonetheless, the small size of the SNLS
-    dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of
-    this test, and only a strong deviation would be noticeable. Extending the
-redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
-allows to increase the SNLS subsample to 31, yet the stretch $p$-values remains
-high (34\%) showing no sign of inconsistency.}
+    stretch and color distributions between the low- and high-end of our
+    fiducial SNLS and PS1 samples, respectively. Nonetheless, the small size of
+    the SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the
+    sensitivity of this test, and only a strong deviation would be noticeable.
+Extending the redshift range to $0.20 < z < 0.40$ (though we have no PS1 data
+above 0.3) allows to increase the SNLS subsample to 31, yet the stretch
+$p$-values remains high (34\%) showing no sign of inconsistency.}
 
@@
 
-\textbf{We finally highlight that the SNe~Ia color is more prone to observational selection
-    effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
-    Fig.~3 of \cite{kessler2017}. Hence, since the comparison of color
-    distributions shows no significant hint of leftover observational selection effect, it
-    further supports our claim that our simple redshift-based selection criteria
-    are sufficient to build the complete SNe~Ia samples required to test the
-    redshift evolution of the stretch distribution.}
+\textbf{We finally highlight that the SNe~Ia color is more prone to
+    observational selection effects than stretch as illustrated in
+    Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of \cite{kessler2017}. Hence,
+    since the comparison of color distributions shows no significant hint of
+    leftover observational selection effect, it further supports our claim that
+    our simple redshift-based selection criteria are sufficient to build the
+    complete SNe~Ia samples required to test the redshift evolution of the
+stretch distribution.}
 
@@
 
-\textbf{In order to ensure that our results are not driven by the incompletely-modeled
-    HST subsample, we recompute $\Delta$AIC for each model excluding this
-    dataset; we find that it does not change $\Delta$AIC by more than few
-    tenths. The consistency of these values with those in Table~\ref{tab:comp}
-show that the HST subsample does not drive our conclusions.}
-
-We report in Table~\ref{tab:bbc} \textbf{our determination of}
-$\mu_0$ and $\sigma_{\pm}$ \textbf{for each sample when implementing an asymmetric Gaussian
+\textbf{In order to ensure that our results are not driven by the
+    incompletely-modeled HST subsample, we recompute $\Delta$AIC for each model
+    excluding this dataset; we find that it does not change $\Delta$AIC by more
+    than few tenths. The consistency of these values with those in
+    Table~\ref{tab:comp} show that the HST subsample does not drive our
+conclusions.}
+
+We report in Table~\ref{tab:bbc} \textbf{our determination of} $\mu_0$ and
+$\sigma_{\pm}$ \textbf{for each sample when implementing an asymmetric Gaussian
 model}, adjusted on the nominally selection-free samples using our fiducial cuts
@@
 derived these model parameters using the full BBC formalism, using numerous
-simulations to model the observational selection effects (see details e.g., Section~3 of
-\citealt{kessler2017}). The agreement between our fit of the asymmetric
-Gaussians on the supposedly selection-free part of the samples and the results
-derived using the BBC formalism supports our approach to \textbf{constructing} a
-sample with negligible observational selection effects. If we were to use \cite{scolnic2016}
-and \cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
-parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
-Base drifting model and the BBC modeling would go even deeper from $-32$ to
-$-47$. We further discuss the consequence of this result for cosmology in
-Section~\ref{sec:discussion}.
+simulations to model the \textbf{observational} selection effects (see details
+e.g., Section~3 of \citealt{kessler2017}). The agreement between our fit of the
+asymmetric Gaussians on the supposedly selection-free part of the samples and
+the results derived using the BBC formalism supports our approach to
+\textbf{constructing} a sample with negligible \textbf{observational} selection
+effects. If we were to use \cite{scolnic2016} and \cite{scolnic2018a} best fit
+values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for SNLS, SDSS and
+PS1, respectively, the $\Delta$AIC between our Base drifting model and the BBC
+modeling would go even deeper from $-32$ to $-47$. We further discuss the
+consequence of this result for cosmology in Section~\ref{sec:discussion}.
 
@@
 indeed appear to share the same underlying high-stretch mode. Furthermore, one
-might wonder whether a low-stretch mode might also exist in the
-young-population (see Fig.~\ref{fig:stretchlssfr}). We \textbf{test for this} by
-allowing this population to also be bimodal, finding the amplitude of
-\textbf{such a} low-stretch mode to be compatible with~0 ($<2\%$) in this young
-population. More generally, this raises the question of \textbf{how
-well a given environmental tracer (here LsSFR) traces the age}. An
-analysis dedicated to this question will be presented in Briday et al.\ in prep.
+might wonder whether a low-stretch mode might also exist in the young-population
+(see Fig.~\ref{fig:stretchlssfr}). We \textbf{test for this} by allowing this
+population to also be bimodal, finding the amplitude of \textbf{such a}
+low-stretch mode to be compatible with~0 ($<2\%$) in this young population. More
+generally, this raises the question of \textbf{how well a given environmental
+tracer (here LsSFR) traces the age}. An analysis dedicated to this question will
+be presented in Briday et al.\ in prep.
 
@@
 dataset (see Section~\ref{sec:modeling}) --- reduces the significance of the
-results presented in this section, as expected. Even so, non-drifting models remain
-strongly disfavored. For instance, the best fitting sample-based Gaussian
-asymmetric model is still $\Delta\mathrm{AIC}<-10$ less representative of the
-data than our Base drifting model.
+results presented in this section, as expected. \textbf{Even so}, non-drifting
+models remain strongly disfavored. \textbf{For} instance, the best
+\textbf{fitting} sample-based Gaussian asymmetric \textbf{model} is still
+$\Delta\mathrm{AIC}<-10$ less representative of the data than our Base drifting
+\textbf{model}.
 
@@
 However, surveys usually do have significant Malmquist bias for the upper half
-of their SN redshift distribution. As a consequence, mismodeling of the
+of their SN redshift distribution. As a consequence, \textbf{mismodeling} of the
 underlying stretch distribution will bias the SN magnitudes derived from such
@@
 Commonly used Malmquist bias correction techniques, such as the BBC-formalism,
-assume per sample asymmetric Gaussian functions to model the underlying
+assume per sample asymmetric Gaussian functions \textbf{to} model the underlying
 stretch and color distributions. Yet, as shown in Section~\ref{sec:results},
-such a sample-based distribution is excluded in comparison to our drifting
-model. Then, contrary to what \citet[][Section~2]{scolnic2016} and
-\citet[][Section~5.4]{scolnic2018a} have suggested --- i.e., that traditional surveys
-span sufficiently limited redshift ranges such that the per-sample approach accounts
-for implicit redshift drifts --- a direct modeling of the redshift drift is more
-appropriate than a sample-based approach. We \textbf{add} here that, as
-measurements of modern surveys try to cover increasingly larger redshift ranges
-in order to reduce calibration systematic uncertainties, this sample-based
-approach \textbf{becomes} less valid, notably for PS1, DES and, soon, LSST.
+such a sample-based distribution is excluded \textbf{in comparison to} our
+drifting model. Then, \textbf{contrary to} what \citet[][Section~2]{scolnic2016}
+and \citet[][Section~5.4]{scolnic2018a} have suggested --- i.e., that
+traditional surveys span \textbf{sufficiently} limited redshift ranges
+\textbf{such that} the per-sample approach accounts for implicit redshift drifts
+--- a direct modeling of the redshift drift is more appropriate than a
+sample-based approach. We \textbf{add} here that, as measurements of modern
+surveys try to cover increasingly larger redshift ranges in order to reduce
+calibration systematic uncertainties, this sample-based approach
+\textbf{becomes} less valid, notably for PS1, DES and, soon, LSST.
 
@@
 lower redshift), the higher the amplitude of the old-specific low-stretch mode.
-This redshift dependency of the underlying stretch distributions is shown running from blue to red in
-Fig.~\ref{fig:bbc_pdf_ps1} for the redshift range covered by PS1. The observed
-$x_1$ histogram follows \textbf{the model we defined using} the sum of
-individual underlying SN-redshift distributions. As expected, the two modeling
-approaches differ mostly in the negative part of the SN stretch distribution.
-The asymmetric Gaussian distribution goes through the middle of the bimodal
-distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
-under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting model
-for typical PS1 SN redshifts. This means that the SN bias-corrected standardized
-magnitude estimated at a redshift suffering from by observational selection effects would be biased
-by an mismodeling of the true underlying stretch distribution.
+This redshift dependency \textbf{of the underlying stretch distributions is
+shown running from} blue to red in Fig.~\ref{fig:bbc_pdf_ps1} for \textbf{the}
+redshift range covered by PS1. The observed $x_1$ histogram follows \textbf{the
+model we defined using} the sum of individual underlying SN-redshift
+distributions. As expected, the two modeling approaches differ mostly in the
+negative part of the SN stretch distribution. The asymmetric Gaussian
+distribution goes through the middle of the bimodal distribution,
+over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and under-estimating it at
+$x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN
+redshifts. This means that the SN bias-corrected standardized magnitude
+estimated at a redshift \textbf{suffering} from by \textbf{observational}
+selection effects would be biased by \textbf{a mismodeling} of the true
+underlying stretch distribution.
 
@@
 redshift drift should be further studied and care should be taken when using
-samples affected by observational selection effects.
+samples affected by \textbf{observational} selection effects.
 
@@
 
-We have presented \textbf{an initial} study of the drift of the underlying SNe~Ia
-stretch distribution as a function of redshift. We built effectively volume-limited
-SN~Ia subsamples from the Pantheon dataset \citep[][SDSS, PS1 and
-SNLS]{scolnic2018a}, to which we added HST and SNfactory data from
-\cite{rigault2020} for the high- and low-redshift bins. We only considered the
-SNe that have been discovered in the redshift range of each survey where
-observational selection effects are negligible, so that the observed SNe~Ia stretches are
-a random sampling of the true underlying distribution. This resulted in a fiducial sample 
-of 569 SNe~Ia (422 SNe when more conservative cuts were applied).
+We have presented \textbf{an initial} study of the drift of the underlying
+SNe~Ia stretch distribution as a function of redshift. We built
+\textbf{effectively volume-limited SN~Ia subsamples} from the Pantheon dataset
+\citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to which we added HST and SNfactory
+data from \cite{rigault2020} for the high- and low-redshift bins. We only
+considered the SNe that have been discovered in the redshift range of each
+survey where \textbf{observational} selection effects are negligible, so that
+the observed SNe~Ia stretches are a random sampling of the true underlying
+distribution. This resulted in a \textbf{fiducial} sample of 569 SNe~Ia (422 SNe
+when more conservative cuts were \textbf{applied}).
 
@@
 
-In addition to this ``base'' model, we have studied various distributions,
-including redshift independent models; we also studied the prediction from a
-per-sample asymmetric Gaussian stretch distribution used, for instance, by the
-Beams with Bias Correction Malmquist bias correction algorithm
+In addition to this ``base'' \textbf{model}, we have studied various
+distributions, including redshift independent models; we also studied the
+prediction from a per-sample asymmetric Gaussian stretch distribution used, for
+instance, by the Beams with Bias Correction Malmquist bias correction algorithm
 \citep{scolnic2016, kessler2017}.
@@
 \begin{enumerate}
+
     \item The underlying SN~Ia stretch distribution is significantly redshift
         dependent, as previously suggested by e.g.~\cite{howell2007}, \textbf{in
-        a way that observational selection effects alone cannot explain}. This result is
-        largely independent of details on each age-population model.
+        a way that observational selection effects alone cannot explain}. This
+        result is largely independent of details on each age-population model.
 
@@
         the relative fraction of younger and older SNe~Ia follows the prediction
-        made in \cite{rigault2020}. This second result further supports the
-        existence of both young and old SN~Ia populations, in agreement with
+        made in \cite{rigault2020}. This second result \textbf{further} supports
+        the existence of both young and old SN~Ia populations, in agreement with
         rate studies \cite{mannucci2005, scannapieco2005, sullivan2006,
@@
 
-    \item Models using survey-based asymmetric Gaussian distributions, e.g., as employed
-        in the current implementation of BBC, are excluded \textbf{as a}
-        good description of the data relative to our drifting model. Hence, the
-        sample-based approach does not accurately account for redshift drift, a problem
-        that will be exacerbated as surveys span increasingly larger redshift ranges.
-        \textbf{As a result, even if the necessary} extra degrees of freedom might be acceptable
-        given the large number of SNe~Ia in cosmological studies, extrapolating
-        the SN property distributions from the volume-limited part of a survey
-        to its Malmquist-biased magnitude-limited part would still be inaccurate
+    \item Models using survey-based asymmetric Gaussian distributions,
+        \textbf{e.g., as employed} in the current implementation of BBC, are
+        excluded \textbf{as a} good description of the data relative to our
+        drifting model. Hence, the sample-based approach does not accurately
+        account for redshift drift, \textbf{a problem that will be exacerbated
+        as surveys} span increasingly larger redshift ranges. \textbf{As a
+    result, even if the necessary} extra degrees of freedom might be acceptable
+    given the large number of SNe~Ia in cosmological studies, extrapolating the
+    SN property distributions from the volume-limited part of a survey to its
+    Malmquist-biased magnitude-limited \textbf{part} would still be inaccurate
     because of the redshift evolution.
@@
 \textbf{The next step in this line of analysis will incorporate our model into
-the SNANA framework \citep{SNANA}, both to more accurately account for observational selection
-functions and to test the impact of our model on the derivation of
-cosmological parameters; this study is underway.}
+the SNANA framework \citep{SNANA}, both to more accurately account for
+observational selection functions and to test the impact of our model on the
+derivation of cosmological parameters; this study is underway.}
 

### 2021-02-08 Nicolas.tex
Two columns for us 8-)
Rid of textbf, 5sigma in abstract, 1 column for referee

--- 2021-01-15 Nicolas.tex
@@
 \abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
-    and as survey statistics increase, the question of astrophysical
-    systematic uncertainties rises, notably that of SN~Ia population
-    evolution. In this paper, we study the dependence with redshift of the SN~Ia
-    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
-    probe its potential redshift drift. The SN stretch has been shown to
-    strongly correlate with the SN environment, notably with stellar age
-    tracers. We model the underlying stretch distribution as a function of
-    redshift, using the evolution of the fraction of young and old SNe~Ia as
-    predicted by \cite{rigault2020}, and assuming constant underlying stretch
-    distribution for each age population \textbf{consisting} of Gaussian
-    mixtures. We test our prediction against published samples \textbf{cut} to
-    have marginal magnitude selection effects, so that any observed change is
-    indeed of astrophysical and not observational in origin. \textbf{In this first
-    study, there are indications} that the underlying SN~Ia stretch distribution
-    is evolving as a function of redshift, and that the young/old drifting model
-    is a better description of the data than any time-constant model, including
-    the sample-based asymmetric distributions often used to correct Malmquist
-    bias. The favored underlying stretch model is the bimodal one derived from
-    \cite{rigault2020}, \textbf{composed of} a high-stretch mode shared by both
-    young and old environments, and a low-stretch mode exclusive to old
-    environments. The precise impact of the redshift evolution of the SN Ia
-    population intrinsic properties on cosmology remains to be studied. Yet, the
-    astrophysical drift of the SN stretch distribution does affect current
-    Malmquist bias corrections and thereby distances derived from SN affected by
-observational selection effects. We highlight that such a bias will increase with surveys
+    and as survey statistics increase, the question of astrophysical systematic
+    uncertainties rises, notably that of SN~Ia population evolution. In this
+    paper, we study the dependence with redshift of the SN~Ia \texttt{SALT2.4}
+    lightcurve stretch, a purely intrinsic SN property, to probe its potential
+    redshift drift. The SN stretch has been shown to strongly correlate with the
+    SN environment, notably with stellar age tracers. We model the underlying
+    stretch distribution as a function of redshift, using the evolution of the
+    fraction of young and old SNe~Ia as predicted by \cite{rigault2020}, and
+    assuming constant underlying stretch distribution for each age population
+    consisting of Gaussian mixtures. We test our prediction against published
+    samples cut to have marginal magnitude selection effects, so that any
+    observed change is indeed of astrophysical and not observational in origin.
+    In this first study, there are indications that the underlying SN~Ia stretch
+    distribution is evolving as a function of redshift, and that the young/old
+    drifting model is a better description of the data than any time-constant
+    model, including the sample-based asymmetric distributions often used to
+    correct Malmquist bias \textbf{at a significance of over 5 $\sigma$}. The
+    favored underlying stretch model is the bimodal one derived from
+    \cite{rigault2020}, composed of a high-stretch mode shared by both young and
+    old environments, and a low-stretch mode exclusive to old environments. The
+    precise impact of the redshift evolution of the SN Ia population intrinsic
+    properties on cosmology remains to be studied. Yet, the astrophysical drift
+    of the SN stretch distribution does affect current Malmquist bias
+    corrections and thereby distances derived from SN affected by observational
+selection effects. We highlight that such a bias will increase with surveys
 covering increasingly larger redshift ranges, which is particularly important
@@
 \citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ derived when
-the SNe~Ia are anchored to Cepheids \citep[the SH0ES project,][]{riess2009,
-riess2016} is $\sim5\sigma$ higher than what is predicted from cosmic microwave
-background (CMB) data measured by Planck assuming the standard $\L$CDM
-\citep{planck2018, riess2019, reid2019}, or when the SN luminosity is anchored
-at intermediate redshift by the baryon acoustic oscillation (BAO) scale
-\citep{feeney2019}. While using the tip of the red giant branch technique in
-place of the Cepheids seem to favor an intermediate value of $H_0$
-\citep{freedman2019, freedman2020}, time delay measurements from strong lensing
-seem to also favor high $H_0$ values \citep{wong2019}.
+the SNe~Ia are anchored to Cepheids \citep[the SH0ES
+project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
+from cosmic microwave background (CMB) data measured by Planck assuming the
+standard $\L$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
+luminosity is anchored at intermediate redshift by the baryon acoustic
+oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
+branch technique in place of the Cepheids seem to favor an intermediate value of
+$H_0$ \citep{freedman2019, freedman2020}, time delay measurements from strong
+lensing seem to also favor high $H_0$ values \citep{wong2019}.
 
@@
 fundamental physics. Yet, no simple solution is able to accommodate this $H_0$
-tension when accounting for all other probes \citep{knox2019}. Alternatively, systematic effects affecting one or several
-of the aforementioned analyses might explain at least some of this tension.
-\cite{rigault2015} suggested that SNe Ia from the Cepheid calibrator sample
-differ by construction from those in the Hubble flow sample, as the former
-strongly favors young stellar populations, while the latter does not. This
-selection effect would impact the derivation of $H_0$ if SNe~Ia from young and
-older environments differ in average standardized magnitudes.
-
-%For the last decades, numerous analyses have studied 
-The relation between SNe~Ia
-and their host galaxy properties has been studied extensively. The first key finding
-was that the standardized SNe~Ia magnitudes
-significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass host 
-galaxies being
-brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
-betoule2014, rigault2020, kim19}. This mass-step correction is currently used in
-cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including for
-deriving $H_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
-between the SNe and their host galaxies remains unclear, particularly when using global properties such
-as the host stellar mass. This in turn raises the question of the accuracy of such
-corrections since host galaxy global properties evolve with redshift. 
-More recently, studies have used the local SN host galaxy environment to probe
-more direct connections between the SNe and their host galaxy environment \citep{rigault2013},
+tension when accounting for all other probes \citep{knox2019}. Alternatively,
+systematic effects affecting one or several of the aforementioned analyses
+might explain at least some of this tension. \cite{rigault2015}
+suggested that SNe Ia from the Cepheid calibrator sample differ by construction
+from those in the Hubble flow sample, as the former strongly favors
+young stellar populations, while the latter does not. This selection
+effect would impact the derivation of $H_0$ if SNe~Ia from young and older
+environments differ in average standardized magnitudes.
+
+The relation between SNe~Ia and their host galaxy properties has been
+studied extensively. The first key finding was that the standardized SNe~Ia
+magnitudes significantly depend on the host galaxy stellar mass, with
+SNe~Ia from high-mass host galaxies being brighter on average
+\cite[e.g.][]{kelly2010, sullivan2010, childress2013, betoule2014, rigault2020,
+kim19}. This mass-step correction is currently used in cosmological analyses
+\citep[e.g.][]{betoule2014, scolnic2018a}, including for deriving $H_0$
+\citep{riess2016, riess2019}. Yet, the underlying connection between the SNe and
+their host galaxies remains , particularlyunclear when using
+global properties such as the host stellar . This in turnmass raises
+the question of the accuracy of such corrections since host galaxy
+global properties evolve with redshift. More recently, studies have used the
+local SN host galaxy environment to probe more direct connections
+between the SNe and their host galaxy environments \citep{rigault2013},
 showing that local age tracers such as the Local specific Star Formation Rate
@@
 magnitude \citep{roman2018, kim18,rigault2020}. Further reinforcing this
-connection is the identification of SN~Ia spectral features that are 
-correlated with LsSFR \citep{nordin2018}. These results suggest age as the driving
+connection is the identification of SN~Ia spectral features that are correlated
+with LsSFR \citep{nordin2018}. These results suggest age as the driving
 parameter underlying the mass-step. If true, this would have a significant
@@
 following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
-model). \cite{rigault2020} used the LsSFR to classify which are the
-younger (those with a high LsSFR) and which are the older (those with low
-LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN stretch has
-been known to be strongly correlated with the SN host galaxy properties
-\citep{hamuy1996, hamuy2000}. This correlation that has been extensively confirmed
-since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
+model). \cite{rigault2020} used the LsSFR to classify which are the younger
+(those with a high LsSFR) and which are the older (those with low LsSFR).
+Furthermore, since the first SNe~Ia host analyses, the SN stretch has been known
+to be strongly correlated with the SN host galaxy properties
+\citep{hamuy1996, hamuy2000}. This correlation that has been extensively
+confirmed since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
 gupta2011, dandrea2011, childress2013, rigault2013, pan2014, kim19}. Following
-the ``A+B'' model and the connection between SN stretch and host galaxy properties,
-\cite{howell2007} first discussed the potential redshift drift of the SN stretch
-distribution. In this paper we revisit this question using the most recent SNe~Ia
-datasets.
-
-In this paper, we take a step aside \mr{from the cosmological analyses} to probe the validity of our modeling of the
-SN population, which we claim to be constituted of two age-populations
-\citep{rigault2013,rigault2015,rigault2020}: one old and one younger, the former
-having on average lower lightcurve stretches and being brighter after
-standardization. We use the correlation between the SN age, as probed by the
-LsSFR, and the SN stretch to model the expected evolution of the underlying SN
-stretch distribution as a function of redshift. This modeling relies on three
-assumptions: (1) there are two distinct populations of SNe~Ia; (2) the relative
-fraction of each of these populations as a function of redshift follows the
-model presented in \cite{rigault2020} and (3) the underlying distribution of
-stretch for each age sample is constant. This paper aims at testing this
-specific model with datasets from the literature. 
-% Note that the progenitor age
-% as traced by the LsSFR seems to capture physical feature intrinsic to the
-%progenitor and/or explosion mechanism that the stretch alone is not capturing
-% \citep{nordin2018}.
+the ``A+B'' model and the connection between SN stretch and host galaxy
+properties, \cite{howell2007} first discussed the potential redshift drift of
+the SN stretch distribution. In this paper we revisit this question
+using the most recent SNe~Ia datasets.
+
+In this paper, we take a step aside from the cosmological analyses to
+probe the validity of our modeling of the SN population, which we claim to be
+constituted of two age-populations \citep{rigault2013,rigault2015,rigault2020}:
+one old and one younger, the former having on average lower lightcurve stretches
+and being brighter after standardization. We use the correlation between the SN
+age, as probed by the LsSFR, and the SN stretch to model the expected evolution
+of the underlying SN stretch distribution as a function of redshift. This
+modeling relies on three assumptions: (1) there are two distinct populations of
+SNe~Ia; (2) the relative fraction of each of these populations as a function of
+redshift follows the model presented in \cite{rigault2020} and (3) the
+underlying distribution of stretch for each age sample is constant. This paper
+aims at testing this specific model with datasets from the literature.
 
@@
 
-The ideal SN~Ia sample for studying this question would be a very deep, large-area,
-volume-limited sample. This would capture the true underlying stretch distribution
-function, and we would then study how it evolves with redshift. No such sample exists, 
-so first we must construct subsamples that are as near to volume-limited as possible 
-from existing high-redshift SN~Ia samples.
+The ideal SN~Ia sample for studying this question would be a very deep,
+    large-area, volume-limited sample. This would capture the true underlying
+    stretch distribution function, and we would then study how it evolves with
+    redshift. No such sample exists, so first we must construct subsamples that
+are as near to volume-limited as possible from existing high-redshift SN~Ia
+samples.
 
@@
 distributions in a few bins of redshift. In practice, however, differential
-observational selection effects will affect the observed SN stretch distributions. Indeed,
-because the observed SN~Ia magnitude correlates with the lightcurve stretch (and
-color), the first SNe~Ia that a magnitude-limited survey will miss are the
-lowest-stretch (and reddest) ones. Consequently, if magnitude-related observational selection
-effects are not accounted for, one might confuse true population drift with
-survey properties, and conversely.
-
-Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring SN types
-and host galaxy redshifts, the observation selection effects of magnitude-limited surveys should be
-negligible below a given redshift at which even the faintest normal SNe~Ia can be
-observed. \textbf{Aiding in the construction of nearly volume-limited 
-subsamples is the fact that the SN~Ia population trails off towards fainter SNe~Ia.
-A complication is that complete spectroscopic follow-up has not always been
-the norm, as discussed below.}
-In contrast, targeted surveys have highly complex observational
-selection functions and so are discarded from our analysis. High-redshift SN cosmology
-samples, such as Pantheon, are predominately from magnitude-limited surveys from
+observational selection effects will affect the observed SN
+stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
+with the lightcurve stretch (and color), the first SNe~Ia that a
+magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
+Consequently, if magnitude-related observational selection effects are
+not accounted for, one might confuse true population drift with survey
+properties, and conversely.
+
+Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring
+SN types and host galaxy redshifts, the observation selection
+effects of magnitude-limited surveys should be negligible below a given redshift
+at which even the faintest normal SNe~Ia can be observed.
+Aiding in the construction of nearly volume-limited subsamples is the
+    fact that the SN~Ia population trails off towards fainter SNe~Ia. A
+    complication is that complete spectroscopic follow-up has not always been
+    the norm, as discussed below. In contrast, targeted surveys have highly
+    complex observational selection functions and so are
+    discarded from our analysis. High-redshift SN cosmology samples,
+    such as Pantheon, are predominately from magnitude-limited surveys from
 which volume-limited SN~Ia subsamples can be constructed.
-%\textbf{which have easier to model observational selection functions}.
 
@@
 c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
-parent distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
-\citet{campbell2013} for similar contours, the second using a more conservative
-$|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is
-$M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
-standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
-along the aforementioned ellipse given the standardization coefficient
-$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is
-that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
-in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one ought to detect
-this object typically 5 days before and a week after peak to build a suitable
-lightcurve, the effective limiting standardized absolute magnitude is
-approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit
-$m_{\lim}$ of a magnitude limited survey, one can derive the maximum redshift
-$z_{\lim}$ above which the faintest SNe~Ia will be missed using the relation
-between apparent magnitude, redshift and absolute magnitude $\mu(z_{\lim}) =
-m_{\lim} - M_{\lim}$.
+parent distribution \citep{guy2007, betoule2014}; see also
+\citet{bazin2011} and \citet{campbell2013} for similar contours, the second
+using a more conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude
+with $x_1=0$ and $c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can
+derive the absolute standardized magnitude at maximum of light $M = M_0 - \alpha
+x_1 + \beta c$ along the aforementioned ellipse given the standardization
+coefficient $\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the
+faintest SN~Ia is that with $(x_1=-1.65, c=0.25)$ and an absolute standardized
+magnitude at peak in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one
+ought to detect this object typically 5 days before and a week after peak to
+build a suitable lightcurve, the effective limiting standardized absolute
+magnitude is approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude
+limit $m_{\lim}$ of a magnitude limited survey, one can derive the maximum
+redshift $z_{\lim}$ above which the faintest SNe~Ia will be missed using the
+relation between apparent magnitude, redshift and absolute magnitude
+$\mu(z_{\lim}) = m_{\lim} - M_{\lim}$.
 
@@
 
-\textbf{We will thus consider a set of cuts that will define a first fiducial
+We will thus consider a set of cuts that will define a first fiducial
     sample, taking the limits as initially suggested by the previous
-    completeness analysis. \textbf{However,} as this solution might be an overly
-    simplified way to create a complete sample (e.g., because it ignores spectroscopic follow-up
-    in efficiency for redshifts below $z_{lim}$), we also consider another set of cuts to define a so-called
-    ``conservative'' sample. The latter is smaller and therefore less statistically
-    constraining, but also even less prone to observational selection effects. If the
-    redshift drift is still significant in the conservative sample, it would be
-    even more meaningful in a carefully-tailored selection-free sample. These
-    samples are adequate for the goal
-    of this study, which is to develop a first implementation of a model for drift in
-    SN Ia properties. If fruitful, the sample selection can later be refined, if needed, with a
-    more detailed
-    %drift modeling before refining it on a more complex, yet more accurate,
-%follow-up analysis also accounting for precise 
-model of the observational selection, e.g., using
-the SNANA package \citep{SNANA}.}
+    completeness analysis. However, as this solution might be an overly
+    simplified way to create a complete sample, e.g., because it ignores
+    spectroscopic follow-up in efficiency for redshifts below $z_{lim}$, we
+    also consider another set of cuts to define a so-called ``conservative''
+    sample. The latter is smaller and therefore less statistically constraining,
+    but also even less prone to observational selection effects. If the redshift
+    drift is still significant in the conservative sample, it would be even more
+    meaningful in a carefully-tailored selection-free sample. These samples are
+    adequate for the goal of this study, which is to develop a first
+    implementation of a model for drift in SN Ia properties. If fruitful, the
+    sample selection can later be refined, if needed, with a more detailed model
+of the observational selection, e.g., using the SNANA package \citep{SNANA}.
 
@@
 \citet[][see their Section~5]{perrett2010} suggests however a lower limit of
-$z_{\lim}=0.55$. \textbf{We will therefore use $z=0.60$ and $z=0.55$ as redshift
-limits for the fiducial and conservative samples, respectively, for SNLS.}
+$z_{\lim}=0.55$. We will therefore use $z=0.60$ and $z=0.55$ as redshift
+limits for the fiducial and conservative samples, respectively, for SNLS.
 
@@
 $5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields $z_{\lim}=0.31$, in
-agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. \textbf{If we were to be
+agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. If we were to be
     conservative, this figure would also suggest of a more stringent
 $z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
-conservative samples, respectively, for PS1.}
+conservative samples, respectively, for PS1.
 
@@
 \citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
-SDSS surveys \textbf{had to contend with limited spectroscopic resources. As
-discussed in} \citet[][Section~2]{kessler2009}, during the first year of SDSS,
+SDSS surveys had to contend with limited spectroscopic resources. As
+discussed in \citet[][Section~2]{kessler2009}, during the first year of SDSS,
 SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
@@
 spectroscopic resources were available, and \cite{kessler2009} and
-\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Based on this,
-\textbf{we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
-fiducial and conservative samples, respectively, for SDSS.}
+\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$.
+Based on this, we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
+fiducial and conservative samples, respectively, for SDSS.
 
@@
 distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
-expected, the selected redshift limits \textbf{are roughly located slightly
-before the peak of these histograms. In Section \ref{ssec:verify} we validate that these
-redshift limits are effective for constructing nearly volume-limited subsamples
-from samples that were initially more closely magnitude-limited in their search
-or spectroscopic follow-up.}
+expected, the selected redshift limits are roughly located slightly
+    before the peak of these histograms. In Section \ref{ssec:verify} we
+validate that these redshift limits are effective for constructing nearly
+volume-limited subsamples from samples that were initially more closely
+magnitude-limited in their search or spectroscopic follow-up.
 
@@
     \caption{Composition of the SNe~Ia dataset used in this analysis.
-    Conservative cuts are indicated in parentheses. \textbf{The SNf limit is set by
-\cite{rigault2020}, see text.}}
+    Conservative cuts are indicated in parentheses. The SNf limit is set
+by \cite{rigault2020}, see text.}
     \label{tab:sample}
@@
         \citealt{scolnic2018a}). The colored parts represent the distribution of
-        SNe~Ia kept in our analysis for they are supposedly free from observational selection
-    bias (see Section~\ref{sec:sample}). The darker (resp. lighter) color
-responds to the conservative (resp. fiducial) selection cut.}
+        SNe~Ia kept in our analysis for they are supposedly free from
+    observational selection bias (see Section~\ref{sec:sample}). The
+darker (resp. lighter) color responds to the conservative (resp. fiducial)
+selection cut.}
     \label{fig:cuts}
@@
 and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
-\citealt{aldering2020}). \textbf{For this dataset, spectroscopic screening was
+\citealt{aldering2020}). For this dataset, spectroscopic screening was
     done for candidates with $r \lesssim 19.5$; redshift cuts were then applied
@@
 < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} for extracting
-local host properties.} These 114 SNfactory SNe~Ia are thus in the
+local host properties. These 114 SNfactory SNe~Ia are thus in the
 volume-limited part of the survey (Aldering et al., in prep.), and are therefore
@@
 Finally, we include the HST sample from Pantheon \citep{strolger04}.
-\textbf{These high-redshift SNe are of great interest as they provide the
+These high-redshift SNe are of great interest as they provide the
     greatest leverage for testing evolution. While at these redshifts the
@@
     that, while compatible with it, our results are not dependent on the
-    inclusion of this dataset.}
+    inclusion of this dataset.
 
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe here
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. We observe here
 that the fraction of low-stretch SNe (typically $x_1 < -1$) appears to decrease as
-a function of redshift; this is confirmed in Fig.~\ref{fig:modelall}, in which}
-the evolution of the mean stretch is shown, \textbf{with} the data split in
+a function of redshift; this is confirmed in Fig.~\ref{fig:modelall}, in which
+the evolution of the mean stretch is shown, with the data split in
 redshift bins of regular sample size. We see that SNe~Ia at higher redshift have
@@
 redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
-stretch distribution is \textbf{evolving with redshift}.
+stretch distribution is evolving with redshift.
 
@@
 
-\textbf{In section~\ref{ssec:cuts}, we have built volume-limited samples from a
-    set of magnitude-limited ones, using simple redshift cuts. This simplified
-    approach is statistically sub-optimal, but should be enough to test
-    our key question: whether redshift evolution of stretch is compatible with the
-    \cite{rigault2020} model. However, the possibility remains that a complex
-    observational selection function related to spectroscopic follow-up efficiencies below our fiducial (or
-    even conservative) redshift cuts could still affect our sample, making it not fully
-volume-limited; this would, in turn, bias our conclusion on the astrophysical
-drift of the SNe~Ia population. We now look at this possibility.}
-
-\textbf{To test for the existence of potential leftover observational selection biases in our
-    sample, we compare the stretch and color distributions of the SNe~Ia
+In section~\ref{ssec:cuts}, we have built volume-limited samples from a set of
+magnitude-limited ones, using simple redshift cuts. This simplified approach is
+statistically sub-optimal, but should be enough to test our key question:
+whether redshift evolution of stretch is compatible with the \cite{rigault2020}
+model. However, the possibility remains that a complex observational selection
+function related to spectroscopic follow-up efficiencies below our fiducial (or
+even conservative) redshift cuts could still affect our sample, making it not
+fully volume-limited; this would, in turn, bias our conclusion on the
+astrophysical drift of the SNe~Ia population. We now look at this possibility.
+
+To test for the existence of potential leftover observational selection biases
+in our sample, we compare the stretch and color distributions of the SNe~Ia
 originating from different datasets having overlapping redshift ranges: these
-    distributions should be similar if they reflect the underlying parent population. 
-    Note that the redshift range has to be
-    narrow enough so that any drift would be negligible.}
-
-\textbf{The two samples that overlap the most in redshift are PS1 and SDSS in
-    the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping subsample
-    consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus
-    the most likely to be affected by residual observational selection effects (see the
-    corresponding discussion in section~\ref{ssec:cuts}). Over that same
+distributions should be similar if they reflect the underlying parent
+population. We note that the redshift range has to be narrow enough so that any
+drift would be negligible.
+
+The two samples that overlap the most in redshift are PS1 and SDSS in the
+redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping
+subsample consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus
+the most likely to be affected by residual observational selection effects (see
+the corresponding discussion in section~\ref{ssec:cuts}). Over that same
 redshift range, PS1 has 52 SNe~Ia, which are in the lowest redshift bins and
 thus unlikely to have any observational selection issue. To identify potential
-    inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib}
-    (upper panels) compares the stretch and color distribution of both these
-    surveys. The resulting Kolmogorov-Smirnov (KS) similarity test $p$-values
-($p >10\%$) do not support any inconsistency, in agreement with the
-visual impression from Fig.~\ref{fig:distrib}.}
-
-\textbf{We perform a similar analysis for PS1 and SNLS over the redshift range
-    $0.20 < z < 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same
-    conclusion can be drawn: there is no substantial sign of discrepancy in the
-    stretch and color distributions between the low- and high-end of our fiducial SNLS
-    and PS1 samples, respectively. Nonetheless, the small size of the SNLS
-    dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of
-    this test, and only a strong deviation would be noticeable. Extending the
-redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
-allows to increase the SNLS subsample to 31, yet the stretch $p$-values remains
-high (34\%) showing no sign of inconsistency.}
+inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib} (upper
+panels) compares the stretch and color distribution of both these surveys. The
+resulting Kolmogorov-Smirnov (KS) similarity test $p$-values ($p >10\%$) do not
+support any inconsistency, in agreement with the visual impression from
+Fig.~\ref{fig:distrib}.
+
+We perform a similar analysis for PS1 and SNLS over the redshift range $0.20 < z
+< 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same conclusion can be
+drawn: there is no substantial sign of discrepancy in the stretch and color
+distributions between the low- and high-end of our fiducial SNLS and PS1
+samples, respectively. Nonetheless, the small size of the SNLS dataset at $z <
+0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of this test, and only a
+strong deviation would be noticeable. Extending the redshift range to $0.20 < z
+< 0.40$ (though we have no PS1 data above 0.3) allows to increase the SNLS
+subsample to 31, yet the stretch $p$-values remains high (34\%) showing no sign
+of inconsistency.
 
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
-    \caption{\textbf{$x_1$ (left) and $c$ (right) distribution histograms of
+    \caption{$x_1$ (left) and $c$ (right) distribution histograms of
             different surveys overlapping in redshift. \textit{Facing up}: SDSS
@@
             distributions are not drawn from the same underlying
-    distributions}}
+    distributions}
     \label{fig:distrib}
@@
 
-\textbf{We finally highlight that the SNe~Ia color is more prone to observational selection
-    effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
-    Fig.~3 of \cite{kessler2017}. Hence, since the comparison of color
-    distributions shows no significant hint of leftover observational selection effect, it
-    further supports our claim that our simple redshift-based selection criteria
-    are sufficient to build the complete SNe~Ia samples required to test the
-    redshift evolution of the stretch distribution.}
+We finally highlight that the SNe~Ia color is more prone to
+    observational selection effects than stretch as illustrated in
+    Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of \cite{kessler2017}. Hence,
+    since the comparison of color distributions shows no significant hint of
+    leftover observational selection effect, it further supports our claim that
+    our simple redshift-based selection criteria are sufficient to build the
+    complete SNe~Ia samples required to test the redshift evolution of the
+stretch distribution.
 
@@
 \end{equation}
-with $K=0.87$. This model is comparable to the evolution \textbf{subsequently}
+with $K=0.87$. This model is comparable to the evolution subsequently
 predicted by \cite{childress2014} based on SN rates in galaxies depending on
@@
 by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
-of the underlying \textbf{stretch} distribution of SNe~Ia as a function of redshift $X_1(z)$ is
+of the underlying stretch distribution of SNe~Ia as a function of redshift $X_1(z)$ is
 given by:
@@
 \end{align}
-\textbf{This is our Base drifting model.}
+This is our Base drifting model.
 
@@
     \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
-        as a function of redshift. Markers show the \textbf{stretch plain mean
-        (with error estimated from scatter)} measured in redshift bins of equal
+        as a function of redshift. Markers show the stretch plain mean
+        (with error estimated from scatter) measured in redshift bins of equal
         sample size, indicated in light gray at the bottom of each redshift bin.
@@
         conservative samples, respectively. The orange horizontal line
-        represents the mean stretch \textbf{for the non-evolving Gaussian model
-        (last line of Table~\ref{tab:comp}) on the fiducial sample}. Best
+        represents the mean stretch for the non-evolving Gaussian model
+        (last line of Table~\ref{tab:comp}) on the fiducial sample. Best
         fits of our Base drifting model are shown as blue, dashed-blue and
@@
 
-\textbf{Note} that, when comparing models adjusted on individual subsamples
-rather than globally, the Bayesian Information Criterion ($\mathrm{BIC} =
--2\ln(L) + k\ln(n)$, with $n$ the number of data points) might be better suited
-than AIC, since it explicitly accounts for the fact that each subsample is
-fitted separately: the sample-based model BIC is rightfully the sum of the BIC
-for each sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the
-sample-based asymmetric Gaussian model as being as pertinent as the Base model.
-
-\textbf{In order to ensure that our results are not driven by the incompletely-modeled
-    HST subsample, we recompute $\Delta$AIC for each model excluding this
-    dataset; we find that it does not change $\Delta$AIC by more than few
-    tenths. The consistency of these values with those in Table~\ref{tab:comp}
-show that the HST subsample does not drive our conclusions.}
-
-We report in Table~\ref{tab:bbc} \textbf{our determination of}
-$\mu_0$ and $\sigma_{\pm}$ \textbf{for each sample when implementing an asymmetric Gaussian
-model}, adjusted on the nominally selection-free samples using our fiducial cuts
+We note that, when comparing models adjusted on individual subsamples rather
+than globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
+k\ln(n)$, with $n$ the number of data points) might be better suited than AIC,
+since it explicitly accounts for the fact that each subsample is fitted
+separately: the sample-based model BIC is rightfully the sum of the BIC for each
+sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the sample-based
+asymmetric Gaussian model as being as pertinent as the Base model.
+
+In order to ensure that our results are not driven by the incompletely-modeled
+HST subsample, we recompute $\Delta$AIC for each model excluding this dataset;
+we find that it does not change $\Delta$AIC by more than few tenths. The
+consistency of these values with those in Table~\ref{tab:comp} show that the HST
+subsample does not drive our conclusions.
+
+We report in Table~\ref{tab:bbc} our determination of $\mu_0$ and
+$\sigma_{\pm}$ for each sample when implementing an asymmetric Gaussian
+model, adjusted on the nominally selection-free samples using our fiducial cuts
 (see Section~\ref{sec:sample}). We find our results in close agreement with
@@
 derived these model parameters using the full BBC formalism, using numerous
-simulations to model the observational selection effects (see details e.g., Section~3 of
-\citealt{kessler2017}). The agreement between our fit of the asymmetric
-Gaussians on the supposedly selection-free part of the samples and the results
-derived using the BBC formalism supports our approach to \textbf{constructing} a
-sample with negligible observational selection effects. If we were to use \cite{scolnic2016}
-and \cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
-parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
-Base drifting model and the BBC modeling would go even deeper from $-32$ to
-$-47$. We further discuss the consequence of this result for cosmology in
-Section~\ref{sec:discussion}.
+simulations to model the observational selection effects (see details
+e.g., Section~3 of \citealt{kessler2017}). The agreement between our fit of the
+asymmetric Gaussians on the supposedly selection-free part of the samples and
+the results derived using the BBC formalism supports our approach to
+constructing a sample with negligible observational selection
+effects. If we were to use \cite{scolnic2016} and \cite{scolnic2018a} best fit
+values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for SNLS, SDSS and
+PS1, respectively, the $\Delta$AIC between our Base drifting model and the BBC
+modeling would go even deeper from $-32$ to $-47$. We further discuss the
+consequence of this result for cosmology in Section~\ref{sec:discussion}.
 
@@
 corresponding fit is not significantly better, with a $\Delta$AIC of $-0.4$.
-This \textbf{reinforces our assumption} that the young and old populations
+This reinforces our assumption that the young and old populations
 indeed appear to share the same underlying high-stretch mode. Furthermore, one
-might wonder whether a low-stretch mode might also exist in the
-young-population (see Fig.~\ref{fig:stretchlssfr}). We \textbf{test for this} by
-allowing this population to also be bimodal, finding the amplitude of
-\textbf{such a} low-stretch mode to be compatible with~0 ($<2\%$) in this young
-population. More generally, this raises the question of \textbf{how
-well a given environmental tracer (here LsSFR) traces the age}. An
-analysis dedicated to this question will be presented in Briday et al.\ in prep.
+might wonder whether a low-stretch mode might also exist in the young-population
+(see Fig.~\ref{fig:stretchlssfr}). We test for this by allowing this
+population to also be bimodal, finding the amplitude of such a
+low-stretch mode to be compatible with~0 ($<2\%$) in this young population. More
+generally, this raises the question of how well a given environmental
+tracer (here LsSFR) traces the age. An analysis dedicated to this question will
+be presented in Briday et al.\ in prep.
 
@@
 dataset (see Section~\ref{sec:modeling}) --- reduces the significance of the
-results presented in this section, as expected. Even so, non-drifting models remain
-strongly disfavored. For instance, the best fitting sample-based Gaussian
-asymmetric model is still $\Delta\mathrm{AIC}<-10$ less representative of the
-data than our Base drifting model.
+results presented in this section, as expected. Even so, non-drifting
+models remain strongly disfavored. For instance, the best
+fitting sample-based Gaussian asymmetric model is still
+$\Delta\mathrm{AIC}<-10$ less representative of the data than our Base drifting
+model.
 
@@
 Phillips relation \citep{phillips1993} standardization parameter $\alpha$ is not
-redshift dependent (a study \textbf{beyond} the scope of this paper, but see
+redshift dependent (a study beyond the scope of this paper, but see
 e.g. \citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
@@
 stretch and color distributions. Yet, as shown in Section~\ref{sec:results},
-such a sample-based distribution is excluded in comparison to our drifting
-model. Then, contrary to what \citet[][Section~2]{scolnic2016} and
-\citet[][Section~5.4]{scolnic2018a} have suggested --- i.e., that traditional surveys
-span sufficiently limited redshift ranges such that the per-sample approach accounts
-for implicit redshift drifts --- a direct modeling of the redshift drift is more
-appropriate than a sample-based approach. We \textbf{add} here that, as
-measurements of modern surveys try to cover increasingly larger redshift ranges
-in order to reduce calibration systematic uncertainties, this sample-based
-approach \textbf{becomes} less valid, notably for PS1, DES and, soon, LSST.
+such a sample-based distribution is excluded in comparison to our
+drifting model. Then, contrary to what \citet[][Section~2]{scolnic2016}
+and \citet[][Section~5.4]{scolnic2018a} have suggested --- i.e., that
+traditional surveys span sufficiently limited redshift ranges
+such that the per-sample approach accounts for implicit redshift drifts
+--- a direct modeling of the redshift drift is more appropriate than a
+sample-based approach. We add here that, as measurements of modern
+surveys try to cover increasingly larger redshift ranges in order to reduce
+calibration systematic uncertainties, this sample-based approach
+becomes less valid, notably for PS1, DES and, soon, LSST.
 
@@
 lower redshift), the higher the amplitude of the old-specific low-stretch mode.
-This redshift dependency of the underlying stretch distributions is shown running from blue to red in
-Fig.~\ref{fig:bbc_pdf_ps1} for the redshift range covered by PS1. The observed
-$x_1$ histogram follows \textbf{the model we defined using} the sum of
-individual underlying SN-redshift distributions. As expected, the two modeling
-approaches differ mostly in the negative part of the SN stretch distribution.
-The asymmetric Gaussian distribution goes through the middle of the bimodal
-distribution, over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
-under-estimating it at $x_1\sim-1.7$ in comparison to our Base drifting model
-for typical PS1 SN redshifts. This means that the SN bias-corrected standardized
-magnitude estimated at a redshift suffering from by observational selection effects would be biased
-by an mismodeling of the true underlying stretch distribution.
+This redshift dependency of the underlying stretch distributions is
+shown running from blue to red in Fig.~\ref{fig:bbc_pdf_ps1} for the
+redshift range covered by PS1. The observed $x_1$ histogram follows the
+model we defined using the sum of individual underlying SN-redshift
+distributions. As expected, the two modeling approaches differ mostly in the
+negative part of the SN stretch distribution. The asymmetric Gaussian
+distribution goes through the middle of the bimodal distribution,
+over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and under-estimating it at
+$x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN
+redshifts. This means that the SN bias-corrected standardized magnitude
+estimated at a redshift suffering from by observational
+selection effects would be biased by a mismodeling of the true
+underlying stretch distribution.
 
@@
 
-\textbf{Assessing t}he amplitude of this magnitude bias for cosmology is beyond
+Assessing the amplitude of this magnitude bias for cosmology is beyond
 the scope of this paper given the complexity of the BBC analysis. It would
@@
 
-In the era of modern cosmology, where we aim \textbf{to measure} $w_0$ at a
-sub-percent level and $w_a$ \textbf{with} ten-percent precision
+In the era of modern cosmology, where we aim to measure $w_0$ at a
+sub-percent level and $w_a$ with ten-percent precision
 \citep[e.g.,][]{lsstpaper}, we stress that correct modeling of potential SN
@@
 
-We have presented \textbf{an initial} study of the drift of the underlying SNe~Ia
-stretch distribution as a function of redshift. We built effectively volume-limited
-SN~Ia subsamples from the Pantheon dataset \citep[][SDSS, PS1 and
-SNLS]{scolnic2018a}, to which we added HST and SNfactory data from
-\cite{rigault2020} for the high- and low-redshift bins. We only considered the
-SNe that have been discovered in the redshift range of each survey where
-observational selection effects are negligible, so that the observed SNe~Ia stretches are
-a random sampling of the true underlying distribution. This resulted in a fiducial sample 
-of 569 SNe~Ia (422 SNe when more conservative cuts were applied).
+We have presented an initial study of the drift of the underlying
+SNe~Ia stretch distribution as a function of redshift. We built
+effectively volume-limited SN~Ia subsamples from the Pantheon dataset
+\citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to which we added HST and SNfactory
+data from \cite{rigault2020} for the high- and low-redshift bins. We only
+considered the SNe that have been discovered in the redshift range of each
+survey where observational selection effects are negligible, so that
+the observed SNe~Ia stretches are a random sampling of the true underlying
+distribution. This resulted in a fiducial sample of 569 SNe~Ia (422 SNe
+when more conservative cuts were applied).
 
@@
 model which depends on the expected fraction of ``young'' and ``old'' SNe~Ia as
-a function of redshift, \textbf{with} each age population having its own
+a function of redshift, with each age population having its own
 underlying stretch distribution.
 
-In addition to this ``base'' model, we have studied various distributions,
-including redshift independent models; we also studied the prediction from a
-per-sample asymmetric Gaussian stretch distribution used, for instance, by the
-Beams with Bias Correction Malmquist bias correction algorithm
+In addition to this ``base'' model, we have studied various
+distributions, including redshift independent models; we also studied the
+prediction from a per-sample asymmetric Gaussian stretch distribution used, for
+instance, by the Beams with Bias Correction Malmquist bias correction algorithm
 \citep{scolnic2016, kessler2017}.
@@
 \begin{enumerate}
+
     \item The underlying SN~Ia stretch distribution is significantly redshift
-        dependent, as previously suggested by e.g.~\cite{howell2007}, \textbf{in
-        a way that observational selection effects alone cannot explain}. This result is
-        largely independent of details on each age-population model.
-
-    \item Redshift-independent models are \textbf{quantitatively} excluded as
+        dependent, as previously suggested by e.g.~\cite{howell2007}, in
+        a way that observational selection effects alone cannot explain. This
+        result is largely independent of details on each age-population model.
+
+    \item Redshift-independent models are quantitatively excluded as
         suitable descriptions of the data relative to our Base model. This model
@@
         the relative fraction of younger and older SNe~Ia follows the prediction
-        made in \cite{rigault2020}. This second result further supports the
-        existence of both young and old SN~Ia populations, in agreement with
+        made in \cite{rigault2020}. This second result further supports
+        the existence of both young and old SN~Ia populations, in agreement with
         rate studies \cite{mannucci2005, scannapieco2005, sullivan2006,
@@
 
-    \item Models using survey-based asymmetric Gaussian distributions, e.g., as employed
-        in the current implementation of BBC, are excluded \textbf{as a}
-        good description of the data relative to our drifting model. Hence, the
-        sample-based approach does not accurately account for redshift drift, a problem
-        that will be exacerbated as surveys span increasingly larger redshift ranges.
-        \textbf{As a result, even if the necessary} extra degrees of freedom might be acceptable
-        given the large number of SNe~Ia in cosmological studies, extrapolating
-        the SN property distributions from the volume-limited part of a survey
-        to its Malmquist-biased magnitude-limited part would still be inaccurate
+    \item Models using survey-based asymmetric Gaussian distributions,
+        e.g., as employed in the current implementation of BBC, are
+        excluded as a good description of the data relative to our
+        drifting model. Hence, the sample-based approach does not accurately
+        account for redshift drift, a problem that will be exacerbated
+        as surveys span increasingly larger redshift ranges. As a
+    result, even if the necessary extra degrees of freedom might be acceptable
+    given the large number of SNe~Ia in cosmological studies, extrapolating the
+    SN property distributions from the volume-limited part of a survey to its
+    Malmquist-biased magnitude-limited part would still be inaccurate
     because of the redshift evolution.
@@
 
-In this paper, we considered a simple \textbf{two-population} Gaussian mixture
+In this paper, we considered a simple two-population Gaussian mixture
 modeling. Additional data free from significant Malmquist bias would enable
@@
 
-\textbf{The next step in this line of analysis will incorporate our model into
-the SNANA framework \citep{SNANA}, both to more accurately account for observational selection
-functions and to test the impact of our model on the derivation of
-cosmological parameters; this study is underway.}
+The next step in this line of analysis will incorporate our model into
+the SNANA framework \citep{SNANA}, both to more accurately account for
+observational selection functions and to test the impact of our model on the
+derivation of cosmological parameters; this study is underway.
 

### 2021-02-26 Nicolas.tex
2 columns. DONE!!
Started final changes for acceptance

@@
     stretch distribution as a function of redshift, using the evolution of the
-    fraction of young and old SNe~Ia as predicted by \cite{rigault2020}, and
-    assuming constant underlying stretch distribution for each age population
-    consisting of Gaussian mixtures. We test our prediction against published
-    samples cut to have marginal magnitude selection effects, so that any
-    observed change is indeed of astrophysical and not observational in origin.
-    In this first study, there are indications that the underlying SN~Ia stretch
-    distribution is evolving as a function of redshift, and that the young/old
-    drifting model is a better description of the data than any time-constant
-    model, including the sample-based asymmetric distributions often used to
-    correct Malmquist bias \textbf{at a significance of over 5 $\sigma$}. The
-    favored underlying stretch model is the bimodal one derived from
-    \cite{rigault2020}, composed of a high-stretch mode shared by both young and
-    old environments, and a low-stretch mode exclusive to old environments. The
-    precise impact of the redshift evolution of the SN Ia population intrinsic
-    properties on cosmology remains to be studied. Yet, the astrophysical drift
-    of the SN stretch distribution does affect current Malmquist bias
-    corrections and thereby distances derived from SN affected by observational
-selection effects. We highlight that such a bias will increase with surveys
-covering increasingly larger redshift ranges, which is particularly important
-for LSST.}
+    fraction of young and old SNe~Ia as predicted \textbf{in a previous study},
+    and assuming constant underlying stretch distribution for each age
+    population consisting of Gaussian mixtures. We test our prediction against
+    published samples cut to have marginal magnitude selection effects, so that
+    any observed change is indeed of astrophysical and not observational in
+    origin. In this first study, there are indications that the underlying
+    SN~Ia stretch distribution is evolving as a function of redshift, and that
+    the young/old drifting model is a better description of the data than any
+    time-constant model, including the sample-based asymmetric distributions
+    often used to correct Malmquist bias at a significance of over 5 $\sigma$.
+    The favored underlying stretch model is \textbf{a bimodal one}, composed of
+    a high-stretch mode shared by both young and old environments, and a
+    low-stretch mode exclusive to old environments. The precise impact of the
+    redshift evolution of the SN Ia population intrinsic properties on cosmology
+    remains to be studied. Yet, the astrophysical drift of the SN stretch
+    distribution does affect current Malmquist bias corrections and thereby
+    distances derived from SN affected by observational selection effects. We
+highlight that such a bias will increase with surveys covering increasingly
+larger redshift ranges, which is particularly important for LSST.}
 
@@
         the existence of both young and old SN~Ia populations, in agreement with
-        rate studies \cite{mannucci2005, scannapieco2005, sullivan2006,
+        rate studies \citep{mannucci2005, scannapieco2005, sullivan2006,
         aubourg2008}.

### 2021-02-26 Nicolas.tex
1 column acceptance

@@
     stretch distribution as a function of redshift, using the evolution of the
-    fraction of young and old SNe~Ia as predicted \textbf{in a previous study},
-    and assuming constant underlying stretch distribution for each age
+    fraction of young and old SNe~Ia as predicted \textbf{using the SNfactory
+    dataset}, and assuming constant underlying stretch distribution for each age
     population consisting of Gaussian mixtures. We test our prediction against
     published samples cut to have marginal magnitude selection effects, so that
-    any observed change is indeed of astrophysical and not observational in
-    origin. In this first study, there are indications that the underlying
-    SN~Ia stretch distribution is evolving as a function of redshift, and that
-    the young/old drifting model is a better description of the data than any
-    time-constant model, including the sample-based asymmetric distributions
-    often used to correct Malmquist bias at a significance of over 5 $\sigma$.
-    The favored underlying stretch model is \textbf{a bimodal one}, composed of
-    a high-stretch mode shared by both young and old environments, and a
+    any observed change is indeed astrophysical and not observational in origin.
+    In this first study, there are indications that the underlying SN~Ia stretch
+    distribution is evolving as a function of redshift, and that the young/old
+    drifting model is a better description of the data than any time-constant
+    model, including the sample-based asymmetric distributions often used to
+    correct Malmquist bias at a significance of over 5 $\sigma$.  The favored
+    underlying stretch model is \textbf{a bimodal one}, composed of a
+    high-stretch mode shared by both young and old environments, and a
     low-stretch mode exclusive to old environments. The precise impact of the
@@
 
-The relation between SNe~Ia and their host galaxy properties has been
-studied extensively. The first key finding was that the standardized SNe~Ia
-magnitudes significantly depend on the host galaxy stellar mass, with
-SNe~Ia from high-mass host galaxies being brighter on average
-\cite[e.g.][]{kelly2010, sullivan2010, childress2013, betoule2014, rigault2020,
-kim19}. This mass-step correction is currently used in cosmological analyses
-\citep[e.g.][]{betoule2014, scolnic2018a}, including for deriving $H_0$
-\citep{riess2016, riess2019}. Yet, the underlying connection between the SNe and
-their host galaxies remains , particularlyunclear when using
-global properties such as the host stellar . This in turnmass raises
-the question of the accuracy of such corrections since host galaxy
-global properties evolve with redshift. More recently, studies have used the
-local SN host galaxy environment to probe more direct connections
-between the SNe and their host galaxy environments \citep{rigault2013},
-showing that local age tracers such as the Local specific Star Formation Rate
-(LsSFR) or the local color are more strongly correlated with the standardized SN
-magnitude \citep{roman2018, kim18,rigault2020}. Further reinforcing this
-connection is the identification of SN~Ia spectral features that are correlated
-with LsSFR \citep{nordin2018}. These results suggest age as the driving
-parameter underlying the mass-step. If true, this would have a significant
-impact for cosmology, since the environmental correction to apply to SN
-standardization could strongly vary with redshift. \citep{rigault2013,
-childress2014, scolnic2018a}. Yet, the importance of local SN environmental
-studies remains highly debated \cite[e.g.][]{jones2015, jones2019} and
-especially the impact of such an astrophysical bias has on the derivation of
-$H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.
+The relation between SNe~Ia and their host galaxy properties has been studied
+extensively. The first key finding was that the standardized SNe~Ia magnitudes
+significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass
+host galaxies being brighter on average \cite[e.g.][]{kelly2010, sullivan2010,
+childress2013, betoule2014, kim19}. This mass-step correction is currently used
+in cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including
+for deriving $H_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
+between the SNe and their host galaxies remains unclear, particularly when using
+global properties such as the host stellar mass. This in turn raises the
+question of the accuracy of such corrections since host galaxy global properties
+evolve with redshift. More recently, studies have used the local SN host galaxy
+environment to probe more direct connections between the SNe and their host
+galaxy environments \citep{rigault2013}, showing that local age tracers such as
+the Local specific Star Formation Rate (LsSFR) or the local color are more
+strongly correlated with the standardized SN magnitude \citep{roman2018, kim18,
+rigault2020}. Further reinforcing this connection is the identification of SN~Ia
+spectral features that are correlated with LsSFR \citep{nordin2018}. These
+results suggest age as the driving parameter underlying the mass-step. If true,
+this would have a significant impact for cosmology, since the environmental
+correction to apply to SN standardization could strongly vary with redshift
+\citep{rigault2013, childress2014, scolnic2018a}. Yet, the importance of local
+SN environmental studies remains highly debated \cite[e.g.][]{jones2015,
+jones2019} and especially the impact of such an astrophysical bias has on the
+derivation of $H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.
 
@@
 Furthermore, since the first SNe~Ia host analyses, the SN stretch has been known
-to be strongly correlated with the SN host galaxy properties
-\citep{hamuy1996, hamuy2000}. This correlation that has been extensively
-confirmed since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
-gupta2011, dandrea2011, childress2013, rigault2013, pan2014, kim19}. Following
-the ``A+B'' model and the connection between SN stretch and host galaxy
-properties, \cite{howell2007} first discussed the potential redshift drift of
-the SN stretch distribution. In this paper we revisit this question
+to be strongly correlated with the SN host galaxy properties \citep{hamuy1996,
+hamuy2000}. This correlation that has been extensively confirmed since
+\citep[e.g.][]{neill2009, lampeitl2010, gupta2011, dandrea2011, pan2014}.
+Following the ``A+B'' model and the connection between SN stretch and host
+galaxy properties, \cite{howell2007} first discussed the potential redshift
+drift of the SN stretch distribution. In this paper we revisit this question
 using the most recent SNe~Ia datasets.
 
-In this paper, we take a step aside from the cosmological analyses to
-probe the validity of our modeling of the SN population, which we claim to be
-constituted of two age-populations \citep{rigault2013,rigault2015,rigault2020}:
-one old and one younger, the former having on average lower lightcurve stretches
-and being brighter after standardization. We use the correlation between the SN
-age, as probed by the LsSFR, and the SN stretch to model the expected evolution
-of the underlying SN stretch distribution as a function of redshift. This
-modeling relies on three assumptions: (1) there are two distinct populations of
-SNe~Ia; (2) the relative fraction of each of these populations as a function of
-redshift follows the model presented in \cite{rigault2020} and (3) the
-underlying distribution of stretch for each age sample is constant. This paper
-aims at testing this specific model with datasets from the literature.
+In this paper, we take a step aside from the cosmological analyses to probe the
+validity of our modeling of the SN population, which we claim to be constituted
+of two age-populations \citep{rigault2013, rigault2015, rigault2020}: one old
+and one younger, the former having on average lower lightcurve stretches and
+being brighter after standardization. We use the correlation between the SN age,
+as probed by the LsSFR, and the SN stretch to model the expected evolution of
+the underlying SN stretch distribution as a function of redshift. This modeling
+relies on three assumptions: (1) there are two distinct populations of SNe~Ia;
+(2) the relative fraction of each of these populations as a function of redshift
+follows the model presented in \cite{rigault2020} and (3) the underlying
+distribution of stretch for each age sample is constant. This paper aims at
+testing this specific model with datasets from the literature.
 
@@
 stretch distribution function, and we would then study how it evolves with
-    redshift. No such sample exists, so first we must construct subsamples that
-are as near to volume-limited as possible from existing high-redshift SN~Ia
-samples.
+redshift. No such sample exists, so first we must construct subsamples that are
+as near to volume-limited as possible from existing high-redshift SN~Ia samples.
 
@@
         corresponding standardized absolute magnitude using the $\alpha$ and
-        $\beta$ coefficients from \cite{scolnic2018a}. The grey diagonal lines
+        $\beta$ coefficients from \cite{scolnic2018a}. The gray diagonal lines
         represent the $(x_1, c)$ evolution for $m = m_{\lim}$, for $z$ between
@@
     observational selection bias (see Section~\ref{sec:sample}). The
-darker (resp. lighter) color responds to the conservative (resp. fiducial)
+darker (resp.\ lighter) color responds to the conservative (resp.\ fiducial)
 selection cut.}
@@
             $0.15$ and $0.25$, respectively. Kolmogorov-Smirnov test $p$-values
-            are indicated on the top (resp. bottom) of each panel showing no
+            are indicated on the top (resp.\ bottom) of each panel showing no
             sign that the SDSS and PS1 (resp. PS1 and SNLS) $x_1$ and $c$
@@
 Hence, if we denote $\delta(z)$ (resp. $\psi(z) = 1-\delta(z)$) the fraction of
-young (resp. old) SNe~Ia in the Universe as a function of redshift, then the
+young (resp.\ old) SNe~Ia in the Universe as a function of redshift, then the
 ratio $\delta/\psi$ is expected to follow the evolution of the specific star
@@
 by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
-of the underlying stretch distribution of SNe~Ia as a function of redshift $X_1(z)$ is
-given by:
+of the underlying stretch distribution of SNe~Ia as a function of redshift
+$X_1(z)$ is given by:
 \begin{align}\label{eq:stretchz}
@@
 Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
-(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
+(resp.\ conservative) sample are given Table~\ref{tab:modelresults}, and the
 predicted redshift evolution of mean stretch (expected $x_1$ given the
@@
         the SNfactory dataset only, respectively; all are compatible. The
-        light-blue band illustrates the amplitude of the error (incl.
+        light-blue band illustrates the amplitude of the error (incl.\
         covariance) of the best fit model when considering the fiducial
@@
 This is our so-called ``Base'' model. However, to test different modeling
-choices, we have implemented a suite of alternative parametrizations that we
+choices, we have implemented a suite of alternative parameterizations that we
 also adjust to the data following the procedure described in
@@
     \caption{Distribution of the PS1 SN~Ia \textsc{\texttt{SALT2.4}} stretch
-        ($x_1$) after the fiducial redshift limit cut (grey histogram). This
+        ($x_1$) after the fiducial redshift limit cut (gray histogram). This
         distribution is supposed to be a random draw from the underlying stretch

### 2021-02-26 Nicolas.tex
2 columns and go
cleaned the usepackages, 1column useless now

@@
-\documentclass[referee]{aa}
+\documentclass[]{aa}
 
 \usepackage{txfonts}
-\usepackage{ulem}
-\usepackage{amsmath, mathtools}
-\usepackage[colorlinks,breaklinks]{hyperref}
-\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}
+\usepackage[colorlinks, breaklinks]{hyperref}
+\hypersetup{linkcolor=blue, citecolor=blue, filecolor=black, urlcolor=blue}
 
@@
 
-\def\l{\lambda}\def\L{\Lambda}
-
-\newcommand{\nn}[1]{\textcolor[rgb]{1.0, 0.6, 0.0}{#1}}
-\newcommand{\mr}[1]{\textcolor[rgb]{0.5, 0.20, 0.3}{#1}}
-\newcommand{\yc}[1]{\textcolor[RGB]{217, 22, 102}{#1}}
 \newcommand{\prob}[2]{\mathcal{P}\left( #1 \mid #2\right)}
@@
 from cosmic microwave background (CMB) data measured by Planck assuming the
-standard $\L$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
+standard $\Lambda$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
 luminosity is anchored at intermediate redshift by the baryon acoustic

### 2021-03-02 Nicolas.tex
That's it now

@@
     stretch distribution as a function of redshift, using the evolution of the
-    fraction of young and old SNe~Ia as predicted \textbf{using the SNfactory
-    dataset}, and assuming constant underlying stretch distribution for each age
+    fraction of young and old SNe~Ia as predicted using the SNfactory dataset,
+    and assuming constant underlying stretch distribution for each age
     population consisting of Gaussian mixtures. We test our prediction against
@@
     correct Malmquist bias at a significance of over 5 $\sigma$. The favored
-    underlying stretch model is \textbf{a bimodal one}, composed of a
-    high-stretch mode shared by both young and old environments, and a
-    low-stretch mode exclusive to old environments. The precise impact of the
-    redshift evolution of the SN Ia population intrinsic properties on cosmology
-    remains to be studied. Yet, the astrophysical drift of the SN stretch
-    distribution does affect current Malmquist bias corrections and thereby
-    distances derived from SN affected by observational selection effects. We
-highlight that such a bias will increase with surveys covering increasingly
-larger redshift ranges, which is particularly important for LSST.}
+    underlying stretch model is a bimodal one, composed of a high-stretch mode
+    shared by both young and old environments, and a low-stretch mode exclusive
+    to old environments. The precise impact of the redshift evolution of the SN
+    Ia population intrinsic properties on cosmology remains to be studied. Yet,
+    the astrophysical drift of the SN stretch distribution does affect current
+    Malmquist bias corrections and thereby distances derived from SN affected by
+    observational selection effects. We highlight that such a bias will increase
+with surveys covering increasingly larger redshift ranges, which is particularly
+important for LSST.}
 
@@
 host galaxies being brighter on average \cite[e.g.][]{kelly2010, sullivan2010,
-childress2013, betoule2014, kim19}. This mass-step correction is currently used
-in cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including
-for deriving $H_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
-between the SNe and their host galaxies remains unclear, particularly when using
-global properties such as the host stellar mass. This in turn raises the
-question of the accuracy of such corrections since host galaxy global properties
-evolve with redshift. More recently, studies have used the local SN host galaxy
-environment to probe more direct connections between the SNe and their host
-galaxy environments \citep{rigault2013}, showing that local age tracers such as
-the Local specific Star Formation Rate (LsSFR) or the local color are more
-strongly correlated with the standardized SN magnitude \citep{roman2018, kim18,
-rigault2020}. Further reinforcing this connection is the identification of SN~Ia
-spectral features that are correlated with LsSFR \citep{nordin2018}. These
-results suggest age as the driving parameter underlying the mass-step. If true,
-this would have a significant impact for cosmology, since the environmental
-correction to apply to SN standardization could strongly vary with redshift
-\citep{rigault2013, childress2014, scolnic2018a}. Yet, the importance of local
-SN environmental studies remains highly debated \cite[e.g.][]{jones2015,
-jones2019} and especially the impact of such an astrophysical bias has on the
-derivation of $H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.
+childress2013, betoule2014, kim19, smith2020}. This mass-step correction is
+currently used in cosmological analyses \citep[e.g.][]{betoule2014,
+scolnic2018a}, including for deriving $H_0$ \citep{riess2016, riess2019}. Yet,
+the underlying connection between the SNe and their host galaxies remains
+unclear, particularly when using global properties such as the host stellar
+mass. This in turn raises the question of the accuracy of such corrections since
+host galaxy global properties evolve with redshift. More recently, studies have
+used the local SN host galaxy environment to probe more direct connections
+between the SNe and their host galaxy environments \citep{rigault2013}, showing
+that local age tracers such as the Local specific Star Formation Rate (LsSFR) or
+the local color are more strongly correlated with the standardized SN magnitude
+\citep{roman2018, kim18, rigault2020}. Further reinforcing this connection is
+the identification of SN~Ia spectral features that are correlated with LsSFR
+\citep{nordin2018}. These results suggest age as the driving parameter
+underlying the mass-step. If true, this would have a significant impact for
+cosmology, since the environmental correction to apply to SN standardization
+could strongly vary with redshift \citep{rigault2013, childress2014,
+scolnic2018a}. Yet, the importance of local SN environmental studies remains
+highly debated \cite[e.g.][]{jones2015, jones2019} and especially the impact of
+such an astrophysical bias has on the derivation of $H_0$ \citep{jones2015,
+riess2016, riess2018, rose2019}.
 
 The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
-\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
+\cite{mannucci2005, scannapieco2005, sullivan2006, smith2012} have shown that
 the relative SNe~Ia rate in galaxies could be explained if two populations
@@
 been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
-and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
-\citealt{aldering2020}). For this dataset, spectroscopic screening was
-    done for candidates with $r \lesssim 19.5$; redshift cuts were then applied
-    when selecting which SN Ia to follow, resulting in a redshift range of $0.02
-< z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} for extracting
-local host properties. These 114 SNfactory SNe~Ia are thus in the
+and~4.2.2 ; see \citealt{aldering2020}). For this dataset, spectroscopic
+screening was done for candidates with $r \lesssim 19.5$; redshift cuts were
+then applied when selecting which SN Ia to follow, resulting in a redshift range
+of $0.02 < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} for
+extracting local host properties. These 114 SNfactory SNe~Ia are thus in the
 volume-limited part of the survey (Aldering et al., in prep.), and are therefore
@@
 sample is particularly useful for studying SN property drift, as it enables us
-to have a large complete SN~Ia sample at $z<0.1$.
-
-Finally, we include the HST sample from Pantheon \citep{strolger04}.
-These high-redshift SNe are of great interest as they provide the
-    greatest leverage for testing evolution. While at these redshifts the
-    supernovae typing is challenging, the target classification was robust
-    enough to include them within the cosmological analysis \citep{scolnic2018a}
-    and we do not impose further cuts. Section~\ref{sec:results} highlights
-    that, while compatible with it, our results are not dependent on the
-    inclusion of this dataset.
+to have a large complete SN~Ia sample at $z<0.1$. Finally, we include the HST
+sample from Pantheon \citep{strolger04}. These high-redshift SNe are of great
+interest as they provide the greatest leverage for testing evolution. While at
+these redshifts the supernovae typing is challenging, the target classification
+was robust enough to include them within the cosmological analysis
+\citep{scolnic2018a} and we do not impose further cuts.
+Section~\ref{sec:results} highlights that, while compatible with it, our results
+are not dependent on the inclusion of this dataset.
 
 We present the stretch distribution and redshift histogram of these five surveys
-up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. We observe here
-that the fraction of low-stretch SNe (typically $x_1 < -1$) appears to decrease as
-a function of redshift; this is confirmed in Fig.~\ref{fig:modelall}, in which
-the evolution of the mean stretch is shown, with the data split in
-redshift bins of regular sample size. We see that SNe~Ia at higher redshift have
-on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
-redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
-stretch distribution is evolving with redshift.
+up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. We observe here that
+the fraction of low-stretch SNe (typically $x_1 < -1$) appears to decrease as a
+function of redshift; this is confirmed in Fig.~\ref{fig:modelall}, in which the
+evolution of the mean stretch is shown, with the data split in redshift bins of
+regular sample size. We see that SNe~Ia at higher redshift have on average
+larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift
+($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch
+distribution is evolving with redshift.
 
@@
 rates and delay time distributions \citep[e.g.,][]{mannucci2005,
-scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
+scannapieco2005, sullivan2006, smith2012, childress2014, maozmannucci2014}.
 In short, it was assumed that the number of ``young'' SNe~Ia follows the star
@@
 
-\bibitem[Saunders et al.(2020)]{saunders2020} Saunders, C., Aldering, G.,
-Antilogus, P., et al.\ 2020, VizieR Online Data Catalog, J/ApJ/869/167
+\bibitem[Saunders et al.(2018)]{saunders2018} Saunders, C., Aldering, G.,
+Antilogus, P., et al.\ 2018, \apj, 869, 167. doi:10.3847/1538-4357/aaec7e
 
@@
 et al.\ 2016, \apj, 827, 1
+
+\bibitem[Smith et al.(2012)]{smith2012} Smith, M., Nichol, R.~C., Dilday, B., et
+al.\ 2012, \apj, 755, 61
+
+\bibitem[Smith et al.(2020)]{smith2020} Smith, M., Sullivan, M., Wiseman, P., et
+al.\ 2020, \mnras, 494, 4426
 

### 2021-03-05 Nicolas.tex
Language editor files

@@
 
+\newcommand{\LEt}[1]{\textcolor{red}{\textbf{ #1}}}
 \newcommand{\orcid}[1]{
@@
 \begin{document}
-\title{Redshift Evolution of the Underlying Type Ia Supernova Stretch
-Distribution}
-
-\titlerunning{Redshift Evolution of the Underlying Type Ia Supernova Stretch
-Distribution}
+
+\title{Redshift evolution of the underlying type Ia supernova stretch
+distribution}
+
+\titlerunning{Redshift evolution of the underlying type Ia supernova stretch
+distribution}
 \authorrunning{N.~Nicolas et al.}
@@
     and as survey statistics increase, the question of astrophysical systematic
-    uncertainties rises, notably that of SN~Ia population evolution. In this
-    paper, we study the dependence with redshift of the SN~Ia \texttt{SALT2.4}
-    lightcurve stretch, a purely intrinsic SN property, to probe its potential
-    redshift drift. The SN stretch has been shown to strongly correlate with the
-    SN environment, notably with stellar age tracers. We model the underlying
+    uncertainties arises, notably that of the evolution of SN~Ia populations. We study the dependence on redshift of the SN~Ia \texttt{SALT2.4}
+    light-curve stretch, which is a purely intrinsic SN property, to probe its potential
+    redshift drift. The SN stretch has been shown to be strongly correlated with the
+    SN environment, notably with stellar age tracers. We modeled the underlying
     stretch distribution as a function of redshift, using the evolution of the
     fraction of young and old SNe~Ia as predicted using the SNfactory dataset,
-    and assuming constant underlying stretch distribution for each age
-    population consisting of Gaussian mixtures. We test our prediction against
-    published samples cut to have marginal magnitude selection effects, so that
+    and assuming a constant underlying stretch distribution for each age
+    population consisting of Gaussian mixtures. We tested our prediction against
+    published samples that were cut to have marginal magnitude selection effects, so that
     any observed change is indeed astrophysical and not observational in origin.
     In this first study, there are indications that the underlying SN~Ia stretch
-    distribution is evolving as a function of redshift, and that the young/old
+    distribution evolves as a function of redshift, and that the young/old\LEt{the slash commonly denotes "ratio", as in "S/N" for "signal-to-noise ratio".
+The way you use it here is ambiguous: do you mean "and" or "or", or maybe a range?\ Please rephrase for clarity}
     drifting model is a better description of the data than any time-constant
-    model, including the sample-based asymmetric distributions often used to
-    correct Malmquist bias at a significance of over 5 $\sigma$. The favored
+    model, including the sample-based asymmetric distributions that are often used to
+    correct Malmquist bias at a significance higher than 5 $\sigma$. The favored
     underlying stretch model is a bimodal one, composed of a high-stretch mode
-    shared by both young and old environments, and a low-stretch mode exclusive
-    to old environments. The precise impact of the redshift evolution of the SN
-    Ia population intrinsic properties on cosmology remains to be studied. Yet,
-    the astrophysical drift of the SN stretch distribution does affect current
-    Malmquist bias corrections and thereby distances derived from SN affected by
-    observational selection effects. We highlight that such a bias will increase
+    shared by both young and old environments, and a low-stretch mode that is exclusive
+    to old environments. The precise effect of the redshift evolution of the intrinsic properties of a SN
+    Ia population on cosmology remains to be studied. The astrophysical drift of the SN stretch distribution does affect current
+    Malmquist bias corrections, however, and thereby the distances that are derived based on SN that are affected by
+    observational selection effects. We highlight that this bias will increase
 with surveys covering increasingly larger redshift ranges, which is particularly
-important for LSST.}
+important for the \LEt{please spell this out here and in the main text, you use the abbreviation only one time each}LSST.}
 
@@
 Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
-enabled the discovery of the acceleration of the Universe's expansion
-\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
-understand the properties of dark energy (DE) as it is the only tool able to
-precisely map the recent expansion rate ($z<0.5$), when DE is driving it
-\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
-Hubble Constant ($H_0$), provided one can calibrate their absolute magnitude
-\citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ derived when
-the SNe~Ia are anchored to Cepheids \citep[the SH0ES
+enabled the discovery that the Universe expands
+\citep{riess1998, perlmutter1999}. They remain a key cosmological probe today for
+understanding the properties of dark energy (DE) because it\LEt{grammatically, "it" here refers to "DE".
+Is this correct, or should this be "they" because you still refer to the SNe Ia, or do you mean "the probe"?
+If you mean the probe, please rephrase to "because this probe" for clarity} is the only tool that can
+precisely map the recent expansion rate ($z<0.5$) when DE is driving it
+\citep[e.g.,][]{scolnicastro2020}. Type Ia supernovae are also the key to directly measuring the
+Hubble constant ($H_0$) if their absolute magnitude can be calibrated
+\citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ tha tis derived when
+the SNe~Ia are anchored to Cepheids \citep[the\LEt{please provide the spelled-out versions of all instruments and surveys at first occurrence.
+If this interrupts the flow of the sentence too much, it may be done as a footnote} SH0ES
 project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
@@
 oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
-branch technique in place of the Cepheids seem to favor an intermediate value of
-$H_0$ \citep{freedman2019, freedman2020}, time delay measurements from strong
-lensing seem to also favor high $H_0$ values \citep{wong2019}.
-
-The $H_0$ tension has received a lot of attention, as it could be a sign of new
-fundamental physics. Yet, no simple solution is able to accommodate this $H_0$
-tension when accounting for all other probes \citep{knox2019}. Alternatively,
+branch technique in place of the Cepheids seems to favor an intermediate value of
+$H_0$ \citep{freedman2019, freedman2020}, time-delay measurements from strong
+lensing also appear to favor high $H_0$ values \citep{wong2019}.
+
+The $H_0$ conundrum has received much attention because it could be a sign of new
+fundamental physics. No simple solution is so far able to accommodate this $H_0$
+conundrum when all other probes are accounted for, however \citep{knox2019}. Alternatively,
 systematic effects affecting one or several of the aforementioned analyses
-might explain at least some of this tension. \cite{rigault2015}
+might explain at least some of this discrepancy. \cite{rigault2015}
 suggested that SNe Ia from the Cepheid calibrator sample differ by construction
 from those in the Hubble flow sample, as the former strongly favors
-young stellar populations, while the latter does not. This selection
-effect would impact the derivation of $H_0$ if SNe~Ia from young and older
-environments differ in average standardized magnitudes.
+young stellar populations while the latter does not. This selection
+effect would affect the derivation of $H_0$ if SNe~Ia from young and older
+environments differed in average standardized magnitudes.
 
@@
 significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass
-host galaxies being brighter on average \cite[e.g.][]{kelly2010, sullivan2010,
+host galaxies being brighter on average \cite[e.g.,][]{kelly2010, sullivan2010,
 childress2013, betoule2014, kim19, smith2020}. This mass-step correction is
-currently used in cosmological analyses \citep[e.g.][]{betoule2014,
-scolnic2018a}, including for deriving $H_0$ \citep{riess2016, riess2019}. Yet,
-the underlying connection between the SNe and their host galaxies remains
-unclear, particularly when using global properties such as the host stellar
-mass. This in turn raises the question of the accuracy of such corrections since
-host galaxy global properties evolve with redshift. More recently, studies have
+currently used in cosmological analyses \citep[e.g.,][]{betoule2014,
+scolnic2018a}, including to derive $H_0$ \citep{riess2016, riess2019}. The underlying connection between the SNe and their host galaxies remains
+unclear, however, particularly when global properties such as the host stellar
+mass are used. This in turn raises the question of the accuracy of such corrections because
+the global properties of the host galaxies evolve with redshift. More recently, studies have
 used the local SN host galaxy environment to probe more direct connections
 between the SNe and their host galaxy environments \citep{rigault2013}, showing
-that local age tracers such as the Local specific Star Formation Rate (LsSFR) or
+that local age tracers such as the local specific star formation rate (LsSFR) or
 the local color are more strongly correlated with the standardized SN magnitude
-\citep{roman2018, kim18, rigault2020}. Further reinforcing this connection is
-the identification of SN~Ia spectral features that are correlated with LsSFR
-\citep{nordin2018}. These results suggest age as the driving parameter
-underlying the mass-step. If true, this would have a significant impact for
-cosmology, since the environmental correction to apply to SN standardization
+\citep{roman2018, kim18, rigault2020}. The identification of SN~Ia spectral features that are correlated with LsSFR
+\citep{nordin2018} further support this connection. These results suggest age as the driving parameter
+underlying the mass step. If this is true, itwould have a significant effect for
+cosmology because the environmental correction that would need to be applied to SN standardization
 could strongly vary with redshift \citep{rigault2013, childress2014,
-scolnic2018a}. Yet, the importance of local SN environmental studies remains
-highly debated \cite[e.g.][]{jones2015, jones2019} and especially the impact of
+scolnic2018a}. The importance of local SN environmental studies remains
+highly debated, however \cite[e.g.,][]{jones2015, jones2019}, and especially the effect of
 such an astrophysical bias has on the derivation of $H_0$ \citep{jones2015,
@@
 
-The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
+The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.\LEt{to properly include these references into the main text, please separate them by commas and set off the last reference additionally with "and".
+This is a LaTeX command error that I cannot fix for you with the program I work with (citet and citep)}
 \cite{mannucci2005, scannapieco2005, sullivan2006, smith2012} have shown that
 the relative SNe~Ia rate in galaxies could be explained if two populations
-existed, one young, following the host star formation activity, and one old
-following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
-model). \cite{rigault2020} used the LsSFR to classify which are the younger
-(those with a high LsSFR) and which are the older (those with low LsSFR).
+existed, a young population that follows the host star formation activity, and an old population that
+followed the host stellar mass (the so-called prompt-and-delayed or A+B
+model). \cite{rigault2020} used the LsSFR to determine the younger
+(those with a high LsSFR) and the older (those with a low LsSFR) populations.
 Furthermore, since the first SNe~Ia host analyses, the SN stretch has been known
 to be strongly correlated with the SN host galaxy properties \citep{hamuy1996,
-hamuy2000}. This correlation that has been extensively confirmed since
-\citep[e.g.][]{neill2009, lampeitl2010, gupta2011, dandrea2011, pan2014}.
-Following the ``A+B'' model and the connection between SN stretch and host
+hamuy2000}. This correlation has been extensively confirmed since then
+\citep[e.g.,][]{neill2009, lampeitl2010, gupta2011, dandrea2011, pan2014}.
+Following the A+B model and the connection between SN stretch and host
 galaxy properties, \cite{howell2007} first discussed the potential redshift
@@
 
-In this paper, we take a step aside from the cosmological analyses to probe the
-validity of our modeling of the SN population, which we claim to be constituted
-of two age-populations \citep{rigault2013, rigault2015, rigault2020}: one old
-and one younger, the former having on average lower lightcurve stretches and
-being brighter after standardization. We use the correlation between the SN age,
-as probed by the LsSFR, and the SN stretch to model the expected evolution of
+We here step aside from the cosmological analyses to probe the
+validity of our modeling of the SN population, which we claim is constituted
+of two age populations \citep{rigault2013, rigault2015, rigault2020}: an older
+and a younger population, the former having lower light-curve stretches on average and
+being brighter after standardization. We use the correlation between the SN age
+as probed by the LsSFR and the SN stretch to model the expected evolution of
 the underlying SN stretch distribution as a function of redshift. This modeling
-relies on three assumptions: (1) there are two distinct populations of SNe~Ia;
+relies on three assumptions: (1) There are two distinct populations of SNe~Ia,
 (2) the relative fraction of each of these populations as a function of redshift
-follows the model presented in \cite{rigault2020} and (3) the underlying
-distribution of stretch for each age sample is constant. This paper aims at
-testing this specific model with datasets from the literature.
-
-We present in Section~\ref{sec:sample} the sample we are using for this
-analysis, derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
-importance of obtaining a ``complete'' sample, i.e.\ representative of the true
-underlying SNe Ia distribution, and how we build one from the Pantheon sample.
-We then present in Section~\ref{sec:modeling} our modeling of the distribution
-of stretch and our results are presented in Section~\ref{sec:results}. In this
+follows the model presented in \cite{rigault2020}, and (3) the underlying
+distribution of stretch for each age sample is constant. This paper tests this specific model with datasets from the literature.
+
+We present the sample we used for this analysis in Section~\ref{sec:sample}. It was derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
+importance of obtaining a ``complete'' sample, that is, a sample that is\ representative of the true
+underlying SNe Ia distribution, and how we built one from the Pantheon sample.
+We then present our modeling of the distribution
+of stretch in Section~\ref{sec:modeling}  and our results in Section~\ref{sec:results}. In this
 section, we test whether the SN stretch distribution evolves as a function of
-redshift and if the aforementioned age model is in good agreement with this
+redshift and determine whether the aforementioned age model agrees well with this
 evolution. We briefly discuss these results in the context of SN cosmology in
-Section~\ref{sec:discussion} and we conclude in Section~\ref{sec:ccl}.
-
-\section{Complete Sample Construction}\label{sec:sample}
+Section~\ref{sec:discussion}, and we conclude in Section~\ref{sec:ccl}.
+
+\section{Complete sample construction}\label{sec:sample}
 
@@
 stretch distribution function, and we would then study how it evolves with
-redshift. No such sample exists, so first we must construct subsamples that are
-as near to volume-limited as possible from existing high-redshift SN~Ia samples.
+redshift. No such sample exists, therefore we must first construct subsamples from existing high-redshift SN~Ia samples that are
+as near to volume-limited as possible.
 
@@
 
-We base our analysis on the most recent comprehensive SNe~Ia compilation, the
-Pantheon catalog from \cite{scolnic2018a}. A naive approach to test the SN
+We based our analysis on the most recent comprehensive SNe~Ia compilation, the
+Pantheon catalog from \cite{scolnic2018a}. A naive approach to testing the SN
 stretch redshift drift would be to simply compare the observed SN stretch
-distributions in a few bins of redshift. In practice, however, differential
+distributions in a few redshift bins. In practice, however, differential
 observational selection effects will affect the observed SN
-stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
-with the lightcurve stretch (and color), the first SNe~Ia that a
-magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
+stretch distributions. Because the observed SN~Ia magnitude correlates
+with the light-curve stretch (and color), the first SNe~Ia that a
+magnitude-limited survey will miss are those at the lowest stretch (which are the reddest).
 Consequently, if magnitude-related observational selection effects are
-not accounted for, one might confuse true population drift with survey
-properties, and conversely.
+not accounted for, true population drift might be confused with survey
+properties, and vice versa.
 
@@
 Aiding in the construction of nearly volume-limited subsamples is the
-    fact that the SN~Ia population trails off towards fainter SNe~Ia. A
+    fact that the SN~Ia population trails off toward fainter SNe~Ia. A
     complication is that complete spectroscopic follow-up has not always been
@@
     discarded from our analysis. High-redshift SN cosmology samples,
-    such as Pantheon, are predominately from magnitude-limited surveys from
+    such as the Pantheon sample, are predominately assembled from magnitude-limited surveys from
 which volume-limited SN~Ia subsamples can be constructed.
 
-We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia
+We present in Fig.~\ref{fig:maglim} the light-curve stretch and color of SNe~Ia
 from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
-Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
+Sky Survey \citep[SDSS][]{frieman2008}, and the SuperNovae Legacy Survey
 \citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
@@
 derive the absolute standardized magnitude at maximum of light $M = M_0 - \alpha
-x_1 + \beta c$ along the aforementioned ellipse given the standardization
-coefficient $\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the
+x_1 + \beta c$ within the aforementioned ellipse given the standardization
+coefficient $\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: The
 faintest SN~Ia is that with $(x_1=-1.65, c=0.25)$ and an absolute standardized
-magnitude at peak in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one
-ought to detect this object typically 5 days before and a week after peak to
-build a suitable lightcurve, the effective limiting standardized absolute
+magnitude at peak in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. This object typically ought to be detected 5 days before and a week after peak to
+build a suitable light curve, the effective limiting standardized absolute
 magnitude is approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude
-limit $m_{\lim}$ of a magnitude limited survey, one can derive the maximum
+limit $m_{\lim}$ of a magnitude limited survey, we can derive the maximum
 redshift $z_{\lim}$ above which the faintest SNe~Ia will be missed using the
-relation between apparent magnitude, redshift and absolute magnitude
+relation between apparent magnitude, redshift, and absolute magnitude
 $\mu(z_{\lim}) = m_{\lim} - M_{\lim}$.
@@
     \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
-        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples of
+        light-curve parameters of SNe~Ia from the SDSS, PS1, and SNLS samples of
         the Pantheon catalog. The individual SNe are shown as blue dots. The
@@
         corresponding standardized absolute magnitude using the $\alpha$ and
-        $\beta$ coefficients from \cite{scolnic2018a}. The gray diagonal lines
-        represent the $(x_1, c)$ evolution for $m = m_{\lim}$, for $z$ between
-        $0.50$ and $z=1.70$ using SNLS's $m_{\lim}$ of $24.8$~mag.}
+        $\beta$ coefficients from \cite{scolnic2018a}. The diagonal gray lines
+        represent the $(x_1, c)$ evolution for $m = m_{\lim}$ for $z$ between
+        $0.50$ and $z=1.70$ using the SNLS $m_{\lim}$ of $24.8$~mag.}
     \label{fig:maglim}
@@
 
-We will thus consider a set of cuts that will define a first fiducial
+We therefore considered a set of cuts that defines a first fiducial
     sample, taking the limits as initially suggested by the previous
     completeness analysis. However, as this solution might be an overly
-    simplified way to create a complete sample, e.g., because it ignores
+    simplified way to create a complete sample, for example, because it ignores
     spectroscopic follow-up in efficiency for redshifts below $z_{lim}$, we
-    also consider another set of cuts to define a so-called ``conservative''
-    sample. The latter is smaller and therefore less statistically constraining,
+    also considered another set of cuts to define a so-called conservative
+    sample. This is smaller and therefore less statistically constraining,
     but also even less prone to observational selection effects. If the redshift
     drift is still significant in the conservative sample, it would be even more
-    meaningful in a carefully-tailored selection-free sample. These samples are
+    meaningful in a carefully tailored selection-free sample. These samples are
     adequate for the goal of this study, which is to develop a first
     implementation of a model for drift in SN Ia properties. If fruitful, the
-    sample selection can later be refined, if needed, with a more detailed model
-of the observational selection, e.g., using the SNANA package \citep{SNANA}.
-
-SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
+    sample selection can later be refined as required with a more detailed model
+of the observational selection, for instance, using the SNANA package \citep{SNANA}.
+
+The SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
 redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
@@
 24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
-final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
-\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
-\citet[][see their Section~5]{perrett2010} suggests however a lower limit of
-$z_{\lim}=0.55$. We will therefore use $z=0.60$ and $z=0.55$ as redshift
-limits for the fiducial and conservative samples, respectively, for SNLS.
+final release website.}}. This converts into a $z_{\lim}=0.60$, in agreement with
+\cite{neill2006}, \cite{perrett2010}, and \cite{bazin2011}. Fig.~14 of
+\citet[][see their Section~5]{perrett2010}, however, suggests a lower limit of
+$z_{\lim}=0.55$. We therefore used $z=0.60$ and $z=0.55$ as redshift
+limits for the fiducial and conservative samples, respectively, for the SNLS.
 
@@
 $5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields $z_{\lim}=0.31$, in
-agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. If we were to be
+agreement with Fig.~6 of \cite{scolnic2018a}, for example. If we were to be
     conservative, this figure would also suggest of a more stringent
-$z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
+$z_{\lim}=0.27$ cut; we therefore used $0.31$ and $0.27$ for our fiducial and
 conservative samples, respectively, for PS1.
 
-In a similar redshift range, SDSS has a limiting magnitude of 22.5
+In a similar redshift range, the SDSS has a limiting magnitude of 22.5
 \citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
@@
 SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
-to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
+to a redshift cut at $0.15$. For the remaining SDSS survey, additional
 spectroscopic resources were available, and \cite{kessler2009} and
-\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$.
-Based on this, we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
-fiducial and conservative samples, respectively, for SDSS.
+\cite{dilday2008} showed a reasonable completeness up to $z_{\lim}=0.2$.
+Based on this, we used $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
+fiducial and conservative samples, respectively, for the SDSS.
 
@@
     before the peak of these histograms. In Section \ref{ssec:verify} we
-validate that these redshift limits are effective for constructing nearly
+confirm that these redshift limits are effective for constructing nearly
 volume-limited subsamples from samples that were initially more closely
-magnitude-limited in their search or spectroscopic follow-up.
+magnitude limited in their search or spectroscopic follow-up.
 
@@
     \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia from the
-        SDSS, PS1 and SNLS dataset respectively (data from Pantheon,
+        SDSS, PS1, and SNLS dataset (data from Pantheon,
         \citealt{scolnic2018a}). The colored parts represent the distribution of
-        SNe~Ia kept in our analysis for they are supposedly free from
+        SNe~Ia that we retained in our analysis because they are assumed to be free from
     observational selection bias (see Section~\ref{sec:sample}). The
-darker (resp.\ lighter) color responds to the conservative (resp.\ fiducial)
+darker (lighter) color responds to the conservative (fiducial)
 selection cut.}
@@
 
-In addition, we use the SNe~Ia from the Nearby Supernova Factory
+In addition, we used the SNe~Ia from the Nearby Supernova Factory
 \citep[SNfactory,][]{aldering2002} published in \cite{rigault2020} and that have
-been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
+been discovered from nontargeted searches (114 SNe~Ia, see their sections~3
 and~4.2.2 ; see \citealt{aldering2020}). For this dataset, spectroscopic
-screening was done for candidates with $r \lesssim 19.5$; redshift cuts were
+screening was performed for candidates with $r \lesssim 19.5$; redshift cuts were
 then applied when selecting which SN Ia to follow, resulting in a redshift range
-of $0.02 < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} for
-extracting local host properties. These 114 SNfactory SNe~Ia are thus in the
+of $0.02 < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} to extract local host properties. These 114 SNfactory SNe~Ia are thus in the
 volume-limited part of the survey (Aldering et al., in prep.), and are therefore
 assumed to be a random sampling of the underlying SN population. The SNfactory
-sample is particularly useful for studying SN property drift, as it enables us
+sample is particularly useful for studying SN property drift because it enables us
 to have a large complete SN~Ia sample at $z<0.1$. Finally, we include the HST
@@
 these redshifts the supernovae typing is challenging, the target classification
-was robust enough to include them within the cosmological analysis
-\citep{scolnic2018a} and we do not impose further cuts.
-Section~\ref{sec:results} highlights that, while compatible with it, our results
+was robust enough to include them in the cosmological analysis
+\citep{scolnic2018a}, and we did not impose further cuts.
+Section~\ref{sec:results} highlights that while compatible with it, our results
 are not dependent on the inclusion of this dataset.
@@
 evolution of the mean stretch is shown, with the data split in redshift bins of
-regular sample size. We see that SNe~Ia at higher redshift have on average
-larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift
+regular sample size. SNe~Ia at higher redshift have a
+larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) on average than those at lower redshift
 ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch
-distribution is evolving with redshift.
+distribution evolves with redshift.
 
@@
     \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_stac_75-lb-cividis.pdf}
-    \caption{\textit{Bottom:} \textsc{\texttt{SALT2.4}} lightcurve stretch as a
+    \caption{\textit{Bottom:} \textsc{\texttt{SALT2.4}} light-curve stretch as a
         function of redshift for each survey considered in this analysis (see
-        legend). Solid (resp.\ open) markers correspond to the conservative
-        (resp.\ fiducial) redshift cuts. \textit{Top:} stacked redshift
-    histograms in dark (resp.\ light) colors for the conservative (resp. \
-fiducial) redshift cuts.}
+        legend). Solid (open) markers correspond to the conservative
+        (fiducial) redshift cuts. \textit{Top:} Stacked redshift
+    histograms in dark (light) colors for the conservative (fiducial) redshift cuts.}
     \label{fig:sample}
@@
 
-In section~\ref{ssec:cuts}, we have built volume-limited samples from a set of
-magnitude-limited ones, using simple redshift cuts. This simplified approach is
-statistically sub-optimal, but should be enough to test our key question:
-whether redshift evolution of stretch is compatible with the \cite{rigault2020}
-model. However, the possibility remains that a complex observational selection
+In section~\ref{ssec:cuts} we have built volume-limited samples from a set of
+magnitude-limited ones using simple redshift cuts. This simplified approach is
+statistically suboptimal, but should suffice to test our key question
+whether redshift evolution of stretch is compatible with the model of \cite{rigault2020}. However, the possibility remains that a complex observational selection
 function related to spectroscopic follow-up efficiencies below our fiducial (or
-even conservative) redshift cuts could still affect our sample, making it not
-fully volume-limited; this would, in turn, bias our conclusion on the
-astrophysical drift of the SNe~Ia population. We now look at this possibility.
+even conservative) redshift cuts might still affect our sample, making it not
+fully volume limited; this would in turn bias our conclusion about the
+astrophysical drift of the SNe~Ia population. We now examine this possibility.
 
 To test for the existence of potential leftover observational selection biases
-in our sample, we compare the stretch and color distributions of the SNe~Ia
+in our sample, we compared the stretch and color distributions of the SNe~Ia
 originating from different datasets having overlapping redshift ranges: these
@@
 
-The two samples that overlap the most in redshift are PS1 and SDSS in the
+The two samples that overlap most in redshift are PS1 and SDSS in the
 redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping
 subsample consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus
-the most likely to be affected by residual observational selection effects (see
+is most likely to be affected by residual observational selection effects (see
 the corresponding discussion in section~\ref{ssec:cuts}). Over that same
-redshift range, PS1 has 52 SNe~Ia, which are in the lowest redshift bins and
+redshift range, PS1 has 52 SNe~Ia that are in the lowest redshift bins and
 thus unlikely to have any observational selection issue. To identify potential
 inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib} (upper
-panels) compares the stretch and color distribution of both these surveys. The
+panels) compares the stretch and color distribution of these two surveys. The
 resulting Kolmogorov-Smirnov (KS) similarity test $p$-values ($p >10\%$) do not
@@
 
-We perform a similar analysis for PS1 and SNLS over the redshift range $0.20 < z
+We performed a similar analysis for PS1 and SNLS over the redshift range $0.20 < z
 < 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same conclusion can be
-drawn: there is no substantial sign of discrepancy in the stretch and color
-distributions between the low- and high-end of our fiducial SNLS and PS1
+drawn: There is no substantial sign of discrepancy in the stretch and color
+distributions between the low and high end of our fiducial SNLS and PS1
 samples, respectively. Nonetheless, the small size of the SNLS dataset at $z <
@@
 strong deviation would be noticeable. Extending the redshift range to $0.20 < z
-< 0.40$ (though we have no PS1 data above 0.3) allows to increase the SNLS
-subsample to 31, yet the stretch $p$-values remains high (34\%) showing no sign
+< 0.40$ (although we have no PS1 data above 0.3) allows increasing the SNLS
+subsample to 31, but the stretch $p$-value remains high (34\%), showing no sign
 of inconsistency.
@@
             different surveys overlapping in redshift. \textit{Facing up}: SDSS
-            and PS1 within the redshift range $0.10 < z < 0.20$; \textit{facing
+            and PS1 within the redshift range $0.10 < z < 0.20$. \textit{Facing
             down}: PS1 and SNLS within the redshift range $0.20 < z < 0.31$.
-            Error bars show the Poisson noise. Our stretch ``Base'' model is
+            Error bars show the Poisson noise. Our stretch base model is
             illustrated in orange at the mean redshift of the redshift ranges,
             $0.15$ and $0.25$, respectively. Kolmogorov-Smirnov test $p$-values
-            are indicated on the top (resp.\ bottom) of each panel showing no
-            sign that the SDSS and PS1 (resp. PS1 and SNLS) $x_1$ and $c$
+            are indicated at the top (bottom) of each panel and show no
+            sign that the SDSS and PS1 (PS1 and SNLS) $x_1$ and $c$
             distributions are not drawn from the same underlying
-    distributions}
+    distributions.}
     \label{fig:distrib}
@@
 We finally highlight that the SNe~Ia color is more prone to
-    observational selection effects than stretch as illustrated in
-    Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of \cite{kessler2017}. Hence,
-    since the comparison of color distributions shows no significant hint of
-    leftover observational selection effect, it further supports our claim that
+    observational selection effects than stretch, as illustrated in
+    Fig.~\ref{fig:maglim}; see also Fig.~3 of \cite{kessler2017}, for instance. Hence,
+    because the comparison of color distributions shows no significant indication of
+    leftover observational selection effect, this further supports our claim that
     our simple redshift-based selection criteria are sufficient to build the
@@
 \cite{rigault2020} presented a model for the evolution of the fraction of
-younger and older SNe~Ia as a function of redshift following former work on
-rates and delay time distributions \citep[e.g.,][]{mannucci2005,
+younger and older SNe~Ia as a function of redshift following previous work on
+rates and delay-time distributions \citep[e.g.,][]{mannucci2005,
 scannapieco2005, sullivan2006, smith2012, childress2014, maozmannucci2014}.
-In short, it was assumed that the number of ``young'' SNe~Ia follows the star
-formation rate (SFR) in the Universe, while the number of ``old'' SNe~Ia follows
-the number of Gyr-old stars in the Universe, i.e. the stellar mass (M$^*$).
-Hence, if we denote $\delta(z)$ (resp. $\psi(z) = 1-\delta(z)$) the fraction of
-young (resp.\ old) SNe~Ia in the Universe as a function of redshift, then the
+In short, it was assumed that the number of young SNe~Ia follows the star
+formation rate (SFR) in the Universe, while the number of old SNe~Ia follows
+the number of billion-old stars in the Universe, that is, the stellar mass (M$^*$).
+Hence, if we denote $\delta(z)$ ( $\psi(z) = 1-\delta(z)$) the fraction of
+young (old) SNe~Ia in the Universe as a function of redshift, then the
 ratio $\delta/\psi$ is expected to follow the evolution of the specific star
@@
 
-\subsection{``Base'' underlying stretch distribution}
+\subsection{Base underlying stretch distribution}
 \label{sec:basemodel}
@@
 To model the evolution of the full SN stretch distribution as a function of
-redshift, given our aforementioned model of the evolution of the fraction of
-younger and older SNe~Ia with cosmic time, we need to model the SN stretch
-distribution for each age subsample.
-
-\cite{rigault2020} presented the relation between SN stretch and LsSFR
+redshift, we need to model the SN stretch
+distribution for each age subsample given our aforementioned model of the evolution of the fraction of
+younger and older SNe~Ia with cosmic time. \cite{rigault2020} presented the relation between SN stretch and LsSFR
 measurement, a progenitor age tracer, using the SNfactory sample. This relation
@@
 current analysis. Given the structure of the stretch-LsSFR scatter plot, our
-model of the underlying SN~Ia stretch distribution is defined as follows:
+model of the underlying SN~Ia stretch distribution is defined as follows\LEt{please reformat so that this becomes proper text, there is no need for a two-item bulleted list}:
 
@@
     \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
-    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$)
-        as a function of the local specific star formation rate (LsSFR) for
-        SNfactory SNe used in this analysis. The color corresponds to the
-        probability, $p_y$, for the SNe~Ia to be young, i.e. to have
+    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} light-curve stretch ($x_1$)
+        as a function of the LsSFR for
+        SNfactory SNe. The color corresponds to the
+        probability, $p_y$, for the SNe~Ia to be young, i.e., to have
         $\log\mathrm{LsSFR} \geq -10.82$ \citep[see][]{rigault2020}.
         \textit{Right}: $p_y$-weighted histogram of the SN stretches, as well as
-        the adjusted Base model; the younger and older population contributions
+        the adjusted base model; contributions of the younger and older population
         are shown in purple and yellow, respectively.}
@@
         \sigma_2{}^2)$, where one mode is the same as for the young population,
-        $a$ representing the relative influence of the two modes.
+        $a$ representing the relative effect of the two modes.
 \end{itemize}
@@
 by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
-But generally, the fraction of young SNe~Ia as a function of redshift is given
-by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
+In general, however, the fraction of young SNe~Ia as a function of redshift is given
+by $\delta(z)$ (see Eq.~\ref{eq:delta}), and therefore our redshift drift model
 of the underlying stretch distribution of SNe~Ia as a function of redshift
-$X_1(z)$ is given by:
+$X_1(z)$ is given by
 \begin{align}\label{eq:stretchz}
@@
 \end{align}
-This is our Base drifting model.
+This is our base drifting model.
 
@@
 
-Given the probability $y^i$ that a given SN is young, and assuming our Base
-model (see Section~\ref{sec:basemodel}), the probability to measure a
-\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
+Given the probability $y^i$ that a given SN is young and assuming our base
+model (see Section~\ref{sec:basemodel}), the probability of measuring a
+\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by
 \begin{align}\label{eq:likelihoodsnf}
@@
     & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
-    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg]
+    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg].
 \end{align}
 
-The maximum-likelihood estimate of the 5~free parameters
+The maximum-likelihood estimate of the five~free parameters
 $\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained
@@
 
-Depending on whether $y^i$ can be estimated directly from LsSFR measurements or
-not, there are two ways to proceed, which we now discuss.
+Depending on whether $y^i$ can be estimated directly from LsSFR measurements, there are two ways to proceed. We discuss them below.
 
@@
 
-For the SNfactory sample, we can readily set $y^i = p^i_y$, the probability to
-have $\log(\textrm{LsSFR}) \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
-minimize Eq.~\ref{eq:likelihood} with respect to $\vec{\theta}$. Results on
+For the SNfactory sample, we can readily set $y^i = p^i_y$, the probability of
+having $\log(\textrm{LsSFR}) \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
+minimize Eq.~\ref{eq:likelihood} with respect to $\vec{\theta}$. Results of
 fitting the SNf SNe with this model are shown Table~\ref{tab:modelresults} and
@@
     \centering
-    \caption{Best fit values of the parameters for the Base stretch distribution
+    \caption{Best-fit values of the parameters for the base stretch distribution
     model when applied to the SNfactory dataset only (114 SNe~Ia), the fiducial
-569 SN~Ia sample or the conservative one (422).}
+569 SN~Ia sample, or the conservative model (422).}
     \label{tab:modelresults}
@@
 
-When lacking direct LsSFR measurements (i.e. $p_y^i$), we can extend the
-analysis to non-SNfactory samples by using the redshift-evolution of the
+When direct LsSFR measurements are lacking (i.e., $p_y^i$), we can extend the
+analysis to non-SNfactory samples by using the redshift evolution of the
 fraction $\delta(z)$ of young SNe~Ia (Eq.~\ref{eq:delta}) as a proxy for the
@@
 $\vec{\theta}\equiv(\mu_1, \mu_2, \sigma_1, \sigma_2, a)$ of the stretch
-distribution $X_1$ (Eq.~\ref{eq:stretchz}), but this time assuming $y^i =
+distribution $X_1$ (Eq.~\ref{eq:stretchz}), but this time, assuming $y^i =
 \delta(z^i)$ for any given SN~$i$. 
@@
 \\
-For the rest of the analysis, we will therefore minimize Eq.~\ref{eq:likelihood}
-using $p_y^i$ --- the probability for the SN $i$ to be young --- when available
-(i.e. for SNfactory dataset), and $\delta(z^i)$ --- the expected fraction of
-young SNe~Ia at the SN redshift $z^i$ --- otherwise.
+For the remaining analysis, we therefore minimized Eq.~\ref{eq:likelihood}
+using $p_y^i$ , the probability for the SN $i$ to be young, when available
+(i.e., for SNfactory dataset), and $\delta(z^i),$ the expected fraction of
+young SNe~Ia at the SN redshift $z^i$  , otherwise.
 
 Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
-(resp.\ conservative) sample are given Table~\ref{tab:modelresults}, and the
+(conservative) sample are given Table~\ref{tab:modelresults}, and the
 predicted redshift evolution of mean stretch (expected $x_1$ given the
 distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
-Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances.
-We see in this figure that the measured mean SN~Ia stretch per redshift bins of
+Fig.~\ref{fig:modelall} accounting for parameter errors and their covariances.
+This figure shows that the measured mean SN~Ia stretch per redshift bin of
 equal sample size closely follows our redshift drift modeling. This is indeed
 what is expected if old environments favor low SN stretches
-\citep[e.g.][]{howell2007} \textit{and} if the fraction of old SNe~Ia declines as
+\citep[e.g.,][]{howell2007} \textit{\textup{and}} if the fraction of old SNe~Ia declines as
 a function of redshift. See Section~\ref{sec:results} for a more quantitative
@@
         as a function of redshift. Markers show the stretch plain mean
-        (with error estimated from scatter) measured in redshift bins of equal
+        (the error is estimated from the scatter) measured in redshift bins of equal
         sample size, indicated in light gray at the bottom of each redshift bin.
-        Full and light markers are used when considering the fiducial or the
-        conservative samples, respectively. The orange horizontal line
-        represents the mean stretch for the non-evolving Gaussian model
+        Full and light markers are used when the fiducial or conservative samples are considered, respectively. The horizontal orange line
+        represents the mean stretch for the nonevolving Gaussian model
         (last line of Table~\ref{tab:comp}) on the fiducial sample. Best
-        fits of our Base drifting model are shown as blue, dashed-blue and
-        gray, when fitted on the fiducial sample, the conservative one or
+        fits of our base drifting model are shown as blue, dashed blue, and
+        gray when fitted on the fiducial sample, the conservative sample, or
         the SNfactory dataset only, respectively; all are compatible. The
-        light-blue band illustrates the amplitude of the error (incl.\
-        covariance) of the best fit model when considering the fiducial
+        light blue band illustrates the amplitude of the error (including\
+        covariance) of the best-fit model when considering the fiducial
     dataset.}
@@
 
-In Section~\ref{sec:basemodel}, we have modeled the underlying stretch
-distribution following \cite{rigault2020}, i.e.\ as a single Gaussian for the
-``young'' SNe~Ia and a mixture of two Gaussians for the ``old'' SNe Ia
+In Section~\ref{sec:basemodel} we have modeled the underlying stretch
+distribution following \cite{rigault2020}, that is,\ as a single Gaussian for the
+young SNe~Ia and a mixture of two Gaussians for the old SNe Ia
 population, one being the same as for the young population, plus another one for
-the fast-declining SNe~Ia that seem to only exist in old local environments.
-This is our so-called ``Base'' model. However, to test different modeling
-choices, we have implemented a suite of alternative parameterizations that we
-also adjust to the data following the procedure described in
+the fast-declining SNe~Ia that appear to only exist in old local environments.
+This is our so-called base model. However, to test different modeling
+choices, we implemented a suite of alternative parameterizations that we
+also adjusted to the data following the procedure described in
 Section~\ref{sec:modelnopy}.
@@
 single normal distribution for each of the young and old populations. We thus
-consider a ``Howell+drift'' model, with one single Gaussian per age group and
+considered a Howell+drift model, with one single Gaussian per age group and
 the $\delta(z)$ drift from Eq.~\ref{eq:delta}.
 
-Alternatively, since we aim at probing the existence of an evolution with
-redshift, we also test constant models by restricting the ``Base'' and
-``Howell'' models to use a supposedly redshift-independent fraction $\delta(z)
-\equiv f$ of young SNe; these models are hereafter labeled ``Base+constant'' and
-``Howell+constant''.
-
-We also consider another intrinsically non-drifting model, the functional form
-developed for Beams with Bias Correction \cite[BBC,][]{scolnic2016,
+Alternatively, because we aim to probe the existence of an evolution with
+redshift, we also tested constant models by restricting the base and
+Howell models to use an assumed redshift-independent fraction $\delta(z)
+\equiv f$ of young SNe; these models are hereafter labeled base+constant and
+Howell+constant.
+
+We also considered another intrinsically nondrifting model, the functional form
+developed for beams with bias correction \cite[BBC,][]{scolnic2016,
 kessler2017}, used in recent SN cosmological analyses
-\cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to account
+\cite[e.g.,][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to account
 for Malmquist biases. The BBC formalism assumes sample-based (hence
-intrinsically non-drifting) asymmetric Gaussian stretch distributions:
+intrinsically nondrifting) asymmetric Gaussian stretch distributions:
 $\mathcal{N}\left(\mu, \sigma_-{}^2\; \text{if} \;x_1<\mu,\; \text{else}
 \;\sigma_+{}^2\right)$. The idea behind this sample-based approach is twofold:
-(1) Malmquist biases are driven by survey properties and (2) since current
+(1) Malmquist biases are driven by survey properties, and (2) because current
 surveys cover limited redshift ranges, having a sample-based approach covers
 some potential redshift evolution information \citep{scolnic2016, scolnic2018a}.
-See further discussion concerning BBC in Section~\ref{sec:discussion}.
-
-Finally, for the sake of completeness, we also consider redshift-independent
+See a more detailed discussion of BBC in Section~\ref{sec:discussion}. Finally, for the sake of completeness, we also considered redshift-independent
 pure and asymmetric Gaussian models.
@@
 We adjusted each of the models described above on both the fiducial and
-conservative samples (cf. Section~\ref{sec:sample}); results are gathered
-in Table~\ref{tab:comp}, and illustrated in Fig.~\ref{fig:mod_comp}.
+conservative samples (cf. Section~\ref{sec:sample}). The results are gathered
+in Table~\ref{tab:comp} and are illustrated in Fig.~\ref{fig:mod_comp}.
 
@@
     \caption{Comparison of the relative ability of each model to describe the
-        data. For each considered model, we report if the model is drifting or
-        not, its number of free parameters and, for both the fiducial and the
+        data. For each considered model, we report whether the model is drifting, its number of free parameters, and for both the fiducial and the
         conservative cuts, $-2\ln(L)$ (see Eq.~\ref{eq:likelihood}), the AIC and
-        the AIC difference ($\Delta$AIC) between this model and the Base model
-        used as reference for it has the lowest AIC.}
+        the AIC difference ($\Delta$AIC) between this model and the base model
+        used as reference because it has the lowest AIC.}
     \label{tab:comp}
@@
     \includegraphics[width=\linewidth]{Article_figures/mod_comp.pdf}
-    \caption{$\Delta$AIC between ``Base'' model (reference) and other models
+    \caption{$\Delta$AIC between base model (reference) and other models
         (see Table~\ref{tab:comp}). Full and open blue markers correspond to
@@
         the results when the analysis is performed on the conservative sample
-        rather than the fiducial one. Color-bands illustrate the validity of
-        the models, from Acceptable ($\Delta\mathrm{AIC} > -5$) to Excluded
+        rather than the fiducial one. Color bands illustrate the validity of
+        the models, from acceptable ($\Delta\mathrm{AIC} > -5$) to excluded
         ($\Delta\mathrm{AIC} < -20$), see text. According to the AIC, all
-        non-drifting models (open symbols) are excluded to be as good a
-        representation of the data as the Base (drifting) model.}
+        nondrifting models (open symbols) are excluded as a poorer representation of the data than the base (drifting) model.}
     \label{fig:mod_comp}
@@
 
-Since the various models have different degrees of freedom, we use the Akaike
-Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare their ability
-to properly describe the observations. This estimator penalizes extra degrees of
-freedom to avoid over-fitting the data, and is defined as follow:
+Because the various models have different degrees of freedom, we used the Akaike
+information criterion \citep[AIC, e.g.,][]{burnham2004} to compare their ability
+to properly describe the observations. This estimator penalizes additional degrees of
+freedom to avoid overfitting the data and is defined as follows:
 \begin{equation}
     \mathrm{AIC} = -2\ln(L) + 2k
-\end{equation}
+,\end{equation}
 where $-2\ln(L)$ is derived by minimizing Eq.~\eqref{eq:likelihood}, and $k$ is
-the number of free parameters to be adjusted. The reference model is the one
-with the smallest AIC; in comparison to this model, the models with
-$\Delta\mathrm{AIC}<5$ are coined acceptable, the ones with
-$5<\Delta\mathrm{AIC}<20$ are unfavored, and those with $\Delta\mathrm{AIC}>20$
-are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
+the number of free parameters to be adjusted. The reference model has the smallest AIC; in comparison to this model, the models with
+$\Delta\mathrm{AIC}<5$ are coined acceptable, those with
+$5<\Delta\mathrm{AIC}<20$ are not favored, and those with $\Delta\mathrm{AIC}>20$
+are deemed excluded. This roughly corresponds to 2, 3, and 5~$\sigma$ limits for
 a Gaussian probability distribution.
 
-The best model (with smallest AIC) is the so-called Base model and thus is our
-reference model; this is true both on the fiducial and conservative samples.
-The Base model also has the smallest $-2\ln(L)$, making it the most likely
-model even ignoring the over-fitting issue accounted for by the AIC formalism.
+The best model (with the smallest AIC) is the so-called base model and thus is our
+reference model; this is true for the fiducial and conservative samples.
+The base model also has the smallest $-2\ln(L)$, making it the most likely
+model even when the overfitting issue that is accounted for by the AIC formalism is ignored.
 
 Furthermore, we find that redshift-independent stretch distributions are all
-excluded as suitable descriptions of the data relative to the Base model. In
-fact, the best non-drifting model (the Asymmetric one) has a very marginal
+excluded as suitable descriptions of the data relative to the base model. The best nondrifting model (the asymmetric model) has a very marginal
 chance ($p \equiv \exp\left(\Delta\mathrm{AIC}/2\right) = 5\times10^{-6}$) to
-describe the data as well as the Base model. This result is just a quantitative
-assessment of qualitative facts clearly visible in Fig.~\ref{fig:modelall}: the
+describe the data as well as the base model. This result is just a quantitative
+assessment of qualitative facts that are clearly visible in Fig.~\ref{fig:modelall}: The
 mean SN stretch per bin of redshift strongly suggests a significant redshift
@@
 implementations of the BBC technique \citep{scolnic2016, kessler2017} has one of
-the highest AIC value in our analysis (see Section~\ref{sec:results}). While its
+the highest AIC values in our analysis (see Section~\ref{sec:results}). While its
 $-2\ln(L)$ is the smallest of all redshift-independent models (but still $-11.5$
-worse than the reference Base model), it is strongly penalized for requiring
-15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the 5~samples of the
+lower than the reference base model), it is strongly penalized for requiring
+15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the five~samples of the
 analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted as a
 probability $p=2\times 10^{-7}$ of being as good a representation of the data
-as the Base model.
-
-We note that, when comparing models adjusted on individual subsamples rather
-than globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
-k\ln(n)$, with $n$ the number of data points) might be better suited than AIC,
-since it explicitly accounts for the fact that each subsample is fitted
+as the base model.
+
+We note that when models are compared that were adjusted on individual subsamples rather
+than globally, the Bayesian information criterion ($\mathrm{BIC} = -2\ln(L) +
+k\ln(n)$, with $n$ the number of data points) might be better suited than AIC because it explicitly accounts for the fact that each subsample is fitted
 separately: the sample-based model BIC is rightfully the sum of the BIC for each
 sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the sample-based
-asymmetric Gaussian model as being as pertinent as the Base model.
-
-In order to ensure that our results are not driven by the incompletely-modeled
-HST subsample, we recompute $\Delta$AIC for each model excluding this dataset;
-we find that it does not change $\Delta$AIC by more than few tenths. The
-consistency of these values with those in Table~\ref{tab:comp} show that the HST
+asymmetric Gaussian model as being as pertinent as the base model.
+
+In order to ensure that our results are not driven by the incompletely modeled
+HST subsample, we recomputed $\Delta$AIC for each model excluding this dataset;
+this did not change $\Delta$AIC by more than few tenths. The
+consistency of these values with those in Table~\ref{tab:comp} shows that the HST
 subsample does not drive our conclusions.
@@
 We report in Table~\ref{tab:bbc} our determination of $\mu_0$ and
-$\sigma_{\pm}$ for each sample when implementing an asymmetric Gaussian
-model, adjusted on the nominally selection-free samples using our fiducial cuts
-(see Section~\ref{sec:sample}). We find our results in close agreement with
-\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
+$\sigma_{\pm}$ for each sample when an asymmetric Gaussian
+model was implmented, and adjusted on the nominally selection-free samples using our fiducial cuts
+(see Section~\ref{sec:sample}). Our results are in close agreement with those of
+\cite{scolnic2016} for the SNLS and SDSS and with the results reported by \cite{scolnic2018a} for PS1, who
 derived these model parameters using the full BBC formalism, using numerous
-simulations to model the observational selection effects (see details
+simulations to model the observational selection effects (see details,
 e.g., Section~3 of \citealt{kessler2017}). The agreement between our fit of the
@@
 constructing a sample with negligible observational selection
-effects. If we were to use \cite{scolnic2016} and \cite{scolnic2018a} best fit
-values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for SNLS, SDSS and
-PS1, respectively, the $\Delta$AIC between our Base drifting model and the BBC
+effects. If we were to use the \cite{scolnic2016} and \cite{scolnic2018a} best-fit
+values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for the SNLS, SDSS and
+PS1, respectively, the $\Delta$AIC between our base drifting model and the BBC
 modeling would go even deeper from $-32$ to $-47$. We further discuss the
@@
 This reinforces our assumption that the young and old populations
-indeed appear to share the same underlying high-stretch mode. Furthermore, one
-might wonder whether a low-stretch mode might also exist in the young-population
-(see Fig.~\ref{fig:stretchlssfr}). We test for this by allowing this
+indeed appear to share the same underlying high-stretch mode. Furthermore, we
+might wonder whether a low-stretch mode might also exist in the young population
+(see Fig.~\ref{fig:stretchlssfr}). We tested for this by allowing this
 population to also be bimodal, finding the amplitude of such a
@@
 
-Finally, ignoring the LsSFR measurements --- available only for the SNfactory
-dataset (see Section~\ref{sec:modeling}) --- reduces the significance of the
-results presented in this section, as expected. Even so, non-drifting
-models remain strongly disfavored. For instance, the best
-fitting sample-based Gaussian asymmetric model is still
-$\Delta\mathrm{AIC}<-10$ less representative of the data than our Base drifting
+Finally, ignoring the LsSFR measurements, which are only available for the SNfactory
+dataset (see Section~\ref{sec:modeling}), reduces the significance of the
+results presented in this section, as expected. Even so, nondrifting
+models remain strongly disfavored. For instance, the best-fitting sample-based Gaussian asymmetric model is still
+$\Delta\mathrm{AIC}<-10,$ which is less representative of the data than our base drifting
 model.
@@
 To the best of our knowledge, a SN~Ia stretch redshift drift modeling has never
-been explicitly used in cosmological analyses, though Bayesian hierarchy
-formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016} or Steve
-\citep{hinton2019} can easily allow it (see e.g., sections~1.3 and 2.5 of
-\cite{rubin2015}). Not doing so is a second order issue for SN cosmology, as it
-only affects the way one accounts for Malmquist bias. Indeed, as long as the
+been explicitly used in cosmological analyses, although a Bayesian hierarchy
+formalism such as UNITY \citep{rubin2015}, BAHAMAS \citep{shariff2016}, or Steve
+\citep{hinton2019} can easily allow it (see, e.g., sections~1.3 and 2.5 of
+\cite{rubin2015}). Not doing so is a second-order issue for SN cosmology because it
+only affects the way in which the Malmquist bias is accounted for. As long as the
 Phillips relation \citep{phillips1993} standardization parameter $\alpha$ is not
-redshift dependent (a study beyond the scope of this paper, but see
-e.g. \citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
-cosmology are blind to the underlying stretch distribution for complete samples.
+redshift dependent (a study that is beyond the scope of this paper, but see,
+e.g., \citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
+cosmology are indeed blind to the underlying stretch distribution for complete samples.
 However, surveys usually do have significant Malmquist bias for the upper half
@@
 Commonly used Malmquist bias correction techniques, such as the BBC-formalism,
-assume per sample asymmetric Gaussian functions to model the underlying
-stretch and color distributions. Yet, as shown in Section~\ref{sec:results},
-such a sample-based distribution is excluded in comparison to our
-drifting model. Then, contrary to what \citet[][Section~2]{scolnic2016}
-and \citet[][Section~5.4]{scolnic2018a} have suggested --- i.e., that
+assume per-sample asymmetric Gaussian functions to model the underlying
+stretch and color distributions. As shown in Section~\ref{sec:results},
+however, such a sample-based distribution is excluded in comparison to our
+drifting model. In contrast to what \citet[][Section~2]{scolnic2016}
+and \citet[][Section~5.4]{scolnic2018a} have suggested, that is, that
 traditional surveys span sufficiently limited redshift ranges
-such that the per-sample approach accounts for implicit redshift drifts
---- a direct modeling of the redshift drift is more appropriate than a
-sample-based approach. We add here that, as measurements of modern
+such that the per-sample approach accounts for implicit redshift drifts,
+a direct modeling of the redshift drift is therefore more appropriate than a
+sample-based approach. We add here that as measurements of modern
 surveys try to cover increasingly larger redshift ranges in order to reduce
 calibration systematic uncertainties, this sample-based approach
-becomes less valid, notably for PS1, DES and, soon, LSST.
+becomes less valid, notably for PS1, DES and, soon, the LSST.
 
@@
 underlying stretch distribution between the per-sample asymmetric modeling and
-our Base drifting model for the PS1 sample. Our model is bimodal and the
+our base drifting model for the PS1 sample. Our model is bimodal, and the
 relative amplitude of each mode depends on the redshift-dependent fraction of
@@
 lower redshift), the higher the amplitude of the old-specific low-stretch mode.
-This redshift dependency of the underlying stretch distributions is
-shown running from blue to red in Fig.~\ref{fig:bbc_pdf_ps1} for the
+This redshift dependence on the underlying stretch distributions is
+shown with colors from blue to red in Fig.~\ref{fig:bbc_pdf_ps1} for the
 redshift range covered by PS1. The observed $x_1$ histogram follows the
@@
 distribution goes through the middle of the bimodal distribution,
-over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and under-estimating it at
-$x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN
+overestimating the number of SNe~Ia at $x_1\sim-0.7$ and underestimating it at
+$x_1\sim-1.7$ in comparison to our base drifting model for typical PS1 SN
 redshifts. This means that the SN bias-corrected standardized magnitude
-estimated at a redshift suffering from by observational
+estimated at a redshift plagued by observational
 selection effects would be biased by a mismodeling of the true
@@
         (its error); the dashed line shows the \cite{scolnic2018a} result. The
-        black line (band) shows our best fitted Base-modeling (its error, see
+        black line (band) shows our best-fit base modeling (its error, see
         Table \ref{tab:modelresults}) that includes redshift drift. For
-        illustration, we show as colored (from blue to red with increasing
+        illustration, we show (colored from blue to red with increasing
         redshifts) the evolution of the underlying stretch distribution as a
@@
 the scope of this paper given the complexity of the BBC analysis. It would
-require a full study using our Base model (Eq.~\ref{eq:stretchz}) in place of
+require a full study using our base model (Eq.~\ref{eq:stretchz}) in place of
 the sample-based asymmetric modeling as part of the BBC simulations. However, we
-already highlight that even if a non-drifting sample-based model could provide
+already highlighted that even if a nondrifting sample-based model could provide
 comparable result in the volume-limited part of the various samples, these
-models would differ when extrapolating to higher redshifts, precisely where the
+models would differ when extrapolated to higher redshifts, precisely where the
 underlying distribution will matter for correcting Malmquist biases.
@@
 In the era of modern cosmology, where we aim to measure $w_0$ at a
-sub-percent level and $w_a$ with ten-percent precision
+subpercent level and $w_a$ with 10\%\ precision
 \citep[e.g.,][]{lsstpaper}, we stress that correct modeling of potential SN
-redshift drift should be further studied and care should be taken when using
-samples affected by observational selection effects.
+redshift drift should be further studied and care should be taken when
+samples are used that are affected by observational selection effects.
 
@@
 effectively volume-limited SN~Ia subsamples from the Pantheon dataset
-\citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to which we added HST and SNfactory
+\citep[][SDSS, PS1, and SNLS]{scolnic2018a}, to which we added HST and SNfactory
 data from \cite{rigault2020} for the high- and low-redshift bins. We only
 considered the SNe that have been discovered in the redshift range of each
-survey where observational selection effects are negligible, so that
+survey in which observational selection effects are negligible, so that
 the observed SNe~Ia stretches are a random sampling of the true underlying
@@
 Following predictions made in \cite{rigault2020}, we introduced a redshift drift
-model which depends on the expected fraction of ``young'' and ``old'' SNe~Ia as
+model that depends on the expected fraction of young and old SNe~Ia as
 a function of redshift, with each age population having its own
@@
 
-In addition to this ``base'' model, we have studied various
-distributions, including redshift independent models; we also studied the
+In addition to this base model, we studied various
+distributions, including redshift-independent models; we also studied the
 prediction from a per-sample asymmetric Gaussian stretch distribution used, for
-instance, by the Beams with Bias Correction Malmquist bias correction algorithm
+instance, by the beams with bias correction Malmquist bias correction algorithm
 \citep{scolnic2016, kessler2017}.
-
-Our conclusions are the following:
+Our conclusions are listed below.
 \begin{enumerate}
@@
     \item The underlying SN~Ia stretch distribution is significantly redshift
-        dependent, as previously suggested by e.g.~\cite{howell2007}, in
+        dependent, as previously suggested by~\cite{howell2007}, for example, in
         a way that observational selection effects alone cannot explain. This
-        result is largely independent of details on each age-population model.
+        result is largely independent of the details of each age-population model.
 
     \item Redshift-independent models are quantitatively excluded as
-        suitable descriptions of the data relative to our Base model. This model
-        assumes that: (1) the younger population has a unimodal Gaussian stretch
-        distribution, while the older population stretch distribution is
-        bimodal, one mode being the same as the young one; (2) the evolution of
+        suitable descriptions of the data relative to our base model. This model
+        assumes that (1) the younger population has a unimodal Gaussian stretch
+        distribution while the older population stretch distribution is
+        bimodal, one mode being the same as the young one, and (2) the evolution of
         the relative fraction of younger and older SNe~Ia follows the prediction
@@
     \item Models using survey-based asymmetric Gaussian distributions,
-        e.g., as employed in the current implementation of BBC, are
+        for instance, as employed in the current implementation of BBC, are
         excluded as a good description of the data relative to our
-        drifting model. Hence, the sample-based approach does not accurately
+        drifting model. This means that the sample-based approach does not accurately
         account for redshift drift, a problem that will be exacerbated
@@
             (1-\delta(z))&\times \left[a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
-            (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2)\right]
+            (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2)\right],
             \tag{\ref{eq:stretchz}}
         \end{align*}
-        with $a=0.51$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
+        with $a=0.51$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$, and
         $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}), and using the
-        age-population drift model \begin{align*}
+        age-population drift model, \begin{align*}
             \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
@@
 
-In this paper, we considered a simple two-population Gaussian mixture
+We considered a simple two-population Gaussian mixture
 modeling. Additional data free from significant Malmquist bias would enable
-us to refine it, as necessary. We note that samples at the low- and high-redshift
+us to refine it as required. We note that samples at the low- and high-redshift
 ends of the Hubble diagram would be particularly helpful for this drifting
-analysis; fortunately this will soon be provided by the Zwicky Transient
-Facility \citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange
+analysis; fortunately, this will soon be provided by the Zwicky Transient
+Facility \citep[low-$z$,][]{bellm2019, graham2019} and the Subaru and SeeChange
 SNe~Ia programs (high-$z$), respectively.
@@
 the SNANA framework \citep{SNANA}, both to more accurately account for
-observational selection functions and to test the impact of our model on the
-derivation of cosmological parameters; this study is underway.
+observational selection functions and to test the effect of our model on the
+derivation of cosmological parameters. This study is currently being conducted.
 

### 2021-03-05 Nicolas.tex
Implemented changes and put own modifications in bf

@@
 
-\newcommand{\LEt}[1]{\textcolor{red}{\textbf{ #1}}}
 \newcommand{\orcid}[1]{
@@
     and as survey statistics increase, the question of astrophysical systematic
-    uncertainties arises, notably that of the evolution of SN~Ia populations. We study the dependence on redshift of the SN~Ia \texttt{SALT2.4}
-    light-curve stretch, which is a purely intrinsic SN property, to probe its potential
-    redshift drift. The SN stretch has been shown to be strongly correlated with the
-    SN environment, notably with stellar age tracers. We modeled the underlying
-    stretch distribution as a function of redshift, using the evolution of the
-    fraction of young and old SNe~Ia as predicted using the SNfactory dataset,
-    and assuming a constant underlying stretch distribution for each age
-    population consisting of Gaussian mixtures. We tested our prediction against
-    published samples that were cut to have marginal magnitude selection effects, so that
-    any observed change is indeed astrophysical and not observational in origin.
-    In this first study, there are indications that the underlying SN~Ia stretch
-    distribution evolves as a function of redshift, and that the young/old\LEt{the slash commonly denotes "ratio", as in "S/N" for "signal-to-noise ratio".
-The way you use it here is ambiguous: do you mean "and" or "or", or maybe a range?\ Please rephrase for clarity}
-    drifting model is a better description of the data than any time-constant
-    model, including the sample-based asymmetric distributions that are often used to
-    correct Malmquist bias at a significance higher than 5 $\sigma$. The favored
+    uncertainties arises, notably that of the evolution of SN~Ia populations. We
+    study the dependence on redshift of the SN~Ia \texttt{SALT2.4} light-curve
+    stretch, which is a purely intrinsic SN property, to probe its potential
+    redshift drift. The SN stretch has been shown to be strongly correlated with
+    the SN environment, notably with stellar age tracers. We modeled the
+    underlying stretch distribution as a function of redshift, using the
+    evolution of the fraction of young and old SNe~Ia as predicted using the
+    SNfactory dataset, and assuming a constant underlying stretch distribution
+    for each age population consisting of Gaussian mixtures. We tested our
+    prediction against published samples that were cut to have marginal
+    magnitude selection effects, so that any observed change is indeed
+    astrophysical and not observational in origin. In this first study, there
+    are indications that the underlying SN~Ia stretch distribution evolves as a
+    function of redshift, and that the \textbf{age} drifting model is a better
+    description of the data than any time-constant model, including the
+    sample-based asymmetric distributions that are often used to correct
+    Malmquist bias at a significance higher than 5 $\sigma$. The favored
     underlying stretch model is a bimodal one, composed of a high-stretch mode
-    shared by both young and old environments, and a low-stretch mode that is exclusive
-    to old environments. The precise effect of the redshift evolution of the intrinsic properties of a SN
-    Ia population on cosmology remains to be studied. The astrophysical drift of the SN stretch distribution does affect current
-    Malmquist bias corrections, however, and thereby the distances that are derived based on SN that are affected by
-    observational selection effects. We highlight that this bias will increase
-with surveys covering increasingly larger redshift ranges, which is particularly
-important for the \LEt{please spell this out here and in the main text, you use the abbreviation only one time each}LSST.}
+    shared by both young and old environments, and a low-stretch mode that is
+    exclusive to old environments. The precise effect of the redshift evolution
+    of the intrinsic properties of a SN Ia population on cosmology remains to be
+    studied. The astrophysical drift of the SN stretch distribution does affect
+    current Malmquist bias corrections, however, and thereby the distances that
+    are derived based on SN that are affected by observational selection
+effects. We highlight that this bias will increase with surveys covering
+increasingly larger redshift ranges, which is particularly important for the
+Large Synoptic Survey Telescope.}
 
@@
 Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
-enabled the discovery that the Universe expands
-\citep{riess1998, perlmutter1999}. They remain a key cosmological probe today for
-understanding the properties of dark energy (DE) because it\LEt{grammatically, "it" here refers to "DE".
-Is this correct, or should this be "they" because you still refer to the SNe Ia, or do you mean "the probe"?
-If you mean the probe, please rephrase to "because this probe" for clarity} is the only tool that can
-precisely map the recent expansion rate ($z<0.5$) when DE is driving it
-\citep[e.g.,][]{scolnicastro2020}. Type Ia supernovae are also the key to directly measuring the
-Hubble constant ($H_0$) if their absolute magnitude can be calibrated
-\citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ tha tis derived when
-the SNe~Ia are anchored to Cepheids \citep[the\LEt{please provide the spelled-out versions of all instruments and surveys at first occurrence.
-If this interrupts the flow of the sentence too much, it may be done as a footnote} SH0ES
+enabled the discovery that the Universe\textbf{'s expansion accelerates}
+\citep{riess1998, perlmutter1999}. They remain a key cosmological probe today
+for understanding the properties of dark energy (DE) because they are the only
+tool that can precisely map the recent expansion rate ($z<0.5$) when DE is
+driving it \citep[e.g.,][]{scolnicastro2020}. Type Ia supernovae are also the
+key to directly measuring the Hubble constant ($H_0$) if their absolute
+magnitude can be calibrated \citep{riess2016, freedman2019}. Interestingly, the
+value of $H_0$ that is derived when the SNe~Ia are anchored to Cepheids
+\citep[the Supernovae, $H_0$, for the Equation of State of dark energy
 project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
@@
 oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
-branch technique in place of the Cepheids seems to favor an intermediate value of
-$H_0$ \citep{freedman2019, freedman2020}, time-delay measurements from strong
+branch technique in place of the Cepheids seems to favor an intermediate value
+of $H_0$ \citep{freedman2019, freedman2020}, time-delay measurements from strong
 lensing also appear to favor high $H_0$ values \citep{wong2019}.
 
-The $H_0$ conundrum has received much attention because it could be a sign of new
-fundamental physics. No simple solution is so far able to accommodate this $H_0$
-conundrum when all other probes are accounted for, however \citep{knox2019}. Alternatively,
-systematic effects affecting one or several of the aforementioned analyses
-might explain at least some of this discrepancy. \cite{rigault2015}
-suggested that SNe Ia from the Cepheid calibrator sample differ by construction
-from those in the Hubble flow sample, as the former strongly favors
-young stellar populations while the latter does not. This selection
-effect would affect the derivation of $H_0$ if SNe~Ia from young and older
-environments differed in average standardized magnitudes.
+The $H_0$ conundrum has received much attention because it could be a sign of
+new fundamental physics. No simple solution is so far able to accommodate this
+$H_0$ conundrum when all other probes are accounted for, however
+\citep{knox2019}. Alternatively, systematic effects affecting one or several of
+the aforementioned analyses might explain at least some of this discrepancy.
+\cite{rigault2015} suggested that SNe Ia from the Cepheid calibrator sample
+differ by construction from those in the Hubble flow sample, as the former
+strongly favors young stellar populations while the latter does not. This
+selection effect would affect the derivation of $H_0$ if SNe~Ia from young and
+older environments differed in average standardized magnitudes.
 
@@
 currently used in cosmological analyses \citep[e.g.,][]{betoule2014,
-scolnic2018a}, including to derive $H_0$ \citep{riess2016, riess2019}. The underlying connection between the SNe and their host galaxies remains
-unclear, however, particularly when global properties such as the host stellar
-mass are used. This in turn raises the question of the accuracy of such corrections because
-the global properties of the host galaxies evolve with redshift. More recently, studies have
-used the local SN host galaxy environment to probe more direct connections
-between the SNe and their host galaxy environments \citep{rigault2013}, showing
-that local age tracers such as the local specific star formation rate (LsSFR) or
-the local color are more strongly correlated with the standardized SN magnitude
-\citep{roman2018, kim18, rigault2020}. The identification of SN~Ia spectral features that are correlated with LsSFR
-\citep{nordin2018} further support this connection. These results suggest age as the driving parameter
-underlying the mass step. If this is true, itwould have a significant effect for
-cosmology because the environmental correction that would need to be applied to SN standardization
-could strongly vary with redshift \citep{rigault2013, childress2014,
-scolnic2018a}. The importance of local SN environmental studies remains
-highly debated, however \cite[e.g.,][]{jones2015, jones2019}, and especially the effect of
-such an astrophysical bias has on the derivation of $H_0$ \citep{jones2015,
-riess2016, riess2018, rose2019}.
-
-The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.\LEt{to properly include these references into the main text, please separate them by commas and set off the last reference additionally with "and".
-This is a LaTeX command error that I cannot fix for you with the program I work with (citet and citep)}
-\cite{mannucci2005, scannapieco2005, sullivan2006, smith2012} have shown that
-the relative SNe~Ia rate in galaxies could be explained if two populations
-existed, a young population that follows the host star formation activity, and an old population that
-followed the host stellar mass (the so-called prompt-and-delayed or A+B
-model). \cite{rigault2020} used the LsSFR to determine the younger
-(those with a high LsSFR) and the older (those with a low LsSFR) populations.
-Furthermore, since the first SNe~Ia host analyses, the SN stretch has been known
-to be strongly correlated with the SN host galaxy properties \citep{hamuy1996,
-hamuy2000}. This correlation has been extensively confirmed since then
-\citep[e.g.,][]{neill2009, lampeitl2010, gupta2011, dandrea2011, pan2014}.
-Following the A+B model and the connection between SN stretch and host
-galaxy properties, \cite{howell2007} first discussed the potential redshift
-drift of the SN stretch distribution. In this paper we revisit this question
-using the most recent SNe~Ia datasets.
+scolnic2018a}, including to derive $H_0$ \citep{riess2016, riess2019}. The
+underlying connection between the SNe and their host galaxies remains unclear,
+however, particularly when global properties such as the host stellar mass are
+used. This in turn raises the question of the accuracy of such corrections
+because the global properties of the host galaxies evolve with redshift. More
+recently, studies have used the local SN host galaxy environment to probe more
+direct connections between the SNe and their host galaxy environments
+\citep{rigault2013}, showing that local age tracers such as the local specific
+star formation rate (LsSFR) or the local color are more strongly correlated with
+the standardized SN magnitude \citep{roman2018, kim18, rigault2020}. The
+identification of SN~Ia spectral features that are correlated with LsSFR
+\citep{nordin2018} further support this connection. These results suggest age as
+the driving parameter underlying the mass step. If this is true, it would have a
+significant effect for cosmology because the environmental correction that would
+need to be applied to SN standardization could strongly vary with redshift
+\citep{rigault2013, childress2014, scolnic2018a}. The importance of local SN
+environmental studies remains highly debated, however \cite[e.g.,][]{jones2015,
+jones2019}, and especially the effect of such an astrophysical bias has on the
+derivation of $H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.
+
+The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
+\cite{mannucci2005}, \cite{scannapieco2005}, \cite{sullivan2006} and
+\cite{smith2012} have shown that the relative SNe~Ia rate in galaxies could be
+explained if two populations existed, a young population that follows the host
+star formation activity, and an old population that followed the host stellar
+mass (the so-called prompt-and-delayed or A+B model). \cite{rigault2020} used
+the LsSFR to determine the younger (those with a high LsSFR) and the older
+(those with a low LsSFR) populations. Furthermore, since the first SNe~Ia host
+analyses, the SN stretch has been known to be strongly correlated with the SN
+host galaxy properties \citep{hamuy1996, hamuy2000}. This correlation has been
+extensively confirmed since then \citep[e.g.,][]{neill2009, lampeitl2010,
+gupta2011, dandrea2011, pan2014}. Following the A+B model and the connection
+between SN stretch and host galaxy properties, \cite{howell2007} first discussed
+the potential redshift drift of the SN stretch distribution. In this paper we
+revisit this question using the most recent SNe~Ia datasets.
 
@@
 
-We present the sample we used for this analysis in Section~\ref{sec:sample}. It was derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
-importance of obtaining a ``complete'' sample, that is, a sample that is\ representative of the true
-underlying SNe Ia distribution, and how we built one from the Pantheon sample.
-We then present our modeling of the distribution
-of stretch in Section~\ref{sec:modeling}  and our results in Section~\ref{sec:results}. In this
-section, we test whether the SN stretch distribution evolves as a function of
-redshift and determine whether the aforementioned age model agrees well with this
-evolution. We briefly discuss these results in the context of SN cosmology in
-Section~\ref{sec:discussion}, and we conclude in Section~\ref{sec:ccl}.
+We present the sample we used for this analysis in Section~\ref{sec:sample}. It
+was derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
+importance of obtaining a ``complete'' sample, that is, a sample that is\
+representative of the true underlying SNe Ia distribution, and how we built one
+from the Pantheon sample. We then present our modeling of the distribution of
+stretch in Section~\ref{sec:modeling} and our results in
+Section~\ref{sec:results}. In this section, we test whether the SN stretch
+distribution evolves as a function of redshift and determine whether the
+aforementioned age model agrees well with this evolution. We briefly discuss
+these results in the context of SN cosmology in Section~\ref{sec:discussion},
+and we conclude in Section~\ref{sec:ccl}.
 
@@
 stretch distribution function, and we would then study how it evolves with
-redshift. No such sample exists, therefore we must first construct subsamples from existing high-redshift SN~Ia samples that are
-as near to volume-limited as possible.
+redshift. No such sample exists, therefore we must first construct subsamples
+from existing high-redshift SN~Ia samples that are as near to volume-limited as
+possible.
 
@@
 distributions in a few redshift bins. In practice, however, differential
-observational selection effects will affect the observed SN
-stretch distributions. Because the observed SN~Ia magnitude correlates
-with the light-curve stretch (and color), the first SNe~Ia that a
-magnitude-limited survey will miss are those at the lowest stretch (which are the reddest).
-Consequently, if magnitude-related observational selection effects are
-not accounted for, true population drift might be confused with survey
-properties, and vice versa.
-
-Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring
-SN types and host galaxy redshifts, the observation selection
-effects of magnitude-limited surveys should be negligible below a given redshift
-at which even the faintest normal SNe~Ia can be observed.
-Aiding in the construction of nearly volume-limited subsamples is the
-    fact that the SN~Ia population trails off toward fainter SNe~Ia. A
-    complication is that complete spectroscopic follow-up has not always been
-    the norm, as discussed below. In contrast, targeted surveys have highly
-    complex observational selection functions and so are
-    discarded from our analysis. High-redshift SN cosmology samples,
-    such as the Pantheon sample, are predominately assembled from magnitude-limited surveys from
+observational selection effects will affect the observed SN stretch
+distributions. Because the observed SN~Ia magnitude correlates with the
+light-curve stretch (and color), the first SNe~Ia that a magnitude-limited
+survey will miss are those at the lowest stretch (which are the reddest).
+Consequently, if magnitude-related observational selection effects are not
+accounted for, true population drift might be confused with survey properties,
+and vice versa.
+
+Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring SN
+types and host galaxy redshifts, the observation selection effects of
+magnitude-limited surveys should be negligible below a given redshift at which
+even the faintest normal SNe~Ia can be observed. Aiding in the construction of
+nearly volume-limited subsamples is the fact that the SN~Ia population trails
+off toward fainter SNe~Ia. A complication is that complete spectroscopic
+follow-up has not always been the norm, as discussed below. In contrast,
+targeted surveys have highly complex observational selection functions and so
+are discarded from our analysis. High-redshift SN cosmology samples, such as the
+Pantheon sample, are predominately assembled from magnitude-limited surveys from
 which volume-limited SN~Ia subsamples can be constructed.
@@
 \citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
-c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
-parent distribution \citep{guy2007, betoule2014}; see also
-\citet{bazin2011} and \citet{campbell2013} for similar contours, the second
-using a more conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude
-with $x_1=0$ and $c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can
-derive the absolute standardized magnitude at maximum of light $M = M_0 - \alpha
-x_1 + \beta c$ within the aforementioned ellipse given the standardization
-coefficient $\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: The
-faintest SN~Ia is that with $(x_1=-1.65, c=0.25)$ and an absolute standardized
-magnitude at peak in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. This object typically ought to be detected 5 days before and a week after peak to
-build a suitable light curve, the effective limiting standardized absolute
-magnitude is approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude
-limit $m_{\lim}$ of a magnitude limited survey, we can derive the maximum
-redshift $z_{\lim}$ above which the faintest SNe~Ia will be missed using the
-relation between apparent magnitude, redshift, and absolute magnitude
-$\mu(z_{\lim}) = m_{\lim} - M_{\lim}$.
+c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full parent
+distribution \citep{guy2007, betoule2014}; see also \citet{bazin2011} and
+\citet{campbell2013} for similar contours, the second using a more conservative
+$|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is
+$M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
+standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
+within the aforementioned ellipse given the standardization coefficient
+$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: The faintest SN~Ia is
+that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
+in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. This object typically ought
+to be detected 5 days before and a week after peak to build a suitable
+light-curve, the effective limiting standardized absolute magnitude is
+approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit
+$m_{\lim}$ of a magnitude limited survey, we can derive the maximum redshift
+$z_{\lim}$ above which the faintest SNe~Ia will be missed using the relation
+between apparent magnitude, redshift, and absolute magnitude $\mu(z_{\lim}) =
+m_{\lim} - M_{\lim}$.
 
@@
 
-We therefore considered a set of cuts that defines a first fiducial
-    sample, taking the limits as initially suggested by the previous
-    completeness analysis. However, as this solution might be an overly
-    simplified way to create a complete sample, for example, because it ignores
-    spectroscopic follow-up in efficiency for redshifts below $z_{lim}$, we
-    also considered another set of cuts to define a so-called conservative
-    sample. This is smaller and therefore less statistically constraining,
-    but also even less prone to observational selection effects. If the redshift
-    drift is still significant in the conservative sample, it would be even more
-    meaningful in a carefully tailored selection-free sample. These samples are
-    adequate for the goal of this study, which is to develop a first
-    implementation of a model for drift in SN Ia properties. If fruitful, the
-    sample selection can later be refined as required with a more detailed model
-of the observational selection, for instance, using the SNANA package \citep{SNANA}.
+We therefore considered a set of cuts that defines a first fiducial sample,
+taking the limits as initially suggested by the previous completeness analysis.
+However, as this solution might be an overly simplified way to create a complete
+sample, for example, because it ignores spectroscopic follow-up in efficiency
+for redshifts below $z_{lim}$, we also considered another set of cuts to define
+a so-called conservative sample. This is smaller and therefore less
+statistically constraining, but also even less prone to observational selection
+effects. If the redshift drift is still significant in the conservative sample,
+it would be even more meaningful in a carefully tailored selection-free sample.
+These samples are adequate for the goal of this study, which is to develop a
+first implementation of a model for drift in SN Ia properties. If fruitful, the
+sample selection can later be refined as required with a more detailed model of
+the observational selection, for instance, using the SNANA package
+\citep{SNANA}.
 
@@
 24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
-final release website.}}. This converts into a $z_{\lim}=0.60$, in agreement with
-\cite{neill2006}, \cite{perrett2010}, and \cite{bazin2011}. Fig.~14 of
+final release website.}}. This converts into a $z_{\lim}=0.60$, in agreement
+with \cite{neill2006}, \cite{perrett2010}, and \cite{bazin2011}. Fig.~14 of
 \citet[][see their Section~5]{perrett2010}, however, suggests a lower limit of
-$z_{\lim}=0.55$. We therefore used $z=0.60$ and $z=0.55$ as redshift
-limits for the fiducial and conservative samples, respectively, for the SNLS.
+$z_{\lim}=0.55$. We therefore used $z=0.60$ and $z=0.55$ as redshift limits for
+the fiducial and conservative samples, respectively, for the SNLS.
 
@@
 agreement with Fig.~6 of \cite{scolnic2018a}, for example. If we were to be
-    conservative, this figure would also suggest of a more stringent
-$z_{\lim}=0.27$ cut; we therefore used $0.31$ and $0.27$ for our fiducial and
-conservative samples, respectively, for PS1.
+conservative, this figure would also suggest of a more stringent $z_{\lim}=0.27$
+cut; we therefore used $0.31$ and $0.27$ for our fiducial and conservative
+samples, respectively, for PS1.
 
@@
 \citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
-SDSS surveys had to contend with limited spectroscopic resources. As
-discussed in \citet[][Section~2]{kessler2009}, during the first year of SDSS,
-SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
-to a redshift cut at $0.15$. For the remaining SDSS survey, additional
-spectroscopic resources were available, and \cite{kessler2009} and
-\cite{dilday2008} showed a reasonable completeness up to $z_{\lim}=0.2$.
-Based on this, we used $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
-fiducial and conservative samples, respectively, for the SDSS.
+SDSS surveys had to contend with limited spectroscopic resources. As discussed
+in \citet[][Section~2]{kessler2009}, during the first year of SDSS, SNe~Ia with
+$r<20.5$~mag were favored for spectroscopic follow-up, corresponding to a
+redshift cut at $0.15$. For the remaining SDSS survey, additional spectroscopic
+resources were available, and \cite{kessler2009} and \cite{dilday2008} showed a
+reasonable completeness up to $z_{\lim}=0.2$. Based on this, we used
+$z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our fiducial and conservative samples,
+respectively, for the SDSS.
 
@@
 distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
-expected, the selected redshift limits are roughly located slightly
-    before the peak of these histograms. In Section \ref{ssec:verify} we
-confirm that these redshift limits are effective for constructing nearly
-volume-limited subsamples from samples that were initially more closely
-magnitude limited in their search or spectroscopic follow-up.
+expected, the selected redshift limits are roughly located slightly before the
+peak of these histograms. In Section \ref{ssec:verify} we confirm that these
+redshift limits are effective for constructing nearly volume-limited subsamples
+from samples that were initially more closely magnitude limited in their search
+or spectroscopic follow-up.
 
@@
         \citealt{scolnic2018a}). The colored parts represent the distribution of
-        SNe~Ia that we retained in our analysis because they are assumed to be free from
-    observational selection bias (see Section~\ref{sec:sample}). The
-darker (lighter) color responds to the conservative (fiducial)
-selection cut.}
+        SNe~Ia that we retained in our analysis because they are assumed to be
+    free from observational selection bias (see Section~\ref{sec:sample}). The
+darker (lighter) color responds to the conservative (fiducial) selection cut.}
     \label{fig:cuts}
@@
 and~4.2.2 ; see \citealt{aldering2020}). For this dataset, spectroscopic
-screening was performed for candidates with $r \lesssim 19.5$; redshift cuts were
-then applied when selecting which SN Ia to follow, resulting in a redshift range
-of $0.02 < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} to extract local host properties. These 114 SNfactory SNe~Ia are thus in the
+screening was performed for candidates with $r \lesssim 19.5$; redshift cuts
+were then applied when selecting which SN Ia to follow, resulting in a redshift
+range of $0.02 < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} to
+extract local host properties. These 114 SNfactory SNe~Ia are thus in the
 volume-limited part of the survey (Aldering et al., in prep.), and are therefore
 assumed to be a random sampling of the underlying SN population. The SNfactory
-sample is particularly useful for studying SN property drift because it enables us
-to have a large complete SN~Ia sample at $z<0.1$. Finally, we include the HST
+sample is particularly useful for studying SN property drift because it enables
+us to have a large complete SN~Ia sample at $z<0.1$. Finally, we include the HST
 sample from Pantheon \citep{strolger04}. These high-redshift SNe are of great
@@
 evolution of the mean stretch is shown, with the data split in redshift bins of
-regular sample size. SNe~Ia at higher redshift have a
-larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) on average than those at lower redshift
-($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch
-distribution evolves with redshift.
+regular sample size. SNe~Ia at higher redshift have a larger stretch ($0.34 \pm
+0.10$ at $z\sim0.65$) on average than those at lower redshift ($-0.17\pm 0.10$
+at $z\sim0.05$), suggesting that the underlying stretch distribution evolves
+with redshift.
 
@@
         function of redshift for each survey considered in this analysis (see
-        legend). Solid (open) markers correspond to the conservative
-        (fiducial) redshift cuts. \textit{Top:} Stacked redshift
-    histograms in dark (light) colors for the conservative (fiducial) redshift cuts.}
+        legend). Solid (open) markers correspond to the conservative (fiducial)
+        redshift cuts. \textit{Top:} Stacked redshift histograms in dark (light)
+    colors for the conservative (fiducial) redshift cuts.}
     \label{fig:sample}
@@
 
-\subsection{Testing the construction of a volume-limited sample}\label{ssec:verify}
+\subsection{Testing the construction of a volume-limited
+sample}\label{ssec:verify}
 
@@
 magnitude-limited ones using simple redshift cuts. This simplified approach is
-statistically suboptimal, but should suffice to test our key question
-whether redshift evolution of stretch is compatible with the model of \cite{rigault2020}. However, the possibility remains that a complex observational selection
-function related to spectroscopic follow-up efficiencies below our fiducial (or
-even conservative) redshift cuts might still affect our sample, making it not
-fully volume limited; this would in turn bias our conclusion about the
-astrophysical drift of the SNe~Ia population. We now examine this possibility.
+statistically suboptimal, but should suffice to test our key question whether
+redshift evolution of stretch is compatible with the model of
+\cite{rigault2020}. However, the possibility remains that a complex
+observational selection function related to spectroscopic follow-up efficiencies
+below our fiducial (or even conservative) redshift cuts might still affect our
+sample, making it not fully volume limited; this would in turn bias our
+conclusion about the astrophysical drift of the SNe~Ia population. We now
+examine this possibility.
 
@@
 
-The two samples that overlap most in redshift are PS1 and SDSS in the
-redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping
-subsample consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus
-is most likely to be affected by residual observational selection effects (see
-the corresponding discussion in section~\ref{ssec:cuts}). Over that same
-redshift range, PS1 has 52 SNe~Ia that are in the lowest redshift bins and
-thus unlikely to have any observational selection issue. To identify potential
-inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib} (upper
-panels) compares the stretch and color distribution of these two surveys. The
-resulting Kolmogorov-Smirnov (KS) similarity test $p$-values ($p >10\%$) do not
-support any inconsistency, in agreement with the visual impression from
+The two samples that overlap most in redshift are PS1 and SDSS in the redshift
+range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping subsample
+consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus is most
+likely to be affected by residual observational selection effects (see the
+corresponding discussion in section~\ref{ssec:cuts}). Over that same redshift
+range, PS1 has 52 SNe~Ia that are in the lowest redshift bins and thus unlikely
+to have any observational selection issue. To identify potential inconsistency
+between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib} (upper panels)
+compares the stretch and color distribution of these two surveys. The resulting
+Kolmogorov-Smirnov (KS) similarity test $p$-values ($p >10\%$) do not support
+any inconsistency, in agreement with the visual impression from
 Fig.~\ref{fig:distrib}.
 
-We performed a similar analysis for PS1 and SNLS over the redshift range $0.20 < z
-< 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same conclusion can be
-drawn: There is no substantial sign of discrepancy in the stretch and color
-distributions between the low and high end of our fiducial SNLS and PS1
-samples, respectively. Nonetheless, the small size of the SNLS dataset at $z <
-0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of this test, and only a
-strong deviation would be noticeable. Extending the redshift range to $0.20 < z
-< 0.40$ (although we have no PS1 data above 0.3) allows increasing the SNLS
-subsample to 31, but the stretch $p$-value remains high (34\%), showing no sign
-of inconsistency.
+We performed a similar analysis for PS1 and SNLS over the redshift range $0.20 <
+z < 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same conclusion can
+be drawn: There is no substantial sign of discrepancy in the stretch and color
+distributions between the low and high end of our fiducial SNLS and PS1 samples,
+respectively. Nonetheless, the small size of the SNLS dataset at $z < 0.31$ (12
+SNe~Ia vs. 90 for PS1) limits the sensitivity of this test, and only a strong
+deviation would be noticeable. Extending the redshift range to $0.20 < z < 0.40$
+(although we have no PS1 data above 0.3) allows increasing the SNLS subsample to
+31, but the stretch $p$-value remains high (34\%), showing no sign of
+inconsistency.
 
@@
 
-We finally highlight that the SNe~Ia color is more prone to
-    observational selection effects than stretch, as illustrated in
-    Fig.~\ref{fig:maglim}; see also Fig.~3 of \cite{kessler2017}, for instance. Hence,
-    because the comparison of color distributions shows no significant indication of
-    leftover observational selection effect, this further supports our claim that
-    our simple redshift-based selection criteria are sufficient to build the
-    complete SNe~Ia samples required to test the redshift evolution of the
-stretch distribution.
+We finally highlight that the SNe~Ia color is more prone to observational
+selection effects than stretch, as illustrated in Fig.~\ref{fig:maglim}; see
+also Fig.~3 of \cite{kessler2017}, for instance. Hence, because the comparison
+of color distributions shows no significant indication of leftover observational
+selection effect, this further supports our claim that our simple redshift-based
+selection criteria are sufficient to build the complete SNe~Ia samples required
+to test the redshift evolution of the stretch distribution.
 
@@
 rates and delay-time distributions \citep[e.g.,][]{mannucci2005,
-scannapieco2005, sullivan2006, smith2012, childress2014, maozmannucci2014}.
-In short, it was assumed that the number of young SNe~Ia follows the star
-formation rate (SFR) in the Universe, while the number of old SNe~Ia follows
-the number of billion-old stars in the Universe, that is, the stellar mass (M$^*$).
-Hence, if we denote $\delta(z)$ ( $\psi(z) = 1-\delta(z)$) the fraction of
-young (old) SNe~Ia in the Universe as a function of redshift, then the
-ratio $\delta/\psi$ is expected to follow the evolution of the specific star
-formation rate (SFR/M$^*$), which goes as $(1+z)^{2.8}$ until $z\sim2$
+scannapieco2005, sullivan2006, smith2012, childress2014, maozmannucci2014}. In
+short, it was assumed that the number of young SNe~Ia follows the star formation
+rate (SFR) in the Universe, while the number of old SNe~Ia follows the number of
+billion-old stars in the Universe, that is, the stellar mass (M$^*$). Hence, if
+we denote $\delta(z)$ ($\psi(z) = 1-\delta(z)$) the fraction of young (old)
+SNe~Ia in the Universe as a function of redshift, then the ratio $\delta/\psi$
+is expected to follow the evolution of the specific star formation rate
+(SFR/M$^*$), which goes as $(1+z)^{2.8}$ until $z\sim2$
 \citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
@@
 expectations \citep{mannucci2006, rodney2014}, \cite{rigault2020} concluded that
-
 \begin{equation}
@@
 To model the evolution of the full SN stretch distribution as a function of
-redshift, we need to model the SN stretch
-distribution for each age subsample given our aforementioned model of the evolution of the fraction of
-younger and older SNe~Ia with cosmic time. \cite{rigault2020} presented the relation between SN stretch and LsSFR
-measurement, a progenitor age tracer, using the SNfactory sample. This relation
-is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory SNe used in the
-current analysis. Given the structure of the stretch-LsSFR scatter plot, our
-model of the underlying SN~Ia stretch distribution is defined as follows\LEt{please reformat so that this becomes proper text, there is no need for a two-item bulleted list}:
+redshift, we need to model the SN stretch distribution for each age subsample
+given our aforementioned model of the evolution of the fraction of younger and
+older SNe~Ia with cosmic time. \cite{rigault2020} presented the relation between
+SN stretch and LsSFR measurement, a progenitor age tracer, using the SNfactory
+sample. This relation is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory
+SNe used in the current analysis. Given the structure of the stretch-LsSFR
+scatter plot, our model of the underlying SN~Ia stretch distribution is defined
+as follows: \textbf{the stretch distribution of the younger population
+($\log(\mathrm{LsSFR})\geq-10.82$) is modeled as a single normal distribution
+$\mathcal{N}(\mu_1, \sigma_1{}^2)$, and the stretch distribution of the older
+population ($\log(\mathrm{LsSFR})<-10.82$) is modeled as a bimodal Gaussian
+mixture $a\times \mathcal{N}(\mu_1, \sigma_1{}^2) + (1-a)\times
+\mathcal{N}(\mu_2, \sigma_2{}^2)$, where one mode is the same as for the young
+population, $a$ representing the relative effect of the two modes.}
 
@@
 
-\begin{itemize}
-    \item for the younger population (i.e., $\log(\mathrm{LsSFR})\geq-10.82$),
-        the stretch distribution is modeled as a single normal distribution
-        $\mathcal{N}(\mu_1, \sigma_1{}^2)$; 
-    \item the older population (i.e., $\log( \mathrm{LsSFR})<-10.82$) stretch
-        distribution is modeled as a bimodal Gaussian mixture $a\times
-        \mathcal{N}(\mu_1, \sigma_1{}^2) + (1-a)\times \mathcal{N}(\mu_2,
-        \sigma_2{}^2)$, where one mode is the same as for the young population,
-        $a$ representing the relative effect of the two modes.
-\end{itemize}
-
 The stretch probability distribution function (pdf) of a given SN will be the
@@
 by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
-In general, however, the fraction of young SNe~Ia as a function of redshift is given
-by $\delta(z)$ (see Eq.~\ref{eq:delta}), and therefore our redshift drift model
-of the underlying stretch distribution of SNe~Ia as a function of redshift
+In general, however, the fraction of young SNe~Ia as a function of redshift is
+given by $\delta(z)$ (see Eq.~\ref{eq:delta}), and therefore our redshift drift
+model of the underlying stretch distribution of SNe~Ia as a function of redshift
 $X_1(z)$ is given by
@@
 
-Given the probability $y^i$ that a given SN is young and assuming our base
-model (see Section~\ref{sec:basemodel}), the probability of measuring a
-\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by
+Given the probability $y^i$ that a given SN is young and assuming our base model
+(see Section~\ref{sec:basemodel}), the probability of measuring a
+\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error $\mathrm{d}x_1^i$ is
+given by
 \begin{align}\label{eq:likelihoodsnf}
@@
 The maximum-likelihood estimate of the five~free parameters
-$\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained
-by minimizing the following:
+$\vec{\theta}\equiv({\mu_1, \mu_2, \sigma_1, \sigma_2,a})$ of the model is
+obtained by minimizing the following:
 \begin{equation}\label{eq:likelihood}
@@
 
-Depending on whether $y^i$ can be estimated directly from LsSFR measurements, there are two ways to proceed. We discuss them below.
+Depending on whether $y^i$ can be estimated directly from LsSFR measurements,
+there are two ways to proceed. We discuss them below.
 
@@
     model when applied to the SNfactory dataset only (114 SNe~Ia), the fiducial
-569 SN~Ia sample, or the conservative model (422).}
+569 SN~Ia sample, or the conservative \textbf{sample} (422).}
     \label{tab:modelresults}
@@
 distribution $X_1$ (Eq.~\ref{eq:stretchz}), but this time, assuming $y^i =
-\delta(z^i)$ for any given SN~$i$. 
-\\
-\\
-For the remaining analysis, we therefore minimized Eq.~\ref{eq:likelihood}
-using $p_y^i$ , the probability for the SN $i$ to be young, when available
-(i.e., for SNfactory dataset), and $\delta(z^i),$ the expected fraction of
-young SNe~Ia at the SN redshift $z^i$  , otherwise.
+\delta(z^i)$ for any given SN~$i$.\bigbreak
+
+\noindent For the remaining analysis, we therefore minimized
+Eq.~\ref{eq:likelihood} using $p_y^i$, the probability for the SN $i$ to be
+young, when available (i.e., for SNfactory dataset), and $\delta(z^i)$, the
+expected fraction of young SNe~Ia at the SN redshift $z^i$, otherwise.
 
 Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
-(conservative) sample are given Table~\ref{tab:modelresults}, and the
-predicted redshift evolution of mean stretch (expected $x_1$ given the
-distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
-Fig.~\ref{fig:modelall} accounting for parameter errors and their covariances.
-This figure shows that the measured mean SN~Ia stretch per redshift bin of
-equal sample size closely follows our redshift drift modeling. This is indeed
-what is expected if old environments favor low SN stretches
-\citep[e.g.,][]{howell2007} \textit{\textup{and}} if the fraction of old SNe~Ia declines as
-a function of redshift. See Section~\ref{sec:results} for a more quantitative
-discussion.
+(conservative) sample are given Table~\ref{tab:modelresults}, and the predicted
+redshift evolution of mean stretch (expected $x_1$ given the distribution of
+Eq.~\ref{eq:stretchz}) illustrated as a blue band in Fig.~\ref{fig:modelall}
+accounting for parameter errors and their covariances. This figure shows that
+the measured mean SN~Ia stretch per redshift bin of equal sample size closely
+follows our redshift drift modeling. This is indeed what is expected if old
+environments favor low SN stretches \citep[e.g.,][]{howell2007} and if the
+fraction of old SNe~Ia declines as a function of redshift. See
+Section~\ref{sec:results} for a more quantitative discussion.
 
@@
     \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
-        as a function of redshift. Markers show the stretch plain mean
-        (the error is estimated from the scatter) measured in redshift bins of equal
+        as a function of redshift. Markers show the stretch plain mean (the
+        error is estimated from the scatter) measured in redshift bins of equal
         sample size, indicated in light gray at the bottom of each redshift bin.
-        Full and light markers are used when the fiducial or conservative samples are considered, respectively. The horizontal orange line
-        represents the mean stretch for the nonevolving Gaussian model
-        (last line of Table~\ref{tab:comp}) on the fiducial sample. Best
-        fits of our base drifting model are shown as blue, dashed blue, and
-        gray when fitted on the fiducial sample, the conservative sample, or
-        the SNfactory dataset only, respectively; all are compatible. The
-        light blue band illustrates the amplitude of the error (including\
-        covariance) of the best-fit model when considering the fiducial
-    dataset.}
+        Full and light markers are used when the fiducial or conservative
+        samples are considered, respectively. The horizontal orange line
+        represents the mean stretch for the nonevolving Gaussian model (last
+        line of Table~\ref{tab:comp}) on the fiducial sample. Best fits of our
+        base drifting model are shown as blue, dashed blue, and gray when fitted
+        on the fiducial sample, the conservative sample, or the SNfactory
+        dataset only, respectively; all are compatible. The light blue band
+        illustrates the amplitude of the error (including\ covariance) of the
+        best-fit model when considering the fiducial dataset.}
     \label{fig:modelall}
@@
 In Section~\ref{sec:basemodel} we have modeled the underlying stretch
-distribution following \cite{rigault2020}, that is,\ as a single Gaussian for the
-young SNe~Ia and a mixture of two Gaussians for the old SNe Ia
-population, one being the same as for the young population, plus another one for
-the fast-declining SNe~Ia that appear to only exist in old local environments.
-This is our so-called base model. However, to test different modeling
-choices, we implemented a suite of alternative parameterizations that we
-also adjusted to the data following the procedure described in
-Section~\ref{sec:modelnopy}.
+distribution following \cite{rigault2020}, that is,\ as a single Gaussian for
+the young SNe~Ia and a mixture of two Gaussians for the old SNe Ia population,
+one being the same as for the young population, plus another one for the
+fast-declining SNe~Ia that appear to only exist in old local environments. This
+is our so-called base model. However, to test different modeling choices, we
+implemented a suite of alternative parameterizations that we also adjusted to
+the data following the procedure described in Section~\ref{sec:modelnopy}.
 
@@
 single normal distribution for each of the young and old populations. We thus
-considered a Howell+drift model, with one single Gaussian per age group and
-the $\delta(z)$ drift from Eq.~\ref{eq:delta}.
+considered a Howell+drift model, with one single Gaussian per age group and the
+$\delta(z)$ drift from Eq.~\ref{eq:delta}.
 
 Alternatively, because we aim to probe the existence of an evolution with
-redshift, we also tested constant models by restricting the base and
-Howell models to use an assumed redshift-independent fraction $\delta(z)
-\equiv f$ of young SNe; these models are hereafter labeled base+constant and
-Howell+constant.
+redshift, we also tested constant models by restricting the base and Howell
+models to use an assumed redshift-independent fraction $\delta(z) \equiv f$ of
+young SNe; these models are hereafter labeled base+constant and Howell+constant.
 
@@
 some potential redshift evolution information \citep{scolnic2016, scolnic2018a}.
-See a more detailed discussion of BBC in Section~\ref{sec:discussion}. Finally, for the sake of completeness, we also considered redshift-independent
-pure and asymmetric Gaussian models.
+See a more detailed discussion of BBC in Section~\ref{sec:discussion}. Finally,
+for the sake of completeness, we also considered redshift-independent pure and
+asymmetric Gaussian models.
 
@@
 We adjusted each of the models described above on both the fiducial and
-conservative samples (cf. Section~\ref{sec:sample}). The results are gathered
-in Table~\ref{tab:comp} and are illustrated in Fig.~\ref{fig:mod_comp}.
+conservative samples (cf. Section~\ref{sec:sample}). The results are gathered in
+Table~\ref{tab:comp} and are illustrated in Fig.~\ref{fig:mod_comp}.
 
@@
     \caption{Comparison of the relative ability of each model to describe the
-        data. For each considered model, we report whether the model is drifting, its number of free parameters, and for both the fiducial and the
-        conservative cuts, $-2\ln(L)$ (see Eq.~\ref{eq:likelihood}), the AIC and
-        the AIC difference ($\Delta$AIC) between this model and the base model
-        used as reference because it has the lowest AIC.}
+        data. For each considered model, we report whether the model is
+        drifting, its number of free parameters, and for both the fiducial and
+        the conservative cuts, $-2\ln(L)$ (see Eq.~\ref{eq:likelihood}), the AIC
+        and the AIC difference ($\Delta$AIC) between this model and the base
+    model used as reference because it has the lowest AIC.}
     \label{tab:comp}
@@
     \includegraphics[width=\linewidth]{Article_figures/mod_comp.pdf}
-    \caption{$\Delta$AIC between base model (reference) and other models
-        (see Table~\ref{tab:comp}). Full and open blue markers correspond to
-        models with and without redshift drift, respectively. Light markers show
-        the results when the analysis is performed on the conservative sample
-        rather than the fiducial one. Color bands illustrate the validity of
-        the models, from acceptable ($\Delta\mathrm{AIC} > -5$) to excluded
+    \caption{$\Delta$AIC between base model (reference) and other models (see
+        Table~\ref{tab:comp}). Full and open blue markers correspond to models
+        with and without redshift drift, respectively. Light markers show the
+        results when the analysis is performed on the conservative sample rather
+        than the fiducial one. Color bands illustrate the validity of the
+        models, from acceptable ($\Delta\mathrm{AIC} > -5$) to excluded
         ($\Delta\mathrm{AIC} < -20$), see text. According to the AIC, all
-        nondrifting models (open symbols) are excluded as a poorer representation of the data than the base (drifting) model.}
+    nondrifting models (open symbols) are excluded as a poorer representation of
+the data than the base (drifting) model.}
     \label{fig:mod_comp}
@@
 information criterion \citep[AIC, e.g.,][]{burnham2004} to compare their ability
-to properly describe the observations. This estimator penalizes additional degrees of
-freedom to avoid overfitting the data and is defined as follows:
+to properly describe the observations. This estimator penalizes additional
+degrees of freedom to avoid overfitting the data and is defined as follows:
 \begin{equation}
-    \mathrm{AIC} = -2\ln(L) + 2k
-,\end{equation}
+    \mathrm{AIC} = -2\ln(L) + 2k,
+\end{equation}
 where $-2\ln(L)$ is derived by minimizing Eq.~\eqref{eq:likelihood}, and $k$ is
-the number of free parameters to be adjusted. The reference model has the smallest AIC; in comparison to this model, the models with
+the number of free parameters to be adjusted. The reference model has the
+smallest AIC; in comparison to this model, the models with
 $\Delta\mathrm{AIC}<5$ are coined acceptable, those with
-$5<\Delta\mathrm{AIC}<20$ are not favored, and those with $\Delta\mathrm{AIC}>20$
-are deemed excluded. This roughly corresponds to 2, 3, and 5~$\sigma$ limits for
-a Gaussian probability distribution.
-
-The best model (with the smallest AIC) is the so-called base model and thus is our
-reference model; this is true for the fiducial and conservative samples.
-The base model also has the smallest $-2\ln(L)$, making it the most likely
-model even when the overfitting issue that is accounted for by the AIC formalism is ignored.
+$5<\Delta\mathrm{AIC}<20$ are not favored, and those with
+$\Delta\mathrm{AIC}>20$ are deemed excluded. This roughly corresponds to 2, 3,
+and 5~$\sigma$ limits for a Gaussian probability distribution.
+
+The best model (with the smallest AIC) is the so-called base model and thus is
+our reference model; this is true for the fiducial and conservative samples.
+The base model also has the smallest $-2\ln(L)$, making it the most likely model
+even when the overfitting issue that is accounted for by the AIC formalism is
+ignored.
 
 Furthermore, we find that redshift-independent stretch distributions are all
-excluded as suitable descriptions of the data relative to the base model. The best nondrifting model (the asymmetric model) has a very marginal
-chance ($p \equiv \exp\left(\Delta\mathrm{AIC}/2\right) = 5\times10^{-6}$) to
-describe the data as well as the base model. This result is just a quantitative
-assessment of qualitative facts that are clearly visible in Fig.~\ref{fig:modelall}: The
-mean SN stretch per bin of redshift strongly suggests a significant redshift
+excluded as suitable descriptions of the data relative to the base model. The
+best nondrifting model (the asymmetric model) has a very marginal chance ($p
+\equiv \exp\left(\Delta\mathrm{AIC}/2\right) = 5\times10^{-6}$) to describe the
+data as well as the base model. This result is just a quantitative assessment of
+qualitative facts that are clearly visible in Fig.~\ref{fig:modelall}: The mean
+SN stretch per bin of redshift strongly suggests a significant redshift
 evolution rather than a constant value, and this evolution is well described by
@@
 implementations of the BBC technique \citep{scolnic2016, kessler2017} has one of
-the highest AIC values in our analysis (see Section~\ref{sec:results}). While its
-$-2\ln(L)$ is the smallest of all redshift-independent models (but still $-11.5$
-lower than the reference base model), it is strongly penalized for requiring
-15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the five~samples of the
-analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted as a
-probability $p=2\times 10^{-7}$ of being as good a representation of the data
-as the base model.
-
-We note that when models are compared that were adjusted on individual subsamples rather
-than globally, the Bayesian information criterion ($\mathrm{BIC} = -2\ln(L) +
-k\ln(n)$, with $n$ the number of data points) might be better suited than AIC because it explicitly accounts for the fact that each subsample is fitted
-separately: the sample-based model BIC is rightfully the sum of the BIC for each
-sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the sample-based
-asymmetric Gaussian model as being as pertinent as the base model.
+the highest AIC values in our analysis (see Section~\ref{sec:results}). While
+its $-2\ln(L)$ is the smallest of all redshift-independent models (but still
+$-11.5$ lower than the reference base model), it is strongly penalized for
+requiring 15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the five~samples
+of the analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be
+interpreted as a probability $p=2\times 10^{-7}$ of being as good a
+representation of the data as the base model.
+
+We note that when models that were adjusted on individual subsamples rather than
+globally \textbf{are compared}, the Bayesian information criterion
+($\mathrm{BIC} = -2\ln(L) + k\ln(n)$, with $n$ the number of data points) might
+be better suited than AIC because it explicitly accounts for the fact that each
+subsample is fitted separately: the sample-based model BIC is rightfully the sum
+of the BIC for each sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the
+sample-based asymmetric Gaussian model as being as pertinent as the base model.
 
@@
 HST subsample, we recomputed $\Delta$AIC for each model excluding this dataset;
-this did not change $\Delta$AIC by more than few tenths. The
-consistency of these values with those in Table~\ref{tab:comp} shows that the HST
-subsample does not drive our conclusions.
-
-We report in Table~\ref{tab:bbc} our determination of $\mu_0$ and
-$\sigma_{\pm}$ for each sample when an asymmetric Gaussian
-model was implmented, and adjusted on the nominally selection-free samples using our fiducial cuts
-(see Section~\ref{sec:sample}). Our results are in close agreement with those of
-\cite{scolnic2016} for the SNLS and SDSS and with the results reported by \cite{scolnic2018a} for PS1, who
-derived these model parameters using the full BBC formalism, using numerous
-simulations to model the observational selection effects (see details,
-e.g., Section~3 of \citealt{kessler2017}). The agreement between our fit of the
-asymmetric Gaussians on the supposedly selection-free part of the samples and
-the results derived using the BBC formalism supports our approach to
-constructing a sample with negligible observational selection
-effects. If we were to use the \cite{scolnic2016} and \cite{scolnic2018a} best-fit
-values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for the SNLS, SDSS and
-PS1, respectively, the $\Delta$AIC between our base drifting model and the BBC
-modeling would go even deeper from $-32$ to $-47$. We further discuss the
-consequence of this result for cosmology in Section~\ref{sec:discussion}.
+this did not change $\Delta$AIC by more than few tenths. The consistency of
+these values with those in Table~\ref{tab:comp} shows that the HST subsample
+does not drive our conclusions.
+
+We report in Table~\ref{tab:bbc} our determination of $\mu_0$ and $\sigma_{\pm}$
+for each sample when an asymmetric Gaussian model was implemented, and adjusted
+on the nominally selection-free samples using our fiducial cuts (see
+Section~\ref{sec:sample}). Our results are in close agreement with those of
+\cite{scolnic2016} for the SNLS and SDSS and with the results reported by
+\cite{scolnic2018a} for PS1, who derived these model parameters using the full
+BBC formalism, using numerous simulations to model the observational selection
+effects (see details, e.g., Section~3 of \citealt{kessler2017}). The agreement
+between our fit of the asymmetric Gaussians on the supposedly selection-free
+part of the samples and the results derived using the BBC formalism supports our
+approach to constructing a sample with negligible observational selection
+effects. If we were to use the \cite{scolnic2016} and \cite{scolnic2018a}
+best-fit values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for the SNLS,
+SDSS and PS1, respectively, the $\Delta$AIC between our base drifting model and
+the BBC modeling would go even deeper from $-32$ to $-47$. We further discuss
+the consequence of this result for cosmology in Section~\ref{sec:discussion}.
 
@@
 corresponding fit is not significantly better, with a $\Delta$AIC of $-0.4$.
-This reinforces our assumption that the young and old populations
-indeed appear to share the same underlying high-stretch mode. Furthermore, we
-might wonder whether a low-stretch mode might also exist in the young population
-(see Fig.~\ref{fig:stretchlssfr}). We tested for this by allowing this
-population to also be bimodal, finding the amplitude of such a
-low-stretch mode to be compatible with~0 ($<2\%$) in this young population. More
-generally, this raises the question of how well a given environmental
-tracer (here LsSFR) traces the age. An analysis dedicated to this question will
-be presented in Briday et al.\ in prep.
-
-Finally, ignoring the LsSFR measurements, which are only available for the SNfactory
-dataset (see Section~\ref{sec:modeling}), reduces the significance of the
-results presented in this section, as expected. Even so, nondrifting
-models remain strongly disfavored. For instance, the best-fitting sample-based Gaussian asymmetric model is still
-$\Delta\mathrm{AIC}<-10,$ which is less representative of the data than our base drifting
-model.
+This reinforces our assumption that the young and old populations indeed appear
+to share the same underlying high-stretch mode. Furthermore, we might wonder
+whether a low-stretch mode might also exist in the young population (see
+Fig.~\ref{fig:stretchlssfr}). We tested for this by allowing this population to
+also be bimodal, finding the amplitude of such a low-stretch mode to be
+compatible with~0 ($<2\%$) in this young population. More generally, this raises
+the question of how well a given environmental tracer (here LsSFR) traces the
+age. An analysis dedicated to this question will be presented in Briday et al.\
+in prep.
+
+Finally, ignoring the LsSFR measurements, which are only available for the
+SNfactory dataset (see Section~\ref{sec:modeling}), reduces the significance of
+the results presented in this section, as expected. Even so, nondrifting models
+remain strongly disfavored. For instance, the best-fitting sample-based Gaussian
+asymmetric model is still $\Delta\mathrm{AIC}<-10$, which is less representative
+of the data than our base drifting model.
 
@@
 \citep{hinton2019} can easily allow it (see, e.g., sections~1.3 and 2.5 of
-\cite{rubin2015}). Not doing so is a second-order issue for SN cosmology because it
-only affects the way in which the Malmquist bias is accounted for. As long as the
-Phillips relation \citep{phillips1993} standardization parameter $\alpha$ is not
-redshift dependent (a study that is beyond the scope of this paper, but see,
+\cite{rubin2015}). Not doing so is a second-order issue for SN cosmology because
+it only affects the way in which the Malmquist bias is accounted for. As long as
+the Phillips relation \citep{phillips1993} standardization parameter $\alpha$ is
+not redshift dependent (a study that is beyond the scope of this paper, but see,
 e.g., \citealt{scolnic2018a}), the stretch-corrected SNe~Ia magnitudes used for
-cosmology are indeed blind to the underlying stretch distribution for complete samples.
-However, surveys usually do have significant Malmquist bias for the upper half
-of their SN redshift distribution. As a consequence, mismodeling of the
-underlying stretch distribution will bias the SN magnitudes derived from such
-surveys.
+cosmology are indeed blind to the underlying stretch distribution for complete
+samples. However, surveys usually do have significant Malmquist bias for the
+upper half of their SN redshift distribution. As a consequence, mismodeling of
+the underlying stretch distribution will bias the SN magnitudes derived from
+such surveys.
 
 Commonly used Malmquist bias correction techniques, such as the BBC-formalism,
-assume per-sample asymmetric Gaussian functions to model the underlying
-stretch and color distributions. As shown in Section~\ref{sec:results},
-however, such a sample-based distribution is excluded in comparison to our
-drifting model. In contrast to what \citet[][Section~2]{scolnic2016}
-and \citet[][Section~5.4]{scolnic2018a} have suggested, that is, that
-traditional surveys span sufficiently limited redshift ranges
-such that the per-sample approach accounts for implicit redshift drifts,
-a direct modeling of the redshift drift is therefore more appropriate than a
-sample-based approach. We add here that as measurements of modern
-surveys try to cover increasingly larger redshift ranges in order to reduce
-calibration systematic uncertainties, this sample-based approach
-becomes less valid, notably for PS1, DES and, soon, the LSST.
+assume per-sample asymmetric Gaussian functions to model the underlying stretch
+and color distributions. As shown in Section~\ref{sec:results}, however, such a
+sample-based distribution is excluded in comparison to our drifting model. In
+contrast to what \citet[][Section~2]{scolnic2016} and
+\citet[][Section~5.4]{scolnic2018a} have suggested, that is, that traditional
+surveys span sufficiently limited redshift ranges such that the per-sample
+approach accounts for implicit redshift drifts, a direct modeling of the
+redshift drift is therefore more appropriate than a sample-based approach. We
+add here that as measurements of modern surveys try to cover increasingly larger
+redshift ranges in order to reduce calibration systematic uncertainties, this
+sample-based approach becomes less valid, notably for PS1, DES and, soon, the
+Large Synoptic Survey Telescope.
 
@@
 lower redshift), the higher the amplitude of the old-specific low-stretch mode.
-This redshift dependence on the underlying stretch distributions is
-shown with colors from blue to red in Fig.~\ref{fig:bbc_pdf_ps1} for the
-redshift range covered by PS1. The observed $x_1$ histogram follows the
-model we defined using the sum of individual underlying SN-redshift
-distributions. As expected, the two modeling approaches differ mostly in the
-negative part of the SN stretch distribution. The asymmetric Gaussian
-distribution goes through the middle of the bimodal distribution,
-overestimating the number of SNe~Ia at $x_1\sim-0.7$ and underestimating it at
-$x_1\sim-1.7$ in comparison to our base drifting model for typical PS1 SN
-redshifts. This means that the SN bias-corrected standardized magnitude
-estimated at a redshift plagued by observational
-selection effects would be biased by a mismodeling of the true
-underlying stretch distribution.
+This redshift dependence on the underlying stretch distributions is shown with
+colors from blue to red in Fig.~\ref{fig:bbc_pdf_ps1} for the redshift range
+covered by PS1. The observed $x_1$ histogram follows the model we defined using
+the sum of individual underlying SN-redshift distributions. As expected, the two
+modeling approaches differ mostly in the negative part of the SN stretch
+distribution. The asymmetric Gaussian distribution goes through the middle of
+the bimodal distribution, overestimating the number of SNe~Ia at $x_1\sim-0.7$
+and underestimating it at $x_1\sim-1.7$ in comparison to our base drifting model
+for typical PS1 SN redshifts. This means that the SN bias-corrected standardized
+magnitude estimated at a redshift plagued by observational selection effects
+would be biased by a mismodeling of the true underlying stretch distribution.
 
@@
 
-Assessing the amplitude of this magnitude bias for cosmology is beyond
-the scope of this paper given the complexity of the BBC analysis. It would
-require a full study using our base model (Eq.~\ref{eq:stretchz}) in place of
-the sample-based asymmetric modeling as part of the BBC simulations. However, we
-already highlighted that even if a nondrifting sample-based model could provide
+Assessing the amplitude of this magnitude bias for cosmology is beyond the scope
+of this paper given the complexity of the BBC analysis. It would require a full
+study using our base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
+asymmetric modeling as part of the BBC simulations. However, we already
+highlighted that even if a nondrifting sample-based model could provide
 comparable result in the volume-limited part of the various samples, these
@@
 
-In the era of modern cosmology, where we aim to measure $w_0$ at a
-subpercent level and $w_a$ with 10\%\ precision
-\citep[e.g.,][]{lsstpaper}, we stress that correct modeling of potential SN
-redshift drift should be further studied and care should be taken when
-samples are used that are affected by observational selection effects.
+In the era of modern cosmology, where we aim to measure $w_0$ at a subpercent
+level and $w_a$ with 10\%\ precision \citep[e.g.,][]{lsstpaper}, we stress that
+correct modeling of potential SN redshift drift should be further studied and
+care should be taken when samples that are affected by observational selection
+effects are used.
 
@@
 
-We have presented an initial study of the drift of the underlying
-SNe~Ia stretch distribution as a function of redshift. We built
-effectively volume-limited SN~Ia subsamples from the Pantheon dataset
-\citep[][SDSS, PS1, and SNLS]{scolnic2018a}, to which we added HST and SNfactory
-data from \cite{rigault2020} for the high- and low-redshift bins. We only
-considered the SNe that have been discovered in the redshift range of each
-survey in which observational selection effects are negligible, so that
-the observed SNe~Ia stretches are a random sampling of the true underlying
-distribution. This resulted in a fiducial sample of 569 SNe~Ia (422 SNe
-when more conservative cuts were applied).
+We have presented an initial study of the drift of the underlying SNe~Ia stretch
+distribution as a function of redshift. We built effectively volume-limited
+SN~Ia subsamples from the Pantheon dataset \citep[][SDSS, PS1, and
+SNLS]{scolnic2018a}, to which we added HST and SNfactory data from
+\cite{rigault2020} for the high- and low-redshift bins. We only considered the
+SNe that have been discovered in the redshift range of each survey in which
+observational selection effects are negligible, so that the observed SNe~Ia
+stretches are a random sampling of the true underlying distribution. This
+resulted in a fiducial sample of 569 SNe~Ia (422 SNe when more conservative cuts
+were applied).
 
 Following predictions made in \cite{rigault2020}, we introduced a redshift drift
-model that depends on the expected fraction of young and old SNe~Ia as
-a function of redshift, with each age population having its own
-underlying stretch distribution.
-
-In addition to this base model, we studied various
-distributions, including redshift-independent models; we also studied the
-prediction from a per-sample asymmetric Gaussian stretch distribution used, for
-instance, by the beams with bias correction Malmquist bias correction algorithm
-\citep{scolnic2016, kessler2017}.
-Our conclusions are listed below.
+model that depends on the expected fraction of young and old SNe~Ia as a
+function of redshift, with each age population having its own underlying stretch
+distribution.
+
+In addition to this base model, we studied various distributions, including
+redshift-independent models; we also studied the prediction from a per-sample
+asymmetric Gaussian stretch distribution used, for instance, by the beams with
+bias correction Malmquist bias correction algorithm \citep{scolnic2016,
+kessler2017}. Our conclusions are listed below.
 \begin{enumerate}
@@
         a way that observational selection effects alone cannot explain. This
-        result is largely independent of the details of each age-population model.
-
-    \item Redshift-independent models are quantitatively excluded as
-        suitable descriptions of the data relative to our base model. This model
-        assumes that (1) the younger population has a unimodal Gaussian stretch
-        distribution while the older population stretch distribution is
-        bimodal, one mode being the same as the young one, and (2) the evolution of
-        the relative fraction of younger and older SNe~Ia follows the prediction
-        made in \cite{rigault2020}. This second result further supports
-        the existence of both young and old SN~Ia populations, in agreement with
+        result is largely independent of the details of each age-population
+        model.
+
+    \item Redshift-independent models are quantitatively excluded as suitable
+        descriptions of the data relative to our base model. This model assumes
+        that (1) the younger population has a unimodal Gaussian stretch
+        distribution while the older population stretch distribution is bimodal,
+        one mode being the same as the young one, and (2) the evolution of the
+        relative fraction of younger and older SNe~Ia follows the prediction
+        made in \cite{rigault2020}. This second result further supports the
+        existence of both young and old SN~Ia populations, in agreement with
         rate studies \citep{mannucci2005, scannapieco2005, sullivan2006,
@@
 
-    \item Models using survey-based asymmetric Gaussian distributions,
-        for instance, as employed in the current implementation of BBC, are
-        excluded as a good description of the data relative to our
-        drifting model. This means that the sample-based approach does not accurately
-        account for redshift drift, a problem that will be exacerbated
-        as surveys span increasingly larger redshift ranges. As a
-    result, even if the necessary extra degrees of freedom might be acceptable
-    given the large number of SNe~Ia in cosmological studies, extrapolating the
-    SN property distributions from the volume-limited part of a survey to its
+    \item Models using survey-based asymmetric Gaussian distributions, for
+        instance, as employed in the current implementation of BBC, are excluded
+        as a good description of the data relative to our drifting model. This
+        means that the sample-based approach does not accurately account for
+        redshift drift, a problem that will be exacerbated as surveys span
+        increasingly larger redshift ranges. As a result, even if the necessary
+        extra degrees of freedom might be acceptable given the large number of
+        SNe~Ia in cosmological studies, extrapolating the SN property
+        distributions from the volume-limited part of a survey to its
         Malmquist-biased magnitude-limited part would still be inaccurate
@@
 
-We considered a simple two-population Gaussian mixture
-modeling. Additional data free from significant Malmquist bias would enable
-us to refine it as required. We note that samples at the low- and high-redshift
-ends of the Hubble diagram would be particularly helpful for this drifting
-analysis; fortunately, this will soon be provided by the Zwicky Transient
-Facility \citep[low-$z$,][]{bellm2019, graham2019} and the Subaru and SeeChange
-SNe~Ia programs (high-$z$), respectively.
-
-The next step in this line of analysis will incorporate our model into
-the SNANA framework \citep{SNANA}, both to more accurately account for
-observational selection functions and to test the effect of our model on the
-derivation of cosmological parameters. This study is currently being conducted.
+We considered a simple two-population Gaussian mixture modeling. Additional data
+free from significant Malmquist bias would enable us to refine it as required.
+We note that samples at the low- and high-redshift ends of the Hubble diagram
+would be particularly helpful for this drifting analysis; fortunately, this will
+soon be provided by the Zwicky Transient Facility \citep[low-$z$,][]{bellm2019,
+graham2019} and the Subaru and SeeChange SNe~Ia programs (high-$z$),
+respectively.
+
+The next step in this line of analysis will incorporate our model into the SNANA
+framework \citep{SNANA}, both to more accurately account for observational
+selection functions and to test the effect of our model on the derivation of
+cosmological parameters. This study is currently being conducted.
 

### 2021-03-05 Nicolas.tex
Not inversing 'are used' and 'are compared' in fine

@@
 
-\date{Submitted to A\&A the 19th of May 2020}
+\date{Submitted to A\&A\ the 19th of May 2020}
 
@@
 
-We note that when models that were adjusted on individual subsamples rather than
-globally \textbf{are compared}, the Bayesian information criterion
+We note that when models are compared that were adjusted on individual
+subsamples rather than globally, the Bayesian information criterion
 ($\mathrm{BIC} = -2\ln(L) + k\ln(n)$, with $n$ the number of data points) might
@@
 correct modeling of potential SN redshift drift should be further studied and
-care should be taken when samples that are affected by observational selection
-effects are used.
+care should be taken when samples are used that are affected by observational
+selection effects.
 

### 2021-03-16 Nicolas.tex
80 char limit in .tex
Wrong date in Rigault 2020
Removed bold

@@
     are indications that the underlying SN~Ia stretch distribution evolves as a
-    function of redshift, and that the \textbf{age} drifting model is a better
+    function of redshift, and that the age drifting model is a better
     description of the data than any time-constant model, including the
@@
 Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
-enabled the discovery that the Universe\textbf{'s expansion accelerates}
+enabled the discovery that the Universe's expansion accelerates
 \citep{riess1998, perlmutter1999}. They remain a key cosmological probe today
@@
 
-We here step aside from the cosmological analyses to probe the
-validity of our modeling of the SN population, which we claim is constituted
-of two age populations \citep{rigault2013, rigault2015, rigault2020}: an older
-and a younger population, the former having lower light-curve stretches on average and
+We here step aside from the cosmological analyses to probe the validity of our
+modeling of the SN population, which we claim is constituted of two age
+populations \citep{rigault2013, rigault2015, rigault2020}: an older and a
+younger population, the former having lower light-curve stretches on average and
 being brighter after standardization. We use the correlation between the SN age
-as probed by the LsSFR and the SN stretch to model the expected evolution of
-the underlying SN stretch distribution as a function of redshift. This modeling
+as probed by the LsSFR and the SN stretch to model the expected evolution of the
+underlying SN stretch distribution as a function of redshift. This modeling
 relies on three assumptions: (1) There are two distinct populations of SNe~Ia,
@@
 follows the model presented in \cite{rigault2020}, and (3) the underlying
-distribution of stretch for each age sample is constant. This paper tests this specific model with datasets from the literature.
+distribution of stretch for each age sample is constant. This paper tests this
+specific model with datasets from the literature.
 
@@
 scatter plot, our model of the underlying SN~Ia stretch distribution is defined
-as follows: \textbf{the stretch distribution of the younger population
+as follows: the stretch distribution of the younger population
 ($\log(\mathrm{LsSFR})\geq-10.82$) is modeled as a single normal distribution
@@
 \mathcal{N}(\mu_2, \sigma_2{}^2)$, where one mode is the same as for the young
-population, $a$ representing the relative effect of the two modes.}
+population, $a$ representing the relative effect of the two modes.
 
@@
     \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
-    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} light-curve stretch ($x_1$)
-        as a function of the LsSFR for
-        SNfactory SNe. The color corresponds to the
-        probability, $p_y$, for the SNe~Ia to be young, i.e., to have
-        $\log\mathrm{LsSFR} \geq -10.82$ \citep[see][]{rigault2020}.
-        \textit{Right}: $p_y$-weighted histogram of the SN stretches, as well as
-        the adjusted base model; contributions of the younger and older population
-        are shown in purple and yellow, respectively.}
+    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} light-curve stretch
+        ($x_1$) as a function of the LsSFR for SNfactory SNe. The color
+        corresponds to the probability, $p_y$, for the SNe~Ia to be young, i.e.,
+        to have $\log\mathrm{LsSFR} \geq -10.82$ \citep[see][]{rigault2020}.
+    \textit{Right}: $p_y$-weighted histogram of the SN stretches, as well as the
+adjusted base model; contributions of the younger and older population are shown
+in purple and yellow, respectively.}
     \label{fig:stretchlssfr}
@@
     model when applied to the SNfactory dataset only (114 SNe~Ia), the fiducial
-569 SN~Ia sample, or the conservative \textbf{sample} (422).}
+569 SN~Ia sample, or the conservative sample (422).}
     \label{tab:modelresults}
@@
 
-%\bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T., et al.\ 2019, \prl, 122, 221301
-
 % Q
@@
 \bibitem[Rigault et al.(2020)]{rigault2020} Rigault, M., Brinnel, V., Aldering,
-G., et al.\ 2029, \aap, 644, A176
+G., et al.\ 2020, \aap, 644, A176
 

### 2021-04-29 Nicolas.tex
Added proof and  corrected refs, signed copright form

@@
     Université Clermont Auvergne, CNRS/IN2P3, Laboratoire de
-    Physique de Clermont, F-63000 Clermont-Ferrand, France.
+    Physique de Clermont, F-63000 Clermont-Ferrand, France
     \and
     Physics Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720 
+    1 Cyclotron Road, Berkeley, CA, 94720, USA
     \and
     Institut fur Physik, Humboldt-Universität zu Berlin, Newtonstr. 15,
-    12489 Berlin
+    12489 Berlin, Germany
     \and
-    University of Southampton: Southampton, GB
+    University of Southampton, Southampton, UK
 }
 
-\date{Submitted to A\&A\ the 19th of May 2020}
+\date{Submitted to A\&A\ the 19th of May 2020 / Accepted 26 February 2021}
 
@@
 
-\bibitem[Brout et al.(2019)]{brout2019} Brout, D., Scolnic, D., Kessler, R., et
-al.\ 2019, \apj, 874, 150
-
 \bibitem[Burnham \& Anderson(2004)]{burnham2004} Burnham, K., Anderson, D., \
@@
 D.~M.\ 2015, \apj, 812, 3 1
-
-\bibitem[Jones et al.(2018)]{jones2018} Jones, D.~O., Riess, A.~G., Scolnic,
-D.~M., et al.\ 2018, \apj, 867, 108
-
-\bibitem[Jones et al.(2018)b]{jones2018b} Jones, D.~O., Scolnic, D.~M., Riess,
-A.~G., et al.\ 2018, \apj, 857, 51
 
@@
 
-\bibitem[Saunders et al.(2018)]{saunders2018} Saunders, C., Aldering, G.,
-Antilogus, P., et al.\ 2018, \apj, 869, 167. doi:10.3847/1538-4357/aaec7e
-
 \bibitem[Scannapieco \& Bildsten(2005)]{scannapieco2005} Scannapieco, E., \&
@@
 
-\bibitem[Strolger et al.(2004)]{strolger04} Strolger, L.-G., Riess, A.~G.,
+\bibitem[Strolger et al.(2004)]{strolger04} Strolger, L.-G., Riess, A.~G., et
+al.\ 2004, \apj, 613, 200
 
